### 2019-12-16 paper_main.tex
added more acknowledgements

% GravExplain
\documentclass[prb,preprint]{revtex4-1} 
% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}

% Note that AJP uses the same style as Phys. Rev. B (prb).
% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
% We look forward to receiving your submission to AJP.
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{float}
\usepackage{mathtools}
\usepackage{hyperref}

\bibliographystyle{apsrev4-2}
% \setlength{\parindent}{0pt}

\begin{document}

\title{GravExplain}

\author{James W. Gardner}
\email{u6069809@anu.edu.au}
\affiliation{Research School of Physics, Australian National University, 60 Mills Rd, Acton ACT 2601 Australia}

\author{Hannah Middleton}
\email{hannah.middleton@unimelb.edu.au}
\author{Andrew Melatos}
\email{amelatos@unimelb.edu.au}
\affiliation{School of Physics, University of Melbourne, Parkville, VIC, 3010, Australia}
\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}

% Summer 2019/202
\date{\today}

% AJP requires an abstract for all regular article submissions.
\begin{abstract}
GravExplain

\end{abstract}

\maketitle

\section{Introduction}
GravExplain

\section{Results}
GravExplain

\section{Conclusion}
GravExplain

\appendix
\section{}
GravExplain

\begin{acknowledgments}
We gratefully acknowledge Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance.

J.~Gardner also acknowledges the ANU PhB program for providing travel funding during the project.

This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.

\end{acknowledgments}

\end{document}

### 2019-12-16 paper/paper_main.tex
added an extra acknowledgement and moved paper tex into own directory

--- 2019-12-16 paper_main.tex
@@
 \section{Introduction}
-GravExplain
+A network of gravitational wave observatories around the world are .....
+
+A Michelson interferometer is commonly used by gravitational wave research groups to demonstrate gravitational wave science to non-specialist audiences. 
+These are typically table-top demonstrations
 
@@
 
+The authors are grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
+
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.

### 2019-12-19 paper/paper_main.tex
added a couple of comments for later
a first go at the introduction text and adding bib file

--- 2019-12-16 paper_main.tex
@@
 \usepackage{mathtools}
+\usepackage[usenames,dvipsnames]{xcolor}
 \usepackage{hyperref}
@@
 % \setlength{\parindent}{0pt}
+
+\newcommand{\han}{\textcolor{orange}}
 
@@
 \section{Introduction}
-GravExplain
+\han{just had a first go at the introduction - not sure where we are aiming yet, so will probably end up re-writing!}
+
+% general gw intro
+In 2015 gravitational waves were observed for the first time from the merger of a binary black hole system. 
+The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe.
+Since then, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi} mergers.
+With the increase in public interest in gravitational wave science, many gravitational wave research groups around the world have produced demonstrations and activities to explain this research to a general audience.
+Activites range from online data analysis tutorials,~\cite{GWOSC:Online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions\han{add refs}, artistic interpretation, and musical performance \han{add refs}.
+
+% what are gws and how do we detect them...
+Gravitational waves are a prediction of Einstein's theory of general relativity. 
+They are distrubances in spacetime caused by the accelleration of massive objects. 
+The effect of them is to change distances; a ``stretch and squeeze'' effect. 
+Observatories such as LIGO, Virgo and KAGRA (\han{reference}) are laser interferometers; they use the interference of laser light to measure changes in distance. 
+The observatories themselves are extremely complex devices, however they are based on the Michelson interferometer. 
+Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate this science to non-specialist audiences~\cite{}\han{refs} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
+
+% continuous gws
+To date the network of gravitational wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018}\han{update this ref and add GWOSC}. 
+However, these observatories are also searching for continuous gravitational waves; persistent, periodic signals which should be present in the data for many months or years. 
+A prime target for these searches are rotating neutron stars. 
+So far many searches have been made of LIGO and Virgo data, however no detection has been made so far.
+
+% overview
+In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous waves to a general audience. 
+The techniques used here are very similar to those used by the real continuous wave searches. 
+In section X we describe...... \han{add this in at the end}
 
@@
 
+\bibliographystyle{myunsrt}
+\bibliography{ifoDemoBib}
+
 \end{document}

### 2019-12-19 paper/paper_main.tex
some rewording in the introduction

@@
 % general gw intro
-In 2015 gravitational waves were observed for the first time from the merger of a binary black hole system. 
+In 2015 gravitational waves were observed for the first time from the merger two black holes in a binary system.~\cite{GW150914} 
 The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe.
 Since then, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi} mergers.
-With the increase in public interest in gravitational wave science, many gravitational wave research groups around the world have produced demonstrations and activities to explain this research to a general audience.
-Activites range from online data analysis tutorials,~\cite{GWOSC:Online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions\han{add refs}, artistic interpretation, and musical performance \han{add refs}.
+In recent years there has been an increased public interest in gravitaitonal-wave science. 
+Many gravitational wave research groups around the world have produce demonstrations and activities to aid in explaining this topic to a general audience
+Activites range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations,\han{ref} phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions\han{add refs}, artistic interpretation, and musical performance \han{add refs}.
 
@@
 In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous waves to a general audience. 
-The techniques used here are very similar to those used by the real continuous wave searches. 
+The techniques used are very similar to those used by researchers to search for continuous waves in LIGO and Virgo data. 
 In section X we describe...... \han{add this in at the end}

### 2020-01-16 paper/paper_main.tex
some small edits in the intro

--- 2019-12-16 paper/paper_main.tex
@@
 \section{Introduction}
-A network of gravitational wave observatories around the world are .....
 
-A Michelson interferometer is commonly used by gravitational wave research groups to demonstrate gravitational wave science to non-specialist audiences. 
-These are typically table-top demonstrations
+% general gw intro
+In 2015 gravitational waves were observed for the first time from the merger of two black holes in a binary system.~\cite{GW150914} 
+The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe.
+Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
+In recent years there has been an increased public interest in gravitaitonal-wave science. 
+Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience
+Activites range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance \han{add refs}.
+
+% what are gws and how do we detect them...
+Gravitational waves are a prediction of Einstein's theory of general relativity. 
+They are distrubances in spacetime caused by the accelleration of massive objects. 
+The effect of gravitational waves is to change distances; a ``stretch and squeeze'' effect. 
+Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
+The observatories themselves are extremely complex devices, however they are based on the Michelson interferometer. 
+Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate this science to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO}\han{add exhibit ifo when available} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
+
+% continuous gws
+To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
+However, these observatories are also searching for continuous gravitational waves; persistent, periodic signals which should be present in the data for many months or years. 
+A prime target for these searches are rotating neutron stars. 
+So far many searches have been made of LIGO and Virgo data, however no detection has been made so far.
+
+% overview
+In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous waves to a general audience. 
+The techniques used are very similar to those used by researchers to search for continuous waves in LIGO and Virgo data. 
+In section X we describe...... \han{add this in at the end}
 

### 2020-01-17 paper/paper_main.tex
48th day of Summer

@@
 
+\newcommand{\jam}{\textcolor{magenta}}
 \newcommand{\han}{\textcolor{orange}}
@@
 
-\title{GravExplain}
+\title{GravExplain: Continuous gravitational wave data analysis in the undergraduate laboratory}
 
@@
 \section{Introduction}
+\jam{there’s this introduction to the domain context of gravitational waves (which is super awesome) but there’s also basic things about michelson interferometers, do we need to explain that?}
+
+\subsection{Gravitational Waves}
+\jam{a previous supervisor hated having subsections right after a section heading, let me know if this displease you as well}
 
@@
 
-\section{Results}
-GravExplain
+\subsection{Interferometry}
+% path difference, virtual thin film interference
 
-\section{Conclusion}
-GravExplain
+\section{Method}
+
+\subsection{Data capture: webcam}
+% how to analyse video, python OpenCV2
+
+\subsection{Advanced data capture: photodiode}
+% leave circuit design to an appendix
+
+\section{Results and analysis}
+% combined/interwoven results, analysis, and discussion
+
+\subsection{Tracking tones: the Viterbi algorithm}
+% explaination and successful application
+\jam{should we introduce viterbi in the introduction, probably?}
+
+\subsection{Hearing harmonies and chords}
+% the best we can do right now
+% write about removing mains noise
+\jam{this is as far as we go right now}
+
+\subsection{Reconstructing music and voice}
+% signal processing required
+% may well be a ``future work'' section at the end
+\jam{! purely hypothetical right now !}
+
+\section{Conclusions}
+% we are able to successfully ...
 
@@
 \section{}
-GravExplain
 

### 2020-01-17 report_main.tex
48th day of Summer

--- 2020-01-17 paper/paper_main.tex
@@
 \newcommand{\jam}{\textcolor{magenta}}
-\newcommand{\han}{\textcolor{orange}}
 
@@
 \section{Introduction}
-\jam{there’s this introduction to the domain context of gravitational waves (which is super awesome) but there’s also basic things about michelson interferometers, do we need to explain that?}
-
-\subsection{Gravitational Waves}
-\jam{a previous supervisor hated having subsections right after a section heading, let me know if this displease you as well}
-
-% general gw intro
-In 2015 gravitational waves were observed for the first time from the merger of two black holes in a binary system.~\cite{GW150914} 
-The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe.
-Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years there has been an increased public interest in gravitaitonal-wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience
-Activites range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance \han{add refs}.
-
-% what are gws and how do we detect them...
-Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are distrubances in spacetime caused by the accelleration of massive objects. 
-The effect of gravitational waves is to change distances; a ``stretch and squeeze'' effect. 
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
-The observatories themselves are extremely complex devices, however they are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate this science to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO}\han{add exhibit ifo when available} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
-
-% continuous gws
-To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-However, these observatories are also searching for continuous gravitational waves; persistent, periodic signals which should be present in the data for many months or years. 
-A prime target for these searches are rotating neutron stars. 
-So far many searches have been made of LIGO and Virgo data, however no detection has been made so far.
-
-% overview
-In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous waves to a general audience. 
-The techniques used are very similar to those used by researchers to search for continuous waves in LIGO and Virgo data. 
-In section X we describe...... \han{add this in at the end}
+% see paper?
 
@@
 
+The authors are grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
+
 J.~Gardner also acknowledges the ANU PhB program for providing travel funding during the project.
-
-The authors are grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
 

### 2020-01-21 paper/paper_main.tex
51st day of Summer
53rd day of Summer

@@
 
-\title{GravExplain: Continuous gravitational wave data analysis in the undergraduate laboratory}
+\title{GravExplain: Continuous gravitational wave data analysis in a table-top experiment}
 
@@
 \section{Introduction}
-\jam{there’s this introduction to the domain context of gravitational waves (which is super awesome) but there’s also basic things about michelson interferometers, do we need to explain that?}
-
-\subsection{Gravitational Waves}
-\jam{a previous supervisor hated having subsections right after a section heading, let me know if this displease you as well}
 
@@
 
-\subsection{Interferometry}
+\section{Method and results}
+\jam{a previous supervisor hated having subsections right after a section heading, let me know if this displeases you as well}
+
+\subsection{Hardware: Michelson interferometry}
 % path difference, virtual thin film interference
-
-\section{Method}
 
@@
 
-\subsection{Advanced data capture: photodiode}
-% leave circuit design to an appendix
-
-\section{Results and analysis}
+% \section{Results and analysis}
 % combined/interwoven results, analysis, and discussion
@@
 
+\subsection{Advanced data capture: photodiode}
+% leave circuit design to an appendix
+
+\subsection{Photodiode viterbi}
+\jam{inject simulation wandering tone}
+
 \subsection{Hearing harmonies and chords}
@@
 
-\subsection{Reconstructing music and voice}
+\subsection{Future work: reconstructing music and voice}
 % signal processing required
@@
 \begin{acknowledgments}
-We gratefully acknowledge Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance.
-
-J.~Gardner also acknowledges the ANU PhB program for providing travel funding during the project.
+We gratefully acknowledge Jude Prezens, Alex Tolotchkoc, Blake Molyneux, Robin Evans, and Bill Moran for their technical advice and generous assistance throughout the project.
 
@@
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
+
+We also acknowledge the ANU PhB program for providing travel funding during the project.
 

### 2020-01-31 paper/paper_main.tex
62nd day of Summer

@@
 \begin{abstract}
-GravExplain
+Continuous gravitational wave searches use sophisticated statistical techniques to search for signals in noisy data.
+We demonstrate some of these techniques in an table-top Michelson interferometer.
+We also create an optical microphone capable of playing back simple recordings of injected audio.
 
@@
 
-\section{Method and results}
-\jam{a previous supervisor hated having subsections right after a section heading, let me know if this displeases you as well}
+% \section{Method and results}
+\section{Table-top gravitational wave science}
 
@@
 
-\subsection{Data capture: webcam}
+\subsection{Method: webcam}
 % how to analyse video, python OpenCV2
@@
 
+\subsection{Advanced method: photodiode and raspberry pi}
+% leave circuit design to an appendix
+% only briefly mention Salen-Key, leave to signals processing section
+
 \subsection{Tracking tones: the Viterbi algorithm}
 % explaination and successful application
-\jam{should we introduce viterbi in the introduction, probably?}
+% \jam{should we introduce viterbi in the introduction, probably?}
 
-\subsection{Advanced data capture: photodiode}
-% leave circuit design to an appendix
+% we can do this with both webcam and photodiode
 
-\subsection{Photodiode viterbi}
-\jam{inject simulation wandering tone}
+\section{Optical microphone}
+% describe aim of optical microphone: play back injected sounds
+% acknowledge problems: fringe counting, transfer function, coupling
 
-\subsection{Hearing harmonies and chords}
-% the best we can do right now
-% write about removing mains noise
-\jam{this is as far as we go right now}
+\subsection{Signals processing}
+% breakdown of model and all of the filters
 
-\subsection{Future work: reconstructing music and voice}
-% signal processing required
-% may well be a ``future work'' section at the end
-\jam{! purely hypothetical right now !}
+\subsection{Reconstructing music and voice}
+% results of processing, likely mixed with above subsection
 
@@
 \appendix
-\section{}
+
+\newpage
+\section{Intensity relation derivation}
+\begin{align}
+    y_0(t) &= A \sin{(k L - \omega t)} \\
+    y_1(t) &= A \sin{(k (L + 2 d \cos{\theta}) - \omega t)} \\
+    Y(t) &= A \sin{(k L - \omega t)} + A \sin{(k (L + 2 d \cos{\theta}) - \omega t)} \\
+    \phi &= 2 k d \cos{\theta} \\
+    Y(t) &= A\; \Im\{e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)}\} \\
+    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \phi/2 - \omega t)} + e^{i (k L + \phi/2 + \phi/2 - \omega t)}\} \\    
+    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \omega t)} (e^{- i \phi/2} + e^{i \phi/2})\} \\    
+    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \omega t)} (2 \cos{(\phi/2)})\} \\ 
+    Y(t) &= 2 A \cos{(\phi/2)}\; \Im\{e^{i (k L + \phi/2 - \omega t)}\} \\  
+    \therefore I(t) &= I_0 \cos^2{(\phi/2)}, \; \phi = \frac{4 \pi \cos{\theta}}{\lambda}\, d(t) \\
+    \frac{dI}{dt} &= \frac{dI}{d\phi}\; \frac{d\phi}{dt}\\
+    \frac{dI}{d\phi} &= - I_0 \cos{(\phi/2)} \sin{(\phi/2)}\\
+    &= - \frac{I_0}{2}\, \sin{(\phi)}\\
+    \delta\phi &= \frac{4 \pi \cos{\theta}}{\lambda}\, h\\
+    \therefore \delta I &= \frac{- 2 \pi I_0 \cos{\theta}}{\lambda}\,\sin{(\phi_0)}\, h
+\end{align}
+
+\section{Circuit design}
+% schematic of breadboard and connections to pi
+% https://www.circuit-diagram.org/editor/
 

### 2020-01-31 paper/paper_main.tex
62nd day of Summer

@@
 % https://www.circuit-diagram.org/editor/
+% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{circuit_diagram.pdf}
+	\caption{Circuit diagram for photo-diode reading.
+Photo-diode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC), that’s finally read by the special purpose input (SPI) pins of a Raspberry Pi}
+	\label{fig:circuit_diagram}
+\end{figure}
 

### 2020-02-03 paper/paper_main.tex
65th day of Summer

@@
 \usepackage{hyperref}
+% \usepackage{siunitx}
 
@@
 
-% Summer 2019/202
+% Summer 2019/2020
 \date{\today}
@@
 Continuous gravitational wave searches use sophisticated statistical techniques to search for signals in noisy data.
-We demonstrate some of these techniques in an table-top Michelson interferometer.
+We demonstrate some of these techniques in a table-top Michelson interferometer.
 We also create an optical microphone capable of playing back simple recordings of injected audio.
@@
 \maketitle
+
+% cover page that attributes all contributions
 
@@
 Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years there has been an increased public interest in gravitaitonal-wave science. 
+In recent years there has been an increased public interest in gravitational-wave science. 
 Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience
-Activites range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance \han{add refs}.
+Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance \han{add refs}.
 
@@
 Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are distrubances in spacetime caused by the accelleration of massive objects. 
+They are disturbances in spacetime caused by the acceleration of massive objects. 
 The effect of gravitational waves is to change distances; a ``stretch and squeeze'' effect. 
@@
 The observatories themselves are extremely complex devices, however they are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate this science to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO}\han{add exhibit ifo when available} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
+Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate this science to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO}\han{add exhibit info when available} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
 
@@
 % \section{Method and results}
-\section{Table-top gravitational wave science}
+\section{Table-top gravitational wave data science}
 
-\subsection{Hardware: Michelson interferometry}
-% path difference, virtual thin film interference
+\subsection{Michelson interferometry}
+% hardware: path difference, virtual thin film interference
 
-\subsection{Method: webcam}
+\subsection{The webcam method}
 % how to analyse video, python OpenCV2
+
+% acknowledge problems: fringe counting, transfer function, coupling
 
@@
 
-\subsection{Advanced method: photodiode and raspberry pi}
-% leave circuit design to an appendix
-% only briefly mention Salen-Key, leave to signals processing section
-
-\subsection{Tracking tones: the Viterbi algorithm}
+\subsection{Tracking tones with the Viterbi algorithm}
 % explaination and successful application
@@
 \section{Optical microphone}
-% describe aim of optical microphone: play back injected sounds
-% acknowledge problems: fringe counting, transfer function, coupling
+% aim of optical microphone: play back injected sounds
 
-\subsection{Signals processing}
+The natural succession to tracking simple tones is to try complex audio signals, such as simple chords, and ultimately, music and speech.
+The Viterbi algorithm is not applicable to these signals, as it is designed to track a single tone signal that slowly, continuously changes frequency.
+Whereas music and speech consist of many tones that jump frequencies.
+As such, the interest now is in being able to record and play-back these complex audio signals.
+Using the interferometer set-up to do makes it an optical microphone, using light to capture sound.
+
+\subsection{The photodiode method}
+% advanced method: explain how we capture data
+
+For an optical microphone to be sensitive to speech and music it needs a frequency response far higher than the 30Hz of a standard webcam.
+Both speech and music need at least kHz response, with speech intelligibility requiring up to 3kHz and music preferably at least 8kHz.
+This can be addressed by a high-speed webcam, or more simply, by a photodiode.
+
+A photodiode is an electrical component that acts as a regular diode when no light is incident on it (blocking any current flow in the reverse direction), but becomes more and more conductive in the reverse direction as the intensity of incident light increases.
+Fundamentally, a current is created in the photodiode by the photo-electric effect.
+Placing a photodiode in reverse-bias over an op-amp creates a photo-detector that will produce a voltage dependent linearly (as a sound approximation) to the incident intensity.
+
+To the interferometer set-up, placing a photodiode anywhere in the interference pattern will record the signal.
+The signal is independent of the position, up to an initial phase difference, since the interference principle means that all peaks become troughs together (and visa versa).
+This experiment placed it just off-centre.
+
+The analog signal from the photodiode is captured by an analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi, in the standard configuration.
+This digitally samples the signal at roughly 16kHz which introduces the first step of signal processing.
+Any existing frequency component of the analog signal above the Nyquist frequency of 8kHz would be aliased down into the detected range.
+To prevent this, an anti-aliasing Sallen-Key filter tuned to 16kHz is included before the ADC.
+This component attenuates any frequencies above 16kHz, before they get digitally sampled.
+
+All together, the photodiode measuring circuit diagram is shown in Figure~\ref{fig:circuit_diagram} in the appendices.
+It’s worth noting that it’s not immediately clear what is limiting the sample rate to 16kHz, but it is likely non-optimal reading by the Pi.
+% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
+It is also worth noting that this method does not solve the problems of fringe counting nor speaker coupling.
+
+\subsection{Signals processing: basics}
+% transfer function, power-spectral-density
+
+\subsection{Signals processing: filters}
 % breakdown of model and all of the filters
+
+% link back, gravitational wave searches also do lots of signal processing
 
@@
 
-\begin{figure}
+\begin{figure}%[H]
 	\includegraphics[width=\textwidth]{circuit_diagram.pdf}
@@
 \begin{acknowledgments}
-We gratefully acknowledge Jude Prezens, Alex Tolotchkoc, Blake Molyneux, Robin Evans, and Bill Moran for their technical advice and generous assistance throughout the project.
+The authors are grateful to Robin Evans, Bill Moran, Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
 
-The authors are grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
+The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
 
@@
 
-We also acknowledge the ANU PhB program for providing travel funding during the project.
+Travel during the project was supported by the Australian National University PhB Science  program.
 

### 2020-02-03 paper/paper_main.tex
65th day of Summer

@@
 
-The analog signal from the photodiode is captured by an analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi, in the standard configuration.
+The analog signal from the photodiode is captured by a 10-bit analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi, in the standard configuration.
 This digitally samples the signal at roughly 16kHz which introduces the first step of signal processing.
@@
 % The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-It is also worth noting that this method does not solve the problems of fringe counting nor speaker coupling.
+The photodiode is found to also be able to replicate the Viterbi results from before, but for a much greater range of frequencies as expected.
+However, this method does not solve the problems of fringe counting nor speaker coupling from before.
 
-\subsection{Signals processing: basics}
-% transfer function, power-spectral-density
+\subsection{Initial recordings}
+% breakdown of all the filters used along the way
 
-\subsection{Signals processing: filters}
-% breakdown of model and all of the filters
+The optical microphone (interferometer and photodiode circuit ensemble) was tested on many recordings of human voice and on music ranging from simple chord melodies to full track songs.
+Each recording was around one minute long but processing was generally done on ten second segments.
+For each trial, the recording was played through the speaker while the Pi script noted the photo-voltage, with minimal activity about the optical table.
+The saved  voltage time-series was then directly converted to a .wav file and played as an audio recording.
 
-% link back, gravitational wave searches also do lots of signal processing
+These initial recordings were extremely noisy and painful to listen to, with a loud bass hum throughout.
+Spectra of all of these recordings showed a dominant mains hum at 50Hz with strong power up to and beyond the 8th harmonic thereof.
+This power is also present with the speaker off, as in Figure~\ref{fig:psd_noise}, and also weakly with the photodiode in complete darkness.
 
-\subsection{Reconstructing music and voice}
-% results of processing, likely mixed with above subsection
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{podo_noise_psd_zoom-cropped.pdf}
+	\caption{Power spectral density (PSD) plot of the optical microphone system with no injected audio (with speaker off), ambience shows high power in the 50Hz mains hum and its harmonics}
+	\label{fig:psd_noise}
+\end{figure}
+
+Early attempts to remove the mains hum included zeroing the channel of each 50Hz harmonic, this is effectively using a rectangular comb filter.
+Applying (multiplying by) any filter in frequency space is equivalent to convolving with the inverse Fourier transform of that filter.
+While this did remove the mains harmonics, it also destroyed the rest of the signal.
+Similarly, low-passing the signal is problematic as the harmonics are all over the 100Hz-1kHz region, which also carries a lot of the signal.
+Low-passing the logarithm of the signal performs better but still doesn’t significantly remove the noise.
+
+Of these simple filters, the best performing was a band-pass Butterworth filter with frequency response shown in Figure~\ref{fig:butterworth}.
+This filter is `maximally flat' within the band region, here 150Hz to 3kHz, and trails off outside of the band with slope dependent on the order of the filter, here of order 5.
+Behind the remaining noise one could hear something distinctly speech-like or musical, just with very low intelligibility (as tested by multiple participants), i.e.
+one couldn’t understand the speech but could recognise that someone was talking.
+
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{butterworth_150_3000-cropped.pdf}
+	\caption{Butterworth bandpass filter frequency response, green vertical lines show the band limits of 150Hz and 3kHz. Any amplitude beyond -3dB is significant attenuation.}
+	\label{fig:butterworth}
+\end{figure}
+
+\subsection{logMMSE estimator}
+% logMMSE
+
+Speech enhancement for a noisy channel is a classic problem in signal processing.
+Hu~\&~Loizou~2006~\cite{SubjectiveComparison} compare various methods and find the best performing statistical technique to be the logMMSE (alt.\ MMSE-LSA) estimator which minimises the mean square error (MSE) of the log-spectral amplitude.
+This means that it takes the logarithm of the spectra of the original signal and finds the fit (or estimate) that has the minimal MSE to that log-spectra.
+Use of the logarithm is a common technique in speech processing as it, somewhat surprisingly, does not significantly change the intelligibility to the human ear.
+
+Applying an existing implementation of logMMSE to the optical microphone recordings produces dramatically cleaner results.
+The time series and spectrum raw from the optical microphone and after the logMMSE filter are shown in Figures~\ref{fig:logMMSE_timeseries}~\&~\ref{fig:logMMSE_spectrum}.
+All of the mains harmonics are significantly attenuated.
+Music out of the optical microphone is varied, loud rhythmic bass and drums are received well and the beat and chords remain recognisable, airy flutes and violins though struggle to be heard in the results.
+Speech is also far better separated from the background, however, the intelligibility remains low with the speech sounding mumbled and indistinct.
+
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{filter_timeseries_aa_melatos-cropped.pdf}
+	\caption{Optical microphone raw voltage time series (top) and with logMMSE filter applied (bottom), showing only one second of one minute recording, the effect at the start of the time series after filtering is the result of a finite signal}
+	\label{fig:logMMSE_timeseries}
+\end{figure}
+
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{filter_spectrum_aa_melatos-cropped.pdf}
+	\caption{Optical microphone raw signal spectrum (top) and with logMMSE filter applied (bottom), showing only up to 2kHz and a detail of the spectrum (before filtering the 50Hz line has a strength of 40 leaving one unable to see the rest of the spectrum)}
+	\label{fig:logMMSE_spectrum}
+\end{figure}
+
+The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
+However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
+An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
+% Visually it is impossible to confirm this for high frequencies, but it remains possible.
+
+\subsection{Future work}
+% full model of transfer function
+% SEGAN and other machine learning
 

### 2020-02-04 paper/paper_main.tex
66th day of Summer

@@
 In recent years there has been an increased public interest in gravitational-wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience
+Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience.
 Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance \han{add refs}.
@@
 % \section{Method and results}
-\section{Table-top gravitational wave data science}
+\section{Table-top gravitational wave science}
+
+Detectable gravitational waves are caused by the most massive, violent events in the universe.
+Even so, the effects they produce are on the order of one part in $10^{21}$ for our best detectors.
+For a table-top demonstration this is impractical, so we demonstrate the same techniques for capturing audio instead.
 
@@
 
+A Michelson interferometer is a well-known optics configuration used in many undergraduate and research labs.
+A beam of coherent (monochromatic and equal phase) laser light is split to travel down and back from a mirror at the end of two arms at right-angles to each other, before returning to interfere at a screen to produce a pattern of concentric, circular fringes.
+See appendix for the exact set-up, lenses are included to focus the beam and view the interference pattern \jam{include Michelson schematic with speaker, webcam, photodiode and picture of interference pattern}.
+
+The appearance of the interference pattern, as shown in Figure~\ref{fig:interference_pattern}, is due to the path difference between one mirror and the virtual image of the other, resulting in a phase difference between the returning beams that meet at a point on the screen.
+As this phase difference only depends on the angle from the centre-line of the beam, the pattern is circularly symmetric for an aligned configuration.
+The intensity, $I$, of this pattern is given by Equation~\ref{eq:intensity_result} below.
+Where A is the amplitude of the electric field of the initial beam, d is the difference in the arm lengths, $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Equation~\ref{eq:intensity_derivation} in the appendix for the full derivation).
+\begin{equation}
+\label{eq:intensity_result}
+I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+\end{equation}
+
+This means that the intensity changes linearly with a small change in separation (again, see Equation~\ref{eq:intensity_derivation}).
+
+\begin{equation}
+	\delta I = \frac{- 8 \pi A^2 \cos{\theta} \sin{\phi}}{\lambda}\, \delta d
+\end{equation}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{webcam_still0.pdf}
+	\caption{Interference pattern from Michelson interferometer, red dot indicates measured point for intensity time series}
+	\label{fig:interference_pattern}
+\end{figure}
+
+% A note that while the interference pattern appears similar to the Airy pattern produced by a circular aperture, they are not in fact the same geometry.
+
 \subsection{The webcam method}
@@
 
-% acknowledge problems: fringe counting, transfer function, coupling
+In this demonstration, a 0.5W speaker is stuck to the back of one of the interferometer mirrors.
+As the speaker cone deforms to play audio, the mirror moves with a change in separation, $\delta d$, as some function of the amplitude and frequency of the injected audio.
+The motion of the fringes follows the motion of the speaker at the same frequency.
+If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected audio but with amplitude given by a transfer function (which accounts for the coupling and resonance of the speaker-mirror system).
+Large fringe motion means that multiple fringes can pass over the measured point in a single speaker deformation, multiple fringes being counted raises the frequency of the intensity time series relative to the injected audio.
 
-% \section{Results and analysis}
-% combined/interwoven results, analysis, and discussion
+We measure the intensity at a point on the screen with a standard USB webcam, sampling at 30Hz.
+This limits the spectral content of signals we can observe to under the Nyquist frequency of 15Hz.
+This isn’t an issue to demonstrating the techniques, but see later Section~\ref{sec:photodiode} for a method to raise this limit to audible frequencies.	While playing a tone under 15Hz through the speaker, we recorded the interference pattern with the webcam.
+Then, at an off-centre pixel, we take the green-channel of the RGB video as approximate to the intensity time series, as the laser is monochromatic green light.
+Taking the Fourier transform of this signal clearly shows the injected tone, as shown in Figure~\ref{fig:webcam_spectrum}.
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{webcam_expt_4_0209-cropped.pdf}
+	\caption{Spectrum of intensity time series for injected tone of 2.09Hz, fundamental and harmonics clearly present}
+	\label{fig:webcam_spectrum}
+\end{figure}
 
@@
 % explaination and successful application
-% \jam{should we introduce viterbi in the introduction, probably?}
+% lean on the continuous gravitational waves angle
 
-% we can do this with both webcam and photodiode
+The signals that continuous gravitational wave searches look for change frequency slowly over time.
+This means that we can’t just look at the spectral content for the full observing run and search for a peak.
+Instead, the common method is to bin the observed time series into chunks of time, compute the spectra for each time interval, and then path-find through the grid of spectra over time for the most likely signal.
+Where the path-finding is restricted to go forwards in time and only move in frequency by a set amount, here 0.11 Hz per second.
+The Fourier amplitude of each frequency is used as the weight at each vertex in the grid of spectra over time.
+
+The Viterbi algorithm is a path-finding algorithm that solves this problem.
+Given a weighted graph, a sequence of subgraphs (the stages or layers to path-find through), and any restrictions on connectivity (the allowed amount of source frequency change over time), Viterbi will find the best path (in terms of the weights) to each node in the next subgraph from the first subgraph at each iteration.
+At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), then called the Viterbi path.
+
+To demonstrate this technique, we inject (play through the speaker) a tone that wanders in frequency over time, remembering to integrate up the phase change over time.
+Alternatively, a function generator can be slowly controlled to achieve the same input.
+Shown in Figure~\ref{fig:viterbi_overlay} is the grid (spectrogram) of the spectra of the observed signal over time, along with the Viterbi path and the injected signal.
+Due to the high signal-to-noise ratio (SNR), the Viterbi algorithm easily recovers the injected frequency over time.
+Note that the start of the signal moves too quickly (more than 0.11 Hz per second) for Viterbi to track, and that there appears to be an anomaly at 150s, likely a disturbance around the interferometer.
+Decreasing the signal-to-noise ratio closer to unity or increasing the allowed change in frequency over time causes Viterbi to lose the signal.
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{expt_overlay_2_viterbi_test_webcam.pdf}
+	\caption{Spectrogram (spectra over time) of observed signal overlaid with white crosses for the Viterbi path and pink dots for the injected signal, notice the high correlation, the signal is able to be recovered}
+	\label{fig:viterbi_overlay}
+\end{figure}
+
+\jam{what more is there to say about Viterbi?}
 
@@
 
+This aim is auxiliary to demonstrating continuous wave techniques, as advanced signal processing is required to extract the signal from noisy data in gravitational wave searches.
+To be able to recover audio with the optical microphone similarly requires signal processing, although of a simpler nature.
+
 \subsection{The photodiode method}
+\label{sec:photodiode}
 % advanced method: explain how we capture data
@@
 % full model of transfer function
+A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
+Although we have the separation-intensity relation as linear for small motions, the other components of this full model are yet to be determined.
+
 % SEGAN and other machine learning
@@
 
-\newpage
+\section{Michelson interferometer schematic}
+\jam{please include}
+
+Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam. We did not find it necessary to demonstrate the results desired.
+
+% \newpage
 \section{Intensity relation derivation}
+
 \begin{align}
-    y_0(t) &= A \sin{(k L - \omega t)} \\
-    y_1(t) &= A \sin{(k (L + 2 d \cos{\theta}) - \omega t)} \\
-    Y(t) &= A \sin{(k L - \omega t)} + A \sin{(k (L + 2 d \cos{\theta}) - \omega t)} \\
-    \phi &= 2 k d \cos{\theta} \\
-    Y(t) &= A\; \Im\{e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)}\} \\
-    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \phi/2 - \omega t)} + e^{i (k L + \phi/2 + \phi/2 - \omega t)}\} \\    
-    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \omega t)} (e^{- i \phi/2} + e^{i \phi/2})\} \\    
-    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \omega t)} (2 \cos{(\phi/2)})\} \\ 
-    Y(t) &= 2 A \cos{(\phi/2)}\; \Im\{e^{i (k L + \phi/2 - \omega t)}\} \\  
-    \therefore I(t) &= I_0 \cos^2{(\phi/2)}, \; \phi = \frac{4 \pi \cos{\theta}}{\lambda}\, d(t) \\
-    \frac{dI}{dt} &= \frac{dI}{d\phi}\; \frac{d\phi}{dt}\\
-    \frac{dI}{d\phi} &= - I_0 \cos{(\phi/2)} \sin{(\phi/2)}\\
-    &= - \frac{I_0}{2}\, \sin{(\phi)}\\
-    \delta\phi &= \frac{4 \pi \cos{\theta}}{\lambda}\, h\\
-    \therefore \delta I &= \frac{- 2 \pi I_0 \cos{\theta}}{\lambda}\,\sin{(\phi_0)}\, h
+\label{eq:intensity_derivation}
+    y_0(t) &= A e^{i (k L - \omega t)} \\
+    y_1(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \\
+    I(t) &= \lvert A e^{i (k L - \omega t)} + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2 \\
+    I(t) &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
+    I(t) &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2) \\
+    I(t) &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
+    I(t) &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda} \\
+    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
+    \frac{dI}{d\phi} &= - 2 A^2 \sin{\phi}\\
+    \delta\phi &= \frac{4 \pi \cos{\theta}}{\lambda}\, \delta d\\
+    \therefore \delta I &= \frac{- 8 \pi A^2 \cos{\theta} \sin{\phi}}{\lambda}\, \delta d
 \end{align}

### 2020-02-05 paper/paper_main.tex
67th day of Summer

--- 2020-02-03 paper/paper_main.tex
@@
 
-\title{GravExplain: Continuous gravitational wave data analysis in a table-top experiment}
+\title{GravExplain: Continuous gravitational wave searches in a table-top experiment}
 
@@
 Continuous gravitational wave searches use sophisticated statistical techniques to search for signals in noisy data.
-We demonstrate some of these techniques in a table-top Michelson interferometer.
-We also create an optical microphone capable of playing back simple recordings of injected audio.
+We demonstrate some of these techniques with a table-top Michelson interferometer, namely the Viterbi algorithm for recovering injected tones and a signal processing filter pipeline.
+We also use the same experiment to demonstrate an optical microphone capable of playing back simple recordings of injected audio.
 
@@
 % \section{Method and results}
-\section{Table-top gravitational wave data science}
+\section{Table-top gravitational wave science}
+
+Detectable gravitational waves are caused by the most violent events in the universe.
+Like two massive black-holes merging with each other.
+Even so, the effects they produce are on the order of one part in $10^{21}$ for our best detectors.
+Given the impossibility of such detection with a table-top experiment, we demonstrate the same techniques but for observing sound waves.
 
@@
 
+A Michelson interferometer is a well-known optics configuration used in many undergraduate and research labs, including in the detectors for gravitational waves.
+A laser beam of coherent (monochromatic and equal phase) light is split into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
+The beams return to interfere at a screen and produce a pattern of concentric, circular fringes.
+
+\jam{See appendix for the exact set-up, lenses are included to focus the beam and view the interference pattern ||| include Michelson schematic with speaker, webcam, and photodiode}.
+
+The interference pattern, as shown in Figure~\ref{fig:interference_pattern}, is due to the path difference between the two arms resulting in a phase difference between the two beams when they meet at the screen.
+This path difference is perhaps better seen as the perpendicular separation between one mirror and the virtual image of the other.
+For an aligned configuration and a fixed separation, the phase difference at a point only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
+The intensity, $I$, of this pattern at a particular angle is given by Equation~\ref{eq:intensity_result} below.
+Where A is the amplitude of the electric field of the beam, d is the separation (or the difference in the arm lengths), $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Appendix~\ref{app:intensity_derivation} for the derivation to get to this equation).
+\begin{equation}
+\label{eq:intensity_result}
+I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+\end{equation}
+
+This equation also means that the intensity changes linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
+
+\begin{equation}
+\label{eq:intensity_change}
+	\delta I = \frac{- 8 \pi A^2 \cos{\theta} \sin{\phi}}{\lambda}\, \delta d
+\end{equation}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{webcam_still0.pdf}
+	\caption{Interference pattern from Michelson interferometer as viewed by webcam from below (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured \jam{do I need to add a scale to this picture?}}
+	\label{fig:interference_pattern}
+\end{figure}
+
+% A note that while the interference pattern appears similar to the Airy pattern produced by a circular aperture, they are not in fact the same geometry.
+
 \subsection{The webcam method}
@@
 
-% acknowledge problems: fringe counting, transfer function, coupling
+We demonstrate a signal source with a 0.5W speaker stuck to the back of one of the mirrors.
+This speaker was obtained from a standard commercial 3.5mm jack input, USB powered pair of speakers.
+Alternatively, a speaker can be driven by a function generator and power plugs.
+As the speaker deforms to play something, the mirror moves with a change in separation, $\delta d$, as some function of the amplitude and frequency of the sound.
+The motion of the fringes follows the motion of the mirror and therefore the speaker at the same frequency.
+If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected (played) audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
+Large fringe motion means that multiple fringes can pass over the measured point in a single speaker deformation, such over-counting would artificially raise the measured frequency.
+As such, any motion of the fringes, and so of the speaker, must be kept small.
 
-% \section{Results and analysis}
-% combined/interwoven results, analysis, and discussion
+We measure the intensity at a point on the screen with a standard commercial USB webcam, sampling at 30Hz (or 30 fps).
+This limits the spectral content of signals we can observe to be under the Nyquist frequency of 15Hz.
+This isn’t an issue to demonstrate what we want, but see later Section~\ref{sec:photodiode} to raise this limit to audible frequencies.
+While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
+Then, at an off-centre pixel as seen in Figure~\ref{fig:interference_pattern}, we take the green-channel of the RGB video as approximate to the total intensity, as the laser is monochromatic green light.
+We can play this time series of intensity over time as an audio recording, and the tone is clearly heard.
+Taking the Fourier transform of this signal also strongly shows the injected tone at the correct frequency, as shown in Figure~\ref{fig:webcam_spectrum}.
+Therefore, we’re confident that we’ll be able to track input frequencies.
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{webcam_expt_4_0209-cropped.pdf}
+	\caption{Spectrum of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz hitting the interferometer, injected fundamental frequency and harmonics thereof clearly present in spectrum}
+	\label{fig:webcam_spectrum}
+\end{figure}
 
@@
 % explaination and successful application
-% \jam{should we introduce viterbi in the introduction, probably?}
+% lean on the continuous gravitational waves angle
 
-% we can do this with both webcam and photodiode
+Continuous gravitational wave searches look for signals that change frequency slowly over time.
+This means that finding the signal isn’t as simple as searching for a peak in the spectral content for the full observing run.
+Instead, the common method is to split the observed time series into bins of time, compute the spectra for each time interval, and then find the most likely path for the signal frequency through the spectra over time.
+The plot of spectra over time is known as a spectrogram.
+This turns the problem of searching for a signal in noisy data into one of finding the best path through a grid of weights (here the Fourier amplitudes).
+We restrict the path-finding to some rate of frequency change of time, usually informed by a physical model, but here just set to 0.11 Hz per second to make it one frequency bin per every time bin.
+
+The Viterbi algorithm is a path-finding algorithm that solves this particular problem of going from one side of a grid to the other.
+See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.
+Abstractly speaking through, given a weighted graph (e.g.\ the spectrogram), a sensible sequence of subgraphs (e.g.\ the spectra at each time bin), and any restrictions on connectivity (e.g.\ the allowed amount of source frequency change over time), the Viterbi algorithm will find the best path to each node in the next subgraph all the way from the first subgraph at each iteration.
+At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path.
+
+To demonstrate this technique, we inject a tone that wanders in frequency over time, remembering to integrate up the phase change over time (see Appendix~\ref{app:phase_gotcha} for why).
+Alternatively, a function generator can be slowly controlled to achieve the same input.
+Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, along with the Viterbi path and the injected signal.
+With great success, the Viterbi algorithm recovers the injected frequency over time, likely due to the high signal-to-noise ratio (SNR).
+Note that the signal initially moves too quickly (more than 0.11 Hz per second) for Viterbi to track, and that there appears to be an anomaly at 150s, likely a disturbance around the interferometer.
+Decreasing the signal-to-noise ratio closer to unity or increasing the allowed change in frequency over time causes Viterbi to more easily lose the signal.
+However, the demonstration clearly shows the ability to recover a continuous wave with the Viterbi algorithm, just as is done in continuous gravitational wave searches.
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{expt_overlay_2_viterbi_test_webcam.pdf}
+	\caption{Spectrogram of observed signal overlaid with white crosses for the Viterbi path and pink dots for the injected signal, notice the high correlation between them indicating that the signal is able to be recovered; also note the deviation at the start when the signal moves too quickly and just before the background anomaly at 150s}
+	\label{fig:viterbi_overlay}
+\end{figure}
+
+\jam{What more is there to say about Viterbi?}
 
@@
 
-The natural succession to tracking simple tones is to try complex audio signals, such as simple chords, and ultimately, music and speech.
-The Viterbi algorithm is not applicable to these signals, as it is designed to track a single tone signal that slowly, continuously changes frequency.
-Whereas music and speech consist of many tones that jump frequencies.
-As such, the interest now is in being able to record and play-back these complex audio signals.
-Using the interferometer set-up to do makes it an optical microphone, using light to capture sound.
+The natural succession to simple tones is to study complex audio signals, such as music and speech.
+The Viterbi algorithm and tracking more broadly is not applicable to these signals.
+Compare a single tone that slowly, continuously changes frequency to music and speech which consist of a broad spectrum that is constantly changing.
+Instead, the interest here is in being able to record and play-back these complex audio signals, rather than track them.
+Using the interferometer to do so makes it, in ensemble with the measuring device, an ‘optical microphone’, using light to capture sound.
+
+Advanced signal processing is required to extract gravitational wave signals from very noisy observational data.
+Recovering audio from an optical microphone requires related signal processing.
+Although far simpler, it serves as another useful demonstration.
 
 \subsection{The photodiode method}
+\label{sec:photodiode}
 % advanced method: explain how we capture data
@@
 For an optical microphone to be sensitive to speech and music it needs a frequency response far higher than the 30Hz of a standard webcam.
-Both speech and music need at least kHz response, with speech intelligibility requiring up to 3kHz and music preferably at least 8kHz.
-This can be addressed by a high-speed webcam, or more simply, by a photodiode.
+Both speech and music need at least kHz range, with speech intelligibility requiring up to 3kHz and music preferably up to and beyond 8kHz.
+This can be addressed by using a high-speed webcam, or more simply, by using a photodiode as the measuring device.
 
-A photodiode is an electrical component that acts as a regular diode when no light is incident on it (blocking any current flow in the reverse direction), but becomes more and more conductive in the reverse direction as the intensity of incident light increases.
+A photodiode is an electrical component that acts as a regular diode (blocking any current flow in the reverse direction) when no light is incident on it, but becomes more and more conductive in the reverse direction as the intensity of incident light increases.
 Fundamentally, a current is created in the photodiode by the photo-electric effect.
-Placing a photodiode in reverse-bias over an op-amp creates a photo-detector that will produce a voltage dependent linearly (as a sound approximation) to the incident intensity.
+Placing a photodiode in reverse-bias over an op-amp creates a photo-detector that will produce a voltage dependent approximately linearly to the incident intensity.
+\jam{How sound is this approximation?} This experiment had the photodiode in the interference pattern at roughly the same spot as the webcam was monitoring previously, just off-centre.
 
-To the interferometer set-up, placing a photodiode anywhere in the interference pattern will record the signal.
-The signal is independent of the position, up to an initial phase difference, since the interference principle means that all peaks become troughs together (and visa versa).
-This experiment placed it just off-centre.
+The voltage signal from the photo-detector is captured by a 10-bit analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi, in the standard configuration.
+This digitally samples the signal at roughly 16kHz.
+As such, any frequency component of the analog signal above the Nyquist frequency of 8kHz would be aliased down into the detected range.
+To prevent this, an anti-aliasing Sallen-Key filter tuned to 16kHz is included before the ADC.
+This component attenuates any frequencies above 8kHz, before they get digitally sampled.
+It’s not immediately clear what’s limiting the sample rate to 16kHz, but it is likely non-optimal reading of the ADC by the Pi script.
+% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
 
-The analog signal from the photodiode is captured by an analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi, in the standard configuration.
-This digitally samples the signal at roughly 16kHz which introduces the first step of signal processing.
-Any existing frequency component of the analog signal above the Nyquist frequency of 8kHz would be aliased down into the detected range.
-To prevent this, an anti-aliasing Sallen-Key filter tuned to 16kHz is included before the ADC.
-This component attenuates any frequencies above 16kHz, before they get digitally sampled.
+The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}.
 
-All together, the photodiode measuring circuit diagram is shown in Figure~\ref{fig:circuit_diagram} in the appendices.
-It’s worth noting that it’s not immediately clear what is limiting the sample rate to 16kHz, but it is likely non-optimal reading by the Pi.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-It is also worth noting that this method does not solve the problems of fringe counting nor speaker coupling.
+\subsection{Initial recordings}
+% breakdown of all the filters used along the way
 
-\subsection{Signals processing: basics}
-% transfer function, power-spectral-density
+The optical microphone was tested on recordings of one human talking and on music ranging from simple melodies to full tracks.
+Each recording was around one minute long with processing done on ten second segments.
+For each observation, the audio was played through the speaker with minimal activity around interferometer as the Pi script recorded the signal.
+The saved time series was then directly converted to a .wav file and played as an audio recording.
 
-\subsection{Signals processing: filters}
-% breakdown of model and all of the filters
+These initial recordings were extremely noisy and painful to listen to, with a loud bass sound throughout.
+Spectra of all of these recordings showed a dominant mains hum at 50Hz with strong power up to and beyond the 8th harmonic thereof.
+This power is also present with the speaker off, as is shown in Figure~\ref{fig:psd_noise}, and also weakly when monitoring with the photodiode in complete darkness.
 
-% link back, gravitational wave searches also do lots of signal processing
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{podo_noise_psd_zoom-cropped.pdf}
+	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), show high power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the ambient lighting), otherwise is a fairly white spectrum except for the broad, shallow maximum at 750Hz}
+	\label{fig:psd_noise}
+\end{figure}
 
-\subsection{Reconstructing music and voice}
-% results of processing, likely mixed with above subsection
+Early attempts to remove the mains hum included zeroing the channel of each 50Hz harmonic, this is effectively using a rectangular comb filter.
+Applying a filter (multiplying by a function) in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
+While this did remove the mains harmonics, it also destroyed the rest of the signal.
+
+Attempting to low-pass the signal did better (to the ear) but was also problematic as the harmonics are all over the 100Hz-1kHz region, which also carries a lot of the speech and music information.
+Low-passing the logarithm of the signal and then converting back did better still but doesn’t significantly remove the mains noise.
+
+Of these simple filters, the best performing was a band-pass Butterworth filter with frequency response shown in Figure~\ref{fig:butterworth}.
+This filter is `maximally flat' within the band region, here 150Hz to 3kHz, and trails off outside of the band with slope dependent on the order of the filter, here of order 5.
+This still suffered from the presence of the higher mains harmonics.
+However, behind the remaining noise one could hear something distinctly speech-like or musical, just with very low intelligibility, you couldn’t understand the speech but could recognise that someone was talking.
+
+Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
+Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
+\jam{Should I show photodiode Viterbi results?}
+
+\begin{figure}%[H]
+	\includegraphics[width=0.8\textwidth]{butterworth_150_3000-cropped.pdf}
+	\caption{Butterworth bandpass filter frequency response, green vertical lines show the band limits of 150Hz and 3kHz.
+Note the very flat response within the band, any amplitude beyond -3dB is significant attenuation.}
+	\label{fig:butterworth}
+\end{figure}
+
+\subsection{logMMSE estimator}
+% logMMSE
+
+Speech enhancement for a noisy channel is a classic problem in signal processing.
+Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare various methods and find the best performing statistical technique to be the logMMSE (alt.\ MMSE-LSA) estimator which minimises the mean square error (MSE) of the log-spectral amplitude.
+This means that it takes the logarithm of the spectrum of the original signal and finds the fit (or estimate) that has the minimal MSE to that log-spectra.
+% Use of the logarithm is a common technique in speech processing as it, somewhat surprisingly, does not significantly change the intelligibility to the human ear.
+
+Applying an existing implementation of logMMSE to the optical microphone recordings produces dramatically cleaner results.
+The time series and spectrum from before and after the logMMSE filter is applied are shown in Figures~\ref{fig:logMMSE_timeseries}~\&~\ref{fig:logMMSE_spectrum}.
+All of the mains harmonics are significantly attenuated.
+And further high-passing removes the low frequency curve beneath 100Hz.
+
+There is varied success with passing music passed through the filters.
+Loud, rhythmic chords and drums are received well and remain recognisable, while thinner, airy flute and violin melodies struggle to be heard in the results.
+It is likely that the lower sounds have higher amplitude speaker oscillations and so pass through the optical microphone better.
+
+However, speech intelligibility remains very low with the voice sounding mumbled and indistinct.
+Enhancement so far has removed most of the background sound, as evidenced by the near silence between words, but has not clarified the speech.
+There is a question here as to where the diction in the speech is lost, at the speaker-mirror coupling, at the photodiode reading, or if it is still there, just obscured in the noisy signal.
+
+% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
+However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
+An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
+% % Visually it is impossible to confirm this for high frequencies, but it remains possible.
+
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{filter_timeseries_aa_melatos-cropped.pdf}
+	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) logMMSE filter is applied, only one second of the one minute recording is shown.
+The bump at the start of the time series is the result of filtering a finite signal, and is expected.}
+	\label{fig:logMMSE_timeseries}
+\end{figure}
+
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{filter_spectrum_aa_melatos-cropped.pdf}
+	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) logMMSE filter is applied, showing only up to 2kHz and a detail of the spectrum (limited to show only up to amplitude 2, while before filtering the 50Hz line has an amplitude of 40)}
+	\label{fig:logMMSE_spectrum}
+\end{figure}
+
+\subsection{Future work}
+% full model of transfer function
+A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
+Although we have the separation-intensity relation as linear for small motions, the other components of this full model are yet to be determined.
+
+Theoretically, if we had the correct model, then signal recovery would only be limited by the SNR. We could also use long recordings of white noise to estimate a matched filter for the system.
+
+For further speech enhancement, the optimal filter would involve a hidden-Markov-model trained to English phonemes. Alternatively, machine learning solutions also exist through-out the field.
+% SEGAN?
+
+% any other wild speculation?
 
@@
 % we are able to successfully ...
+Using a Michelson interferometer, we able to successfully demonstrate the Viterbi algorithm recovering an injected signal.
+Just as is done is continuous gravitational wave searches.
+We are also able to play-back chordal or rhythmic music from an optical microphone (using the same interferometer), while recordings of simple speech are audible but unintelligible.
 
@@
 
-\newpage
+\section{Open-source code}
+
+\jam{do I need to include access to sample .mp4 and .wav recordings: source, raw, and processed?}
+\jam{this should be neatened, make all .py and move all .ipynb in here}
+
+Current build found at:
+\url{https://github.com/daccordeon/gravexplain}
+
+\begin{verbatim}
+Directory structure:
+gravexplain/
+    source/
+        [py executable script versions of all .ipynb notebooks]
+        raspberry_pi_photodiode_adc.py
+            (reads photodiode signal through ADC by SPI pins, saves to .csv)
+    paper/
+        paper_main.tex
+        ifoDemoBib.bib
+        myunsrt.bst
+
+    mock_continuous_signal.ipynb
+        (simulates applying the Viterbi algorithm to a noisy signal)
+    experiment_viterbi.ipynb
+        (applies the Viterbi algorithm to an .mp4 from the webcam method)
+    experiment_photodiode.ipynb
+        (filters signal from optical microphone produced by raspberry_pi_photodiode_adc.py,
+        also applies the Viterbi algorithm)
+
+    tracker_time_series.ipynb
+        (defunct; saves signal and spectrum plots from a .csv file)
+    open_cv.ipynb
+        (defunct; various functions produced while learning the openCV library)
+
+    .gitignore
+    LICENSE
+    README.txt
+\end{verbatim}
+
+\section{Michelson interferometer schematic}
+\jam{please include}
+
+Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
+We did not find it necessary to include it to be able to demonstrate the results desired.
+
+% \newpage
 \section{Intensity relation derivation}
+\label{app:intensity_derivation}
+
+We want the change in intensity, $\delta I$, at a point in the Michelson interference pattern, given a small change, $\delta d$, in the perpendicular path difference of the mirrors (or in the separation of one and virtual image of the other).
+The point is at angle $\theta$ on the screen above the central ray, noting circular symmetry.
+At this point, two rays meet, one from each mirror, with the total electric field (\jam{parallel to the screen?}) being $E_1(t) + E_2(t)$.
+Given an initial separation of $d$, the path difference between the mirrors is $2 d \cos{\theta}$ as the longer ray must travel to and from the further mirror, and both move up at the angle $\theta$ to meet at the screen.
+
 \begin{align}
-    y_0(t) &= A \sin{(k L - \omega t)} \\
-    y_1(t) &= A \sin{(k (L + 2 d \cos{\theta}) - \omega t)} \\
-    Y(t) &= A \sin{(k L - \omega t)} + A \sin{(k (L + 2 d \cos{\theta}) - \omega t)} \\
-    \phi &= 2 k d \cos{\theta} \\
-    Y(t) &= A\; \Im\{e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)}\} \\
-    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \phi/2 - \omega t)} + e^{i (k L + \phi/2 + \phi/2 - \omega t)}\} \\    
-    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \omega t)} (e^{- i \phi/2} + e^{i \phi/2})\} \\    
-    Y(t) &= A\; \Im\{e^{i (k L + \phi/2 - \omega t)} (2 \cos{(\phi/2)})\} \\ 
-    Y(t) &= 2 A \cos{(\phi/2)}\; \Im\{e^{i (k L + \phi/2 - \omega t)}\} \\  
-    \therefore I(t) &= I_0 \cos^2{(\phi/2)}, \; \phi = \frac{4 \pi \cos{\theta}}{\lambda}\, d(t) \\
-    \frac{dI}{dt} &= \frac{dI}{d\phi}\; \frac{d\phi}{dt}\\
-    \frac{dI}{d\phi} &= - I_0 \cos{(\phi/2)} \sin{(\phi/2)}\\
-    &= - \frac{I_0}{2}\, \sin{(\phi)}\\
-    \delta\phi &= \frac{4 \pi \cos{\theta}}{\lambda}\, h\\
-    \therefore \delta I &= \frac{- 2 \pi I_0 \cos{\theta}}{\lambda}\,\sin{(\phi_0)}\, h
+\label{eq:intensity_derivation}
+    E_1(t) &= A e^{i (k L - \omega t)} \\
+    E_2(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \\
+    I(t) &= \lvert A e^{i (k L - \omega t)} + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2
 \end{align}
 
+Collapse the phase difference to $\phi$ and expand the norm squared by expanding the complex exponentials.
+From there is it trigonometric manipulation to reach the result of intensity in terms of separation.
+
+\begin{align}    
+    I(t) &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
+    I(t) &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2) \\
+    I(t) &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
+    I(d) &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+\end{align}
+
+Then take derivatives and find that the intensity changes linearly for small changes in separation.
+
+\begin{align}    
+    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
+    \frac{dI}{d\phi} &= - 2 A^2 \sin{\phi}\\
+    \delta\phi &= \frac{4 \pi \cos{\theta}}{\lambda}\, \delta d\\
+    \therefore \delta I &= \frac{- 8 \pi A^2 \cos{\theta} \sin{\phi}}{\lambda}\, \delta d
+\end{align}
+
+\section{The Viterbi Algorithm}
+\label{app:viterbi}
+
+To find the Viterbi path through a spectrogram (spectra over time), given an M x N grid of M frequency rows and N time columns and a rate limit for the change of frequency over time that we convert into the number of allowed rows we can move while stepping from each column to the next (this limit was $\pm 1$ row/column in our case).
+The path must start from the first column and end at the last, taking a single step in each column along the way.
+Each element in the grid is the norm of the Fourier component of a frequency at a time.
+These are normalised between $(0, 1)$ to use as multiplicative weights, i.e.\ the total weight of a path is the product of all weights along that path.
+In fact, we take the logarithm and so the sum of all weights to avoid underflow.
+The best path has the highest weight.
+
+With the grid prepared, the Viterbi algorithm starts in the second column.
+In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, taking the greatest cumulative weight of those available, ‘nearby’ cells (the value of which was discovered in the previous iteration).
+For each cell in the current column, the algorithm notes this weight plus the cell’s own weight, along with the row index of which cell in the previous column it came from.
+Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
+That best path is the Viterbi path.
+
+\section{Common changing frequency mistake}
+\label{app:phase_gotcha}
+
+We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
+The common mistake is to just appeal to the standard form $\sin{(2 \pi f t)}$, where f would be a constant frequency, and write down $\sin{(2 \pi f(t) t)}$.
+However, accumulation of phase means that this produces a curve that looks very different to that desired.
+Instead, one should see that the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
+Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
+
 \section{Circuit design}
+\label{app:circuit_diagram}
 % schematic of breadboard and connections to pi
@@
 
+Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}
+
 \begin{figure}%[H]
 	\includegraphics[width=\textwidth]{circuit_diagram.pdf}
-	\caption{Circuit diagram for photo-diode reading.
-Photo-diode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC), that’s finally read by the special purpose input (SPI) pins of a Raspberry Pi}
+	\caption{Circuit diagram for photodiode reading.
+Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC), that’s finally read by the special purpose input (SPI) pins of a Raspberry Pi}
 	\label{fig:circuit_diagram}

### 2020-02-05 source/raspberry_pi_photodiode_adc.py

@@ -1,0 +1,70 @@
+#!/usr/bin/python
+"""raspberry_pi_photodiode_adc.py
+James Gardner 2020
+
+reads ADC signal and saves to .csv
+"""
+
+import spidev
+import time
+#import os
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Define sensor channels
+light_channel = 0
+# Define delay between readings
+delay = 0.2
+
+# Open SPI bus
+spi = spidev.SpiDev()
+spi.open(0,0)
+spi.max_speed_hz=1000000
+
+# Function to read SPI data from MCP3008 chip
+# Channel must be an integer 0-7
+def read_channel(channel):
+    adc = spi.xfer2([1, (8 + channel) << 4, 0])
+    data = ((adc[1] & 3) << 8) + adc[2]
+    return data
+
+# Function to convert data to voltage level,
+# rounded to specified number of decimal places.
+def convert_volts(data, places):
+    volts = (3.3*data) / float(1023)
+    volts = round(volts, places)
+    return volts
+
+def live_plot():
+    delay = 1e-7
+    t0 = time.time()
+    x, y = [], []
+
+    while True:
+        t1 = time.time()
+        x.append(t1-t0)
+        y.append(read_channel(light_channel))
+        plt.plot(x, y, 'k')
+        plt.pause(delay)
+
+    plt.show()
+
+def save_stream(duration=1, out_file='adc_time_series.csv', show_plot=False):
+    with open(out_file, 'w') as f:
+        t0 = time.time()
+        while True:
+            t1 = time.time()
+            t = t1-t0
+            light_level = read_channel(light_channel)
+            f.write('{},{}\n'.format(t,light_level))
+            if t > duration:
+                break
+
+    if show_plot:
+        time_series = np.genfromtxt(out_file,delimiter=',')
+        print('resolution = {:.3f}Hz'.format(len(time_series)/duration))
+        plt.plot(time_series[:,0], time_series[:,1],'-k')
+        plt.show()
+
+if __name__ == '__main__':
+    save_stream(duration=1)

### 2020-02-06 paper/paper_main.tex
68th day of Summer

@@
 \begin{document}
+
+% cover page
+\section*{Project contributions}
+% contributions clarify what’s to be assessed
+Hannah Middleton: introduction of paper, design and making of interferometer
+
+James W. Gardner: all other paper sections, design and making of photodiode circuit, and all of the source code
+
+\jam{do we need to repeat those people in the Acknowledgements?}
+
+\newpage
 
@@
 % overview
-In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous waves to a general audience. 
+In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous gravitational waves to a general audience. 
 The techniques used are very similar to those used by researchers to search for continuous waves in LIGO and Virgo data.
-In section X we describe...... \han{add this in at the end}
+In Section~\ref{sec:ifo} we detail the interferometer set-up, Section~\ref{sec:single_tone} demonstrates observing a single tone, Section~\ref{sec:viterbi_wandering} details tracking a wandering frequency tone, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
 
@@
 \section{Table-top gravitational wave science}
+\label{sec:ifo}
 
@@
 Like two massive black-holes merging with each other.
-Even so, the effects they produce are on the order of one part in $10^{21}$ for our best detectors.
+Even so, the effects they produce are on the order of one part in $10^{21}$ for our best detectors\cite{GW150914}.
 Given the impossibility of such detection with a table-top experiment, we demonstrate the same techniques but for observing sound waves.
@@
 The beams return to interfere at a screen and produce a pattern of concentric, circular fringes.
+The configuration is shown in Figure~\ref{fig:ifo_schematic_webcam}.
 
-\jam{See appendix for the exact set-up, lenses are included to focus the beam and view the interference pattern ||| include Michelson schematic with speaker, webcam, and photodiode}.
+\begin{figure}
+	\includegraphics[width=0.8\textwidth]{ifo_schematic_webcam.pdf}
+	\caption{Michelson interferometer schematic, interference pattern is recorded by a webcam}
+	\label{fig:ifo_schematic_webcam}
+\end{figure}
+% Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
+We did not find it necessary to include it to be able to demonstrate the results desired.
 
@@
 	\includegraphics[width=\textwidth]{webcam_still0.pdf}
-	\caption{Interference pattern from Michelson interferometer as viewed by webcam from below (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured \jam{do I need to add a scale to this picture?}}
+	\caption{Interference pattern from Michelson interferometer as viewed by webcam from below (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured, the central fringe is approximately 5mm across}
 	\label{fig:interference_pattern}
@@
 
-\subsection{The webcam method}
+\section{Observing a single tone}
+\label{sec:single_tone}
+% the webcam method
 % how to analyse video, python OpenCV2
+
+\subsection{Method}
 
@@
 As such, any motion of the fringes, and so of the speaker, must be kept small.
+
+\subsection{Results}
 
@@
 
-\subsection{Tracking tones with the Viterbi algorithm}
+\section{Observing a wandering tone}
+\label{sec:viterbi_wandering}
+% Tracking tones with the Viterbi algorithm
 % explaination and successful application
 % lean on the continuous gravitational waves angle
+
+\jam{need to cite the various searches here}
 
@@
 
+\subsection{The Viterbi algorithm}
+
 The Viterbi algorithm is a path-finding algorithm that solves this particular problem of going from one side of a grid to the other.
@@
 
+\subsection{Results}
+
 To demonstrate this technique, we inject a tone that wanders in frequency over time, remembering to integrate up the phase change over time (see Appendix~\ref{app:phase_gotcha} for why).
@@
 Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, along with the Viterbi path and the injected signal.
-With great success, the Viterbi algorithm recovers the injected frequency over time, likely due to the high signal-to-noise ratio (SNR).
+To the eye, the Viterbi algorithm recovers the injected frequency over time, likely due to the high signal-to-noise ratio (SNR).
+Quantifying this, the Viterbi path stays within one frequency bin (approximately 0.3Hz) of the injected tone for 94\% of the run.
 Note that the signal initially moves too quickly (more than 0.11 Hz per second) for Viterbi to track, and that there appears to be an anomaly at 150s, likely a disturbance around the interferometer.
@@
 
-\jam{What more is there to say about Viterbi?}
-
-\section{Optical microphone}
+\section{Complex audio}
+\label{sec:optical_microphone}
+% Optical microphone
 % aim of optical microphone: play back injected sounds
@@
 As such, any frequency component of the analog signal above the Nyquist frequency of 8kHz would be aliased down into the detected range.
-To prevent this, an anti-aliasing Sallen-Key filter tuned to 16kHz is included before the ADC.
-This component attenuates any frequencies above 8kHz, before they get digitally sampled.
+To prevent this, an anti-aliasing Sallen-Key filter tuned to 16kHz is included before the ADC. This component attenuates any frequencies above 8kHz, before they get digitally sampled.
 It’s not immediately clear what’s limiting the sample rate to 16kHz, but it is likely non-optimal reading of the ADC by the Pi script.
@@
 
-The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}.
+The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}. The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo}.
 
-\subsection{Initial recordings}
+\begin{figure}
+	\includegraphics[width=0.8\textwidth]{ifo_schematic_photodiode.pdf}
+	\caption{Michelson interferometer schematic, a part of the interference pattern is recorded by a photodiode}
+	\label{fig:ifo_schematic_podo}
+\end{figure}
+
+\subsection{Initial results}
 % breakdown of all the filters used along the way
@@
 For each observation, the audio was played through the speaker with minimal activity around interferometer as the Pi script recorded the signal.
-The saved time series was then directly converted to a .wav file and played as an audio recording.
+The saved time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{scipy.io.wavfile.write}\cite{scipy} implementation in Python\cite{python}.
 
@@
 \end{figure}
+
+\subsection{Filters applied}
 
@@
 
-Applying an existing implementation of logMMSE to the optical microphone recordings produces dramatically cleaner results.
+Applying an existing implementation of logMMSE\cite{logmmse} to the optical microphone recordings produces dramatically cleaner results.
 The time series and spectrum from before and after the logMMSE filter is applied are shown in Figures~\ref{fig:logMMSE_timeseries}~\&~\ref{fig:logMMSE_spectrum}.
@@
 
-\subsection{Future work}
-% full model of transfer function
+\section{Future work}
+
 A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
@@
 
+\jam{should this be more broad a future work section?}
+
 \section{Conclusions}
-% we are able to successfully ...
+
 Using a Michelson interferometer, we able to successfully demonstrate the Viterbi algorithm recovering an injected signal.
@@
 
+\newpage
 \appendix
@@
 
-\jam{do I need to include access to sample .mp4 and .wav recordings: source, raw, and processed?}
-\jam{this should be neatened, make all .py and move all .ipynb in here}
+\jam{update this to the current README.txt}
+
+This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.
 
@@
 \end{verbatim}
-
-\section{Michelson interferometer schematic}
-\jam{please include}
-
-Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
-We did not find it necessary to include it to be able to demonstrate the results desired.
 
@@
 \begin{figure}%[H]
-	\includegraphics[width=\textwidth]{circuit_diagram.pdf}
+	\includegraphics[width=\textwidth]{circuit_diagram_2.pdf}
 	\caption{Circuit diagram for photodiode reading.

### 2020-02-07 paper/paper_main.tex
69th day of Summer

@@
 \begin{figure}
-	\includegraphics[width=0.8\textwidth]{ifo_schematic_webcam.pdf}
+	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_webcam.pdf}
 	\caption{Michelson interferometer schematic, interference pattern is recorded by a webcam}
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{webcam_still0.pdf}
+	\includegraphics[width=\textwidth]{figures/webcam_still0.pdf}
 	\caption{Interference pattern from Michelson interferometer as viewed by webcam from below (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured, the central fringe is approximately 5mm across}
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{webcam_expt_4_0209-cropped.pdf}
+	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
 	\caption{Spectrum of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz hitting the interferometer, injected fundamental frequency and harmonics thereof clearly present in spectrum}
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{expt_overlay_2_viterbi_test_webcam.pdf}
+	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
 	\caption{Spectrogram of observed signal overlaid with white crosses for the Viterbi path and pink dots for the injected signal, notice the high correlation between them indicating that the signal is able to be recovered; also note the deviation at the start when the signal moves too quickly and just before the background anomaly at 150s}
@@
 \begin{figure}
-	\includegraphics[width=0.8\textwidth]{ifo_schematic_photodiode.pdf}
+	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_photodiode.pdf}
 	\caption{Michelson interferometer schematic, a part of the interference pattern is recorded by a photodiode}
@@
 \begin{figure}%[H]
-	\includegraphics[width=\textwidth]{podo_noise_psd_zoom-cropped.pdf}
+	\includegraphics[width=\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
 	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), show high power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the ambient lighting), otherwise is a fairly white spectrum except for the broad, shallow maximum at 750Hz}
@@
 \begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{butterworth_150_3000-cropped.pdf}
+	\includegraphics[width=0.8\textwidth]{figures/butterworth_150_3000-cropped.pdf}
 	\caption{Butterworth bandpass filter frequency response, green vertical lines show the band limits of 150Hz and 3kHz.
@@
 \begin{figure}%[H]
-	\includegraphics[width=\textwidth]{filter_timeseries_aa_melatos-cropped.pdf}
+	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
 	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) logMMSE filter is applied, only one second of the one minute recording is shown.
@@
 \begin{figure}%[H]
-	\includegraphics[width=\textwidth]{filter_spectrum_aa_melatos-cropped.pdf}
+	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
 	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) logMMSE filter is applied, showing only up to 2kHz and a detail of the spectrum (limited to show only up to amplitude 2, while before filtering the 50Hz line has an amplitude of 40)}
@@
 \begin{figure}%[H]
-	\includegraphics[width=\textwidth]{circuit_diagram_2.pdf}
+	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
 	\caption{Circuit diagram for photodiode reading.

### 2020-02-07 source/experiment_viterbi.py

@@ -1,232 +1,0 @@
-#!/usr/bin/env python3
-"""experiment_viterbi.py
-James Gardner 2019
-ANU / Melbourne Uni
-
-applies viterbi analysis pipeline to experiment video of
-michelson interferometer with a mirror driven at a changing frequency
-takes signal as time series of intensity at the video's centre
-"""
-
-import cv2
-import numpy as np
-import matplotlib.pyplot as plt
-from mpl_toolkits.axes_grid1 import make_axes_locatable
-from tqdm import tqdm_notebook as tqdm
-
-def fourier_spectrum(signal, fps, return_spectrum=False,
-                     produce_plot=False, out_plot_name='tmp.pdf', out_plot_title=''):
-    """finds fourier spectrum of signal time series as numpy array,
-    has functionality to return and/or plot the sectrum (both default off),
-    built from code originally found in tracker_time_series.ipynb
-    """   
-    signal_frames = len(signal)
-    # will drop two frames later, fps: frames per second
-    total_time = (signal_frames-2)/fps
-    t = np.linspace(0,total_time,signal_frames)
-    dt = t[1] - t[0]
-
-    yf = np.fft.fft(signal)
-    # normalised-absolute value of FT'd signal
-    nrm_abs_yf = 2/signal_frames*np.abs(yf)
-    # values at the centre of each frequency bin
-    freq_scale = np.fft.fftfreq(len(yf),dt)
-    # real signals are symmetric about 0 in frequency domain
-    freq_scale_positive = freq_scale[:signal_frames//2]
-    # frequency distribution values on positive side
-    freq_prob = nrm_abs_yf[:signal_frames//2]
-    
-    if produce_plot:
-        fig, (ax0,ax1) = plt.subplots(2,figsize=(14,14))
-        ax0.plot(t,signal)
-        ax0.set(title='signal: {}'.format(out_plot_title),ylabel='signal strength',xlabel='time, t')
-        # signal average value gives magnitude of frequency = 0 term
-        # simple fix is to drop first two bins, otherwise need to shift signal
-        ax1.plot(freq_scale_positive[2:],freq_prob[2:])
-        ax1.set(title='discrete FFT',ylabel='freq strength in signal',xlabel='frequency, f')
-        plt.savefig(out_plot_name,bbox_inches='tight')
-        plt.clf()
-    
-    if return_spectrum:
-        return freq_prob[2:], freq_scale_positive[2:]
-    
[…]

### 2020-02-07 source/main.py

@@ -1,0 +1,24 @@
+#!/usr/bin/env python3
+"""main.py
+James Gardner 2020
+ANU / Melbourne Uni
+
+performs entire analysis pipeline if executed
+i.e. run the following command in terminal:
+python3 main.py
+with the following files present in cwd:
+expt_viterbi_test_webcam.mp4 expt_4_0209.mp4 aa_melatos.csv aa_jam_track.csv
+this will produce the respective plots used in the GravExplain paper
+"""
+
+import viterbi_video_analysis
+import optical_microphone_filters
+
+if __name__ == "__main__":
+    viterbi_video_analysis.webcam_video_analysis('expt_4_0209.mp4', point=(230,240))
+    viterbi_video_analysis.PointViterbi('expt_viterbi_test_webcam.mp4', scanning_range=1).plot()
+    
+    optical_microphone_filters.psd_plot()
+    optical_microphone_filters.butter_filter_plot()
+    optical_microphone_filters.logmmse_filter('aa_melatos.csv')
+    optical_microphone_filters.logmmse_filter('aa_jam_track.csv')    

### 2020-02-07 source/optical_microphone_filters.py

@@ -1,0 +1,700 @@
+#!/usr/bin/env python3
+"""experiment_viterbi.py
+James Gardner 2020
+ANU / Melbourne Uni
+
+various assorted pieces of analysis of data from photodiode
+reading the changing pattern of the optical microphone interferometer
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+import scipy.io.wavfile as wavfile
+import scipy.signal as ssignal
+from logmmse import logmmse
+
+def fourier_spectrum_2(signal, fps, return_spectrum=False, cutter=2, remove_mains=False,
+                     produce_plot=False, out_plot_name='tmp.pdf', out_plot_title=''):
+    """finds fourier spectrum of signal time series as numpy array,
+    has functionality to return and/or plot the sectrum (both default off),
+    built from code originally found in tracker_time_series.ipynb
+    
+    note: based off of fourier_spectrum, careful with older scripts
+    """   
+    signal_frames = len(signal)
+    
+    # will drop two frames later, fps: frames per second
+    total_time = (signal_frames-cutter)/fps
+    t = np.linspace(0,total_time,signal_frames)
+    dt = t[1] - t[0]
+
+    yf = np.fft.fft(signal)
+    # normalised-absolute value of FT'd signal
+    nrm_abs_yf = 2/signal_frames*np.abs(yf)
+    # values at the centre of each frequency bin
+    freq_scale = np.fft.fftfreq(len(yf),dt)
+    # real signals are symmetric about 0 in frequency domain
+    freq_scale_positive = freq_scale[cutter:signal_frames//2]
+    # frequency distribution values on positive side
+    freq_prob = nrm_abs_yf[cutter:signal_frames//2]
+    
+    # freqscale 0, +inf, -inf, -0
+    #print(freq_scale[:10], freq_prob[:10])#,freq_scale[-5])
+    
+    if produce_plot:
+        
+        # tune out 50Hz mains noise
+        # closest value in scale is index 5
+        #to_go = None#(5,)#tuple(range(10))#(5,)
+        # mains_mark
+        to_go = ()
+        
+        if remove_mains:
+            mm_0, mm_1 = None, None
+            for i, f in enumerate(freq_scale_positive):
+                if f > 50:
+                    mm_0, mm_1 = i-1, i
+                    break
[…]

### 2020-02-07 source/viterbi_video_analysis.py

@@ -1,0 +1,323 @@
+#!/usr/bin/env python3
+"""experiment_viterbi.py
+James Gardner 2020
+ANU / Melbourne Uni
+
+applies viterbi analysis pipeline to experiment video of
+michelson interferometer with a mirror driven at a changing frequency
+takes signal as time series of intensity at the video's centre
+"""
+
+import cv2
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+def fourier_spectrum(signal, fps, return_spectrum=False,
+                     produce_plot=False, out_plot_name='tmp.pdf', out_plot_title=''):
+    """finds fourier spectrum of signal time series as numpy array,
+    has functionality to return and/or plot the sectrum (both default off),
+    built from code originally found in tracker_time_series.ipynb
+    """   
+    signal_frames = len(signal)
+    # will drop two frames later, fps: frames per second
+    total_time = (signal_frames-2)/fps
+    t = np.linspace(0,total_time,signal_frames)
+    dt = t[1] - t[0]
+
+    yf = np.fft.fft(signal)
+    # normalised-absolute value of FT'd signal
+    nrm_abs_yf = 2/signal_frames*np.abs(yf)
+    # values at the centre of each frequency bin
+    freq_scale = np.fft.fftfreq(len(yf),dt)
+    # real signals are symmetric about 0 in frequency domain
+    freq_scale_positive = freq_scale[:signal_frames//2]
+    # frequency distribution values on positive side
+    freq_prob = nrm_abs_yf[:signal_frames//2]
+    
+    if produce_plot:
+        fig, (ax0,ax1) = plt.subplots(2,figsize=(14,14))
+        ax0.plot(t,signal)
+        ax0.set(title='signal: {}'.format(out_plot_title),ylabel='signal strength',xlabel='time, t')
+        # signal average value gives magnitude of frequency = 0 term
+        # simple fix is to drop first two bins, otherwise need to shift signal
+        ax1.plot(freq_scale_positive[2:],freq_prob[2:])
+        ax1.set(title='discrete FFT',ylabel='freq strength in signal',xlabel='frequency, f')
+        plt.savefig(out_plot_name,bbox_inches='tight')
+        plt.clf()
+    
+    if return_spectrum:
+        return freq_prob[2:], freq_scale_positive[2:]
+
+def series_at_point(filename, point=None, return_series=False,
+                    produce_plot=False, out_plot_name='tmp.pdf'):
[…]

### 2020-02-12 paper/paper_main.tex
74th day of Summer

@@
 % contributions clarify what’s to be assessed
-Hannah Middleton: introduction of paper, design and making of interferometer
+James W. Gardner: all paper sections other than the introduction, design and making of the photodiode circuit, and all of the source code at \url{https://github.com/daccordeon/gravexplain}
 
-James W. Gardner: all other paper sections, design and making of photodiode circuit, and all of the source code
+Hannah Middleton: introduction of paper, design and making of the interferometer
 
-\jam{do we need to repeat those people in the Acknowledgements?}
+Andrew Melatos, Robin Evans, William Moran: vital advice and comments throughout the project
 
@@
 
-\author{Hannah Middleton}
+\author{Hannah Middleton$^{(i)}$}
 \email{hannah.middleton@unimelb.edu.au}
-\author{Andrew Melatos}
+\author{Andrew Melatos$^{(i)}$}
 \email{amelatos@unimelb.edu.au}
-\affiliation{School of Physics, University of Melbourne, Parkville, VIC, 3010, Australia}
+\author{Robin Evans$^{(ii)}$}
+\email{robinje@unimelb.edu.au}
+\author{William Moran$^{(ii)}$}
+\email{wmoran@unimelb.edu.au}
+\affiliation{School of Physics$^{(i)}$ and Department of Electrical and Electronic Engineering$^{(ii)}$, University of Melbourne, Parkville, VIC, 3010, Australia}
 \affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}
@@
 
-This equation also means that the intensity changes linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
+This equation means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
+This makes the problem of extracting the original signal more difficult.
 
@@
 \label{eq:intensity_change}
-	\delta I = \frac{- 8 \pi A^2 \cos{\theta} \sin{\phi}}{\lambda}\, \delta d
+	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
 \end{equation}
@@
 
-The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}. The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo}.
+The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}.
+The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the experiment, showing the interferometer, the circuit, and the Pi, is shown in Figure~\ref{fig:setup_pic2}.
 
@@
 	\label{fig:ifo_schematic_podo}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.8\textwidth]{figures/setup_pic2.pdf}
+	\caption{Picture taken of the optical microphone set-up, showing the Michelson interferometer, the photodiode (in an insert), the circuit, and the Raspberry Pi.}
+	\label{fig:setup_pic2}
 \end{figure}
@@
 A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-Although we have the separation-intensity relation as linear for small motions, the other components of this full model are yet to be determined.
+Although we have the separation-intensity relation as non-linear for small motions, the other components of this full model are yet to be determined.
 
@@
 \section{Open-source code}
-
-\jam{update this to the current README.txt}
-
 This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.
 
-Current build found at:
+The current build can be found at:
 \url{https://github.com/daccordeon/gravexplain}
-
-\begin{verbatim}
-Directory structure:
-gravexplain/
-    source/
-        [py executable script versions of all .ipynb notebooks]
-        raspberry_pi_photodiode_adc.py
-            (reads photodiode signal through ADC by SPI pins, saves to .csv)
-    paper/
-        paper_main.tex
-        ifoDemoBib.bib
-        myunsrt.bst
-
-    mock_continuous_signal.ipynb
-        (simulates applying the Viterbi algorithm to a noisy signal)
-    experiment_viterbi.ipynb
-        (applies the Viterbi algorithm to an .mp4 from the webcam method)
-    experiment_photodiode.ipynb
-        (filters signal from optical microphone produced by raspberry_pi_photodiode_adc.py,
-        also applies the Viterbi algorithm)
-
-    tracker_time_series.ipynb
-        (defunct; saves signal and spectrum plots from a .csv file)
-    open_cv.ipynb
-        (defunct; various functions produced while learning the openCV library)
-
-    .gitignore
-    LICENSE
-    README.txt
-\end{verbatim}
+% \begin{verbatim}
+% README
+% \end{verbatim}
 
@@
 
-Then take derivatives and find that the intensity changes linearly for small changes in separation.
+Then take derivatives and find that the intensity change is non-linear.
 
@@
     \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
-    \frac{dI}{d\phi} &= - 2 A^2 \sin{\phi}\\
-    \delta\phi &= \frac{4 \pi \cos{\theta}}{\lambda}\, \delta d\\
-    \therefore \delta I &= \frac{- 8 \pi A^2 \cos{\theta} \sin{\phi}}{\lambda}\, \delta d
+    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
+    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
+    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
 \end{align}
@@
 \begin{acknowledgments}
-The authors are grateful to Robin Evans, Bill Moran, Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
 

### 2020-02-12 source/main.py
74th day of Summer

@@ -11,14 +11,23 @@ expt_viterbi_test_webcam.mp4 expt_4_0209.mp4 aa_melatos.csv aa_jam_track.csv
 this will produce the respective plots used in the GravExplain paper
 """
 
+import os.path
 import viterbi_video_analysis
 import optical_microphone_filters
 
 if __name__ == "__main__":
-    viterbi_video_analysis.webcam_video_analysis('expt_4_0209.mp4', point=(230,240))
-    viterbi_video_analysis.PointViterbi('expt_viterbi_test_webcam.mp4', scanning_range=1).plot()
     
-    optical_microphone_filters.psd_plot()
-    optical_microphone_filters.butter_filter_plot()
-    optical_microphone_filters.logmmse_filter('aa_melatos.csv')
-    optical_microphone_filters.logmmse_filter('aa_jam_track.csv')    
+    if (os.path.isfile('expt_4_0209.mp4') and os.path.isfile('expt_viterbi_test_webcam.mp4')
+        and os.path.isfile('aa_melatos.csv') and os.path.isfile('aa_jam_track.csv')):
+        print('Welcome to GravExplain main.py, all sample files are present, beginning analysis now')
+        
+        viterbi_video_analysis.webcam_video_analysis('expt_4_0209.mp4', point=(230,240))
+        viterbi_video_analysis.PointViterbi('expt_viterbi_test_webcam.mp4', scanning_range=1).plot()
+
+        optical_microphone_filters.psd_plot()
+        optical_microphone_filters.butter_filter_plot()
+        optical_microphone_filters.logmmse_filter('aa_melatos.csv')
+        optical_microphone_filters.logmmse_filter('aa_jam_track.csv')           
+        
+    else:
+        print ('Welcome to GravExplain main.py, some or all of the sample files are missing, please consult README.txt')     

### 2020-02-12 source/optical_microphone_filters.py

@@ -3,8 +3,8 @@
 James Gardner 2020
 ANU / Melbourne Uni
 
-various assorted pieces of analysis of data from photodiode
-reading the changing pattern of the optical microphone interferometer
+analysis of data from photodiode reading the pattern of
+the optical microphone interferometer
 """
 
 import numpy as np
@@ -19,8 +19,6 @@ def fourier_spectrum_2(signal, fps, return_spectrum=False, cutter=2, remove_main
     """finds fourier spectrum of signal time series as numpy array,
     has functionality to return and/or plot the sectrum (both default off),
     built from code originally found in tracker_time_series.ipynb
-    
-    note: based off of fourier_spectrum, careful with older scripts
     """   
     signal_frames = len(signal)
     
@@ -97,8 +95,7 @@ def fourier_spectrum_2(signal, fps, return_spectrum=False, cutter=2, remove_main
     
 def cut_mains(signal, fps, cutter=2):
     """returns signal with the 50Hz mains cut out
-    
-    NB: this is a bad signal processing technique and should not be used
+    NB: this is a bad signal processing technique and should not be used!
     """
     signal_frames = len(signal)
     
@@ -150,8 +147,6 @@ def viterbi_pathfinder(grid, scanning_range=3):
     """find the highest scoring path through the grid, left-to-right,
     as by the viterbi algorithm, with connections plus-minus the scanning_range;
     returns score grid for best path to each node and a bitmap of the total best path
-    
-    NB: this is ripped from previous scripts
     """
     # normalised grid, algorithm goal is to maximise product of values
     ngrid  = grid/np.max(grid)
@@ -210,14 +205,13 @@ def viterbi_pathfinder(grid, scanning_range=3):
     return score_grid, path_grid
 
 def absmax(axis):
-    """returns the maximum absolute value along an np.axis"""
+    """returns the maximum absolute value along an axis of a np.array"""
     return max(abs(np.min(axis)), abs(np.max(axis)))
 
 def photodiode_experiment_viterbi(filename='podo_viterbi_test.csv',
                                   filetag='viterbi_test'):
     """performs viterbi analysis on window of .wav recording from photodiode"""
     # read in photodiode (aka podo) data as time series
-
     mega = np.genfromtxt(filename,delimiter=',')
     times = mega[:,0]
[…]

### 2020-02-12 source/raspberry_pi_photodiode_adc.py

@@ -1,41 +1,46 @@
 #!/usr/bin/python
 """raspberry_pi_photodiode_adc.py
 James Gardner 2020
+built from existing scripts online since lost,
+gratitude to those original authors
 
-reads ADC signal and saves to .csv
+to be run on a Rasperry Pi,
+reads the signal from an ADC and saves to a .csv file
 """
 
 import spidev
 import time
-#import os
 import numpy as np
 import matplotlib.pyplot as plt
 
-# Define sensor channels
+# define sensor channels
 light_channel = 0
-# Define delay between readings
+# define delay between readings
 delay = 0.2
 
-# Open SPI bus
+# open SPI bus
 spi = spidev.SpiDev()
 spi.open(0,0)
 spi.max_speed_hz=1000000
 
-# Function to read SPI data from MCP3008 chip
-# Channel must be an integer 0-7
 def read_channel(channel):
+    """returns a single value from a channel of the ADC,
+    takes SPI data from MCP3008 chip,
+    channel must be an integer 0-7"""
     adc = spi.xfer2([1, (8 + channel) << 4, 0])
     data = ((adc[1] & 3) << 8) + adc[2]
     return data
 
-# Function to convert data to voltage level,
-# rounded to specified number of decimal places.
 def convert_volts(data, places):
+    """function to convert data to voltage level,
+    rounded to specified number of decimal places"""
     volts = (3.3*data) / float(1023)
     volts = round(volts, places)
     return volts
 
 def live_plot():
+    """displays a live plot of the channel reading,
+    can get very slow if left for too long"""
     delay = 1e-7
     t0 = time.time()
     x, y = [], []
@@ -50,6 +55,9 @@ def live_plot():
     plt.show()
 
 def save_stream(duration=1, out_file='adc_time_series.csv', show_plot=False):
+    """saves a data stream from the ADC for the given duration to a .csv files,
+    sample rate is just above 16kHz,
+    show_plot is not recommend due to the large number of points"""
     with open(out_file, 'w') as f:
         t0 = time.time()
         while True:
@@ -67,4 +75,9 @@ def save_stream(duration=1, out_file='adc_time_series.csv', show_plot=False):
         plt.show()
 
 if __name__ == '__main__':
-    save_stream(duration=1)
+    print('Welcome to raspberry_pi_photodiode_adc.py, attempting to save a one second stream now')
+    try:
+        save_stream(duration=1)
+    except:
[…]

### 2020-02-12 source/viterbi_video_analysis.py

@@ -3,7 +3,7 @@
 James Gardner 2020
 ANU / Melbourne Uni
 
-applies viterbi analysis pipeline to experiment video of
+applies viterbi analysis pipeline to experiment video (as .mp4) of
 michelson interferometer with a mirror driven at a changing frequency
 takes signal as time series of intensity at the video's centre
 """
@@ -151,16 +151,11 @@ def inj_wandering_signal(duration=300, fps=16000,
                          meander_amp=9, meander_decay=0.01, meander_freq=0.005,                     
                          filetag='webcam', return_freq_series=False, save_to_csv=False,
                          save_wav_recording=False, plot_wandering_signal = False):
-    """create injected meandering signal to test viteri analysis
-    # webcam options
-    meander_amp, meander_decay, meander_freq = 9, 0.01, 0.005
-    filetag = 'webcam'
-
-    # podo options (podo=photodiode)
+    """creates wandering frequency, FM-modulated signal to test viteri analysis,
+    default arguments are for testing the webcam, else see below for photodiode arguments:
     meander_amp, meander_decay, meander_freq = 300, 0.01, 0.01
     filetag = 'podo'  
     """
-
     # 5 minutes duration
     long_timesteps = fps*duration
 
@@ -287,7 +282,9 @@ class PointViterbi(object):
             plt.clf()
 
 def webcam_video_analysis(filename, point=None):
-    """filename must be a .mp4, saves the opening still and a spectrum of a point in the video"""
+    """saves the opening still of video and a spectrum of a point in the video,
+    filename must be an .mp4
+    """
     cap = cv2.VideoCapture(filename)
     ret, frame = cap.read()
     cv2.imwrite('tmp_webcam_still.png',frame)

### 2020-02-12 paper/paper_main.tex
74th day of Summer

@@
 % contributions clarify what’s to be assessed
-James W. Gardner: all paper sections other than the introduction, design and making of the photodiode circuit, and all of the source code at \url{https://github.com/daccordeon/gravexplain}
+James W. Gardner:
+\begin{itemize}
+\item all paper sections other than the introduction, i.e.\ from Section~\ref{sec:ifo} onwards, and all figures in the paper
+\item addition of the injection speaker, design of photodiode circuit from standard examples, assembly of circuit on breadboard, and set-up of Raspberry Pi to measure output
+\item all of the source code at \url{https://github.com/daccordeon/gravexplain}, that reads the data from the Raspberry Pi, analyses it, and creates the figures used in the paper
+\item talks presented were based on paper and figures therein
+\end{itemize}
 
-Hannah Middleton: introduction of paper, design and making of the interferometer
+Hannah Middleton:
+\begin{itemize}
+\item introduction of paper, i.e.\ Section~\ref{sec:introduction}
+\item design and assembly of the interferometer
+\end{itemize}
 
 Andrew Melatos, Robin Evans, William Moran: vital advice and comments throughout the project
+
+\subsection*{Statement of achievements}
+% one paragraph recount of what’s been done for two months, like the abstract
 
@@
 Continuous gravitational wave searches use sophisticated statistical techniques to search for signals in noisy data.
-We demonstrate some of these techniques with a table-top Michelson interferometer, namely the Viterbi algorithm for recovering injected tones and a signal processing filter pipeline.
-We also use the same experiment to demonstrate an optical microphone capable of playing back simple recordings of injected audio.
+We demonstrate some of these techniques with a table-top Michelson interferometer, namely the Viterbi algorithm for recovering a tone with wandering frequency.
+We also use the same experiment to demonstrate an optical microphone capable of playing back simple audio recordings of injected music.
+However, played-back speech is currently unintelligible so we investigate preliminary methods of speech enhancement.
+
+\jam{Andrew says that this should be a bit more detailed!}
 
@@
 
-% cover page that attributes all contributions
-
 \section{Introduction}
+\label{sec:introduction}
 
@@
 In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous gravitational waves to a general audience. 
-The techniques used are very similar to those used by researchers to search for continuous waves in LIGO and Virgo data.
+The techniques used are very similar to those used to search for continuous waves in LIGO and Virgo data.
 In Section~\ref{sec:ifo} we detail the interferometer set-up, Section~\ref{sec:single_tone} demonstrates observing a single tone, Section~\ref{sec:viterbi_wandering} details tracking a wandering frequency tone, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
+We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
 
@@
 
-Detectable gravitational waves are caused by the most violent events in the universe.
-Like two massive black-holes merging with each other.
+Detectable gravitational waves are caused by some of the most violent events in the universe.
+Like two massive black-holes merging with each other\cite{GW150914}.
 Even so, the effects they produce are on the order of one part in $10^{21}$ for our best detectors\cite{GW150914}.
-Given the impossibility of such detection with a table-top experiment, we demonstrate the same techniques but for observing sound waves.
+For a table-top experiment these effects would be far too small, so we demonstrate observing a far louder source, sound waves.
+Although the analogue to gravitational waves is tenuous, we can still demonstrate some of the continuous wave analysis satisfactorily.
+The other benefit of observing sound is that we can use existing signal processing techniques designed for audio signals.
 
@@
 A Michelson interferometer is a well-known optics configuration used in many undergraduate and research labs, including in the detectors for gravitational waves.
-A laser beam of coherent (monochromatic and equal phase) light is split into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
-The beams return to interfere at a screen and produce a pattern of concentric, circular fringes.
+A laser beam of coherent (monochromatic and equal phase) light is split evenly by a 50/50 beamsplitter into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
+The beams recombine and interfere at the beamsplitter before hitting a screen to produce a pattern of concentric, circular fringes.
 The configuration is shown in Figure~\ref{fig:ifo_schematic_webcam}.
+In the detectors for gravitational waves, these arms are kilometres long and many other complexities (such as power recycling and quantum squeezing)\cite{GW150914} are necessary to observe a signal.
 
@@
 
-The interference pattern, as shown in Figure~\ref{fig:interference_pattern}, is due to the path difference between the two arms resulting in a phase difference between the two beams when they meet at the screen.
+The interference pattern, as shown in Figure~\ref{fig:interference_pattern}, is due to the path difference between the two arms of unequal lengths (if the lengths were equal, then only a single, bright dot would be seen).
+This results in a phase difference between the two beams when they meet at the screen.
 This path difference is perhaps better seen as the perpendicular separation between one mirror and the virtual image of the other.
 For an aligned configuration and a fixed separation, the phase difference at a point only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-The intensity, $I$, of this pattern at a particular angle is given by Equation~\ref{eq:intensity_result} below.
-Where A is the amplitude of the electric field of the beam, d is the separation (or the difference in the arm lengths), $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Appendix~\ref{app:intensity_derivation} for the derivation to get to this equation).
-\begin{equation}
-\label{eq:intensity_result}
-I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{equation}
-
-This equation means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
-This makes the problem of extracting the original signal more difficult.
-
-\begin{equation}
-\label{eq:intensity_change}
-	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{equation}
 
@@
 	\includegraphics[width=\textwidth]{figures/webcam_still0.pdf}
-	\caption{Interference pattern from Michelson interferometer as viewed by webcam from below (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured, the central fringe is approximately 5mm across}
+	\caption{Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
+For scale, the central fringe is approximately 5mm across}
 	\label{fig:interference_pattern}
@@
 
+Continuous gravitational wave searches are, by the nature of the astrophysical sources that they look for, always searching for periodic, or even sinusoidal, signals.
+Often the frequency of this periodicity can change, as addressed later, but sometimes it is (near) constant.
+As such, the audio analogue is to observe a single tone (a sinusoidal note of one frequency).
+
 \subsection{Method}
 
-We demonstrate a signal source with a 0.5W speaker stuck to the back of one of the mirrors.
-This speaker was obtained from a standard commercial 3.5mm jack input, USB powered pair of speakers.
-Alternatively, a speaker can be driven by a function generator and power plugs.
-As the speaker deforms to play something, the mirror moves with a change in separation, $\delta d$, as some function of the amplitude and frequency of the sound.
+We use a 0.5W speaker stuck to the back of one of the mirrors as a source for our demonstration, as shown in Figure~\ref{fig:ifo_schematic_webcam}.
+This speaker was part of a commercial 3.5mm-jack, USB powered pair of speakers and is driven by a PC.
+Previously, we used a function generator to accomplish the same signal.
+As the speaker deforms to play sound, the mirror moves with a change in separation, $\delta d$, as some coupling function of the amplitude and frequency of the sound.
 The motion of the fringes follows the motion of the mirror and therefore the speaker at the same frequency.
@@
 Large fringe motion means that multiple fringes can pass over the measured point in a single speaker deformation, such over-counting would artificially raise the measured frequency.
-As such, any motion of the fringes, and so of the speaker, must be kept small.
+As such, any motion of the fringes, and so of the speaker, must be kept small by playing sound softly through the speaker.
 
@@
 
-We measure the intensity at a point on the screen with a standard commercial USB webcam, sampling at 30Hz (or 30 fps).
+We measure the intensity at a point on the screen with a commercial USB webcam placed as to view the screen (Figure~\ref{fig:ifo_schematic_webcam}), sampling at 30Hz (or 30 fps, frames-per-second).
 This limits the spectral content of signals we can observe to be under the Nyquist frequency of 15Hz.
-This isn’t an issue to demonstrate what we want, but see later Section~\ref{sec:photodiode} to raise this limit to audible frequencies.
+See later Section~\ref{sec:photodiode} where we raise this limit into to audible frequencies (20 Hz to 20kHz) with a photodiode.
 While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
-Then, at an off-centre pixel as seen in Figure~\ref{fig:interference_pattern}, we take the green-channel of the RGB video as approximate to the total intensity, as the laser is monochromatic green light.
-We can play this time series of intensity over time as an audio recording, and the tone is clearly heard.
-Taking the Fourier transform of this signal also strongly shows the injected tone at the correct frequency, as shown in Figure~\ref{fig:webcam_spectrum}.
-Therefore, we’re confident that we’ll be able to track input frequencies.
+Then, at an off-centre pixel as seen in Figure~\ref{fig:interference_pattern}, we take the green-channel of the colour (RGB) video as approximate to the total intensity, as the laser is monochromatic green light.
+We can play this time series of intensity over time as an audio recording (we used the \textbf{scipy.io.wavfile.write}\cite{scipy} function in Python\cite{python}) and the tone is clearly heard.
+Taking the Fourier transform of this signal shows the injected tone, as seen in Figure~\ref{fig:webcam_spectrum} where we record the highest peak Fourier amplitude at $(2.099\pm 0.008)\mathrm{Hz}$ for an injected tone of 2.09Hz (a result within two standard deviations).
+Therefore, we’re able to observe a single tone quite confidently.
 
@@
 	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Spectrum of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz hitting the interferometer, injected fundamental frequency and harmonics thereof clearly present in spectrum}
+	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the highest peak is measured at $(2.099\pm 0.008) \mathrm{Hz}$ for an injected tone of 2.09Hz.
+In fact, the actual highest peak is at 0Hz, but that corresponds only to a non-zero average value, so we don’t display or consider it.}
 	\label{fig:webcam_spectrum}
@@
 
-\jam{need to cite the various searches here}
+Continuous gravitational wave searches often look for signals that wander (change) in frequency slowly over time\cite{2019PhRvD.100l2002A}.
+This means that finding the signal isn’t as simple as searching for a peak in the frequency content for the full observing run, especially when the signal-to-noise ratio is particularly low.
 
-Continuous gravitational wave searches look for signals that change frequency slowly over time.
-This means that finding the signal isn’t as simple as searching for a peak in the spectral content for the full observing run.
-Instead, the common method is to split the observed time series into bins of time, compute the spectra for each time interval, and then find the most likely path for the signal frequency through the spectra over time.
-The plot of spectra over time is known as a spectrogram.
-This turns the problem of searching for a signal in noisy data into one of finding the best path through a grid of weights (here the Fourier amplitudes).
-We restrict the path-finding to some rate of frequency change of time, usually informed by a physical model, but here just set to 0.11 Hz per second to make it one frequency bin per every time bin.
+One method (among many) is to split the observed time series into bins of time, compute the Fourier transform for each time interval, and then create a grid (known as a spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time.
+The method then treats the problem as one of finding the best path through a weighted graph, where here the weights are given by the Fourier amplitudes.
+The method takes in a parameter of the allowed frequency wandering, some rate of frequency change over time, usually informed by a physical model.
+Here, this parameter is set to 0.11 Hz per second to make it simply one frequency bin per every time bin, which will prove poorly adjusted to the signal.
 
@@
 
-The Viterbi algorithm is a path-finding algorithm that solves this particular problem of going from one side of a grid to the other.
+The Viterbi algorithm is a path-finding algorithm that solves this problem.
+Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns holding spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
+Then, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph.
+At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path.
 See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.
-Abstractly speaking through, given a weighted graph (e.g.\ the spectrogram), a sensible sequence of subgraphs (e.g.\ the spectra at each time bin), and any restrictions on connectivity (e.g.\ the allowed amount of source frequency change over time), the Viterbi algorithm will find the best path to each node in the next subgraph all the way from the first subgraph at each iteration.
-At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path.
 
@@
 
-To demonstrate this technique, we inject a tone that wanders in frequency over time, remembering to integrate up the phase change over time (see Appendix~\ref{app:phase_gotcha} for why).
-Alternatively, a function generator can be slowly controlled to achieve the same input.
-Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, along with the Viterbi path and the injected signal.
-To the eye, the Viterbi algorithm recovers the injected frequency over time, likely due to the high signal-to-noise ratio (SNR).
+We demonstrate this method of using the Viterbi algorithm to find a signal, by injecting a wandering frequency signal into the speaker and recording the output via the webcam.
+Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
+
+Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, overlaid with the injected signal and the recovered Viterbi path.
+To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment.
 Quantifying this, the Viterbi path stays within one frequency bin (approximately 0.3Hz) of the injected tone for 94\% of the run.
-Note that the signal initially moves too quickly (more than 0.11 Hz per second) for Viterbi to track, and that there appears to be an anomaly at 150s, likely a disturbance around the interferometer.
-Decreasing the signal-to-noise ratio closer to unity or increasing the allowed change in frequency over time causes Viterbi to more easily lose the signal.
-However, the demonstration clearly shows the ability to recover a continuous wave with the Viterbi algorithm, just as is done in continuous gravitational wave searches.
+Note that the signal initially moves faster than the chosen frequency wander parameter, causing a discrepancy between the injected and recovered signals.
+To fix this, the parameter should be doubled, although this would then cause the recovered path to stray more often.
+There also appears to be an anomaly at 150s, likely from a disturbance (such as someone walking) around the interferometer.
+
+Empirically, decreasing the signal-to-noise ratio closer to unity also causes the Viterbi algorithm to more easily stray from the signal to reach nearby fluctuations.
+However, the only correlation through the grid is the signal, so we always expect the Viterbi path to lie near to the signal no matter how high the noise is.
 
@@
 \begin{figure}
-	\includegraphics[width=0.8\textwidth]{figures/setup_pic2.pdf}
+	\includegraphics[width=\textwidth]{figures/setup_pic2.pdf}
 	\caption{Picture taken of the optical microphone set-up, showing the Michelson interferometer, the photodiode (in an insert), the circuit, and the Raspberry Pi.}
@@
 \section{Future work}
+\label{sec:future_work}
 
-A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-Although we have the separation-intensity relation as non-linear for small motions, the other components of this full model are yet to be determined.
+\jam{BATTERY POWER}
 
@@
 
-\jam{should this be more broad a future work section?}
+\subsection{Transfer function}
+
+The intensity, $I$, of the interference pattern at a particular angle is given by Equation~\ref{eq:intensity_result} below.
+Where A is the amplitude of the electric field of the beam, d is the separation (or the difference in the arm lengths), $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Appendix~\ref{app:intensity_derivation} for the derivation to get to this equation).
+\begin{equation}
+\label{eq:intensity_result}
+I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+\end{equation}
+
+This Equation~\ref{eq:intensity_result} means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
+This makes the problem of extracting the original signal more difficult.
+
+\begin{equation}
+\label{eq:intensity_change}
+	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
+\end{equation}
+
+A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
+Although we have the separation-intensity relation as non-linear for small motions (see above), the other components of this full model are yet to be determined.
+Figure~\ref{fig:pipeline_highlighted} shows this full pipeline, with the currently determined section highlighted.
+
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{figures/pipeline_highlighted.pdf}
+	\caption{???}
+	\label{fig:pipeline_highlighted}
+\end{figure}
 
 \section{Conclusions}
+\label{sec:conclusions}
 
@@
 
-% \newpage
+\section{Guide to reproducing results}
+\label{app:reproducing_results}
+% as for a 3rd year undergraduate
+\jam{Add this in, lots of advice and detail!}
+
+\section{The Viterbi Algorithm}
+\label{app:viterbi}
+
+To find the Viterbi path through a spectrogram (spectra over time), given an M x N grid of M frequency rows and N time columns and a rate limit for the change of frequency over time that we convert into the number of allowed rows we can move while stepping from each column to the next (this limit was $\pm 1$ row/column in our case).
+The path must start from the first column and end at the last, taking a single step in each column along the way.
+Each element in the grid is the norm of the Fourier component of a frequency at a time.
+These are normalised between $(0, 1)$ to use as multiplicative weights, i.e.\ the total weight of a path is the product of all weights along that path.
+In fact, we take the logarithm and so the sum of all weights to avoid underflow.
+The best path has the highest weight.
+
+With the grid prepared, the Viterbi algorithm starts in the second column.
+In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, taking the greatest cumulative weight of those available, ‘nearby’ cells (the value of which was discovered in the previous iteration).
+For each cell in the current column, the algorithm notes this weight plus the cell’s own weight, along with the row index of which cell in the previous column it came from.
+Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
+That best path is the Viterbi path.
+
+\section{Common wandering frequency mistake}
+\label{app:phase_gotcha}
+
+We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
+The common mistake is to just appeal to the standard form $\sin{(2 \pi f t)}$, where f would be a constant frequency, and write down $\sin{(2 \pi f(t) t)}$.
+However, accumulation of phase means that this produces a curve that looks very different to that desired.
+Instead, one should see that the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
+Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
+
+\section{Circuit design}
+\label{app:circuit_diagram}
+% schematic of breadboard and connections to pi
+% https://www.circuit-diagram.org/editor/
+% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
+
+Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}
+
+\begin{figure}%[H]
+	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
+	\caption{Circuit diagram for photodiode reading.
+Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC), that’s finally read by the special purpose input (SPI) pins of a Raspberry Pi}
+	\label{fig:circuit_diagram}
+\end{figure}
+
 \section{Intensity relation derivation}
@@
 
-\section{The Viterbi Algorithm}
-\label{app:viterbi}
-
-To find the Viterbi path through a spectrogram (spectra over time), given an M x N grid of M frequency rows and N time columns and a rate limit for the change of frequency over time that we convert into the number of allowed rows we can move while stepping from each column to the next (this limit was $\pm 1$ row/column in our case).
-The path must start from the first column and end at the last, taking a single step in each column along the way.
-Each element in the grid is the norm of the Fourier component of a frequency at a time.
-These are normalised between $(0, 1)$ to use as multiplicative weights, i.e.\ the total weight of a path is the product of all weights along that path.
-In fact, we take the logarithm and so the sum of all weights to avoid underflow.
-The best path has the highest weight.
-
-With the grid prepared, the Viterbi algorithm starts in the second column.
-In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, taking the greatest cumulative weight of those available, ‘nearby’ cells (the value of which was discovered in the previous iteration).
-For each cell in the current column, the algorithm notes this weight plus the cell’s own weight, along with the row index of which cell in the previous column it came from.
-Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
-That best path is the Viterbi path.
-
-\section{Common changing frequency mistake}
-\label{app:phase_gotcha}
-
-We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
-The common mistake is to just appeal to the standard form $\sin{(2 \pi f t)}$, where f would be a constant frequency, and write down $\sin{(2 \pi f(t) t)}$.
-However, accumulation of phase means that this produces a curve that looks very different to that desired.
-Instead, one should see that the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
-Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
-
-\section{Circuit design}
-\label{app:circuit_diagram}
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-
-Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
-	\caption{Circuit diagram for photodiode reading.
-Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC), that’s finally read by the special purpose input (SPI) pins of a Raspberry Pi}
-	\label{fig:circuit_diagram}
-\end{figure}
-
 \begin{acknowledgments}

### 2020-02-12 paper/paper_main.tex
75th day of Summer

@@
 % one paragraph recount of what’s been done for two months, like the abstract
+
+\jam{write this!}
 
@@
 The techniques used are very similar to those used to search for continuous waves in LIGO and Virgo data.
-In Section~\ref{sec:ifo} we detail the interferometer set-up, Section~\ref{sec:single_tone} demonstrates observing a single tone, Section~\ref{sec:viterbi_wandering} details tracking a wandering frequency tone, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
+As an overview, Section~\ref{sec:ifo} details the interferometer, Section~\ref{sec:single_tone} describes observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
@@
 Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-\jam{Should I show photodiode Viterbi results?}
 
@@
 % logMMSE
+
+\jam{Explain MMSE as in the second talk, be honest in some way}
 
@@
 An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
-% % Visually it is impossible to confirm this for high frequencies, but it remains possible.
+% Visually it is impossible to confirm this for high frequencies, but it remains possible.
 
@@
 \jam{BATTERY POWER}
-
-Theoretically, if we had the correct model, then signal recovery would only be limited by the SNR. We could also use long recordings of white noise to estimate a matched filter for the system.
-
-For further speech enhancement, the optimal filter would involve a hidden-Markov-model trained to English phonemes. Alternatively, machine learning solutions also exist through-out the field.
-% SEGAN?
-
 % any other wild speculation?
@@
 
+Theoretically, if we had the correct model, then signal recovery would only be limited by the SNR. We could also use long recordings of white noise to estimate a matched filter for the system.
+
+For further speech enhancement, the optimal filter would involve a hidden-Markov-model trained to English phonemes. Alternatively, machine learning solutions also exist through-out the field.
+
 \begin{figure}%[H]
@@
 \section{Open-source code}
+\label{app:code}
 This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.
@@
 
-\section{Guide to reproducing results}
+\section{Additional guide to reproducing results}
 \label{app:reproducing_results}
 % as for a 3rd year undergraduate
+
+Besides the directions given in the rest of this article, there are some observations and technical details useful to someone wanting to reproduce the results shown herein.
+For the analysis and signal processing performed the best reference is Appendix~\ref{app:code} for the open-source code.
+
+For the experiment, since a Michelson interferometer is such a standard configuration, we’ll not detail it further here, see any optics manual for more information.
+To obtain an independent speaker cone we recommend carefully dismantling a budget commercial speaker system to get at the speaker within.
+To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place.
+If the other speaker in the pair remains, keep it as far away from the experiment and as damped as possible else it may act as a second, interfering source.
+% we used a Class 2 LDS5 ThorLabs laser at 532nm
+
+For the photodiode circuit, we used a Raspberry Pi and an MCP3008 ADC to monitor the analog signal.
+This could also be done with the analog pins on an Arduino.
+Be careful of the gain through the LM358 op-amps in the circuit though, else the ADC in either case could threshold the signal.
+Also be ready to change the values in the Sallen-Key filter to match your own sampling rate.
+The photodiode was an OSRAM BPW21, we mounted it through a discarded piece of speaker mesh (similar to black cloth over a frame) and placed another piece over the face to not over expose and threshold the signal.
+For the rest of the circuit, we used standard components assembled as in \ref{app:circuit_diagram}.
+
 \jam{Add this in, lots of advice and detail!}

### 2020-02-13 paper/paper_main.tex
adding some more text and references to introduction

@@
 Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience.
-Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance \han{add refs}.
+Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
 
@@
 They are disturbances in spacetime caused by the acceleration of massive objects. 
-The effect of gravitational waves is to change distances; a ``stretch and squeeze'' effect. 
+The effect of gravitational waves is a change in distances; a ``stretch and squeeze'' effect. 
 Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
-The observatories themselves are extremely complex devices, however they are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate this science to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO}\han{add exhibit info when available} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
+Gravitational-wave observatories are extremely complex devices, however they are based on the Michelson interferometer. 
+Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
 
@@
 To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-However, these observatories are also searching for continuous gravitational waves; persistent, periodic signals which should be present in the data for many months or years. 
-A prime target for these searches are rotating neutron stars. 
-So far many searches have been made of LIGO and Virgo data, however no detection has been made so far.
+These observatories are also searching for continuous gravitational waves (CW); persistent periodic signals at near-monochromatic frequencis. 
+One of the prime targets of CW searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
+LMXBs are binary systms containing a compact object (such as a neturon star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
+The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
+It is a prime target for CW searches from LMXBs and numerous searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
 
-% overview
-In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous gravitational waves to a general audience. 
-The techniques used are very similar to those used to search for continuous waves in LIGO and Virgo data.
-As an overview, Section~\ref{sec:ifo} details the interferometer, Section~\ref{sec:single_tone} describes observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
+% sound analogy
+Here we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
+The analogy of listening to the universe is often used to explain gravitational-wave observations; transient gravitational-wave signals have been converted into audio.
+Continuous waves may emitt at constant frequencies, or they may wander slowly in frequency over time. 
+This can be demonstrated with a table-top interferometer by playing audio of constant or slowly changing frequency tones to produce a simulation of a continuous wave signal. 
+The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data, particularly for the wandering frequency demonstration~\cite{ScoX1ViterbiO2,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
+
+In this work we detail our table-top intererometer design in Section~\ref{sec:ifo}, in Section~\ref{sec:single_tone} we describe observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.

### 2020-02-13 paper/paper_main.tex
small change in intro text

@@
 To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-These observatories are also searching for continuous gravitational waves (CW); persistent periodic signals at near-monochromatic frequencis. 
-One of the prime targets of CW searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
+These observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencis. 
+One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
 LMXBs are binary systms containing a compact object (such as a neturon star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
 The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for CW searches from LMXBs and numerous searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
+It is a prime target for continuous wave searches from LMXBs and numerous searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
 

### 2020-02-13 paper/paper_main.tex
setting up merge conflict

--- 2020-02-12 paper/paper_main.tex
@@
 \item all of the source code at \url{https://github.com/daccordeon/gravexplain}, that reads the data from the Raspberry Pi, analyses it, and creates the figures used in the paper
-\item talks presented were based on paper and figures therein
+\item talks presented were based on the paper and figures therein
 \end{itemize}
@@
 % one paragraph recount of what’s been done for two months, like the abstract
-
-\jam{write this!}
+The project spanned three weeks in December, four weeks in January, and two weeks in February.
+The first two weeks were spent writing the initial scripts and achieving the Viterbi results with the webcam set-up.
+The third week was spent setting up the photodiode circuit and the Raspberry Pi.
+All of January was spent progressing through the optical microphone experiment, and was largely driven by suggestions made at the signal processing meetings with the electrical engineers.
+By the end of the third week of January we had a working optical microphone but low speech intelligibility.
+The fourth week of January was spent on attempting methods of speech enhancement.
+Finally, the two weeks in February were spent writing and drafting the paper/report, and giving presentations.
 
@@
 \begin{abstract}
-Continuous gravitational wave searches use sophisticated statistical techniques to search for signals in noisy data.
-We demonstrate some of these techniques with a table-top Michelson interferometer, namely the Viterbi algorithm for recovering a tone with wandering frequency.
-We also use the same experiment to demonstrate an optical microphone capable of playing back simple audio recordings of injected music.
-However, played-back speech is currently unintelligible so we investigate preliminary methods of speech enhancement.
-
-\jam{Andrew says that this should be a bit more detailed!}
+Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
+We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
+In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
+We also demonstrate how the inclusion of a photodiode and Raspberry Pi allows the interferometer to act as an optical microphone.
+We are able to play-back simple audio recordings from the photodiode data.
+However, played-back speech is unintelligible and so we investigate techniques of speech enhancement.
 
@@
 
-\section{Complex audio}
+\section{Playing back complex audio}
 \label{sec:optical_microphone}
@@
 
-The natural succession to simple tones is to study complex audio signals, such as music and speech.
+The natural succession to observing simple tones is to study complex audio signals, such as music and speech.
 The Viterbi algorithm and tracking more broadly is not applicable to these signals.
@@
 
-For an optical microphone to be sensitive to speech and music it needs a frequency response far higher than the 30Hz of a standard webcam.
-Both speech and music need at least kHz range, with speech intelligibility requiring up to 3kHz and music preferably up to and beyond 8kHz.
+For an optical microphone to be sensitive to speech and music it needs a frequency response (a sampling rate) far higher than the 30Hz of a standard webcam.
+Both speech and music need at least kHz range, with speech intelligibility requiring up to 3kHz and music preferably up to and beyond 8kHz\cite{speech_intelligibility}.
 This can be addressed by using a high-speed webcam, or more simply, by using a photodiode as the measuring device.
@@
 As such, any frequency component of the analog signal above the Nyquist frequency of 8kHz would be aliased down into the detected range.
-To prevent this, an anti-aliasing Sallen-Key filter tuned to 16kHz is included before the ADC. This component attenuates any frequencies above 8kHz, before they get digitally sampled.
+To prevent this, an anti-aliasing Sallen-Key filter\cite{sallen_key_filter} tuned to 16kHz is included before the ADC.
+This component attenuates any frequencies above 8kHz, before they get digitally sampled.
 It’s not immediately clear what’s limiting the sample rate to 16kHz, but it is likely non-optimal reading of the ADC by the Pi script.
@@
 	\includegraphics[width=\textwidth]{figures/setup_pic2.pdf}
-	\caption{Picture taken of the optical microphone set-up, showing the Michelson interferometer, the photodiode (in an insert), the circuit, and the Raspberry Pi.}
+	\caption{Picture taken of the optical microphone set-up, showing the Michelson interferometer, the photodiode (in an insert), the circuit, and the Raspberry Pi.
+For a closer photo of the circuit, see Figure~\ref{fig:circuit_pic3}.}
 	\label{fig:setup_pic2}
@@
 
-\subsection{Filters applied}
+\subsection{Applying filters to the data}
 
@@
 
-Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
+% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
 Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
@@
 
-\subsection{logMMSE estimator}
+\subsection{The logMMSE estimator}
 % logMMSE
@@
 
-\subsection{Transfer function}
+\subsection{Finding the transfer function}
 
@@
 The photodiode was an OSRAM BPW21, we mounted it through a discarded piece of speaker mesh (similar to black cloth over a frame) and placed another piece over the face to not over expose and threshold the signal.
-For the rest of the circuit, we used standard components assembled as in \ref{app:circuit_diagram}.
+For the rest of the circuit, we used standard components assembled on a breadboard as detailed in Appendix~\ref{app:circuit_diagram}.
 
-\jam{Add this in, lots of advice and detail!}
-
-\section{The Viterbi Algorithm}
+\section{The Viterbi algorithm}
 \label{app:viterbi}
@@
 
-\section{Common wandering frequency mistake}
+\section{A common mistake with wandering frequency signals}
 \label{app:phase_gotcha}
@@
 
-Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}
+Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
+A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
 
@@
 	\caption{Circuit diagram for photodiode reading.
-Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC), that’s finally read by the special purpose input (SPI) pins of a Raspberry Pi}
+Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC).
+The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.}
 	\label{fig:circuit_diagram}
+\end{figure}
+
+\begin{figure}%[H]
+	\includegraphics[width=0.8\textwidth]{figures/circuit_pic3.pdf}
+	\caption{Photo of circuit for photodiode reading assembled on breadboard.}
+	\label{fig:circuit_pic3}
 \end{figure}

### 2020-02-13 paper/paper_main.tex

@@
 To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-However, these observatories are also searching for continuous gravitational waves; persistent, periodic signals which should be present in the data for many months or years. 
-A prime target for these searches are rotating neutron stars. 
-So far many searches have been made of LIGO and Virgo data, however no detection has been made so far.
+These observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencis. 
+One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
+LMXBs are binary systms containing a compact object (such as a neturon star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
+The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
+It is a prime target for continuous wave searches from LMXBs and numerous searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
 
-% overview
-In this article we describe how a Michelson interferometer can be used to demonstrate the search for continuous gravitational waves to a general audience. 
-The techniques used are very similar to those used to search for continuous waves in LIGO and Virgo data.
-As an overview, Section~\ref{sec:ifo} details the interferometer, Section~\ref{sec:single_tone} describes observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
+% sound analogy
+Here we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
+The analogy of listening to the universe is often used to explain gravitational-wave observations; transient gravitational-wave signals have been converted into audio.
+Continuous waves may emitt at constant frequencies, or they may wander slowly in frequency over time. 
+This can be demonstrated with a table-top interferometer by playing audio of constant or slowly changing frequency tones to produce a simulation of a continuous wave signal. 
+The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data, particularly for the wandering frequency demonstration~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
+
+In this work we detail our table-top intererometer design in Section~\ref{sec:ifo}, in Section~\ref{sec:single_tone} we describe observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.

### 2020-02-13 paper/paper_main.tex
2nd draft

@@
 We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
+In particular, we use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
 We also demonstrate how the inclusion of a photodiode and Raspberry Pi allows the interferometer to act as an optical microphone.
-We are able to play-back simple audio recordings from the photodiode data.
-However, played-back speech is unintelligible and so we investigate techniques of speech enhancement.
+That is, we are able to play back simple audio recordings from the photodiode data.
+However, more complex recordings, of speech, are unintelligible when played back, despite preliminary efforts towards speech enhancement.
 
@@
 
-Advanced signal processing is required to extract gravitational wave signals from very noisy observational data.
+Advanced signal processing is required to extract gravitational wave signals from noisy observational data.
 Recovering audio from an optical microphone requires related signal processing.
@@
 A photodiode is an electrical component that acts as a regular diode (blocking any current flow in the reverse direction) when no light is incident on it, but becomes more and more conductive in the reverse direction as the intensity of incident light increases.
-Fundamentally, a current is created in the photodiode by the photo-electric effect.
-Placing a photodiode in reverse-bias over an op-amp creates a photo-detector that will produce a voltage dependent approximately linearly to the incident intensity.
-\jam{How sound is this approximation?} This experiment had the photodiode in the interference pattern at roughly the same spot as the webcam was monitoring previously, just off-centre.
+% Fundamentally, a current is created in the photodiode by the photo-electric effect.
+Placing a photodiode in reverse-bias over an op-amp creates a photo-detector that will produce a voltage dependent on the incident intensity.
+This experiment had the photodiode in the interference pattern at roughly the same spot as the webcam was monitoring previously, just off-centre in the position of the red dot in Figure~\ref{fig:interference_pattern}.
 
-The voltage signal from the photo-detector is captured by a 10-bit analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi, in the standard configuration.
+The voltage signal from the photo-detector is captured by a 10-bit analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi Model 3 v1.2, in the standard configuration.
 This digitally samples the signal at roughly 16kHz.
@@
 The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the experiment, showing the interferometer, the circuit, and the Pi, is shown in Figure~\ref{fig:setup_pic2}.
+The only change to the interferometer schematic (Figure~\ref{fig:ifo_schematic_podo}) is the substitution of the webcam for the photodiode.
 
@@
 
-The optical microphone was tested on recordings of one human talking and on music ranging from simple melodies to full tracks.
-Each recording was around one minute long with processing done on ten second segments.
+The optical microphone was tested on recordings of speech from different people and on music ranging from simple melodies to full track songs.
+Each recording was around one minute long with processing often done on ten second segments.
 For each observation, the audio was played through the speaker with minimal activity around interferometer as the Pi script recorded the signal.
-The saved time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{scipy.io.wavfile.write}\cite{scipy} implementation in Python\cite{python}.
+The time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{scipy.io.wavfile.write}\cite{scipy} implementation in Python\cite{python}.
 
-These initial recordings were extremely noisy and painful to listen to, with a loud bass sound throughout.
-Spectra of all of these recordings showed a dominant mains hum at 50Hz with strong power up to and beyond the 8th harmonic thereof.
-This power is also present with the speaker off, as is shown in Figure~\ref{fig:psd_noise}, and also weakly when monitoring with the photodiode in complete darkness.
+When played back and listen to, these initial recordings were extremely noisy and had a loud bass note throughout.
+Each Fourier spectrum showed a dominant AC mains hum with power ranging from the fundamental of 50Hz up to and beyond the 8th harmonic.
+With the speaker off this mains power was still present, as shown the power spectral density (effectively the norm squared of the Fourier spectrum) in Figure~\ref{fig:psd_noise}.
+And even with the speaker off and the photodiode in complete darkness, the mains hum was still present (with distribution similar to but weaker than that shown in Figure~\ref{fig:psd_noise}).
 
@@
 	\includegraphics[width=\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
-	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), show high power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the ambient lighting), otherwise is a fairly white spectrum except for the broad, shallow maximum at 750Hz}
+	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), shows strong power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and cooling), otherwise is a fairly white spectrum except for the broad, shallow maximum at 750Hz}
 	\label{fig:psd_noise}
@@
 \subsection{Applying filters to the data}
+\label{sec:simple_filters}
 
-Early attempts to remove the mains hum included zeroing the channel of each 50Hz harmonic, this is effectively using a rectangular comb filter.
-Applying a filter (multiplying by a function) in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
-While this did remove the mains harmonics, it also destroyed the rest of the signal.
+Early attempts to remove the mains hum included zeroing the channel of each 50Hz harmonic in the Fourier spectrum.
+Effectively, this is multiplying the spectrum by a rectangular comb filter.
+While this did remove the mains harmonics, it also audibly ruined the rest of the signal.
+This is because applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
+And the inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complex sinc-like function that when convolved with significantly damages the actual signal.
 
-Attempting to low-pass the signal did better (to the ear) but was also problematic as the harmonics are all over the 100Hz-1kHz region, which also carries a lot of the speech and music information.
-Low-passing the logarithm of the signal and then converting back did better still but doesn’t significantly remove the mains noise.
+To fix this, we tried using a smoother comb filter but this suffered from too significantly attenuating the rest of the spectrum between each harmonic.
 
-Of these simple filters, the best performing was a band-pass Butterworth filter with frequency response shown in Figure~\ref{fig:butterworth}.
-This filter is `maximally flat' within the band region, here 150Hz to 3kHz, and trails off outside of the band with slope dependent on the order of the filter, here of order 5.
-This still suffered from the presence of the higher mains harmonics.
-However, behind the remaining noise one could hear something distinctly speech-like or musical, just with very low intelligibility, you couldn’t understand the speech but could recognise that someone was talking.
+Applying a high-pass filter to the signal worked well at removing the 50 and 100Hz tones but was problematic as the mains harmonics extend far past 100Hz.
+This region above 100Hz carries a lot of the fundamental frequencies for speech and music\cite{speech_intelligibility} and if the cut-off frequency is placed as to remove all the mains harmonics, then the resultant signal is unrecognisable.
+
+We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass.
+
+Band-passing the signal similarly doesn’t address the remaining harmonics but does low-pass away other noise at higher frequencies.
+The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over a band from 150Hz to 3kHz (beyond 2kHz the information content drops\cite{speech_intelligibility}).
+For a low-pass Butterworth filter at cut-off frequency $\omega_c$ and gain $\epsilon$ the frequency response (attenuation at each frequency) follows Equation~\ref{eq:butterworth}.
+\begin{equation}
+\label{eq:butterworth}
+H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\frac{\omega}{\omega_c})^{2n}}}
+\end{equation}
+For a band-pass filter, the above low-pass is combined with a high-pass filter of similar form.
+The response of the filter used is shown in Figure~\ref{fig:butterworth}.
+Despite improving, the recording with the filter applied still had very low intelligibility, it could be recognised as speech but was not understandable.
+
+The ideal filter would be one that completely attenuates the undesired parts of the spectrum, has no changing attenuation (is flat) on the rest of the spectrum, and has the smoothest edges as to not damage the signal.
+These three conditions are contradictory and any filter must compromise between them.
+The above simple filters fail at one or more of each of these conditions.
+The combs ruin the signal and the high-passes don’t remove the noise.
+The Butterworth filter performed the best audibly as it was maximally flat within the band region, prioritising the second condition.
 
@@
 	\caption{Butterworth bandpass filter frequency response, green vertical lines show the band limits of 150Hz and 3kHz.
-Note the very flat response within the band, any amplitude beyond -3dB is significant attenuation.}
+Note the flat response within the band, any amplitude beyond -3dB is significant attenuation (half power).}
 	\label{fig:butterworth}
@@
 
-\jam{Explain MMSE as in the second talk, be honest in some way}
+Speech enhancement for a noisy channel is a classic problem in signal processing.
+Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compared various methods and found the best performing statistical technique among those at the time to be a logMMSE (alt.\ MMSE-LSA) estimator.
+This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
+Although this technique historically minimised with respect to the spectral amplitude\cite{EphraimSpeechEU_MMSE}, throughout speech processing use of the logarithm seems produce better results as perceived by the human ear\cite{SubjectiveComparison}.
 
-Speech enhancement for a noisy channel is a classic problem in signal processing.
-Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare various methods and find the best performing statistical technique to be the logMMSE (alt.\ MMSE-LSA) estimator which minimises the mean square error (MSE) of the log-spectral amplitude.
-This means that it takes the logarithm of the spectrum of the original signal and finds the fit (or estimate) that has the minimal MSE to that log-spectra.
-% Use of the logarithm is a common technique in speech processing as it, somewhat surprisingly, does not significantly change the intelligibility to the human ear.
+We use an existing implementation\cite{logmmse} of the logMMSE estimator.
+Applying this estimator to the optical microphone recordings produces dramatically clean results.
+The time series (in Figure~\ref{fig:logMMSE_timeseries}) and spectrum (in Figure~\ref{fig:logMMSE_spectrum}) from before and after applying the logMMSE filter are shown.
+Comparing the spectra in Figure~\ref{fig:logMMSE_spectrum} shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.
 
-Applying an existing implementation of logMMSE\cite{logmmse} to the optical microphone recordings produces dramatically cleaner results.
-The time series and spectrum from before and after the logMMSE filter is applied are shown in Figures~\ref{fig:logMMSE_timeseries}~\&~\ref{fig:logMMSE_spectrum}.
-All of the mains harmonics are significantly attenuated.
-And further high-passing removes the low frequency curve beneath 100Hz.
+For recordings of music listened to after the estimator is applied, simple chords and drums are heard well, while more complex melodies are often completely absent in the results.
+By observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
+This could be a perceptual effect of listening to the lossy recording, but it seems more likely to be a frequency dependence somewhere in the optical microphone system.
+Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies.
 
-There is varied success with passing music passed through the filters.
-Loud, rhythmic chords and drums are received well and remain recognisable, while thinner, airy flute and violin melodies struggle to be heard in the results.
-It is likely that the lower sounds have higher amplitude speaker oscillations and so pass through the optical microphone better.
-
-However, speech intelligibility remains very low with the voice sounding mumbled and indistinct.
-Enhancement so far has removed most of the background sound, as evidenced by the near silence between words, but has not clarified the speech.
-There is a question here as to where the diction in the speech is lost, at the speaker-mirror coupling, at the photodiode reading, or if it is still there, just obscured in the noisy signal.
-
+After applying the estimator to speech recordings, intelligibility still remains low with the voice sounding mumbled and indistinct.
+The estimator removes most of the background sound, but does not clarify the speech.
+% There is a question here as to where the diction in the speech is lost, at the speaker-mirror coupling, at the photodiode reading, or if it is still there, just obscured in the noisy signal.
 % The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
@@
 An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
-% Visually it is impossible to confirm this for high frequencies, but it remains possible.
 
@@
 	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) logMMSE filter is applied, only one second of the one minute recording is shown.
-The bump at the start of the time series is the result of filtering a finite signal, and is expected.}
+The rise at the start of the time series is an expected result from filtering a finite signal.}
 	\label{fig:logMMSE_timeseries}
@@
 \label{sec:future_work}
+% wild speculation
+For better speech enhancement, we are currently working towards two future advances.
 
-\jam{BATTERY POWER}
-% any other wild speculation?
+The first would be to better isolate the optical microphone from the mains supply.
+The Raspberry Pi is USB powered and requires low power so it could even be run off a commercial external phone battery.
+Everything else in the circuit, including the photodiode, is connected to the Pi, so this would isolate them electrically as well.
+The laser is DC powered through a converter, so there is no issue there.
+With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
 
-\subsection{Finding the transfer function}
+The second advancement would be to achieve a full model of the transfer function of the system.
+Starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading, see Figure~\ref{fig:pipeline_highlighted} for this full sequence.
+Theoretically, if we had the correct transfer function, then signal recovery would only be limited by the signal-to-noise ratio as we could construct an optimal matched filter.
 
-The intensity, $I$, of the interference pattern at a particular angle is given by Equation~\ref{eq:intensity_result} below.
-Where A is the amplitude of the electric field of the beam, d is the separation (or the difference in the arm lengths), $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Appendix~\ref{app:intensity_derivation} for the derivation to get to this equation).
-\begin{equation}
-\label{eq:intensity_result}
-I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{equation}
+In our investigation, we have currently only derived part of this sequence, the relation between a deflection of the mirror and the change in the intensity of the interference pattern at some point (highlighted in Figure~\ref{fig:pipeline_highlighted}).
+We find this relation to be non-linear, see Appendix~\ref{app:intensity_derivation} for the full discussion and derivation.
 
-This Equation~\ref{eq:intensity_result} means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
-This makes the problem of extracting the original signal more difficult.
-
-\begin{equation}
-\label{eq:intensity_change}
-	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{equation}
-
-A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-Although we have the separation-intensity relation as non-linear for small motions (see above), the other components of this full model are yet to be determined.
-Figure~\ref{fig:pipeline_highlighted} shows this full pipeline, with the currently determined section highlighted.
-
-Theoretically, if we had the correct model, then signal recovery would only be limited by the SNR. We could also use long recordings of white noise to estimate a matched filter for the system.
-
-For further speech enhancement, the optimal filter would involve a hidden-Markov-model trained to English phonemes. Alternatively, machine learning solutions also exist through-out the field.
+Asides from these two main avenues, other techniques are available to better process the recording.
+In particular, for extracting the speech there exists much literature\cite{HMM_english} on using a hidden Markov model train to English phonemes to recognise speech.
+Alternatively, machine learning solutions also exist throughout the field that are able to compete\cite{HMM_english} with statistical techniques like the logMMSE estimator.
+More simply, to remove any remaining mains hum we could monitor it separately from the interferometer and subtract the sinusoid with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
+This could also be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
 
@@
 	\includegraphics[width=\textwidth]{figures/pipeline_highlighted.pdf}
-	\caption{???}
+	\caption{Sequence of relations for optical microphone transfer function}
 	\label{fig:pipeline_highlighted}
@@
 
-Using a Michelson interferometer, we able to successfully demonstrate the Viterbi algorithm recovering an injected signal.
-Just as is done is continuous gravitational wave searches.
-We are also able to play-back chordal or rhythmic music from an optical microphone (using the same interferometer), while recordings of simple speech are audible but unintelligible.
+% Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
+We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
+In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
+
+We use sound and a table-top Michelson interferometer as an analogue to gravitational wave detectors to demonstrate signal processing techniques.
+Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer.
+We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.
+
+Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the intensity time series of the interference pattern into an audio recording and the interferometer into an optical microphone.
+We demonstrated that this optical microphone is capable of playing back simple recordings well and speech recordings unintelligibly.
+Dominant 50Hz mains hum in the recordings led to an exploration of speech enhancement techniques to remove it, including the well performing logMMSE estimate.
+From here we currently have many plans to further remove the mains hum and improve the intelligibility.
+
+These demonstrations have the potential to be adapted into the undergraduate laboratory.
+There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches, but also in the signal processing of speech enhancement as an open problem for engineering focused students.
+See Appendix~\ref{app:reproducing_results} for advice aimed at the undergraduate student for replicating the results.
 
@@
 The common mistake is to just appeal to the standard form $\sin{(2 \pi f t)}$, where f would be a constant frequency, and write down $\sin{(2 \pi f(t) t)}$.
-However, accumulation of phase means that this produces a curve that looks very different to that desired.
+However, accumulation of phase produces the wrong curve.
 Instead, one should see that the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
@@
 \label{app:intensity_derivation}
+
+The intensity, $I$, of the interference pattern at a particular angle is given by Equation~\ref{eq:intensity_result} below.
+Where A is the amplitude of the electric field of the beam, d is the separation (or the difference in the arm lengths), $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Appendix~\ref{app:intensity_derivation} for the derivation to get to this equation).
+\begin{equation}
+\label{eq:intensity_result}
+I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+\end{equation}
+
+This Equation~\ref{eq:intensity_result} means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
+This makes the problem of extracting the original signal more difficult.
+
+\begin{equation}
+\label{eq:intensity_change}
+	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
+\end{equation}
 

### 2020-02-14 paper/paper_main.tex
PhB final submission

@@
 % \usepackage{siunitx}
+\usepackage{textcomp}
 
@@
 \begin{itemize}
-\item all paper sections other than the introduction, i.e.\ from Section~\ref{sec:ifo} onwards, and all figures in the paper
+\item all paper sections other than the introduction, i.e.\ the abstract and from Section~\ref{sec:ifo} onwards, and all figures in the paper
 \item addition of the injection speaker, design of photodiode circuit from standard examples, assembly of circuit on breadboard, and set-up of Raspberry Pi to measure output
-\item all of the source code at \url{https://github.com/daccordeon/gravexplain}, that reads the data from the Raspberry Pi, analyses it, and creates the figures used in the paper
-\item talks presented were based on the paper and figures therein
+\item all of the source code at \url{https://github.com/daccordeon/gravexplain} (containing over 1000 lines of Python scripts), that reads the data from the Raspberry Pi, performs all of the analysis on it, and creates all the figures used in the paper
+\item presentation of two talks based on the paper and the figures therein
 \end{itemize}
@@
 \begin{itemize}
-\item introduction of paper, i.e.\ Section~\ref{sec:introduction}
+\item the introduction of the paper, i.e.\ Section~\ref{sec:introduction}
 \item design and assembly of the interferometer
+\item continuous support and advice throughout the project
 \end{itemize}
 
-Andrew Melatos, Robin Evans, William Moran: vital advice and comments throughout the project
+Andrew Melatos, Robin Evans, William Moran: other vital advice and comments throughout the project
 
-\subsection*{Statement of achievements}
+\newpage
+\section*{Statement of achievements}
 % one paragraph recount of what’s been done for two months, like the abstract
@@
 The first two weeks were spent writing the initial scripts and achieving the Viterbi results with the webcam set-up.
-The third week was spent setting up the photodiode circuit and the Raspberry Pi.
-All of January was spent progressing through the optical microphone experiment, and was largely driven by suggestions made at the signal processing meetings with the electrical engineers.
-By the end of the third week of January we had a working optical microphone but low speech intelligibility.
-The fourth week of January was spent on attempting methods of speech enhancement.
-Finally, the two weeks in February were spent writing and drafting the paper/report, and giving presentations.
+This was relatively straight forward and the greatest hurdle was understanding the algorithm.
+
+The third week was spent setting up the photodiode circuit and the Raspberry Pi which involved constant testing and changing of the components and design of the circuit.
+All of January was spent on the optical microphone.
+Much difficultly was had combating 50Hz mains hum in the experiment.
+Eventually through many discussions at the signal processing meetings, we had a working optical microphone but low speech intelligibility.
+The last week of January was spent on methods of speech enhancement.
+If the investigation had continued we had much future work planned and possible solutions in mind.
+
+Finally, the start of February was spent writing the paper and giving presentations (one nine minute general talk and one 50 minute technical follow-up talk).
 
@@
 
-\title{GravExplain: Continuous gravitational wave searches in a table-top experiment}
+\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
 
@@
 \email{u6069809@anu.edu.au}
-\affiliation{Research School of Physics, Australian National University, 60 Mills Rd, Acton ACT 2601 Australia}
+\affiliation{College of Science, Australian National University, Acton, ACT, 2601, Australia}
 
@@
 Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-We also demonstrate how the inclusion of a photodiode and Raspberry Pi allows the interferometer to act as an optical microphone.
-That is, we are able to play back simple audio recordings from the photodiode data.
-However, more complex recordings, of speech, are unintelligible when played back, despite preliminary efforts towards speech enhancement.
+We demonstrate these techniques by observing sound with a table-top Michelson interferometer.
+In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
+We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings.
+However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
 
@@
 
+\begin{figure}
+	\includegraphics[width=0.9\textwidth]{figures/ligo_pic.pdf}
+	\caption{Aerial photo of the LIGO Hanford facility and its 4km long beam arms, courtesy of LIGO/Caltech/MIT}
+	\label{fig:ligo_pic}
+\end{figure}
+
 % general gw intro
-In 2015 gravitational waves were observed for the first time from the merger of two black holes in a binary system.~\cite{GW150914} 
-The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe.
+In 2015, gravitational waves were observed for the first time from the merger of two black holes in a binary system~\cite{GW150914}. 
+The observation, made by the Laser Interferometer Gravitational wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
 Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years there has been an increased public interest in gravitational-wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience.
-Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
+In recent years, there has been increased public interest in gravitational wave science. 
+Many gravitational wave research groups around the world have produced demonstrations and activities to explain the topic to a general audience.
+Activities range from hands-on demonstrations, exhibitions, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online}, video games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
 
@@
 Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are disturbances in spacetime caused by the acceleration of massive objects. 
-The effect of gravitational waves is a change in distances; a ``stretch and squeeze'' effect. 
+They are disturbances in space-time caused by the acceleration of massive objects. 
+The effect of gravitational waves is a change in lengths; a `stretching and squeezing' of the very distance between objects.
 Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance.
-Gravitational-wave observatories are extremely complex devices, however they are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
+As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
+These observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
+Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets}.
 
 % continuous gws
-To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-These observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencis. 
+To date, while the network of gravitational wave observatories have only observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}, these observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencies.
+These continuous waves may emit at constant frequencies, or they may wander slowly in frequency over time. 
 One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
-LMXBs are binary systms containing a compact object (such as a neturon star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
+LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
 The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for continuous wave searches from LMXBs and numerous searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
+It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
 
 % sound analogy
-Here we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
-The analogy of listening to the universe is often used to explain gravitational-wave observations; transient gravitational-wave signals have been converted into audio.
-Continuous waves may emitt at constant frequencies, or they may wander slowly in frequency over time. 
-This can be demonstrated with a table-top interferometer by playing audio of constant or slowly changing frequency tones to produce a simulation of a continuous wave signal. 
-The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data, particularly for the wandering frequency demonstration~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
+Here, we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience; the analogy of listening to the universe is often used to explain gravitational wave observations.
+ %; transient gravitational wave signals have been converted into audio.
+This can be demonstrated with a table-top interferometer by playing sound close enough to move the mirrors and change the interference pattern.
+Using audio of slowly wandering frequency tones to produce an analogue of a continuous gravitational wave signal.
+The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}, particularly for the wandering frequency demonstration.
 
-In this work we detail our table-top intererometer design in Section~\ref{sec:ifo}, in Section~\ref{sec:single_tone} we describe observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
+In this work we detail our table-top interferometer design in Section~\ref{sec:ifo} and describe observing a single note from a speaker in Section~\ref{sec:single_tone}.
+In Section~\ref{sec:viterbi_wandering}, we demonstrate the analysis techniques for observing a wandering frequency signal.
+In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech.
+This demonstration of an `optical microphone' serves as a more general exhibition of signal processing.
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
@@
 
-Detectable gravitational waves are caused by some of the most violent events in the universe.
-Like two massive black-holes merging with each other\cite{GW150914}.
-Even so, the effects they produce are on the order of one part in $10^{21}$ for our best detectors\cite{GW150914}.
-For a table-top experiment these effects would be far too small, so we demonstrate observing a far louder source, sound waves.
-Although the analogue to gravitational waves is tenuous, we can still demonstrate some of the continuous wave analysis satisfactorily.
-The other benefit of observing sound is that we can use existing signal processing techniques designed for audio signals.
+Even though gravitational waves are caused by some of the most massive events in the universe, their observable effects on Earth are vanishingly small | on the order of only one part in $10^{21}$ for our best detectors~\cite{GW150914}.
+Clearly, the measurement of such effects is far beyond any table-top experiment.
+But, the fundamental principle of gravitational wave detection remains that changes in the lengths of the beam arms of an interferometer leads to a change in the interference pattern observed.
+By observing sound waves that vibrate the mirrors of a table-top interferometer and so change the lengths of the beam arms we have an analogue of this principle.
 
-\subsection{Michelson interferometry}
+\subsection{Method of Michelson interferometry to observe sound}
 % hardware: path difference, virtual thin film interference
 
-A Michelson interferometer is a well-known optics configuration used in many undergraduate and research labs, including in the detectors for gravitational waves.
-A laser beam of coherent (monochromatic and equal phase) light is split evenly by a 50/50 beamsplitter into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
+In a Michelson interferometer, a coherent (monochromatic and equal phase) source of laser light is split evenly by a beamsplitter into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
 The beams recombine and interfere at the beamsplitter before hitting a screen to produce a pattern of concentric, circular fringes.
-The configuration is shown in Figure~\ref{fig:ifo_schematic_webcam}.
-In the detectors for gravitational waves, these arms are kilometres long and many other complexities (such as power recycling and quantum squeezing)\cite{GW150914} are necessary to observe a signal.
+The configuration used in this experiment is shown diagrammatically in Figure~\ref{fig:ifo_schematic_webcam} and as part of a photo in Figure~\ref{fig:setup_pic2}.
 
-\begin{figure}
+A sample interference pattern is shown in Figure~\ref{fig:interference_pattern}.
+This pattern is due to the path difference between the two arms of unequal lengths (if the lengths were equal, then only a single, bright dot would be seen).
+This results in a phase difference between the two beams when they meet at the screen.
+This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
+For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
+When the path difference changes, the fringes stay circular but move radially in or out.
+
+\begin{figure}[h]
 	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_webcam.pdf}
-	\caption{Michelson interferometer schematic, interference pattern is recorded by a webcam}
+	\caption{Michelson interferometer schematic with interference pattern recorded by a commercial webcam.
+We used a Class 2 532nm laser~\cite{ThorLabsIFO} and a commercial speaker.
+The beam is focused at a converging lens before the 50/50 beamsplitter and the interference pattern is made visible by a diverging lens after the beamsplitter.
+The arms lengths between the speakers are roughly 3cm different and the entire interferometer (without the screen) fits within the dimensions of an A4 piece of paper.
+We did not include a compensating plate with the beamsplitter, as is often done.}
 	\label{fig:ifo_schematic_webcam}
 \end{figure}
+
+We used a 0.5W speaker stuck to the back of one of the mirrors as an audio source for our demonstration.
+This speaker was part of a commercial 3.5mm-jack, USB powered pair of speakers and was driven by a PC.
+To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place.
+We kept the other speaker in the pair face-down and as far away from the experiment as the cables would allow to reduce it acting as a second source.
+
+As the speaker deforms to play sound, the mirror moves with a change in separation as some coupling function of the amplitude and frequency of the sound.
+The motion of the fringes follows this motion of the mirror and therefore the speaker at the same frequency.
+If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
+If the fringe motion was large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting would artificially raise the measured frequency.
+As such, any motion of the fringes must be kept small by playing sound softly through the speaker.
+
+For later sections, we derive the non-linear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen, this can be found in Appendix~\ref{app:intensity_derivation}.
+
 % Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
@@
 
-The interference pattern, as shown in Figure~\ref{fig:interference_pattern}, is due to the path difference between the two arms of unequal lengths (if the lengths were equal, then only a single, bright dot would be seen).
-This results in a phase difference between the two beams when they meet at the screen.
-This path difference is perhaps better seen as the perpendicular separation between one mirror and the virtual image of the other.
-For an aligned configuration and a fixed separation, the phase difference at a point only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-
-\begin{figure}
+\begin{figure}[h]
 	\includegraphics[width=\textwidth]{figures/webcam_still0.pdf}
@@
 
-Continuous gravitational wave searches are, by the nature of the astrophysical sources that they look for, always searching for periodic, or even sinusoidal, signals.
-Often the frequency of this periodicity can change, as addressed later, but sometimes it is (near) constant.
-As such, the audio analogue is to observe a single tone (a sinusoidal note of one frequency).
+As mentioned in Section~\ref{sec:introduction}, continuous gravitational wave searches look for near monochromatic signals. The audio analogue of this is a simple sinusoidal tone; a constant note.
 
-\subsection{Method}
+We measured the intensity at a point on the screen with a commercial USB webcam placed as to view the screen as in Figure~\ref{fig:ifo_schematic_webcam}.
+This webcam samples at 30Hz, which limits the spectral content of signals we can observe to below 15Hz, the Nyquist frequency.
+% See later Section~\ref{sec:photodiode} where we raise this limit into audible frequencies (20 Hz to 20kHz) with a photodiode.
+While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
+Then, at an off-centre pixel (shown as a red dot in Figure~\ref{fig:interference_pattern}), we took the green-channel of the video as approximate to the total intensity, as the laser is monochromatic green light (at 532nm).
 
-We use a 0.5W speaker stuck to the back of one of the mirrors as a source for our demonstration, as shown in Figure~\ref{fig:ifo_schematic_webcam}.
-This speaker was part of a commercial 3.5mm-jack, USB powered pair of speakers and is driven by a PC.
-Previously, we used a function generator to accomplish the same signal.
-As the speaker deforms to play sound, the mirror moves with a change in separation, $\delta d$, as some coupling function of the amplitude and frequency of the sound.
-The motion of the fringes follows the motion of the mirror and therefore the speaker at the same frequency.
-If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected (played) audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
-Large fringe motion means that multiple fringes can pass over the measured point in a single speaker deformation, such over-counting would artificially raise the measured frequency.
-As such, any motion of the fringes, and so of the speaker, must be kept small by playing sound softly through the speaker.
+We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}.
+The two tones played together sound similar and don’t produce any noticeable beating.
+To quantify the precision, we recorded a signal from an injected tone of 2.09Hz.
+From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\mathrm{Hz}$, a result within two standard deviations.
+This showed that the experiment observes a single tone confidently.
 
-\subsection{Results}
-
-We measure the intensity at a point on the screen with a commercial USB webcam placed as to view the screen (Figure~\ref{fig:ifo_schematic_webcam}), sampling at 30Hz (or 30 fps, frames-per-second).
-This limits the spectral content of signals we can observe to be under the Nyquist frequency of 15Hz.
-See later Section~\ref{sec:photodiode} where we raise this limit into to audible frequencies (20 Hz to 20kHz) with a photodiode.
-While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
-Then, at an off-centre pixel as seen in Figure~\ref{fig:interference_pattern}, we take the green-channel of the colour (RGB) video as approximate to the total intensity, as the laser is monochromatic green light.
-We can play this time series of intensity over time as an audio recording (we used the \textbf{scipy.io.wavfile.write}\cite{scipy} function in Python\cite{python}) and the tone is clearly heard.
-Taking the Fourier transform of this signal shows the injected tone, as seen in Figure~\ref{fig:webcam_spectrum} where we record the highest peak Fourier amplitude at $(2.099\pm 0.008)\mathrm{Hz}$ for an injected tone of 2.09Hz (a result within two standard deviations).
-Therefore, we’re able to observe a single tone quite confidently.
-
-\begin{figure}
+\begin{figure}[h]
 	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the highest peak is measured at $(2.099\pm 0.008) \mathrm{Hz}$ for an injected tone of 2.09Hz.
-In fact, the actual highest peak is at 0Hz, but that corresponds only to a non-zero average value, so we don’t display or consider it.}
+	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the amplitude 17.5 peak was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
+The actual maximum is achieved at 0Hz, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
 	\label{fig:webcam_spectrum}
@@
 
-Continuous gravitational wave searches often look for signals that wander (change) in frequency slowly over time\cite{2019PhRvD.100l2002A}.
-This means that finding the signal isn’t as simple as searching for a peak in the frequency content for the full observing run, especially when the signal-to-noise ratio is particularly low.
+As also mentioned in Section~\ref{sec:introduction}, continuous wave searches often look for signals that wander slowly in frequency~\cite{2019PhRvD.100l2002A}.
+The audio analogue of this is a tone that wanders in frequency; a note that changes pitch.
+Unlike the above Section~\ref{sec:single_tone} though, finding the signal isn’t as simple as searching for a peak in the spectrum for the full observing run.
 
-One method (among many) is to split the observed time series into bins of time, compute the Fourier transform for each time interval, and then create a grid (known as a spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time.
-The method then treats the problem as one of finding the best path through a weighted graph, where here the weights are given by the Fourier amplitudes.
-The method takes in a parameter of the allowed frequency wandering, some rate of frequency change over time, usually informed by a physical model.
-Here, this parameter is set to 0.11 Hz per second to make it simply one frequency bin per every time bin, which will prove poorly adjusted to the signal.
+One method (among many) for continuous wave searches is to split the observed signal into intervals of time, compute the Fourier transform for each time interval, and then create a grid (known as a spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time.
+This method turns the problem into finding the best path through the spectrogram grid as if it were a weighted graph, where here the weights are the Fourier amplitudes.
+
+When finding the path, the method requires a limit for the allowed frequency change over time, usually informed by a physical model.
+Here, this frequency wander limit was set to 0.11 Hz per second to make it simply one frequency bin per every time bin, which will prove poorly adjusted to the signal.
 
@@
 
-The Viterbi algorithm is a path-finding algorithm that solves this problem.
-Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns holding spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
+The Viterbi algorithm is a path-finding algorithm that is used in continuous wave searches~\cite{viterbi_application} to solve this problem.
+Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
 Then, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph.
@@
 
-We demonstrate this method of using the Viterbi algorithm to find a signal, by injecting a wandering frequency signal into the speaker and recording the output via the webcam.
+We demonstrate this method of using the Viterbi algorithm by injecting a wandering frequency signal into the speaker and recording the output via the webcam.
 Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
 
-Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, overlaid with the injected signal and the recovered Viterbi path.
+\begin{figure}[h]
+	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
+	\caption{Spectrogram of observed signal overlaid with pink dots for the injected signal and white crosses for the Viterbi path.
+Note the high correlation between them.
+Also note the deviations at the start time when the signal moves too quickly and just before the background anomaly at 150s.}
+	\label{fig:viterbi_overlay}
+\end{figure}
+
+Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, overlaid with the injected signal (the pink dots) and the recovered Viterbi path (the white crosses).
 To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment.
 Quantifying this, the Viterbi path stays within one frequency bin (approximately 0.3Hz) of the injected tone for 94\% of the run.
-Note that the signal initially moves faster than the chosen frequency wander parameter, causing a discrepancy between the injected and recovered signals.
-To fix this, the parameter should be doubled, although this would then cause the recovered path to stray more often.
+Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals.
+To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often.
 There also appears to be an anomaly at 150s, likely from a disturbance (such as someone walking) around the interferometer.
 
-Empirically, decreasing the signal-to-noise ratio closer to unity also causes the Viterbi algorithm to more easily stray from the signal to reach nearby fluctuations.
-However, the only correlation through the grid is the signal, so we always expect the Viterbi path to lie near to the signal no matter how high the noise is.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
-	\caption{Spectrogram of observed signal overlaid with white crosses for the Viterbi path and pink dots for the injected signal, notice the high correlation between them indicating that the signal is able to be recovered; also note the deviation at the start when the signal moves too quickly and just before the background anomaly at 150s}
-	\label{fig:viterbi_overlay}
-\end{figure}
+Empirically, decreasing the signal-to-noise ratio closer to unity (by making the speaker quieter) causes the Viterbi algorithm to stray more from the signal.
+However, since the only correlation through the grid is the signal, we always expect the Viterbi path to lie near to the signal no matter how high the noise is, and this is indeed what we see.
 
@@
 Compare a single tone that slowly, continuously changes frequency to music and speech which consist of a broad spectrum that is constantly changing.
-Instead, the interest here is in being able to record and play-back these complex audio signals, rather than track them.
-Using the interferometer to do so makes it, in ensemble with the measuring device, an ‘optical microphone’, using light to capture sound.
+Rather than track them, the interest here is in being able to record and play-back these complex audio signals.
+Using the interferometer to do so makes it an ‘optical microphone’, using light to capture sound.
+This device has precedence in the laser microphones~\cite{laser_microphone} of the defence industry which operate on a variety of related principles.
 
-Advanced signal processing is required to extract gravitational wave signals from noisy observational data.
-Recovering audio from an optical microphone requires related signal processing.
-Although far simpler, it serves as another useful demonstration.
+Although this is indirectly related to the advanced signal processing required to extract gravitational wave signals, using an interferometer to capture sound can serve as an interesting demonstration to a broader physics and engineering audience.
+As such this could be introduced as a separate demonstration.
 
@@
 
-For an optical microphone to be sensitive to speech and music it needs a frequency response (a sampling rate) far higher than the 30Hz of a standard webcam.
-Both speech and music need at least kHz range, with speech intelligibility requiring up to 3kHz and music preferably up to and beyond 8kHz\cite{speech_intelligibility}.
-This can be addressed by using a high-speed webcam, or more simply, by using a photodiode as the measuring device.
+For the optical microphone to be sensitive to speech and music it needs a frequency response (a sampling rate) far higher than the 30Hz of the webcam.
+In the context of audible frequencies (roughly 20Hz to 20kHz), speech intelligibility requires frequencies up to 3kHz and music requires frequencies up to and beyond 8kHz~\cite{speech_intelligibility}.
+Therefore the sampling rate of the optical microphone must be at least 16kHz to capture both speech and music.
+We achieved this with a photodiode.
 
-A photodiode is an electrical component that acts as a regular diode (blocking any current flow in the reverse direction) when no light is incident on it, but becomes more and more conductive in the reverse direction as the intensity of incident light increases.
+A photodiode is an electrical component that acts as a regular diode, blocking any current flow in the `reverse' direction, when no light is incident on it.
+But as the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.
 % Fundamentally, a current is created in the photodiode by the photo-electric effect.
-Placing a photodiode in reverse-bias over an op-amp creates a photo-detector that will produce a voltage dependent on the incident intensity.
-This experiment had the photodiode in the interference pattern at roughly the same spot as the webcam was monitoring previously, just off-centre in the position of the red dot in Figure~\ref{fig:interference_pattern}.
 
-The voltage signal from the photo-detector is captured by a 10-bit analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi Model 3 v1.2, in the standard configuration.
-This digitally samples the signal at roughly 16kHz.
-As such, any frequency component of the analog signal above the Nyquist frequency of 8kHz would be aliased down into the detected range.
-To prevent this, an anti-aliasing Sallen-Key filter\cite{sallen_key_filter} tuned to 16kHz is included before the ADC.
-This component attenuates any frequencies above 8kHz, before they get digitally sampled.
-It’s not immediately clear what’s limiting the sample rate to 16kHz, but it is likely non-optimal reading of the ADC by the Pi script.
+We placed a (OSRAM BPW21) photodiode in reverse-bias over an (LM358) op-amp to create a photo-detector that would produce a voltage dependent on the incident intensity.
+We placed the photodiode in the interference pattern at roughly the same off-centre spot as the webcam was monitoring previously (although we saw no difference when monitoring at other positions).
+The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.
+
+The voltage signal from the photo-detector was captured by a (MCP3008) 10-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2.
+Together, these sampled the signal at roughly 16kHz, our desired limit.
+
+However, by sampling, any frequency component of the analog signal above the Nyquist frequency of 8kHz would have been aliased down into the detected range.
+To prevent this, we included an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to 16kHz before the ADC.
+This component attenuated any frequencies above 8kHz, before they got digitally sampled.
+We also had to place another cloth screen over the face of the photodiode, as the reading without it, after gaining through the op-amps, was too high for the ADC.
+% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
 % The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
 
-The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}.
-The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the experiment, showing the interferometer, the circuit, and the Pi, is shown in Figure~\ref{fig:setup_pic2}.
-The only change to the interferometer schematic (Figure~\ref{fig:ifo_schematic_podo}) is the substitution of the webcam for the photodiode.
+A diagram and a photo of the circuit are shown in Appendix~\ref{app:circuit_diagram}.
+The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the entire optical microphone is shown in Figure~\ref{fig:setup_pic2}.
 
@@
 	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_photodiode.pdf}
-	\caption{Michelson interferometer schematic, a part of the interference pattern is recorded by a photodiode}
+	\caption{Michelson interferometer schematic with the intensity of the interference pattern at a point measured by a photodiode.
+See Figure~\ref{fig:ifo_schematic_webcam} for details on the optics configuration.}
 	\label{fig:ifo_schematic_podo}
@@
 	\includegraphics[width=\textwidth]{figures/setup_pic2.pdf}
-	\caption{Picture taken of the optical microphone set-up, showing the Michelson interferometer, the photodiode (in an insert), the circuit, and the Raspberry Pi.
-For a closer photo of the circuit, see Figure~\ref{fig:circuit_pic3}.}
+	\caption{Picture taken of the optical microphone set-up.
+Showing the Michelson interferometer, the photodiode (behind the cloth screen and in an insert), the circuit, and the Raspberry Pi.
+See Appendix~\ref{app:circuit_diagram} for a diagram and closer photo of the circuit.}
 	\label{fig:setup_pic2}
@@
 
-\subsection{Initial results}
+\subsection{Initial results of the optical microphone}
 % breakdown of all the filters used along the way
 
-The optical microphone was tested on recordings of speech from different people and on music ranging from simple melodies to full track songs.
-Each recording was around one minute long with processing often done on ten second segments.
-For each observation, the audio was played through the speaker with minimal activity around interferometer as the Pi script recorded the signal.
-The time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{scipy.io.wavfile.write}\cite{scipy} implementation in Python\cite{python}.
+The optical microphone was tested on varied recordings of speech from different people and on music ranging from simple melodies to full track songs.
+Each recording was around one minute long with processing done on ten second segments.
+For each observation, the audio was played through the speaker as the Pi script recorded the signal from the photodiode.
+Care was taken to have minimal activity around interferometer while this was happening.
+The time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.
 
-When played back and listen to, these initial recordings were extremely noisy and had a loud bass note throughout.
-Each Fourier spectrum showed a dominant AC mains hum with power ranging from the fundamental of 50Hz up to and beyond the 8th harmonic.
+When played back, these initial recordings were extremely noisy and had a loud bass hum throughout.
+Each Fourier spectrum showed dominant AC mains noise with power ranging from the fundamental of 50Hz up to and beyond the 8th harmonic thereof.
 With the speaker off this mains power was still present, as shown the power spectral density (effectively the norm squared of the Fourier spectrum) in Figure~\ref{fig:psd_noise}.
@@
 
-\begin{figure}%[H]
+\begin{figure}[h]%[H]
 	\includegraphics[width=\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
-	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), shows strong power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and cooling), otherwise is a fairly white spectrum except for the broad, shallow maximum at 750Hz}
+	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), shows strong power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and even cooling), but otherwise is a fairly white spectrum.}
+	% except for the broad, shallow maximum at 750Hz}
 	\label{fig:psd_noise}
@@
 
-Early attempts to remove the mains hum included zeroing the channel of each 50Hz harmonic in the Fourier spectrum.
-Effectively, this is multiplying the spectrum by a rectangular comb filter.
+The first attempt made to remove the mains hum was to zero the channel of each 50Hz harmonic in the Fourier spectrum.
+Effectively, this multiplied the spectrum by a rectangular comb filter.
 While this did remove the mains harmonics, it also audibly ruined the rest of the signal.
 This is because applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
-And the inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complex sinc-like function that when convolved with significantly damages the actual signal.
+And the inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function that when convolved with significantly damages the signal.
 
-To fix this, we tried using a smoother comb filter but this suffered from too significantly attenuating the rest of the spectrum between each harmonic.
+To fix this, we tried using a smoother comb filter but this too significantly attenuated the rest of the spectrum between each harmonic to hear the signal at all.
 
-Applying a high-pass filter to the signal worked well at removing the 50 and 100Hz tones but was problematic as the mains harmonics extend far past 100Hz.
-This region above 100Hz carries a lot of the fundamental frequencies for speech and music\cite{speech_intelligibility} and if the cut-off frequency is placed as to remove all the mains harmonics, then the resultant signal is unrecognisable.
+Applying a high-pass filter with cut-off frequency 150Hz to the signal worked well at removing the 50Hz and 100Hz tones but was problematic as the mains harmonics extend far past 100Hz and so remain after filtering.
+This region above 100Hz carries a lot of the fundamental frequencies for speech and music~\cite{speech_intelligibility} and if the cut-off frequency is placed as to remove more of the mains harmonics, then the resultant signal is unrecognisable.
 
@@
 
-Band-passing the signal similarly doesn’t address the remaining harmonics but does low-pass away other noise at higher frequencies.
-The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over a band from 150Hz to 3kHz (beyond 2kHz the information content drops\cite{speech_intelligibility}).
+\begin{figure}[h]%[H]
+	\includegraphics[width=0.8\textwidth]{figures/butterworth_150_3000-cropped.pdf}
+	\caption{Butterworth bandpass filter frequency response, any amplitude beyond -3dB is significant attenuation (half power).
+The green vertical lines show the band limits of 150Hz and 3kHz.
+Note the flat response within the band.}
+	\label{fig:butterworth}
+\end{figure}
+
+Band-passing the signal similarly doesn’t address the remaining harmonics above 100Hz but does low-pass away other noise at higher frequencies.
+The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over a band from 150Hz to 3kHz (beyond 2kHz speech information content drops significantly~\cite{speech_intelligibility}).
 For a low-pass Butterworth filter at cut-off frequency $\omega_c$ and gain $\epsilon$ the frequency response (attenuation at each frequency) follows Equation~\ref{eq:butterworth}.
@@
 \end{equation}
-For a band-pass filter, the above low-pass is combined with a high-pass filter of similar form.
+For a Butterworth band-pass filter, the above low-pass filter is combined with a high-pass filter of similar form.
 The response of the filter used is shown in Figure~\ref{fig:butterworth}.
-Despite improving, the recording with the filter applied still had very low intelligibility, it could be recognised as speech but was not understandable.
+This filter improved on the above simpler filters but recordings of speech still had very low intelligibility, they could be recognised as speech but not understood.
 
-The ideal filter would be one that completely attenuates the undesired parts of the spectrum, has no changing attenuation (is flat) on the rest of the spectrum, and has the smoothest edges as to not damage the signal.
+The ideal filter would be one that completely attenuates the undesired parts of the spectrum, does not change the rest of the spectrum, and has smooth edges as to not damage the signal through convolution.
 These three conditions are contradictory and any filter must compromise between them.
 The above simple filters fail at one or more of each of these conditions.
-The combs ruin the signal and the high-passes don’t remove the noise.
-The Butterworth filter performed the best audibly as it was maximally flat within the band region, prioritising the second condition.
+The combs ruin the signal and the high-passes do not remove the noise.
+The Butterworth filter performed the best audibly as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
 
@@
 Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-
-\begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{figures/butterworth_150_3000-cropped.pdf}
-	\caption{Butterworth bandpass filter frequency response, green vertical lines show the band limits of 150Hz and 3kHz.
-Note the flat response within the band, any amplitude beyond -3dB is significant attenuation (half power).}
-	\label{fig:butterworth}
-\end{figure}
 
@@
 This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
-Although this technique historically minimised with respect to the spectral amplitude\cite{EphraimSpeechEU_MMSE}, throughout speech processing use of the logarithm seems produce better results as perceived by the human ear\cite{SubjectiveComparison}.
+Although this technique historically minimised with respect to the spectral amplitude~\cite{EphraimSpeechEU_MMSE}, throughout speech processing use of the logarithm seems produce better results as perceived by the human ear~\cite{SubjectiveComparison}.
 
-We use an existing implementation\cite{logmmse} of the logMMSE estimator.
-Applying this estimator to the optical microphone recordings produces dramatically clean results.
+We used an existing implementation~\cite{logmmse} of the logMMSE estimator.
+Applying this estimator to the optical microphone recordings produced dramatically clean results.
 The time series (in Figure~\ref{fig:logMMSE_timeseries}) and spectrum (in Figure~\ref{fig:logMMSE_spectrum}) from before and after applying the logMMSE filter are shown.
-Comparing the spectra in Figure~\ref{fig:logMMSE_spectrum} shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.
+Comparing the before and after spectra in Figure~\ref{fig:logMMSE_spectrum} shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.
 
-For recordings of music listened to after the estimator is applied, simple chords and drums are heard well, while more complex melodies are often completely absent in the results.
+After applying the estimator, simple chords and drums are heard well in played back music.
+While more complex melodies are sometimes completely absent.
 By observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
-This could be a perceptual effect of listening to the lossy recording, but it seems more likely to be a frequency dependence somewhere in the optical microphone system.
-Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies.
+This could be a perceptual effect of listening to the lossy recording, but seems more likely to be a frequency dependence somewhere in the optical microphone system.
+Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies and so instruments like electric bass and drums are louder in the results.
 
 After applying the estimator to speech recordings, intelligibility still remains low with the voice sounding mumbled and indistinct.
-The estimator removes most of the background sound, but does not clarify the speech.
-% There is a question here as to where the diction in the speech is lost, at the speaker-mirror coupling, at the photodiode reading, or if it is still there, just obscured in the noisy signal.
+The estimator removes most of the background noise, but does not clarify the speech.
+
+To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
+This requires a better understanding of the system and further improvements to the experiment if true.
 % The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
@@
 	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
-	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) logMMSE filter is applied, only one second of the one minute recording is shown.
-The rise at the start of the time series is an expected result from filtering a finite signal.}
+	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) the logMMSE estimator is applied, showing one second of the one minute recording.
+The rise at the start of the bottom	 time series is an expected result from filtering a finite signal.}
 	\label{fig:logMMSE_timeseries}
@@
 	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
-	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) logMMSE filter is applied, showing only up to 2kHz and a detail of the spectrum (limited to show only up to amplitude 2, while before filtering the 50Hz line has an amplitude of 40)}
+	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) the logMMSE estimator is applied, showing only up to 2kHz and to amplitude 2 to see greater detail.}
 	\label{fig:logMMSE_spectrum}
@@
 % wild speculation
-For better speech enhancement, we are currently working towards two future advances.
+We are currently working towards two future advancements for the optical microphone.
 
-The first would be to better isolate the optical microphone from the mains supply.
+The first would be to better isolate the experiment from the mains supply.
 The Raspberry Pi is USB powered and requires low power so it could even be run off a commercial external phone battery.
@@
 The second advancement would be to achieve a full model of the transfer function of the system.
-Starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading, see Figure~\ref{fig:pipeline_highlighted} for this full sequence.
 Theoretically, if we had the correct transfer function, then signal recovery would only be limited by the signal-to-noise ratio as we could construct an optimal matched filter.
+This function would have to start from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
+This sequence that the signal takes is shown in Figure~\ref{fig:pipeline_highlighted}.
 
-In our investigation, we have currently only derived part of this sequence, the relation between a deflection of the mirror and the change in the intensity of the interference pattern at some point (highlighted in Figure~\ref{fig:pipeline_highlighted}).
-We find this relation to be non-linear, see Appendix~\ref{app:intensity_derivation} for the full discussion and derivation.
+In our investigation, we have currently only derived part of this sequence, the relation between a deflection of the mirror and the change in the intensity of the interference pattern at some point.
+We find this relation to be non-linear, see Appendix~\ref{app:intensity_derivation} for the derivation.
 
-Asides from these two main avenues, other techniques are available to better process the recording.
-In particular, for extracting the speech there exists much literature\cite{HMM_english} on using a hidden Markov model train to English phonemes to recognise speech.
-Alternatively, machine learning solutions also exist throughout the field that are able to compete\cite{HMM_english} with statistical techniques like the logMMSE estimator.
-More simply, to remove any remaining mains hum we could monitor it separately from the interferometer and subtract the sinusoid with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
-This could also be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
+Asides from the transfer function, for extracting the speech there exists much literature~\cite{HMM_english} on using a hidden Markov model train to English phonemes to recognise speech.
+Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
+
+To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
+This could be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
 
@@
 	\includegraphics[width=\textwidth]{figures/pipeline_highlighted.pdf}
-	\caption{Sequence of relations for optical microphone transfer function}
+	\caption{Sequence of relations for optical microphone transfer function. The currently derived relation is highlighted.}
 	\label{fig:pipeline_highlighted}
@@
 
+\newpage
 \section{Conclusions}
@@
 
-We use sound and a table-top Michelson interferometer as an analogue to gravitational wave detectors to demonstrate signal processing techniques.
+We use sound and a table-top Michelson interferometer as an analogue to a gravitational wave detector.
+% This serves to demonstrate signal processing techniques.
+
 Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer.
@@
 
-Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the intensity time series of the interference pattern into an audio recording and the interferometer into an optical microphone.
-We demonstrated that this optical microphone is capable of playing back simple recordings well and speech recordings unintelligibly.
+Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the interferometer into an optical microphone.
+We demonstrated that this optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech.
 Dominant 50Hz mains hum in the recordings led to an exploration of speech enhancement techniques to remove it, including the well performing logMMSE estimate.
-From here we currently have many plans to further remove the mains hum and improve the intelligibility.
+From here we have many plans to further remove the mains hum and improve the intelligibility.
 
-These demonstrations have the potential to be adapted into the undergraduate laboratory.
-There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches, but also in the signal processing of speech enhancement as an open problem for engineering focused students.
-See Appendix~\ref{app:reproducing_results} for advice aimed at the undergraduate student for replicating the results.
+These demonstrations have good potential to be adapted into the undergraduate laboratory.
+There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches.
+Most of all, the signal processing of speech enhancement is an interesting problem that’s relevant to digital communications.
 
+% \bigskip
 \newpage
@@
 \label{app:code}
-This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.
+This project was implemented in python3~\cite{python} scripts as well as jupyter notebooks~\cite{jupyter}~\cite{ipython} and used the packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
 
-The current build can be found at:
+The current build and sample recordings of the optical microphone can be found at:
 \url{https://github.com/daccordeon/gravexplain}
@@
 
-\section{Additional guide to reproducing results}
-\label{app:reproducing_results}
-% as for a 3rd year undergraduate
-
-Besides the directions given in the rest of this article, there are some observations and technical details useful to someone wanting to reproduce the results shown herein.
-For the analysis and signal processing performed the best reference is Appendix~\ref{app:code} for the open-source code.
-
-For the experiment, since a Michelson interferometer is such a standard configuration, we’ll not detail it further here, see any optics manual for more information.
-To obtain an independent speaker cone we recommend carefully dismantling a budget commercial speaker system to get at the speaker within.
-To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place.
-If the other speaker in the pair remains, keep it as far away from the experiment and as damped as possible else it may act as a second, interfering source.
-% we used a Class 2 LDS5 ThorLabs laser at 532nm
-
-For the photodiode circuit, we used a Raspberry Pi and an MCP3008 ADC to monitor the analog signal.
-This could also be done with the analog pins on an Arduino.
-Be careful of the gain through the LM358 op-amps in the circuit though, else the ADC in either case could threshold the signal.
-Also be ready to change the values in the Sallen-Key filter to match your own sampling rate.
-The photodiode was an OSRAM BPW21, we mounted it through a discarded piece of speaker mesh (similar to black cloth over a frame) and placed another piece over the face to not over expose and threshold the signal.
-For the rest of the circuit, we used standard components assembled on a breadboard as detailed in Appendix~\ref{app:circuit_diagram}.
-
-\section{The Viterbi algorithm}
+\section{Implementation of the Viterbi algorithm}
 \label{app:viterbi}
 
-To find the Viterbi path through a spectrogram (spectra over time), given an M x N grid of M frequency rows and N time columns and a rate limit for the change of frequency over time that we convert into the number of allowed rows we can move while stepping from each column to the next (this limit was $\pm 1$ row/column in our case).
-The path must start from the first column and end at the last, taking a single step in each column along the way.
-Each element in the grid is the norm of the Fourier component of a frequency at a time.
-These are normalised between $(0, 1)$ to use as multiplicative weights, i.e.\ the total weight of a path is the product of all weights along that path.
+We used the Viterbi algorithm to find a signal in a spectrogram. Here we detail the implementation used.
+
+We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
+These are normalised between $(0, 1)$ to use as multiplicative weights.
 In fact, we take the logarithm and so the sum of all weights to avoid underflow.
-The best path has the highest weight.
+The best path has the highest weight among all paths.
 
-With the grid prepared, the Viterbi algorithm starts in the second column.
-In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, taking the greatest cumulative weight of those available, ‘nearby’ cells (the value of which was discovered in the previous iteration).
-For each cell in the current column, the algorithm notes this weight plus the cell’s own weight, along with the row index of which cell in the previous column it came from.
+With the grid prepared, the algorithm starts in the second column.
+In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, selecting the previous cell with the greatest cumulative weight among those ‘nearby’ (the value of which was discovered in the previous iteration).
+Where the cumulative weight is the weight of the best path to reach that previous cell.
+The nearness of a previous cell is given by the limit on the allowed frequency change over time.
+
+For each cell in the current column, the algorithm calculates the cumulative weight to reach that cell by adding its weight to the cumulative weight to reach it.
+The algorithm also notes the row index of which cell in the previous column it best connects to.
 Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
-That best path is the Viterbi path.
+That discovered best path is the Viterbi path.
 
-\section{A common mistake with wandering frequency signals}
+\section{A common mistake for those new to wandering frequency signals}
 \label{app:phase_gotcha}
@@
 We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
-The common mistake is to just appeal to the standard form $\sin{(2 \pi f t)}$, where f would be a constant frequency, and write down $\sin{(2 \pi f(t) t)}$.
-However, accumulation of phase produces the wrong curve.
-Instead, one should see that the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
+A common mistake is to appeal to the standard form $\sin{(2 \pi f t)}$, where f is a constant frequency, and guess $\sin{(2 \pi f(t) t)}$.
+However, this fails to address the phase accumulated earlier in the signal.
+Instead, the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
 Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
+% This formula is also useful when performing frequency modulation.
 
-\section{Circuit design}
+\section{Photodiode circuit design}
 \label{app:circuit_diagram}
@@
 A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
+This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
 
-\begin{figure}%[H]
+\begin{figure}[h]%[H]
 	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
@@
 	\includegraphics[width=0.8\textwidth]{figures/circuit_pic3.pdf}
-	\caption{Photo of circuit for photodiode reading assembled on breadboard.}
+	\caption{Photo of photodiode circuit assembled on breadboard.}
 	\label{fig:circuit_pic3}
@@
 
-\section{Intensity relation derivation}
+\section{Interferometric intensity relation derivation}
 \label{app:intensity_derivation}
 
-The intensity, $I$, of the interference pattern at a particular angle is given by Equation~\ref{eq:intensity_result} below.
-Where A is the amplitude of the electric field of the beam, d is the separation (or the difference in the arm lengths), $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Appendix~\ref{app:intensity_derivation} for the derivation to get to this equation).
-\begin{equation}
-\label{eq:intensity_result}
-I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{equation}
+To obtain part of the transfer function for the optical microphone, we want to know the resultant change in intensity, $\delta I$, at a point in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
+Let the point be at angle $\theta$ on the screen (in the far-field) above the central ray as measured from the mirrors.
+Noting circular symmetry as to not specify exactly where in the fringe the point is.
 
-This Equation~\ref{eq:intensity_result} means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
-This makes the problem of extracting the original signal more difficult.
+At this point, two rays meet, one from each mirror, with total electric field amplitude $E_1(t) + E_2(t)$.
+Given an initial difference between the arm lengths of $d$, the path difference between the two beams at the screen is $2 d \cos{\theta}$ as the longer beam must travel back and forth over the extra distance, and both travel at angle $\theta$ to meet the point~\cite{fringes:online}.
 
-\begin{equation}
-\label{eq:intensity_change}
-	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{equation}
-
-We want the change in intensity, $\delta I$, at a point in the Michelson interference pattern, given a small change, $\delta d$, in the perpendicular path difference of the mirrors (or in the separation of one and virtual image of the other).
-The point is at angle $\theta$ on the screen above the central ray, noting circular symmetry.
-At this point, two rays meet, one from each mirror, with the total electric field (\jam{parallel to the screen?}) being $E_1(t) + E_2(t)$.
-Given an initial separation of $d$, the path difference between the mirrors is $2 d \cos{\theta}$ as the longer ray must travel to and from the further mirror, and both move up at the angle $\theta$ to meet at the screen.
+The intensity at the point is then the norm squared of the total electric field amplitude as expressed in Equation~\ref{eq:intensity_derivation} below.
 
 \begin{align}
-\label{eq:intensity_derivation}
     E_1(t) &= A e^{i (k L - \omega t)} \\
@@
     I(t) &= \lvert A e^{i (k L - \omega t)} + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2
+\label{eq:intensity_derivation}    
 \end{align}
 
-Collapse the phase difference to $\phi$ and expand the norm squared by expanding the complex exponentials.
-From there is it trigonometric manipulation to reach the result of intensity in terms of separation.
+Collapse the phase difference to some value $\phi = 2 k d \cos{\theta}$.
+Then expand the norm squared using Euler’s relation, $e^{i \theta} = \cos{\theta} + i \sin{\theta}$.
+Here we stop considering intensity as a function of time as the time dependence will vanish.
 
 \begin{align}    
-    I(t) &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
-    I(t) &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2) \\
-    I(t) &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
-    I(d) &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+    I &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
+    I &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2)
 \end{align}
 
-Then take derivatives and find that the intensity change is non-linear.
+Expand and collect like terms, pull out a factor of two, and then recognise the form of $\cos{(\alpha-\beta)}$.
+
+\begin{align}
+    I &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
+    I &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+\end{align}
+
+Finally, changing derivatives and expand $\phi$ to find that the change in intensity given a change in distance is non-linear, that ${\delta I}/{\delta d}$ is non-constant.
 
@@
 \end{align}
+
+This non-linearity in part of the transfer function will make the reverse problem of extracting the original signal more difficult.
 

### 2020-04-14 paper/ifo-appendix.tex
some editing throughout the paper draft

--- 2020-02-14 paper/paper_main.tex
@@
-% GravExplain
-\documentclass[prb,preprint]{revtex4-1} 
-% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
-
-% Note that AJP uses the same style as Phys. Rev. B (prb).
-% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
-% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
-% We look forward to receiving your submission to AJP.
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb,amsthm}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{mathtools}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{hyperref}
-% \usepackage{siunitx}
-\usepackage{textcomp}
-
-\bibliographystyle{apsrev4-2}
-% \setlength{\parindent}{0pt}
-
-\newcommand{\jam}{\textcolor{magenta}}
-\newcommand{\han}{\textcolor{orange}}
+\documentclass[paper-main.tex]{subfiles}
 
 \begin{document}
-
-% cover page
-\section*{Project contributions}
-% contributions clarify what’s to be assessed
-James W. Gardner:
-\begin{itemize}
-\item all paper sections other than the introduction, i.e.\ the abstract and from Section~\ref{sec:ifo} onwards, and all figures in the paper
-\item addition of the injection speaker, design of photodiode circuit from standard examples, assembly of circuit on breadboard, and set-up of Raspberry Pi to measure output
-\item all of the source code at \url{https://github.com/daccordeon/gravexplain} (containing over 1000 lines of Python scripts), that reads the data from the Raspberry Pi, performs all of the analysis on it, and creates all the figures used in the paper
-\item presentation of two talks based on the paper and the figures therein
-\end{itemize}
-
-Hannah Middleton:
-\begin{itemize}
-\item the introduction of the paper, i.e.\ Section~\ref{sec:introduction}
-\item design and assembly of the interferometer
-\item continuous support and advice throughout the project
-\end{itemize}
-
-Andrew Melatos, Robin Evans, William Moran: other vital advice and comments throughout the project
-
-\newpage
-\section*{Statement of achievements}
-% one paragraph recount of what’s been done for two months, like the abstract
-The project spanned three weeks in December, four weeks in January, and two weeks in February.
-The first two weeks were spent writing the initial scripts and achieving the Viterbi results with the webcam set-up.
-This was relatively straight forward and the greatest hurdle was understanding the algorithm.
-
-The third week was spent setting up the photodiode circuit and the Raspberry Pi which involved constant testing and changing of the components and design of the circuit.
-All of January was spent on the optical microphone.
-Much difficultly was had combating 50Hz mains hum in the experiment.
-Eventually through many discussions at the signal processing meetings, we had a working optical microphone but low speech intelligibility.
-The last week of January was spent on methods of speech enhancement.
-If the investigation had continued we had much future work planned and possible solutions in mind.
-
-Finally, the start of February was spent writing the paper and giving presentations (one nine minute general talk and one 50 minute technical follow-up talk).
-
-\newpage
-
-\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
-
-\author{James W. Gardner}
-\email{u6069809@anu.edu.au}
-\affiliation{College of Science, Australian National University, Acton, ACT, 2601, Australia}
-
-\author{Hannah Middleton$^{(i)}$}
-\email{hannah.middleton@unimelb.edu.au}
-\author{Andrew Melatos$^{(i)}$}
-\email{amelatos@unimelb.edu.au}
-\author{Robin Evans$^{(ii)}$}
-\email{robinje@unimelb.edu.au}
-\author{William Moran$^{(ii)}$}
-\email{wmoran@unimelb.edu.au}
-\affiliation{School of Physics$^{(i)}$ and Department of Electrical and Electronic Engineering$^{(ii)}$, University of Melbourne, Parkville, VIC, 3010, Australia}
-\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}
-
-% Summer 2019/2020
-\date{\today}
-
-% AJP requires an abstract for all regular article submissions.
-\begin{abstract}
-Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings.
-However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
-
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-\begin{figure}
-	\includegraphics[width=0.9\textwidth]{figures/ligo_pic.pdf}
-	\caption{Aerial photo of the LIGO Hanford facility and its 4km long beam arms, courtesy of LIGO/Caltech/MIT}
-	\label{fig:ligo_pic}
-\end{figure}
-
-% general gw intro
-In 2015, gravitational waves were observed for the first time from the merger of two black holes in a binary system~\cite{GW150914}. 
-The observation, made by the Laser Interferometer Gravitational wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
-Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years, there has been increased public interest in gravitational wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to explain the topic to a general audience.
-Activities range from hands-on demonstrations, exhibitions, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online}, video games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
-
-% what are gws and how do we detect them...
-Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are disturbances in space-time caused by the acceleration of massive objects. 
-The effect of gravitational waves is a change in lengths; a `stretching and squeezing' of the very distance between objects.
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance.
-As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
-These observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets}.
-
-% continuous gws
-To date, while the network of gravitational wave observatories have only observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}, these observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencies.
-These continuous waves may emit at constant frequencies, or they may wander slowly in frequency over time. 
-One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
-LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
-The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
-
-% sound analogy
-Here, we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience; the analogy of listening to the universe is often used to explain gravitational wave observations.
- %; transient gravitational wave signals have been converted into audio.
-This can be demonstrated with a table-top interferometer by playing sound close enough to move the mirrors and change the interference pattern.
-Using audio of slowly wandering frequency tones to produce an analogue of a continuous gravitational wave signal.
-The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}, particularly for the wandering frequency demonstration.
-
-In this work we detail our table-top interferometer design in Section~\ref{sec:ifo} and describe observing a single note from a speaker in Section~\ref{sec:single_tone}.
-In Section~\ref{sec:viterbi_wandering}, we demonstrate the analysis techniques for observing a wandering frequency signal.
-In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech.
-This demonstration of an `optical microphone' serves as a more general exhibition of signal processing.
-We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
-
-% \section{Method and results}
-\section{Table-top gravitational wave science}
-\label{sec:ifo}
-
-Even though gravitational waves are caused by some of the most massive events in the universe, their observable effects on Earth are vanishingly small | on the order of only one part in $10^{21}$ for our best detectors~\cite{GW150914}.
-Clearly, the measurement of such effects is far beyond any table-top experiment.
-But, the fundamental principle of gravitational wave detection remains that changes in the lengths of the beam arms of an interferometer leads to a change in the interference pattern observed.
-By observing sound waves that vibrate the mirrors of a table-top interferometer and so change the lengths of the beam arms we have an analogue of this principle.
-
-\subsection{Method of Michelson interferometry to observe sound}
-% hardware: path difference, virtual thin film interference
-
-In a Michelson interferometer, a coherent (monochromatic and equal phase) source of laser light is split evenly by a beamsplitter into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
-The beams recombine and interfere at the beamsplitter before hitting a screen to produce a pattern of concentric, circular fringes.
-The configuration used in this experiment is shown diagrammatically in Figure~\ref{fig:ifo_schematic_webcam} and as part of a photo in Figure~\ref{fig:setup_pic2}.
-
-A sample interference pattern is shown in Figure~\ref{fig:interference_pattern}.
-This pattern is due to the path difference between the two arms of unequal lengths (if the lengths were equal, then only a single, bright dot would be seen).
-This results in a phase difference between the two beams when they meet at the screen.
-This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
-For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-When the path difference changes, the fringes stay circular but move radially in or out.
-
-\begin{figure}[h]
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_webcam.pdf}
-	\caption{Michelson interferometer schematic with interference pattern recorded by a commercial webcam.
-We used a Class 2 532nm laser~\cite{ThorLabsIFO} and a commercial speaker.
-The beam is focused at a converging lens before the 50/50 beamsplitter and the interference pattern is made visible by a diverging lens after the beamsplitter.
-The arms lengths between the speakers are roughly 3cm different and the entire interferometer (without the screen) fits within the dimensions of an A4 piece of paper.
-We did not include a compensating plate with the beamsplitter, as is often done.}
-	\label{fig:ifo_schematic_webcam}
-\end{figure}
-
-We used a 0.5W speaker stuck to the back of one of the mirrors as an audio source for our demonstration.
-This speaker was part of a commercial 3.5mm-jack, USB powered pair of speakers and was driven by a PC.
-To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place.
-We kept the other speaker in the pair face-down and as far away from the experiment as the cables would allow to reduce it acting as a second source.
-
-As the speaker deforms to play sound, the mirror moves with a change in separation as some coupling function of the amplitude and frequency of the sound.
-The motion of the fringes follows this motion of the mirror and therefore the speaker at the same frequency.
-If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
-If the fringe motion was large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting would artificially raise the measured frequency.
-As such, any motion of the fringes must be kept small by playing sound softly through the speaker.
-
-For later sections, we derive the non-linear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen, this can be found in Appendix~\ref{app:intensity_derivation}.
-
-% Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
-We did not find it necessary to include it to be able to demonstrate the results desired.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/webcam_still0.pdf}
-	\caption{Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
-For scale, the central fringe is approximately 5mm across}
-	\label{fig:interference_pattern}
-\end{figure}
-
-% A note that while the interference pattern appears similar to the Airy pattern produced by a circular aperture, they are not in fact the same geometry.
-
-\section{Observing a single tone}
-\label{sec:single_tone}
-% the webcam method
-% how to analyse video, python OpenCV2
-
-As mentioned in Section~\ref{sec:introduction}, continuous gravitational wave searches look for near monochromatic signals. The audio analogue of this is a simple sinusoidal tone; a constant note.
-
-We measured the intensity at a point on the screen with a commercial USB webcam placed as to view the screen as in Figure~\ref{fig:ifo_schematic_webcam}.
-This webcam samples at 30Hz, which limits the spectral content of signals we can observe to below 15Hz, the Nyquist frequency.
-% See later Section~\ref{sec:photodiode} where we raise this limit into audible frequencies (20 Hz to 20kHz) with a photodiode.
-While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
-Then, at an off-centre pixel (shown as a red dot in Figure~\ref{fig:interference_pattern}), we took the green-channel of the video as approximate to the total intensity, as the laser is monochromatic green light (at 532nm).
-
-We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}.
-The two tones played together sound similar and don’t produce any noticeable beating.
-To quantify the precision, we recorded a signal from an injected tone of 2.09Hz.
-From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\mathrm{Hz}$, a result within two standard deviations.
-This showed that the experiment observes a single tone confidently.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the amplitude 17.5 peak was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
-The actual maximum is achieved at 0Hz, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
-	\label{fig:webcam_spectrum}
-\end{figure}
-
-\section{Observing a wandering tone}
-\label{sec:viterbi_wandering}
-% Tracking tones with the Viterbi algorithm
-% explaination and successful application
-% lean on the continuous gravitational waves angle
-
-As also mentioned in Section~\ref{sec:introduction}, continuous wave searches often look for signals that wander slowly in frequency~\cite{2019PhRvD.100l2002A}.
-The audio analogue of this is a tone that wanders in frequency; a note that changes pitch.
-Unlike the above Section~\ref{sec:single_tone} though, finding the signal isn’t as simple as searching for a peak in the spectrum for the full observing run.
-
-One method (among many) for continuous wave searches is to split the observed signal into intervals of time, compute the Fourier transform for each time interval, and then create a grid (known as a spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time.
-This method turns the problem into finding the best path through the spectrogram grid as if it were a weighted graph, where here the weights are the Fourier amplitudes.
-
-When finding the path, the method requires a limit for the allowed frequency change over time, usually informed by a physical model.
-Here, this frequency wander limit was set to 0.11 Hz per second to make it simply one frequency bin per every time bin, which will prove poorly adjusted to the signal.
-
-\subsection{The Viterbi algorithm}
-
-The Viterbi algorithm is a path-finding algorithm that is used in continuous wave searches~\cite{viterbi_application} to solve this problem.
-Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
-Then, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph.
-At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path.
-See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.
-
-\subsection{Results}
-
-We demonstrate this method of using the Viterbi algorithm by injecting a wandering frequency signal into the speaker and recording the output via the webcam.
-Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
-	\caption{Spectrogram of observed signal overlaid with pink dots for the injected signal and white crosses for the Viterbi path.
-Note the high correlation between them.
-Also note the deviations at the start time when the signal moves too quickly and just before the background anomaly at 150s.}
-	\label{fig:viterbi_overlay}
-\end{figure}
-
-Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, overlaid with the injected signal (the pink dots) and the recovered Viterbi path (the white crosses).
-To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment.
-Quantifying this, the Viterbi path stays within one frequency bin (approximately 0.3Hz) of the injected tone for 94\% of the run.
-Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals.
-To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often.
-There also appears to be an anomaly at 150s, likely from a disturbance (such as someone walking) around the interferometer.
-
-Empirically, decreasing the signal-to-noise ratio closer to unity (by making the speaker quieter) causes the Viterbi algorithm to stray more from the signal.
-However, since the only correlation through the grid is the signal, we always expect the Viterbi path to lie near to the signal no matter how high the noise is, and this is indeed what we see.
-
-\section{Playing back complex audio}
-\label{sec:optical_microphone}
-% Optical microphone
-% aim of optical microphone: play back injected sounds
-
-The natural succession to observing simple tones is to study complex audio signals, such as music and speech.
-The Viterbi algorithm and tracking more broadly is not applicable to these signals.
-Compare a single tone that slowly, continuously changes frequency to music and speech which consist of a broad spectrum that is constantly changing.
-Rather than track them, the interest here is in being able to record and play-back these complex audio signals.
-Using the interferometer to do so makes it an ‘optical microphone’, using light to capture sound.
-This device has precedence in the laser microphones~\cite{laser_microphone} of the defence industry which operate on a variety of related principles.
-
-Although this is indirectly related to the advanced signal processing required to extract gravitational wave signals, using an interferometer to capture sound can serve as an interesting demonstration to a broader physics and engineering audience.
-As such this could be introduced as a separate demonstration.
-
-\subsection{The photodiode method}
-\label{sec:photodiode}
-% advanced method: explain how we capture data
-
-For the optical microphone to be sensitive to speech and music it needs a frequency response (a sampling rate) far higher than the 30Hz of the webcam.
-In the context of audible frequencies (roughly 20Hz to 20kHz), speech intelligibility requires frequencies up to 3kHz and music requires frequencies up to and beyond 8kHz~\cite{speech_intelligibility}.
-Therefore the sampling rate of the optical microphone must be at least 16kHz to capture both speech and music.
-We achieved this with a photodiode.
-
-A photodiode is an electrical component that acts as a regular diode, blocking any current flow in the `reverse' direction, when no light is incident on it.
-But as the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.
-% Fundamentally, a current is created in the photodiode by the photo-electric effect.
-
-We placed a (OSRAM BPW21) photodiode in reverse-bias over an (LM358) op-amp to create a photo-detector that would produce a voltage dependent on the incident intensity.
-We placed the photodiode in the interference pattern at roughly the same off-centre spot as the webcam was monitoring previously (although we saw no difference when monitoring at other positions).
-The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.
-
-The voltage signal from the photo-detector was captured by a (MCP3008) 10-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2.
-Together, these sampled the signal at roughly 16kHz, our desired limit.
-
-However, by sampling, any frequency component of the analog signal above the Nyquist frequency of 8kHz would have been aliased down into the detected range.
-To prevent this, we included an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to 16kHz before the ADC.
-This component attenuated any frequencies above 8kHz, before they got digitally sampled.
-We also had to place another cloth screen over the face of the photodiode, as the reading without it, after gaining through the op-amps, was too high for the ADC.
-% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-
-A diagram and a photo of the circuit are shown in Appendix~\ref{app:circuit_diagram}.
-The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the entire optical microphone is shown in Figure~\ref{fig:setup_pic2}.
-
-\begin{figure}
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_photodiode.pdf}
-	\caption{Michelson interferometer schematic with the intensity of the interference pattern at a point measured by a photodiode.
-See Figure~\ref{fig:ifo_schematic_webcam} for details on the optics configuration.}
-	\label{fig:ifo_schematic_podo}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/setup_pic2.pdf}
-	\caption{Picture taken of the optical microphone set-up.
-Showing the Michelson interferometer, the photodiode (behind the cloth screen and in an insert), the circuit, and the Raspberry Pi.
-See Appendix~\ref{app:circuit_diagram} for a diagram and closer photo of the circuit.}
-	\label{fig:setup_pic2}
-\end{figure}
-
-\subsection{Initial results of the optical microphone}
-% breakdown of all the filters used along the way
-
-The optical microphone was tested on varied recordings of speech from different people and on music ranging from simple melodies to full track songs.
-Each recording was around one minute long with processing done on ten second segments.
-For each observation, the audio was played through the speaker as the Pi script recorded the signal from the photodiode.
-Care was taken to have minimal activity around interferometer while this was happening.
-The time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.
-
-When played back, these initial recordings were extremely noisy and had a loud bass hum throughout.
-Each Fourier spectrum showed dominant AC mains noise with power ranging from the fundamental of 50Hz up to and beyond the 8th harmonic thereof.
-With the speaker off this mains power was still present, as shown the power spectral density (effectively the norm squared of the Fourier spectrum) in Figure~\ref{fig:psd_noise}.
-And even with the speaker off and the photodiode in complete darkness, the mains hum was still present (with distribution similar to but weaker than that shown in Figure~\ref{fig:psd_noise}).
-
-\begin{figure}[h]%[H]
-	\includegraphics[width=\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
-	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), shows strong power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and even cooling), but otherwise is a fairly white spectrum.}
-	% except for the broad, shallow maximum at 750Hz}
-	\label{fig:psd_noise}
-\end{figure}
-
-\subsection{Applying filters to the data}
-\label{sec:simple_filters}
-
-The first attempt made to remove the mains hum was to zero the channel of each 50Hz harmonic in the Fourier spectrum.
-Effectively, this multiplied the spectrum by a rectangular comb filter.
-While this did remove the mains harmonics, it also audibly ruined the rest of the signal.
-This is because applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
-And the inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function that when convolved with significantly damages the signal.
-
-To fix this, we tried using a smoother comb filter but this too significantly attenuated the rest of the spectrum between each harmonic to hear the signal at all.
-
-Applying a high-pass filter with cut-off frequency 150Hz to the signal worked well at removing the 50Hz and 100Hz tones but was problematic as the mains harmonics extend far past 100Hz and so remain after filtering.
-This region above 100Hz carries a lot of the fundamental frequencies for speech and music~\cite{speech_intelligibility} and if the cut-off frequency is placed as to remove more of the mains harmonics, then the resultant signal is unrecognisable.
-
-We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass.
-
-\begin{figure}[h]%[H]
-	\includegraphics[width=0.8\textwidth]{figures/butterworth_150_3000-cropped.pdf}
-	\caption{Butterworth bandpass filter frequency response, any amplitude beyond -3dB is significant attenuation (half power).
-The green vertical lines show the band limits of 150Hz and 3kHz.
-Note the flat response within the band.}
-	\label{fig:butterworth}
-\end{figure}
-
-Band-passing the signal similarly doesn’t address the remaining harmonics above 100Hz but does low-pass away other noise at higher frequencies.
-The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over a band from 150Hz to 3kHz (beyond 2kHz speech information content drops significantly~\cite{speech_intelligibility}).
-For a low-pass Butterworth filter at cut-off frequency $\omega_c$ and gain $\epsilon$ the frequency response (attenuation at each frequency) follows Equation~\ref{eq:butterworth}.
-\begin{equation}
-\label{eq:butterworth}
-H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\frac{\omega}{\omega_c})^{2n}}}
-\end{equation}
-For a Butterworth band-pass filter, the above low-pass filter is combined with a high-pass filter of similar form.
-The response of the filter used is shown in Figure~\ref{fig:butterworth}.
-This filter improved on the above simpler filters but recordings of speech still had very low intelligibility, they could be recognised as speech but not understood.
-
-The ideal filter would be one that completely attenuates the undesired parts of the spectrum, does not change the rest of the spectrum, and has smooth edges as to not damage the signal through convolution.
-These three conditions are contradictory and any filter must compromise between them.
-The above simple filters fail at one or more of each of these conditions.
-The combs ruin the signal and the high-passes do not remove the noise.
-The Butterworth filter performed the best audibly as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
-
-% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
-Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-
-\subsection{The logMMSE estimator}
-% logMMSE
-
-Speech enhancement for a noisy channel is a classic problem in signal processing.
-Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compared various methods and found the best performing statistical technique among those at the time to be a logMMSE (alt.\ MMSE-LSA) estimator.
-This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
-Although this technique historically minimised with respect to the spectral amplitude~\cite{EphraimSpeechEU_MMSE}, throughout speech processing use of the logarithm seems produce better results as perceived by the human ear~\cite{SubjectiveComparison}.
-
-We used an existing implementation~\cite{logmmse} of the logMMSE estimator.
-Applying this estimator to the optical microphone recordings produced dramatically clean results.
-The time series (in Figure~\ref{fig:logMMSE_timeseries}) and spectrum (in Figure~\ref{fig:logMMSE_spectrum}) from before and after applying the logMMSE filter are shown.
-Comparing the before and after spectra in Figure~\ref{fig:logMMSE_spectrum} shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.
-
-After applying the estimator, simple chords and drums are heard well in played back music.
-While more complex melodies are sometimes completely absent.
-By observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
-This could be a perceptual effect of listening to the lossy recording, but seems more likely to be a frequency dependence somewhere in the optical microphone system.
-Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies and so instruments like electric bass and drums are louder in the results.
-
-After applying the estimator to speech recordings, intelligibility still remains low with the voice sounding mumbled and indistinct.
-The estimator removes most of the background noise, but does not clarify the speech.
-
-To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-This requires a better understanding of the system and further improvements to the experiment if true.
-% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
-However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
-An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
-	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) the logMMSE estimator is applied, showing one second of the one minute recording.
-The rise at the start of the bottom	 time series is an expected result from filtering a finite signal.}
-	\label{fig:logMMSE_timeseries}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
-	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) the logMMSE estimator is applied, showing only up to 2kHz and to amplitude 2 to see greater detail.}
-	\label{fig:logMMSE_spectrum}
-\end{figure}
-
-\section{Future work}
-\label{sec:future_work}
-% wild speculation
-We are currently working towards two future advancements for the optical microphone.
-
-The first would be to better isolate the experiment from the mains supply.
-The Raspberry Pi is USB powered and requires low power so it could even be run off a commercial external phone battery.
-Everything else in the circuit, including the photodiode, is connected to the Pi, so this would isolate them electrically as well.
-The laser is DC powered through a converter, so there is no issue there.
-With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
-
-The second advancement would be to achieve a full model of the transfer function of the system.
-Theoretically, if we had the correct transfer function, then signal recovery would only be limited by the signal-to-noise ratio as we could construct an optimal matched filter.
-This function would have to start from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-This sequence that the signal takes is shown in Figure~\ref{fig:pipeline_highlighted}.
-
-In our investigation, we have currently only derived part of this sequence, the relation between a deflection of the mirror and the change in the intensity of the interference pattern at some point.
-We find this relation to be non-linear, see Appendix~\ref{app:intensity_derivation} for the derivation.
-
-Asides from the transfer function, for extracting the speech there exists much literature~\cite{HMM_english} on using a hidden Markov model train to English phonemes to recognise speech.
-Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
-
-To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
-This could be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/pipeline_highlighted.pdf}
-	\caption{Sequence of relations for optical microphone transfer function. The currently derived relation is highlighted.}
-	\label{fig:pipeline_highlighted}
-\end{figure}
-
-\newpage
-\section{Conclusions}
-\label{sec:conclusions}
-
-% Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-
-We use sound and a table-top Michelson interferometer as an analogue to a gravitational wave detector.
-% This serves to demonstrate signal processing techniques.
-
-Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer.
-We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.
-
-Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the interferometer into an optical microphone.
-We demonstrated that this optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech.
-Dominant 50Hz mains hum in the recordings led to an exploration of speech enhancement techniques to remove it, including the well performing logMMSE estimate.
-From here we have many plans to further remove the mains hum and improve the intelligibility.
-
-These demonstrations have good potential to be adapted into the undergraduate laboratory.
-There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches.
-Most of all, the signal processing of speech enhancement is an interesting problem that’s relevant to digital communications.
-
-% \bigskip
-\newpage
-\appendix
 
@@
 
-\begin{figure}[h]%[H]
-	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
+\begin{figure*}
+	\includegraphics[width=.65\textwidth]{figures/circuit_diagram_2.pdf}
 	\caption{Circuit diagram for photodiode reading.
@@
 	\label{fig:circuit_diagram}
-\end{figure}
+\end{figure*}
 
-\begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{figures/circuit_pic3.pdf}
+\begin{figure*}%[H]
+	\includegraphics[width=0.5\textwidth]{figures/circuit_pic3.pdf}
 	\caption{Photo of photodiode circuit assembled on breadboard.}
 	\label{fig:circuit_pic3}
-\end{figure}
+\end{figure*}
 
@@
 \begin{align}    
-    I &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
-    I &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2)
+    I =& A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
+    I =& A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 \nonumber\\ &+ (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2)
 \end{align}
@@
 \begin{align}
-    I &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
-    I &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+    I =& A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} \nonumber\\&+ 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
+    I =& 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
 \end{align}
@@
 
-\begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
-
-The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
-
-This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
-
-Travel during the project was supported by the Australian National University PhB Science  program.
-
-\end{acknowledgments}
-
-\bibliographystyle{myunsrt}
-\bibliography{ifoDemoBib}
-
 \end{document}

### 2020-04-14 paper/ifo-complexAudio.tex
some editing throughout the paper draft

\documentclass[paper-main.tex]{subfiles}

\begin{document}

% Optical microphone
% aim of optical microphone: play back injected sounds

A natural succession to observing simple tones is to use complex audio signal, such as music and speech. 
The Viterbi algorithm and tracking more broadly is not applicable to these signals---music and speech are constantly changing over a broad spectrum. 
Instead of tracking a changing frequency over time, here we are interested in being able to record the interference pattern and play back these complex audio signal. 
We use the interferometer as an `optical microphone', using light to capture sound. 
The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.

This section is indirectly related to the advanced signal processing required to extract gravitational-wave signals. 
The optical microphone provides an interesting independent demonstration for a broader physics and engineering audience, particularly for the undergraduate lab.

We describe additional components to the demonstration in Section~\ref{sec:photodiode} and our initial results in Section~\ref{sec:initialResultsOpMic}.
Filter and speech enhancement techniques applied to the data are detailed in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse} respectively.

\subsection{The photodiode method}
\label{sec:photodiode}
% advanced method: explain how we capture data

For the optical microphone to be sensitive to speech and music it needs a sampling rate far higher than the $30\,{\rm Hz}$ of the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering}. 
Audible frequencies are in the range $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
Speech intelligibility requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$~\cite{speech_intelligibility}. 
Therefore the optical microphone data requires a sample rate of at least $16\,{\rm kHz}$ to capture both speach and music. 
We achieved this with a photodiode.

A photodiode is an electrical component that acts as a regular diode wheen no light is incident on it, blocking any current flow in the `reverse' direction.
As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction. 
% Fundamentally, a current is created in the photodiode by the photo-electric effect.
We place a (OSRAM BPW21) photodiod in reverse-bias over an (LM358) op-amp. 
This creates a photo-detector that produces a voltage dependent on the incident intensity. 
The photodiode recored the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} (however we saw no difference when monitoring at other positions in the interference pattern).
The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.

The voltage signal from the photo-detector is captured by a (MCP3008) $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
Together, these sampled the signal at roughly $16\,{\rm kHz}$, our desired limit.

Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing into the detected range. 
To prevent this, we include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC. 
This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
We also place another cloth screen over the face of the photodiode, as the reading without it, after gaining through the op-amps, was too high for the ADC.
% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).

An updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a photograph of the entire optical microphone in Figure~\ref{fig:setup_pic2}.
See Appendix~\ref{app:circuit_diagram} for a diagram and photograph of the circuit.

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_photodiode.pdf}
	\caption{Michelson interferometer schematic with the intensity of the interference pattern at a point measured by a photodiode. See Figure~\ref{fig:ifo_schematic_webcam} for details on the optical configuration.}
	\label{fig:ifo_schematic_podo}
\end{figure}

\begin{figure*}
	\includegraphics[width=.6\textwidth]{figures/setup_pic2.pdf}
\caption{Photograph of the optical microphone set-up. The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. The interferometer is laid out as in Figure~\ref{fig:ifo_schematic_podo}. In the main image, the photodiode is placed behind the cloth screen. The inset shows a face-on view of the photodiode with the covering screen removed. See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.}
	\label{fig:setup_pic2}
\end{figure*}

\subsection{Initial results of the optical microphone}
\label{sec:initialResultsOpMic}
% breakdown of all the filters used along the way

The optical microphone is tested on varied recordings of speech from different people and on music ranging from simple melodies to full track songs. 
Each recording is around one minute long with processing applied to ten second segments. 
For each observation, the audio is played through the speaker whilst the Pi script records the signal from the photodiode. 
During recordings, care is taken to ensure minimal activity around the interferometer to reduce environmental noise. 
The timeseries is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.

We find that initial recordings are extremely noisy with a loud bass hum throughout.
Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond the $8$th harmonic thereof.
We find that the mains power signal is present in data recorded without an injected audio signal as shown by the power spectral density (effectively the norm squared of the Fourier spectrum) in Figure~\ref{fig:psd_noise}.
The mains power is also present in data taken with the photodiode in complete darkness with a similar but weaker distribution as that shown in Figure~\ref{fig:psd_noise}).
The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off). We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and even cooling), but otherwise is a fairly white spectrum.}
	% except for the broad, shallow maximum at 750Hz}
	\label{fig:psd_noise}
\end{figure}

\subsection{Applying filters to the data}
\label{sec:simple_filters}

Here we describe several approaches to remove the $50\,{\rm Hz}$ mains power signal and its harmonics.

Simply zeroing the frequency channel of each $50\,{\rm Hz}$ harmonic in the Fourier spectrum is unsuccessful. 
Effectively, this multiplies the spectrum by a rectangular comb filter. 
It does remove the mains harmonics, however it audibly ruins the rest of the signal. 
Applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter. 
The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function, which significantly damages the signal.

A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible.

Applying a high-pass filter with cut-off frequency $150\,{\rm Hz}$ to the signal works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics.
However, mains harmonics $> 100\,{\rm Hz}$ and remain after filtering. 
The region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}. 
Therefore, although a high-pass filter with a higher cut-off removes more harmonics, the resultant signal is unrecognisable. 
Applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.

%We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass.

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000-cropped.pdf}
	\caption{Butterworth bandpass filter frequency response, any amplitude beyond -3dB is significant attenuation (half power). The green vertical lines show the band limits of 150Hz and 3kHz. Note the flat response within the band.}
	\label{fig:butterworth}
\end{figure}

Band-passing the signal similarly does not address the remaining harmonics above $100\,{\rm Hz}$. 
However, it does low-pass away other noise at higher frequencies. 
The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}). 
For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\epsilon$, the frequency response (attenuation at each frequency) is,
\begin{equation}
\label{eq:butterworth}
H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\frac{\omega}{\omega_c})^{2n}}},
\end{equation}
as a function of frequency $\omega$. 
To construct a Butterworth band-pass filter, the above low-pass filter is combined with a high-pass filter of similar form. 
The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
This filter improves on the above simpler filters but recordings of speech still have very low intelligibility, they can be recognised as speech but not understood.

The ideal filter would be one that i.) completely attenuates the undesired parts of the spectrum, ii.) does not change the rest of the spectrum, and iii.) has smooth edges so as to not damage the signal through convolution. 
These three conditions are contradictory and any filter must compromise between them. 
The above simple filters fail at one or more of each of these conditions. 
The combs ruin the signal and the high-passes do not remove the noise. 
The Butterworth filter performs the best audibly as it is optimised to be `maximally flat' within the band region, prioritising the second condition.

\han{Is it worth also including some plots to show how the different filters perform? E.g. a PSD with comb filter, high pass filter, Butterworth filter?}

% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal. Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.

\subsection{The logMMSE estimator}
\label{sec:logmmse}
% logMMSE

Speech enhancement for a noisy channel is a classic problem in signal processing. 
Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods. 
They find the best performing statistical technique considered to be a logMMSE (alt.\ MMSE-LSA) estimator. 
This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
Historically this technique minimises with respect to the spectral amplitude~\cite{EphraimSpeechEU_MMSE}. 
However in speech processing, use of the logarithm seems to produce better results as percieved by the human ear~\cite{SubjectiveComparison}.

In this work we used an existing implementation~\cite{logmmse} of the logMMSE estimator. 
Applying this estimator to the optical microphone recordings produced dramatically cleaner results.

In Figures~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} we present results from playing a speech recording through the optical microphone. 
Figure~\ref{fig:logMMSE_timeseries} shows the time series from before (top panel) and after (bottom panel) applying the logMMSE filter. 
Figure~\ref{fig:logMMSE_spectrum} shows the corresponding spectra before (top panel) and after (bottom panel) the logMMSE filter. 
Comparing the before and after spectra shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.

\begin{figure*}
	\includegraphics[width=.8\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
	\caption{Time series of the optical microphone recording of an adult male voice (saying ``A cathode ...'') shown before (top panel) and after (bottom panel) the logMMSE estimator is applied. One second of the one minute recording is shown. The rise at the start of the bottom time series is an expected result from filtering a finite signal.}
	\label{fig:logMMSE_timeseries}
\end{figure*}

\begin{figure*}
	\includegraphics[width=.8\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
	\caption{Spectrum of optical microphone recording of adult male voice shown before (top panel) and after (bottom panel) the logMMSE estimator is applied. The spectrum shows up to $2\,{\rm kHz}$ and to amplitude $2$ to see greater detail.}
	\label{fig:logMMSE_spectrum}
\end{figure*}

We also find improvement with music recordings.
After applying the estimator, simple chords and drums can be heard clearly. 
However more complex melodies are sometimes completely absent from the resulting time series. 
From observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
This could be a perceptual effect of listening to the lossy recording, but seems more likely to be a frequency dependence somewhere in the optical microphone system. 
Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies and so instruments like electric bass and drums are louder in the results.

Overall, we find that the intelligibility of speech signals remains low. 
The voice sounds mumbled and indistinct. 
The estimator removes most of the background noise, but does not clarify the speech. 
To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all. 
This requires a better understanding of the system and further improvements to the experiment if true.
% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio. However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference. An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.

\end{document}

### 2020-04-14 paper/ifo-conclusions.tex
some editing throughout the paper draft

\documentclass[paper-main.tex]{subfiles}

\begin{document}

% Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data. We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer. In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.

In this work we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector. 
Signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirror. 
We consider monochromatic and slowly-wandering frequency signals as an analogue to continuous wave signals as well as more complex audio recordings of speech and music. 
% This serves to demonstrate signal processing techniques.

In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of a monochromatic and wandering frequency signal. 
A webcam is used to track the interference pattern and the injected signal is recovered succesffully with the Viterbi algorithm. 
These sections direcly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
%Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer. We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.

Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the interferometer into an optical microphone.
We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 
Dominant $50\,{\rm Hz}$ mains hum in the recordings led to an exploration of speech enhancement techniques to remove it, including the well performing logMMSE estimate. 
Further avenues of exploration for this work include changes to the demonstration hardware to further remove the mains harmonics as well as investigating a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.

The demonstration presented here has good potential to be adapted for use in the undergraduate laboratory. 
It can be used to demonstrate topics of interest to astrophysics and digital communications including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement. 
%These demonstrations have good potential to be adapted into the undergraduate laboratory. There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches. Most of all, the signal processing of speech enhancement is an interesting problem that’s relevant to digital communications.

\end{document}

### 2020-04-14 paper/ifo-futureWork.tex
some editing throughout the paper draft

--- 2020-02-13 paper/paper_main.tex
@@
-% GravExplain
-\documentclass[prb,preprint]{revtex4-1} 
-% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
-
-% Note that AJP uses the same style as Phys. Rev. B (prb).
-% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
-% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
-% We look forward to receiving your submission to AJP.
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb,amsthm}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{mathtools}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{hyperref}
-% \usepackage{siunitx}
-
-\bibliographystyle{apsrev4-2}
-% \setlength{\parindent}{0pt}
-
-\newcommand{\jam}{\textcolor{magenta}}
-\newcommand{\han}{\textcolor{orange}}
+\documentclass[paper-main.tex]{subfiles}
 
 \begin{document}
+\han{add more text here from Changrong's ideas}
 
-% cover page
-\section*{Project contributions}
-% contributions clarify what’s to be assessed
-James W. Gardner:
-\begin{itemize}
-\item all paper sections other than the introduction, i.e.\ from Section~\ref{sec:ifo} onwards, and all figures in the paper
-\item addition of the injection speaker, design of photodiode circuit from standard examples, assembly of circuit on breadboard, and set-up of Raspberry Pi to measure output
-\item all of the source code at \url{https://github.com/daccordeon/gravexplain}, that reads the data from the Raspberry Pi, analyses it, and creates the figures used in the paper
-\item talks presented were based on the paper and figures therein
-\end{itemize}
+% wild speculation
+We are currently working towards two future advancements for the optical microphone.
 
-Hannah Middleton:
-\begin{itemize}
-\item introduction of paper, i.e.\ Section~\ref{sec:introduction}
-\item design and assembly of the interferometer
-\end{itemize}
+The first is to better isolate the experiment from the mains supply. 
+The Raspberry Pi is USB powered and requires low power so it could be run off a commercial external phone battery. 
+Everything else in the circuit, including the photodiode, is connected to the Pi, so this would isolate them electrically as well. 
+The laser is DC powered through a converter, so should not introduce mains harmonics itself. 
+Using the interferometer in a dark room and a sealed container could better isolate if from ambient light and air currents.
+These changes may reduce or remove the $50\,{\rm Hz}$ harmonics from the data.
+%With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
 
-Andrew Melatos, Robin Evans, William Moran: vital advice and comments throughout the project
+The second advancement would be to achieve a full model of the transfer function of the system. 
+Theoretically, if we had the correct transfer function, then signal recovery would only be limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
+The transfer function would need to start from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
+The sequence that the signal takes is shown in Figure~\ref{fig:pipeline_highlighted}.
+In our investigation, we have currently only derived part of this sequence, the relation between a deflection of the mirror and the change in the intensity of the interference pattern at some point. 
+We find this relation to be non-linear, see Appendix~\ref{app:intensity_derivation} for the derivation.
 
-\subsection*{Statement of achievements}
-% one paragraph recount of what’s been done for two months, like the abstract
-The project spanned three weeks in December, four weeks in January, and two weeks in February.
-The first two weeks were spent writing the initial scripts and achieving the Viterbi results with the webcam set-up.
-The third week was spent setting up the photodiode circuit and the Raspberry Pi.
-All of January was spent progressing through the optical microphone experiment, and was largely driven by suggestions made at the signal processing meetings with the electrical engineers.
-By the end of the third week of January we had a working optical microphone but low speech intelligibility.
-The fourth week of January was spent on attempting methods of speech enhancement.
-Finally, the two weeks in February were spent writing and drafting the paper/report, and giving presentations.
+Asides from the transfer function, for extracting the speech there exists much literature~\cite{HMM_english} on using a hidden Markov model train to English phonemes to recognise speech.
+Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
 
-\newpage
-
-\title{GravExplain: Continuous gravitational wave searches in a table-top experiment}
-
-\author{James W. Gardner}
-\email{u6069809@anu.edu.au}
-\affiliation{Research School of Physics, Australian National University, 60 Mills Rd, Acton ACT 2601 Australia}
-
-\author{Hannah Middleton$^{(i)}$}
-\email{hannah.middleton@unimelb.edu.au}
-\author{Andrew Melatos$^{(i)}$}
-\email{amelatos@unimelb.edu.au}
-\author{Robin Evans$^{(ii)}$}
-\email{robinje@unimelb.edu.au}
-\author{William Moran$^{(ii)}$}
-\email{wmoran@unimelb.edu.au}
-\affiliation{School of Physics$^{(i)}$ and Department of Electrical and Electronic Engineering$^{(ii)}$, University of Melbourne, Parkville, VIC, 3010, Australia}
-\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}
-
-% Summer 2019/2020
-\date{\today}
-
-% AJP requires an abstract for all regular article submissions.
-\begin{abstract}
-Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-We also demonstrate how the inclusion of a photodiode and Raspberry Pi allows the interferometer to act as an optical microphone.
-We are able to play-back simple audio recordings from the photodiode data.
-However, played-back speech is unintelligible and so we investigate techniques of speech enhancement.
-
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-% general gw intro
-In 2015 gravitational waves were observed for the first time from the merger of two black holes in a binary system.~\cite{GW150914} 
-The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe.
-Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years there has been an increased public interest in gravitational-wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience.
-Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
-
-% what are gws and how do we detect them...
-Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are disturbances in spacetime caused by the acceleration of massive objects. 
-The effect of gravitational waves is a change in distances; a ``stretch and squeeze'' effect. 
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
-Gravitational-wave observatories are extremely complex devices, however they are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
-
-% continuous gws
-To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-These observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencis. 
-One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
-LMXBs are binary systms containing a compact object (such as a neturon star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
-The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for continuous wave searches from LMXBs and numerous searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
-
-% sound analogy
-Here we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
-The analogy of listening to the universe is often used to explain gravitational-wave observations; transient gravitational-wave signals have been converted into audio.
-Continuous waves may emitt at constant frequencies, or they may wander slowly in frequency over time. 
-This can be demonstrated with a table-top interferometer by playing audio of constant or slowly changing frequency tones to produce a simulation of a continuous wave signal. 
-The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data, particularly for the wandering frequency demonstration~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
-
-In this work we detail our table-top intererometer design in Section~\ref{sec:ifo}, in Section~\ref{sec:single_tone} we describe observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
-We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
-
-% \section{Method and results}
-\section{Table-top gravitational wave science}
-\label{sec:ifo}
-
-Detectable gravitational waves are caused by some of the most violent events in the universe.
-Like two massive black-holes merging with each other\cite{GW150914}.
-Even so, the effects they produce are on the order of one part in $10^{21}$ for our best detectors\cite{GW150914}.
-For a table-top experiment these effects would be far too small, so we demonstrate observing a far louder source, sound waves.
-Although the analogue to gravitational waves is tenuous, we can still demonstrate some of the continuous wave analysis satisfactorily.
-The other benefit of observing sound is that we can use existing signal processing techniques designed for audio signals.
-
-\subsection{Michelson interferometry}
-% hardware: path difference, virtual thin film interference
-
-A Michelson interferometer is a well-known optics configuration used in many undergraduate and research labs, including in the detectors for gravitational waves.
-A laser beam of coherent (monochromatic and equal phase) light is split evenly by a 50/50 beamsplitter into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
-The beams recombine and interfere at the beamsplitter before hitting a screen to produce a pattern of concentric, circular fringes.
-The configuration is shown in Figure~\ref{fig:ifo_schematic_webcam}.
-In the detectors for gravitational waves, these arms are kilometres long and many other complexities (such as power recycling and quantum squeezing)\cite{GW150914} are necessary to observe a signal.
+To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
+This could be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
 
 \begin{figure}
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_webcam.pdf}
-	\caption{Michelson interferometer schematic, interference pattern is recorded by a webcam}
-	\label{fig:ifo_schematic_webcam}
-\end{figure}
-% Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
-We did not find it necessary to include it to be able to demonstrate the results desired.
-
-The interference pattern, as shown in Figure~\ref{fig:interference_pattern}, is due to the path difference between the two arms of unequal lengths (if the lengths were equal, then only a single, bright dot would be seen).
-This results in a phase difference between the two beams when they meet at the screen.
-This path difference is perhaps better seen as the perpendicular separation between one mirror and the virtual image of the other.
-For an aligned configuration and a fixed separation, the phase difference at a point only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/webcam_still0.pdf}
-	\caption{Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
-For scale, the central fringe is approximately 5mm across}
-	\label{fig:interference_pattern}
-\end{figure}
-
-% A note that while the interference pattern appears similar to the Airy pattern produced by a circular aperture, they are not in fact the same geometry.
-
-\section{Observing a single tone}
-\label{sec:single_tone}
-% the webcam method
-% how to analyse video, python OpenCV2
-
-Continuous gravitational wave searches are, by the nature of the astrophysical sources that they look for, always searching for periodic, or even sinusoidal, signals.
-Often the frequency of this periodicity can change, as addressed later, but sometimes it is (near) constant.
-As such, the audio analogue is to observe a single tone (a sinusoidal note of one frequency).
-
-\subsection{Method}
-
-We use a 0.5W speaker stuck to the back of one of the mirrors as a source for our demonstration, as shown in Figure~\ref{fig:ifo_schematic_webcam}.
-This speaker was part of a commercial 3.5mm-jack, USB powered pair of speakers and is driven by a PC.
-Previously, we used a function generator to accomplish the same signal.
-As the speaker deforms to play sound, the mirror moves with a change in separation, $\delta d$, as some coupling function of the amplitude and frequency of the sound.
-The motion of the fringes follows the motion of the mirror and therefore the speaker at the same frequency.
-If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected (played) audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
-Large fringe motion means that multiple fringes can pass over the measured point in a single speaker deformation, such over-counting would artificially raise the measured frequency.
-As such, any motion of the fringes, and so of the speaker, must be kept small by playing sound softly through the speaker.
-
-\subsection{Results}
-
-We measure the intensity at a point on the screen with a commercial USB webcam placed as to view the screen (Figure~\ref{fig:ifo_schematic_webcam}), sampling at 30Hz (or 30 fps, frames-per-second).
-This limits the spectral content of signals we can observe to be under the Nyquist frequency of 15Hz.
-See later Section~\ref{sec:photodiode} where we raise this limit into to audible frequencies (20 Hz to 20kHz) with a photodiode.
-While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
-Then, at an off-centre pixel as seen in Figure~\ref{fig:interference_pattern}, we take the green-channel of the colour (RGB) video as approximate to the total intensity, as the laser is monochromatic green light.
-We can play this time series of intensity over time as an audio recording (we used the \textbf{scipy.io.wavfile.write}\cite{scipy} function in Python\cite{python}) and the tone is clearly heard.
-Taking the Fourier transform of this signal shows the injected tone, as seen in Figure~\ref{fig:webcam_spectrum} where we record the highest peak Fourier amplitude at $(2.099\pm 0.008)\mathrm{Hz}$ for an injected tone of 2.09Hz (a result within two standard deviations).
-Therefore, we’re able to observe a single tone quite confidently.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the highest peak is measured at $(2.099\pm 0.008) \mathrm{Hz}$ for an injected tone of 2.09Hz.
-In fact, the actual highest peak is at 0Hz, but that corresponds only to a non-zero average value, so we don’t display or consider it.}
-	\label{fig:webcam_spectrum}
-\end{figure}
-
-\section{Observing a wandering tone}
-\label{sec:viterbi_wandering}
-% Tracking tones with the Viterbi algorithm
-% explaination and successful application
-% lean on the continuous gravitational waves angle
-
-Continuous gravitational wave searches often look for signals that wander (change) in frequency slowly over time\cite{2019PhRvD.100l2002A}.
-This means that finding the signal isn’t as simple as searching for a peak in the frequency content for the full observing run, especially when the signal-to-noise ratio is particularly low.
-
-One method (among many) is to split the observed time series into bins of time, compute the Fourier transform for each time interval, and then create a grid (known as a spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time.
-The method then treats the problem as one of finding the best path through a weighted graph, where here the weights are given by the Fourier amplitudes.
-The method takes in a parameter of the allowed frequency wandering, some rate of frequency change over time, usually informed by a physical model.
-Here, this parameter is set to 0.11 Hz per second to make it simply one frequency bin per every time bin, which will prove poorly adjusted to the signal.
-
-\subsection{The Viterbi algorithm}
-
-The Viterbi algorithm is a path-finding algorithm that solves this problem.
-Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns holding spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
-Then, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph.
-At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path.
-See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.
-
-\subsection{Results}
-
-We demonstrate this method of using the Viterbi algorithm to find a signal, by injecting a wandering frequency signal into the speaker and recording the output via the webcam.
-Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
-
-Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, overlaid with the injected signal and the recovered Viterbi path.
-To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment.
-Quantifying this, the Viterbi path stays within one frequency bin (approximately 0.3Hz) of the injected tone for 94\% of the run.
-Note that the signal initially moves faster than the chosen frequency wander parameter, causing a discrepancy between the injected and recovered signals.
-To fix this, the parameter should be doubled, although this would then cause the recovered path to stray more often.
-There also appears to be an anomaly at 150s, likely from a disturbance (such as someone walking) around the interferometer.
-
-Empirically, decreasing the signal-to-noise ratio closer to unity also causes the Viterbi algorithm to more easily stray from the signal to reach nearby fluctuations.
-However, the only correlation through the grid is the signal, so we always expect the Viterbi path to lie near to the signal no matter how high the noise is.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
-	\caption{Spectrogram of observed signal overlaid with white crosses for the Viterbi path and pink dots for the injected signal, notice the high correlation between them indicating that the signal is able to be recovered; also note the deviation at the start when the signal moves too quickly and just before the background anomaly at 150s}
-	\label{fig:viterbi_overlay}
-\end{figure}
-
-\section{Playing back complex audio}
-\label{sec:optical_microphone}
-% Optical microphone
-% aim of optical microphone: play back injected sounds
-
-The natural succession to observing simple tones is to study complex audio signals, such as music and speech.
-The Viterbi algorithm and tracking more broadly is not applicable to these signals.
-Compare a single tone that slowly, continuously changes frequency to music and speech which consist of a broad spectrum that is constantly changing.
-Instead, the interest here is in being able to record and play-back these complex audio signals, rather than track them.
-Using the interferometer to do so makes it, in ensemble with the measuring device, an ‘optical microphone’, using light to capture sound.
-
-Advanced signal processing is required to extract gravitational wave signals from very noisy observational data.
-Recovering audio from an optical microphone requires related signal processing.
-Although far simpler, it serves as another useful demonstration.
-
-\subsection{The photodiode method}
-\label{sec:photodiode}
-% advanced method: explain how we capture data
-
-For an optical microphone to be sensitive to speech and music it needs a frequency response (a sampling rate) far higher than the 30Hz of a standard webcam.
-Both speech and music need at least kHz range, with speech intelligibility requiring up to 3kHz and music preferably up to and beyond 8kHz\cite{speech_intelligibility}.
-This can be addressed by using a high-speed webcam, or more simply, by using a photodiode as the measuring device.
-
-A photodiode is an electrical component that acts as a regular diode (blocking any current flow in the reverse direction) when no light is incident on it, but becomes more and more conductive in the reverse direction as the intensity of incident light increases.
-Fundamentally, a current is created in the photodiode by the photo-electric effect.
-Placing a photodiode in reverse-bias over an op-amp creates a photo-detector that will produce a voltage dependent approximately linearly to the incident intensity.
-\jam{How sound is this approximation?} This experiment had the photodiode in the interference pattern at roughly the same spot as the webcam was monitoring previously, just off-centre.
-
-The voltage signal from the photo-detector is captured by a 10-bit analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi, in the standard configuration.
-This digitally samples the signal at roughly 16kHz.
-As such, any frequency component of the analog signal above the Nyquist frequency of 8kHz would be aliased down into the detected range.
-To prevent this, an anti-aliasing Sallen-Key filter\cite{sallen_key_filter} tuned to 16kHz is included before the ADC.
-This component attenuates any frequencies above 8kHz, before they get digitally sampled.
-It’s not immediately clear what’s limiting the sample rate to 16kHz, but it is likely non-optimal reading of the ADC by the Pi script.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-
-The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}.
-The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the experiment, showing the interferometer, the circuit, and the Pi, is shown in Figure~\ref{fig:setup_pic2}.
-
-\begin{figure}
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_photodiode.pdf}
-	\caption{Michelson interferometer schematic, a part of the interference pattern is recorded by a photodiode}
-	\label{fig:ifo_schematic_podo}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/setup_pic2.pdf}
-	\caption{Picture taken of the optical microphone set-up, showing the Michelson interferometer, the photodiode (in an insert), the circuit, and the Raspberry Pi.
-For a closer photo of the circuit, see Figure~\ref{fig:circuit_pic3}.}
-	\label{fig:setup_pic2}
-\end{figure}
-
-\subsection{Initial results}
-% breakdown of all the filters used along the way
-
-The optical microphone was tested on recordings of one human talking and on music ranging from simple melodies to full tracks.
-Each recording was around one minute long with processing done on ten second segments.
-For each observation, the audio was played through the speaker with minimal activity around interferometer as the Pi script recorded the signal.
-The saved time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{scipy.io.wavfile.write}\cite{scipy} implementation in Python\cite{python}.
-
-These initial recordings were extremely noisy and painful to listen to, with a loud bass sound throughout.
-Spectra of all of these recordings showed a dominant mains hum at 50Hz with strong power up to and beyond the 8th harmonic thereof.
-This power is also present with the speaker off, as is shown in Figure~\ref{fig:psd_noise}, and also weakly when monitoring with the photodiode in complete darkness.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
-	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), show high power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the ambient lighting), otherwise is a fairly white spectrum except for the broad, shallow maximum at 750Hz}
-	\label{fig:psd_noise}
-\end{figure}
-
-\subsection{Applying filters to the data}
-
-Early attempts to remove the mains hum included zeroing the channel of each 50Hz harmonic, this is effectively using a rectangular comb filter.
-Applying a filter (multiplying by a function) in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
-While this did remove the mains harmonics, it also destroyed the rest of the signal.
-
-Attempting to low-pass the signal did better (to the ear) but was also problematic as the harmonics are all over the 100Hz-1kHz region, which also carries a lot of the speech and music information.
-Low-passing the logarithm of the signal and then converting back did better still but doesn’t significantly remove the mains noise.
-
-Of these simple filters, the best performing was a band-pass Butterworth filter with frequency response shown in Figure~\ref{fig:butterworth}.
-This filter is `maximally flat' within the band region, here 150Hz to 3kHz, and trails off outside of the band with slope dependent on the order of the filter, here of order 5.
-This still suffered from the presence of the higher mains harmonics.
-However, behind the remaining noise one could hear something distinctly speech-like or musical, just with very low intelligibility, you couldn’t understand the speech but could recognise that someone was talking.
-
-% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
-Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-
-\begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{figures/butterworth_150_3000-cropped.pdf}
-	\caption{Butterworth bandpass filter frequency response, green vertical lines show the band limits of 150Hz and 3kHz.
-Note the very flat response within the band, any amplitude beyond -3dB is significant attenuation.}
-	\label{fig:butterworth}
-\end{figure}
-
-\subsection{The logMMSE estimator}
-% logMMSE
-
-\jam{Explain MMSE as in the second talk, be honest in some way}
-
-Speech enhancement for a noisy channel is a classic problem in signal processing.
-Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare various methods and find the best performing statistical technique to be the logMMSE (alt.\ MMSE-LSA) estimator which minimises the mean square error (MSE) of the log-spectral amplitude.
-This means that it takes the logarithm of the spectrum of the original signal and finds the fit (or estimate) that has the minimal MSE to that log-spectra.
-% Use of the logarithm is a common technique in speech processing as it, somewhat surprisingly, does not significantly change the intelligibility to the human ear.
-
-Applying an existing implementation of logMMSE\cite{logmmse} to the optical microphone recordings produces dramatically cleaner results.
-The time series and spectrum from before and after the logMMSE filter is applied are shown in Figures~\ref{fig:logMMSE_timeseries}~\&~\ref{fig:logMMSE_spectrum}.
-All of the mains harmonics are significantly attenuated.
-And further high-passing removes the low frequency curve beneath 100Hz.
-
-There is varied success with passing music passed through the filters.
-Loud, rhythmic chords and drums are received well and remain recognisable, while thinner, airy flute and violin melodies struggle to be heard in the results.
-It is likely that the lower sounds have higher amplitude speaker oscillations and so pass through the optical microphone better.
-
-However, speech intelligibility remains very low with the voice sounding mumbled and indistinct.
-Enhancement so far has removed most of the background sound, as evidenced by the near silence between words, but has not clarified the speech.
-There is a question here as to where the diction in the speech is lost, at the speaker-mirror coupling, at the photodiode reading, or if it is still there, just obscured in the noisy signal.
-
-% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
-However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
-An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
-% Visually it is impossible to confirm this for high frequencies, but it remains possible.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
-	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) logMMSE filter is applied, only one second of the one minute recording is shown.
-The bump at the start of the time series is the result of filtering a finite signal, and is expected.}
-	\label{fig:logMMSE_timeseries}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
-	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) logMMSE filter is applied, showing only up to 2kHz and a detail of the spectrum (limited to show only up to amplitude 2, while before filtering the 50Hz line has an amplitude of 40)}
-	\label{fig:logMMSE_spectrum}
-\end{figure}
-
-\section{Future work}
-\label{sec:future_work}
-
-\jam{BATTERY POWER}
-% any other wild speculation?
-
-\subsection{Finding the transfer function}
-
-The intensity, $I$, of the interference pattern at a particular angle is given by Equation~\ref{eq:intensity_result} below.
-Where A is the amplitude of the electric field of the beam, d is the separation (or the difference in the arm lengths), $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Appendix~\ref{app:intensity_derivation} for the derivation to get to this equation).
-\begin{equation}
-\label{eq:intensity_result}
-I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{equation}
-
-This Equation~\ref{eq:intensity_result} means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
-This makes the problem of extracting the original signal more difficult.
-
-\begin{equation}
-\label{eq:intensity_change}
-	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{equation}
-
-A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-Although we have the separation-intensity relation as non-linear for small motions (see above), the other components of this full model are yet to be determined.
-Figure~\ref{fig:pipeline_highlighted} shows this full pipeline, with the currently determined section highlighted.
-
-Theoretically, if we had the correct model, then signal recovery would only be limited by the SNR. We could also use long recordings of white noise to estimate a matched filter for the system.
-
-For further speech enhancement, the optimal filter would involve a hidden-Markov-model trained to English phonemes. Alternatively, machine learning solutions also exist through-out the field.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/pipeline_highlighted.pdf}
-	\caption{???}
+	\includegraphics[width=.49\textwidth]{figures/pipeline_highlighted.pdf}
+	\caption{Sequence of relations for optical microphone transfer function. The currently derived relation is highlighted.}
 	\label{fig:pipeline_highlighted}
@@
 
-\section{Conclusions}
-\label{sec:conclusions}
-
-Using a Michelson interferometer, we able to successfully demonstrate the Viterbi algorithm recovering an injected signal.
-Just as is done is continuous gravitational wave searches.
-We are also able to play-back chordal or rhythmic music from an optical microphone (using the same interferometer), while recordings of simple speech are audible but unintelligible.
-
-\newpage
-\appendix
-
-\section{Open-source code}
-\label{app:code}
-This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.
-
-The current build can be found at:
-\url{https://github.com/daccordeon/gravexplain}
-% \begin{verbatim}
-% README
-% \end{verbatim}
-
-\section{Additional guide to reproducing results}
-\label{app:reproducing_results}
-% as for a 3rd year undergraduate
-
-Besides the directions given in the rest of this article, there are some observations and technical details useful to someone wanting to reproduce the results shown herein.
-For the analysis and signal processing performed the best reference is Appendix~\ref{app:code} for the open-source code.
-
-For the experiment, since a Michelson interferometer is such a standard configuration, we’ll not detail it further here, see any optics manual for more information.
-To obtain an independent speaker cone we recommend carefully dismantling a budget commercial speaker system to get at the speaker within.
-To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place.
-If the other speaker in the pair remains, keep it as far away from the experiment and as damped as possible else it may act as a second, interfering source.
-% we used a Class 2 LDS5 ThorLabs laser at 532nm
-
-For the photodiode circuit, we used a Raspberry Pi and an MCP3008 ADC to monitor the analog signal.
-This could also be done with the analog pins on an Arduino.
-Be careful of the gain through the LM358 op-amps in the circuit though, else the ADC in either case could threshold the signal.
-Also be ready to change the values in the Sallen-Key filter to match your own sampling rate.
-The photodiode was an OSRAM BPW21, we mounted it through a discarded piece of speaker mesh (similar to black cloth over a frame) and placed another piece over the face to not over expose and threshold the signal.
-For the rest of the circuit, we used standard components assembled on a breadboard as detailed in Appendix~\ref{app:circuit_diagram}.
-
-\section{The Viterbi algorithm}
-\label{app:viterbi}
-
-To find the Viterbi path through a spectrogram (spectra over time), given an M x N grid of M frequency rows and N time columns and a rate limit for the change of frequency over time that we convert into the number of allowed rows we can move while stepping from each column to the next (this limit was $\pm 1$ row/column in our case).
-The path must start from the first column and end at the last, taking a single step in each column along the way.
-Each element in the grid is the norm of the Fourier component of a frequency at a time.
-These are normalised between $(0, 1)$ to use as multiplicative weights, i.e.\ the total weight of a path is the product of all weights along that path.
-In fact, we take the logarithm and so the sum of all weights to avoid underflow.
-The best path has the highest weight.
-
-With the grid prepared, the Viterbi algorithm starts in the second column.
-In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, taking the greatest cumulative weight of those available, ‘nearby’ cells (the value of which was discovered in the previous iteration).
-For each cell in the current column, the algorithm notes this weight plus the cell’s own weight, along with the row index of which cell in the previous column it came from.
-Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
-That best path is the Viterbi path.
-
-\section{A common mistake with wandering frequency signals}
-\label{app:phase_gotcha}
-
-We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
-The common mistake is to just appeal to the standard form $\sin{(2 \pi f t)}$, where f would be a constant frequency, and write down $\sin{(2 \pi f(t) t)}$.
-However, accumulation of phase means that this produces a curve that looks very different to that desired.
-Instead, one should see that the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
-Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
-
-\section{Circuit design}
-\label{app:circuit_diagram}
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-
-Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
-	\caption{Circuit diagram for photodiode reading.
-Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC).
-The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.}
-	\label{fig:circuit_diagram}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{figures/circuit_pic3.pdf}
-	\caption{Photo of circuit for photodiode reading assembled on breadboard.}
-	\label{fig:circuit_pic3}
-\end{figure}
-
-\section{Intensity relation derivation}
-\label{app:intensity_derivation}
-
-We want the change in intensity, $\delta I$, at a point in the Michelson interference pattern, given a small change, $\delta d$, in the perpendicular path difference of the mirrors (or in the separation of one and virtual image of the other).
-The point is at angle $\theta$ on the screen above the central ray, noting circular symmetry.
-At this point, two rays meet, one from each mirror, with the total electric field (\jam{parallel to the screen?}) being $E_1(t) + E_2(t)$.
-Given an initial separation of $d$, the path difference between the mirrors is $2 d \cos{\theta}$ as the longer ray must travel to and from the further mirror, and both move up at the angle $\theta$ to meet at the screen.
-
-\begin{align}
-\label{eq:intensity_derivation}
-    E_1(t) &= A e^{i (k L - \omega t)} \\
-    E_2(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \\
-    I(t) &= \lvert A e^{i (k L - \omega t)} + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2
-\end{align}
-
-Collapse the phase difference to $\phi$ and expand the norm squared by expanding the complex exponentials.
-From there is it trigonometric manipulation to reach the result of intensity in terms of separation.
-
-\begin{align}    
-    I(t) &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
-    I(t) &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2) \\
-    I(t) &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
-    I(d) &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{align}
-
-Then take derivatives and find that the intensity change is non-linear.
-
-\begin{align}    
-    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
-    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
-    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
-    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{align}
-
-\begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
-
-The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
-
-This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
-
-Travel during the project was supported by the Australian National University PhB Science  program.
-
-\end{acknowledgments}
-
-\bibliographystyle{myunsrt}
-\bibliography{ifoDemoBib}
-
 \end{document}

### 2020-04-14 paper/ifo-introduction.tex
some editing throughout the paper draft

--- 2020-02-13 paper/paper_main.tex
@@
-% GravExplain
-\documentclass[prb,preprint]{revtex4-1} 
-% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
-
-% Note that AJP uses the same style as Phys. Rev. B (prb).
-% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
-% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
-% We look forward to receiving your submission to AJP.
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb,amsthm}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{mathtools}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{hyperref}
-% \usepackage{siunitx}
-
-\bibliographystyle{apsrev4-2}
-% \setlength{\parindent}{0pt}
-
-\newcommand{\jam}{\textcolor{magenta}}
-\newcommand{\han}{\textcolor{orange}}
+\documentclass[paper-main.tex]{subfiles}
 
@@
 
-% cover page
-\section*{Project contributions}
-% contributions clarify what’s to be assessed
-James W. Gardner:
-\begin{itemize}
-\item all paper sections other than the introduction, i.e.\ from Section~\ref{sec:ifo} onwards, and all figures in the paper
-\item addition of the injection speaker, design of photodiode circuit from standard examples, assembly of circuit on breadboard, and set-up of Raspberry Pi to measure output
-\item all of the source code at \url{https://github.com/daccordeon/gravexplain}, that reads the data from the Raspberry Pi, analyses it, and creates the figures used in the paper
-\item talks presented were based on the paper and figures therein
-\end{itemize}
-
-Hannah Middleton:
-\begin{itemize}
-\item introduction of paper, i.e.\ Section~\ref{sec:introduction}
-\item design and assembly of the interferometer
-\end{itemize}
-
-Andrew Melatos, Robin Evans, William Moran: vital advice and comments throughout the project
-
-\subsection*{Statement of achievements}
-% one paragraph recount of what’s been done for two months, like the abstract
-The project spanned three weeks in December, four weeks in January, and two weeks in February.
-The first two weeks were spent writing the initial scripts and achieving the Viterbi results with the webcam set-up.
-The third week was spent setting up the photodiode circuit and the Raspberry Pi.
-All of January was spent progressing through the optical microphone experiment, and was largely driven by suggestions made at the signal processing meetings with the electrical engineers.
-By the end of the third week of January we had a working optical microphone but low speech intelligibility.
-The fourth week of January was spent on attempting methods of speech enhancement.
-Finally, the two weeks in February were spent writing and drafting the paper/report, and giving presentations.
-
-\newpage
-
-\title{GravExplain: Continuous gravitational wave searches in a table-top experiment}
-
-\author{James W. Gardner}
-\email{u6069809@anu.edu.au}
-\affiliation{Research School of Physics, Australian National University, 60 Mills Rd, Acton ACT 2601 Australia}
-
-\author{Hannah Middleton$^{(i)}$}
-\email{hannah.middleton@unimelb.edu.au}
-\author{Andrew Melatos$^{(i)}$}
-\email{amelatos@unimelb.edu.au}
-\author{Robin Evans$^{(ii)}$}
-\email{robinje@unimelb.edu.au}
-\author{William Moran$^{(ii)}$}
-\email{wmoran@unimelb.edu.au}
-\affiliation{School of Physics$^{(i)}$ and Department of Electrical and Electronic Engineering$^{(ii)}$, University of Melbourne, Parkville, VIC, 3010, Australia}
-\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}
-
-% Summer 2019/2020
-\date{\today}
-
-% AJP requires an abstract for all regular article submissions.
-\begin{abstract}
-Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-We also demonstrate how the inclusion of a photodiode and Raspberry Pi allows the interferometer to act as an optical microphone.
-We are able to play-back simple audio recordings from the photodiode data.
-However, played-back speech is unintelligible and so we investigate techniques of speech enhancement.
-
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
 % general gw intro
-In 2015 gravitational waves were observed for the first time from the merger of two black holes in a binary system.~\cite{GW150914} 
-The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and provides a new means to observe the universe.
+In 2015, gravitational waves were observed for the first time from the merger of two black holes in a binary system~\cite{GW150914}. 
+The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
 Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years there has been an increased public interest in gravitational-wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to aid in explaining this topic to a general audience.
-Activities range from online data analysis tutorials,~\cite{GWOSC:online,LOSC:2015} hands-on demonstrations, phone apps,~\cite{LaserLabs:online,SciVR:online} and online games,~\cite{BlackHoleHunter:online} to exhibitions,\cite{L2URSSE} artistic interpretation, and musical performance~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
+In recent years, there has been increased public interest in gravitational wave science. 
+Many gravitational wave research groups around the world have produced demonstrations and activities to explain the topic to a general audience.
+Activities range from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}%\han{look up Sarah Farmer's project}.
 
@@
 Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are disturbances in spacetime caused by the acceleration of massive objects. 
-The effect of gravitational waves is a change in distances; a ``stretch and squeeze'' effect. 
+They are disturbances in space-time caused by the acceleration of massive objects. 
+The effect of gravitational waves is a change in lengths; a `stretching and squeezing' of the distance between objects.
 Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
-Gravitational-wave observatories are extremely complex devices, however they are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used by gravitational-wave research groups to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets} and they are often used in undergraduate lab experiments~\cite{UgoliniEtAl:2019}.
+%As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
+These observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
+Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
 
 % continuous gws
-To date the network of gravitational-wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-These observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencis. 
+To date the network of gravitational wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
+However, the network is also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencies.
+Continuous waves may be emitted at constant frequencies, or may wander slowly in frequency over time. 
 One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
-LMXBs are binary systms containing a compact object (such as a neturon star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
+LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
 The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for continuous wave searches from LMXBs and numerous searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
+It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
 
 % sound analogy
-Here we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
-The analogy of listening to the universe is often used to explain gravitational-wave observations; transient gravitational-wave signals have been converted into audio.
-Continuous waves may emitt at constant frequencies, or they may wander slowly in frequency over time. 
-This can be demonstrated with a table-top interferometer by playing audio of constant or slowly changing frequency tones to produce a simulation of a continuous wave signal. 
-The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data, particularly for the wandering frequency demonstration~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
+Here, we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
+The analogy of sound is commonly used in explaining gravitational wave observations. 
+%; the analogy of listening to the universe is often used to explain gravitational wave observations.
+%; transient gravitational wave signals have been converted into audio.
+In this work we utilise this analogy by playing sound close to a Michelson interferometer, simulating a gravitational-wave signal. 
+Audio of constant or slowly wandering frequency tones provide an analogue of a continuous wave signal.
+Continuous wave data analysis techniques are also demonstrated. 
+The techniques used in this work are similar to those used to search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}, particularly for the wandering frequency demonstration.
 
-In this work we detail our table-top intererometer design in Section~\ref{sec:ifo}, in Section~\ref{sec:single_tone} we describe observing a single tone signal, Section~\ref{sec:viterbi_wandering} describes observing a wandering frequency signal, and Section~\ref{sec:optical_microphone} extends the set-up to be able to capture complex audio, such as music and speech.
+This demonstration can be used to explain the search for continuous waves to a general audience as well as undergraduate students. 
+The analysis methods applied here follow signal processing techniques used in some of the most recent searches for continuous waves~\cite{ScoX1O2Viterbi:2019}. 
+As an undergraduate lab experiment, this demonstration covers topics of interest to both physics and engineering students.
+
+In this work we detail our table-top interferometer design in Section~\ref{sec:ifo} and describe observing a single note from a speaker in Section~\ref{sec:single_tone}.
+In Section~\ref{sec:viterbi_wandering}, we demonstrate the analysis techniques for observing a wandering frequency signal.
+In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech.
+This demonstration of an `optical microphone' serves as a more general exhibition of signal processing.
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
 
-% \section{Method and results}
-\section{Table-top gravitational wave science}
-\label{sec:ifo}
-
-Detectable gravitational waves are caused by some of the most violent events in the universe.
-Like two massive black-holes merging with each other\cite{GW150914}.
-Even so, the effects they produce are on the order of one part in $10^{21}$ for our best detectors\cite{GW150914}.
-For a table-top experiment these effects would be far too small, so we demonstrate observing a far louder source, sound waves.
-Although the analogue to gravitational waves is tenuous, we can still demonstrate some of the continuous wave analysis satisfactorily.
-The other benefit of observing sound is that we can use existing signal processing techniques designed for audio signals.
-
-\subsection{Michelson interferometry}
-% hardware: path difference, virtual thin film interference
-
-A Michelson interferometer is a well-known optics configuration used in many undergraduate and research labs, including in the detectors for gravitational waves.
-A laser beam of coherent (monochromatic and equal phase) light is split evenly by a 50/50 beamsplitter into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
-The beams recombine and interfere at the beamsplitter before hitting a screen to produce a pattern of concentric, circular fringes.
-The configuration is shown in Figure~\ref{fig:ifo_schematic_webcam}.
-In the detectors for gravitational waves, these arms are kilometres long and many other complexities (such as power recycling and quantum squeezing)\cite{GW150914} are necessary to observe a signal.
-
-\begin{figure}
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_webcam.pdf}
-	\caption{Michelson interferometer schematic, interference pattern is recorded by a webcam}
-	\label{fig:ifo_schematic_webcam}
-\end{figure}
-% Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
-We did not find it necessary to include it to be able to demonstrate the results desired.
-
-The interference pattern, as shown in Figure~\ref{fig:interference_pattern}, is due to the path difference between the two arms of unequal lengths (if the lengths were equal, then only a single, bright dot would be seen).
-This results in a phase difference between the two beams when they meet at the screen.
-This path difference is perhaps better seen as the perpendicular separation between one mirror and the virtual image of the other.
-For an aligned configuration and a fixed separation, the phase difference at a point only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/webcam_still0.pdf}
-	\caption{Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
-For scale, the central fringe is approximately 5mm across}
-	\label{fig:interference_pattern}
-\end{figure}
-
-% A note that while the interference pattern appears similar to the Airy pattern produced by a circular aperture, they are not in fact the same geometry.
-
-\section{Observing a single tone}
-\label{sec:single_tone}
-% the webcam method
-% how to analyse video, python OpenCV2
-
-Continuous gravitational wave searches are, by the nature of the astrophysical sources that they look for, always searching for periodic, or even sinusoidal, signals.
-Often the frequency of this periodicity can change, as addressed later, but sometimes it is (near) constant.
-As such, the audio analogue is to observe a single tone (a sinusoidal note of one frequency).
-
-\subsection{Method}
-
-We use a 0.5W speaker stuck to the back of one of the mirrors as a source for our demonstration, as shown in Figure~\ref{fig:ifo_schematic_webcam}.
-This speaker was part of a commercial 3.5mm-jack, USB powered pair of speakers and is driven by a PC.
-Previously, we used a function generator to accomplish the same signal.
-As the speaker deforms to play sound, the mirror moves with a change in separation, $\delta d$, as some coupling function of the amplitude and frequency of the sound.
-The motion of the fringes follows the motion of the mirror and therefore the speaker at the same frequency.
-If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected (played) audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
-Large fringe motion means that multiple fringes can pass over the measured point in a single speaker deformation, such over-counting would artificially raise the measured frequency.
-As such, any motion of the fringes, and so of the speaker, must be kept small by playing sound softly through the speaker.
-
-\subsection{Results}
-
-We measure the intensity at a point on the screen with a commercial USB webcam placed as to view the screen (Figure~\ref{fig:ifo_schematic_webcam}), sampling at 30Hz (or 30 fps, frames-per-second).
-This limits the spectral content of signals we can observe to be under the Nyquist frequency of 15Hz.
-See later Section~\ref{sec:photodiode} where we raise this limit into to audible frequencies (20 Hz to 20kHz) with a photodiode.
-While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
-Then, at an off-centre pixel as seen in Figure~\ref{fig:interference_pattern}, we take the green-channel of the colour (RGB) video as approximate to the total intensity, as the laser is monochromatic green light.
-We can play this time series of intensity over time as an audio recording (we used the \textbf{scipy.io.wavfile.write}\cite{scipy} function in Python\cite{python}) and the tone is clearly heard.
-Taking the Fourier transform of this signal shows the injected tone, as seen in Figure~\ref{fig:webcam_spectrum} where we record the highest peak Fourier amplitude at $(2.099\pm 0.008)\mathrm{Hz}$ for an injected tone of 2.09Hz (a result within two standard deviations).
-Therefore, we’re able to observe a single tone quite confidently.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the highest peak is measured at $(2.099\pm 0.008) \mathrm{Hz}$ for an injected tone of 2.09Hz.
-In fact, the actual highest peak is at 0Hz, but that corresponds only to a non-zero average value, so we don’t display or consider it.}
-	\label{fig:webcam_spectrum}
-\end{figure}
-
-\section{Observing a wandering tone}
-\label{sec:viterbi_wandering}
-% Tracking tones with the Viterbi algorithm
-% explaination and successful application
-% lean on the continuous gravitational waves angle
-
-Continuous gravitational wave searches often look for signals that wander (change) in frequency slowly over time\cite{2019PhRvD.100l2002A}.
-This means that finding the signal isn’t as simple as searching for a peak in the frequency content for the full observing run, especially when the signal-to-noise ratio is particularly low.
-
-One method (among many) is to split the observed time series into bins of time, compute the Fourier transform for each time interval, and then create a grid (known as a spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time.
-The method then treats the problem as one of finding the best path through a weighted graph, where here the weights are given by the Fourier amplitudes.
-The method takes in a parameter of the allowed frequency wandering, some rate of frequency change over time, usually informed by a physical model.
-Here, this parameter is set to 0.11 Hz per second to make it simply one frequency bin per every time bin, which will prove poorly adjusted to the signal.
-
-\subsection{The Viterbi algorithm}
-
-The Viterbi algorithm is a path-finding algorithm that solves this problem.
-Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns holding spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
-Then, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph.
-At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path.
-See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.
-
-\subsection{Results}
-
-We demonstrate this method of using the Viterbi algorithm to find a signal, by injecting a wandering frequency signal into the speaker and recording the output via the webcam.
-Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
-
-Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, overlaid with the injected signal and the recovered Viterbi path.
-To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment.
-Quantifying this, the Viterbi path stays within one frequency bin (approximately 0.3Hz) of the injected tone for 94\% of the run.
-Note that the signal initially moves faster than the chosen frequency wander parameter, causing a discrepancy between the injected and recovered signals.
-To fix this, the parameter should be doubled, although this would then cause the recovered path to stray more often.
-There also appears to be an anomaly at 150s, likely from a disturbance (such as someone walking) around the interferometer.
-
-Empirically, decreasing the signal-to-noise ratio closer to unity also causes the Viterbi algorithm to more easily stray from the signal to reach nearby fluctuations.
-However, the only correlation through the grid is the signal, so we always expect the Viterbi path to lie near to the signal no matter how high the noise is.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
-	\caption{Spectrogram of observed signal overlaid with white crosses for the Viterbi path and pink dots for the injected signal, notice the high correlation between them indicating that the signal is able to be recovered; also note the deviation at the start when the signal moves too quickly and just before the background anomaly at 150s}
-	\label{fig:viterbi_overlay}
-\end{figure}
-
-\section{Playing back complex audio}
-\label{sec:optical_microphone}
-% Optical microphone
-% aim of optical microphone: play back injected sounds
-
-The natural succession to observing simple tones is to study complex audio signals, such as music and speech.
-The Viterbi algorithm and tracking more broadly is not applicable to these signals.
-Compare a single tone that slowly, continuously changes frequency to music and speech which consist of a broad spectrum that is constantly changing.
-Instead, the interest here is in being able to record and play-back these complex audio signals, rather than track them.
-Using the interferometer to do so makes it, in ensemble with the measuring device, an ‘optical microphone’, using light to capture sound.
-
-Advanced signal processing is required to extract gravitational wave signals from very noisy observational data.
-Recovering audio from an optical microphone requires related signal processing.
-Although far simpler, it serves as another useful demonstration.
-
-\subsection{The photodiode method}
-\label{sec:photodiode}
-% advanced method: explain how we capture data
-
-For an optical microphone to be sensitive to speech and music it needs a frequency response (a sampling rate) far higher than the 30Hz of a standard webcam.
-Both speech and music need at least kHz range, with speech intelligibility requiring up to 3kHz and music preferably up to and beyond 8kHz\cite{speech_intelligibility}.
-This can be addressed by using a high-speed webcam, or more simply, by using a photodiode as the measuring device.
-
-A photodiode is an electrical component that acts as a regular diode (blocking any current flow in the reverse direction) when no light is incident on it, but becomes more and more conductive in the reverse direction as the intensity of incident light increases.
-Fundamentally, a current is created in the photodiode by the photo-electric effect.
-Placing a photodiode in reverse-bias over an op-amp creates a photo-detector that will produce a voltage dependent approximately linearly to the incident intensity.
-\jam{How sound is this approximation?} This experiment had the photodiode in the interference pattern at roughly the same spot as the webcam was monitoring previously, just off-centre.
-
-The voltage signal from the photo-detector is captured by a 10-bit analog-to-digital converter (ADC) connected to the special purpose input (SPI) pins of a Raspberry Pi, in the standard configuration.
-This digitally samples the signal at roughly 16kHz.
-As such, any frequency component of the analog signal above the Nyquist frequency of 8kHz would be aliased down into the detected range.
-To prevent this, an anti-aliasing Sallen-Key filter\cite{sallen_key_filter} tuned to 16kHz is included before the ADC.
-This component attenuates any frequencies above 8kHz, before they get digitally sampled.
-It’s not immediately clear what’s limiting the sample rate to 16kHz, but it is likely non-optimal reading of the ADC by the Pi script.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-
-The entire circuit diagram is detailed and shown in Appendix~\ref{app:circuit_diagram}.
-The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the experiment, showing the interferometer, the circuit, and the Pi, is shown in Figure~\ref{fig:setup_pic2}.
-
-\begin{figure}
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_photodiode.pdf}
-	\caption{Michelson interferometer schematic, a part of the interference pattern is recorded by a photodiode}
-	\label{fig:ifo_schematic_podo}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/setup_pic2.pdf}
-	\caption{Picture taken of the optical microphone set-up, showing the Michelson interferometer, the photodiode (in an insert), the circuit, and the Raspberry Pi.
-For a closer photo of the circuit, see Figure~\ref{fig:circuit_pic3}.}
-	\label{fig:setup_pic2}
-\end{figure}
-
-\subsection{Initial results}
-% breakdown of all the filters used along the way
-
-The optical microphone was tested on recordings of one human talking and on music ranging from simple melodies to full tracks.
-Each recording was around one minute long with processing done on ten second segments.
-For each observation, the audio was played through the speaker with minimal activity around interferometer as the Pi script recorded the signal.
-The saved time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{scipy.io.wavfile.write}\cite{scipy} implementation in Python\cite{python}.
-
-These initial recordings were extremely noisy and painful to listen to, with a loud bass sound throughout.
-Spectra of all of these recordings showed a dominant mains hum at 50Hz with strong power up to and beyond the 8th harmonic thereof.
-This power is also present with the speaker off, as is shown in Figure~\ref{fig:psd_noise}, and also weakly when monitoring with the photodiode in complete darkness.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
-	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), show high power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the ambient lighting), otherwise is a fairly white spectrum except for the broad, shallow maximum at 750Hz}
-	\label{fig:psd_noise}
-\end{figure}
-
-\subsection{Applying filters to the data}
-
-Early attempts to remove the mains hum included zeroing the channel of each 50Hz harmonic, this is effectively using a rectangular comb filter.
-Applying a filter (multiplying by a function) in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
-While this did remove the mains harmonics, it also destroyed the rest of the signal.
-
-Attempting to low-pass the signal did better (to the ear) but was also problematic as the harmonics are all over the 100Hz-1kHz region, which also carries a lot of the speech and music information.
-Low-passing the logarithm of the signal and then converting back did better still but doesn’t significantly remove the mains noise.
-
-Of these simple filters, the best performing was a band-pass Butterworth filter with frequency response shown in Figure~\ref{fig:butterworth}.
-This filter is `maximally flat' within the band region, here 150Hz to 3kHz, and trails off outside of the band with slope dependent on the order of the filter, here of order 5.
-This still suffered from the presence of the higher mains harmonics.
-However, behind the remaining noise one could hear something distinctly speech-like or musical, just with very low intelligibility, you couldn’t understand the speech but could recognise that someone was talking.
-
-% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
-Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-
-\begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{figures/butterworth_150_3000-cropped.pdf}
-	\caption{Butterworth bandpass filter frequency response, green vertical lines show the band limits of 150Hz and 3kHz.
-Note the very flat response within the band, any amplitude beyond -3dB is significant attenuation.}
-	\label{fig:butterworth}
-\end{figure}
-
-\subsection{The logMMSE estimator}
-% logMMSE
-
-\jam{Explain MMSE as in the second talk, be honest in some way}
-
-Speech enhancement for a noisy channel is a classic problem in signal processing.
-Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare various methods and find the best performing statistical technique to be the logMMSE (alt.\ MMSE-LSA) estimator which minimises the mean square error (MSE) of the log-spectral amplitude.
-This means that it takes the logarithm of the spectrum of the original signal and finds the fit (or estimate) that has the minimal MSE to that log-spectra.
-% Use of the logarithm is a common technique in speech processing as it, somewhat surprisingly, does not significantly change the intelligibility to the human ear.
-
-Applying an existing implementation of logMMSE\cite{logmmse} to the optical microphone recordings produces dramatically cleaner results.
-The time series and spectrum from before and after the logMMSE filter is applied are shown in Figures~\ref{fig:logMMSE_timeseries}~\&~\ref{fig:logMMSE_spectrum}.
-All of the mains harmonics are significantly attenuated.
-And further high-passing removes the low frequency curve beneath 100Hz.
-
-There is varied success with passing music passed through the filters.
-Loud, rhythmic chords and drums are received well and remain recognisable, while thinner, airy flute and violin melodies struggle to be heard in the results.
-It is likely that the lower sounds have higher amplitude speaker oscillations and so pass through the optical microphone better.
-
-However, speech intelligibility remains very low with the voice sounding mumbled and indistinct.
-Enhancement so far has removed most of the background sound, as evidenced by the near silence between words, but has not clarified the speech.
-There is a question here as to where the diction in the speech is lost, at the speaker-mirror coupling, at the photodiode reading, or if it is still there, just obscured in the noisy signal.
-
-% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
-However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
-An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
-% Visually it is impossible to confirm this for high frequencies, but it remains possible.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
-	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) logMMSE filter is applied, only one second of the one minute recording is shown.
-The bump at the start of the time series is the result of filtering a finite signal, and is expected.}
-	\label{fig:logMMSE_timeseries}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
-	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) logMMSE filter is applied, showing only up to 2kHz and a detail of the spectrum (limited to show only up to amplitude 2, while before filtering the 50Hz line has an amplitude of 40)}
-	\label{fig:logMMSE_spectrum}
-\end{figure}
-
-\section{Future work}
-\label{sec:future_work}
-
-\jam{BATTERY POWER}
-% any other wild speculation?
-
-\subsection{Finding the transfer function}
-
-The intensity, $I$, of the interference pattern at a particular angle is given by Equation~\ref{eq:intensity_result} below.
-Where A is the amplitude of the electric field of the beam, d is the separation (or the difference in the arm lengths), $\theta$ is the angle from the centre-line to the point on the screen, and $\lambda$ is the wavelength, 532nm, of the laser (see Appendix~\ref{app:intensity_derivation} for the derivation to get to this equation).
-\begin{equation}
-\label{eq:intensity_result}
-I = 2 A^2 (1+\cos{\phi}), \; \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{equation}
-
-This Equation~\ref{eq:intensity_result} means that the intensity changes non-linearly with a small change in separation, as in Equation~\ref{eq:intensity_change} (see Appendix~\ref{app:intensity_derivation} again for the derivation).
-This makes the problem of extracting the original signal more difficult.
-
-\begin{equation}
-\label{eq:intensity_change}
-	\frac{\delta I}{\delta d} = \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{equation}
-
-A full model of the transfer function would detail the system starting from the voltage sent to the speaker and ending with the value recorded by the Pi, including any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-Although we have the separation-intensity relation as non-linear for small motions (see above), the other components of this full model are yet to be determined.
-Figure~\ref{fig:pipeline_highlighted} shows this full pipeline, with the currently determined section highlighted.
-
-Theoretically, if we had the correct model, then signal recovery would only be limited by the SNR. We could also use long recordings of white noise to estimate a matched filter for the system.
-
-For further speech enhancement, the optimal filter would involve a hidden-Markov-model trained to English phonemes. Alternatively, machine learning solutions also exist through-out the field.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/pipeline_highlighted.pdf}
-	\caption{???}
-	\label{fig:pipeline_highlighted}
-\end{figure}
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-Using a Michelson interferometer, we able to successfully demonstrate the Viterbi algorithm recovering an injected signal.
-Just as is done is continuous gravitational wave searches.
-We are also able to play-back chordal or rhythmic music from an optical microphone (using the same interferometer), while recordings of simple speech are audible but unintelligible.
-
-\newpage
-\appendix
-
-\section{Open-source code}
-\label{app:code}
-This project is implemented in python3\cite{python} inside of jupyter notebooks\cite{jupyter}\cite{ipython} and uses the packages of numpy\cite{numpy}, scipy\cite{scipy}, matplotlib\cite{matplotlib}, tqdm\cite{tqdm}, and logmmse\cite{logmmse}.
-
-The current build can be found at:
-\url{https://github.com/daccordeon/gravexplain}
-% \begin{verbatim}
-% README
-% \end{verbatim}
-
-\section{Additional guide to reproducing results}
-\label{app:reproducing_results}
-% as for a 3rd year undergraduate
-
-Besides the directions given in the rest of this article, there are some observations and technical details useful to someone wanting to reproduce the results shown herein.
-For the analysis and signal processing performed the best reference is Appendix~\ref{app:code} for the open-source code.
-
-For the experiment, since a Michelson interferometer is such a standard configuration, we’ll not detail it further here, see any optics manual for more information.
-To obtain an independent speaker cone we recommend carefully dismantling a budget commercial speaker system to get at the speaker within.
-To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place.
-If the other speaker in the pair remains, keep it as far away from the experiment and as damped as possible else it may act as a second, interfering source.
-% we used a Class 2 LDS5 ThorLabs laser at 532nm
-
-For the photodiode circuit, we used a Raspberry Pi and an MCP3008 ADC to monitor the analog signal.
-This could also be done with the analog pins on an Arduino.
-Be careful of the gain through the LM358 op-amps in the circuit though, else the ADC in either case could threshold the signal.
-Also be ready to change the values in the Sallen-Key filter to match your own sampling rate.
-The photodiode was an OSRAM BPW21, we mounted it through a discarded piece of speaker mesh (similar to black cloth over a frame) and placed another piece over the face to not over expose and threshold the signal.
-For the rest of the circuit, we used standard components assembled on a breadboard as detailed in Appendix~\ref{app:circuit_diagram}.
-
-\section{The Viterbi algorithm}
-\label{app:viterbi}
-
-To find the Viterbi path through a spectrogram (spectra over time), given an M x N grid of M frequency rows and N time columns and a rate limit for the change of frequency over time that we convert into the number of allowed rows we can move while stepping from each column to the next (this limit was $\pm 1$ row/column in our case).
-The path must start from the first column and end at the last, taking a single step in each column along the way.
-Each element in the grid is the norm of the Fourier component of a frequency at a time.
-These are normalised between $(0, 1)$ to use as multiplicative weights, i.e.\ the total weight of a path is the product of all weights along that path.
-In fact, we take the logarithm and so the sum of all weights to avoid underflow.
-The best path has the highest weight.
-
-With the grid prepared, the Viterbi algorithm starts in the second column.
-In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, taking the greatest cumulative weight of those available, ‘nearby’ cells (the value of which was discovered in the previous iteration).
-For each cell in the current column, the algorithm notes this weight plus the cell’s own weight, along with the row index of which cell in the previous column it came from.
-Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
-That best path is the Viterbi path.
-
-\section{A common mistake with wandering frequency signals}
-\label{app:phase_gotcha}
-
-We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
-The common mistake is to just appeal to the standard form $\sin{(2 \pi f t)}$, where f would be a constant frequency, and write down $\sin{(2 \pi f(t) t)}$.
-However, accumulation of phase means that this produces a curve that looks very different to that desired.
-Instead, one should see that the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
-Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
-
-\section{Circuit design}
-\label{app:circuit_diagram}
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-
-Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
-	\caption{Circuit diagram for photodiode reading.
-Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC).
-The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.}
-	\label{fig:circuit_diagram}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{figures/circuit_pic3.pdf}
-	\caption{Photo of circuit for photodiode reading assembled on breadboard.}
-	\label{fig:circuit_pic3}
-\end{figure}
-
-\section{Intensity relation derivation}
-\label{app:intensity_derivation}
-
-We want the change in intensity, $\delta I$, at a point in the Michelson interference pattern, given a small change, $\delta d$, in the perpendicular path difference of the mirrors (or in the separation of one and virtual image of the other).
-The point is at angle $\theta$ on the screen above the central ray, noting circular symmetry.
-At this point, two rays meet, one from each mirror, with the total electric field (\jam{parallel to the screen?}) being $E_1(t) + E_2(t)$.
-Given an initial separation of $d$, the path difference between the mirrors is $2 d \cos{\theta}$ as the longer ray must travel to and from the further mirror, and both move up at the angle $\theta$ to meet at the screen.
-
-\begin{align}
-\label{eq:intensity_derivation}
-    E_1(t) &= A e^{i (k L - \omega t)} \\
-    E_2(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \\
-    I(t) &= \lvert A e^{i (k L - \omega t)} + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2
-\end{align}
-
-Collapse the phase difference to $\phi$ and expand the norm squared by expanding the complex exponentials.
-From there is it trigonometric manipulation to reach the result of intensity in terms of separation.
-
-\begin{align}    
-    I(t) &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
-    I(t) &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2) \\
-    I(t) &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
-    I(d) &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{align}
-
-Then take derivatives and find that the intensity change is non-linear.
-
-\begin{align}    
-    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
-    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
-    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
-    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{align}
-
-\begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
-
-The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
-
-This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
-
-Travel during the project was supported by the Australian National University PhB Science  program.
-
-\end{acknowledgments}
-
-\bibliographystyle{myunsrt}
-\bibliography{ifoDemoBib}
-
 \end{document}

### 2020-04-14 paper/ifo-singleTone.tex
some editing throughout the paper draft

\documentclass[paper-main.tex]{subfiles}

\begin{document}

% the webcam method
% how to analyse video, python OpenCV2

Continuous wave searches look for signals which are close to monochromatic. 
In this section we consider a simple sinusoidal tone; a constant note. 
As described in Section~\ref{sec:ifo}, the audio is played through a speaker fixed to the back of one of the mirrors.

The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam placed to view the screen (see Figure~\ref{fig:ifo_schematic_webcam}). 
The webcam's sample rate is $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
A tone under $15\,{\rm Hz}$ is played through the speaker for a minute and the interference pattern recorded with the webcam.

The resulting data is a video capture of the whole interference pattern. 
As only the frequency of the interference ring motion is needed, we take data from a single off-centre pixel indicated by the red dot in Figure~\ref{fig:interference_pattern}. 
We take the green-channel of the video as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).

We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
The two tones played together sound similar and don’t produce any noticeable beating.
To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
The experiment is able to recover a single tone confidently. 
%This showed that the experiment observes a single tone confidently.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of $2.09\,{\rm Hz}$. The peak (with amplitude $17.5$) was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
	\label{fig:webcam_spectrum}
\end{figure}

\end{document}

\begin{comment}
As described in Section~\ref{sec:introduction}, continuous gravitational wave searches look for near monochromatic signals. 
The audio analogue of this is a simple sinusoidal tone; a constant note.

The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam placed to view the screen (see Figure~\ref{fig:ifo_schmatic_webcam}). 
The webcam's sample rate is $30\,{\rm Hz}$, which limits the spectral content of observable signals to below 15Hz, the Nyquist frequency.
A tone under $15\,{\rm Hz}$ is played through the speaker for a mintue and the interference pattern recorded with the webcam.

The data is taken from a single off-centre pixelm, indicated by the red dot in Figure~\ref{fig:interference_pattern}. 
We take the green-channel of the video as an approximation to the total intensity as the laser is monochromatic green light (at $532\,{\rm nm}$).

We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. The two tones played together sound similar and don’t produce any noticeable beating.
To quantify the precision, we recorded a signal from an injected tone of 2.09Hz. From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\mathrm{Hz}$, a result within two standard deviations. This showed that the experiment observes a single tone confidently.

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the amplitude 17.5 peak was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at 0Hz, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
	\label{fig:webcam_spectrum}
\end{figure*}
\end{comment}

### 2020-04-14 paper/ifo-tableTopGWs.tex
some editing throughout the paper draft

--- 2020-02-14 paper/paper_main.tex
@@
-% GravExplain
-\documentclass[prb,preprint]{revtex4-1} 
-% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
-
-% Note that AJP uses the same style as Phys. Rev. B (prb).
-% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
-% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
-% We look forward to receiving your submission to AJP.
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb,amsthm}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{mathtools}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{hyperref}
-% \usepackage{siunitx}
-\usepackage{textcomp}
-
-\bibliographystyle{apsrev4-2}
-% \setlength{\parindent}{0pt}
-
-\newcommand{\jam}{\textcolor{magenta}}
-\newcommand{\han}{\textcolor{orange}}
+\documentclass[paper-main.tex]{subfiles}
 
@@
 
-% cover page
-\section*{Project contributions}
-% contributions clarify what’s to be assessed
-James W. Gardner:
-\begin{itemize}
-\item all paper sections other than the introduction, i.e.\ the abstract and from Section~\ref{sec:ifo} onwards, and all figures in the paper
-\item addition of the injection speaker, design of photodiode circuit from standard examples, assembly of circuit on breadboard, and set-up of Raspberry Pi to measure output
-\item all of the source code at \url{https://github.com/daccordeon/gravexplain} (containing over 1000 lines of Python scripts), that reads the data from the Raspberry Pi, performs all of the analysis on it, and creates all the figures used in the paper
-\item presentation of two talks based on the paper and the figures therein
-\end{itemize}
+The basic phyiscs of gravitational-wave detection can be demonstrated with a table-top Michelson iterferometer: a common optical configuration which can be used to measure changes in distance. 
+A simple Michelson interferometer comprises of a laser, a beamsplitter, and two mirrors. 
+Laser light is split evenly by a beamsplitter into two arms at right angles to each other. 
+The beams travel to the end of each arm where they are reflected by mirrors. 
+The beams then travel back to the beamsplitter where they recombine. 
+The combination of the two beams produces an interference pattern which is dependent on the relative distance traveled by each of the beams. 
+A change in arm length results in a change in the interference pattern.
 
-Hannah Middleton:
-\begin{itemize}
-\item the introduction of the paper, i.e.\ Section~\ref{sec:introduction}
-\item design and assembly of the interferometer
-\item continuous support and advice throughout the project
-\end{itemize}
+Gravitational-wave detectors are large, complex interferometer.
+The LIGO and Virgo detectors have $4\,{\rm km}$ and $3\,{\rm km}$ arms respectively.
+They are sensitive to miniscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$~\cite{GW150914}.
+However, their design is based on the Michelson interferometer. 
+Detection and analysis concepts can be demonstrated with table-top optical equipment.
 
-Andrew Melatos, Robin Evans, William Moran: other vital advice and comments throughout the project
-
-\newpage
-\section*{Statement of achievements}
-% one paragraph recount of what’s been done for two months, like the abstract
-The project spanned three weeks in December, four weeks in January, and two weeks in February.
-The first two weeks were spent writing the initial scripts and achieving the Viterbi results with the webcam set-up.
-This was relatively straight forward and the greatest hurdle was understanding the algorithm.
-
-The third week was spent setting up the photodiode circuit and the Raspberry Pi which involved constant testing and changing of the components and design of the circuit.
-All of January was spent on the optical microphone.
-Much difficultly was had combating 50Hz mains hum in the experiment.
-Eventually through many discussions at the signal processing meetings, we had a working optical microphone but low speech intelligibility.
-The last week of January was spent on methods of speech enhancement.
-If the investigation had continued we had much future work planned and possible solutions in mind.
-
-Finally, the start of February was spent writing the paper and giving presentations (one nine minute general talk and one 50 minute technical follow-up talk).
-
-\newpage
-
-\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
-
-\author{James W. Gardner}
-\email{u6069809@anu.edu.au}
-\affiliation{College of Science, Australian National University, Acton, ACT, 2601, Australia}
-
-\author{Hannah Middleton$^{(i)}$}
-\email{hannah.middleton@unimelb.edu.au}
-\author{Andrew Melatos$^{(i)}$}
-\email{amelatos@unimelb.edu.au}
-\author{Robin Evans$^{(ii)}$}
-\email{robinje@unimelb.edu.au}
-\author{William Moran$^{(ii)}$}
-\email{wmoran@unimelb.edu.au}
-\affiliation{School of Physics$^{(i)}$ and Department of Electrical and Electronic Engineering$^{(ii)}$, University of Melbourne, Parkville, VIC, 3010, Australia}
-\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}
-
-% Summer 2019/2020
-\date{\today}
-
-% AJP requires an abstract for all regular article submissions.
-\begin{abstract}
-Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings.
-However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
-
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
+A commonly-used analogy in explaining gravitational-wave science is sound, although it is important to note that gravitational waves are not sound. 
+Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}. 
+Sound vibrations are a simple means to move the components of a demonstration interferometer. 
+Audio files can be used as a simulated gravitational-wave signal in a table-top experiment. 
+The vibrations move the mirrors, changing the length of the interferometer arms, and the resulting interference pattern. 
+We use sound throughout this work as a means of demonstrating gravitaitonal-wave signals and the data analysis and signal processing techniques used to detect them.
 
 \begin{figure}
-	\includegraphics[width=0.9\textwidth]{figures/ligo_pic.pdf}
-	\caption{Aerial photo of the LIGO Hanford facility and its 4km long beam arms, courtesy of LIGO/Caltech/MIT}
-	\label{fig:ligo_pic}
+	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam.pdf}
+	\caption{Michelson interferometer schematic with interference pattern recorded by a commercial webcam.
+The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}.
+The beam is first reflected from a mirror a right angle (to save space on the optical breadboard) and then focused at a converging lens before the $50$/$50$ beamsplitter.
+The split beam is reflected from mirrors M1 and M2 at the ends of the interferometer arms and recombined at the beamsplitter.
+The interference pattern is enlarged by a diverging lens after the beamsplitter and projected onto a screen.
+A commertial speaker is fixed to the back of M2, as the signal input.
+The interferometer arms lengths are roughly $3\,{\rm cm}$ different and the entire interferometer (without the screen) fits within the dimensions of an A4 piece of paper.
+We did not include a compensating plate with the beamsplitter, as is often done.}
+	\label{fig:ifo_schematic_webcam}
 \end{figure}
 
-% general gw intro
-In 2015, gravitational waves were observed for the first time from the merger of two black holes in a binary system~\cite{GW150914}. 
-The observation, made by the Laser Interferometer Gravitational wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
-Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years, there has been increased public interest in gravitational wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to explain the topic to a general audience.
-Activities range from hands-on demonstrations, exhibitions, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online}, video games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
+The Michelson interferometer configuration used in this work is shown schematically in Figure~\ref{fig:ifo_schematic_webcam}.
+It is constructed on a $450 \times 450 \,{\rm mm} $ optical breadboard. 
+The laser (shown in the top-left in the figure) has a wavelength of $532\,{\rm nm}$.
+In this configuration, the laser beam is first reflected from a mirror which turn the beam $90^{\circ}$ (to save space on the optical breadboard). 
+The beam passes through a a converging lens and then to the beamsplitter, which is $50$/$50$ reflection/transmission. 
+The beams are reflected by mirrors at the end of each of the arm (\han{$\approx XX\,{\rm cm}$ long}) and are recombined at the beamsplitter. 
+After the beamsplitter there is a diverging lens to aid viewing of the interference pattern when it is projected onto a screen as shown.
 
-% what are gws and how do we detect them...
-Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are disturbances in space-time caused by the acceleration of massive objects. 
-The effect of gravitational waves is a change in lengths; a `stretching and squeezing' of the very distance between objects.
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance.
-As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
-These observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets}.
+A photograph of the interference pattern is shown in Figure~\ref{fig:interference_pattern}. 
+If the arm lengths were equal, then the interference pattern would be a single dot. 
+An intentional offset in the arm lengths of the interferometer produces a pattern with a ring of interference fringes. 
+This results in a phase difference between the two beams when they meet at the screen.
+This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
+For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
+When the path difference changes, the fringes stay circular but move radially in or out.
+A ring interference pattern is often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily seen.
 
-% continuous gws
-To date, while the network of gravitational wave observatories have only observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}, these observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencies.
-These continuous waves may emit at constant frequencies, or they may wander slowly in frequency over time. 
-One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
-LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
-The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
+\begin{figure}
+ \begin{center}
+  \includegraphics[width=.45\textwidth]{figures/webcam_still0_crop.pdf}
+ \end{center}
+ \caption{\label{fig:interference_pattern}
+ Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
+For scale, the central fringe is approximately $5\,{\rm mm}$ across.}
+\end{figure}
 
-% sound analogy
-Here, we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience; the analogy of listening to the universe is often used to explain gravitational wave observations.
- %; transient gravitational wave signals have been converted into audio.
-This can be demonstrated with a table-top interferometer by playing sound close enough to move the mirrors and change the interference pattern.
-Using audio of slowly wandering frequency tones to produce an analogue of a continuous gravitational wave signal.
-The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}, particularly for the wandering frequency demonstration.
+A $0.5\,{\rm W}$ speaker is fixed to the back of one of the mirrors (M2 in Figure~\ref{fig:ifo_schematic_webcam}), providing the audio source for the set-up. 
+The speaker is part of a commercial $3.5\,{\rm mm}$-jack, USB powered pair of speakers and was driven by a PC. 
+It is attached to the back of the mirror using commercial adhesive putty. 
+The other speaker in the pair is kept face-down and as far away from the interferometer as the cables allow to reduce it acting as a second noise source.
 
-In this work we detail our table-top interferometer design in Section~\ref{sec:ifo} and describe observing a single note from a speaker in Section~\ref{sec:single_tone}.
-In Section~\ref{sec:viterbi_wandering}, we demonstrate the analysis techniques for observing a wandering frequency signal.
-In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech.
-This demonstration of an `optical microphone' serves as a more general exhibition of signal processing.
-We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
+As the speaker deforms to play sound, the mirror moves producing a change in arm length related to some coupling function of the amplitude and frequency of the audio. 
+The motion of the interference fringes follows the motion of the mirror, and therefore the speaker, at the same frequency. 
+If the fringe motion is small, then changes in the intensity at some point on the screen follows the frequency of the injected audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
+If the fringe motion is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting artificially raises the measured frequency.
+As such, any motion of the fringes must be kept small by playing sound softly through the speaker.
+For later sections, we derive the non-linear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen, this can be found in Appendix~\ref{app:intensity_derivation}.
+%\han{move over!}
 
-% \section{Method and results}
-\section{Table-top gravitational wave science}
-\label{sec:ifo}
+Data can be recorded via a webcam or photo-diode. 
+In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used as shown in Figure~\ref{fig:ifo_schematic_webcam}. 
+In Section~\ref{sec:optical_microphone}, which considers more complex audio, a photodiode is used as it provides a higher sample rate than the webcam.
 
+% Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
+We did not find it necessary to include it to be able to demonstrate the results desired.
+
+\begin{comment}
 Even though gravitational waves are caused by some of the most massive events in the universe, their observable effects on Earth are vanishingly small | on the order of only one part in $10^{21}$ for our best detectors~\cite{GW150914}.
@@
 When the path difference changes, the fringes stay circular but move radially in or out.
-
-\begin{figure}[h]
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_webcam.pdf}
-	\caption{Michelson interferometer schematic with interference pattern recorded by a commercial webcam.
-We used a Class 2 532nm laser~\cite{ThorLabsIFO} and a commercial speaker.
-The beam is focused at a converging lens before the 50/50 beamsplitter and the interference pattern is made visible by a diverging lens after the beamsplitter.
-The arms lengths between the speakers are roughly 3cm different and the entire interferometer (without the screen) fits within the dimensions of an A4 piece of paper.
-We did not include a compensating plate with the beamsplitter, as is often done.}
-	\label{fig:ifo_schematic_webcam}
-\end{figure}
 
@@
 \end{figure}
-
-% A note that while the interference pattern appears similar to the Airy pattern produced by a circular aperture, they are not in fact the same geometry.
-
-\section{Observing a single tone}
-\label{sec:single_tone}
-% the webcam method
-% how to analyse video, python OpenCV2
-
-As mentioned in Section~\ref{sec:introduction}, continuous gravitational wave searches look for near monochromatic signals. The audio analogue of this is a simple sinusoidal tone; a constant note.
-
-We measured the intensity at a point on the screen with a commercial USB webcam placed as to view the screen as in Figure~\ref{fig:ifo_schematic_webcam}.
-This webcam samples at 30Hz, which limits the spectral content of signals we can observe to below 15Hz, the Nyquist frequency.
-% See later Section~\ref{sec:photodiode} where we raise this limit into audible frequencies (20 Hz to 20kHz) with a photodiode.
-While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
-Then, at an off-centre pixel (shown as a red dot in Figure~\ref{fig:interference_pattern}), we took the green-channel of the video as approximate to the total intensity, as the laser is monochromatic green light (at 532nm).
-
-We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}.
-The two tones played together sound similar and don’t produce any noticeable beating.
-To quantify the precision, we recorded a signal from an injected tone of 2.09Hz.
-From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\mathrm{Hz}$, a result within two standard deviations.
-This showed that the experiment observes a single tone confidently.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the amplitude 17.5 peak was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
-The actual maximum is achieved at 0Hz, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
-	\label{fig:webcam_spectrum}
-\end{figure}
-
-\section{Observing a wandering tone}
-\label{sec:viterbi_wandering}
-% Tracking tones with the Viterbi algorithm
-% explaination and successful application
-% lean on the continuous gravitational waves angle
-
-As also mentioned in Section~\ref{sec:introduction}, continuous wave searches often look for signals that wander slowly in frequency~\cite{2019PhRvD.100l2002A}.
-The audio analogue of this is a tone that wanders in frequency; a note that changes pitch.
-Unlike the above Section~\ref{sec:single_tone} though, finding the signal isn’t as simple as searching for a peak in the spectrum for the full observing run.
-
-One method (among many) for continuous wave searches is to split the observed signal into intervals of time, compute the Fourier transform for each time interval, and then create a grid (known as a spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time.
-This method turns the problem into finding the best path through the spectrogram grid as if it were a weighted graph, where here the weights are the Fourier amplitudes.
-
-When finding the path, the method requires a limit for the allowed frequency change over time, usually informed by a physical model.
-Here, this frequency wander limit was set to 0.11 Hz per second to make it simply one frequency bin per every time bin, which will prove poorly adjusted to the signal.
-
-\subsection{The Viterbi algorithm}
-
-The Viterbi algorithm is a path-finding algorithm that is used in continuous wave searches~\cite{viterbi_application} to solve this problem.
-Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
-Then, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph.
-At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path.
-See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.
-
-\subsection{Results}
-
-We demonstrate this method of using the Viterbi algorithm by injecting a wandering frequency signal into the speaker and recording the output via the webcam.
-Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
-	\caption{Spectrogram of observed signal overlaid with pink dots for the injected signal and white crosses for the Viterbi path.
-Note the high correlation between them.
-Also note the deviations at the start time when the signal moves too quickly and just before the background anomaly at 150s.}
-	\label{fig:viterbi_overlay}
-\end{figure}
-
-Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, overlaid with the injected signal (the pink dots) and the recovered Viterbi path (the white crosses).
-To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment.
-Quantifying this, the Viterbi path stays within one frequency bin (approximately 0.3Hz) of the injected tone for 94\% of the run.
-Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals.
-To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often.
-There also appears to be an anomaly at 150s, likely from a disturbance (such as someone walking) around the interferometer.
-
-Empirically, decreasing the signal-to-noise ratio closer to unity (by making the speaker quieter) causes the Viterbi algorithm to stray more from the signal.
-However, since the only correlation through the grid is the signal, we always expect the Viterbi path to lie near to the signal no matter how high the noise is, and this is indeed what we see.
-
-\section{Playing back complex audio}
-\label{sec:optical_microphone}
-% Optical microphone
-% aim of optical microphone: play back injected sounds
-
-The natural succession to observing simple tones is to study complex audio signals, such as music and speech.
-The Viterbi algorithm and tracking more broadly is not applicable to these signals.
-Compare a single tone that slowly, continuously changes frequency to music and speech which consist of a broad spectrum that is constantly changing.
-Rather than track them, the interest here is in being able to record and play-back these complex audio signals.
-Using the interferometer to do so makes it an ‘optical microphone’, using light to capture sound.
-This device has precedence in the laser microphones~\cite{laser_microphone} of the defence industry which operate on a variety of related principles.
-
-Although this is indirectly related to the advanced signal processing required to extract gravitational wave signals, using an interferometer to capture sound can serve as an interesting demonstration to a broader physics and engineering audience.
-As such this could be introduced as a separate demonstration.
-
-\subsection{The photodiode method}
-\label{sec:photodiode}
-% advanced method: explain how we capture data
-
-For the optical microphone to be sensitive to speech and music it needs a frequency response (a sampling rate) far higher than the 30Hz of the webcam.
-In the context of audible frequencies (roughly 20Hz to 20kHz), speech intelligibility requires frequencies up to 3kHz and music requires frequencies up to and beyond 8kHz~\cite{speech_intelligibility}.
-Therefore the sampling rate of the optical microphone must be at least 16kHz to capture both speech and music.
-We achieved this with a photodiode.
-
-A photodiode is an electrical component that acts as a regular diode, blocking any current flow in the `reverse' direction, when no light is incident on it.
-But as the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.
-% Fundamentally, a current is created in the photodiode by the photo-electric effect.
-
-We placed a (OSRAM BPW21) photodiode in reverse-bias over an (LM358) op-amp to create a photo-detector that would produce a voltage dependent on the incident intensity.
-We placed the photodiode in the interference pattern at roughly the same off-centre spot as the webcam was monitoring previously (although we saw no difference when monitoring at other positions).
-The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.
-
-The voltage signal from the photo-detector was captured by a (MCP3008) 10-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2.
-Together, these sampled the signal at roughly 16kHz, our desired limit.
-
-However, by sampling, any frequency component of the analog signal above the Nyquist frequency of 8kHz would have been aliased down into the detected range.
-To prevent this, we included an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to 16kHz before the ADC.
-This component attenuated any frequencies above 8kHz, before they got digitally sampled.
-We also had to place another cloth screen over the face of the photodiode, as the reading without it, after gaining through the op-amps, was too high for the ADC.
-% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-
-A diagram and a photo of the circuit are shown in Appendix~\ref{app:circuit_diagram}.
-The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the entire optical microphone is shown in Figure~\ref{fig:setup_pic2}.
-
-\begin{figure}
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_photodiode.pdf}
-	\caption{Michelson interferometer schematic with the intensity of the interference pattern at a point measured by a photodiode.
-See Figure~\ref{fig:ifo_schematic_webcam} for details on the optics configuration.}
-	\label{fig:ifo_schematic_podo}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/setup_pic2.pdf}
-	\caption{Picture taken of the optical microphone set-up.
-Showing the Michelson interferometer, the photodiode (behind the cloth screen and in an insert), the circuit, and the Raspberry Pi.
-See Appendix~\ref{app:circuit_diagram} for a diagram and closer photo of the circuit.}
-	\label{fig:setup_pic2}
-\end{figure}
-
-\subsection{Initial results of the optical microphone}
-% breakdown of all the filters used along the way
-
-The optical microphone was tested on varied recordings of speech from different people and on music ranging from simple melodies to full track songs.
-Each recording was around one minute long with processing done on ten second segments.
-For each observation, the audio was played through the speaker as the Pi script recorded the signal from the photodiode.
-Care was taken to have minimal activity around interferometer while this was happening.
-The time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.
-
-When played back, these initial recordings were extremely noisy and had a loud bass hum throughout.
-Each Fourier spectrum showed dominant AC mains noise with power ranging from the fundamental of 50Hz up to and beyond the 8th harmonic thereof.
-With the speaker off this mains power was still present, as shown the power spectral density (effectively the norm squared of the Fourier spectrum) in Figure~\ref{fig:psd_noise}.
-And even with the speaker off and the photodiode in complete darkness, the mains hum was still present (with distribution similar to but weaker than that shown in Figure~\ref{fig:psd_noise}).
-
-\begin{figure}[h]%[H]
-	\includegraphics[width=\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
-	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), shows strong power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and even cooling), but otherwise is a fairly white spectrum.}
-	% except for the broad, shallow maximum at 750Hz}
-	\label{fig:psd_noise}
-\end{figure}
-
-\subsection{Applying filters to the data}
-\label{sec:simple_filters}
-
-The first attempt made to remove the mains hum was to zero the channel of each 50Hz harmonic in the Fourier spectrum.
-Effectively, this multiplied the spectrum by a rectangular comb filter.
-While this did remove the mains harmonics, it also audibly ruined the rest of the signal.
-This is because applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
-And the inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function that when convolved with significantly damages the signal.
-
-To fix this, we tried using a smoother comb filter but this too significantly attenuated the rest of the spectrum between each harmonic to hear the signal at all.
-
-Applying a high-pass filter with cut-off frequency 150Hz to the signal worked well at removing the 50Hz and 100Hz tones but was problematic as the mains harmonics extend far past 100Hz and so remain after filtering.
-This region above 100Hz carries a lot of the fundamental frequencies for speech and music~\cite{speech_intelligibility} and if the cut-off frequency is placed as to remove more of the mains harmonics, then the resultant signal is unrecognisable.
-
-We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass.
-
-\begin{figure}[h]%[H]
-	\includegraphics[width=0.8\textwidth]{figures/butterworth_150_3000-cropped.pdf}
-	\caption{Butterworth bandpass filter frequency response, any amplitude beyond -3dB is significant attenuation (half power).
-The green vertical lines show the band limits of 150Hz and 3kHz.
-Note the flat response within the band.}
-	\label{fig:butterworth}
-\end{figure}
-
-Band-passing the signal similarly doesn’t address the remaining harmonics above 100Hz but does low-pass away other noise at higher frequencies.
-The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over a band from 150Hz to 3kHz (beyond 2kHz speech information content drops significantly~\cite{speech_intelligibility}).
-For a low-pass Butterworth filter at cut-off frequency $\omega_c$ and gain $\epsilon$ the frequency response (attenuation at each frequency) follows Equation~\ref{eq:butterworth}.
-\begin{equation}
-\label{eq:butterworth}
-H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\frac{\omega}{\omega_c})^{2n}}}
-\end{equation}
-For a Butterworth band-pass filter, the above low-pass filter is combined with a high-pass filter of similar form.
-The response of the filter used is shown in Figure~\ref{fig:butterworth}.
-This filter improved on the above simpler filters but recordings of speech still had very low intelligibility, they could be recognised as speech but not understood.
-
-The ideal filter would be one that completely attenuates the undesired parts of the spectrum, does not change the rest of the spectrum, and has smooth edges as to not damage the signal through convolution.
-These three conditions are contradictory and any filter must compromise between them.
-The above simple filters fail at one or more of each of these conditions.
-The combs ruin the signal and the high-passes do not remove the noise.
-The Butterworth filter performed the best audibly as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
-
-% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
-Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-
-\subsection{The logMMSE estimator}
-% logMMSE
-
-Speech enhancement for a noisy channel is a classic problem in signal processing.
-Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compared various methods and found the best performing statistical technique among those at the time to be a logMMSE (alt.\ MMSE-LSA) estimator.
-This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
-Although this technique historically minimised with respect to the spectral amplitude~\cite{EphraimSpeechEU_MMSE}, throughout speech processing use of the logarithm seems produce better results as perceived by the human ear~\cite{SubjectiveComparison}.
-
-We used an existing implementation~\cite{logmmse} of the logMMSE estimator.
-Applying this estimator to the optical microphone recordings produced dramatically clean results.
-The time series (in Figure~\ref{fig:logMMSE_timeseries}) and spectrum (in Figure~\ref{fig:logMMSE_spectrum}) from before and after applying the logMMSE filter are shown.
-Comparing the before and after spectra in Figure~\ref{fig:logMMSE_spectrum} shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.
-
-After applying the estimator, simple chords and drums are heard well in played back music.
-While more complex melodies are sometimes completely absent.
-By observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
-This could be a perceptual effect of listening to the lossy recording, but seems more likely to be a frequency dependence somewhere in the optical microphone system.
-Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies and so instruments like electric bass and drums are louder in the results.
-
-After applying the estimator to speech recordings, intelligibility still remains low with the voice sounding mumbled and indistinct.
-The estimator removes most of the background noise, but does not clarify the speech.
-
-To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-This requires a better understanding of the system and further improvements to the experiment if true.
-% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
-However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
-An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
-	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) the logMMSE estimator is applied, showing one second of the one minute recording.
-The rise at the start of the bottom	 time series is an expected result from filtering a finite signal.}
-	\label{fig:logMMSE_timeseries}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
-	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) the logMMSE estimator is applied, showing only up to 2kHz and to amplitude 2 to see greater detail.}
-	\label{fig:logMMSE_spectrum}
-\end{figure}
-
-\section{Future work}
-\label{sec:future_work}
-% wild speculation
-We are currently working towards two future advancements for the optical microphone.
-
-The first would be to better isolate the experiment from the mains supply.
-The Raspberry Pi is USB powered and requires low power so it could even be run off a commercial external phone battery.
-Everything else in the circuit, including the photodiode, is connected to the Pi, so this would isolate them electrically as well.
-The laser is DC powered through a converter, so there is no issue there.
-With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
-
-The second advancement would be to achieve a full model of the transfer function of the system.
-Theoretically, if we had the correct transfer function, then signal recovery would only be limited by the signal-to-noise ratio as we could construct an optimal matched filter.
-This function would have to start from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-This sequence that the signal takes is shown in Figure~\ref{fig:pipeline_highlighted}.
-
-In our investigation, we have currently only derived part of this sequence, the relation between a deflection of the mirror and the change in the intensity of the interference pattern at some point.
-We find this relation to be non-linear, see Appendix~\ref{app:intensity_derivation} for the derivation.
-
-Asides from the transfer function, for extracting the speech there exists much literature~\cite{HMM_english} on using a hidden Markov model train to English phonemes to recognise speech.
-Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
-
-To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
-This could be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/pipeline_highlighted.pdf}
-	\caption{Sequence of relations for optical microphone transfer function. The currently derived relation is highlighted.}
-	\label{fig:pipeline_highlighted}
-\end{figure}
-
-\newpage
-\section{Conclusions}
-\label{sec:conclusions}
-
-% Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-
-We use sound and a table-top Michelson interferometer as an analogue to a gravitational wave detector.
-% This serves to demonstrate signal processing techniques.
-
-Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer.
-We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.
-
-Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the interferometer into an optical microphone.
-We demonstrated that this optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech.
-Dominant 50Hz mains hum in the recordings led to an exploration of speech enhancement techniques to remove it, including the well performing logMMSE estimate.
-From here we have many plans to further remove the mains hum and improve the intelligibility.
-
-These demonstrations have good potential to be adapted into the undergraduate laboratory.
-There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches.
-Most of all, the signal processing of speech enhancement is an interesting problem that’s relevant to digital communications.
-
-% \bigskip
-\newpage
-\appendix
-
-\section{Open-source code}
-\label{app:code}
-This project was implemented in python3~\cite{python} scripts as well as jupyter notebooks~\cite{jupyter}~\cite{ipython} and used the packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
-
-The current build and sample recordings of the optical microphone can be found at:
-\url{https://github.com/daccordeon/gravexplain}
-% \begin{verbatim}
-% README
-% \end{verbatim}
-
-\section{Implementation of the Viterbi algorithm}
-\label{app:viterbi}
-
-We used the Viterbi algorithm to find a signal in a spectrogram. Here we detail the implementation used.
-
-We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
-These are normalised between $(0, 1)$ to use as multiplicative weights.
-In fact, we take the logarithm and so the sum of all weights to avoid underflow.
-The best path has the highest weight among all paths.
-
-With the grid prepared, the algorithm starts in the second column.
-In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, selecting the previous cell with the greatest cumulative weight among those ‘nearby’ (the value of which was discovered in the previous iteration).
-Where the cumulative weight is the weight of the best path to reach that previous cell.
-The nearness of a previous cell is given by the limit on the allowed frequency change over time.
-
-For each cell in the current column, the algorithm calculates the cumulative weight to reach that cell by adding its weight to the cumulative weight to reach it.
-The algorithm also notes the row index of which cell in the previous column it best connects to.
-Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
-That discovered best path is the Viterbi path.
-
-\section{A common mistake for those new to wandering frequency signals}
-\label{app:phase_gotcha}
-
-We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
-A common mistake is to appeal to the standard form $\sin{(2 \pi f t)}$, where f is a constant frequency, and guess $\sin{(2 \pi f(t) t)}$.
-However, this fails to address the phase accumulated earlier in the signal.
-Instead, the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
-Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
-% This formula is also useful when performing frequency modulation.
-
-\section{Photodiode circuit design}
-\label{app:circuit_diagram}
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-
-Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
-This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
-
-\begin{figure}[h]%[H]
-	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
-	\caption{Circuit diagram for photodiode reading.
-Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC).
-The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.}
-	\label{fig:circuit_diagram}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{figures/circuit_pic3.pdf}
-	\caption{Photo of photodiode circuit assembled on breadboard.}
-	\label{fig:circuit_pic3}
-\end{figure}
-
-\section{Interferometric intensity relation derivation}
-\label{app:intensity_derivation}
-
-To obtain part of the transfer function for the optical microphone, we want to know the resultant change in intensity, $\delta I$, at a point in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
-Let the point be at angle $\theta$ on the screen (in the far-field) above the central ray as measured from the mirrors.
-Noting circular symmetry as to not specify exactly where in the fringe the point is.
-
-At this point, two rays meet, one from each mirror, with total electric field amplitude $E_1(t) + E_2(t)$.
-Given an initial difference between the arm lengths of $d$, the path difference between the two beams at the screen is $2 d \cos{\theta}$ as the longer beam must travel back and forth over the extra distance, and both travel at angle $\theta$ to meet the point~\cite{fringes:online}.
-
-The intensity at the point is then the norm squared of the total electric field amplitude as expressed in Equation~\ref{eq:intensity_derivation} below.
-
-\begin{align}
-    E_1(t) &= A e^{i (k L - \omega t)} \\
-    E_2(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \\
-    I(t) &= \lvert A e^{i (k L - \omega t)} + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2
-\label{eq:intensity_derivation}    
-\end{align}
-
-Collapse the phase difference to some value $\phi = 2 k d \cos{\theta}$.
-Then expand the norm squared using Euler’s relation, $e^{i \theta} = \cos{\theta} + i \sin{\theta}$.
-Here we stop considering intensity as a function of time as the time dependence will vanish.
-
-\begin{align}    
-    I &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
-    I &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2)
-\end{align}
-
-Expand and collect like terms, pull out a factor of two, and then recognise the form of $\cos{(\alpha-\beta)}$.
-
-\begin{align}
-    I &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
-    I &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{align}
-
-Finally, changing derivatives and expand $\phi$ to find that the change in intensity given a change in distance is non-linear, that ${\delta I}/{\delta d}$ is non-constant.
-
-\begin{align}    
-    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
-    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
-    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
-    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{align}
-
-This non-linearity in part of the transfer function will make the reverse problem of extracting the original signal more difficult.
-
-\begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
-
-The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
-
-This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
-
-Travel during the project was supported by the Australian National University PhB Science  program.
-
-\end{acknowledgments}
-
-\bibliographystyle{myunsrt}
-\bibliography{ifoDemoBib}
+\end{comment}
 

### 2020-04-14 paper/ifo-wanderingTone.tex
some editing throughout the paper draft

\documentclass[paper-main.tex]{subfiles}

\begin{document}

% Tracking tones with the Viterbi algorithm
% explaination and successful application
% lean on the continuous gravitational waves angle
Continuous wave signals may wander slowly in frequency over time. 
%As also mentioned in Section~\ref{sec:introduction}, continuous wave searches often look for signals that wander slowly in frequency~\cite{ScoX1O2Viterbi:2019}. 
The audio analogue of this is a tone that wanders in frequency; a note that changes pitch. 
Unlike the single tone described in Section~\ref{sec:single_tone}, finding the signal isn’t as simple as searching for a peak in the spectrum for the full timeseries data.%observing run.

\han{should we add some more astro details here about LMXBs and why we think they wander in frequency?}

The data from a gravitaitonal wave observation run can be many months to a year. 
One method to search for a slowly wandering continuous wave signal is to split the data into several shorter intervals. 
Each chunk is analysed seperately. 
In our project we take a Fourier transform of each interval and then create a grid (or spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time. 
(In continuous wave searches, a detection statistic is used to calculate the probability of a signal being present in a particular frequency bin.)
The task is then to find the best path through the spectrogram grid as if it where a weighted graph, where here the weights are the Fourier amplitudes. 
The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method find the best path through the grid. 
It is used in contiuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.

\subsection{The Viterbi algorithm}

%Here we overview the Viterbi algorithm and describe its usage in this work.

Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander). 
n, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph. 
At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path. 
See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.

% moved from above
The method requires a limit for the allowed frequency change over time, usually informed by a physical model. 
Here, the frequency wander limit is set to $0.11\,{\rm Hz}$ which corresponds to one frequency bin per every time bin. 
In some cases, we see this proves poorly adjusted to the signal.

\subsection{Results}

A wandering frequency signal is played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
	\caption{Results of playing a wandering tone into the interferometer. The spectrogram shows the observed freuqncy amplitudes at each time-frequency step. The overlaid pink dots and white crosses show the injected signal and recovered Viterbi path respetively. At the start (before $\sim 15\,{\rm s}$) the signal changes frequency too quickly for the Viterbi algorithm to recover. At $150\,{\rm s}$ there the data appears anomalous, which may be due to some background noise. }
	\label{fig:viterbi_overlay}
\end{figure*}

Our results are shown in Figure~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
The overlaid pick-dot and white-cross markers show the injected signal and recovered Viterbi path respectively. 
To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment \han{What was the SNR? / can we quantify this?}.
The Viterbi path stays within one frequency bin (approximately $0.3\,{\rm Hz}$) of the injected tone for $94\%$ or the run. \han{Should within one bin be the same as the wander limit above?}
Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals. 
To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often. 
There also appears to be an anomaly at $150\,{\rm s}$, likely from a disturbance (such as someone walking) around the interferometer.

Empirically, decreasing the signal-to-noise ratio closer to unity (by making the speaker quieter) causes the Viterbi algorithm to stray more from the signal. 
However, since the only correlation through the grid is the signal, we always expect the Viterbi path to lie near to the signal no matter how high the noise is, and this is indeed what we see. \han{Should we show results for this?}

\end{document}

### 2020-04-14 paper/paper_main.tex
some editing throughout the paper draft

@@
 % GravExplain
-\documentclass[prb,preprint]{revtex4-1} 
+\documentclass[prb,reprint]{revtex4-1} 
 % \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
@@
 \usepackage{textcomp}
+\usepackage{subfiles}
 
@@
 \begin{document}
+
+\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
+
+\author{James W. Gardner}
+\email{u6069809@anu.edu.au}
+\affiliation{%
+College of Science, Australian National University, Acton, ACT, 2601, Australia
+}%
+
+\author{Changrong Liu}
+\email{}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+
+\author{Hannah Middleton}
+\email{hannah.middleton@unimelb.edu.au}
+\affiliation{%
+ School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
+}
+
+\author{Andrew Melatos}
+\email{amelatos@unimelb.edu.au}
+\affiliation{%
+ School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
+}
+
+\author{Robin Evans}
+\email{robinje@unimelb.edu.au}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
+}
+
+\author{William Moran}
+\email{wmoran@unimelb.edu.au}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+\affiliation{
+}
+
+\author{\han{also invite the people who helped us with interferometer design to be authors??}}
+\email{wmoran@unimelb.edu.au}
+\affiliation{%
+Adelaide, Adelaide-OzGrav
+}
+
+% Summer 2019/2020
+\date{\today}
+
+\begin{abstract}
+Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy data. 
+We demonstrate these techniques by observing sound with a table-top Michelson interferometer. 
+In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
+We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings. 
+However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
+%Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data. 
+%We demonstrate these techniques by observing sound with a table-top Michelson interferometer. 
+%In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
+%We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings. 
+%However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
+\end{abstract}
+
+\maketitle
+
+\section{Introduction}
+\label{sec:introduction}
+\subfile{ifo-introduction.tex}
+
+\section{Table-top gravitational wave science}
+\label{sec:ifo}
+\subfile{ifo-tableTopGWs.tex}
+
+\section{Observing a single tone}
+\label{sec:single_tone}
+\subfile{ifo-singleTone.tex}
+
+\section{Observing a wandering tone}
+\label{sec:viterbi_wandering}
+\subfile{ifo-wanderingTone.tex}
+
+\section{Playing back complex audio}
+\label{sec:optical_microphone}
+\subfile{ifo-complexAudio.tex}
+
+\section{Future work}
+\label{sec:future_work}
+\subfile{ifo-futureWork.tex}
+
+\section{Conclusions}
+\label{sec:conclusions}
+\subfile{ifo-conclusions.tex}
+
+\begin{acknowledgments}
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
+The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer. 
+This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
+Travel during the project was supported by the Australian National University PhB Science program.
+
+\end{acknowledgments}
+
+\bibliographystyle{myunsrt}
+\bibliography{ifoDemoBib}
+
+\newpage
+\appendix
+\subfile{ifo-appendix.tex}
+
+\end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \email{u6069809@anu.edu.au}
-\affiliation{College of Science, Australian National University, Acton, ACT, 2601, Australia}
+\affiliation{%
+College of Science, Australian National University, Acton, ACT, 2601, Australia
+}%
 
-\author{Hannah Middleton$^{(i)}$}
+\author{Hannah Middleton}
 \email{hannah.middleton@unimelb.edu.au}
-\author{Andrew Melatos$^{(i)}$}
+\affiliation{%
+ School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
+}
+
+\author{Andrew Melatos}
 \email{amelatos@unimelb.edu.au}
-\author{Robin Evans$^{(ii)}$}
+\affiliation{%
+ School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
+}
+
+\author{Robin Evans}
 \email{robinje@unimelb.edu.au}
-\author{William Moran$^{(ii)}$}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
+}
+
+\author{William Moran}
 \email{wmoran@unimelb.edu.au}
-\affiliation{School of Physics$^{(i)}$ and Department of Electrical and Electronic Engineering$^{(ii)}$, University of Melbourne, Parkville, VIC, 3010, Australia}
-\affiliation{OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
 
@@
 % A note that while the interference pattern appears similar to the Airy pattern produced by a circular aperture, they are not in fact the same geometry.
-
-\section{Observing a single tone}
-\label{sec:single_tone}
-% the webcam method
-% how to analyse video, python OpenCV2
-
-As mentioned in Section~\ref{sec:introduction}, continuous gravitational wave searches look for near monochromatic signals. The audio analogue of this is a simple sinusoidal tone; a constant note.
-
-We measured the intensity at a point on the screen with a commercial USB webcam placed as to view the screen as in Figure~\ref{fig:ifo_schematic_webcam}.
-This webcam samples at 30Hz, which limits the spectral content of signals we can observe to below 15Hz, the Nyquist frequency.
-% See later Section~\ref{sec:photodiode} where we raise this limit into audible frequencies (20 Hz to 20kHz) with a photodiode.
-While playing a tone under 15Hz through the speaker for a minute, we recorded the interference pattern with the webcam.
-Then, at an off-centre pixel (shown as a red dot in Figure~\ref{fig:interference_pattern}), we took the green-channel of the video as approximate to the total intensity, as the laser is monochromatic green light (at 532nm).
-
-We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}.
-The two tones played together sound similar and don’t produce any noticeable beating.
-To quantify the precision, we recorded a signal from an injected tone of 2.09Hz.
-From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\mathrm{Hz}$, a result within two standard deviations.
-This showed that the experiment observes a single tone confidently.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the amplitude 17.5 peak was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
-The actual maximum is achieved at 0Hz, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
-	\label{fig:webcam_spectrum}
-\end{figure}
 

### 2020-05-05 paper/ifo-complexAudio.tex
some text editing and schematic editing

@@
 The Viterbi algorithm and tracking more broadly is not applicable to these signals---music and speech are constantly changing over a broad spectrum. 
-Instead of tracking a changing frequency over time, here we are interested in being able to record the interference pattern and play back these complex audio signal. 
+Instead of tracking a changing frequency over time, here we are interested in being able to record the interference pattern and play back these complex audio signals. 
 We use the interferometer as an `optical microphone', using light to capture sound. 
@@
 
-\subsection{The photodiode method}
+\subsection{Recording data via photodiode}
 \label{sec:photodiode}
@@
 
-For the optical microphone to be sensitive to speech and music it needs a sampling rate far higher than the $30\,{\rm Hz}$ of the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering}.
+Speech and music recordings require a higher sample rate than that of the webcam ($30\,{\rm Hz}$) used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering}. 
 Audible frequencies are in the range $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
 Speech intelligibility requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$~\cite{speech_intelligibility}. 
-Therefore the optical microphone data requires a sample rate of at least $16\,{\rm kHz}$ to capture both speach and music. 
+Therefore the optical microphone data requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music. 
 We achieved this with a photodiode.
@@
 This creates a photo-detector that produces a voltage dependent on the incident intensity. 
-The photodiode recored the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} (however we saw no difference when monitoring at other positions in the interference pattern).
+The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} (however we see no difference when monitoring at other positions in the interference pattern).
 The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.
@@
 The voltage signal from the photo-detector is captured by a (MCP3008) $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
-Together, these sampled the signal at roughly $16\,{\rm kHz}$, our desired limit.
+Together, these sampled the signal at roughly $16\,{\rm kHz}$.
 
@@
 \begin{figure}
-	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_photodiode.pdf}
-	\caption{Michelson interferometer schematic with the intensity of the interference pattern at a point measured by a photodiode.
-See Figure~\ref{fig:ifo_schematic_webcam} for details on the optical configuration.}
+	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_photodiode_edit.pdf}
+	\caption{
+Michelson interferometer schematic. 
+The schematic is identical to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
+}
 	\label{fig:ifo_schematic_podo}
@@
 	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000-cropped.pdf}
-	\caption{Butterworth bandpass filter frequency response, any amplitude beyond -3dB is significant attenuation (half power).
-The green vertical lines show the band limits of 150Hz and 3kHz.
+	\caption{Butterworth bandpass filter frequency response, any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power).
+The green vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$.
 Note the flat response within the band.}

### 2020-05-05 paper/ifo-introduction.tex
some text editing and schematic editing

@@
 Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years, there has been increased public interest in gravitational wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to explain the topic to a general audience.
+Gravitational-wave detection has led to increased public and media interest in gravitational-wave science. 
+Gravitational-wave research groups around the world have produced demonstrations and activities to explain their research to general audiences. 
 Activities range from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}%\han{look up Sarah Farmer's project}.
@@
 % what are gws and how do we detect them...
-Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are disturbances in space-time caused by the acceleration of massive objects. 
-The effect of gravitational waves is a change in lengths; a `stretching and squeezing' of the distance between objects.
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
+Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}. 
+They are disturbances in spacetime caused by the acceleration of massive objects. 
+The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects.
+Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers, they use the interference of laser light to measure changes in distance. 
 %As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
-These observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
+Gravitational-wave observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
 Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
@@
 Continuous waves may be emitted at constant frequencies, or may wander slowly in frequency over time. 
-One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
+Rotating neutron stars in low mass X-ray binaries (LMXBs) are one of the targets for continuous-wave searches. 
 LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
-The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
+Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
 It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
@@
 % sound analogy
-Here, we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
-The analogy of sound is commonly used in explaining gravitational wave observations. 
-%; the analogy of listening to the universe is often used to explain gravitational wave observations.
-%; transient gravitational wave signals have been converted into audio.
-In this work we utilise this analogy by playing sound close to a Michelson interferometer, simulating a gravitational-wave signal. 
-Audio of constant or slowly wandering frequency tones provide an analogue of a continuous wave signal.
-Continuous wave data analysis techniques are also demonstrated. 
-The techniques used in this work are similar to those used to search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}, particularly for the wandering frequency demonstration.
+We use the analogy of sound to demonstrate gravitational-wave science with a table-top experiment (see also Section~\ref{sec:ifo}). 
+We use similar data analysis techniques to those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
+As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to signal processing and speech enhancement. 
+It can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.
 
-This demonstration can be used to explain the search for continuous waves to a general audience as well as undergraduate students. 
-The analysis methods applied here follow signal processing techniques used in some of the most recent searches for continuous waves~\cite{ScoX1O2Viterbi:2019}. 
-As an undergraduate lab experiment, this demonstration covers topics of interest to both physics and engineering students.
-
-In this work we detail our table-top interferometer design in Section~\ref{sec:ifo} and describe observing a single note from a speaker in Section~\ref{sec:single_tone}.
-In Section~\ref{sec:viterbi_wandering}, we demonstrate the analysis techniques for observing a wandering frequency signal.
+In this work we present a selection of demonstrations using different types of audio signals ranging from simple tones, analogous to continuous wave signals, to more complex audio of music and speech.
+In Section~\ref{sec:ifo} we detail our table-top interferometer design. 
+We demonstrate observing a single note from a speaker in Section~\ref{sec:single_tone}.
+In Section~\ref{sec:viterbi_wandering} we replace the single note with a slowly changing frequency and analyse the signal using techniques from continuous-wave searches. 
 In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech. 

### 2020-05-05 paper/ifo-singleTone.tex
some text editing and schematic editing

@@
 
-The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam placed to view the screen (see Figure~\ref{fig:ifo_schematic_webcam}). 
+The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam placed to view the screen (see Fig.~\ref{fig:ifo_schematic_webcam}). 
 The webcam's sample rate is $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
@@
 The resulting data is a video capture of the whole interference pattern. 
-As only the frequency of the interference ring motion is needed, we take data from a single off-centre pixel indicated by the red dot in Figure~\ref{fig:interference_pattern}. 
+As only the frequency of the interference ring motion is needed, we take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
 We take the green-channel of the video as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).
 
-We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
-The two tones played together sound similar and don’t produce any noticeable beating.
-To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
-From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
-The experiment is able to recover a single tone confidently. 
+As an initial inspection, we play the timeseries of the intensity using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
+The two tones sound similar by ear and don't produce any noticeable beating. 
+The result is quantified in Fig.~\ref{fig:webcam_spectrum}. 
+A tone of $2.09\,{\rm Hz}$ is played through the speaker and the data recorded via webcam. 
+Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording. 
+We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ \han{[James: I added in that this is the $1\sigma$ uncertainty, is that correct?] (where the error represents the $1\sigma$ uncertainty).}
+The demonstration is able to recover a single frequency tone to within two standard deviations. 
+%We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
+%The two tones played together sound similar and don’t produce any noticeable beating.
+%To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
+%From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
+%The experiment is able to recover a single tone confidently. 
 %This showed that the experiment observes a single tone confidently.
@@
 	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of $2.09\,{\rm Hz}$.
-The peak (with amplitude $17.5$) was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
-The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
-	\label{fig:webcam_spectrum}
+	\caption{\label{fig:webcam_spectrum}
+Single injected tone results. 
+The Fourier amplitude of the intensity timeseres recorded by webcam. 
+The frequency of the injection is $2.09\,{\rm Hz}$. 
+The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
+%The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
+The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.
+}
+
 \end{figure}

### 2020-05-05 paper/ifo-tableTopGWs.tex
some text editing and schematic editing

@@
 
-The basic phyiscs of gravitational-wave detection can be demonstrated with a table-top Michelson iterferometer: a common optical configuration which can be used to measure changes in distance. 
+The basic physics of gravitational-wave detection can be demonstrated with a table-top Michelson interferometer, an optical configuration commonly used in undergraduate labs. 
 A simple Michelson interferometer comprises of a laser, a beamsplitter, and two mirrors. 
-Laser light is split evenly by a beamsplitter into two arms at right angles to each other. 
-The beams travel to the end of each arm where they are reflected by mirrors. 
-The beams then travel back to the beamsplitter where they recombine. 
+Laser light is split by a beamsplitter into two arms at right angles to each other. 
+At the end of each arm is a mirror which reflects the beams back to the beamsplitter where they recombine. 
 The combination of the two beams produces an interference pattern which is dependent on the relative distance traveled by each of the beams. 
@@
 
-Gravitational-wave detectors are large, complex interferometer.
+Gravitational-wave detectors are large, complex interferometers.
 The LIGO and Virgo detectors have $4\,{\rm km}$ and $3\,{\rm km}$ arms respectively.
-They are sensitive to miniscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$~\cite{GW150914}.
+They are sensitive to miniscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$ ~\cite{GW150914}.
 However, their design is based on the Michelson interferometer. 
@@
 
-A commonly-used analogy in explaining gravitational-wave science is sound, although it is important to note that gravitational waves are not sound. 
-Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}. 
-Sound vibrations are a simple means to move the components of a demonstration interferometer. 
-Audio files can be used as a simulated gravitational-wave signal in a table-top experiment. 
+A commonly-used analogy in explaining gravitational-wave science is sound. 
+Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online} (however it is important to note that gravitational waves are not sound).
+Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
 The vibrations move the mirrors, changing the length of the interferometer arms, and the resulting interference pattern. 
-We use sound throughout this work as a means of demonstrating gravitaitonal-wave signals and the data analysis and signal processing techniques used to detect them.
+We use sound throughout this work as a means of simulating gravitational-wave-like signals and the data analysis and signal processing techniques used to detect them.
 
 \begin{figure}
-	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam.pdf}
-	\caption{Michelson interferometer schematic with interference pattern recorded by a commercial webcam.
+% 	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam.pdf}
+	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam_edit.pdf}
+	\caption{\label{fig:ifo_schematic_webcam}
+Schematic of the table-top Michelson interferometer. 
+The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}. 
+The beam is first reflected from a mirror at a right angle and the focused at a converging lens before the beamsplitter which is $50$/$50$ in reflection and transmission. 
+The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and are recombined at the beamsplitter. 
+The beam is enlarged by a diverging lens after the beamsplitter. 
+The interference pattern is projected onto a screen and recorded using a commercial webcam connected to a personal computer (PC). 
+A speaker fixed to the back of M2 provides the signal input from the PC.  
+%Michelson interferometer schematic with interference pattern recorded by a commercial webcam.
 The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}.
@@
 We did not include a compensating plate with the beamsplitter, as is often done.}
-	\label{fig:ifo_schematic_webcam}
+    }
 \end{figure}
 
-The Michelson interferometer configuration used in this work is shown schematically in Figure~\ref{fig:ifo_schematic_webcam}.
+The Michelson interferometer configuration used in this work is shown schematically in Fig.~\ref{fig:ifo_schematic_webcam}.
 It is constructed on a $450 \times 450 \,{\rm mm} $ optical breadboard. 
-The laser (shown in the top-left in the figure) has a wavelength of $532\,{\rm nm}$.
-In this configuration, the laser beam is first reflected from a mirror which turn the beam $90^{\circ}$ (to save space on the optical breadboard). 
-The beam passes through a a converging lens and then to the beamsplitter, which is $50$/$50$ reflection/transmission. 
-The beams are reflected by mirrors at the end of each of the arm (\han{$\approx XX\,{\rm cm}$ long}) and are recombined at the beamsplitter. 
-After the beamsplitter there is a diverging lens to aid viewing of the interference pattern when it is projected onto a screen as shown.
+The laser (top-left) has a wavelength of $532\,{\rm nm}$.
+In the configuration shown in Fig.~\ref{fig:ifo_schematic_webcam} the laser beam is first reflected from a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard). 
+The beam passes through a a converging lens and then to the beamsplitter, which is $50$/$50$ reflection and transmission. 
+The beams are reflected by mirrors at the end of each of the two arms (\han{$\approx XX\,{\rm cm}$ and $\approx XX\,{\rm cm}$ long}) and are recombined at the beamsplitter. 
+The resulting interference pattern is projected onto a screen and recorded by a webcam.
 
-A photograph of the interference pattern is shown in Figure~\ref{fig:interference_pattern}. 
-If the arm lengths were equal, then the interference pattern would be a single dot. 
-An intentional offset in the arm lengths of the interferometer produces a pattern with a ring of interference fringes. 
-This results in a phase difference between the two beams when they meet at the screen.
+The interference pattern is a ring of interference fringes as shown by the photograph in Fig.~\ref{fig:interference_pattern}. 
+The pattern is enlarged by a diverging lens between the beamsplitter and the screen to aid in viewing. 
+A change in the relative length of the interferometer arms results in the interference rings moving radially inwards or outwards. 
+Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
+%This results in a phase difference between the two beams when they meet at the screen.
 This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
@@
 When the path difference changes, the fringes stay circular but move radially in or out.
-A ring interference pattern is often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily seen.
+%A ring interference pattern is often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily seen.
 
@@
 
-A $0.5\,{\rm W}$ speaker is fixed to the back of one of the mirrors (M2 in Figure~\ref{fig:ifo_schematic_webcam}), providing the audio source for the set-up. 
+A $0.5\,{\rm W}$ speaker is fixed to the back of one of the mirrors (M2 in Fig.~\ref{fig:ifo_schematic_webcam}), providing the audio source for the set-up. 
 The speaker is part of a commercial $3.5\,{\rm mm}$-jack, USB powered pair of speakers and was driven by a PC. 
@@
 
-Data can be recorded via a webcam or photo-diode. 
-In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used as shown in Figure~\ref{fig:ifo_schematic_webcam}. 
+Data can be recorded via a webcam (as shown in Fig.~\ref{fig:ifo_schematic_webcam}) or photo-diode. 
+In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used. 
 In Section~\ref{sec:optical_microphone}, which considers more complex audio, a photodiode is used as it provides a higher sample rate than the webcam.

### 2020-05-05 paper/ifo-wanderingTone.tex
some text editing and schematic editing

@@
 The audio analogue of this is a tone that wanders in frequency; a note that changes pitch. 
-Unlike the single tone described in Section~\ref{sec:single_tone}, finding the signal isn’t as simple as searching for a peak in the spectrum for the full timeseries data.%observing run.
+%Unlike the single tone described in Section~\ref{sec:single_tone}, finding the signal isn’t as simple as searching for a peak in the spectrum for the full timeseries data.%observing run.
 
-\han{should we add some more astro details here about LMXBs and why we think they wander in frequency?}
+The analysis techniques used here are inspired by those used to search for continuous waves from spinning neutron in LIGO and Virgo data.
+Continuous wave searches are performed on long datasets (months to years in duration). 
+The frequency of gravitational-wave emission wanders significantly over the observation period.
 
-The data from a gravitaitonal wave observation run can be many months to a year. 
-One method to search for a slowly wandering continuous wave signal is to split the data into several shorter intervals. 
-Each chunk is analysed seperately. 
-In our project we take a Fourier transform of each interval and then create a grid (or spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time. 
-(In continuous wave searches, a detection statistic is used to calculate the probability of a signal being present in a particular frequency bin.)
-The task is then to find the best path through the spectrogram grid as if it where a weighted graph, where here the weights are the Fourier amplitudes. 
+One method to search for a wandering signal is to split the data into several shorter segments which are analysed seperately initially. 
+The resulting segments are combined to form a grid in time and frequency (spectogram). 
+The value at each spectrogram point determins the likelihood that a signal is present at a particular time and frequency. 
+In gravitational-wave analysis, a detection statistic is used to calculate the likelihood of a signal (the statistic used depends on the type of target\han{refs}). 
+In this work we take a Fourier transfprm of each time segment.
+The resulting spectrogram gives the Fourier amplitude for each frequency and time bin. 
+The task is then to find a wandering frequency signal through the spectogram. 
 The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method find the best path through the grid. 
@@
 
+In section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work and in section~\ref{sec:wanderingResults} we present our results for recovery of a wandering audio frequency.
+
+%The data from a gravitaitonal wave observation run can be many months to a year. 
+%One method to search for a slowly wandering continuous wave signal is to split the data into several shorter intervals. 
+%Each chunk is analysed seperately. 
+%In our project we take a Fourier transform of each interval and then create a grid (or spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time. 
+%(In continuous wave searches, a detection statistic is used to calculate the likelihood of a signal being present in a particular frequency bin.)
+%The task is then to find the best path through the spectrogram grid as if it where a weighted graph, where here the weights are the Fourier amplitudes. 
+%The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method find the best path through the grid. 
+%It is used in contiuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
+
 \subsection{The Viterbi algorithm}
-
+\label{sec:viterbi}
 %Here we overview the Viterbi algorithm and describe its usage in this work.
@@
 Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
-n, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph. 
+At each iteration, the algorithm will find the best path to each node in the next sub-graph. 
+This is repeated to the final sub-graph. 
+%n, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph. 
 At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path. 
@@
 \subsection{Results}
+\label{sec:wanderingResults}
 
@@
 	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
-	\caption{Results of playing a wandering tone into the interferometer.
-The spectrogram shows the observed freuqncy amplitudes at each time-frequency step.
-The overlaid pink dots and white crosses show the injected signal and recovered Viterbi path respetively.
-At the start (before $\sim 15\,{\rm s}$) the signal changes frequency too quickly for the Viterbi algorithm to recover.
-At $150\,{\rm s}$ there the data appears anomalous, which may be due to some background noise.
-}
-	\label{fig:viterbi_overlay}
+	\caption{\label{fig:viterbi_overlay}
+Recovery of a wandering tone. 
+The spectrogram shows the observed frequency amplitdue at each time-frequency step. 
+The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path respetively. 
+On the left, before $\sim 15\,{\rm s}$, the signal changes frequency too quickly for the Viterbi algorithm to recover. 
+At $150\,{\rm s}$ the data appears anomalous, which may be due to some background noise. }
 \end{figure*}
@@
 To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment \han{What was the SNR? / can we quantify this?}.
-The Viterbi path stays within one frequency bin (approximately $0.3\,{\rm Hz}$) of the injected tone for $94\%$ or the run. \han{Should within one bin be the same as the wander limit above?}
+The Viterbi path stays within one frequency bin (approximately $0.3\,{\rm Hz}$) of the injected tone for $94\%$ or the run. \han{to check}
 Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals. 
@@
 
-Empirically, decreasing the signal-to-noise ratio closer to unity (by making the speaker quieter) causes the Viterbi algorithm to stray more from the signal. 
-However, since the only correlation through the grid is the signal, we always expect the Viterbi path to lie near to the signal no matter how high the noise is, and this is indeed what we see.
+%Empirically, decreasing the signal-to-noise ratio closer to unity (by making the speaker quieter) causes the Viterbi algorithm to stray more from the signal. 
+%However, since the only correlation through the grid is the signal, we always expect the Viterbi path to lie near to the signal no matter how high the noise is, and this is indeed what we see.
 \han{Should we show results for this?}

### 2020-05-05 paper/paper_main.tex
some text editing and schematic editing

@@
 
-\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
+%\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
+\title{Title ideas...
+\\\hspace{2em}\newline
+Demonstrating continuous gravitational wave searches in the lab with an optical microphone
+\\\hspace{2em}\newline
+An optical microphone analogy to continuous gravitational wave searches 
+\\\hspace{2em}\newline
+Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone
+\\\hspace{2em}\newline
+}
 
@@
 }%
-
-\author{Changrong Liu}
-\email{}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
 
@@
 }
-\affiliation{
+\affiliation{%
+\han{check all affiliations}
+}
+
+\author{Changrong Liu}
+\email{}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+\affiliation{%
+\han{check all affiliations}
 }
@@
 \begin{abstract}
-Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy data. 
-We demonstrate these techniques by observing sound with a table-top Michelson interferometer. 
-In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings. 
-However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
+Gravitational wave observatories around the world are searching for continuous waves: persistent signals from spinning neutron stars. 
+Searches use sophisticated statistical techniques to look for weak signals in noisy data. 
+We show how continuous wave searches can be demonstrated with a table-top model gravitational-wave detector (a Michelson interferometer). 
+Sound is used as an analogue to gravitational-wave signals in our model detector. 
+Using signal processing techniques from continuous wave searches we demonstrate the recovery of a constant and changing frequency tone. 
+Finally we investigate the performance of the demonstration as an `optical microphone'. 
+We test its recovery of music and speech. 
+Simple chords and drums are easily recovered, however speech recordings are very noisy. 
+Application of an existing speech enhancement technique provides some noise reduction, however the audio remains unintelligible. 
+This demonstration has scope for use in the physics and electrical engineering undergraduate lab. 
+It can also be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences. 
+%Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy data. 
+%We demonstrate these techniques by observing sound with a table-top Michelson interferometer. 
+%In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
+%We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings. 
+%However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
 %Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data. 
@@
 
-\section{Observing a single tone}
+\section{Single tone}
 \label{sec:single_tone}
@@
 
-\section{Observing a wandering tone}
+\section{Wandering tone}
 \label{sec:viterbi_wandering}
@@
 
-\section{Playing back complex audio}
+\section{Complex audio: music and speech}
 \label{sec:optical_microphone}

### 2020-05-13 paper/paper_main.tex
updating title

@@
 %\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
-\title{Title ideas...
-\\\hspace{2em}\newline
-Demonstrating continuous gravitational wave searches in the lab with an optical microphone
-\\\hspace{2em}\newline
-An optical microphone analogy to continuous gravitational wave searches 
-\\\hspace{2em}\newline
-Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone
-\\\hspace{2em}\newline
-}
+\title{Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
 

### 2020-05-27 paper/ifo-singleTone.tex
a few more edits to the draft

@@
 The webcam's sample rate is $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
-A tone under $15\,{\rm Hz}$ is played through the speaker for a minute and the interference pattern recorded with the webcam.
+A tone under $15\,{\rm Hz}$ is played through the speaker for one minute and the interference pattern recorded with the webcam.
 
@@
 Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording.
-We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ \han{[James: I added in that this is the $1\sigma$ uncertainty, is that correct?] (where the error represents the $1\sigma$ uncertainty).}
-The demonstration is able to recover a single frequency tone to within two standard deviations. 
+\han{[to do - rewrite once James has the new numbers (also Fig 3 caption)] 
+We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ (where the error represents the $1\sigma$ uncertainty).
+The demonstration is able to recover a single frequency tone to within two standard deviations. }
 %We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
@@
 The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
+\han{[to do: add new numbers from James]}
 %The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$.

### 2020-05-27 paper/ifo-tableTopGWs.tex
a few more edits to the draft

@@
 The beam passes through a a converging lens and then to the beamsplitter, which is $50$/$50$ reflection and transmission. 
-The beams are reflected by mirrors at the end of each of the two arms (\han{$\approx XX\,{\rm cm}$ and $\approx XX\,{\rm cm}$ long}) and are recombined at the beamsplitter. 
+The beams are reflected by mirrors at the end of each of the two arms ($\approx 7.5\,{\rm cm}$ and $\approx 10.0\,{\rm cm}$ long) and are recombined at the beamsplitter. 
 The resulting interference pattern is projected onto a screen and recorded by a webcam.

### 2020-05-27 paper/ifo-wanderingTone.tex
a few more edits to the draft

@@
 
-The analysis techniques used here are inspired by those used to search for continuous waves from spinning neutron in LIGO and Virgo data.
+The analysis techniques used here are inspired by those used to search for continuous waves from spinning neutron in LIGO and Virgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
 Continuous wave searches are performed on long datasets (months to years in duration). 
@@
 
-One method to search for a wandering signal is to split the data into several shorter segments which are analysed seperately initially. 
+One method to search for a wandering signal is to split the data into several shorter segments which are analysed separately initially. 
 The resulting segments are combined to form a grid in time and frequency (spectogram). 
-The value at each spectrogram point determins the likelihood that a signal is present at a particular time and frequency. 
-In gravitational-wave analysis, a detection statistic is used to calculate the likelihood of a signal (the statistic used depends on the type of target\han{refs}). 
-In this work we take a Fourier transfprm of each time segment.
+The value at each spectrogram point determines the likelihood that a signal is present at a particular time and frequency. 
+In gravitational-wave analysis, a detection statistic is used to calculate the likelihood of a signal (the statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}). 
+In this work we take a Fourier transform of each time segment.
 The resulting spectrogram gives the Fourier amplitude for each frequency and time bin. 
 The task is then to find a wandering frequency signal through the spectogram.
-The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method find the best path through the grid. 
-It is used in contiuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
 
-In section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work and in section~\ref{sec:wanderingResults} we present our results for recovery of a wandering audio frequency.
+The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method to find the most likely path of a signal given an observation sequence. 
+It is used in continuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
+
+In Section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work and in Section~\ref{sec:wanderingResults} we present our results for recovery of a wandering audio frequency.
 
@@
 %Here we overview the Viterbi algorithm and describe its usage in this work. 
+%Thinking about this - should we add some maths?
+%Our starting point is the spectrogram grid in time $t$ and frequency $f$. 
+%Each element in the grid contains the Fourier amplitude for a particular $f$ and $t$. 
+%Our objective is to find the most likely through the grid which maximises the summed Fourier amplitudes along the path. 
+%In our implementation we sum the log of the normalised 
+%We label the time-frequency grid points as $t_i$ and $f_j$ respectively, where $i=1,2,...N_t$ and $j=1,2,...,N_f$ (and $N_t$ and $N_f$ are the total number of time and frequency bins respectively). 
+%Starting at $t_{i=2}$, the algorithm iterates through $t_i$, at each stage finding the best path to every $f_j$ from $t_{i=0}$.
 
-Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
+Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount of frequency wander).
 At each iteration, the algorithm will find the best path to each node in the next sub-graph. 
@@
 %n, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph. 
+The Viterbi path is then the overall best path from the first sub-graph to the last. 
+The overall best path from the first sub-graph to the last is then selected, which is called the Viterbi path. 
 At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path. 
@@
 The method requires a limit for the allowed frequency change over time, usually informed by a physical model. 
-Here, the frequency wander limit is set to $0.11\,{\rm Hz}$ which corresponds to one frequency bin per every time bin. 
+In this work the frequency $f$ path can move in one of three ways at each time bin: a) up one frequency bin; b) stay in the same frequency bin; or c) down one frequency bin. 
+We set the frequency bin width to be $\Delta f = 0.3\,{\rm Hz}$ which corresponds to a maximum frequency wander of $0.11\,{\rm Hz}{\rm s}^{-1}$.
+%In this work, the frequency wander limit is set to $0.11\,{\rm Hz} {\rm s}^{-1}$ which corresponds to one frequency bin per every time bin. 
 In some cases, we see this proves poorly adjusted to the signal.
@@
 
-Our results are shown in Figure~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
+Our results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
 The overlaid pick-dot and white-cross markers show the injected signal and recovered Viterbi path respectively. 
-To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment \han{What was the SNR? / can we quantify this?}.
-The Viterbi path stays within one frequency bin (approximately $0.3\,{\rm Hz}$) of the injected tone for $94\%$ or the run. \han{to check}
-Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals. 
-To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often. 
-There also appears to be an anomaly at $150\,{\rm s}$, likely from a disturbance (such as someone walking) around the interferometer.
+The Viterbi path provides a good match to the injected frequency. 
+%To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment \han{What was the SNR? / can we quantify this?}.
+Quantitatively, the Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time. 
+%The Viterbi path stays within one frequency bin (approximately $0.3\,{\rm Hz}$) of the injected tone for $94\%$ of the run. 
+Initially the injected signal moves faster then the frequency wander limit, leading to the discrepancy for $t\lesssim 10\,{\rm s}$. 
+%Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals. 
+%To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often. 
+There is also an anomaly at $150\,{\rm s}$, which is likely due to a disturbance around the interferometer (such as someone walking nearby).
+%There also appears to be an anomaly at $150\,{\rm s}$, likely from a disturbance (such as someone walking) around the interferometer.
 

### 2020-06-16 paper/ifo-introduction.tex
a few more edits to the draft
working on paper comments from Andrew

@@
 % general gw intro
-In 2015, gravitational waves were observed for the first time from the merger of two black holes in a binary system~\cite{GW150914}. 
+In 2015, gravitational waves were observed for the first time from the merger of two black holes~\cite{GW150914}. 
 The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
@@
 Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}. 
-They are disturbances in spacetime caused by the acceleration of massive objects. 
+They are disturbances in spacetime caused by the acceleration of asymmetric massive objects. 
 The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects.
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers, they use the interference of laser light to measure changes in distance. 
+Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers.
+They use the interference of laser light to measure changes in distance. 
 %As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
-Gravitational-wave observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
+Gravitational-wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
 Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
@@
 % continuous gws
-To date the network of gravitational wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-However, the network is also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencies.
+To date the network of gravitational wave observatories has observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
+However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals.
 Continuous waves may be emitted at constant frequencies, or may wander slowly in frequency over time. 
-Rotating neutron stars in low mass X-ray binaries (LMXBs) are one of the targets for continuous-wave searches. 
-LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
+For example, rotating neutron stars in low mass X-ray binaries (LMXBs) are one of the targets for continuous-wave searches. 
+LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion, where the rotation frequency of the compact object wanders over time due to variable accretion of matter (and hence angular momentum transfer) from the companion~\cite{xraybinaries:1997}.
 Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
@@
 % sound analogy
-We use the analogy of sound to demonstrate gravitational-wave science with a table-top experiment (see also Section~\ref{sec:ifo}). 
-We use similar data analysis techniques to those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
-As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to signal processing and speech enhancement. 
-It can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.
+In this paper, we present an \han{inexpensive} table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves.
+As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to electronics, signal processing, and speech enhancement. 
+It gives students in courses such as physics and electrical engineering the freedom to explore the response of an accessible yet nontrivial optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
+The demonstration can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.
 
+The paper presents 
 In this work we present a selection of demonstrations using different types of audio signals ranging from simple tones, analogous to continuous wave signals, to more complex audio of music and speech.
-In Section~\ref{sec:ifo} we detail our table-top interferometer design. 
+In Section~\ref{sec:ifo} we detail the table-top interferometer design. 
 We demonstrate observing a single note from a speaker in Section~\ref{sec:single_tone}.
-In Section~\ref{sec:viterbi_wandering} we replace the single note with a slowly changing frequency and analyse the signal using techniques from continuous-wave searches. 
+In Section~\ref{sec:viterbi_wandering} we replace the single note with a slowly changing frequency and analyse the signal using a hidden Markov model technique from continuous-wave searches. 
 In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech. 
@@
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
+Appendix~\ref{app:code} collects the software and scripts used to produce this work.
 

### 2020-06-16 paper/paper_main.tex
a few more edits to the draft
working on paper comments from Andrew

@@
 \affiliation{%
- School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
+ School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
 }
 \affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
 }
@@
 \affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
 }
@@
 Searches use sophisticated statistical techniques to look for weak signals in noisy data. 
-We show how continuous wave searches can be demonstrated with a table-top model gravitational-wave detector (a Michelson interferometer). 
-Sound is used as an analogue to gravitational-wave signals in our model detector. 
-Using signal processing techniques from continuous wave searches we demonstrate the recovery of a constant and changing frequency tone. 
-Finally we investigate the performance of the demonstration as an `optical microphone'. 
-We test its recovery of music and speech. 
-Simple chords and drums are easily recovered, however speech recordings are very noisy. 
-Application of an existing speech enhancement technique provides some noise reduction, however the audio remains unintelligible. 
+We show how continuous wave searches can be demonstrated with a table-top model gravitational-wave detector (a Michelson interferometer), where sound plays the role of gravitational waves. 
+Using signal processing techniques from continuous wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
+Finally we investigate the performance of the demonstration as an `optical microphone' for recovering music and speech. 
+Simple chords and drums are easily recovered, but speech is more challenging. 
+%however speech recordings are very noisy. 
+%Application of an existing speech enhancement technique provides some noise reduction, however the audio remains unintelligible. 
 This demonstration has scope for use in the physics and electrical engineering undergraduate lab. 

### 2020-06-16 paper/ifo-appendix.tex
going through a few more comments from Andrew

@@
 \label{app:code}
-This project was implemented in python3~\cite{python} scripts as well as jupyter notebooks~\cite{jupyter}~\cite{ipython} and used the packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
-
-The current build and sample recordings of the optical microphone can be found at:
+This project is implemented in python3~\cite{python} scripts as well as jupyter notebooks~\cite{jupyter}~\cite{ipython} and used the packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
+The current build and sample data can be found at:
 \url{https://github.com/daccordeon/gravexplain}
@@
 
-\section{Implementation of the Viterbi algorithm}
-\label{app:viterbi}
-
-We used the Viterbi algorithm to find a signal in a spectrogram. Here we detail the implementation used.
-
-We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
-These are normalised between $(0, 1)$ to use as multiplicative weights.
-In fact, we take the logarithm and so the sum of all weights to avoid underflow.
-The best path has the highest weight among all paths.
-
-With the grid prepared, the algorithm starts in the second column.
-In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, selecting the previous cell with the greatest cumulative weight among those ‘nearby’ (the value of which was discovered in the previous iteration).
-Where the cumulative weight is the weight of the best path to reach that previous cell.
-The nearness of a previous cell is given by the limit on the allowed frequency change over time.
-
-For each cell in the current column, the algorithm calculates the cumulative weight to reach that cell by adding its weight to the cumulative weight to reach it.
-The algorithm also notes the row index of which cell in the previous column it best connects to.
-Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
-That discovered best path is the Viterbi path.
-
-\section{A common mistake for those new to wandering frequency signals}
-\label{app:phase_gotcha}
-
-We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
-A common mistake is to appeal to the standard form $\sin{(2 \pi f t)}$, where f is a constant frequency, and guess $\sin{(2 \pi f(t) t)}$.
-However, this fails to address the phase accumulated earlier in the signal.
-Instead, the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
-Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
-% This formula is also useful when performing frequency modulation.
-
-\section{Photodiode circuit design}
-\label{app:circuit_diagram}
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-
-Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
-This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
-
-\begin{figure*}
-	\includegraphics[width=.65\textwidth]{figures/circuit_diagram_2.pdf}
-	\caption{Circuit diagram for photodiode reading.
-Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC).
-The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.}
-	\label{fig:circuit_diagram}
-\end{figure*}
-
-\begin{figure*}%[H]
-	\includegraphics[width=0.5\textwidth]{figures/circuit_pic3.pdf}
-	\caption{Photo of photodiode circuit assembled on breadboard.}
-	\label{fig:circuit_pic3}
-\end{figure*}
-
-\section{Interferometric intensity relation derivation}
+\section{Interferometric relation between path length and optical intensity}
 \label{app:intensity_derivation}
@@
 
+\section{Implementation of the Viterbi algorithm}
+\label{app:viterbi}
+
+We used the Viterbi algorithm to find a signal in a spectrogram. Here we detail the implementation used.
+
+We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
+These are normalised between $(0, 1)$ to use as multiplicative weights.
+In fact, we take the logarithm and so the sum of all weights to avoid underflow.
+The best path has the highest weight among all paths.
+
+With the grid prepared, the algorithm starts in the second column.
+In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, selecting the previous cell with the greatest cumulative weight among those ‘nearby’ (the value of which was discovered in the previous iteration).
+Where the cumulative weight is the weight of the best path to reach that previous cell.
+The nearness of a previous cell is given by the limit on the allowed frequency change over time.
+
+For each cell in the current column, the algorithm calculates the cumulative weight to reach that cell by adding its weight to the cumulative weight to reach it.
+The algorithm also notes the row index of which cell in the previous column it best connects to.
+Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
+That discovered best path is the Viterbi path.
+
+%\section{A common mistake for those new to wandering frequency signals}
+%\label{app:phase_gotcha}
+%We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
+A common mistake is to appeal to the standard form $\sin{(2 \pi f t)}$, where f is a constant frequency, and guess $\sin{(2 \pi f(t) t)}$.
+However, this fails to address the phase accumulated earlier in the signal.
+Instead, the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
+Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
+%% This formula is also useful when performing frequency modulation.
+
+\section{Photodiode circuit design}
+\label{app:circuit_diagram}
+% schematic of breadboard and connections to pi
+% https://www.circuit-diagram.org/editor/
+% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
+
+Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
+It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
+A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
+This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
+
+\begin{figure*}
+	\includegraphics[width=.65\textwidth]{figures/circuit_diagram_2.pdf}
+	\caption{\label{fig:circuit_diagram}
+Circuit diagram for reading the photodiode. 
+The photodiode is connected in reverse-bias across an op-amp. 
+The analog op-amp output is passed through a Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC). 
+The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.
+}
+\end{figure*}
+
+\begin{figure*}%[H]
+	\includegraphics[width=0.5\textwidth]{figures/circuit_pic3.pdf}
+	\caption{\label{fig:circuit_pic3}
+Photodiode circuit assembled on breadboard.
+}
+\end{figure*}
+
 \end{document}

### 2020-06-16 paper/ifo-complexAudio.tex
going through a few more comments from Andrew

@@
 % aim of optical microphone: play back injected sounds
+A natural successor to simple tomes is complex audio, such as music and speech. 
+Complex audio signals are not quasimonochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{} is not applicable directly.
+Instead, in this section, we explore how to use the interferometer as an `optical microphone' to detect and play back complex audio signals, using light to capture sound. 
+In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{} and Appendix~\ref{}.
+The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
+It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}. 
+%It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae. [refs; my last sentence can be improved]
 
-A natural succession to observing simple tones is to use complex audio signal, such as music and speech. 
-The Viterbi algorithm and tracking more broadly is not applicable to these signals---music and speech are constantly changing over a broad spectrum. 
-Instead of tracking a changing frequency over time, here we are interested in being able to record the interference pattern and play back these complex audio signals. 
-We use the interferometer as an `optical microphone', using light to capture sound. 
-The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
-
+We describe additional components to the demonstration in Section~\ref{sec:photodiode} and our initial results in Section~\ref{sec:initialResultsOpMic}.
+Filter and speech enhancement techniques applied to the data are detailed in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse} respectively.
 This section is indirectly related to the advanced signal processing required to extract gravitational-wave signals. 
@@
 
-We describe additional components to the demonstration in Section~\ref{sec:photodiode} and our initial results in Section~\ref{sec:initialResultsOpMic}.
-Filter and speech enhancement techniques applied to the data are detailed in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse} respectively.
-
-\subsection{Recording data via photodiode}
+\subsection{Photodiode}
 \label{sec:photodiode}
@@
 
-A photodiode is an electrical component that acts as a regular diode wheen no light is incident on it, blocking any current flow in the `reverse' direction.
+A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the `reverse' direction.
 As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction. 
 % Fundamentally, a current is created in the photodiode by the photo-electric effect.
-We place a (OSRAM BPW21) photodiod in reverse-bias over an (LM358) op-amp. 
+We place a OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp. 
 This creates a photo-detector that produces a voltage dependent on the incident intensity. 
-The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} (however we see no difference when monitoring at other positions in the interference pattern).
-The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.
+The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
+We see no difference when monitoring at other positions in the interference pattern.
+The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.
 
-The voltage signal from the photo-detector is captured by a (MCP3008) $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
-Together, these sampled the signal at roughly $16\,{\rm kHz}$.
+The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
+Together, these sample the signal at roughly $16\,{\rm kHz}$.
 
@@
 This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
-We also place another cloth screen over the face of the photodiode, as the reading without it, after gaining through the op-amps, was too high for the ADC.
+We also place another cloth screen over the face of the photodiode, as the reading without it is too high for the ADC after gaining through the op-amps. 
 % It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
@@
 
-\subsection{Initial results of the optical microphone}
+\subsection{Anti-aliased output}
 \label{sec:initialResultsOpMic}
@@
 
-The optical microphone is tested on varied recordings of speech from different people and on music ranging from simple melodies to full track songs. 
+The optical microphone is tested on recordings of speech from different people and on music ranging from simple melodies to full track songs. 
 Each recording is around one minute long with processing applied to ten second segments. 
@@
 We find that initial recordings are extremely noisy with a loud bass hum throughout.
-Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond the $8$th harmonic thereof.
-We find that the mains power signal is present in data recorded without an injected audio signal as shown by the power spectral density (effectively the norm squared of the Fourier spectrum) in Figure~\ref{fig:psd_noise}.
+Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic.
+We find that the mains power signal is present in data recorded without an injected audio signal as seen in the power spectral density in Figure~\ref{fig:psd_noise}.
 The mains power is also present in data taken with the photodiode in complete darkness with a similar but weaker distribution as that shown in Figure~\ref{fig:psd_noise}).
@@
 	\includegraphics[width=.49\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
-	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off).
-We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and even cooling), but otherwise is a fairly white spectrum.}
-	% except for the broad, shallow maximum at 750Hz}
-	\label{fig:psd_noise}
+	\caption{\label{fig:psd_noise}
+Power spectral density (PSD) of the optical microphone  with no injected audio (i.e.\ with the speaker off). 
+We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). 
+Otherwise the PSD is fairly white. 
+}
 \end{figure}
 
-\subsection{Applying filters to the data}
+\subsection{Additional filters}
 \label{sec:simple_filters}
 
-Here we describe several approaches to remove the $50\,{\rm Hz}$ mains power signal and its harmonics.
+Here we describe several approaches to remove the $50\,{\rm Hz}$ mains hum and its harmonics.
 
@@
 Applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter. 
-The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function, which significantly damages the signal.
+The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function, which significantly corrupts the signal.
 
@@
 Applying a high-pass filter with cut-off frequency $150\,{\rm Hz}$ to the signal works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics.
-However, mains harmonics $> 100\,{\rm Hz}$ and remain after filtering. 
+However, mains harmonics above $100\,{\rm Hz}$ remain. 
 The region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}. 
-Therefore, although a high-pass filter with a higher cut-off removes more harmonics, the resultant signal is unrecognisable. 
+Therefore, although a high-pass filter with a higher cut-off removes more harmonics, the played-back signal is unrecognisable. 
 Applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.
@@
 Band-passing the signal similarly does not address the remaining harmonics above $100\,{\rm Hz}$. 
-However, it does low-pass away other noise at higher frequencies. 
-The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}).
+However, it does suppress other noise at higher frequencies. 
+The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}).
 For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\epsilon$, the frequency response (attenuation at each frequency) is,
@@
 The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
-This filter improves on the above simpler filters but recordings of speech still have very low intelligibility, they can be recognised as speech but not understood.
+This filter improves on the above simpler filters but recordings of speech still have low intelligibility.
+They can be recognised as speech without being understood.
 
-The ideal filter would be one that i.) completely attenuates the undesired parts of the spectrum, ii.) does not change the rest of the spectrum, and iii.) has smooth edges so as to not damage the signal through convolution.
+The ideal filter would be one that i) completely attenuates the undesired parts of the spectrum, ii) does not change the rest of the spectrum, and iii) has smooth edges so as to not damage the signal after convolution.
 These three conditions are contradictory and any filter must compromise between them. 
-The above simple filters fail at one or more of each of these conditions. 
-The combs ruin the signal and the high-passes do not remove the noise. 
-The Butterworth filter performs the best audibly as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
+The above simple filters fail one or more of i)--iii). 
+The Butterworth filter performs the best as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
 
@@
 
-\subsection{The logMMSE estimator}
+\subsection{logMMSE estimator}
 \label{sec:logmmse}
@@
 Speech enhancement for a noisy channel is a classic problem in signal processing. 
-Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods. 
-They find the best performing statistical technique considered to be a logMMSE (alt.\ MMSE-LSA) estimator. 
+Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
 This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
-Historically this technique minimises with respect to the spectral amplitude~\cite{EphraimSpeechEU_MMSE}. 
-However in speech processing, use of the logarithm seems to produce better results as percieved by the human ear~\cite{SubjectiveComparison}.
+The logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
 
-In this work we used an existing implementation~\cite{logmmse} of the logMMSE estimator. 
-Applying this estimator to the optical microphone recordings produced dramatically cleaner results.
+In this work we use an existing implementation~\cite{logmmse} of the logMMSE estimator. 
+Applying this estimator to the optical microphone output produced dramatically cleaner results.
 
 In Figures~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} we present results from playing a speech recording through the optical microphone. 
-Figure~\ref{fig:logMMSE_timeseries} shows the time series from before (top panel) and after (bottom panel) applying the logMMSE filter. 
+Figure~\ref{fig:logMMSE_timeseries} shows the anti-aliased photodiode time series from before (top panel) and after (bottom panel) applying the logMMSE filter. 
 Figure~\ref{fig:logMMSE_spectrum} shows the corresponding spectra before (top panel) and after (bottom panel) the logMMSE filter. 
@@
 After applying the estimator, simple chords and drums can be heard clearly. 
-However more complex melodies are sometimes completely absent from the resulting time series. 
+However more complex melodies are sometimes completely absent. 
 From observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
-This could be a perceptual effect of listening to the lossy recording, but seems more likely to be a frequency dependence somewhere in the optical microphone system. 
-Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies and so instruments like electric bass and drums are louder in the results.
+This could be a perceptual effect or a frequency dependence somewhere in the optical microphone. 
+Speculating, perhaps the speaker-mirror oscillates more at at low frequencies and so instruments like electric bass and drums are louder in the results.
 
-Overall, we find that the intelligibility of speech signals remains low. 
-The voice sounds mumbled and indistinct. 
+Overall, we find that the intelligibility of speech remains low. 
+The played-back voice sounds mumbled and indistinct. 
 The estimator removes most of the background noise, but does not clarify the speech. 
 To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-This requires a better understanding of the system and further improvements to the experiment if true.
+This requires a better understanding of the system.
 % The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.

### 2020-06-16 paper/ifo-conclusions.tex
going through a few more comments from Andrew

@@
 
-In this work we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector. 
-Signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirror. 
-We consider monochromatic and slowly-wandering frequency signals as an analogue to continuous wave signals as well as more complex audio recordings of speech and music. 
+In this paper we present an educational demonstration in which an \han{inexpensive}, table top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
+Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
+We consider quasimonochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g.
+from supernovae.
 % This serves to demonstrate signal processing techniques.
 
-In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of a monochromatic and wandering frequency signal. 
-A webcam is used to track the interference pattern and the injected signal is recovered succesffully with the Viterbi algorithm. 
+In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of signals with constant and slowly wandering frequency. 
+A webcam sampling at \han{XX}$\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered succesfully with the Viterbi algorithm. 
 These sections direcly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
@@
 
-Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the interferometer into an optical microphone.
+In Section~\ref{sec:complex_audio} we demonstrate recovery of more complex signals of music and speech. 
+Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
 We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 
-Dominant $50\,{\rm Hz}$ mains hum in the recordings led to an exploration of speech enhancement techniques to remove it, including the well performing logMMSE estimate. 
-Further avenues of exploration for this work include changes to the demonstration hardware to further remove the mains harmonics as well as investigating a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.
+Mains hum at $50\,{\rm Hz}$ and its harmonics dominates the photodiode output. 
+This offers students the opportunity to explore speech enhancement techniques to remove it, including the well performing logMMSE estimator. 
+Further avenues of exploration include a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.
 
 The demonstration presented here has good potential to be adapted for use in the undergraduate laboratory. 
-It can be used to demonstrate topics of interest to astrophysics and digital communications including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement.
+It can be used to demonstrate topics of interest to physicists and electrical engineers including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement.
 %These demonstrations have good potential to be adapted into the undergraduate laboratory.

### 2020-06-16 paper/ifo-singleTone.tex
going through a few more comments from Andrew

@@
 Continuous wave searches look for signals which are close to monochromatic. 
-In this section we consider a simple sinusoidal tone; a constant note. 
+In this section we consider a simple sinusoidal tone; a time at a single constant frequency. 
 As described in Section~\ref{sec:ifo}, the audio is played through a speaker fixed to the back of one of the mirrors.
@@
 
-The resulting data is a video capture of the whole interference pattern. 
-As only the frequency of the interference ring motion is needed, we take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
-We take the green-channel of the video as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).
-
+%The resulting data is a video capture of the whole interference pattern. 
+%As only the frequency of the interference ring motion is needed, w
+We take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
+The green-channel of the video is used as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).
 As an initial inspection, we play the timeseries of the intensity using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
@@
 	\caption{\label{fig:webcam_spectrum}
-Single injected tone results. 
-The Fourier amplitude of the intensity timeseres recorded by webcam. 
+Injected tone at a constant frequency: Fourier amplitude of the intensity time series recorded by the webcam. 
+%Single injected tone results. 
+%The Fourier amplitude of the intensity timeseres recorded by webcam. 
 The frequency of the injection is $2.09\,{\rm Hz}$. 

### 2020-06-16 paper/ifo-tableTopGWs.tex
going through a few more comments from Andrew

@@
 The basic physics of gravitational-wave detection can be demonstrated with a table-top Michelson interferometer, an optical configuration commonly used in undergraduate labs. 
-A simple Michelson interferometer comprises of a laser, a beamsplitter, and two mirrors. 
-Laser light is split by a beamsplitter into two arms at right angles to each other. 
+A simple Michelson interferometer comprises a laser, a beamsplitter, and two mirrors. 
+Laser light is split by a beamsplitter into two arms at right angles. 
 At the end of each arm is a mirror which reflects the beams back to the beamsplitter where they recombine. 
-The combination of the two beams produces an interference pattern which is dependent on the relative distance traveled by each of the beams. 
+The combination of the two beams produces an interference pattern which is dependent on the relative distance travelled by each of the beams. 
 A change in arm length results in a change in the interference pattern.
@@
 They are sensitive to miniscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$ ~\cite{GW150914}.
-However, their design is based on the Michelson interferometer. 
+However, their design is based fundamentally on the Michelson interferometer. 
 Detection and analysis concepts can be demonstrated with table-top optical equipment.
@@
 A commonly-used analogy in explaining gravitational-wave science is sound. 
-Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online} (however it is important to note that gravitational waves are not sound).
+Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}~\footnote{When introducing the acoustic analogy to lay audiences, it ins important to emphasize that gravitational waves are not sound.
+For example, gravitational waves can propagate in a vacuum, whereas sound cannot.}.
 Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
@@
 The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}. 
-The beam is first reflected from a mirror at a right angle and the focused at a converging lens before the beamsplitter which is $50$/$50$ in reflection and transmission. 
+The beam is first reflected from a mirror at a right angle and then focused at a converging lens before the beamsplitter which is $50$/$50$ in reflection and transmission. 
 The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and are recombined at the beamsplitter. 
@@
 In the configuration shown in Fig.~\ref{fig:ifo_schematic_webcam} the laser beam is first reflected from a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard). 
-The beam passes through a a converging lens and then to the beamsplitter, which is $50$/$50$ reflection and transmission. 
+The beam passes through a converging lens and then to the beamsplitter, which is $50$/$50$ reflection and transmission. 
 The beams are reflected by mirrors at the end of each of the two arms ($\approx 7.5\,{\rm cm}$ and $\approx 10.0\,{\rm cm}$ long) and are recombined at the beamsplitter. 
@@
  \caption{\label{fig:interference_pattern}
- Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
-For scale, the central fringe is approximately $5\,{\rm mm}$ across.}
+Michelson interference pattern as viewed by an off-axis webcam (which is why the pattern is elliptical). 
+The red dot marks where the intensity time series is measured. 
+The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
+% Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
+For scale, the central fringe is approximately $5\,{\rm mm}$ across.
+}
 \end{figure}
 
-A $0.5\,{\rm W}$ speaker is fixed to the back of one of the mirrors (M2 in Fig.~\ref{fig:ifo_schematic_webcam}), providing the audio source for the set-up. 
-The speaker is part of a commercial $3.5\,{\rm mm}$-jack, USB powered pair of speakers and was driven by a PC. 
+A $0.5\,{\rm W}$ speaker is fixed to the back of one of the mirrors (M2 in Fig.~\ref{fig:ifo_schematic_webcam}), providing a controllable source of vibrations. 
+The speaker is one half of a pair of commercial, USB-powered speakers fed by a $3.5\,{\rm mm}$ jack and driven by a PC. 
 It is attached to the back of the mirror using commercial adhesive putty. 
-The other speaker in the pair is kept face-down and as far away from the interferometer as the cables allow to reduce it acting as a second noise source.
+The other speaker in the pair is kept face-down and as far away from the interferometer as the cables to reduce its impact as a second source of vibrations.
 
-As the speaker deforms to play sound, the mirror moves producing a change in arm length related to some coupling function of the amplitude and frequency of the audio. 
-The motion of the interference fringes follows the motion of the mirror, and therefore the speaker, at the same frequency. 
-If the fringe motion is small, then changes in the intensity at some point on the screen follows the frequency of the injected audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
+%As the speaker deforms to play sound, the mirror moves producing a change in arm length related to some coupling function of the amplitude and frequency of the audio. 
+The motion of the interference fringes follows the motion of the mirror, and therefore the speaker. 
+If the fringe motion is small, then changes in the intensity at some point on the screen occur at the same frequency as the injected audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
 If the fringe motion is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting artificially raises the measured frequency.
 As such, any motion of the fringes must be kept small by playing sound softly through the speaker.
-For later sections, we derive the non-linear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen, this can be found in Appendix~\ref{app:intensity_derivation}.
-%\han{move over!}
+Even without over-counting, large fringe motions introduce a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome - but physically interesting!
+- phenomena like frequency doubling.
+We derive the non-linear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen in Appendix~\ref{app:intensity_derivation}.
 
@@
 In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used. 
-In Section~\ref{sec:optical_microphone}, which considers more complex audio, a photodiode is used as it provides a higher sample rate than the webcam.
+In Section~\ref{sec:optical_microphone}, which considers more complex audio, we turn to a photodiode because it offers a higher sample rate than the webcam.
 

### 2020-06-16 paper/ifo-wanderingTone.tex
going through a few more comments from Andrew

@@
 % lean on the continuous gravitational waves angle
-Continuous wave signals may wander slowly in frequency over time. 
+Continuous gravitational wave signals may wander slowly in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars~\cite{MelatosDouglassSimula:2015,Jones:2010} or stochastic accretion flows in binary neutron stars~\cite{BildstenTB:1998}.
 %As also mentioned in Section~\ref{sec:introduction}, continuous wave searches often look for signals that wander slowly in frequency~\cite{ScoX1O2Viterbi:2019}. 
@@
 The frequency of gravitational-wave emission wanders significantly over the observation period. 
+In this context, ``significantly'' means ``across multiple frequency bins'', where the typical width of a frequency bin is the reciprocal of the total observing time~\cite{JKS:1998,ScoX1O2Viterbi:2019}
 
-One method to search for a wandering signal is to split the data into several shorter segments which are analysed separately initially. 
-The resulting segments are combined to form a grid in time and frequency (spectogram). 
+One method to search for a wandering signal is to split the data into several shorter segments which are analysed individually. 
+The detection statistics in every segment are combined to form a grid in time and frequency (spectogram). 
 The value at each spectrogram point determines the likelihood that a signal is present at a particular time and frequency. 
-In gravitational-wave analysis, a detection statistic is used to calculate the likelihood of a signal (the statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}). 
-In this work we take a Fourier transform of each time segment.
+In gravitational-wave data analysis, a detection statistic is used to calculate the likelihood of a signal given the antenna beam pattern of the detector, which varies with as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
+The detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}). 
+In this work the detection statistic is the amplitude (modulus) of the Fourier transform of the data at each time segement, which maximizes the likelihood of detecting a sinusiod in Gaussian noise, as described in Appendix\han{to do}.
 The resulting spectrogram gives the Fourier amplitude for each frequency and time bin. 
-The task is then to find a wandering frequency signal through the spectogram.
-
+The task is then to find the most likely wandering frequency signal through the spectogram. 
 The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method to find the most likely path of a signal given an observation sequence. 
@@
 
-In Section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work and in Section~\ref{sec:wanderingResults} we present our results for recovery of a wandering audio frequency.
+In Section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work. 
+In Section~\ref{sec:wanderingResults} we present our results for recovery of a wandering audio frequency.
 
@@
 \label{sec:viterbi}
+
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{/home/hannahm/Documents/viterbi/quickTest/circlesTest}
+\caption{\label{fig:viterbi}
+A schematic diagram of the Viterbi algorithm. 
+Each box represents an element in the time-frequency grid starting from $t_0$ on the left and ending at $t_4$ on the right. 
+The color of the box indicates the likelihood that a signal is present in the box, where a lighter color corresponds to higher likelihood. 
+Each box has a single path leading to it indicated by the dotted lines. 
+The Viterbi path is highlighted in orange. 
+At each time step, the paths can move up one $f$ box, move down one $f$ box, or stay in the same $f$ box. 
+See the main text for a full description. 
+}
+\end{figure}
+
 %Here we overview the Viterbi algorithm and describe its usage in this work. 
-%Thinking about this - should we add some maths?
-%Our starting point is the spectrogram grid in time $t$ and frequency $f$. 
-%Each element in the grid contains the Fourier amplitude for a particular $f$ and $t$. 
-%Our objective is to find the most likely through the grid which maximises the summed Fourier amplitudes along the path. 
-%In our implementation we sum the log of the normalised 
-%We label the time-frequency grid points as $t_i$ and $f_j$ respectively, where $i=1,2,...N_t$ and $j=1,2,...,N_f$ (and $N_t$ and $N_f$ are the total number of time and frequency bins respectively). 
-%Starting at $t_{i=2}$, the algorithm iterates through $t_i$, at each stage finding the best path to every $f_j$ from $t_{i=0}$.
+%Thinking about this 
+Our starting point is the spectrogram grid indexed by time $t$ and frequency $f$ where there are $N_t$ and $N_f$ time and frequency bins respectively as shown in Fig.~\ref{fig:viterbi}.
+We label the elements of the grid as $t_i$ and $f_j$ with $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
+The value in each grid element is a detection statistic based on the time series data observed. 
+It indicates the likelihood that a signal is present in the element. 
+We label the normalised Fourier amplitude as $F(t_i,f_j)$. 
+%In continuous wave searches the detection statistic depends on the type of gravitational wave source. 
+%In this work we use the normalised Fourier amplitude, which we label $F(t_i,f_j)$.
+
+Our objective is to find the most likely path through the grid given the data and any probabilistic constraints on how the frequency wanders from $t_i$ to $t_{i+1}$. 
+In continuous wave searches, a physical model is used to decide how much the frequency of the signal is allowed to wander at each time step, which depends on the type of target. 
+In this work, at each time step we allow the frequency to either a) stay in the same bin; b) move up a single frequency bin; or c) move down a single frequency bin. 
+The probability of transitioning from $f_j$ at $t_i$ to any of $\{f_{j+1},f_j,f_{j-1}\}$ at $t_{i+1}$ is called the transition probability $A(f_k,f_m)$. 
+In this work the three options have equal probability $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise
+
+A path through the grid has a likelihood dependent on the detection statistic (the observation) and the transition probability.
+If the starting probabilities are all equal, i.e. $Pr[f(t_0)] = 1/N_f$, the the probability of the frequency equalling $f(t_{i+1})$ at $t = t_{i+1}$ is defined recursively by
+%If there is equal probability of starting $f_j$ at $t=t_0$, the probability of path $Q$ given the observation $O$ is 
+\begin{eqnarray}
+%P(Q|O) =& A(f_j) F(t=N_t,f_j), \nonumber \\ &~~ \times \dots, A(f_j) F (t=t_0,f_j).
+Pr[f(t_{i+1})=f_j] =~& F[t_{i+1},f(t_{i+1})] \nonumber \\
+                     &*A[f(t_{i+1}),f(t_i)]  \nonumber \\
+                     & *Pr[f(t_i)].
+\label{eqn:recursiveViterbi}
+\end{eqnarray}
+The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes $Pr[f(t_{N_t})] = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
+We can the maximize the latter quantity over $0 \leq j \leq N_f$ to find the optimal Viterbi path overall, i.e. terminating in any frequency bin.
+
+The Viterbi algorithm is an efficient method to find the Viterbi path. 
+Here we describe the algorithm as psudeo-code referring to the schematic in Fig.~\ref{fig:viterbi}, where the circles indicate the elements of the spectrogram grid (lighter colours correspond to higher likelihood).
+The implementation used in this work is available online (see Appendix~\ref{app:code}).
+\begin{enumerate}
+\item Starting at $t_1$, each $f_j$ element has three possible paths it could originate from at $t_0$, indicated by the lines in Fig.~\ref{fig:viterbi}. 
+At each $f_j$ element, the highest $F[t_0,f_j] A[f(t_0),f(t_1)]$ from the previous time step is selected, shown by the solid lines in Fig.~\ref{fig:viterbi}. 
+For example for the element labelled (a), the highest value is directly behind it, therefore the transition from $\{t_0,f_0\}$ to $\{t_1,f_0\}$ is selected. 
+Each element has now selected the best path from the previous time step. 
+In Fig.~\ref{fig:viterbi}, black lines show the best paths and grey lines show all other paths. 
+\item Moving to $t_2$, the path with the highest value of the recursive relation in equation~\ref{eqn:recursiveViterbi} is selected for each $f_j$ as shown by the solid lines from $t_1$ to $t_2$ in Fig.~\ref{fig:viterbi}.
+\item Steps 2 is repeated until the end of the grid is reached. 
+\item At $t=t_{N_t}$, each $f_j$ has a value $Pr[f(t_{N_t})=f_j]$ (equation~\ref{eqn:recursiveViterbi}). 
+The highest likelihood path for each $f_j$ can be seen by tracing the solid lines from right to left in Fig.~\ref{fig:viterbi}. 
+The $f_j$ with the maximum $Pr[f(t_{N_t})=f_j]$ is labelled (b) in Fig.~\ref{fig:viterbi}. 
+\item Tracing the p
+At the end of the grid, each $f_j$ has a unique path associated with it (seen by tracing the dotted lines from right to left in Fig.~\ref{fig:viterbi}).
+The element with the highest accumulated probability is selected, in the case of Fig.~\ref{fig:viterbi}, this is labelled (b).
+\item Tracing the path from right to left in Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c).
+This is the Viterbi path, which is highlighted in orange in Fig.~\ref{fig:viterbi}.
+\end{enumerate}
+
+\begin{comment}
+The route through the grid which maximises the overall probability
+
+Constraints are set on the allowable paths. 
+In continuous wave searches, a physical model is used to decide how much the frequency can wander at each timestep, depends on the type of target. 
+In this work we allow the frequency to either a) stay in the same bin; b) move up a single frequency bin; or c) move down a single frequency bin. 
+We label the probability of transitioning from $f_j$ at $t_{i}$ to $f_{j+1,j,j-1}$ at $t_{i+1}$ as $A_{i,j}$. 
+In this work the three transition option have equal probability.
+
+Our objective is now to find the most likely path through the grid given the data and transition constraints. 
+This is the route which maximises the overall probability, which we call the `Viterbi path'.
+In practice we sum the logarithm of the Fourier amplitude $\log O_{i,j}$ along the path.
+
+The Viterbi algorithm begins at $t_{i=1}$. 
+The algorithm selects the highest probability route for each $f_j$. 
+\begin{equation}
+A_{i,j} O_{i,j+1}
+A_{i,j} O_{i,j}
+A_{i,j} O_{i,j-1}
+\end{equation}
+At $t_{i=2}$ it then selects the best route to this point based on the accumulated probability along the paths so far. 
+For example, at
+
+(a.) The grid point labelled (a) has $f_{j=3}$ and $t_{i=1}$. 
+Looking back at the previous time, there are three options for where the path could have started from at $t_{i=0}$. These are $f_{j=4}$, $f_{j=3}$, or $f_{j=2}$. 
+The algorithm selects $f_{j=3}$ as it has the highest value. 
+(b.) The grid point labelled (b.) is not the best starting position for any of the cells at $t_{i=1}$, no paths starting from this position will be considered from now on.
+
+The probability of a path is given by the accumulated probability along the path. 
+\begin{equation}
+
+\end{equation}
+
+The value of each element in the grid is the logarithm of the normalised Fourier amplitude, which we label $O_{i,j}$.
+
+Our objective is to find the most likely path through the grid which maximises the , the `best path'. 
+In practice, we sum the logarithm of the Fourier amplitudes along the path.
+
+We label the time-frequency grid points as $t_i$ and $f_j$ respectively, where $i=1,2,...N_t$ and $j=1,2,...,N_f$ (and $N_t$ and $N_f$ are the total number of time and frequency bins respectively). 
+Starting at $t_{i=2}$, the algorithm iterates calculates the best path to each $f_j$ from $t_{i=1}$ to $t_=2$.  through $t_i$, at each stage finding the best path to every $f_j$ from $t_{i=0}$.
 
@@
 In some cases, we see this proves poorly adjusted to the signal. 
+\end{comment}
 
-\subsection{Results}
+\subsection{Sample output}
 \label{sec:wanderingResults}
 
-A wandering frequency signal is played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
-Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
+To test the Viterbi algorithm, a signal with wandering frequency is played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
+%Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
 
@@
 Recovery of a wandering tone. 
-The spectrogram shows the observed frequency amplitdue at each time-frequency step. 
-The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path respetively. 
+The spectrogram shows the observed frequency amplitudue at each time-frequency step. 
+The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path respectively. 
 On the left, before $\sim 15\,{\rm s}$, the signal changes frequency too quickly for the Viterbi algorithm to recover. 
@@
 
-Our results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
-The overlaid pick-dot and white-cross markers show the injected signal and recovered Viterbi path respectively. 
+The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
+The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path respectively maximized over all terminating frequency bins. 
 The Viterbi path provides a good match to the injected frequency. 
@@
 %The Viterbi path stays within one frequency bin (approximately $0.3\,{\rm Hz}$) of the injected tone for $94\%$ of the run. 
-Initially the injected signal moves faster then the frequency wander limit, leading to the discrepancy for $t\lesssim 10\,{\rm s}$. 
+Initially the injected signal moves faster then the frequency wander limit, leading to the discrepancy observed for $t\lesssim 10\,{\rm s}$. 
 %Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals. 
 %To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often. 
-There is also an anomaly at $150\,{\rm s}$, which is likely due to a disturbance around the interferometer (such as someone walking nearby).
+There is also an anomaly at $150\,{\rm s}$, which illustrates what happens when the interferometer is disturbed, e.g. when someone walks nearby. 
+The Viterbi algorithm is robust against such disturbances, recovering the optimal path promptly, as Fig.~\ref{fig:viterbi_overlay} shows. 
+%is likely due to a disturbance around the interferometer (such as someone walking nearby).
 %There also appears to be an anomaly at $150\,{\rm s}$, likely from a disturbance (such as someone walking) around the interferometer.

### 2020-06-17 paper/ifo-appendix.tex
working on the viterbi sectoin and further comments from Andrew

@@
 
-To obtain part of the transfer function for the optical microphone, we want to know the resultant change in intensity, $\delta I$, at a point in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
-Let the point be at angle $\theta$ on the screen (in the far-field) above the central ray as measured from the mirrors.
-Noting circular symmetry as to not specify exactly where in the fringe the point is.
+Part of the transfer function for the optical microphone relates change in intensity, $\delta I$, at a point $P$ in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
+Let $\theta$ be the angle between the optical axis and the ray from the centre of the relevant mirror to the point $P$. 
+%Let the point be at angle $\theta$ on the screen (in the far-field) above the central ray as measured from the mirrors. 
+%Noting circular symmetry as to not specify exactly where in the fringe the point is.
+At $P$, two rays meet, one from each mirror, with total electric field amplitude $E_1(t) + E_2(t)$ and 
+\begin{align}
+    E_1(t) &= A e^{i (k L - \omega t)},                       \label{eqn:E1} \\
+    E_2(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)}\,  \label{eqn:E2}, 
+\end{align}
+where $A$ is the amplitude, $L$ is the unperturbed arm length, $k$ is the wavenumber and $\omega$ is the angular frequency of the laser. 
+If the unperturbed difference between the arm lengths is $d$, then the path difference between the two rays is $2 d \cos{\theta}$~\cite{fringes:online}.
 
-At this point, two rays meet, one from each mirror, with total electric field amplitude $E_1(t) + E_2(t)$.
-Given an initial difference between the arm lengths of $d$, the path difference between the two beams at the screen is $2 d \cos{\theta}$ as the longer beam must travel back and forth over the extra distance, and both travel at angle $\theta$ to meet the point~\cite{fringes:online}.
+The intensity $I(t)$ at $P$ is then the norm squared of the total electric field amplitude,
+\begin{equation}
+I(t) = \lvert A e^{i (k L - \omega t)}  + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2\,.
+\label{eqn:intensity1}    
+\end{equation}
+Some elementary trigonometry yields
+\begin{equation}
+%I(t) = 2 A^2 (1 + \cos{\phi})
+I(t) = 2 A^2 \left[ 1 + \cos{\left( 2 k d(t) \cos{\theta} \right) } \right]  \label{eqn:intensity2}\,.
+\end{equation}
+Equation~\ref{eqn:intensity2} shows that $I(t)$ and $d(t)$ are related nonlinearly. 
+Near a bright fringe, where one has $\cos{[2kd\cos{(\theta)}]} \approx 1$, the nonlinearity is quadratic, with $\delta I \propto (\delta d)^2$ for small $\delta d$. 
+Consequently, an oscillation of the form $\delta d \propto \exp{( -i \Omega t )}$ in the arm length difference produces a DC signal and a frequency-double signal $\propto \exp{(-2i\Omega t)}$ in $\delta I$.
+This is discussed further in Section~\ref{}\han{to do}.
 
-The intensity at the point is then the norm squared of the total electric field amplitude as expressed in Equation~\ref{eq:intensity_derivation} below.
-
-\begin{align}
-    E_1(t) &= A e^{i (k L - \omega t)} \\
-    E_2(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \\
-    I(t) &= \lvert A e^{i (k L - \omega t)} + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2
-\label{eq:intensity_derivation}    
-\end{align}
-
-Collapse the phase difference to some value $\phi = 2 k d \cos{\theta}$.
+%Collapse the phase difference to some value $\phi = 2 k d \cos{\theta}$.
 Then expand the norm squared using Euler’s relation, $e^{i \theta} = \cos{\theta} + i \sin{\theta}$.
@@
 
-\begin{align}    
-    I =& A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
-    I =& A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 \nonumber\\ &+ (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2)
-\end{align}
+%\begin{align}    
+%    I =& A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
+%    I =& A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 \nonumber\\ &+ (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2)
+%\end{align}
 
-Expand and collect like terms, pull out a factor of two, and then recognise the form of $\cos{(\alpha-\beta)}$.
+%Expand and collect like terms, pull out a factor of two, and then recognise the form of $\cos{(\alpha-\beta)}$.
 
-\begin{align}
-    I =& A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} \nonumber\\&+ 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
-    I =& 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{align}
+%\begin{align}
+%    I =& A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} \nonumber\\&+ 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
+%    I =& 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
+%\end{align}
 
-Finally, changing derivatives and expand $\phi$ to find that the change in intensity given a change in distance is non-linear, that ${\delta I}/{\delta d}$ is non-constant.
+%Finally, changing derivatives and expand $\phi$ to find that the change in intensity given a change in distance is non-linear, that ${\delta I}/{\delta d}$ is non-constant.
 
-\begin{align}    
-    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
-    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
-    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
-    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{align}
+%\begin{align}    
+%    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
+%    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
+%    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
+%    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
+%\end{align}
 
-This non-linearity in part of the transfer function will make the reverse problem of extracting the original signal more difficult.
+%This non-linearity in part of the transfer function will make the reverse problem of extracting the original signal more difficult.
 
-\section{Implementation of the Viterbi algorithm}
+\section{Viterbi algorithm}
 \label{app:viterbi}
 
-We used the Viterbi algorithm to find a signal in a spectrogram. Here we detail the implementation used.
+This appendix contains some details regarding implementation of the Viterbi algorithm described in Section~\ref{sec:viterbi}. 
+The Viterbi algorithm is a classic method in signal processing, whose theoretical underpinnings and implementation are easily accessible to undergraduate students. 
+The interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001} to learn more.
 
-We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
+\han{to do! psudo code here}
+%We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
 These are normalised between $(0, 1)$ to use as multiplicative weights.
@@
 
-With the grid prepared, the algorithm starts in the second column.
+%With the grid prepared, the algorithm starts in the second column.
 In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, selecting the previous cell with the greatest cumulative weight among those ‘nearby’ (the value of which was discovered in the previous iteration).
@@
 
-For each cell in the current column, the algorithm calculates the cumulative weight to reach that cell by adding its weight to the cumulative weight to reach it.
+%For each cell in the current column, the algorithm calculates the cumulative weight to reach that cell by adding its weight to the cumulative weight to reach it.
 The algorithm also notes the row index of which cell in the previous column it best connects to.
-Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
+%Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
 That discovered best path is the Viterbi path.
@@
 
+%\section{Optical Components List}
+%\label{app:opComponents}
+%In this appendix we include a parts list for the optical components of the Michelson interferometer. 
+%The interferometer described in this paper uses parts from ThorLabs~\cite{}, with a total price of AU\$ 1120 at the time of purchase (July 2019). 
+%We also refer the reader to other examples of table-top interferometers at various price points~\cite{TTExhibit:2020,TTExhibit:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999}.
+
+%\begin{table*}
+%\begin{tabular}{llll}
+%\hline \hline
+%Item & Component & Quantity & Part number \\ 
+%\hline
+%1    & Aluminium optical breadboard $450\,{\rm mm} \times 450\,{\rm mm}$ & 1 & MB4545/M \\
+%\hline \hline
+%\end{tabular}
+%\end{table*}
+
 \end{document}

### 2020-06-17 paper/ifo-complexAudio.tex
working on the viterbi sectoin and further comments from Andrew

@@
 
-We describe additional components to the demonstration in Section~\ref{sec:photodiode} and our initial results in Section~\ref{sec:initialResultsOpMic}.
+We describe additional components to the demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}.
 Filter and speech enhancement techniques applied to the data are detailed in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse} respectively.
@@
 	\includegraphics[width=.6\textwidth]{figures/setup_pic2.pdf}
-\caption{Photograph of the optical microphone set-up.
+\caption{Photograph of the optical microphone.
 The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right.
@@
 One second of the one minute recording is shown.
-The rise at the start of the bottom time series is an expected result from filtering a finite signal.}
+The rise at the start of the bottom panel is expected when filtering a finite signal.}
 	\label{fig:logMMSE_timeseries}
@@
 	\includegraphics[width=.8\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
-	\caption{Spectrum of optical microphone recording of adult male voice shown before (top panel) and after (bottom panel) the logMMSE estimator is applied.
+	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} shown before (top panel) and after (bottom panel) the logMMSE estimator is applied.
 The spectrum shows up to $2\,{\rm kHz}$ and to amplitude $2$ to see greater detail.}

### 2020-06-17 paper/ifo-conclusions.tex
working on the viterbi sectoin and further comments from Andrew

@@
 In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of signals with constant and slowly wandering frequency. 
-A webcam sampling at \han{XX}$\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered succesfully with the Viterbi algorithm. 
+A webcam sampling at $30\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered succesfully with the Viterbi algorithm. 
 These sections direcly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
@@
 It can be used to demonstrate topics of interest to physicists and electrical engineers including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement.
+The detection of gravitational waves in 2015 led to increased excitement and public interest in gravitational wave science. 
+The table-top demonstration presented here can also be adapted as a tool to explain gravitational wave research developments for a wider non-specialist audience. 
+The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves.
+
 %These demonstrations have good potential to be adapted into the undergraduate laboratory.

### 2020-06-17 paper/ifo-futureWork.tex
working on the viterbi sectoin and further comments from Andrew

@@
 % wild speculation
-We are currently working towards two future advancements for the optical microphone.
+We are currently working towards two improvements of the optical microphone.
 
@@
 The Raspberry Pi is USB powered and requires low power so it could be run off a commercial external phone battery. 
-Everything else in the circuit, including the photodiode, is connected to the Pi, so this would isolate them electrically as well. 
+Everything else in the circuit, including the photodiode, is connected to the Pi. 
 The laser is DC powered through a converter, so should not introduce mains harmonics itself. 
-Using the interferometer in a dark room and a sealed container could better isolate if from ambient light and air currents.
-These changes may reduce or remove the $50\,{\rm Hz}$ harmonics from the data.
 %With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
 
-The second advancement would be to achieve a full model of the transfer function of the system. 
-Theoretically, if we had the correct transfer function, then signal recovery would only be limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
-The transfer function would need to start from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-The sequence that the signal takes is shown in Figure~\ref{fig:pipeline_highlighted}.
-In our investigation, we have currently only derived part of this sequence, the relation between a deflection of the mirror and the change in the intensity of the interference pattern at some point. 
-We find this relation to be non-linear, see Appendix~\ref{app:intensity_derivation} for the derivation.
+The second improvement is to measure the full model of the transfer function of the system. 
+Theoretically, with the correct transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
+The transfer function starts from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation Appendix~\ref{app:intensity_derivation}, and the photodiode reading.
+A schematic of the full optical microphone transfer function is shown in Figure~\ref{fig:pipeline_highlighted}.
 
-Asides from the transfer function, for extracting the speech there exists much literature~\cite{HMM_english} on using a hidden Markov model train to English phonemes to recognise speech.
+A voluminous literature exists about using a hidden Markov model trained against phonemes to recognize speech~\cite{HMM_english}. 
+Interestingly, this connects back to the Viterbi algorithm in Section IV, which is also underpinned by a hidden Markov model, albeit of a different type. 
 Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
 
-To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
-This could be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
+%To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
+%This could be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
 
-\begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/pipeline_highlighted.pdf}
-	\caption{Sequence of relations for optical microphone transfer function. The currently derived relation is highlighted.}
+\begin{figure*}
+	\includegraphics[width=.99\textwidth]{figures/pipeline_highlighted.pdf}
+	\caption{
+Flowchart for the optical microphone transfer function. 
+The highlighted stage is analysised in Appendix XX, the other stages are left to future work. 
+}
 	\label{fig:pipeline_highlighted}
-\end{figure}
+\end{figure*}
 

### 2020-06-17 paper/ifo-wanderingTone.tex
working on the viterbi sectoin and further comments from Andrew

@@
 \begin{figure}
-\includegraphics[width=0.49\textwidth]{/home/hannahm/Documents/viterbi/quickTest/circlesTest}
+\includegraphics[width=0.49\textwidth]{figures/viterbiDiagram.pdf}
 \caption{\label{fig:viterbi}
 A schematic diagram of the Viterbi algorithm. 
-Each box represents an element in the time-frequency grid starting from $t_0$ on the left and ending at $t_4$ on the right. 
-The color of the box indicates the likelihood that a signal is present in the box, where a lighter color corresponds to higher likelihood. 
-Each box has a single path leading to it indicated by the dotted lines. 
-The Viterbi path is highlighted in orange. 
-At each time step, the paths can move up one $f$ box, move down one $f$ box, or stay in the same $f$ box. 
-See the main text for a full description. 
+Each box represents an element in the time-frequency grid starting from the instant $t_0$ on the left and ending at $t_{N_t}$ on the right. 
+Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
+The color of each circle indicates the likelihood that a signal is present in the box, where a lighter color corresponds to higher likelihood. 
+AT each time step, the paths can move up one $f$ step, move down one $f$ step, or stay at the same $f$ step (shown by lines in the diagram). 
+Each circles has three paths leading to it indicated by the lines. 
+The best path to each circle at each $t_i$ is highlighted in black. 
+Some routes through the grid are found to be dead ends, such as the path ending at $t_1$ marked (a). 
+At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by equation~\ref{eqn:recursiveViterbi}, marked (b). 
+The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
 }
@@
 \end{eqnarray}
-The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes $Pr[f(t_{N_t})] = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
+The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes $Pr[f(t_{N_t}) = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
 We can the maximize the latter quantity over $0 \leq j \leq N_f$ to find the optimal Viterbi path overall, i.e. terminating in any frequency bin.
@@
 \item Starting at $t_1$, each $f_j$ element has three possible paths it could originate from at $t_0$, indicated by the lines in Fig.~\ref{fig:viterbi}. 
-At each $f_j$ element, the highest $F[t_0,f_j] A[f(t_0),f(t_1)]$ from the previous time step is selected, shown by the solid lines in Fig.~\ref{fig:viterbi}. 
+At each $f_j$ element, the highest $F[t_0,f_j] A[f(t_0),f(t_1)]$ from the previous time step is selected, highlighted by the black lines in Fig.~\ref{fig:viterbi}. 
 For example for the element labelled (a), the highest value is directly behind it, therefore the transition from $\{t_0,f_0\}$ to $\{t_1,f_0\}$ is selected. 
@@
 \item Moving to $t_2$, the path with the highest value of the recursive relation in equation~\ref{eqn:recursiveViterbi} is selected for each $f_j$ as shown by the solid lines from $t_1$ to $t_2$ in Fig.~\ref{fig:viterbi}.
-\item Steps 2 is repeated until the end of the grid is reached. 
+\item Steps 2 is repeated until the end of the grid ($t=t_{N_t}$) is reached. 
+Only the best paths are stored. 
 \item At $t=t_{N_t}$, each $f_j$ has a value $Pr[f(t_{N_t})=f_j]$ (equation~\ref{eqn:recursiveViterbi}). 
-The highest likelihood path for each $f_j$ can be seen by tracing the solid lines from right to left in Fig.~\ref{fig:viterbi}. 
-The $f_j$ with the maximum $Pr[f(t_{N_t})=f_j]$ is labelled (b) in Fig.~\ref{fig:viterbi}. 
-\item Tracing the p
-At the end of the grid, each $f_j$ has a unique path associated with it (seen by tracing the dotted lines from right to left in Fig.~\ref{fig:viterbi}).
-The element with the highest accumulated probability is selected, in the case of Fig.~\ref{fig:viterbi}, this is labelled (b).
-\item Tracing the path from right to left in Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c).
-This is the Viterbi path, which is highlighted in orange in Fig.~\ref{fig:viterbi}.
+The algorithm selects the terminating frequency $f(t_{N_t})$ with the highest value, marked (b) in Fig.~\ref{fig:viterbi}.
+\item The Viterbi path is then found by backtracking along the best paths at each time step. 
+In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c). 
+The Viterbi path is highlighed in orange in fig.~\ref{fig:viterbi}.
 \end{enumerate}
+The Viterbi algorithm is an example of a dynamical programming algorithm, where a computation can be broken down into a series of sub-computations.
+The Viterbi path for a time series sequences spanning $t_0$ to $t_{N_t}$ contains the Viterbi path for sub-sequences of that time series. 
+It is also very computationally efficient.
 

### 2020-06-18 paper/ifo-tableTopGWs.tex
adding a few more references

@@
 The LIGO and Virgo detectors have $4\,{\rm km}$ and $3\,{\rm km}$ arms respectively.
-They are sensitive to miniscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$ ~\cite{GW150914}.
+They are sensitive to minuscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$ ~\cite{GW150914}.
 However, their design is based fundamentally on the Michelson interferometer. 
@@
 
+\han{to do: add some text referring to other demo designs and references:~\cite{TTExhibit:2020,TTExhibit:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999}. }
 % Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.

### 2020-06-24 paper/ifo-complexAudio.tex
a few more edits in complex audio section

@@
 Here we describe several approaches to remove the $50\,{\rm Hz}$ mains hum and its harmonics. 
+We test these approaches on $1\,{\rm s}$ of data throughout this section as an illustrative example.
 
-Simply zeroing the frequency channel of each $50\,{\rm Hz}$ harmonic in the Fourier spectrum is unsuccessful. 
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{figures/freqSpectrumOriginalNoiseSubtracted.jpg}
+\end{center}
+\caption{\label{fig:noiseSubtract}
+Frequency spectrum of a $1\,{\rm s}$ signal before the use of filters. 
+The first panel shows the original frequency spectrum of the human voice recording (saying ``A cathode ...''). 
+The second panel shows the spectrum recorded after the signal passes through the optical microphone (sampled at $16\,000\,{\rm Hz}$).
+The third panel shows a noise-only recording from the optical microphone when no audio signal is being played. 
+The fourth panel shows the result of subtracting the noise-only spectrum in the third panel from the recording in the second panel. 
+We see no obvious improvement from this method. 
+}
+\end{figure*}
+
+Fig.~\ref{fig:noiseSubtract} shows the frequency spectrum of $1\,{\rm s}$ of recorded speech. 
+The first and second panels show the original signal before passing through the interferometer and the optical microphone recording of the signal respectively. 
+An intuitive means to remove noise is to subtract the noise spectrum of the instrument directly from the signal spectrum. 
+The third panel of Fig.~\ref{fig:noiseSubtract} shows a recording from the optical microphone with no audio signal. 
+We see that the output noise is coloured, which reflects the non-linearity of the optical microphone.
+The fourth panel of Fig.~\ref{fig:noiseSubtract} shows the frequency spectrum after the noise spectrum has been subtracted. 
+We see no obvious improvement, which may be due to time-variant noise.
+
+The next step is to apply filters to the data. 
+The ideal filter would be one that 
+\begin{itemize}
+\item[i)] completely attenuates the undesired parts of the spectrum, 
+\item[ii)] does not change the rest of the spectrum, and 
+\item[iii)] has smooth edges so as to not damage the signal after convolution. 
+\end{itemize}
+These three conditions are contradictory and any filter must compromise between them.
+
+We note that simply zeroing the frequency channel of each $50\,{\rm Hz}$ in the frequency spectrum is unsuccessful. 
 Effectively, this multiplies the spectrum by a rectangular comb filter. 
@@
 
-A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible.
+%A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible.
 
@@
 
-%We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass.
+%We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass
 
@@
 \label{eq:butterworth}
-H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\frac{\omega}{\omega_c})^{2n}}},
+H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\omega/\omega_c)^{2n}}},
 \end{equation}
 as a function of frequency $\omega$. 
-To construct a Butterworth band-pass filter, the above low-pass filter is combined with a high-pass filter of similar form. 
+To construct a Butterworth band-pass filter, the low-pass filter in Eq.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
 The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
+The Butterworth filter performs the well as it is optimised to be `maximally flat' within the band region, prioritising the condition ii) .
 This filter improves on the above simpler filters but recordings of speech still have low intelligibility.
 They can be recognised as speech without being understood.
+Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique.
 
-The ideal filter would be one that i) completely attenuates the undesired parts of the spectrum, ii) does not change the rest of the spectrum, and iii) has smooth edges so as to not damage the signal after convolution.
-These three conditions are contradictory and any filter must compromise between them. 
-The above simple filters fail one or more of i)--iii). 
-The Butterworth filter performs the best as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{figures/timeSeriesNotchWiener.jpg}
+\end{center}
+\caption{\label{fig:timeNotchWiener}
+}
+\end{figure*}
 
-\han{Is it worth also including some plots to show how the different filters perform? E.g. a PSD with comb filter, high pass filter, Butterworth filter?}
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{figures/freqNotchWiener.jpg}
+\end{center}
+\caption{\label{fig:freqNotchWiener}
+}
+\end{figure*}
+
+We also consider a combined notch filter and Wiener filter. 
+Timeseries and frequency spectrum results are shown in Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively. 
+The original signal is identical to that shown in Fig.~\ref{fig:noiseSubtract}.
+The first and second panels of Fig.~\ref{fig:timeNotchWiener} show the original audio signal and the recording after passing thought the optical microphone (i.e.
+the timeseries of the data shown in the first two panels of Fig.~\ref{fig:noiseSubtract}).
+
+A notch filter attenuates signals in specific frequency bands. 
+We apply a $15$ cascaded notched filter to remove the mains noise and its harmonics up to $750\,{\rm Hz}$. 
+\han{should we include here a mathematical description of the $15$ cascade notch filter?}
+The timeseries and frequency spectrum results are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}. 
+We seen that the mains harmonics are significantly attenuated in comparison to the analysis in Fig.~\ref{fig:noiseSubtract}.
+
+The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
+It minimizes the mean square error $E(\hat{x}-x)^2$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
+\han{should we add more details here too?}
+The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
+The timeseries is visibly cleaner, however there is a strong noise hum.
+
+The combination of the notch filter and the Wiener filter results are shown in the 
+%The ideal filter would be one that i) completely attenuates the undesired parts of the spectrum, ii) does not change the rest of the spectrum, and iii) has smooth edges so as to not damage the signal after convolution.
+%These three conditions are contradictory and any filter must compromise between them. 
+%The above simple filters fail one or more of i)--iii). 
+%The Butterworth filter performs the best as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
+%\han{see note from Andrew and ask James about this}
+
+%\han{Is it worth also including some plots to show how the different filters perform? E.g. a PSD with comb filter, high pass filter, Butterworth filter?}
 
@@
 \begin{figure*}
-	\includegraphics[width=.8\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
+	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
 	\caption{Time series of the optical microphone recording of an adult male voice (saying ``A cathode ...'') shown before (top panel) and after (bottom panel) the logMMSE estimator is applied.
@@
 \begin{figure*}
-	\includegraphics[width=.8\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
+	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
 	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} shown before (top panel) and after (bottom panel) the logMMSE estimator is applied.

### 2020-06-29 paper/ifo-complexAudio.tex
figures from Changrong and editing complex audio

@@
 The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
-The Butterworth filter performs the well as it is optimised to be `maximally flat' within the band region, prioritising the condition ii) .
+The Butterworth filter performs the well as it is optimised to be `maximally flat' within the band region (prioritising condition (ii) above).
 This filter improves on the above simpler filters but recordings of speech still have low intelligibility.
@@
 
-The combination of the notch filter and the Wiener filter results are shown in the 
+The combination of the notch filter and the Wiener filter produces a cleaner signal than either individually. 
+The time series and spectrum results are shown in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively.
+\han{todo - ask Changrong for some concluding remarks on intelligibility of the result using notch and wiener filter}
+
 %The ideal filter would be one that i) completely attenuates the undesired parts of the spectrum, ii) does not change the rest of the spectrum, and iii) has smooth edges so as to not damage the signal after convolution.

### 2020-06-29 paper/ifo-complexAudio.tex
adding some future work text from Changrongs notes

@@
 A natural successor to simple tomes is complex audio, such as music and speech. 
-Complex audio signals are not quasimonochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{} is not applicable directly.
+Complex audio signals are not quasimonochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not applicable directly.
 Instead, in this section, we explore how to use the interferometer as an `optical microphone' to detect and play back complex audio signals, using light to capture sound. 
-In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{} and Appendix~\ref{}.
+In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{sec:single_tone} and \han{to do: Appendix~\ref{}}.
 The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
-It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}. 
+It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
 %It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae. [refs; my last sentence can be improved]

### 2020-06-29 paper/ifo-futureWork.tex
a few more text changes
adding some future work text from Changrongs notes

@@
 \begin{document}
-\han{add more text here from Changrong's ideas}
 
 % wild speculation
-We are currently working towards two improvements of the optical microphone.
+%We are currently working towards two improvements of the optical microphone.
 
-The first is to better isolate the experiment from the mains supply. 
+The optical microphone experiment describe here can be used to convey a variety of topics as a teaching tool. 
+Here we consider a selection of further extensions to the experiment touching on noise isolation, transfer function analysis, speech recognition, and control systems.
+
+An initial improvement is to better isolate the experiment from interference, such as the mains supply. 
 The Raspberry Pi is USB powered and requires low power so it could be run off a commercial external phone battery. 
@@
 The laser is DC powered through a converter, so should not introduce mains harmonics itself. 
+Transferring the circuit from a breadboard to a printed circuit board will also reduce interference~\cite{elfekey2013design}
+
 %With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
 
-The second improvement is to measure the full model of the transfer function of the system. 
+A second improvement is to measure the full model of the transfer function of the system. 
 Theoretically, with the correct transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
@@
 
-A voluminous literature exists about using a hidden Markov model trained against phonemes to recognize speech~\cite{HMM_english}. 
+A voluminous literature exists about using a hidden Markov model (HMM) trained against phonemes to recognize speech~\cite{HMM_english}. 
 Interestingly, this connects back to the Viterbi algorithm in Section IV, which is also underpinned by a hidden Markov model, albeit of a different type. 
 Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
+
+Other traditional signal processing methods for extracting spectrogram like wavelet transform (WT) \citep{nason1995stationary} can also be implemented and compared with HMM.
+WT provides both time and frequency information, making it easier to pinpoint the origin of noise respect to time.
+In Refs. \cite{tufekci2000feature,agbinya1996discrete}, WT was also proposed for speech recognition.
+
+Finally, we can also consider how to improve the stability of the optical microphone components.
+A feedback loop control system can be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}.
 

### 2020-06-30 paper/ifo-appendix.tex
winter comments part 2

@@
 Consequently, an oscillation of the form $\delta d \propto \exp{( -i \Omega t )}$ in the arm length difference produces a DC signal and a frequency-double signal $\propto \exp{(-2i\Omega t)}$ in $\delta I$.
-This is discussed further in Section~\ref{}\han{to do}.
+This is discussed further in Section~\ref{}\han{to do}. \jam{[not quite sure how much further we can take discussion about this, are we wanting results that show the doubling?]}
 
@@
 
+\section{Likelihood of sinusoids in Gaussian noise}
+\label{app:sinusoid_likelihood}
+\jam{[to do!
+from Andrew: ``Write a short appendix proving that maximizing the likelihood exp(-n||n/2) with respect to A given a signal A cos (wt + phi) is the Fourier transform.
+Classic sig proc result useful for teachers to know.'']}
+
 \section{Viterbi algorithm}
@@
 
-\han{to do! psudo code here}
+\han{to do! pseudo-code here}
+\jam{[Hannah has already explained a lot about Viterbi, at this point is it worth just citing existing documentation for further details?
+I used Wikipedia in writing code, should we cite that?
+Andrew mentions going through the practical aspects, e.g.
+log to avoid underflow, is that worth writing up?
+My thought is only that all of this is technically available in the open source.]}
 %We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
@@
 A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
-This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
+This design was based off \jam{standard examples (I have no idea where I or Alex originally got the designs from, I'm confident though that they are common enough to not need citing given that we also give the diagram)} of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
 

### 2020-06-30 paper/ifo-complexAudio.tex
winter comments part 2

@@
 A natural successor to simple tomes is complex audio, such as music and speech. 
-Complex audio signals are not quasimonochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not applicable directly.
+Complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not applicable directly.
 Instead, in this section, we explore how to use the interferometer as an `optical microphone' to detect and play back complex audio signals, using light to capture sound. 
-In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{sec:single_tone} and \han{to do: Appendix~\ref{}}.
+In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood} \han{to do}.
 The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
@@
 The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2.
+\jam{[Hannah is right, the Pi first appears here.
+If we want to comment on its convenience and usage this is the place.]}
 Together, these sample the signal at roughly $16\,{\rm kHz}$.
@@
 The optical microphone is tested on recordings of speech from different people and on music ranging from simple melodies to full track songs. 
-Each recording is around one minute long with processing applied to ten second segments. 
-For each observation, the audio is played through the speaker whilst the Pi script records the signal from the photodiode. 
+Each recording is around one minute long with \jam{processing [this is poor, each recording was functionally ten seconds long, the fact that the sources were minutes long doesn't really matter]} applied to ten second segments.
+For each observation\jam{[again, poor wording, this means that for each 10 s recording we did the following]}, the audio is played through the speaker whilst the Pi script records the signal from the photodiode.
 During recordings, care is taken to ensure minimal activity around the interferometer to reduce environmental noise. 
@@
 
-We find that initial recordings are extremely noisy with a loud bass hum throughout.
+We find that initial recordings \jam{[Hannah's interpretation is correct here: "We find that the raw output of the optical microphone, collected without special precautions except anti-aliasing, is noisy, with a loud..."?]} are extremely noisy with a loud bass hum throughout.
 Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic.
+\jam{[non-linear appendix, Andrew says to cite Feynman lectures on non-linear oscillations]}
 We find that the mains power signal is present in data recorded without an injected audio signal as seen in the power spectral density in Figure~\ref{fig:psd_noise}.
@@
 The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination. 
+\jam{[No idea about the hump at 750 Hz. It is near to the main speech content, but is too broad for me to have any idea]}
 
@@
 Applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter. 
-The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function, which significantly corrupts the signal.
+The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is some combination of sinc functions, which significantly corrupt the signal.
 
 %A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible. 
+\jam{[re:comb filters.
+When I tried to comb filter it either was too sharp and so jumbled the whole signal in convolution.
+Or if smooth attenuated the signal away into noise.
+At least, that's what I remember.
+Changrong has a comb filter section anyway so I am happy dropping my unsuccessful attempts.]}
 
@@
 The green vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$.
-Note the flat response within the band.}
+Note the flat response within the band.
+	\jam{[to do: change green lines to dashed and red. add panel with PSD after applying Butterworth bandpass]}}
 	\label{fig:butterworth}
@@
 as a function of frequency $\omega$. 
+\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
 To construct a Butterworth band-pass filter, the low-pass filter in Eq.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
@@
 The original signal is identical to that shown in Fig.~\ref{fig:noiseSubtract}.
-The first and second panels of Fig.~\ref{fig:timeNotchWiener} show the original audio signal and the recording after passing thought the optical microphone (i.e.
+The first and second panels of Fig.~\ref{fig:timeNotchWiener} show the original audio signal and the recording after passing through the optical microphone (i.e.
 the timeseries of the data shown in the first two panels of Fig.~\ref{fig:noiseSubtract}).
@@
 \han{should we include here a mathematical description of the $15$ cascade notch filter?}
+\jam{[Yes, I have no idea what a cascaded notched filter is and I'm an author!]}
 The timeseries and frequency spectrum results are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}. 
@@
 The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
-It minimizes the mean square error $E(\hat{x}-x)^2$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
-\han{should we add more details here too?}
+It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
+\han{should we add more details here too?} \jam{[seconded]}
 The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
@@
 One second of the one minute recording is shown.
-The rise at the start of the bottom panel is expected when filtering a finite signal.}
+The rise at the start of the bottom panel is expected when filtering a signal of finite duration.\jam{[add more panels showing original time series, like in Changrong's figures]}}
 	\label{fig:logMMSE_timeseries}
@@
 	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} shown before (top panel) and after (bottom panel) the logMMSE estimator is applied.
-The spectrum shows up to $2\,{\rm kHz}$ and to amplitude $2$ to see greater detail.}
+The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.}
 	\label{fig:logMMSE_spectrum}

### 2020-06-30 paper/ifo-conclusions.tex
adding some future work text from Changrongs notes
winter comments part 2

@@
 
-In this paper we present an educational demonstration in which an \han{inexpensive}, table top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
+In this paper we present an educational demonstration in which an \han{inexpensive} \jam{[see comments in introduction]}, table top Michelson interferometer serves as an analogue to a gravitational-wave detector.
 Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
-We consider quasimonochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g.
+We consider quasi-monochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g.
 from supernovae.
@@
 
-In Section~\ref{sec:complex_audio} we demonstrate recovery of more complex signals of music and speech. 
+In Section~\ref{sec:optical_microphone} we demonstrate recovery of more complex signals of music and speech. 
 Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
+\jam{[Confirming Hannah's comment, the Pi was only necessary to read the photodiode, the webcam video I just saved onto PC and analysed directly]}
 We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 

### 2020-06-30 paper/ifo-futureWork.tex
winter comments part 2

@@
 A second improvement is to measure the full model of the transfer function of the system. 
-Theoretically, with the correct transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
-The transfer function starts from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation Appendix~\ref{app:intensity_derivation}, and the photodiode reading.
+With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
+\jam{[check with Rob et al. about the veracity of this]}
+The transfer function starts from the voltage sent to the speaker and ends with the voltage recorded by the Pi.
+It includes any non-linearities in the speaker itself, the speaker-mirror coupling, the separation-intensity relation Appendix~\ref{app:intensity_derivation}, and the photodiode reading.
 A schematic of the full optical microphone transfer function is shown in Figure~\ref{fig:pipeline_highlighted}.
@@
 WT provides both time and frequency information, making it easier to pinpoint the origin of noise respect to time.
-In Refs. \cite{tufekci2000feature,agbinya1996discrete}, WT was also proposed for speech recognition.
+In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT was also proposed for speech recognition.
 
@@
 Flowchart for the optical microphone transfer function. 
-The highlighted stage is analysised in Appendix XX, the other stages are left to future work. 
+The highlighted stage is analysed in Appendix \jam{XX}, the other stages are left to future work. 
+\jam{[to do: change grey to something more readable and reconsider labels. add editable version to git]}
 }

### 2020-06-30 paper/ifo-introduction.tex
winter comments part 2

@@
 % sound analogy
-In this paper, we present an \han{inexpensive} table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves.
+In this paper, we present an \han{inexpensive} \jam{["Inexpensive" is relative.
+The experiment is cheap compared to LIGO, but out of budget for a high-school.
+Suggestions: "modest", "simple", or just cut the adjective.]} table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves.
 As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to electronics, signal processing, and speech enhancement. 
-It gives students in courses such as physics and electrical engineering the freedom to explore the response of an accessible yet nontrivial optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
+It gives students in courses such as physics and electrical engineering the freedom to explore the response of an accessible yet nontrivial opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
 The demonstration can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.

### 2020-06-30 paper/ifo-singleTone.tex
winter comments part 2

@@
 The demonstration is able to recover a single frequency tone to within two standard deviations. }
+\jam{[For the single tone result, the FWHM is 0.033 Hz. This puts the injected 2.09 Hz well within the uncertainty in the measured value of 2.099 Hz.]}
 %We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
@@
 %This showed that the experiment observes a single tone confidently.
+
+\jam{[re: quantifying the non-linear response, the first two harmonics (at 4.19 and 6.28 Hz) are 0.103 and 0.098 the height of the main peak respectively.]}
 
@@
 \han{[to do: add new numbers from James]}
+\jam{[For the single tone result, the FWHM is 0.033 Hz. This puts the injected 2.09 Hz well within the uncertainty in the measured value of 2.099 Hz.]}
 %The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
@@
 \end{figure}
-
-\end{document}
 
@@
 \end{comment}
+
+\end{document}

### 2020-06-30 paper/ifo-tableTopGWs.tex
winter comments part 2

@@
 A commonly-used analogy in explaining gravitational-wave science is sound. 
-Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}~\footnote{When introducing the acoustic analogy to lay audiences, it ins important to emphasize that gravitational waves are not sound.
+Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
 For example, gravitational waves can propagate in a vacuum, whereas sound cannot.}.
@@
 
+\jam{[I don't know what these other designs are (haven't seen the references), is this something that the Adelaide people would have used in their design?]}
+
 \begin{comment}

### 2020-06-30 paper/ifo-wanderingTone.tex
winter comments part 2

@@
 One method to search for a wandering signal is to split the data into several shorter segments which are analysed individually. 
-The detection statistics in every segment are combined to form a grid in time and frequency (spectogram). 
+The detection statistics in every segment are combined to form a grid in time and frequency (spectrogram). 
 The value at each spectrogram point determines the likelihood that a signal is present at a particular time and frequency. 
@@
 The detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}). 
-In this work the detection statistic is the amplitude (modulus) of the Fourier transform of the data at each time segement, which maximizes the likelihood of detecting a sinusiod in Gaussian noise, as described in Appendix\han{to do}.
+In this work the detection statistic is the amplitude (modulus) of the Fourier transform of the data at each time segment, which maximizes the likelihood of detecting a sinusoid in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood} \han{to do} \jam{[makes sense but is beyond me]}.
 The resulting spectrogram gives the Fourier amplitude for each frequency and time bin. 
-The task is then to find the most likely wandering frequency signal through the spectogram. 
+The task is then to find the most likely wandering frequency signal through the spectrogram. 
 The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method to find the most likely path of a signal given an observation sequence. 
@@
 In Section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work. 
-In Section~\ref{sec:wanderingResults} we present our results for recovery of a wandering audio frequency.
+In Section~\ref{sec:wanderingResults} we present the results for recovery of a wandering audio frequency.
 
@@
 Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
-The color of each circle indicates the likelihood that a signal is present in the box, where a lighter color corresponds to higher likelihood. 
+The colour of each circle indicates the likelihood that a signal is present in the box, where a lighter colour corresponds to higher likelihood. 
 AT each time step, the paths can move up one $f$ step, move down one $f$ step, or stay at the same $f$ step (shown by lines in the diagram). 
@@
 %Thinking about this 
-Our starting point is the spectrogram grid indexed by time $t$ and frequency $f$ where there are $N_t$ and $N_f$ time and frequency bins respectively as shown in Fig.~\ref{fig:viterbi}.
+A spectrogram is a grid indexed by time $t$ and frequency $f$ where here there are $N_t$ and $N_f$ time and frequency bins respectively, as shown in Fig.~\ref{fig:viterbi}.
 We label the elements of the grid as $t_i$ and $f_j$ with $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
@@
 
-Our objective is to find the most likely path through the grid given the data and any probabilistic constraints on how the frequency wanders from $t_i$ to $t_{i+1}$. 
+The objective is to find the most likely path through the grid given the data and any probabilistic constraints on how the frequency wanders from $t_i$ to $t_{i+1}$. 
 In continuous wave searches, a physical model is used to decide how much the frequency of the signal is allowed to wander at each time step, which depends on the type of target. 
@@
 The probability of transitioning from $f_j$ at $t_i$ to any of $\{f_{j+1},f_j,f_{j-1}\}$ at $t_{i+1}$ is called the transition probability $A(f_k,f_m)$. 
-In this work the three options have equal probability $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise
+In this work the three options have equal probability $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
 
@@
 The Viterbi algorithm is an efficient method to find the Viterbi path. 
-Here we describe the algorithm as psudeo-code referring to the schematic in Fig.~\ref{fig:viterbi}, where the circles indicate the elements of the spectrogram grid (lighter colours correspond to higher likelihood).
+Here we describe the algorithm as pseudo-code referring to the schematic in Fig.~\ref{fig:viterbi}, where the circles indicate the elements of the spectrogram grid (lighter colours correspond to higher likelihood).
 The implementation used in this work is available online (see Appendix~\ref{app:code}).
@@
 In Fig.~\ref{fig:viterbi}, black lines show the best paths and grey lines show all other paths.
+
 \item Moving to $t_2$, the path with the highest value of the recursive relation in equation~\ref{eqn:recursiveViterbi} is selected for each $f_j$ as shown by the solid lines from $t_1$ to $t_2$ in Fig.~\ref{fig:viterbi}.
+
 \item Steps 2 is repeated until the end of the grid ($t=t_{N_t}$) is reached. 
 Only the best paths are stored.
+
 \item At $t=t_{N_t}$, each $f_j$ has a value $Pr[f(t_{N_t})=f_j]$ (equation~\ref{eqn:recursiveViterbi}). 
 The algorithm selects the terminating frequency $f(t_{N_t})$ with the highest value, marked (b) in Fig.~\ref{fig:viterbi}.
+
 \item The Viterbi path is then found by backtracking along the best paths at each time step.
+\jam{[Andrew mentioned providing a closed form here, which I'm not sure exists.
+The code works by storing three arrays: the weights, the score for the best path to a given vertex, and the index of the previous vertex along that path.
+Backtracking comprises of looping through that third array from the vertex in the last row with the highest score from the second array.
+I don't think we need to explain it any further than that.]}
 In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c). 
-The Viterbi path is highlighed in orange in fig.~\ref{fig:viterbi}.
+The Viterbi path is highlighted in orange in fig.~\ref{fig:viterbi}.
 \end{enumerate}
@@
 The Viterbi path for a time series sequences spanning $t_0$ to $t_{N_t}$ contains the Viterbi path for sub-sequences of that time series. 
-It is also very computationally efficient.
+It is also very computationally efficient. \jam{[Justify this claim to efficiency? Cite something about it being $O(n^2)$ or whatever it is?]}
 
@@
 To test the Viterbi algorithm, a signal with wandering frequency is played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
+\jam{[The injected frequency wandered along a decaying sinusoid, it was not random.]}
 %Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
@@
 Recovery of a wandering tone. 
-The spectrogram shows the observed frequency amplitudue at each time-frequency step. 
+The spectrogram shows the observed frequency amplitude at each time-frequency step. 
 The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path respectively. 
@@
 Quantitatively, the Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time. 
+\jam{[Andrew wanted something about ``percentage root mean-square-error'' along the path, can you explain what he wants me to calculate?]}
+
 %The Viterbi path stays within one frequency bin (approximately $0.3\,{\rm Hz}$) of the injected tone for $94\%$ of the run. 

### 2020-06-30 paper/paper_main.tex
going through a few more comments from Andrew
a few more text changes
winter comments part 1
winter comments part 2

@@
 % GravExplain
-\documentclass[prb,reprint]{revtex4-1} 
+%\documentclass[prb,reprint,nofootinbib]{revtex4-1} 
+\documentclass[pra,superscriptaddress,reprint,amsmath,amssymb,nofootinbib]{revtex4-1}
 % \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
@@
 \usepackage{subfiles}
+\usepackage[bottom]{footmisc}
+\usepackage{comment}
 
-\bibliographystyle{apsrev4-2}
+%\usepackage{dcolumn}
+%\usepackage{bm}
+%\usepackage[pdftex,plainpages=false,colorlinks=true]{hyperref}
+%\usepackage[toc,page,title,titletoc,header]{appendix} 
+%\usepackage{multirow}
+%\usepackage{tabularx}
+%\usepackage[bottom]{footmisc}
+%\usepackage{caption}
+%\usepackage{subcaption}
+%\usepackage[bottom]{footmisc}
+%\usepackage{soul}
+
+%\bibliographystyle{apsrev4-2}
 % \setlength{\parindent}{0pt}
@@
 
-\author{\han{also invite the people who helped us with interferometer design to be authors??}}
+\author{\han{(to do: ask other authors?)} \jam{[Other authors and affiliations I leave to you to contact and clarify]}}
 \email{wmoran@unimelb.edu.au}
@@
 
-\section{Single tone}
+\section{Constant tone}
 \label{sec:single_tone}
@@
 The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
+\jam{[update if turned into authors]} 
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% I feel like we can just delete everything that follows. It is in earlier commits and distributed throughout the subfiles.
+
+\begin{comment}
 
@@
 \end{document}
+
+\end{comment}

### 2020-07-01 paper/ifo-complexAudio.tex
Butterworth plots

@@
 \begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
+	\includegraphics[width=.49\textwidth]{figures/psd_butterworth_14_6.pdf}
 	\caption{\label{fig:psd_noise}
-Power spectral density (PSD) of the optical microphone  with no injected audio (i.e.\ with the speaker off). 
-We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). 
-Otherwise the PSD is fairly white. 
+Power spectral density (PSD) of background noise (top) from the optical microphone (with the speaker off), and the PSD after applying a Butterworth bandpass filter (bottom) between the two frequencies marked with red, dashed lines.
+We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise the PSD is fairly white. 
+After filtering we see strong attenuation of all frequencies outside the band, but little change to the harmonics within the band.
 }
@@
 \begin{figure}
-	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000-cropped.pdf}
+	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000.pdf}
 	\caption{Butterworth bandpass filter frequency response, any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power).
-The green vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$.
-Note the flat response within the band.
-	\jam{[to do: change green lines to dashed and red. add panel with PSD after applying Butterworth bandpass]}}
+The red, dashed vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$.
+Note the flat response within the band.}
 	\label{fig:butterworth}

### 2020-07-01 source/optical_microphone_filters.py

@@ -465,7 +465,7 @@ def psd_plot(filename):
     ax.xaxis.label.set_size(26)
     ax.yaxis.set_tick_params(labelsize=24)
     ax.yaxis.label.set_size(26) 
-    plt.savefig(filename[-4]+'.pdf', bbox_inches='tight')
+    plt.savefig(filename[:-4]+'.pdf', bbox_inches='tight')
     plt.close(fig)
     
 def butter_filter_recording(filename, filetag=None, produce_plots=False,

### 2020-07-01 paper/ifo-conclusions.tex
intro and conclusion comments

@@
 
-In this paper we present an educational demonstration in which an \han{inexpensive} \jam{[see comments in introduction]}, table top Michelson interferometer serves as an analogue to a gravitational-wave detector.
+In this paper we present an educational demonstration in which a table top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
 Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 

### 2020-07-01 paper/ifo-introduction.tex
intro and conclusion comments

@@
 % sound analogy
-In this paper, we present an \han{inexpensive} \jam{["Inexpensive" is relative.
-The experiment is cheap compared to LIGO, but out of budget for a high-school.
-Suggestions: "modest", "simple", or just cut the adjective.]} table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves.
+In this paper, we present a table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves.
 As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to electronics, signal processing, and speech enhancement. 

### 2020-07-01 paper/ifo-tableTopGWs.tex
intro and conclusion comments

@@
 Detection and analysis concepts can be demonstrated with table-top optical equipment. 
+See Refs.~\cite{TTExhibit:2020,TTExhibit:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of table-top Michelson interferometer designs used for science communication.
 
@@
 
-\han{to do: add some text referring to other demo designs and references:~\cite{TTExhibit:2020,TTExhibit:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999}. }
-% Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
-We did not find it necessary to include it to be able to demonstrate the results desired.
-
-\jam{[I don't know what these other designs are (haven't seen the references), is this something that the Adelaide people would have used in their design?]}
-
 \begin{comment}

### 2020-07-02 paper/ifo-singleTone.tex
working new numbers into text

--- 2020-06-16 paper/ifo-singleTone.tex
@@
 Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording.
-\han{[to do - rewrite once James has the new numbers (also Fig 3 caption)] 
-We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ (where the error represents the $1\sigma$ uncertainty).
-The demonstration is able to recover a single frequency tone to within two standard deviations. }
+We measure a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$. 
+Two harmonics can also be seen in Fig.~\ref{fig:webcam_spectra} at integer harmonics of the main peak. 
+The $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ peaks have amplitudes of $0.103$ and $0.098$ respectively as a fraction of the height of the main peak. 
+They are probably due to the nonlinear system response discussed in Section~\ref{} and Appendix~\ref{}.
+
+%\han{[to do - rewrite once James has the new numbers (also Fig 3 caption)] 
+%We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ (where the error represents the $1\sigma$ uncertainty).
+%The demonstration is able to recover a single frequency tone to within two standard deviations. }
+%\jam{[For the single tone result, the FWHM is 0.033 Hz. This puts the injected 2.09 Hz well within the uncertainty in the measured value of 2.099 Hz.]}
 %We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
@@
 %This showed that the experiment observes a single tone confidently.
+
+%\jam{[re: quantifying the non-linear response, the first two harmonics (at 4.19 and 6.28 Hz) are 0.103 and 0.098 the height of the main peak respectively.]}
 
@@
 The frequency of the injection is $2.09\,{\rm Hz}$. 
-The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
-\han{[to do: add new numbers from James]}
+%The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
+The recovered signal peaks at $2.099\,{\rm Hz}$ with a FWHM of $0.033\,{\rm Hz}$. 
+There are also two visible harmonic peaks at integer multiples of the main peak. 
+Their amplitudes as a fraction of the height of the main peak are  $0.103$ and $0.098$ for the $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ peaks respectively. 
+%\han{[to do: add new numbers from James]}
+%\jam{[For the single tone result, the FWHM is 0.033 Hz. This puts the injected 2.09 Hz well within the uncertainty in the measured value of 2.099 Hz.]}
 %The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$.

### 2020-07-02 paper/ifo-complexAudio.tex
adding general raspberry pi ref and edits to complex audio

@@
 
-The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2.
-\jam{[Hannah is right, the Pi first appears here.
-If we want to comment on its convenience and usage this is the place.]}
+The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data.
+\han{[from hannah: James does this sound ok? And I have referenced the Raspberry Pi general website, but could you add any appropriate references you used]}
 Together, these sample the signal at roughly $16\,{\rm kHz}$.
@@
 % breakdown of all the filters used along the way
-
+\han{James, if you could rewrite these sentences based on your comments below to reflect exactly what you did, that would be helpful - I don't want to make guesses based on your comments}
 The optical microphone is tested on recordings of speech from different people and on music ranging from simple melodies to full track songs. 
@@
 \jam{[No idea about the hump at 750 Hz. It is near to the main speech content, but is too broad for me to have any idea]}
+\han{[from hannah: okay, perhaps we should just say that we don't know the origin?]}
 

### 2020-07-02 paper/ifo-complexAudio.tex
added source panels

@@
 The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}).
-For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\epsilon$, the frequency response (attenuation at each frequency) is,
+For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\varepsilon$, the frequency response (attenuation at each frequency) as a function of frequency $\omega$ is:
 \begin{equation}
 \label{eq:butterworth}
-H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\omega/\omega_c)^{2n}}},
+H(\omega) = \left(1+\varepsilon^2 \left( \frac{\omega}{\omega_c} \right)^{2n}\right)^{-\frac{1}{2}}
 \end{equation}
-as a function of frequency $\omega$. 
+
 \jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
@@
 \begin{figure*}
-	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
-	\caption{Time series of the optical microphone recording of an adult male voice (saying ``A cathode ...'') shown before (top panel) and after (bottom panel) the logMMSE estimator is applied.
+	\includegraphics[width=\textwidth]{figures/combined_timeseries_melatos.pdf}
+	\caption{Time series of an adult male voice (saying ``A cathode ...'') showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
 One second of the one minute recording is shown.
-The rise at the start of the bottom panel is expected when filtering a signal of finite duration.\jam{[add more panels showing original time series, like in Changrong's figures]}}
+The rise at the start of the bottom panel is expected when filtering a signal of finite duration.}
 	\label{fig:logMMSE_timeseries}
@@
 \begin{figure*}
-	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
-	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} shown before (top panel) and after (bottom panel) the logMMSE estimator is applied.
+	\includegraphics[width=\textwidth]{figures/combined_spectrum_melatos.pdf}
+	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
 The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.}

### 2020-07-02 paper/ifo-futureWork.tex
Butterworth plots
changes to transfer function flowchart

@@
 \begin{figure*}
-	\includegraphics[width=.99\textwidth]{figures/pipeline_highlighted.pdf}
+	\includegraphics[height=.99\textwidth, angle=-90]{figures/pipeline.pdf}
 	\caption{
 Flowchart for the optical microphone transfer function. 
-The highlighted stage is analysed in Appendix \jam{XX}, the other stages are left to future work. 
-\jam{[to do: change grey to something more readable and reconsider labels. add editable version to git]}
+The stage inside the red, dashed box is analysed in Appendix~\ref{app:intensity_derivation}. Understanding the other stages is one part of future work. 
 }

### 2020-07-03 paper/ifo-complexAudio.tex
section VB first para

@@
 The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data.
-\han{[from hannah: James does this sound ok? And I have referenced the Raspberry Pi general website, but could you add any appropriate references you used]}
+\han{[from hannah: James does this sound ok?
+And I have referenced the Raspberry Pi general website, but could you add any appropriate references you used]} \jam{[from James: no problem, sounds good.
+no obvious references spring to mind that aren't readily available for anyone trying to reproduce the experiment]}
 Together, these sample the signal at roughly $16\,{\rm kHz}$.
@@
 % breakdown of all the filters used along the way
-\han{James, if you could rewrite these sentences based on your comments below to reflect exactly what you did, that would be helpful - I don't want to make guesses based on your comments}
-The optical microphone is tested on recordings of speech from different people and on music ranging from simple melodies to full track songs. 
-Each recording is around one minute long with \jam{processing [this is poor, each recording was functionally ten seconds long, the fact that the sources were minutes long doesn't really matter]} applied to ten second segments.
-For each observation\jam{[again, poor wording, this means that for each 10 s recording we did the following]}, the audio is played through the speaker whilst the Pi script records the signal from the photodiode.
-During recordings, care is taken to ensure minimal activity around the interferometer to reduce environmental noise. 
-The timeseries is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.
+The optical microphone was tested on recordings of varying speech from different people and on music ranging from simple melodies and rhythms to full songs. 
+The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources).
+When processing the results, we often only looked at the first ten seconds of each observation for efficiency.
+During recordings, care was taken to ensure minimal activity around the interferometer to reduce environmental noise. 
+The timeseries was then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}, as before.
 
-We find that initial recordings \jam{[Hannah's interpretation is correct here: "We find that the raw output of the optical microphone, collected without special precautions except anti-aliasing, is noisy, with a loud..."?]} are extremely noisy with a loud bass hum throughout.
+We find that the raw output of the optical microphone, collected without any further processing (except anti-aliasing), is noisy, with a loud bass hum throughout.
 Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic.
@@
 The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination. 
-\jam{[No idea about the hump at 750 Hz. It is near to the main speech content, but is too broad for me to have any idea]}
-\han{[from hannah: okay, perhaps we should just say that we don't know the origin?]}
+Outside the mains noise, the power appears to peak broadly around 750 Hz, the origin or significance of which we don't know.
 
@@
 %A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible. 
-\jam{[re:comb filters.
+% \jam{[re:comb filters.
 When I tried to comb filter it either was too sharp and so jumbled the whole signal in convolution.

### 2020-07-06 paper/ifo-complexAudio.tex
adding material from Changrong to complex audio

@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/freqSpectrumOriginalNoiseSubtracted.jpg}
+% trim order is <left> <lower> <right> <upper>
+\includegraphics[width=\textwidth,trim={0cm 8.5cm 0 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/timeSeriesNotchWiener.jpg}
+\includegraphics[width=\textwidth,trim={0cm 8.7cm 0 8.3cm},clip]{figures/timeSeriesNotchWiener.pdf}
 \end{center}
 \caption{\label{fig:timeNotchWiener}
+Timeseries results for the notch and Wiener filters. 
+The first panel shows the original speech recording which was played from the speaker. 
+The second panel shows the data from the optical microphone before any filters are applied. 
+The third and fourth panels show results after applying the notch and Wiener filter respectively. 
+Finally, the fifth panel shows the result of applying both the notch and Wiener filter. 
 }
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/freqNotchWiener.jpg}
+\includegraphics[width=\textwidth,trim={0cm 8.2cm 0 8cm},clip]{figures/freqNotchWiener.pdf}
 \end{center}
 \caption{\label{fig:freqNotchWiener}
+Frequency spectrum results for the notch and Wiener filters. 
+The first and second panels show the results of applying the notch and Wiener filter respectively.
+The third panel shows the spectrum after applying both the notch and Wiener filters. 
+The first, second, and third panels show the spectrum of the timeseries show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
 }
@@
 
-A notch filter attenuates signals in specific frequency bands. 
-We apply a $15$ cascaded notched filter to remove the mains noise and its harmonics up to $750\,{\rm Hz}$. 
-\han{should we include here a mathematical description of the $15$ cascade notch filter?}
-\jam{[Yes, I have no idea what a cascaded notched filter is and I'm an author!]}
-The timeseries and frequency spectrum results are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}. 
+\subsubsection{Cascade notched filter}
+A straightforward method for removing hum noise and its harmonics is to use a sequence of notch filters centered at each of the frequency bins we want to remove. 
+A typical transfer function of infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
+\begin{equation}
+    \label{eqn:notch}
+    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}
+\end{equation}
+where $\alpha$ and $\beta$ are self-tuned parameters which adjust the locations of poles and zeros. 
+By allowing the numerator and demoninator of $H(z)$ to be zero, we can find the corresponding locations of zeros and poles of the filter. 
+Notice zeros are at $e^{\pm jw_0}, \text{where } w_0=\cos^{-1}(\beta)$, and poles are at $re^{\pm j\phi}, \text{where } r=\sqrt{\alpha},\phi=\cos^{-1}(\beta(1+\alpha)/(2\sqrt{\alpha}))$.
+At $w_0$, the magnitude response will be zero. 
+Thus, we can calculate $\beta$ setting $w_0$ to be the frequency we want to remove. 
+The parameter $\alpha$ affects the bandwidth ($B_w$) of the filter and has the relation $B_w=\cos^{-1}(\frac{2\alpha}{1+\alpha^2})$. \\
+In our paper, we first decide $\beta$ for each filter. 
+We locate hum noise spectrum at around $50k\,{\rm Hz}$ where the integer $k=1,2,\dots,15$. 
+For each $k$, we build a notch filter to filter out the specific frequency bin approximately at $50k \, {\rm Hz}$. 
+We then alter the parameter $\alpha$ to determine the bandwidth of each filter. 
+On one side, the bandwidth should be as narrow as possible to avoid disturbing useful signals. 
+This can be achieved through increasing orders of the filter (add more poles and zeros). 
+On the other side, the bandwidth should have enough residues to account for the uncertainty of our pinpointing the location of hum noise. 
+After considering these two aspects, we are now able to get the ideal transfer function for each filter. 
+The whole cascaded filter is achieved by multiplying them together, 
+\begin{equation}
+    \label{eqn:notch15}
+    H(z)=H_1(z)H_2(z)\dots H_k(z),k=1,2,\dots 15\,,
+\end{equation}
+where $H(z)$ denotes the transfer function of the full cascaded notch filter, and $H_k(z)$ denotes one single notch filter for removing $50k \, {\rm Hz}$ frequency. 
+We use the built-in MATLAB filter design toolbox to help us realize a well performed cascaded notch filter (each filter has order 6).  
+The magnitude response of the first five notch filters are shown in Fig.~\ref{fig:notchMagResponse}. 
+\begin{figure}
+\begin{center}
+\includegraphics[width=.49\textwidth]{figures/1111}
+\end{center}
+\caption{\label{fig:notchMagResponse}
+Magnitude response for the first five notches of the $15$ cascaded notch filter described by Eqns.~\ref{eqn:notch} and ~\ref{eqn:notch15}. 
+}
+\end{figure}
+
+The timeseries and frequency spectrum results of the notch filter are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}. 
 We seen that the mains harmonics are significantly attenuated in comparison to the analysis in Fig.~\ref{fig:noiseSubtract}.
 
-The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
-It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
-\han{should we add more details here too?} \jam{[seconded]}
-The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
-The timeseries is visibly cleaner, however there is a strong noise hum.
+%A notch filter attenuates signals in specific frequency bands. 
+%We apply a $15$ cascaded notched filter to remove the mains noise and its harmonics up to $750\,{\rm Hz}$. 
+%\han{should we include here a mathematical description of the $15$ cascade notch filter?}
+%\jam{[Yes, I have no idea what a cascaded notched filter is and I'm an author!]}
 
-The combination of the notch filter and the Wiener filter produces a cleaner signal than either individually. 
-The time series and spectrum results are shown in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively.
-\han{todo - ask Changrong for some concluding remarks on intelligibility of the result using notch and wiener filter}
+\subsubsection{Weiner Filter}
+
+We use vector $\textbf{x}=\{x(0),\dots x(N-1)\}$ to denote the noisy speech signal sequence.
+Assume original clean signal is $\textbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\textbf{w}=\{w(0),\dots,w(N-1)\}$.
+We then have
+\begin{equation}
+    \textbf{x}=\textbf{s}+\textbf{w}
+\end{equation}
+Based on the observation data$\textbf{x}$, our goal is to estimate s, such that the error between $\textbf{x}$ and estimated $\mathit{\hat{s}}$ is minimized in the mean square sense, we write
+\begin{equation}
+    \hat{\textbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\textbf{x}-\textbf{s}||_2^2
+\end{equation}
+Under the assumptions that observed data is wide sense stationary and noise sequence is stationary, uncorrelated with signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$.
+The famous Wiener-Hopf equation \citep{noble1959methods} provides us a way to determine $\textbf{h}$, which can be expressed as 
+\begin{equation}
+\label{winer-hopf}
+\begin{bmatrix}  
+r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
+r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
+\vdots&\vdots&\ddots&\vdots\\
+r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
+\end{bmatrix}
+\begin{bmatrix}
+h[0]\\
+h[1]\\
+\vdots\\
+h[n]
+\end{bmatrix}=
+\begin{bmatrix}
+r_{ss}[0]\\
+r_{ss}[1]\\
+\vdots\\
+r_{ss}[n]
+\end{bmatrix}
+\end{equation}
+
+\begin{equation*}
+\left\{
+\begin{array}
+r_{xx}[n]=E[x[i]x[i+n]] \\ 
+r_{ss}[n]=E[s[i]s[i+n]]
+\end{array} \right\}  
+\end{equation*}
+where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
+If further, we do not need to guarantee the causality of Wiener filter, in other words, we estimate current signal based on both past and future observations, Eq.
+\ref{winer-hopf} can then be represented in the frequency domain, which is
+\begin{equation}
+\label{wiener}
+    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}
+\end{equation}
+where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of noisy observations, clean signal and pure noise respectively.
+Intuitively, from Eq.
+\ref{wiener}, we can see Wiener filter amplifies input signal where signal to noise ratio (SNR) is high, and weaken the signal where SNR is low.
+Notice that since this filter is non-causal, which may not be suitable for real time signal processing.
+More detailed analysis can be found in \citep{10.5555/151045}.
+In our case, we construct a causal Wiener filter based on Eq.
+\ref{winer-hopf}, and set the length of the filter to be 100.
+Large n smooths the input signal further, while cause more delay and requires more memory meanwhile.
+$n=100$ is a relatively reasonable choice for balancing these effects.
+
+%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
+%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
+%\han{should we add more details here too?} \jam{[seconded]}
+%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
+%The timeseries is visibly cleaner, however there is a strong noise hum.
+
+\subsubsection{Combined Notch and Weiner filter}
+The Wiener filter makes use of statistical information from the speech data and noise. 
+It amplifies the part of the signal with high SNR, while suppressing the part with low SNR. 
+It is implemented in the form of a finite response filter, which ensures linear phase response and stability at the cost of high orders. 
+By comparison, the notch filter is based on directly removing the unwanted frequency components. 
+It is implemented in the form of an infinite impulse response fileter. 
+Although it significantly decreases the order of the filter, it unavoidably introduces nonlinear phase and instability.
+
+Combining the notch and Wiener filter, we have a trade off between the two filters, which shows better performance in filtering out different noise components (white noise plus hum noise). 
+The results from the combination of notch and Wiener filters are visibly cleaner as shown in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} for the timeseries and spectrum respectively
+
+%The combination of the notch filter and the Wiener filter produces a cleaner signal than either individually. 
+%The time series and spectrum results are shown in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively.
 

### 2020-07-06 paper/ifo-appendix.tex
editing wiener filter text from Changrong

@@
 My thought is only that all of this is technically available in the open source.]}
+\han{[from Hannah: I think it is worth having a go at pseudocode write-up of the implementation as Andrew suggests - James can you add some text based on your code and we can see how it looks?]}
 %We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
@@
 A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
-This design was based off \jam{standard examples (I have no idea where I or Alex originally got the designs from, I'm confident though that they are common enough to not need citing given that we also give the diagram)} of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
+This design was based off \jam{standard examples (I have no idea where I or Alex originally got the designs from, I'm confident though that they are common enough to not need citing given that we also give the diagram)} \han{[from hannah: hmm, okay - I think if we refer to `standard examples' we need to say what that means.
+Are there any equivalent references you can find which include the same information, even if you can't find exactly the one you used?
+If we think about someone who may like to reproduce this work, this is a very useful piece of information]} of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
 

### 2020-07-06 paper/ifo-complexAudio.tex
editing wiener filter text from Changrong

@@
 
-We use vector $\textbf{x}=\{x(0),\dots x(N-1)\}$ to denote the noisy speech signal sequence.
-Assume original clean signal is $\textbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\textbf{w}=\{w(0),\dots,w(N-1)\}$.
-We then have
+A Wiener filter is used to compute an estimate of the original signal. 
+We label the noisy speech signal sequence as  $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence. 
+It is the sum of the original clean signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$, 
 \begin{equation}
-    \textbf{x}=\textbf{s}+\textbf{w}
+    \mathbf{x}=\mathbf{s}+\mathbf{w}
 \end{equation}
-Based on the observation data$\textbf{x}$, our goal is to estimate s, such that the error between $\textbf{x}$ and estimated $\mathit{\hat{s}}$ is minimized in the mean square sense, we write
+Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
 \begin{equation}
-    \hat{\textbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\textbf{x}-\textbf{s}||_2^2
+    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
 \end{equation}
-Under the assumptions that observed data is wide sense stationary and noise sequence is stationary, uncorrelated with signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$.
-The famous Wiener-Hopf equation \citep{noble1959methods} provides us a way to determine $\textbf{h}$, which can be expressed as 
+Assuming that $\textbf{x}$ is wide-sense stationary \han{[perhaps a simple definition of wide-sense stationary should be included in brackets here?
+to do; ask Changrong]} and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$.
+The famous Wiener-Hopf equation~\citep{noble1959methods} provides us a way to determine $\textbf{h}$, as
 \begin{equation}
-\label{winer-hopf}
+\label{eqn:winer-hopf}
 \begin{bmatrix}  
@@
 \end{equation}
-
-\begin{equation*}
-\left\{
-\begin{array}
-r_{xx}[n]=E[x[i]x[i+n]] \\ 
-r_{ss}[n]=E[s[i]s[i+n]]
-\end{array} \right\}  
-\end{equation*}
-where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
-If further, we do not need to guarantee the causality of Wiener filter, in other words, we estimate current signal based on both past and future observations, Eq.
-\ref{winer-hopf} can then be represented in the frequency domain, which is
+where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$, 
+%\begin{equation*}
+%\left\{ 
+\begin{eqnarray} 
+r_{xx}[n] &~=~& E[x(i)~x(i+n)] \,,\\
+r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
+%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
+%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
+\end{eqnarray} 
+%\right\}  
+%\end{equation*}
+%where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
+If we do not need to guarantee the causality of the Wiener filter, i.e.
+we can estimate the current signal based on both past and future observations, then Eqn.~\ref{eqn:winer-hopf} can be represented in the frequency domain as
 \begin{equation}
-\label{wiener}
+\label{eqn:wiener}
     H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}
 \end{equation}
-where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of noisy observations, clean signal and pure noise respectively.
-Intuitively, from Eq.
-\ref{wiener}, we can see Wiener filter amplifies input signal where signal to noise ratio (SNR) is high, and weaken the signal where SNR is low.
-Notice that since this filter is non-causal, which may not be suitable for real time signal processing.
-More detailed analysis can be found in \citep{10.5555/151045}.
-In our case, we construct a causal Wiener filter based on Eq.
-\ref{winer-hopf}, and set the length of the filter to be 100.
-Large n smooths the input signal further, while cause more delay and requires more memory meanwhile.
-$n=100$ is a relatively reasonable choice for balancing these effects.
+where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 
+Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the Wiener filter amplifies the input signal where signal to noise ratio (SNR) is high, and weakens the signal where SNR is low.
+We note that this filter is non-causal, which may not be suitable for real time signal processing. 
+More detailed analysis can be found in Ref.~\citep{10.5555/151045}. 
+In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:winer-hopf}, and set the length of the filter to $n=100$. 
+A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
+The choice of $n=100$ is a reasonably selection to balance smoothing and efficiency.
 
+The timeseries and frequency spectrum results of the Wiener fileter are are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
+There is improvement which is particularly visible in the timeseries, however a strong noise hum persists. 
 %The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 

### 2020-07-07 paper/ifo-appendix.tex
adding app c, general text editing and tidying

@@
 
-\section{Likelihood of sinusoids in Gaussian noise}
+\section{Detecting a sinusoidal signal in Gaussian noise}
 \label{app:sinusoid_likelihood}
-\jam{[to do!
+%\jam{[to do!
 from Andrew: ``Write a short appendix proving that maximizing the likelihood exp(-n||n/2) with respect to A given a signal A cos (wt + phi) is the Fourier transform.
 Classic sig proc result useful for teachers to know.'']}
+In this appendix we demonstrate the the modulus of the Fourier Transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
+We describe the data as
+\begin{equation}
+x(t) = s(t) + n(t)\,, 
+\end{equation}
+where $s(t)$ and $n(t)$ are the signal and noise respectively. 
+The we describe the signal as 
+\begin{equation}
+s(t) = A \cos{\omega t + \phi}\,,
+\end{equation}
+where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal respectively. 
+The log likelihood assuming Gaussian noise is given by
+\begin{eqnarray}
+\log \mathcal{L} &=& -\frac{1}{2}(n|n)\,, \\
+                 &=& -\frac{1}{2}(x-s|x-s)\,, \\ 
+                 &=& -\frac{1}{2}(x|x) - \frac{A^2}{2} (\cos[\omega t + \phi] | \cos[\omega t + \phi]) \\ &&+ A (x | \cos[\omega t +\phi])\,, \\ 
+                 &=& -\frac{1}{2}(x|x) - \frac{A^2}{4} + A (x | \cos[\omega t +\phi])\,, \label{eqn:logLike}\\ 
+\end{eqnarray}
+where we define the inner product between $u(t)$ and $v(t)$ as $(u|v) = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data (and the last line assumes that $T>>1/\omega$ so that $(\cos[\omega t + \phi] | \cos[\omega t + \phi])\approx{1/2}$).
+
+Maximising Eqn.~\ref{eqn:logLike} with respect to $A$ shows that the maximum likelihood is when $A = 2\left( x | \cos[\omega t+\phi]\right)$. 
+The maximum likelihood is then, 
+\begin{eqnarray}
+\log \mathcal{L}_\mathrm{max} = - \frac{1}{2} (x|x) + ( x | cos[\omega t + \phi)^2\,. \label{eqn:maxLogLike}
+\end{eqnarray}
+Focusing on the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, baring the constant factor $(x|x)$.
 

### 2020-07-07 paper/ifo-complexAudio.tex
adding app c, general text editing and tidying

@@
 Instead, in this section, we explore how to use the interferometer as an `optical microphone' to detect and play back complex audio signals, using light to capture sound. 
-In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood} \han{to do}.
+In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
 The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
@@
 And I have referenced the Raspberry Pi general website, but could you add any appropriate references you used]} \jam{[from James: no problem, sounds good.
-no obvious references spring to mind that aren't readily available for anyone trying to reproduce the experiment]}
+no obvious references spring to mind that aren't readily available for anyone trying to reproduce the experiment]} \han{[could you add a couple of useful resources?
+We should try our best to help out the reader in terms of reproducibility]}
 Together, these sample the signal at roughly $16\,{\rm kHz}$.
@@
 \end{equation}
-Assuming that $\textbf{x}$ is wide-sense stationary \han{[perhaps a simple definition of wide-sense stationary should be included in brackets here?
-to do; ask Changrong]} and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$.
+Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
+A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
+$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
+$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
+only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.}
+and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
 The famous Wiener-Hopf equation~\citep{noble1959methods} provides us a way to determine $\textbf{h}$, as

### 2020-07-07 paper/ifo-singleTone.tex
changes to transfer function flowchart
adding app c, general text editing and tidying

@@
 We measure a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$. 
-Two harmonics can also be seen in Fig.~\ref{fig:webcam_spectra} at integer harmonics of the main peak. 
+Two harmonics can also be seen in Fig.~\ref{fig:webcam_spectrum} at integer harmonics of the main peak. 
 The $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ peaks have amplitudes of $0.103$ and $0.098$ respectively as a fraction of the height of the main peak. 
-They are probably due to the nonlinear system response discussed in Section~\ref{} and Appendix~\ref{}.
+They are probably due to the nonlinear system response discussed in Section~\ref{sec:ifo} and Appendix~\ref{app:intensity_derivation}.
 

### 2020-07-07 paper/ifo-wanderingTone.tex
adding app c, general text editing and tidying

@@
 The detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}). 
-In this work the detection statistic is the amplitude (modulus) of the Fourier transform of the data at each time segment, which maximizes the likelihood of detecting a sinusoid in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood} \han{to do} \jam{[makes sense but is beyond me]}.
+In this work the detection statistic is the amplitude (modulus) of the Fourier transform of the data at each time segment, which maximizes the likelihood of detecting a sinusoid in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}.
 The resulting spectrogram gives the Fourier amplitude for each frequency and time bin. 

### 2020-07-09 paper/ifo-complexAudio.tex
adding in some comments from Changrong on complex audio section

@@
 
-\subsubsection{Weiner Filter}
+\subsubsection{Wiener Filter}
 
@@
 \begin{equation}
-    \mathbf{x}=\mathbf{s}+\mathbf{w}
+    \mathbf{x}=\mathbf{s}+\mathbf{w}\,.
 \end{equation}
-Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
+Given the observation data $\textbf{x}$, our goal is to filter the signal $\textbf{s}$ from the noise $\textbf{w}$, and get an estimated $\hat{\textbf{s}}$, where the difference between $\hat{\textbf{s}}$ and true signal $\textbf{s}$ is minimized in the mean square error sense, which is represented as
 \begin{equation}
-    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
+\text{BMSE}(\hat{\textbf{s}})=E[(\textbf{s}-\hat{\textbf{s}})^2]\,,
 \end{equation}
-Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
+where $\text{BMSE}$ is the Bayesian mean square error. 
+Assuming that (i) $\textbf{x}$ is wide sense stationary~\footnote{
 A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
@@
 only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} 
-and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
+with zero mean; (ii) the signal $\textbf{s}$ is zero mean; (iii) the noise processes $\textbf{w}$ are uncorrelated with the signal $\textbf{s}$, we can further express the estimated $\hat{\textbf{s}}$ to be a linear combination of present and past input observed data
+\begin{equation}
+\hat{{s}}[n]=\sum_{k=0}^{n}h[k]x[n-k]\,,
+\end{equation}
+where $\textbf{h}=\{h(0),\dots h(n)\}$ represents the coefficients of an $n$th order Wiener filter.
+%Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
+%\begin{equation}
+%    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
+%\end{equation}
+%Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
+%A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
+%$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
+%$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
+%only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.}
+%and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
 The famous Wiener-Hopf equation~\citep{noble1959methods} provides us a way to determine $\textbf{h}$, as
@@
 r_{ss}[n]
-\end{bmatrix}
+\end{bmatrix}\,,
 \end{equation}
@@
 
-\subsubsection{Combined Notch and Weiner filter}
+\subsubsection{Combined Notch and Wiener filter}
 The Wiener filter makes use of statistical information from the speech data and noise. 

### 2020-07-10 paper/ifo-appendix.tex
tasks 1, 3, and 5

@@
 A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
-This design was based off \jam{standard examples (I have no idea where I or Alex originally got the designs from, I'm confident though that they are common enough to not need citing given that we also give the diagram)} \han{[from hannah: hmm, okay - I think if we refer to `standard examples' we need to say what that means.
-Are there any equivalent references you can find which include the same information, even if you can't find exactly the one you used?
-If we think about someone who may like to reproduce this work, this is a very useful piece of information]} of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
+The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an ADC to a Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
+Useful throughout was the Pi documentation~\cite{pidocumentation:online} and the pinoutXYZ reference~\cite{pipinout:online}.
 

### 2020-07-10 paper/ifo-complexAudio.tex
tasks 1, 3, and 5

@@
 The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data.
-\han{[from hannah: James does this sound ok?
-And I have referenced the Raspberry Pi general website, but could you add any appropriate references you used]} \jam{[from James: no problem, sounds good.
-no obvious references spring to mind that aren't readily available for anyone trying to reproduce the experiment]} \han{[could you add a couple of useful resources?
-We should try our best to help out the reader in terms of reproducibility]}
-Together, these sample the signal at roughly $16\,{\rm kHz}$.
+Together, these sample the signal at roughly $16\,{\rm kHz}$. Resources for using the Pi are found in Appendix~\ref{app:circuit_diagram}.
 
@@
 \jam{[non-linear appendix, Andrew says to cite Feynman lectures on non-linear oscillations]}
-We find that the mains power signal is present in data recorded without an injected audio signal as seen in the power spectral density in Figure~\ref{fig:psd_noise}.
-The mains power is also present in data taken with the photodiode in complete darkness with a similar but weaker distribution as that shown in Figure~\ref{fig:psd_noise}).
+This mains power signal is present in the power spectral density of background noise as shown in the top panel of Figure~\ref{fig:psd_noise}.
+The mains power signal is far weaker but still present in the background noise taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution.
 The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination. 
@@
 
-Band-passing the signal similarly does not address the remaining harmonics above $100\,{\rm Hz}$. 
-However, it does suppress other noise at higher frequencies. 
-The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}).
 For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\varepsilon$, the frequency response (attenuation at each frequency) as a function of frequency $\omega$ is:
@@
 To construct a Butterworth band-pass filter, the low-pass filter in Eq.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
-The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
-The Butterworth filter performs the well as it is optimised to be `maximally flat' within the band region (prioritising condition (ii) above).
-This filter improves on the above simpler filters but recordings of speech still have low intelligibility.
-They can be recognised as speech without being understood.
-Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique.
+The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}).
+The response of the filter used is shown in Figure~\ref{fig:butterworth} and its effect on the power spectral density of background noise is shown in the bottom panel of Figure~\ref{fig:psd_noise}.
+Band-passing the signal does not address the remaining harmonics above $100\,{\rm Hz}$. 
+However, it does suppress other noise at higher frequencies. 
+The Butterworth filter performs well at maintaining source intelligibility as it is optimised to be `maximally flat' within the band region (prioritising condition (ii) above).
+However recordings of speech still have low intelligibility (the speech can't be understood) as the filter does not improve beyond the source.
+% They can be recognised as speech without being understood.
+Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique to address this.
 

### 2020-07-10 paper/ifo-conclusions.tex
tasks 1, 3, and 5

@@
 In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of signals with constant and slowly wandering frequency. 
-A webcam sampling at $30\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered succesfully with the Viterbi algorithm. 
-These sections direcly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
+A webcam sampling at $30\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered successfully with the Viterbi algorithm. 
+These sections directly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
 %Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer.

### 2020-07-10 paper/ifo-wanderingTone.tex
tasks 1, 3, and 5

@@
 
-To test the Viterbi algorithm, a signal with wandering frequency is played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
-\jam{[The injected frequency wandered along a decaying sinusoid, it was not random.]}
+To test the Viterbi algorithm, a sinusoidal signal with wandering frequency was played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
+The injected frequency itself followed a sinusoid with decaying amplitude.
+This choice was arbitrary, any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
 %Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.

### 2020-07-10 paper/ifo-appendix.tex
task 4

@@
 
-This appendix contains some details regarding implementation of the Viterbi algorithm described in Section~\ref{sec:viterbi}. 
+This appendix contains some details regarding the implementation of the Viterbi algorithm described in Section~\ref{sec:viterbi}. 
 The Viterbi algorithm is a classic method in signal processing, whose theoretical underpinnings and implementation are easily accessible to undergraduate students. 
@@
 
-\han{to do! pseudo-code here}
-\jam{[Hannah has already explained a lot about Viterbi, at this point is it worth just citing existing documentation for further details?
-I used Wikipedia in writing code, should we cite that?
-Andrew mentions going through the practical aspects, e.g.
-log to avoid underflow, is that worth writing up?
-My thought is only that all of this is technically available in the open source.]}
-\han{[from Hannah: I think it is worth having a go at pseudocode write-up of the implementation as Andrew suggests - James can you add some text based on your code and we can see how it looks?]}
+\subsection{Pseudocode}
+Here we give pseudocode of the implementation used in Section~\ref{sec:viterbi}.
+More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}.
+We use Fourier amplitudes normalised between $(0, 1)$ as multiplicative weights, but to avoid underflow we take the logarithm and consider them as additive weights.
+
+Let $X$ be grid of $j=0,\ldots,N_f$ rows and $i=0,\ldots,N_t$ columns of additive weights for each node.
+Let Y and Z be grids of the same shape to store the weight of the best path to each node and the row index of the previous node on that path, respectively.
+Let W be a length $N_t+1$ array to store the final sequence of row indices.
+We restrict paths to only move one cell up or down at a time (or stay constant).
+
+\begin{algorithmic}
+\Function{Viterbi}{$X$}
+    % \State $a \gets a+1$
+    % \State \Return $a$
+
+    \For{each row $j=0,\ldots,N_f$}
+		\State $Y_{0,j} \gets X_{0,j}$
+    \EndFor
+
+    \For{each column $i=1,\ldots,N_t$}
+	    \For{each row $j=0,\ldots,N_f$}
+
+	    	\State $Y_{i,j} \gets \underset{k \in \{-1,0,1\}}{\max} (Y_{i-1,j+k} + X_{i,j})$
+	    	\State $Z_{i,j} \gets j + \underset{k \in \{-1,0,1\}}{\arg\max} (Y_{i-1,j+k} + X_{i,j})$
+
+	    \EndFor
+    \EndFor
+
+    \State $W_{N_t} \gets \underset{j=0,\ldots,N_f}{\arg\max} (Y_{N_t,j})$
+
+    \For{each col $i=N_t-1,\ldots,0$}
+
+		\State $W_i \gets Z_{i+1, W_{i+1}}$
+
+    \EndFor
+
+    \State \Return $W$
+\EndFunction
+\end{algorithmic}
+
 %We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.

### 2020-07-10 paper/ifo-wanderingTone.tex
task 4

@@
 Quantitatively, the Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time. 
-\jam{[Andrew wanted something about ``percentage root mean-square-error'' along the path, can you explain what he wants me to calculate?]}
+We also consider the root mean square (RMS) fractional error along the path.
+Given $i=0,1,...,N_f$ columns of the injected frequency $I_i$ and recovered (Viterbi) frequency $R_i$, the RMS fractional error is given by:
+\begin{equation}
+\sqrt{ \frac{1}{N_f+1} \sum_{i=0}^{N_f} \frac{(I_i - R_i)^2}{{I_i}^2} }
+\end{equation}
+Computing this along the path, we end up with the value \jam{???}, which indicates a good recovery of the injected signal.
 

### 2020-08-16 paper/ifo-complexAudio.tex
some general text editing and addressing comments

@@
 \begin{figure*}
-	\includegraphics[width=.6\textwidth]{figures/setup_pic2.pdf}
+	\includegraphics[width=.7\textwidth]{figures/setup_pic2.pdf}
 \caption{Photograph of the optical microphone.
@@
 % trim order is <left> <lower> <right> <upper>
-\includegraphics[width=\textwidth,trim={0cm 8.5cm 0 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
+\includegraphics[width=\textwidth,trim={1cm 8.5cm 1cm 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth,trim={0cm 8.7cm 0 8.3cm},clip]{figures/timeSeriesNotchWiener.pdf}
+\includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/timeSeriesNotchWiener.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth,trim={0cm 8.2cm 0 8cm},clip]{figures/freqNotchWiener.pdf}
+\includegraphics[width=\textwidth,trim={0cm 8.2cm 0cm 8cm},clip]{figures/freqNotchWiener.pdf}
 \end{center}

### 2020-08-16 paper/ifo-introduction.tex
some general text editing and addressing comments

@@
 In 2015, gravitational waves were observed for the first time from the merger of two black holes~\cite{GW150914}. 
-The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
-Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
+The observation, made by the Advanced Laser Interferometer Gravitational-wave Observatory~(Advanced LIGO \citep{AdvancedLIGO:2015}), marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
+Since 2015, the Advanced LIGO and Advanced Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
 Gravitational-wave detection has led to increased public and media interest in gravitational-wave science. 
@@
 The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects.
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers.
+Observatories such as LIGO, Virgo, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} are laser interferometers.
 They use the interference of laser light to measure changes in distance. 
@@
 Gravitational-wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
+Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
 
 % continuous gws
-To date the network of gravitational wave observatories has observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
+To date the network of gravitational-wave observatories has observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals.
@@
 LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion, where the rotation frequency of the compact object wanders over time due to variable accretion of matter (and hence angular momentum transfer) from the companion~\cite{xraybinaries:1997}.
-Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
+Scorpius X-1 is an LMXB which is particularly bright in X-rays making it a prime target for continuous wave searches from LMXBs. 
+Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017,ScoX1ViterbiO1:2017,SearchRadiometerO1:2017,SearchCrossCorrO1:2017,MiddletonEtAlO2LMXBs:2020,SearchTwoSpecS6:2017}.
 
@@
 
-The paper presents 
 In this work we present a selection of demonstrations using different types of audio signals ranging from simple tones, analogous to continuous wave signals, to more complex audio of music and speech.

### 2020-08-16 paper/ifo-tableTopGWs.tex
some general text editing and addressing comments

@@
 Detection and analysis concepts can be demonstrated with table-top optical equipment. 
-See Refs.~\cite{TTExhibit:2020,TTExhibit:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of table-top Michelson interferometer designs used for science communication.
+See Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of table-top Michelson interferometer designs used for science communication.
 
@@
 
-The interference pattern is a ring of interference fringes as shown by the photograph in Fig.~\ref{fig:interference_pattern}. 
-The pattern is enlarged by a diverging lens between the beamsplitter and the screen to aid in viewing. 
-A change in the relative length of the interferometer arms results in the interference rings moving radially inwards or outwards. 
-Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
-%This results in a phase difference between the two beams when they meet at the screen.
-This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
-For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-When the path difference changes, the fringes stay circular but move radially in or out.
-%A ring interference pattern is often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily seen.
-
 \begin{figure}
@@
 \end{figure}
+
+The interference pattern is a ring of interference fringes as shown by the photograph in Fig.~\ref{fig:interference_pattern}. 
+The pattern is enlarged by a diverging lens between the beamsplitter and the screen to aid in viewing. 
+A change in the relative length of the interferometer arms results in the interference rings moving radially inwards or outwards. 
+Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
+%This results in a phase difference between the two beams when they meet at the screen.
+This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
+For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
+When the path difference changes, the fringes stay circular but move radially in or out.
+%A ring interference pattern is often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily seen.
 

### 2020-08-16 paper/ifo-wanderingTone.tex
task 2 => tasks complete
some general text editing and addressing comments

@@
 
-% Tracking tones with the Viterbi algorithm
-% explaination and successful application
-% lean on the continuous gravitational waves angle
 Continuous gravitational wave signals may wander slowly in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars~\cite{MelatosDouglassSimula:2015,Jones:2010} or stochastic accretion flows in binary neutron stars~\cite{BildstenTB:1998}.
@@
 In Section~\ref{sec:wanderingResults} we present the results for recovery of a wandering audio frequency.
-
-%The data from a gravitaitonal wave observation run can be many months to a year. 
-%One method to search for a slowly wandering continuous wave signal is to split the data into several shorter intervals. 
-%Each chunk is analysed seperately. 
-%In our project we take a Fourier transform of each interval and then create a grid (or spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time. 
-%(In continuous wave searches, a detection statistic is used to calculate the likelihood of a signal being present in a particular frequency bin.)
-%The task is then to find the best path through the spectrogram grid as if it where a weighted graph, where here the weights are the Fourier amplitudes. 
-%The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method find the best path through the grid. 
-%It is used in contiuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
 
@@
 The Viterbi path for a time series sequences spanning $t_0$ to $t_{N_t}$ contains the Viterbi path for sub-sequences of that time series. 
-It is also very computationally efficient. \jam{[Justify this claim to efficiency? Cite something about it being $O(n^2)$ or whatever it is?]}
-
-\begin{comment}
-The route through the grid which maximises the overall probability
-
-Constraints are set on the allowable paths. 
-In continuous wave searches, a physical model is used to decide how much the frequency can wander at each timestep, depends on the type of target. 
-In this work we allow the frequency to either a) stay in the same bin; b) move up a single frequency bin; or c) move down a single frequency bin. 
-We label the probability of transitioning from $f_j$ at $t_{i}$ to $f_{j+1,j,j-1}$ at $t_{i+1}$ as $A_{i,j}$. 
-In this work the three transition option have equal probability.
-
-Our objective is now to find the most likely path through the grid given the data and transition constraints. 
-This is the route which maximises the overall probability, which we call the `Viterbi path'.
-In practice we sum the logarithm of the Fourier amplitude $\log O_{i,j}$ along the path.
-
-The Viterbi algorithm begins at $t_{i=1}$. 
-The algorithm selects the highest probability route for each $f_j$. 
-\begin{equation}
-A_{i,j} O_{i,j+1}
-A_{i,j} O_{i,j}
-A_{i,j} O_{i,j-1}
-\end{equation}
-At $t_{i=2}$ it then selects the best route to this point based on the accumulated probability along the paths so far. 
-For example, at
-
-(a.) The grid point labelled (a) has $f_{j=3}$ and $t_{i=1}$. 
-Looking back at the previous time, there are three options for where the path could have started from at $t_{i=0}$. These are $f_{j=4}$, $f_{j=3}$, or $f_{j=2}$. 
-The algorithm selects $f_{j=3}$ as it has the highest value. 
-(b.) The grid point labelled (b.) is not the best starting position for any of the cells at $t_{i=1}$, no paths starting from this position will be considered from now on.
-
-The probability of a path is given by the accumulated probability along the path. 
-\begin{equation}
-
-\end{equation}
-
-The value of each element in the grid is the logarithm of the normalised Fourier amplitude, which we label $O_{i,j}$.
-
-Our objective is to find the most likely path through the grid which maximises the , the `best path'. 
-In practice, we sum the logarithm of the Fourier amplitudes along the path.
-
-We label the time-frequency grid points as $t_i$ and $f_j$ respectively, where $i=1,2,...N_t$ and $j=1,2,...,N_f$ (and $N_t$ and $N_f$ are the total number of time and frequency bins respectively). 
-Starting at $t_{i=2}$, the algorithm iterates calculates the best path to each $f_j$ from $t_{i=1}$ to $t_=2$.  through $t_i$, at each stage finding the best path to every $f_j$ from $t_{i=0}$.
-
-Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount of frequency wander).
-At each iteration, the algorithm will find the best path to each node in the next sub-graph. 
-This is repeated to the final sub-graph. 
-%n, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph. 
-The Viterbi path is then the overall best path from the first sub-graph to the last. 
-The overall best path from the first sub-graph to the last is then selected, which is called the Viterbi path. 
-At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path. 
-See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.
-
-% moved from above
-The method requires a limit for the allowed frequency change over time, usually informed by a physical model. 
-In this work the frequency $f$ path can move in one of three ways at each time bin: a) up one frequency bin; b) stay in the same frequency bin; or c) down one frequency bin. 
-We set the frequency bin width to be $\Delta f = 0.3\,{\rm Hz}$ which corresponds to a maximum frequency wander of $0.11\,{\rm Hz}{\rm s}^{-1}$.
-%In this work, the frequency wander limit is set to $0.11\,{\rm Hz} {\rm s}^{-1}$ which corresponds to one frequency bin per every time bin. 
-In some cases, we see this proves poorly adjusted to the signal. 
-\end{comment}
+The algorithm is computationally efficient, at every time step all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}).
 
@@
 \label{sec:wanderingResults}
-
-To test the Viterbi algorithm, a sinusoidal signal with wandering frequency was played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
-The injected frequency itself followed a sinusoid with decaying amplitude.
-This choice was arbitrary, any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
-%Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
 
@@
 
+To test the Viterbi algorithm, a sinusoidal signal with wandering frequency was played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
+The injected frequency itself followed a sinusoid with decaying amplitude.
+This choice was arbitrary, any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
+%Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
+
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
@@
 \end{equation}
-Computing this along the path, we end up with the value \jam{???}, which indicates a good recovery of the injected signal.
+Computing this along the Viterbi path we find an error of 0.082, which indicates fair but not total recovery of the injected signal.
 

### 2020-08-16 paper/paper_main.tex
adding app c, general text editing and tidying
tasks 1, 3, and 5
task 4
some general text editing and addressing comments

@@
 \usepackage{comment}
+\usepackage{algpseudocode}
+% \usepackage{algorithm}
+% \usepackage{algorithmic}
+
+\usepackage{lineno}
+\linenumbers
 
@@
 }%
+\affiliation{%
+ School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
 
@@
 }
-\affiliation{%
-\han{check all affiliations}
-}
 
@@
 }
-\affiliation{%
-\han{check all affiliations}
-}
 
-\author{\han{(to do: ask other authors?)} \jam{[Other authors and affiliations I leave to you to contact and clarify]}}
-\email{wmoran@unimelb.edu.au}
+\author{...}
+\email{to do}
 \affiliation{%
@@
 \begin{abstract}
-Gravitational wave observatories around the world are searching for continuous waves: persistent signals from spinning neutron stars. 
+Gravitational-wave observatories around the world are searching for continuous waves: persistent signals from spinning neutron stars. 
 Searches use sophisticated statistical techniques to look for weak signals in noisy data. 
-We show how continuous wave searches can be demonstrated with a table-top model gravitational-wave detector (a Michelson interferometer), where sound plays the role of gravitational waves. 
-Using signal processing techniques from continuous wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
+We show how continuous-wave searches can be demonstrated with a table-top model gravitational-wave detector (a Michelson interferometer), where sound plays the role of gravitational waves. 
+Using signal processing techniques from continuous-wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
 Finally we investigate the performance of the demonstration as an `optical microphone' for recovering music and speech. 
@@
 The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project. 
-The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
-\jam{[update if turned into authors]} 
-This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
-Travel during the project was supported by the Australian National University PhB Science program.
+This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
+This work also received financial support from the Institute of Physics International Member Grant and travel support from the Australian National University PhB Science program.
 
@@
 
+\appendix
+\subfile{ifo-appendix.tex}
+
+\nocite{*}
 \bibliographystyle{myunsrt}
 \bibliography{ifoDemoBib}
-
-\newpage
-\appendix
-\subfile{ifo-appendix.tex}
 

### 2020-08-16 paper/ifo-singleTone.tex
some general text editing and addressing comments
small tidying in the text

@@
 
-% the webcam method
-% how to analyse video, python OpenCV2
+\begin{figure}
+	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
+	\caption{\label{fig:webcam_spectrum}
+Injected tone at a constant frequency: Fourier amplitude of the intensity time series recorded by the webcam. 
+The frequency of the injection is $2.09\,{\rm Hz}$. 
+The recovered signal peaks at $2.099\,{\rm Hz}$ with a FWHM of $0.033\,{\rm Hz}$. 
+There are also two visible harmonic peaks at integer multiples of the main peak. 
+Their amplitudes as a fraction of the height of the main peak are  $0.103$ and $0.098$ for the $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ peaks respectively. 
+}
 
-Continuous wave searches look for signals which are close to monochromatic. 
-In this section we consider a simple sinusoidal tone; a time at a single constant frequency. 
+\end{figure}
+
+Continuous-wave searches look for signals which are close to monochromatic. 
+In this section we consider a simple sinusoidal tone; a note at a single constant frequency. 
 As described in Section~\ref{sec:ifo}, the audio is played through a speaker fixed to the back of one of the mirrors.
@@
 
-%The resulting data is a video capture of the whole interference pattern. 
-%As only the frequency of the interference ring motion is needed, w
 We take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
@@
 
-%\han{[to do - rewrite once James has the new numbers (also Fig 3 caption)] 
-%We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ (where the error represents the $1\sigma$ uncertainty).
-%The demonstration is able to recover a single frequency tone to within two standard deviations. }
-%\jam{[For the single tone result, the FWHM is 0.033 Hz. This puts the injected 2.09 Hz well within the uncertainty in the measured value of 2.099 Hz.]}
-%We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
-%The two tones played together sound similar and don’t produce any noticeable beating.
-%To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
-%From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
-%The experiment is able to recover a single tone confidently. 
-%This showed that the experiment observes a single tone confidently.
-
-%\jam{[re: quantifying the non-linear response, the first two harmonics (at 4.19 and 6.28 Hz) are 0.103 and 0.098 the height of the main peak respectively.]}
-
-\begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{\label{fig:webcam_spectrum}
-Injected tone at a constant frequency: Fourier amplitude of the intensity time series recorded by the webcam. 
-%Single injected tone results. 
-%The Fourier amplitude of the intensity timeseres recorded by webcam. 
-The frequency of the injection is $2.09\,{\rm Hz}$. 
-%The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
-The recovered signal peaks at $2.099\,{\rm Hz}$ with a FWHM of $0.033\,{\rm Hz}$. 
-There are also two visible harmonic peaks at integer multiples of the main peak. 
-Their amplitudes as a fraction of the height of the main peak are  $0.103$ and $0.098$ for the $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ peaks respectively. 
-%\han{[to do: add new numbers from James]}
-%\jam{[For the single tone result, the FWHM is 0.033 Hz. This puts the injected 2.09 Hz well within the uncertainty in the measured value of 2.099 Hz.]}
-%The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
-The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.
-}
-
-\end{figure}
-
-\begin{comment}
-As described in Section~\ref{sec:introduction}, continuous gravitational wave searches look for near monochromatic signals. 
-The audio analogue of this is a simple sinusoidal tone; a constant note.
-
-The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam placed to view the screen (see Figure~\ref{fig:ifo_schmatic_webcam}). 
-The webcam's sample rate is $30\,{\rm Hz}$, which limits the spectral content of observable signals to below 15Hz, the Nyquist frequency.
-A tone under $15\,{\rm Hz}$ is played through the speaker for a mintue and the interference pattern recorded with the webcam.
-
-The data is taken from a single off-centre pixelm, indicated by the red dot in Figure~\ref{fig:interference_pattern}. 
-We take the green-channel of the video as an approximation to the total intensity as the laser is monochromatic green light (at $532\,{\rm nm}$).
-
-We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}.
-The two tones played together sound similar and don’t produce any noticeable beating.
-To quantify the precision, we recorded a signal from an injected tone of 2.09Hz.
-From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\mathrm{Hz}$, a result within two standard deviations.
-This showed that the experiment observes a single tone confidently.
-
-\begin{figure*}
-	\includegraphics[width=\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
-	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of 2.09Hz, the amplitude 17.5 peak was measured at $(2.099\pm 0.008) \mathrm{Hz}$.
-The actual maximum is achieved at 0Hz, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
-	\label{fig:webcam_spectrum}
-\end{figure*}
-\end{comment}
-
 \end{document}

### 2020-08-16 paper/ifo-tableTopGWs.tex
small tidying in the text

@@
 \begin{figure}
-% 	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam.pdf}
 	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam_edit.pdf}
@@
 The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}. 
-The beam is first reflected from a mirror at a right angle and then focused at a converging lens before the beamsplitter which is $50$/$50$ in reflection and transmission. 
+The beam is first reflected from a mirror at a right angle and then focused at a converging lens before reaching the beamsplitter which is $50$/$50$ in reflection and transmission. 
 The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and are recombined at the beamsplitter. 
@@
 A speaker fixed to the back of M2 provides the signal input from the PC.  
-%Michelson interferometer schematic with interference pattern recorded by a commercial webcam.
-The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}.
-The beam is first reflected from a mirror a right angle (to save space on the optical breadboard) and then focused at a converging lens before the $50$/$50$ beamsplitter.
-The split beam is reflected from mirrors M1 and M2 at the ends of the interferometer arms and recombined at the beamsplitter.
-The interference pattern is enlarged by a diverging lens after the beamsplitter and projected onto a screen.
-A commertial speaker is fixed to the back of M2, as the signal input.
-The interferometer arms lengths are roughly $3\,{\rm cm}$ different and the entire interferometer (without the screen) fits within the dimensions of an A4 piece of paper.
-We did not include a compensating plate with the beamsplitter, as is often done.}
     }
@@
 The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
-% Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
-For scale, the central fringe is approximately $5\,{\rm mm}$ across.
 }
@@
 Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed.
-%This results in a phase difference between the two beams when they meet at the screen.
-This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
-For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-When the path difference changes, the fringes stay circular but move radially in or out.
-%A ring interference pattern is often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily seen.
 
@@
 It is attached to the back of the mirror using commercial adhesive putty. 
-The other speaker in the pair is kept face-down and as far away from the interferometer as the cables to reduce its impact as a second source of vibrations.
+The other speaker in the pair is kept face-down and as far away from the interferometer as the cables allow to reduce its impact as a second source of vibrations.
 
-%As the speaker deforms to play sound, the mirror moves producing a change in arm length related to some coupling function of the amplitude and frequency of the audio. 
 The motion of the interference fringes follows the motion of the mirror, and therefore the speaker. 
@@
 As such, any motion of the fringes must be kept small by playing sound softly through the speaker.
-Even without over-counting, large fringe motions introduce a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome - but physically interesting!
-- phenomena like frequency doubling.
-We derive the non-linear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen in Appendix~\ref{app:intensity_derivation}.
+Even without over-counting, large fringe motions introduce a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting!---phenomena like frequency doubling.
+We derive the nonlinear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen in Appendix~\ref{app:intensity_derivation}.
 
-Data can be recorded via a webcam (as shown in Fig.~\ref{fig:ifo_schematic_webcam}) or photo-diode. 
+Data can be recorded via a webcam (as shown in Fig.~\ref{fig:ifo_schematic_webcam}) or photodiode. 
 In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used. 
@@
 
-\begin{comment}
-Even though gravitational waves are caused by some of the most massive events in the universe, their observable effects on Earth are vanishingly small | on the order of only one part in $10^{21}$ for our best detectors~\cite{GW150914}.
-Clearly, the measurement of such effects is far beyond any table-top experiment.
-But, the fundamental principle of gravitational wave detection remains that changes in the lengths of the beam arms of an interferometer leads to a change in the interference pattern observed.
-By observing sound waves that vibrate the mirrors of a table-top interferometer and so change the lengths of the beam arms we have an analogue of this principle.
-
-\subsection{Method of Michelson interferometry to observe sound}
-% hardware: path difference, virtual thin film interference
-
-In a Michelson interferometer, a coherent (monochromatic and equal phase) source of laser light is split evenly by a beamsplitter into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
-The beams recombine and interfere at the beamsplitter before hitting a screen to produce a pattern of concentric, circular fringes.
-The configuration used in this experiment is shown diagrammatically in Figure~\ref{fig:ifo_schematic_webcam} and as part of a photo in Figure~\ref{fig:setup_pic2}.
-
-A sample interference pattern is shown in Figure~\ref{fig:interference_pattern}.
-This pattern is due to the path difference between the two arms of unequal lengths (if the lengths were equal, then only a single, bright dot would be seen).
-This results in a phase difference between the two beams when they meet at the screen.
-This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
-For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-When the path difference changes, the fringes stay circular but move radially in or out.
-
-We used a 0.5W speaker stuck to the back of one of the mirrors as an audio source for our demonstration.
-This speaker was part of a commercial 3.5mm-jack, USB powered pair of speakers and was driven by a PC.
-To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place.
-We kept the other speaker in the pair face-down and as far away from the experiment as the cables would allow to reduce it acting as a second source.
-
-As the speaker deforms to play sound, the mirror moves with a change in separation as some coupling function of the amplitude and frequency of the sound.
-The motion of the fringes follows this motion of the mirror and therefore the speaker at the same frequency.
-If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
-If the fringe motion was large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting would artificially raise the measured frequency.
-As such, any motion of the fringes must be kept small by playing sound softly through the speaker.
-
-For later sections, we derive the non-linear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen, this can be found in Appendix~\ref{app:intensity_derivation}.
-
-% Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
-We did not find it necessary to include it to be able to demonstrate the results desired.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/webcam_still0.pdf}
-	\caption{Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
-For scale, the central fringe is approximately 5mm across}
-	\label{fig:interference_pattern}
-\end{figure}
-\end{comment}
-
 \end{document}

### 2020-08-16 paper/ifo-wanderingTone.tex
small tidying in the text

@@
 The value at each spectrogram point determines the likelihood that a signal is present at a particular time and frequency. 
-In gravitational-wave data analysis, a detection statistic is used to calculate the likelihood of a signal given the antenna beam pattern of the detector, which varies with as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
-The detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}). 
+In gravitational-wave data analysis, a detection statistic is used to calculate the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
+The detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}. 
 In this work the detection statistic is the amplitude (modulus) of the Fourier transform of the data at each time segment, which maximizes the likelihood of detecting a sinusoid in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}.
 The resulting spectrogram gives the Fourier amplitude for each frequency and time bin.
-The task is then to find the most likely wandering frequency signal through the spectrogram. 
+
+In continuous wave searches a hidden Markov model is used to describe the system. 
+The frequency (the state) of the signal is unknown (hidden) and the signal is Markovian in that its current state depends only on the previous state. 
+The hidden Markov model method relates the observable quantity (the detection statistic) to the signal frequency (the hidden state). 
+The task is then to find the most likely wandering frequency signal given the observable of the spectrogram. 
 The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method to find the most likely path of a signal given an observation sequence. 
-It is used in continuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
+This method is used in continuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,JonesSun:2020,MiddletonEtAlO2LMXBs:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
 
@@
 
-\subsection{The Viterbi algorithm}
+\subsection{The hidden Markov model and Viterbi algorithm}
 \label{sec:viterbi}
@@
 
-%Here we overview the Viterbi algorithm and describe its usage in this work. 
-%Thinking about this 
-A spectrogram is a grid indexed by time $t$ and frequency $f$ where here there are $N_t$ and $N_f$ time and frequency bins respectively, as shown in Fig.~\ref{fig:viterbi}.
+The spectrogram grid is indexed by time $t$ and frequency $f$ where there are $N_t$ and $N_f$ time and frequency bins respectively, as shown in Fig.~\ref{fig:viterbi}.
 We label the elements of the grid as $t_i$ and $f_j$ with $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-The value in each grid element is a detection statistic based on the time series data observed. 
-It indicates the likelihood that a signal is present in the element. 
-We label the normalised Fourier amplitude as $F(t_i,f_j)$. 
-%In continuous wave searches the detection statistic depends on the type of gravitational wave source. 
-%In this work we use the normalised Fourier amplitude, which we label $F(t_i,f_j)$.
+The normalised Fourier amplitude $F(t_i,f_j)$ is the detection statistic indicating the likelihood that a signal is present in each grid element.
 
@@
 In this work, at each time step we allow the frequency to either a) stay in the same bin; b) move up a single frequency bin; or c) move down a single frequency bin. 
-The probability of transitioning from $f_j$ at $t_i$ to any of $\{f_{j+1},f_j,f_{j-1}\}$ at $t_{i+1}$ is called the transition probability $A(f_k,f_m)$. 
+In a hidden Markov model, the probability of transitioning from $f_j$ at $t_i$ to any of $\{f_{j+1},f_j,f_{j-1}\}$ at $t_{i+1}$ is called the transition probability $A(f_k,f_m)$. 
 In this work the three options have equal probability $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
@@
 A path through the grid has a likelihood dependent on the detection statistic (the observation) and the transition probability.
-If the starting probabilities are all equal, i.e. $Pr[f(t_0)] = 1/N_f$, the the probability of the frequency equalling $f(t_{i+1})$ at $t = t_{i+1}$ is defined recursively by
-%If there is equal probability of starting $f_j$ at $t=t_0$, the probability of path $Q$ given the observation $O$ is 
+If the starting probabilities are all equal, i.e. ${\rm Pr}[f(t_0)] = 1/N_f$, the the probability of the frequency equalling $f(t_{i+1})$ at $t = t_{i+1}$ is defined recursively by
 \begin{eqnarray}
-%P(Q|O) =& A(f_j) F(t=N_t,f_j), \nonumber \\ &~~ \times \dots, A(f_j) F (t=t_0,f_j).
-Pr[f(t_{i+1})=f_j] =~& F[t_{i+1},f(t_{i+1})] \nonumber \\
-                     &*A[f(t_{i+1}),f(t_i)]  \nonumber \\
-                     & *Pr[f(t_i)].
+{\rm Pr}[f(t_{i+1})=f_j] =~& F[t_{i+1},f(t_{i+1})] \nonumber \\
+                     &\times A[f(t_{i+1}),f(t_i)]  \nonumber \\
+                     &\times {\rm Pr}[f(t_i)].
 \label{eqn:recursiveViterbi}
 \end{eqnarray}
-The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes $Pr[f(t_{N_t}) = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
+The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes ${\rm Pr}[f(t_{N_t}) = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
 We can the maximize the latter quantity over $0 \leq j \leq N_f$ to find the optimal Viterbi path overall, i.e. terminating in any frequency bin.
@@
 
-\item At $t=t_{N_t}$, each $f_j$ has a value $Pr[f(t_{N_t})=f_j]$ (equation~\ref{eqn:recursiveViterbi}). 
+\item At $t=t_{N_t}$, each $f_j$ has a value ${\rm Pr}[f(t_{N_t})=f_j]$ (equation~\ref{eqn:recursiveViterbi}). 
 The algorithm selects the terminating frequency $f(t_{N_t})$ with the highest value, marked (b) in Fig.~\ref{fig:viterbi}.

### 2020-08-18 paper/ifo-complexAudio.tex
some editing in wandering tone and complex audio paper tex

@@
 % aim of optical microphone: play back injected sounds
-A natural successor to simple tomes is complex audio, such as music and speech. 
+A natural successor to single and wandering tones is complex audio, such as music and speech. 
 Complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not applicable directly.
@@
 The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
-It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
+%It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
 %It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae. [refs; my last sentence can be improved]
@@
 We find that the raw output of the optical microphone, collected without any further processing (except anti-aliasing), is noisy, with a loud bass hum throughout.
-Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic.
-\jam{[non-linear appendix, Andrew says to cite Feynman lectures on non-linear oscillations]}
+Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic, likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}).
 This mains power signal is present in the power spectral density of background noise as shown in the top panel of Figure~\ref{fig:psd_noise}.
@@
 The next step is to apply filters to the data. 
-The ideal filter would be one that 
-\begin{itemize}
-\item[i)] completely attenuates the undesired parts of the spectrum, 
-\item[ii)] does not change the rest of the spectrum, and 
-\item[iii)] has smooth edges so as to not damage the signal after convolution. 
-\end{itemize}
+The ideal filter would be one that:
+%\begin{itemize}
+i) completely attenuates the undesired parts of the spectrum, 
+ii) does not change the rest of the spectrum, and 
+iii) has smooth edges so as to not damage the signal after convolution. 
+%\end{itemize}
 These three conditions are contradictory and any filter must compromise between them.
@@
 
-For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\varepsilon$, the frequency response (attenuation at each frequency) as a function of frequency $\omega$ is:
+We also apply a Butterworth filter. 
+A low-pass Butterworth filter with cut-off frequency $f_c$ and gain $\varepsilon$, has a frequency response (attenuation at each frequency)
 \begin{equation}
 \label{eq:butterworth}
-H(\omega) = \left(1+\varepsilon^2 \left( \frac{\omega}{\omega_c} \right)^{2n}\right)^{-\frac{1}{2}}
+H(\omega) = \left(1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right)^{-\frac{1}{2}}
 \end{equation}
-
-\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
-To construct a Butterworth band-pass filter, the low-pass filter in Eq.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
+% commenting to send to co-authors
+%\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
+To construct a Butterworth band-pass filter, the low-pass filter in Eqn.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
 The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}).
-The response of the filter used is shown in Figure~\ref{fig:butterworth} and its effect on the power spectral density of background noise is shown in the bottom panel of Figure~\ref{fig:psd_noise}.
+The effect of the Butterworth band-pass filter on the power spectral density of the background noise is shown in the bottom panel of Fig.~\ref{fig:psd_noise}, where we see that it reduces the amplitude of the mains harmonics below $150\,{\rm Hz}$
+The response of the filter is shown in Fig.~\ref{fig:butterworth}.
 Band-passing the signal does not address the remaining harmonics above $100\,{\rm Hz}$. 
 However, it does suppress other noise at higher frequencies. 
-The Butterworth filter performs well at maintaining source intelligibility as it is optimised to be `maximally flat' within the band region (prioritising condition (ii) above).
-However recordings of speech still have low intelligibility (the speech can't be understood) as the filter does not improve beyond the source.
-% They can be recognised as speech without being understood.
+The Butterworth filter performs well at maintaining source intelligibility as it is optimised to be `maximally flat' within the band region (prioritising condition ii) above).
+However recordings of speech still have low intelligibility (the speech can't be understood).
 Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique to address this.
@@
 We also consider a combined notch filter and Wiener filter. 
-Timeseries and frequency spectrum results are shown in Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively. 
-The original signal is identical to that shown in Fig.~\ref{fig:noiseSubtract}.
-The first and second panels of Fig.~\ref{fig:timeNotchWiener} show the original audio signal and the recording after passing through the optical microphone (i.e.
+Here we describe our use of these methods with timeseries and frequency spectrum results shown in Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively. 
+The original signal is identical to that shown in Fig.~\ref{fig:noiseSubtract} where the first and second panels of Fig.~\ref{fig:timeNotchWiener} show the original audio signal and the recording after passing through the optical microphone respectively (i.e.
 the timeseries of the data shown in the first two panels of Fig.~\ref{fig:noiseSubtract}).

### 2020-08-18 paper/ifo-wanderingTone.tex
some editing in wandering tone and complex audio paper tex

@@
 Some routes through the grid are found to be dead ends, such as the path ending at $t_1$ marked (a). 
-At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by equation~\ref{eqn:recursiveViterbi}, marked (b). 
+At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:recursiveViterbi}, marked (b). 
 The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
@@
 
-\item Moving to $t_2$, the path with the highest value of the recursive relation in equation~\ref{eqn:recursiveViterbi} is selected for each $f_j$ as shown by the solid lines from $t_1$ to $t_2$ in Fig.~\ref{fig:viterbi}.
+\item Moving to $t_2$, the path with the highest value of the recursive relation in Eqn.~\ref{eqn:recursiveViterbi} is selected for each $f_j$ as shown by the solid lines from $t_1$ to $t_2$ in Fig.~\ref{fig:viterbi}.
+The origins are stored for each grid point.
 
@@
 
-\item At $t=t_{N_t}$, each $f_j$ has a value ${\rm Pr}[f(t_{N_t})=f_j]$ (equation~\ref{eqn:recursiveViterbi}). 
+\item At $t=t_{N_t}$, each $f_j$ has a value ${\rm Pr}[f(t_{N_t})=f_j]$ (Eqn.~\ref{eqn:recursiveViterbi}). 
 The algorithm selects the terminating frequency $f(t_{N_t})$ with the highest value, marked (b) in Fig.~\ref{fig:viterbi}.
 
-\item The Viterbi path is then found by backtracking along the best paths at each time step.
-\jam{[Andrew mentioned providing a closed form here, which I'm not sure exists.
+\item The Viterbi path is then found by backtracking along the stored best paths at each time step (see also Appendix~\ref{app:viterbi}). 
+% come back to this. 
+%\jam{[Andrew mentioned providing a closed form here, which I'm not sure exists.
 The code works by storing three arrays: the weights, the score for the best path to a given vertex, and the index of the previous vertex along that path.
@@
 Quantitatively, the Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time. 
-We also consider the root mean square (RMS) fractional error along the path.
-Given $i=0,1,...,N_f$ columns of the injected frequency $I_i$ and recovered (Viterbi) frequency $R_i$, the RMS fractional error is given by:
+We also consider the root mean square (RMS) fractional error $E_{\rm rms}$ along the path.
+Given $i=0,1,...,N_f$ columns of the injected frequency $I_i$ and recovered (Viterbi) frequency $R_i$, $E_{\rm rms}$ is given by
 \begin{equation}
-\sqrt{ \frac{1}{N_f+1} \sum_{i=0}^{N_f} \frac{(I_i - R_i)^2}{{I_i}^2} }
+E_{\rm rms} = \sqrt{ \frac{1}{N_f+1} \sum_{i=0}^{N_f} \frac{(I_i - R_i)^2}{{I_i}^2} }
 \end{equation}
-Computing this along the Viterbi path we find an error of 0.082, which indicates fair but not total recovery of the injected signal.
+Computing this along the Viterbi path we find $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
 

### 2020-08-21 paper/ifo-conclusions.tex
editing conclusion and future work

@@
 Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
-\jam{[Confirming Hannah's comment, the Pi was only necessary to read the photodiode, the webcam video I just saved onto PC and analysed directly]}
 We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 

### 2020-08-21 paper/ifo-futureWork.tex
tasks 1, 3, and 5
some general text editing and addressing comments
editing conclusion and future work

@@
 
-% wild speculation
-%We are currently working towards two improvements of the optical microphone.
+\begin{figure*}
+	\includegraphics[height=.99\textwidth, angle=-90]{figures/pipeline.pdf}
+	\caption{
+Flowchart for the optical microphone transfer function. 
+The stage inside the red-dash box is analysed in Appendix~\ref{app:intensity_derivation}. Understanding the other stages is one aim of future work. 
+}
+	\label{fig:pipeline_highlighted}
+\end{figure*}
 
@@
 
-A second improvement is to measure the full model of the transfer function of the system. 
-With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
-\jam{[check with Rob et al. about the veracity of this]}
+A second improvement is to measure the full model of the transfer function of the system, providing more information about how the audio signal couples with the interferometer.
+% commented out whilst James is checking this
+%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
+%\jam{[check with Rob et al. about the veracity of this]}
 The transfer function starts from the voltage sent to the speaker and ends with the voltage recorded by the Pi.
-It includes any non-linearities in the speaker itself, the speaker-mirror coupling, the separation-intensity relation Appendix~\ref{app:intensity_derivation}, and the photodiode reading.
+It includes any non-linearities in the speaker itself, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode reading.
 A schematic of the full optical microphone transfer function is shown in Figure~\ref{fig:pipeline_highlighted}.
@@
 
-\begin{figure*}
-	\includegraphics[height=.99\textwidth, angle=-90]{figures/pipeline.pdf}
-	\caption{
-Flowchart for the optical microphone transfer function. 
-The stage inside the red, dashed box is analysed in Appendix~\ref{app:intensity_derivation}. Understanding the other stages is one part of future work. 
-}
-	\label{fig:pipeline_highlighted}
-\end{figure*}
-
 \end{document}

### 2020-08-28 paper/ifo-complexAudio.tex
editing text in complex audio and wandering tone

@@
 
-We describe additional components to the demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}.
+We describe additional hardware components to the demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}.
 Filter and speech enhancement techniques applied to the data are detailed in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse} respectively.
@@
 
-\subsection{Photodiode}
+\subsection{Hardware modifications for the optical microphone}
 \label{sec:photodiode}
@@
 Therefore the optical microphone data requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music. 
-We achieved this with a photodiode.
-
+We achieved this with a photodiode\footnote{
 A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the `reverse' direction.
 As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.
-% Fundamentally, a current is created in the photodiode by the photo-electric effect.
-We place a OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp. 
-This creates a photo-detector that produces a voltage dependent on the incident intensity. 
+}
+
+We place a OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which makes a photo-detector that produces a voltage dependent on the incident intensity. 
 The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
@@
 The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath. 
-
-The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data.
-Together, these sample the signal at roughly $16\,{\rm kHz}$. Resources for using the Pi are found in Appendix~\ref{app:circuit_diagram}.
+The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi (Pi) Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data.
+Together, these sample the signal at roughly $16\,{\rm kHz}$. 
+Resources for using the Pi and photdiode are detailed in Appendix~\ref{app:circuit_diagram}.
 
@@
 
-An updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a photograph of the entire optical microphone in Figure~\ref{fig:setup_pic2}.
+An updated schematic is shown in Fig.~\ref{fig:ifo_schematic_podo} and a photograph of the entire optical microphone in Fig.~\ref{fig:setup_pic2}.
 See Appendix~\ref{app:circuit_diagram} for a diagram and photograph of the circuit.
@@
 \label{sec:initialResultsOpMic}
-% breakdown of all the filters used along the way
-The optical microphone was tested on recordings of varying speech from different people and on music ranging from simple melodies and rhythms to full songs. 
-The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources).
-When processing the results, we often only looked at the first ten seconds of each observation for efficiency.
+
+We test the optical microphone with a variety of recordings including speech from different people and music ranging from simple melodies and rhythms to full songs. 
+%The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources). 
 During recordings, care was taken to ensure minimal activity around the interferometer to reduce environmental noise. 
 The timeseries was then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}, as before.
+When processing the results, we generally only analysed the first $10\,{\rm s}$ of each observation for efficiency.
 
-We find that the raw output of the optical microphone, collected without any further processing (except anti-aliasing), is noisy, with a loud bass hum throughout.
-Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic, likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}).
-This mains power signal is present in the power spectral density of background noise as shown in the top panel of Figure~\ref{fig:psd_noise}.
-The mains power signal is far weaker but still present in the background noise taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution.
-The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination. 
-Outside the mains noise, the power appears to peak broadly around 750 Hz, the origin or significance of which we don't know.
+The raw output of the optical microphone (with only anti-aliasing) is noisy with a loud bass hum throughout. 
+Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic, likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}).
+This mains power signal is present in the power spectral density of background noise, with the speaker switched off, as shown in the top panel of Fig.~\ref{fig:psd_noise}.
+The mains signal is still present (but far weaker) in measurements with the photodiode in darkness, suggesting that ambient lighting has a large contribution.
+Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode. 
+There is also a broad peak visible at around $750\,{\rm Hz}$ in Fig.~\ref{fig:psd_noise}, which we have not been able to identify an origin for.
 
@@
 An intuitive means to remove noise is to subtract the noise spectrum of the instrument directly from the signal spectrum. 
-The third panel of Fig.~\ref{fig:noiseSubtract} shows a recording from the optical microphone with no audio signal. 
+The third panel of Fig.~\ref{fig:noiseSubtract} shows a recording from the optical microphone with no input audio signal. 
 We see that the output noise is coloured, which reflects the non-linearity of the optical microphone.
@@
 
-We also apply a Butterworth filter. 
+We next apply a Butterworth filter. 
 A low-pass Butterworth filter with cut-off frequency $f_c$ and gain $\varepsilon$, has a frequency response (attenuation at each frequency)
@@
 \label{eq:butterworth}
-H(\omega) = \left(1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right)^{-\frac{1}{2}}
+H(\omega) = \left[1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right]^{-1/2}
 \end{equation}
@@
 However, it does suppress other noise at higher frequencies. 
-The Butterworth filter performs well at maintaining source intelligibility as it is optimised to be `maximally flat' within the band region (prioritising condition ii) above).
+In general, the Butterworth filter performs well at maintaining source intelligibility as it is optimised to be `maximally flat' within the band region (prioritising condition ii) above).
 However recordings of speech still have low intelligibility (the speech can't be understood).
@@
 \end{equation}
-where $\alpha$ and $\beta$ are self-tuned parameters which adjust the locations of poles and zeros. 
+where $\alpha$ and $\beta$ are self-tuned parameters which adjust the locations of poles and zeros and $z$ is the complex frequency. 
 By allowing the numerator and demoninator of $H(z)$ to be zero, we can find the corresponding locations of zeros and poles of the filter. 
-Notice zeros are at $e^{\pm jw_0}, \text{where } w_0=\cos^{-1}(\beta)$, and poles are at $re^{\pm j\phi}, \text{where } r=\sqrt{\alpha},\phi=\cos^{-1}(\beta(1+\alpha)/(2\sqrt{\alpha}))$.
+There are zeros are at $e^{\pm jw_0}, \text{where } w_0=\cos^{-1}(\beta)$, and poles are at $re^{\pm j\phi}, \text{where } r=\sqrt{\alpha} \text{ and }\phi=\cos^{-1}[\beta(1+\alpha)/2\sqrt{\alpha}]$.
 At $w_0$, the magnitude response will be zero. 
 Thus, we can calculate $\beta$ setting $w_0$ to be the frequency we want to remove. 
-The parameter $\alpha$ affects the bandwidth ($B_w$) of the filter and has the relation $B_w=\cos^{-1}(\frac{2\alpha}{1+\alpha^2})$. \\
-In our paper, we first decide $\beta$ for each filter. 
-We locate hum noise spectrum at around $50k\,{\rm Hz}$ where the integer $k=1,2,\dots,15$. 
-For each $k$, we build a notch filter to filter out the specific frequency bin approximately at $50k \, {\rm Hz}$. 
+The parameter $\alpha$ affects the bandwidth $B_w$ of the filter where $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$. 
+%In our paper, we first decide $\beta$ for each filter.
+
+In this work we find a cascade filter with $15$ notches works well. 
+We decide $\beta$ first by building a notch filter at each of the frequency bins located at $50k\,{\rm Hz}$ where the integer has values $k=1,2,\dots,15$. 
+%The mains noise is located at $50k\,{\rm Hz}$ where the integer has values $k=1,2,\dots,15$. 
+%For each $k$, we build a notch filter at the specific frequency bin at $50k \, {\rm Hz}$. 
 We then alter the parameter $\alpha$ to determine the bandwidth of each filter. 
@@
 On the other side, the bandwidth should have enough residues to account for the uncertainty of our pinpointing the location of hum noise. 
-After considering these two aspects, we are now able to get the ideal transfer function for each filter. 
+After considering these two aspects, we find the ideal transfer function for each filter. 
 The whole cascaded filter is achieved by multiplying them together, 
@@
     \label{eqn:notch15}
-    H(z)=H_1(z)H_2(z)\dots H_k(z),k=1,2,\dots 15\,,
+    H(z) = H_{k=1}(z) H_{k=2}(z)\dots H_{k=15}(z)
+%    H(z)=H_1(z)H_2(z)\dots H_k(z),k=1,2,\dots 15\,,
 \end{equation}
@@
 The magnitude response of the first five notch filters are shown in Fig.~\ref{fig:notchMagResponse}. 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-\includegraphics[width=.49\textwidth]{figures/1111}
+\includegraphics[width=.7\textwidth]{figures/1111}
 \end{center}
@@
 }
-\end{figure}
+\end{figure*}
 
-The timeseries and frequency spectrum results of the notch filter are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}. 
-We seen that the mains harmonics are significantly attenuated in comparison to the analysis in Fig.~\ref{fig:noiseSubtract}.
+The timeseries and frequency spectrum results of the notch filter are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
+We see that the mains harmonics are significantly attenuated in comparison to the analysis in Fig.~\ref{fig:noiseSubtract}.
 
@@
 
-A Wiener filter is used to compute an estimate of the original signal. 
+Here we use a Wiener filter to compute an estimate of the original signal. 
 We label the noisy speech signal sequence as  $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence. 
@@
 \end{equation}
-Given the observation data $\textbf{x}$, our goal is to filter the signal $\textbf{s}$ from the noise $\textbf{w}$, and get an estimated $\hat{\textbf{s}}$, where the difference between $\hat{\textbf{s}}$ and true signal $\textbf{s}$ is minimized in the mean square error sense, which is represented as
+Given the observation data $\textbf{x}$, our goal is to filter the signal $\textbf{s}$ from the noise $\textbf{w}$ to find an estimated $\hat{\textbf{s}}$, where the difference between $\hat{\textbf{s}}$ and true signal $\textbf{s}$ is minimized in the mean square error sense, which is represented as
 \begin{equation}
@@
 where $\text{BMSE}$ is the Bayesian mean square error. 
-Assuming that (i) $\textbf{x}$ is wide sense stationary~\footnote{
+Assuming that: i) $\textbf{x}$ is wide sense stationary~\footnote{
 A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
@@
 only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} 
-with zero mean; (ii) the signal $\textbf{s}$ is zero mean; (iii) the noise processes $\textbf{w}$ are uncorrelated with the signal $\textbf{s}$, we can further express the estimated $\hat{\textbf{s}}$ to be a linear combination of present and past input observed data
+with zero mean; ii) the signal $\textbf{s}$ is zero mean; and iii) the noise processes $\textbf{w}$ are uncorrelated with the signal $\textbf{s}$, we can further express the estimated $\hat{\textbf{s}}$ to be a linear combination of present and past input observed data
 \begin{equation}

### 2020-08-28 paper/ifo-wanderingTone.tex
editing text in complex audio and wandering tone

@@
 Continuous gravitational wave signals may wander slowly in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars~\cite{MelatosDouglassSimula:2015,Jones:2010} or stochastic accretion flows in binary neutron stars~\cite{BildstenTB:1998}.
-%As also mentioned in Section~\ref{sec:introduction}, continuous wave searches often look for signals that wander slowly in frequency~\cite{ScoX1O2Viterbi:2019}. 
 The audio analogue of this is a tone that wanders in frequency; a note that changes pitch.
-%Unlike the single tone described in Section~\ref{sec:single_tone}, finding the signal isn’t as simple as searching for a peak in the spectrum for the full timeseries data.%observing run.
 
@@
 In continuous wave searches a hidden Markov model is used to describe the system. 
-The frequency (the state) of the signal is unknown (hidden) and the signal is Markovian in that its current state depends only on the previous state. 
+The frequency (the state) of the signal is unknown (hidden) and the signal is Markovian in that the state at any time depends only on the state in had at the previous time step. 
 The hidden Markov model method relates the observable quantity (the detection statistic) to the signal frequency (the hidden state). 
-The task is then to find the most likely wandering frequency signal given the observable of the spectrogram. 
+The task is to find the most likely wandering frequency signal given the observable of the spectrogram. 
 The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method to find the most likely path of a signal given an observation sequence. 
@@
 \label{sec:viterbi}
-
-\begin{figure}
-\includegraphics[width=0.49\textwidth]{figures/viterbiDiagram.pdf}
-\caption{\label{fig:viterbi}
-A schematic diagram of the Viterbi algorithm. 
-Each box represents an element in the time-frequency grid starting from the instant $t_0$ on the left and ending at $t_{N_t}$ on the right. 
-Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
-The colour of each circle indicates the likelihood that a signal is present in the box, where a lighter colour corresponds to higher likelihood. 
-AT each time step, the paths can move up one $f$ step, move down one $f$ step, or stay at the same $f$ step (shown by lines in the diagram). 
-Each circles has three paths leading to it indicated by the lines. 
-The best path to each circle at each $t_i$ is highlighted in black. 
-Some routes through the grid are found to be dead ends, such as the path ending at $t_1$ marked (a). 
-At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:recursiveViterbi}, marked (b). 
-The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
-}
-\end{figure}
 
@@
 In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c). 
-The Viterbi path is highlighted in orange in fig.~\ref{fig:viterbi}.
+The Viterbi path is highlighted in orange in Fig.~\ref{fig:viterbi}.
 \end{enumerate}
@@
 The algorithm is computationally efficient, at every time step all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}).
+
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{figures/viterbiDiagram.pdf}
+\caption{\label{fig:viterbi}
+A schematic diagram of the Viterbi algorithm. 
+Each box represents an element in the time-frequency grid starting from the instant $t_0$ on the left and ending at $t_{N_t}$ on the right. 
+Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
+The colour of each circle indicates the likelihood that a signal is present in the box, where a lighter colour corresponds to higher likelihood. 
+AT each time step, the paths can move up one $f$ step, move down one $f$ step, or stay at the same $f$ step (shown by lines in the diagram). 
+Each circles has three paths leading to it indicated by the lines. 
+The best path to each circle at each $t_i$ is highlighted in black. 
+Some routes through the grid are found to be dead ends, such as the path ending at $t_1$ marked (a). 
+At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:recursiveViterbi}, marked (b). 
+The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
+}
+\end{figure}
 
@@
 
-To test the Viterbi algorithm, a sinusoidal signal with wandering frequency was played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
+To test the Viterbi algorithm, a sinusoidal signal with wandering frequency was played through the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}. 
 The injected frequency itself followed a sinusoid with decaying amplitude. 
 This choice was arbitrary, any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
-%Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
 
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
-The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path respectively maximized over all terminating frequency bins. 
-The Viterbi path provides a good match to the injected frequency. 
-%To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment \han{What was the SNR? / can we quantify this?}.
-Quantitatively, the Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time. 
-We also consider the root mean square (RMS) fractional error $E_{\rm rms}$ along the path.
-Given $i=0,1,...,N_f$ columns of the injected frequency $I_i$ and recovered (Viterbi) frequency $R_i$, $E_{\rm rms}$ is given by
-\begin{equation}
-E_{\rm rms} = \sqrt{ \frac{1}{N_f+1} \sum_{i=0}^{N_f} \frac{(I_i - R_i)^2}{{I_i}^2} }
-\end{equation}
-Computing this along the Viterbi path we find $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
-
-%The Viterbi path stays within one frequency bin (approximately $0.3\,{\rm Hz}$) of the injected tone for $94\%$ of the run. 
-Initially the injected signal moves faster then the frequency wander limit, leading to the discrepancy observed for $t\lesssim 10\,{\rm s}$. 
-%Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals. 
-%To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often. 
+The overlaid pink-dot markers show the injected signal. 
+The white-cross markers show the recovered Viterbi path maximised over all terminating frequency bins, which qualitatively provides a good match to the injection. 
+Initially, the injected signal moves faster than the frequency wander limit, leading to the discrepancy observed for $t\lesssim 10\,{\rm s}$. 
 There is also an anomaly at $150\,{\rm s}$, which illustrates what happens when the interferometer is disturbed, e.g. when someone walks nearby. 
 The Viterbi algorithm is robust against such disturbances, recovering the optimal path promptly, as Fig.~\ref{fig:viterbi_overlay} shows.
-%is likely due to a disturbance around the interferometer (such as someone walking nearby).
-%There also appears to be an anomaly at $150\,{\rm s}$, likely from a disturbance (such as someone walking) around the interferometer.
 
-%Empirically, decreasing the signal-to-noise ratio closer to unity (by making the speaker quieter) causes the Viterbi algorithm to stray more from the signal. 
-%However, since the only correlation through the grid is the signal, we always expect the Viterbi path to lie near to the signal no matter how high the noise is, and this is indeed what we see.
-\han{Should we show results for this?}
+Quantitatively, the Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time. 
+Another figure of merit is the root mean square (RMS) fractional error $E_{\rm rms}$ along the path defined as 
+\han{}
+\begin{equation}
+E_{\rm rms} = \left[ \frac{1}{N_t+1} \sum_{i=0}^{N_t} \frac{(I_i - R_i)^2}{{I_i}^2} \right] ^{1/2}
+\end{equation}
+where $I_i$ and $R_i$ are the injected and recovered (Viterbi) frequency paths. 
+Computing this for the result shown in Fig.~\ref{fig:viterbi_overlay} we find $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
 

### 2020-08-28 paper/paper_main.tex
small tidying in the text
some editing in wandering tone and complex audio paper tex
editing text in complex audio and wandering tone

@@
 
+\author{Changrong Liu}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+
 \author{Robin Evans}
-\email{robinje@unimelb.edu.au}
 \affiliation{%
@@
 \author{William Moran}
-\email{wmoran@unimelb.edu.au}
 \affiliation{%
@@
 
-\author{Changrong Liu}
-\email{}
+\author{Deeksha Beniwal}
 \affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+Department of Physics and The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
 }
 
-\author{...}
-\email{to do}
+\author{Huy Tuong Cao}
 \affiliation{%
-Adelaide, Adelaide-OzGrav
+Department of Physics and The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Craig Ingram}
+\affiliation{%
+Department of Physics and The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Daniel Brown}
+\affiliation{%
+Department of Physics and The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Sebastian Ng}
+\affiliation{%
+Department of Physics and The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{\han{to do: check names - initials?}}
+\affiliation{%
+\han{to do: check all affiliations}
 }
@@
 Simple chords and drums are easily recovered, but speech is more challenging. 
-%however speech recordings are very noisy. 
-%Application of an existing speech enhancement technique provides some noise reduction, however the audio remains unintelligible. 
 This demonstration has scope for use in the physics and electrical engineering undergraduate lab. 
 It can also be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences. 
-%Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy data. 
-%We demonstrate these techniques by observing sound with a table-top Michelson interferometer. 
-%In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-%We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings. 
-%However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
-%Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data. 
-%We demonstrate these techniques by observing sound with a table-top Michelson interferometer. 
-%In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-%We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings. 
-%However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
 \end{abstract}
@@
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% I feel like we can just delete everything that follows. It is in earlier commits and distributed throughout the subfiles.
-
-\begin{comment}
-
-% cover page
-\section*{Project contributions}
-% contributions clarify what’s to be assessed
-James W. Gardner:
-\begin{itemize}
-\item all paper sections other than the introduction, i.e.\ the abstract and from Section~\ref{sec:ifo} onwards, and all figures in the paper
-\item addition of the injection speaker, design of photodiode circuit from standard examples, assembly of circuit on breadboard, and set-up of Raspberry Pi to measure output
-\item all of the source code at \url{https://github.com/daccordeon/gravexplain} (containing over 1000 lines of Python scripts), that reads the data from the Raspberry Pi, performs all of the analysis on it, and creates all the figures used in the paper
-\item presentation of two talks based on the paper and the figures therein
-\end{itemize}
-
-Hannah Middleton:
-\begin{itemize}
-\item the introduction of the paper, i.e.\ Section~\ref{sec:introduction}
-\item design and assembly of the interferometer
-\item continuous support and advice throughout the project
-\end{itemize}
-
-Andrew Melatos, Robin Evans, William Moran: other vital advice and comments throughout the project
-
-\newpage
-\section*{Statement of achievements}
-% one paragraph recount of what’s been done for two months, like the abstract
-The project spanned three weeks in December, four weeks in January, and two weeks in February.
-The first two weeks were spent writing the initial scripts and achieving the Viterbi results with the webcam set-up.
-This was relatively straight forward and the greatest hurdle was understanding the algorithm.
-
-The third week was spent setting up the photodiode circuit and the Raspberry Pi which involved constant testing and changing of the components and design of the circuit.
-All of January was spent on the optical microphone.
-Much difficultly was had combating 50Hz mains hum in the experiment.
-Eventually through many discussions at the signal processing meetings, we had a working optical microphone but low speech intelligibility.
-The last week of January was spent on methods of speech enhancement.
-If the investigation had continued we had much future work planned and possible solutions in mind.
-
-Finally, the start of February was spent writing the paper and giving presentations (one nine minute general talk and one 50 minute technical follow-up talk).
-
-\newpage
-
-\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
-
-\author{James W. Gardner}
-\email{u6069809@anu.edu.au}
-\affiliation{%
-College of Science, Australian National University, Acton, ACT, 2601, Australia
-}%
-
-\author{Hannah Middleton}
-\email{hannah.middleton@unimelb.edu.au}
-\affiliation{%
- School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
-}
-
-\author{Andrew Melatos}
-\email{amelatos@unimelb.edu.au}
-\affiliation{%
- School of Physics, University of Melbourne, Parkville, Vic, 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
-}
-
-\author{Robin Evans}
-\email{robinje@unimelb.edu.au}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery
-}
-
-\author{William Moran}
-\email{wmoran@unimelb.edu.au}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
-
-% Summer 2019/2020
-\date{\today}
-
-% AJP requires an abstract for all regular article submissions.
-\begin{abstract}
-Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we used the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-We also describe how the inclusion of a photodiode and Raspberry Pi allowed the interferometer to act as an `optical microphone' with sufficient fidelity to play back simple audio recordings.
-However, more complex recordings, of speech, remain unintelligible when played back, despite preliminary efforts made towards speech enhancement.
-
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-\begin{figure}
-	\includegraphics[width=0.9\textwidth]{figures/ligo_pic.pdf}
-	\caption{Aerial photo of the LIGO Hanford facility and its 4km long beam arms, courtesy of LIGO/Caltech/MIT}
-	\label{fig:ligo_pic}
-\end{figure}
-
-% general gw intro
-In 2015, gravitational waves were observed for the first time from the merger of two black holes in a binary system~\cite{GW150914}. 
-The observation, made by the Laser Interferometer Gravitational wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
-Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-In recent years, there has been increased public interest in gravitational wave science. 
-Many gravitational wave research groups around the world have produced demonstrations and activities to explain the topic to a general audience.
-Activities range from hands-on demonstrations, exhibitions, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online}, video games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
-
-% what are gws and how do we detect them...
-Gravitational waves are a prediction of Einstein's theory of general relativity. 
-They are disturbances in space-time caused by the acceleration of massive objects. 
-The effect of gravitational waves is a change in lengths; a `stretching and squeezing' of the very distance between objects.
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance.
-As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
-These observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,ExhibitIFO,LIGOIFOGlue,LIGOIFOMagnets}.
-
-% continuous gws
-To date, while the network of gravitational wave observatories have only observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}, these observatories are also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencies.
-These continuous waves may emit at constant frequencies, or they may wander slowly in frequency over time. 
-One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
-LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
-The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{2019PhRvD.100l2002A,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
-
-% sound analogy
-Here, we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience; the analogy of listening to the universe is often used to explain gravitational wave observations.
- %; transient gravitational wave signals have been converted into audio.
-This can be demonstrated with a table-top interferometer by playing sound close enough to move the mirrors and change the interference pattern.
-Using audio of slowly wandering frequency tones to produce an analogue of a continuous gravitational wave signal.
-The analysis techniques used in this work are similar to those used to search for continuous gravitational waves in LIGO and Virgo data~\cite{2019PhRvD.100l2002A,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}, particularly for the wandering frequency demonstration.
-
-In this work we detail our table-top interferometer design in Section~\ref{sec:ifo} and describe observing a single note from a speaker in Section~\ref{sec:single_tone}.
-In Section~\ref{sec:viterbi_wandering}, we demonstrate the analysis techniques for observing a wandering frequency signal.
-In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech.
-This demonstration of an `optical microphone' serves as a more general exhibition of signal processing.
-We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
-
-% \section{Method and results}
-\section{Table-top gravitational wave science}
-\label{sec:ifo}
-
-Even though gravitational waves are caused by some of the most massive events in the universe, their observable effects on Earth are vanishingly small | on the order of only one part in $10^{21}$ for our best detectors~\cite{GW150914}.
-Clearly, the measurement of such effects is far beyond any table-top experiment.
-But, the fundamental principle of gravitational wave detection remains that changes in the lengths of the beam arms of an interferometer leads to a change in the interference pattern observed.
-By observing sound waves that vibrate the mirrors of a table-top interferometer and so change the lengths of the beam arms we have an analogue of this principle.
-
-\subsection{Method of Michelson interferometry to observe sound}
-% hardware: path difference, virtual thin film interference
-
-In a Michelson interferometer, a coherent (monochromatic and equal phase) source of laser light is split evenly by a beamsplitter into two arms at right-angles to each other, travelling down and back from a mirror at the end of each arm.
-The beams recombine and interfere at the beamsplitter before hitting a screen to produce a pattern of concentric, circular fringes.
-The configuration used in this experiment is shown diagrammatically in Figure~\ref{fig:ifo_schematic_webcam} and as part of a photo in Figure~\ref{fig:setup_pic2}.
-
-A sample interference pattern is shown in Figure~\ref{fig:interference_pattern}.
-This pattern is due to the path difference between the two arms of unequal lengths (if the lengths were equal, then only a single, bright dot would be seen).
-This results in a phase difference between the two beams when they meet at the screen.
-This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other.
-For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric.
-When the path difference changes, the fringes stay circular but move radially in or out.
-
-\begin{figure}[h]
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_webcam.pdf}
-	\caption{Michelson interferometer schematic with interference pattern recorded by a commercial webcam.
-We used a Class 2 532nm laser~\cite{ThorLabsIFO} and a commercial speaker.
-The beam is focused at a converging lens before the 50/50 beamsplitter and the interference pattern is made visible by a diverging lens after the beamsplitter.
-The arms lengths between the speakers are roughly 3cm different and the entire interferometer (without the screen) fits within the dimensions of an A4 piece of paper.
-We did not include a compensating plate with the beamsplitter, as is often done.}
-	\label{fig:ifo_schematic_webcam}
-\end{figure}
-
-We used a 0.5W speaker stuck to the back of one of the mirrors as an audio source for our demonstration.
-This speaker was part of a commercial 3.5mm-jack, USB powered pair of speakers and was driven by a PC.
-To attach the speaker to one of the mirrors, we wedged it into part of the mirror mount and used commercial adhesive putty to keep it in place.
-We kept the other speaker in the pair face-down and as far away from the experiment as the cables would allow to reduce it acting as a second source.
-
-As the speaker deforms to play sound, the mirror moves with a change in separation as some coupling function of the amplitude and frequency of the sound.
-The motion of the fringes follows this motion of the mirror and therefore the speaker at the same frequency.
-If the fringe motion is small, then changes in the intensity at some point on the screen follow the frequency of the injected audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
-If the fringe motion was large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting would artificially raise the measured frequency.
-As such, any motion of the fringes must be kept small by playing sound softly through the speaker.
-
-For later sections, we derive the non-linear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen, this can be found in Appendix~\ref{app:intensity_derivation}.
-
-% Note that we do not include the common compensating plate that equalises the amount of time spent in glass for each beam.
-We did not find it necessary to include it to be able to demonstrate the results desired.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/webcam_still0.pdf}
-	\caption{Interference pattern from Michelson interferometer as viewed by webcam from off-centre (hence why the pattern appears non-circular), red dot indicates point where intensity time series was measured from.
-For scale, the central fringe is approximately 5mm across}
-	\label{fig:interference_pattern}
-\end{figure}
-
-% A note that while the interference pattern appears similar to the Airy pattern produced by a circular aperture, they are not in fact the same geometry.
-
-\section{Observing a wandering tone}
-\label{sec:viterbi_wandering}
-% Tracking tones with the Viterbi algorithm
-% explaination and successful application
-% lean on the continuous gravitational waves angle
-
-As also mentioned in Section~\ref{sec:introduction}, continuous wave searches often look for signals that wander slowly in frequency~\cite{2019PhRvD.100l2002A}.
-The audio analogue of this is a tone that wanders in frequency; a note that changes pitch.
-Unlike the above Section~\ref{sec:single_tone} though, finding the signal isn’t as simple as searching for a peak in the spectrum for the full observing run.
-
-One method (among many) for continuous wave searches is to split the observed signal into intervals of time, compute the Fourier transform for each time interval, and then create a grid (known as a spectrogram) where each cell is the Fourier amplitude of a particular frequency at a particular time.
-This method turns the problem into finding the best path through the spectrogram grid as if it were a weighted graph, where here the weights are the Fourier amplitudes.
-
-When finding the path, the method requires a limit for the allowed frequency change over time, usually informed by a physical model.
-Here, this frequency wander limit was set to 0.11 Hz per second to make it simply one frequency bin per every time bin, which will prove poorly adjusted to the signal.
-
-\subsection{The Viterbi algorithm}
-
-The Viterbi algorithm is a path-finding algorithm that is used in continuous wave searches~\cite{viterbi_application} to solve this problem.
-Abstractly speaking, the algorithm is given a weighted graph (e.g.\ the spectrogram grid), a sensible sequence of subgraphs (e.g.\ the columns of the spectrum at each time), and restrictions on connectivity (e.g.\ the allowed amount frequency wander).
-Then, at each iteration, it will find the best path to each node in the next subgraph, all the way from the first subgraph.
-At the end, it selects the overall best path from the first to the last subgraph (here, from the start to end time), this overall best path is called the Viterbi path.
-See Appendix~\ref{app:viterbi} for the full details of the algorithm in this particular implementation.
-
-\subsection{Results}
-
-We demonstrate this method of using the Viterbi algorithm by injecting a wandering frequency signal into the speaker and recording the output via the webcam.
-Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{figures/expt_overlay_2_viterbi_test_webcam.pdf}
-	\caption{Spectrogram of observed signal overlaid with pink dots for the injected signal and white crosses for the Viterbi path.
-Note the high correlation between them.
-Also note the deviations at the start time when the signal moves too quickly and just before the background anomaly at 150s.}
-	\label{fig:viterbi_overlay}
-\end{figure}
-
-Shown in Figure~\ref{fig:viterbi_overlay} is the resultant spectrogram of the observed signal, overlaid with the injected signal (the pink dots) and the recovered Viterbi path (the white crosses).
-To the eye, the Viterbi algorithm easily recovers the injected frequency over time due to the high signal-to-noise ratio (SNR) of this experiment.
-Quantifying this, the Viterbi path stays within one frequency bin (approximately 0.3Hz) of the injected tone for 94\% of the run.
-Note that the signal initially moves faster than the frequency wander limit, causing a discrepancy between the injected and recovered signals.
-To fix this, the limit should be doubled, although this would then cause the recovered path to stray more often.
-There also appears to be an anomaly at 150s, likely from a disturbance (such as someone walking) around the interferometer.
-
-Empirically, decreasing the signal-to-noise ratio closer to unity (by making the speaker quieter) causes the Viterbi algorithm to stray more from the signal.
-However, since the only correlation through the grid is the signal, we always expect the Viterbi path to lie near to the signal no matter how high the noise is, and this is indeed what we see.
-
-\section{Playing back complex audio}
-\label{sec:optical_microphone}
-% Optical microphone
-% aim of optical microphone: play back injected sounds
-
-The natural succession to observing simple tones is to study complex audio signals, such as music and speech.
-The Viterbi algorithm and tracking more broadly is not applicable to these signals.
-Compare a single tone that slowly, continuously changes frequency to music and speech which consist of a broad spectrum that is constantly changing.
-Rather than track them, the interest here is in being able to record and play-back these complex audio signals.
-Using the interferometer to do so makes it an ‘optical microphone’, using light to capture sound.
-This device has precedence in the laser microphones~\cite{laser_microphone} of the defence industry which operate on a variety of related principles.
-
-Although this is indirectly related to the advanced signal processing required to extract gravitational wave signals, using an interferometer to capture sound can serve as an interesting demonstration to a broader physics and engineering audience.
-As such this could be introduced as a separate demonstration.
-
-\subsection{The photodiode method}
-\label{sec:photodiode}
-% advanced method: explain how we capture data
-
-For the optical microphone to be sensitive to speech and music it needs a frequency response (a sampling rate) far higher than the 30Hz of the webcam.
-In the context of audible frequencies (roughly 20Hz to 20kHz), speech intelligibility requires frequencies up to 3kHz and music requires frequencies up to and beyond 8kHz~\cite{speech_intelligibility}.
-Therefore the sampling rate of the optical microphone must be at least 16kHz to capture both speech and music.
-We achieved this with a photodiode.
-
-A photodiode is an electrical component that acts as a regular diode, blocking any current flow in the `reverse' direction, when no light is incident on it.
-But as the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.
-% Fundamentally, a current is created in the photodiode by the photo-electric effect.
-
-We placed a (OSRAM BPW21) photodiode in reverse-bias over an (LM358) op-amp to create a photo-detector that would produce a voltage dependent on the incident intensity.
-We placed the photodiode in the interference pattern at roughly the same off-centre spot as the webcam was monitoring previously (although we saw no difference when monitoring at other positions).
-The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.
-
-The voltage signal from the photo-detector was captured by a (MCP3008) 10-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2.
-Together, these sampled the signal at roughly 16kHz, our desired limit.
-
-However, by sampling, any frequency component of the analog signal above the Nyquist frequency of 8kHz would have been aliased down into the detected range.
-To prevent this, we included an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to 16kHz before the ADC.
-This component attenuated any frequencies above 8kHz, before they got digitally sampled.
-We also had to place another cloth screen over the face of the photodiode, as the reading without it, after gaining through the op-amps, was too high for the ADC.
-% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-
-A diagram and a photo of the circuit are shown in Appendix~\ref{app:circuit_diagram}.
-The updated schematic is shown in Figure~\ref{fig:ifo_schematic_podo} and a picture taken of the entire optical microphone is shown in Figure~\ref{fig:setup_pic2}.
-
-\begin{figure}
-	\includegraphics[width=0.8\textwidth]{figures/ifo_schematic_photodiode.pdf}
-	\caption{Michelson interferometer schematic with the intensity of the interference pattern at a point measured by a photodiode.
-See Figure~\ref{fig:ifo_schematic_webcam} for details on the optics configuration.}
-	\label{fig:ifo_schematic_podo}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{figures/setup_pic2.pdf}
-	\caption{Picture taken of the optical microphone set-up.
-Showing the Michelson interferometer, the photodiode (behind the cloth screen and in an insert), the circuit, and the Raspberry Pi.
-See Appendix~\ref{app:circuit_diagram} for a diagram and closer photo of the circuit.}
-	\label{fig:setup_pic2}
-\end{figure}
-
-\subsection{Initial results of the optical microphone}
-% breakdown of all the filters used along the way
-
-The optical microphone was tested on varied recordings of speech from different people and on music ranging from simple melodies to full track songs.
-Each recording was around one minute long with processing done on ten second segments.
-For each observation, the audio was played through the speaker as the Pi script recorded the signal from the photodiode.
-Care was taken to have minimal activity around interferometer while this was happening.
-The time series was then directly converted to a .wav file and played as an audio recording, using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.
-
-When played back, these initial recordings were extremely noisy and had a loud bass hum throughout.
-Each Fourier spectrum showed dominant AC mains noise with power ranging from the fundamental of 50Hz up to and beyond the 8th harmonic thereof.
-With the speaker off this mains power was still present, as shown the power spectral density (effectively the norm squared of the Fourier spectrum) in Figure~\ref{fig:psd_noise}.
-And even with the speaker off and the photodiode in complete darkness, the mains hum was still present (with distribution similar to but weaker than that shown in Figure~\ref{fig:psd_noise}).
-
-\begin{figure}[h]%[H]
-	\includegraphics[width=\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
-	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off), shows strong power in the 50Hz mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and even cooling), but otherwise is a fairly white spectrum.}
-	% except for the broad, shallow maximum at 750Hz}
-	\label{fig:psd_noise}
-\end{figure}
-
-\subsection{Applying filters to the data}
-\label{sec:simple_filters}
-
-The first attempt made to remove the mains hum was to zero the channel of each 50Hz harmonic in the Fourier spectrum.
-Effectively, this multiplied the spectrum by a rectangular comb filter.
-While this did remove the mains harmonics, it also audibly ruined the rest of the signal.
-This is because applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
-And the inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function that when convolved with significantly damages the signal.
-
-To fix this, we tried using a smoother comb filter but this too significantly attenuated the rest of the spectrum between each harmonic to hear the signal at all.
-
-Applying a high-pass filter with cut-off frequency 150Hz to the signal worked well at removing the 50Hz and 100Hz tones but was problematic as the mains harmonics extend far past 100Hz and so remain after filtering.
-This region above 100Hz carries a lot of the fundamental frequencies for speech and music~\cite{speech_intelligibility} and if the cut-off frequency is placed as to remove more of the mains harmonics, then the resultant signal is unrecognisable.
-
-We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass.
-
-\begin{figure}[h]%[H]
-	\includegraphics[width=0.8\textwidth]{figures/butterworth_150_3000-cropped.pdf}
-	\caption{Butterworth bandpass filter frequency response, any amplitude beyond -3dB is significant attenuation (half power).
-The green vertical lines show the band limits of 150Hz and 3kHz.
-Note the flat response within the band.}
-	\label{fig:butterworth}
-\end{figure}
-
-Band-passing the signal similarly doesn’t address the remaining harmonics above 100Hz but does low-pass away other noise at higher frequencies.
-The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over a band from 150Hz to 3kHz (beyond 2kHz speech information content drops significantly~\cite{speech_intelligibility}).
-For a low-pass Butterworth filter at cut-off frequency $\omega_c$ and gain $\epsilon$ the frequency response (attenuation at each frequency) follows Equation~\ref{eq:butterworth}.
-\begin{equation}
-\label{eq:butterworth}
-H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\frac{\omega}{\omega_c})^{2n}}}
-\end{equation}
-For a Butterworth band-pass filter, the above low-pass filter is combined with a high-pass filter of similar form.
-The response of the filter used is shown in Figure~\ref{fig:butterworth}.
-This filter improved on the above simpler filters but recordings of speech still had very low intelligibility, they could be recognised as speech but not understood.
-
-The ideal filter would be one that completely attenuates the undesired parts of the spectrum, does not change the rest of the spectrum, and has smooth edges as to not damage the signal through convolution.
-These three conditions are contradictory and any filter must compromise between them.
-The above simple filters fail at one or more of each of these conditions.
-The combs ruin the signal and the high-passes do not remove the noise.
-The Butterworth filter performed the best audibly as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
-
-% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
-Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-
-\subsection{The logMMSE estimator}
-% logMMSE
-
-Speech enhancement for a noisy channel is a classic problem in signal processing.
-Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compared various methods and found the best performing statistical technique among those at the time to be a logMMSE (alt.\ MMSE-LSA) estimator.
-This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
-Although this technique historically minimised with respect to the spectral amplitude~\cite{EphraimSpeechEU_MMSE}, throughout speech processing use of the logarithm seems produce better results as perceived by the human ear~\cite{SubjectiveComparison}.
-
-We used an existing implementation~\cite{logmmse} of the logMMSE estimator.
-Applying this estimator to the optical microphone recordings produced dramatically clean results.
-The time series (in Figure~\ref{fig:logMMSE_timeseries}) and spectrum (in Figure~\ref{fig:logMMSE_spectrum}) from before and after applying the logMMSE filter are shown.
-Comparing the before and after spectra in Figure~\ref{fig:logMMSE_spectrum} shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.
-
-After applying the estimator, simple chords and drums are heard well in played back music.
-While more complex melodies are sometimes completely absent.
-By observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
-This could be a perceptual effect of listening to the lossy recording, but seems more likely to be a frequency dependence somewhere in the optical microphone system.
-Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies and so instruments like electric bass and drums are louder in the results.
-
-After applying the estimator to speech recordings, intelligibility still remains low with the voice sounding mumbled and indistinct.
-The estimator removes most of the background noise, but does not clarify the speech.
-
-To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-This requires a better understanding of the system and further improvements to the experiment if true.
-% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
-However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
-An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_timeseries_aa_melatos-cropped.pdf}
-	\caption{Optical microphone recording of adult male voice (saying ``A cathode ...'') shown before (top) and after (bottom) the logMMSE estimator is applied, showing one second of the one minute recording.
-The rise at the start of the bottom	 time series is an expected result from filtering a finite signal.}
-	\label{fig:logMMSE_timeseries}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/filter_spectrum_aa_melatos-cropped.pdf}
-	\caption{Spectrum of optical microphone recording of adult male voice shown before (top) and after (bottom) the logMMSE estimator is applied, showing only up to 2kHz and to amplitude 2 to see greater detail.}
-	\label{fig:logMMSE_spectrum}
-\end{figure}
-
-\section{Future work}
-\label{sec:future_work}
-% wild speculation
-We are currently working towards two future advancements for the optical microphone.
-
-The first would be to better isolate the experiment from the mains supply.
-The Raspberry Pi is USB powered and requires low power so it could even be run off a commercial external phone battery.
-Everything else in the circuit, including the photodiode, is connected to the Pi, so this would isolate them electrically as well.
-The laser is DC powered through a converter, so there is no issue there.
-With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
-
-The second advancement would be to achieve a full model of the transfer function of the system.
-Theoretically, if we had the correct transfer function, then signal recovery would only be limited by the signal-to-noise ratio as we could construct an optimal matched filter.
-This function would have to start from the voltage sent to the speaker and end with the value recorded by the Pi, and include any non-linearities in the speaker audio production, the speaker-mirror coupling, the separation-intensity relation, and the photodiode reading.
-This sequence that the signal takes is shown in Figure~\ref{fig:pipeline_highlighted}.
-
-In our investigation, we have currently only derived part of this sequence, the relation between a deflection of the mirror and the change in the intensity of the interference pattern at some point.
-We find this relation to be non-linear, see Appendix~\ref{app:intensity_derivation} for the derivation.
-
-Asides from the transfer function, for extracting the speech there exists much literature~\cite{HMM_english} on using a hidden Markov model train to English phonemes to recognise speech.
-Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
-
-To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
-This could be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
-
-\begin{figure}%[H]
-	\includegraphics[width=\textwidth]{figures/pipeline_highlighted.pdf}
-	\caption{Sequence of relations for optical microphone transfer function. The currently derived relation is highlighted.}
-	\label{fig:pipeline_highlighted}
-\end{figure}
-
-\newpage
-\section{Conclusions}
-\label{sec:conclusions}
-
-% Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
-
-We use sound and a table-top Michelson interferometer as an analogue to a gravitational wave detector.
-% This serves to demonstrate signal processing techniques.
-
-Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer.
-We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.
-
-Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the interferometer into an optical microphone.
-We demonstrated that this optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech.
-Dominant 50Hz mains hum in the recordings led to an exploration of speech enhancement techniques to remove it, including the well performing logMMSE estimate.
-From here we have many plans to further remove the mains hum and improve the intelligibility.
-
-These demonstrations have good potential to be adapted into the undergraduate laboratory.
-There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches.
-Most of all, the signal processing of speech enhancement is an interesting problem that’s relevant to digital communications.
-
-% \bigskip
-\newpage
-\appendix
-
-\section{Open-source code}
-\label{app:code}
-This project was implemented in python3~\cite{python} scripts as well as jupyter notebooks~\cite{jupyter}~\cite{ipython} and used the packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
-
-The current build and sample recordings of the optical microphone can be found at:
-\url{https://github.com/daccordeon/gravexplain}
-% \begin{verbatim}
-% README
-% \end{verbatim}
-
-\section{Implementation of the Viterbi algorithm}
-\label{app:viterbi}
-
-We used the Viterbi algorithm to find a signal in a spectrogram. Here we detail the implementation used.
-
-We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
-These are normalised between $(0, 1)$ to use as multiplicative weights.
-In fact, we take the logarithm and so the sum of all weights to avoid underflow.
-The best path has the highest weight among all paths.
-
-With the grid prepared, the algorithm starts in the second column.
-In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, selecting the previous cell with the greatest cumulative weight among those ‘nearby’ (the value of which was discovered in the previous iteration).
-Where the cumulative weight is the weight of the best path to reach that previous cell.
-The nearness of a previous cell is given by the limit on the allowed frequency change over time.
-
-For each cell in the current column, the algorithm calculates the cumulative weight to reach that cell by adding its weight to the cumulative weight to reach it.
-The algorithm also notes the row index of which cell in the previous column it best connects to.
-Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
-That discovered best path is the Viterbi path.
-
-\section{A common mistake for those new to wandering frequency signals}
-\label{app:phase_gotcha}
-
-We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
-A common mistake is to appeal to the standard form $\sin{(2 \pi f t)}$, where f is a constant frequency, and guess $\sin{(2 \pi f(t) t)}$.
-However, this fails to address the phase accumulated earlier in the signal.
-Instead, the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
-Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
-% This formula is also useful when performing frequency modulation.
-
-\section{Photodiode circuit design}
-\label{app:circuit_diagram}
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-
-Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
-This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.
-
-\begin{figure}[h]%[H]
-	\includegraphics[width=\textwidth]{figures/circuit_diagram_2.pdf}
-	\caption{Circuit diagram for photodiode reading.
-Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC).
-The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.}
-	\label{fig:circuit_diagram}
-\end{figure}
-
-\begin{figure}%[H]
-	\includegraphics[width=0.8\textwidth]{figures/circuit_pic3.pdf}
-	\caption{Photo of photodiode circuit assembled on breadboard.}
-	\label{fig:circuit_pic3}
-\end{figure}
-
-\section{Interferometric intensity relation derivation}
-\label{app:intensity_derivation}
-
-To obtain part of the transfer function for the optical microphone, we want to know the resultant change in intensity, $\delta I$, at a point in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
-Let the point be at angle $\theta$ on the screen (in the far-field) above the central ray as measured from the mirrors.
-Noting circular symmetry as to not specify exactly where in the fringe the point is.
-
-At this point, two rays meet, one from each mirror, with total electric field amplitude $E_1(t) + E_2(t)$.
-Given an initial difference between the arm lengths of $d$, the path difference between the two beams at the screen is $2 d \cos{\theta}$ as the longer beam must travel back and forth over the extra distance, and both travel at angle $\theta$ to meet the point~\cite{fringes:online}.
-
-The intensity at the point is then the norm squared of the total electric field amplitude as expressed in Equation~\ref{eq:intensity_derivation} below.
-
-\begin{align}
-    E_1(t) &= A e^{i (k L - \omega t)} \\
-    E_2(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \\
-    I(t) &= \lvert A e^{i (k L - \omega t)} + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2
-\label{eq:intensity_derivation}    
-\end{align}
-
-Collapse the phase difference to some value $\phi = 2 k d \cos{\theta}$.
-Then expand the norm squared using Euler’s relation, $e^{i \theta} = \cos{\theta} + i \sin{\theta}$.
-Here we stop considering intensity as a function of time as the time dependence will vanish.
-
-\begin{align}    
-    I &= A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
-    I &= A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 + (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2)
-\end{align}
-
-Expand and collect like terms, pull out a factor of two, and then recognise the form of $\cos{(\alpha-\beta)}$.
-
-\begin{align}
-    I &= A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} + 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
-    I &= 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-\end{align}
-
-Finally, changing derivatives and expand $\phi$ to find that the change in intensity given a change in distance is non-linear, that ${\delta I}/{\delta d}$ is non-constant.
-
-\begin{align}    
-    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
-    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
-    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
-    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-\end{align}
-
-This non-linearity in part of the transfer function will make the reverse problem of extracting the original signal more difficult.
-
-\begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project.
-
-The authors are also grateful to Deeksha Beniwal, Sebastian Ng, and Craig Ingram for their advice and work in designing the interferometer.
-
-This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004) and the Institute of Physics International Member Grant.
-
-Travel during the project was supported by the Australian National University PhB Science  program.
-
-\end{acknowledgments}
-
-\bibliographystyle{myunsrt}
-\bibliography{ifoDemoBib}
-
-\end{document}
-
-\end{comment}

### 2020-08-28 paper/ifo-complexAudio.tex
editing the future work and conclusions

@@
 \end{equation}
-where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 
+where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ are the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 
 Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the Wiener filter amplifies the input signal where signal to noise ratio (SNR) is high, and weakens the signal where SNR is low.
@@
 A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
-The choice of $n=100$ is a reasonably selection to balance smoothing and efficiency.
+The choice of $n=100$ is a reasonable selection to balance smoothing and efficiency.
 
-The timeseries and frequency spectrum results of the Wiener fileter are are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
+The timeseries and frequency spectrum results of the Wiener filter are are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
 There is improvement which is particularly visible in the timeseries, however a strong noise hum persists. 
@@
 Speech enhancement for a noisy channel is a classic problem in signal processing. 
-Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
-This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
+In Ref.~\cite{SubjectiveComparison}, $13$ speech enhancement methods are compared and the best found to be a log minimum means-square error (logMMSE) estimator. 
+This type of estimator is based on work in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
+%Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
+%This type of estimator is based on work Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
 The logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
+We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}).
 
-In this work we use an existing implementation~\cite{logmmse} of the logMMSE estimator. 
-Applying this estimator to the optical microphone output produced dramatically cleaner results.
+%In this work we use an existing implementation~\cite{logmmse} of the logMMSE estimator. 
+%Applying this estimator to the optical microphone output produced dramatically cleaner results.
 
-In Figures~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} we present results from playing a speech recording through the optical microphone. 
-Figure~\ref{fig:logMMSE_timeseries} shows the anti-aliased photodiode time series from before (top panel) and after (bottom panel) applying the logMMSE filter. 
-Figure~\ref{fig:logMMSE_spectrum} shows the corresponding spectra before (top panel) and after (bottom panel) the logMMSE filter. 
-Comparing the before and after spectra shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.
+The timeseries and frequency results of the logMMSE are shown in Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} respectively. 
+The top two panels of each figure again show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison. 
+The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} show the timeseries and corresponding spectra after applying the logMMSE filter.
 
@@
 
-We also find improvement with music recordings.
-After applying the estimator, simple chords and drums can be heard clearly. 
-However more complex melodies are sometimes completely absent. 
+The results show significant attenuation of the mains harmonics as well as a general smoothing of the spectrum.
+However, we find that the intelligibility of speech remains low with the voice sounding mumbled and indistinct. 
+The logMMSE removes most of the background noise, but does not clarify the speech.
+
+We find some improvement with music recordings. 
+After applying the logMMSE, simple chords and drums can be heard clearly, however more complex melodies are sometimes completely absent. 
 From observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
@@
 
-Overall, we find that the intelligibility of speech remains low. 
-The played-back voice sounds mumbled and indistinct. 
-The estimator removes most of the background noise, but does not clarify the speech. 
-To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-This requires a better understanding of the system.
+To address these problems we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
+This requires a better understanding of the system as discussed in Section~\ref{sec:future_work}.
+
+%Overall, we find that the intelligibility of speech remains low. 
+%The played-back voice sounds mumbled and indistinct. 
+%The estimator removes most of the background noise, but does not clarify the speech. 
+%To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
+%This requires a better understanding of the system.
 % The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.

### 2020-08-28 paper/ifo-conclusions.tex
editing the future work and conclusions

@@
 
-In this paper we present an educational demonstration in which a table top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
+In this paper we present an educational demonstration in which a table-top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
 Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
@@
 Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
-We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 
+We demonstrate that the optical microphone is capable of playing back simple recordings well but produces unintelligible recordings of speech. 
 Mains hum at $50\,{\rm Hz}$ and its harmonics dominates the photodiode output. 
-This offers students the opportunity to explore speech enhancement techniques to remove it, including the well performing logMMSE estimator. 
+This demonstration offers students the opportunity to explore a variety of filtering and speech enhancement techniques. 
 Further avenues of exploration include a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.
@@
 It can be used to demonstrate topics of interest to physicists and electrical engineers including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement.
-The detection of gravitational waves in 2015 led to increased excitement and public interest in gravitational wave science. 
-The table-top demonstration presented here can also be adapted as a tool to explain gravitational wave research developments for a wider non-specialist audience. 
+Given the increased excitement and public interest in gravitational-wave science since the detection in 2015, this table-top demonstration presented here can also be adapted as a tool to explain gravitational-wave research developments for a wider non-specialist audience.
 The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves.

### 2020-08-28 paper/ifo-futureWork.tex
editing the future work and conclusions

@@
 An initial improvement is to better isolate the experiment from interference, such as the mains supply. 
-The Raspberry Pi is USB powered and requires low power so it could be run off a commercial external phone battery. 
+The Raspberry Pi is USB powered and requires low power so it can instead be powered by a commercial external phone battery rather than the mains. 
 Everything else in the circuit, including the photodiode, is connected to the Pi. 
 The laser is DC powered through a converter, so should not introduce mains harmonics itself. 
-Transferring the circuit from a breadboard to a printed circuit board will also reduce interference~\cite{elfekey2013design}
+Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}
 
@@
 
-A second improvement is to measure the full model of the transfer function of the system, providing more information about how the audio signal couples with the interferometer.
+A second improvement is to measure the full model of the transfer function of the system, providing information about how the audio signal couples with the interferometer.
 % commented out whilst James is checking this
@@
 It includes any non-linearities in the speaker itself, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode reading. 
-A schematic of the full optical microphone transfer function is shown in Figure~\ref{fig:pipeline_highlighted}.
+A schematic of the full optical microphone transfer function is shown in Fig.~\ref{fig:pipeline_highlighted}.
 
 A voluminous literature exists about using a hidden Markov model (HMM) trained against phonemes to recognize speech~\cite{HMM_english}. 
-Interestingly, this connects back to the Viterbi algorithm in Section IV, which is also underpinned by a hidden Markov model, albeit of a different type. 
+Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type. 
 Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
 
-Other traditional signal processing methods for extracting spectrogram like wavelet transform (WT) \citep{nason1995stationary} can also be implemented and compared with HMM.
+Other traditional signal processing methods for extracting spectrogram-like wavelet transform (WT) \citep{nason1995stationary} can also be implemented and compared with HMM. 
 WT provides both time and frequency information, making it easier to pinpoint the origin of noise respect to time. 
-In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT was also proposed for speech recognition.
+In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT is also proposed for speech recognition.
 

### 2020-08-28 paper/paper_main.tex
editing the future work and conclusions

@@
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
-This work also received financial support from the Institute of Physics International Member Grant and travel support from the Australian National University PhB Science program.
+This work also received financial support for hardware from the Institute of Physics International Member Grant and for travel from the Australian National University PhB Science program.
 

### 2020-09-09 paper/ifo-introduction.tex

--- 2020-07-01 paper/ifo-introduction.tex
@@
 % general gw intro
-In 2015, gravitational waves were observed for the first time from the merger of two black holes~\cite{GW150914}. 
-The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
-Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-Gravitational-wave detection has led to increased public and media interest in gravitational-wave science. 
-Gravitational-wave research groups around the world have produced demonstrations and activities to explain their research to general audiences. 
-Activities range from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}%\han{look up Sarah Farmer's project}.
-
+In 2015, the first detection of gravitational waves from the merger of two black holes~\cite{GW150914} by the Advanced Laser Interferometer Gravitational-wave Observatory~(aLIGO \citep{AdvancedLIGO:2015}) marked a breakthrough in modern astrophysics and opened up a new window onto the universe.
 % what are gws and how do we detect them...
-Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}. 
-They are disturbances in spacetime caused by the acceleration of asymmetric massive objects. 
-The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects.
-Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers.
-They use the interference of laser light to measure changes in distance. 
-%As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
-Gravitational-wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
+Gravitational waves, predicted by Albert Einstein in his theory of general relativity~\cite{Einstein:1916}, are disturbances in `the fabric of' spacetime caused by the acceleration of massive objects.
+% technically incorrect, but do we need to mention antiaxial asymmetry here? 
+Observatories such as aLIGO, aVirgo, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} detect these waves as tiny changes in length: a `stretching and squeezing' of the distance between objects.
+In particular, they use the interference of light to measure changes in the lengths of two perpendicular arms, kilometres long.
 
 % continuous gws
-To date the network of gravitational wave observatories has observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals.
-Continuous waves may be emitted at constant frequencies, or may wander slowly in frequency over time. 
-For example, rotating neutron stars in low mass X-ray binaries (LMXBs) are one of the targets for continuous-wave searches. 
-LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion, where the rotation frequency of the compact object wanders over time due to variable accretion of matter (and hence angular momentum transfer) from the companion~\cite{xraybinaries:1997}.
-Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
-It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.
+To date, the network of gravitational-wave observatories has detected numerous signals from binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers, short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}.
+However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, but is as yet unsuccessful~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017,ScoX1ViterbiO1:2017,SearchRadiometerO1:2017,SearchCrossCorrO1:2017,MiddletonEtAlO2LMXBs:2020,SearchTwoSpecS6:2017}.
+Predicted sources for such continuous (gravitational) waves include rotating neutron stars in orbit around low mass stellar companions (e.g.\ low mass X-ray binaries, LMXBs).
+Continuous waves from such sources may be emitted at constant frequency or may wander slowly in frequency over time.
+For example, the rotation frequency of a neutron star in a LMXB can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}, and this rotation frequency is predicted to determine the frequency of the continuous waves.
+The prime target for continuous wave searches from LMXBs is Scorpius X-1, an LMXB particularly bright in X-rays.
 
+% table-top
+Gravitational-wave observatories are extremely complex but are fundamentally based on the classical Michelson interferometer. 
+Table-top Michelson interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
 % sound analogy
-In this paper, we present a table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves.
-As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to electronics, signal processing, and speech enhancement. 
-It gives students in courses such as physics and electrical engineering the freedom to explore the response of an accessible yet nontrivial opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
-The demonstration can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.
+In this paper, we use a table-top Michelson interferometer as a `toy' gravitational-wave observatory designed to detect sound instead of gravitational waves.
+That is, instead of detecting changes in the spacetime along the arms, we detect Newtonian motion of the mirrors driven by acoustic vibrations in the air around them.
+The design of the interferometer is detailed in Section~\ref{sec:ifo}.
+By playing different sounds through a speaker aimed at the interferometer, capturing the resultant interference pattern, and applying suitable signal processing, we can play-back the original signal and as such turn the interferometer into an `optical microphone', using light to capture sound.
+We observe a constant frequency signal in Section~\ref{sec:single_tone}, a wandering frequency signal with application of the Viterbi algorithm (a hidden Markov model technique) from continuous wave searches in Section~\ref{sec:viterbi_wandering}, and more complex audio, such as music and speech, in Section~\ref{sec:optical_microphone}, to varying degrees of success.
+The latter being less about applying gravitation wave techniques and more about exhibiting general signal processing.
+We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
+% Appendix~\ref{app:code} collects the software and scripts used to produce this work.
 
-The paper presents 
-In this work we present a selection of demonstrations using different types of audio signals ranging from simple tones, analogous to continuous wave signals, to more complex audio of music and speech.
-In Section~\ref{sec:ifo} we detail the table-top interferometer design. 
-We demonstrate observing a single note from a speaker in Section~\ref{sec:single_tone}.
-In Section~\ref{sec:viterbi_wandering} we replace the single note with a slowly changing frequency and analyse the signal using a hidden Markov model technique from continuous-wave searches. 
-In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech. 
-This demonstration of an `optical microphone' serves as a more general exhibition of signal processing. 
-We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
-Appendix~\ref{app:code} collects the software and scripts used to produce this work.
+% broader future applications
+This demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to electronics, signal processing, and speech enhancement. 
+Additionally, it allows students in physics and electrical engineering courses to explore the response of an accessible, yet nontrivial, opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, up to those used in the search for continuous waves in aLIGO and aVirgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
+
+Moreover, gravitational-wave detection has led to increased public and media interest in gravitational-wave science. 
+Research groups around the world have produced demonstrations and activities to explain their research to general audiences, with efforts ranging from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.%\han{look up Sarah Farmer's project}
+This demonstration can be adapted to an effective outreach tool to join those above in explaining gravitational wave science to a non-specialist audience.
 

### 2020-09-09 paper/paper_main.tex

@@
 Gravitational-wave observatories around the world are searching for continuous waves: persistent signals from spinning neutron stars. 
-Searches use sophisticated statistical techniques to look for weak signals in noisy data. 
-We show how continuous-wave searches can be demonstrated with a table-top model gravitational-wave detector (a Michelson interferometer), where sound plays the role of gravitational waves. 
+These searches use sophisticated statistical techniques to look for weak signals in noisy data. 
+In this paper, we demonstrate these techniques using a table-top model gravitational-wave detector (a Michelson interferometer), where sound plays the role of gravitational waves. 
 Using signal processing techniques from continuous-wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
-Finally we investigate the performance of the demonstration as an `optical microphone' for recovering music and speech. 
-Simple chords and drums are easily recovered, but speech is more challenging. 
-This demonstration has scope for use in the physics and electrical engineering undergraduate lab. 
+Finally we use the interferometer as an `optical microphone' to capture music and speech and apply filtering techniques to recover the original signals from the noisy data.
+Such techniques include simple bandpass and notch filters, as well as comparing a Weiner filter to the statistical logMMSE method.
+Simple chords and drums are easily recovered, but complex music and speech are more challenging. 
+This demonstration has scope for use in physics and electrical engineering undergraduate laboratories. 
 It can also be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences. 

### 2020-09-09 paper/ifo-introduction.tex

@@
 % continuous gws
-To date, the network of gravitational-wave observatories has detected numerous signals from binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers, short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}.
+To date, the network of gravitational-wave observatories has detected at least ten signals from binary black hole mergers~\cite{GW151226,GW170104,GW170814,GWTC-1:2018} and at least one from a binary neutron star merger~\cite{GW170817,GW170817multi,GW190425, GWTC-1:2018}, all short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, but is as yet unsuccessful~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017,ScoX1ViterbiO1:2017,SearchRadiometerO1:2017,SearchCrossCorrO1:2017,MiddletonEtAlO2LMXBs:2020,SearchTwoSpecS6:2017}.

### 2020-09-09 paper/paper_main.tex
affiliations: IPAS split from Department of Physics

@@
 \affiliation{%
-Department of Physics and The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
 }

### 2020-09-10 paper/ifo-introduction.tex
restructuring the introduction

--- 2020-08-16 paper/ifo-introduction.tex
@@
 
-% general gw intro
-In 2015, gravitational waves were observed for the first time from the merger of two black holes~\cite{GW150914}. 
-The observation, made by the Advanced Laser Interferometer Gravitational-wave Observatory~(Advanced LIGO \citep{AdvancedLIGO:2015}), marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
-Since 2015, the Advanced LIGO and Advanced Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers.
-Gravitational-wave detection has led to increased public and media interest in gravitational-wave science. 
+In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics an revealed a new means to observe the universe~\cite{GW150914}.
+This first observation and the many to follow it has led to increased public and media interest in gravitational-wave science. 
 Gravitational-wave research groups around the world have produced demonstrations and activities to explain their research to general audiences. 
-Activities range from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}%\han{look up Sarah Farmer's project}.
+Activities range from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
+%\han{look up Sarah Farmer's project}
+Here we describe the application of a table-top demonstration commonly used by gravitational-wave groups, to explain gravitational wave searches and signal processing analysis.
 
-% what are gws and how do we detect them...
+% What are gws and gw detectors 
 Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}. 
@@
 The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects. 
-Observatories such as LIGO, Virgo, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} are laser interferometers.
-They use the interference of laser light to measure changes in distance. 
-%As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
+Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
 Gravitational-wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
-Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
+Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
 
-% continuous gws
-To date the network of gravitational-wave observatories has observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
-However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals.
-Continuous waves may be emitted at constant frequencies, or may wander slowly in frequency over time. 
-For example, rotating neutron stars in low mass X-ray binaries (LMXBs) are one of the targets for continuous-wave searches. 
-LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion, where the rotation frequency of the compact object wanders over time due to variable accretion of matter (and hence angular momentum transfer) from the companion~\cite{xraybinaries:1997}.
-Scorpius X-1 is an LMXB which is particularly bright in X-rays making it a prime target for continuous wave searches from LMXBs. 
-Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017,ScoX1ViterbiO1:2017,SearchRadiometerO1:2017,SearchCrossCorrO1:2017,MiddletonEtAlO2LMXBs:2020,SearchTwoSpecS6:2017}.
+% CW 
+To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of 
+binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018}, 
+binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018} and recently an observation of a black hole with an object that could either have been a neutron star or a black hole~\cite{GW190814}.
+However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
+Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
+Rotating neutron stars are prime candidates for continuous wave emission. 
+An example target are rotating neutron stars in low mass X-ray binaries (LMXB), where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
+The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
+Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches from LMXBs. 
+Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.
 
-% sound analogy
-In this paper, we present a table-top experiment, where a standard Michelson interferometer plays the role of a `toy' gravitational-wave detector, which detects acoustic vibrations (such as music) instead of gravitational waves.
-As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to electronics, signal processing, and speech enhancement. 
-It gives students in courses such as physics and electrical engineering the freedom to explore the response of an accessible yet nontrivial opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
-The demonstration can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.
+In this paper, present a table-top experiment, where a standard Michelson interferometer acts as a `toy' gravitational-wave detector, which detects sound instead of gravitational waves. 
+The demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
+It allows students in courses such as physics and electrical engineering the freedom to explore the response of an accessible, yet nontrivial opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
+The demonstration also has potential to be adapted for use as an outreach tool to explain gravitational-wave science to a non-speacialist audience.
 
-In this work we present a selection of demonstrations using different types of audio signals ranging from simple tones, analogous to continuous wave signals, to more complex audio of music and speech.
-In Section~\ref{sec:ifo} we detail the table-top interferometer design. 
-We demonstrate observing a single note from a speaker in Section~\ref{sec:single_tone}.
-In Section~\ref{sec:viterbi_wandering} we replace the single note with a slowly changing frequency and analyse the signal using a hidden Markov model technique from continuous-wave searches. 
+The design of the interferometer is detailed in Section~\ref{sec:ifo} and we demonstrate observing a single tone from a speaker in Section~\ref{sec:single_tone}. 
+In Section~\ref{sec:viterbi_wandering} we replace the single note with a slowly changing frequency and analyse the signal using a hidden Markov model technique from continuous wave searches. 
 In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech. 
-This demonstration of an `optical microphone' serves as a more general exhibition of signal processing. 
-We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.
+This demonstration of an ‘optical microphone’ serves as a more general exhibition of signal processing. 
+We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in ~\ref{sec:conclusions}. 
 Appendix~\ref{app:code} collects the software and scripts used to produce this work.

### 2020-09-11 paper/ifo-introduction.tex
re: restructuring the introduction

@@
 
-In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics an revealed a new means to observe the universe~\cite{GW150914}.
-This first observation and the many to follow it has led to increased public and media interest in gravitational-wave science. 
+In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
+This first observation and the many that followed it have led to increased public and media interest in gravitational-wave science. 
 Gravitational-wave research groups around the world have produced demonstrations and activities to explain their research to general audiences. 
@@
 %\han{look up Sarah Farmer's project}
-Here we describe the application of a table-top demonstration commonly used by gravitational-wave groups, to explain gravitational wave searches and signal processing analysis.
+% Here we describe the application of a table-top demonstration commonly used by gravitational-wave groups, to explain gravitational wave searches and signal processing analysis. 
+%% We mention our experiment later, not here.
 
 % What are gws and gw detectors 
-Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}. 
-They are disturbances in spacetime caused by the acceleration of asymmetric massive objects. 
+Gravitational waves, predicted by Albert Einstein in his theory of general relativity~\cite{Einstein:1916}, are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
 The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects. 
-Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
+Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (aLIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
 Gravitational-wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
@@
 Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
-Rotating neutron stars are prime candidates for continuous wave emission. 
-An example target are rotating neutron stars in low mass X-ray binaries (LMXB), where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
+Rotating neutron stars are prime candidates for continuous wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
+%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
+% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
 The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
-Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches from LMXBs. 
+Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
 Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.
 
-In this paper, present a table-top experiment, where a standard Michelson interferometer acts as a `toy' gravitational-wave detector, which detects sound instead of gravitational waves. 
+In this paper, we present a table-top Michelson interferometer as a `toy' gravitational-wave detector designed to detect sound instead of gravitational waves. 
 The demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
-It allows students in courses such as physics and electrical engineering the freedom to explore the response of an accessible, yet nontrivial opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
-The demonstration also has potential to be adapted for use as an outreach tool to explain gravitational-wave science to a non-speacialist audience.
+%% I don't like the use of ``freedom to explore''.
+Additionally, it allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
+This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
 
 The design of the interferometer is detailed in Section~\ref{sec:ifo} and we demonstrate observing a single tone from a speaker in Section~\ref{sec:single_tone}. 
-In Section~\ref{sec:viterbi_wandering} we replace the single note with a slowly changing frequency and analyse the signal using a hidden Markov model technique from continuous wave searches. 
-In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech. 
-This demonstration of an ‘optical microphone’ serves as a more general exhibition of signal processing. 
+In Section~\ref{sec:viterbi_wandering} we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous wave searches.
+In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech, to varying degrees of success.
+This demonstration of an ‘optical microphone’ (using light to capture sound) serves as a more general exhibition of signal processing. 
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in ~\ref{sec:conclusions}. 

### 2020-09-11 paper/ifo-introduction.tex
re: re: restructuring introduction

@@
 In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
-This first observation and the many that followed it have led to increased public and media interest in gravitational-wave science. 
-Gravitational-wave research groups around the world have produced demonstrations and activities to explain their research to general audiences. 
-Activities range from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}.
-%\han{look up Sarah Farmer's project}
-% Here we describe the application of a table-top demonstration commonly used by gravitational-wave groups, to explain gravitational wave searches and signal processing analysis. 
-%% We mention our experiment later, not here.
-
 % What are gws and gw detectors 
@@
 Additionally, it allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
-This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
+% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
+This demonstration also has the potential to be used as an outreach tool alongside other exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational wave research groups around the world.
+These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational wave science to a non-specialist audience.
+%\han{look up Sarah Farmer's project}
 

### 2020-09-11 paper/ifo-singleTone.tex
Deeksha's comments: single and wandering tone

@@
 	\caption{\label{fig:webcam_spectrum}
-Injected tone at a constant frequency: Fourier amplitude of the intensity time series recorded by the webcam. 
-The frequency of the injection is $2.09\,{\rm Hz}$. 
-The recovered signal peaks at $2.099\,{\rm Hz}$ with a FWHM of $0.033\,{\rm Hz}$. 
-There are also two visible harmonic peaks at integer multiples of the main peak. 
-Their amplitudes as a fraction of the height of the main peak are  $0.103$ and $0.098$ for the $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ peaks respectively. 
+Recovery of a tone at a constant frequency: Here we plot the Fourier amplitude of the intensity pattern against frequency.
+The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has a FWHM of $0.033\,{\rm Hz}$.
+The plot also shows two harmonics at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ with normalised amplitude to the primary peak of $0.103$ and $0.098$ respectively.
 }	
-
 \end{figure}
 
-Continuous-wave searches look for signals which are close to monochromatic. 
-In this section we consider a simple sinusoidal tone; a note at a single constant frequency. 
-As described in Section~\ref{sec:ifo}, the audio is played through a speaker fixed to the back of one of the mirrors.
+Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone; a signal at a single constant frequency.
 
-The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam placed to view the screen (see Fig.~\ref{fig:ifo_schematic_webcam}). 
-The webcam's sample rate is $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
-A tone under $15\,{\rm Hz}$ is played through the speaker for one minute and the interference pattern recorded with the webcam.
+As described in Section~\ref{sec:ifo}, the audio signal is played through a speaker fixed to the back of M2.
+The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam (see Fig.~\ref{fig:ifo_schematic_webcam}).
+The webcam samples at a rate of $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
+A tone with a frequency of 2.09Hz is played through the speaker for one minute and the interference pattern is recorded by the webcam.
 
-We take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
-The green-channel of the video is used as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).
-As an initial inspection, we play the timeseries of the intensity using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
-The two tones sound similar by ear and don't produce any noticeable beating. 
-The result is quantified in Fig.~\ref{fig:webcam_spectrum}. 
-A tone of $2.09\,{\rm Hz}$ is played through the speaker and the data recorded via webcam. 
-Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording.
-We measure a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$. 
-Two harmonics can also be seen in Fig.~\ref{fig:webcam_spectrum} at integer harmonics of the main peak. 
-The $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ peaks have amplitudes of $0.103$ and $0.098$ respectively as a fraction of the height of the main peak. 
-They are probably due to the nonlinear system response discussed in Section~\ref{sec:ifo} and Appendix~\ref{app:intensity_derivation}.
+We take data from an off-centre pixel as indicated by the pink star in Fig.~\ref{fig:interference_pattern}.
+The pixel intensity in the green-channel of the video (the webcam records in three colour channels: red, blue, and green) is used to approximate the total intensity of the interference pattern.
+This is because the laser emits monochromatic green light at $532\,{\rm nm}$.
+The amplitude of the Fourier transform of the recorded signal is shown in Fig.~\ref{fig:webcam_spectrum}.
+We measured a strong signal at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$.
+Although this range doesn't contain our injected frequency, the error is not substantial and cannot be heard by the (untrained) human ear.
+Two harmonics can also be seen at integer multiples of the peak frequency.
+The peaks at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ have amplitudes of $0.103$ and $0.098$ as a fraction of the height of the main peak, respectively.
+They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo} and Appendix~\ref{app:intensity_derivation}.
+
+In the following section, we use a hidden Markov model technique, commonly used in continuous wave searches, to recover a slowly wandering frequency signal.
 

### 2020-09-11 paper/ifo-tableTopGWs.tex
Deeksha's comments: single and wandering tone

@@
 
-The basic physics of gravitational-wave detection can be demonstrated with a table-top Michelson interferometer, an optical configuration commonly used in undergraduate labs. 
-A simple Michelson interferometer comprises a laser, a beamsplitter, and two mirrors. 
-Laser light is split by a beamsplitter into two arms at right angles. 
-At the end of each arm is a mirror which reflects the beams back to the beamsplitter where they recombine. 
-The combination of the two beams produces an interference pattern which is dependent on the relative distance travelled by each of the beams. 
-A change in arm length results in a change in the interference pattern.
+Gravitational-wave detectors such as aLIGO and aVirgo are large, complex experiments.
+However, their design is based fundamentally on a Michelson interferometer, an optical configuration commonly used in undergraduate laboratories.
+It consists of a beam splitter which splits light from a laser down two long perpendicular ``arms'' ($4\,{\rm km}$ long at aLIGO, $3\,{\rm km}$ at aVirgo).
+Mirrors at the end of each arm reflect the light back towards the beam splitter, where they recombine and produce an interference pattern.
+This pattern is dependent on the relative distance travelled by each beam and thus on the relative length of each arm.
+The current generation gravitational-wave observatories use this technique, in combination with others, to detect minuscule changes in distance of two arms due to a passing gravitational wave.
+For example, the first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller with the width of a proton.
 
-Gravitational-wave detectors are large, complex interferometers.
-The LIGO and Virgo detectors have $4\,{\rm km}$ and $3\,{\rm km}$ arms respectively.
-They are sensitive to minuscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$ ~\cite{GW150914}.
-However, their design is based fundamentally on the Michelson interferometer. 
-Detection and analysis concepts can be demonstrated with table-top optical equipment. 
-See Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of table-top Michelson interferometer designs used for science communication.
-
-A commonly-used analogy in explaining gravitational-wave science is sound. 
-Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
-For example, gravitational waves can propagate in a vacuum, whereas sound cannot.}.
-Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
-The vibrations move the mirrors, changing the length of the interferometer arms, and the resulting interference pattern. 
-We use sound throughout this work as a means of simulating gravitational-wave-like signals and the data analysis and signal processing techniques used to detect them.
+The complex detection and analysis techniques used in the gravitational wave community can be demonstrated using a table-top experiment, such as the one presented here.
+Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} provide a selection of other table-top Michelson interferometer designs used for science communication.
+In addition to this, gravitational-wave signals from binary black hole and neutron star mergers are often converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
+For example, gravitational waves can propagate in a vacuum, whereas sound cannot.} to aid in explanations.~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
+This analogue allows scientists to use sound vibrations as a simple means to move the components of a demonstration interferometer.
+The vibrations move the mirrors, changing the length of the interferometer arms, and thus changing in the interference pattern.
+We use audio signals throughout this work to simulate gravitational-wave-like signals and to demonstrate different data analysis and signal processing techniques used to recover such signals.
 
@@
 Schematic of the table-top Michelson interferometer. 
-The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}. 
-The beam is first reflected from a mirror at a right angle and then focused at a converging lens before reaching the beamsplitter which is $50$/$50$ in reflection and transmission. 
+The output of a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO} (top-left) is reflected from a mirror at a right angle.
+The beam then passes through a converging lens before reaching the beamsplitter which is $50$/$50$ in reflection and transmission.
 The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and are recombined at the beamsplitter. 
-The beam is enlarged by a diverging lens after the beamsplitter. 
-The interference pattern is projected onto a screen and recorded using a commercial webcam connected to a personal computer (PC). 
-A speaker fixed to the back of M2 provides the signal input from the PC.  
+The output interference pattern is enlarged using a diverging lens, projected onto a screen, and recorded using a commercial webcam connected to a personal computer (PC). 
+A speaker fixed to the back of M2 is used to inject audio signals from the PC into the interferometer.
     }
@@
 
-The Michelson interferometer configuration used in this work is shown schematically in Fig.~\ref{fig:ifo_schematic_webcam}.
-It is constructed on a $450 \times 450 \,{\rm mm} $ optical breadboard. 
-The laser (top-left) has a wavelength of $532\,{\rm nm}$.
-In the configuration shown in Fig.~\ref{fig:ifo_schematic_webcam} the laser beam is first reflected from a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard). 
-The beam passes through a converging lens and then to the beamsplitter, which is $50$/$50$ reflection and transmission. 
-The beams are reflected by mirrors at the end of each of the two arms ($\approx 7.5\,{\rm cm}$ and $\approx 10.0\,{\rm cm}$ long) and are recombined at the beamsplitter. 
-The resulting interference pattern is projected onto a screen and recorded by a webcam.
+The optical configuration of the Michelson interferometer used in this work is shown in Fig.~\ref{fig:ifo_schematic_webcam}.
+It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses as a source a green eye-safe laser (top-left) with peak emission at $532\,{\rm nm}$.
+The output of the laser is first reflected off a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
+After passing through a converging lens of focal length (FL = \jam{???}), the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2).
+A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations.
+This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a PC.
+The other speaker in the pair is kept face-down and as far away from the interferometer as possible to prevent it from acting as a second source.
+The beams are reflected by mirrors M1 and M2, located at the end of (approximately) $7.5\,{\rm cm}$ and $10.0\,{\rm cm}$ long arms.
+The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging lens of FL = \jam{???}, and projected onto a screen.
+The interference pattern, as shown by the photograph in Fig.~\ref{fig:interference_pattern}, is a set of concentric light and dark rings (fringes). 
+A change in the relative length of the arms causes these rings to move radially in or outwards.
+% Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
+An intensity time series of some point in the interference pattern is recorded by a webcam (or later, a photodiode) for data analysis to extract the signal injected at the speaker.
 
@@
  \begin{center}
-  \includegraphics[width=.45\textwidth]{figures/webcam_still0_crop.pdf}
+  \includegraphics[height=.45\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
  \end{center}
  \caption{\label{fig:interference_pattern}
-Michelson interference pattern as viewed by an off-axis webcam (which is why the pattern is elliptical). 
-The red dot marks where the intensity time series is measured. 
+The interference pattern produced by the table-top Michelson interferometer.
+This image was taken with the webcam off the beam axis, which may explain why the (predicted spherical) pattern appears elliptical.
+The pink star indicates the point where the intensity time series measurements were taken.
 The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
@@
 
-The interference pattern is a ring of interference fringes as shown by the photograph in Fig.~\ref{fig:interference_pattern}. 
-The pattern is enlarged by a diverging lens between the beamsplitter and the screen to aid in viewing. 
-A change in the relative length of the interferometer arms results in the interference rings moving radially inwards or outwards. 
-Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed.
+The motion of the interference fringes follows the motion of M2, and therefore of the speaker.
+The amplitude of these motions is given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
+If the relative change in the length of the arms is kept small enough, the intensity of the interference pattern at any point on the screen oscillates at the same frequency as the injected audio.
+But if the relative change in the length of the arms is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting (artificially) raises the measured frequency.
+As such, any motion of the fringes must be kept small by playing the sound softly through the speaker.
+Even without over-counting, large fringe motions display a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting---phenomena (e.g frequency doubling).
+We derive this nonlinear relationship between the length difference of the arms and the intensity of the pattern at some point on the screen in Appendix~\ref{app:intensity_derivation}.
 
-A $0.5\,{\rm W}$ speaker is fixed to the back of one of the mirrors (M2 in Fig.~\ref{fig:ifo_schematic_webcam}), providing a controllable source of vibrations. 
-The speaker is one half of a pair of commercial, USB-powered speakers fed by a $3.5\,{\rm mm}$ jack and driven by a PC. 
-It is attached to the back of the mirror using commercial adhesive putty. 
-The other speaker in the pair is kept face-down and as far away from the interferometer as the cables allow to reduce its impact as a second source of vibrations.
+In the following section, we start simply by injecting and recovering a audio signal of constant frequency (a single tone).
 
-The motion of the interference fringes follows the motion of the mirror, and therefore the speaker. 
-If the fringe motion is small, then changes in the intensity at some point on the screen occur at the same frequency as the injected audio but with amplitude given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
-If the fringe motion is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting artificially raises the measured frequency.
-As such, any motion of the fringes must be kept small by playing sound softly through the speaker.
-Even without over-counting, large fringe motions introduce a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting!---phenomena like frequency doubling.
-We derive the nonlinear relation between the separation of the mirrors and the intensity of the pattern at some point on the screen in Appendix~\ref{app:intensity_derivation}.
-
-Data can be recorded via a webcam (as shown in Fig.~\ref{fig:ifo_schematic_webcam}) or photodiode. 
-In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used. 
-In Section~\ref{sec:optical_microphone}, which considers more complex audio, we turn to a photodiode because it offers a higher sample rate than the webcam.
+% Data can be recorded via a webcam (as shown in Fig.~\ref{fig:ifo_schematic_webcam}) or photodiode. 
+% In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used. 
+% In Section~\ref{sec:optical_microphone}, which considers more complex audio, we turn to a photodiode because it offers a higher sample rate than the webcam.
 

### 2020-09-11 paper/ifo-wanderingTone.tex
Deeksha's comments: single and wandering tone

@@
 
-Continuous gravitational wave signals may wander slowly in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars~\cite{MelatosDouglassSimula:2015,Jones:2010} or stochastic accretion flows in binary neutron stars~\cite{BildstenTB:1998}.
-The audio analogue of this is a tone that wanders in frequency; a note that changes pitch.
+Continuous wave searches, such as in Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017,SearchTwoSpecS6:2017,SunEtAlSNR:2018,JonesSun:2020}, are often (not always) performed on long datasets (months to years in duration).
+The frequency of a continuous wave signals may wander significantly over this timescale, due to stochastic internal processes in the superfluid interior of isolated neutron stars~\cite{MelatosDouglassSimula:2015,Jones:2010} or variable accretion flows onto a neutron star from a stellar companion (e.g.
+as in LMXBs)~\cite{BildstenTB:1998}.
+In this context, wandering ``significantly'' means ``across multiple frequency bins'', where the typical width of a frequency bin is the reciprocal of the total observing time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
+The audio analogue to these continuous wave signals is a tone that wanders in frequency; a note that changes pitch.
+Here, we adapt the data analysis technique used to search for slowly wandering continuous waves signals from spinning neutron stars in aLIGO and aVirgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
 
-The analysis techniques used here are inspired by those used to search for continuous waves from spinning neutron in LIGO and Virgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
-Continuous wave searches are performed on long datasets (months to years in duration). 
-The frequency of gravitational-wave emission wanders significantly over the observation period. 
-In this context, ``significantly'' means ``across multiple frequency bins'', where the typical width of a frequency bin is the reciprocal of the total observing time~\cite{JKS:1998,ScoX1O2Viterbi:2019}
-
-One method to search for a wandering signal is to split the data into several shorter segments which are analysed individually. 
+One method to search for a wandering signal is to split the time series into several shorter segments, analyse these individually, then recombine them (under some continuity assumption) to coherently analyse the whole.
 The detection statistics in every segment are combined to form a grid in time and frequency (spectrogram).
-The value at each spectrogram point determines the likelihood that a signal is present at a particular time and frequency. 
+The value at each grid point gives the likelihood that a signal is present at that particular time and frequency.
 In gravitational-wave data analysis, a detection statistic is used to calculate the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
 The detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
+
 In this work the detection statistic is the amplitude (modulus) of the Fourier transform of the data at each time segment, which maximizes the likelihood of detecting a sinusoid in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}.
 The resulting spectrogram gives the Fourier amplitude for each frequency and time bin.
-
 In continuous wave searches a hidden Markov model is used to describe the system.
-The frequency (the state) of the signal is unknown (hidden) and the signal is Markovian in that the state at any time depends only on the state in had at the previous time step. 
+The frequency (the state) of the signal is unknown (hidden) and the signal is Markovian in that the state at any time depends only on the state at the previous time step.
 The hidden Markov model method relates the observable quantity (the detection statistic) to the signal frequency (the hidden state).
@@
 The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method to find the most likely path of a signal given an observation sequence.
+
 This method is used in continuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,JonesSun:2020,MiddletonEtAlO2LMXBs:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.

### 2020-09-12 paper/ifo-tableTopGWs.tex
added: focal lengths

@@
 The output of a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO} (top-left) is reflected from a mirror at a right angle.
-The beam then passes through a converging lens before reaching the beamsplitter which is $50$/$50$ in reflection and transmission.
+The beam then passes through a converging lens (FL = $125.0\,{\rm mm}$) before reaching the beamsplitter which is $50$/$50$ in reflection and transmission.
 The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and are recombined at the beamsplitter. 
-The output interference pattern is enlarged using a diverging lens, projected onto a screen, and recorded using a commercial webcam connected to a personal computer (PC). 
+The output interference pattern is enlarged using a diverging lens (FL = $-25.0\,{\rm mm}$), projected onto a screen, and recorded using a commercial webcam connected to a personal computer (PC). 
 A speaker fixed to the back of M2 is used to inject audio signals from the PC into the interferometer.
@@
 The output of the laser is first reflected off a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
-After passing through a converging lens of focal length (FL = \jam{???}), the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2).
+After passing through a converging (plano-convex) lens of focal length (FL = $125.0\,{\rm mm}$), the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2).
 A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations.
@@
 The beams are reflected by mirrors M1 and M2, located at the end of (approximately) $7.5\,{\rm cm}$ and $10.0\,{\rm cm}$ long arms.
-The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging lens of FL = \jam{???}, and projected onto a screen.
+The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging (bi-concave) lens of FL = $-25.0\,{\rm mm}$, and projected onto a screen.
 The interference pattern, as shown by the photograph in Fig.~\ref{fig:interference_pattern}, is a set of concentric light and dark rings (fringes). 

### 2020-09-28 paper/paper_main.tex
re: restructuring the introduction
Deeksha's comments: single and wandering tone
edits to the acknowledgements and tidying up author list comments to get ready for circulation

@@
 
-\author{\han{to do: check names - initials?}}
-\affiliation{%
-\han{to do: check all affiliations}
-}
+%\author{\han{to do: check names - initials?}}
+%\affiliation{%
+%\han{to do: check all affiliations}
+%}
 
@@
 Finally we use the interferometer as an `optical microphone' to capture music and speech and apply filtering techniques to recover the original signals from the noisy data.
-Such techniques include simple bandpass and notch filters, as well as comparing a Weiner filter to the statistical logMMSE method.
+Such techniques include simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
 Simple chords and drums are easily recovered, but complex music and speech are more challenging. 
@@
 
-\section{Constant tone}
+\section{Constant frequency signal}
 \label{sec:single_tone}
@@
 
-\section{Wandering tone}
+\section{Wandering frequency signal}
 \label{sec:viterbi_wandering}
@@
 \begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance throughout the project. 
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; and Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction.
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
-This work also received financial support for hardware from the Institute of Physics International Member Grant and for travel from the Australian National University PhB Science program.
+This work received financial support towards hardware from the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme and travel support was received from the Australian National PhB Science program.
+This work has been assigned LIGO document number P2000386.
 

### 2020-09-28 paper/paper_main.tex
small rewording in acknowledgements

@@
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
-This work received financial support towards hardware from the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme and travel support was received from the Australian National PhB Science program.
+This work received financial support towards hardware from the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme and travel support from the Australian National PhB Science program.
 This work has been assigned LIGO document number P2000386.

### 2020-10-01 paper/ifo-complexAudio.tex
minor cleaning of figures dir
Deeksha's comments - last few sections

@@
 % aim of optical microphone: play back injected sounds
-A natural successor to single and wandering tones is complex audio, such as music and speech. 
-Complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not applicable directly.
-Instead, in this section, we explore how to use the interferometer as an `optical microphone' to detect and play back complex audio signals, using light to capture sound. 
-In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
-The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
+A natural successor to constant and wandering tones is complex audio, such as music and speech.
+As complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not directly applicable here.
+Instead, in this section, we use the interferometer as an `optical microphone' to detect and play back complex audio signals (i.e.\ using light to capture sound).
+In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
+The device has precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defence industry and that operate on a variety of related principles.
 %It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
@@
 
-We describe additional hardware components to the demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}.
-Filter and speech enhancement techniques applied to the data are detailed in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse} respectively.
-This section is indirectly related to the advanced signal processing required to extract gravitational-wave signals. 
-The optical microphone provides an interesting independent demonstration for a broader physics and engineering audience, particularly for the undergraduate lab.
+We describe the additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}.
+Filter and speech enhancement techniques applied to the data are outlined in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse}, respectively.
+This section is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals.
+In addition to this, the optical microphone serves as an interesting independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories.
 
@@
 
-Speech and music recordings require a higher sample rate than that of the webcam ($30\,{\rm Hz}$) used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering}. 
-Audible frequencies are in the range $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
+Audible frequencies are in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$.
 Speech intelligibility requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$~\cite{speech_intelligibility}.
-Therefore the optical microphone data requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music. 
-We achieved this with a photodiode\footnote{
-A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the `reverse' direction.
-As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.
-}
+Therefore the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music.
+This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$.
+To overcome this issue, we use a photodiode\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the `reverse' direction.
+As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.} at the output of the interferometer to achieve a sampling rate of $16\,{\rm kHz}$.
 
-We place a OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which makes a photo-detector that produces a voltage dependent on the incident intensity. 
-The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
-We see no difference when monitoring at other positions in the interference pattern.
+We place a OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage dependent on the incident intensity. 
+The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
 The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath. 
 The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi (Pi) Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data.
-Together, these sample the signal at roughly $16\,{\rm kHz}$. 
-Resources for using the Pi and photdiode are detailed in Appendix~\ref{app:circuit_diagram}.
+Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. Resources for using the Pi and photodiode are described in Appendix~\ref{app:circuit_diagram}.
 
-Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing into the detected range. 
-To prevent this, we include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC. 
+Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing\footnote{folding of higher frequencies above half the sampling rate} into the detected range.
+We include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
 This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
-We also place another cloth screen over the face of the photodiode, as the reading without it is too high for the ADC after gaining through the op-amps. 
+We also place a cloth screen over the face of the photodiode to reduce the intensity of the voltage signal as to avoid saturating the ADC.
 % It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
@@
 
-We test the optical microphone with a variety of recordings including speech from different people and music ranging from simple melodies and rhythms to full songs. 
 %The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources). 
-During recordings, care was taken to ensure minimal activity around the interferometer to reduce environmental noise. 
-The timeseries was then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}, as before.
-When processing the results, we generally only analysed the first $10\,{\rm s}$ of each observation for efficiency.
+We test the optical microphone with a variety of recordings including speech from different people and music ranging from simple melodies and rhythms to songs.
+During recordings, care must be taken to minimise activity around the interferometer to reduce the amount of environmental noise coupling into the interferometer.
+The time series data is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function.
+When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.
 
-The raw output of the optical microphone (with only anti-aliasing) is noisy with a loud bass hum throughout. 
-Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic, likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}).
-This mains power signal is present in the power spectral density of background noise, with the speaker switched off, as shown in the top panel of Fig.~\ref{fig:psd_noise}.
-The mains signal is still present (but far weaker) in measurements with the photodiode in darkness, suggesting that ambient lighting has a large contribution.
-Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode. 
-There is also a broad peak visible at around $750\,{\rm Hz}$ in Fig.~\ref{fig:psd_noise}, which we have not been able to identify an origin for.
+The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass hum throughout.
+This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), as shown in the top panel of Fig.~\ref{fig:psd_noise}.
+The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ mains signal up to and beyond $8$th harmonic thereof.
+The mains signal is also present (but far weaker) in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution.
+Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself.
+The appearance of harmonics of the mains noise is likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}).
+The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
 
@@
 
-\subsection{Additional filters}
+\subsection{A hierarchy of filters}
 \label{sec:simple_filters}
 
-Here we describe several approaches to remove the $50\,{\rm Hz}$ mains hum and its harmonics. 
-We test these approaches on $1\,{\rm s}$ of data throughout this section as an illustrative example.
+Here, we describe several filters used to remove the $50\,{\rm Hz}$ mains hum and its harmonics, but also to ultimately improve the speech intelligibility of the recording (to be able to understand the words).
+We start with some naive approaches and then move to traditional signal processing filters like band-passing and cascaded notches.
+We finish by combining these with advanced statistical techniques like a Wiener filter and the logMMSE estimator.
+As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
 
@@
 
-Fig.~\ref{fig:noiseSubtract} shows the frequency spectrum of $1\,{\rm s}$ of recorded speech. 
-The first and second panels show the original signal before passing through the interferometer and the optical microphone recording of the signal respectively. 
-An intuitive means to remove noise is to subtract the noise spectrum of the instrument directly from the signal spectrum. 
-The third panel of Fig.~\ref{fig:noiseSubtract} shows a recording from the optical microphone with no input audio signal. 
-We see that the output noise is coloured, which reflects the non-linearity of the optical microphone.
-The fourth panel of Fig.~\ref{fig:noiseSubtract} shows the frequency spectrum after the noise spectrum has been subtracted. 
-We see no obvious improvement, which may be due to time-variant noise.
+The Fourier spectrum of the source signal is shown in the first panel of Fig.~\ref{fig:noiseSubtract}.
+The second panel shows the spectrum of the raw optical microphone recording of the signal.
+The third panel of Fig.~\ref{fig:noiseSubtract} shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.
 
-The next step is to apply filters to the data. 
-The ideal filter would be one that:
+Given that we have access to the background noise spectrum, an intuitive way to remove noise is via subtracting the noise spectrum from the recorded spectrum.
+The fourth panel of Fig.~\ref{fig:noiseSubtract} shows the spectrum obtained after subtracting the background noise spectrum.
+We see no obvious improvement, which may be attributed to a time-variant noise spectrum.
+Why the noise sources are time-varying is a fact yet to be determined.
+
+Beyond this naive approach, we need to apply filters to the data.
+In signals processing, the ideal filter would be one that:
 %\begin{itemize}
@@
 ii) does not change the rest of the spectrum, and 
-iii) has smooth edges so as to not damage the signal after convolution. 
+iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
 %\end{itemize}
-These three conditions are contradictory and any filter must compromise between them.
+However, these three conditions are cannot all hold at once and any filter must compromise between them.
 
-We note that simply zeroing the frequency channel of each $50\,{\rm Hz}$ in the frequency spectrum is unsuccessful. 
-Effectively, this multiplies the spectrum by a rectangular comb filter. 
-It does remove the mains harmonics, however it audibly ruins the rest of the signal. 
-Applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter. 
+\subsubsection{Comb filter}
+
+Simply zeroing the frequency bins corresponding to the harmonics of the $50\,{\rm Hz}$ mains signal is unsuccessful.
+This effectively multiplies the spectrum by a rectangular comb filter.
+It does remove the mains harmonics, but audibly ruins the rest of the signal due to lack of smoothness.
+This is because applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
 The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is some combination of sinc functions, which significantly corrupt the signal.
@@
 
-Applying a high-pass filter with cut-off frequency $150\,{\rm Hz}$ to the signal works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics.
-However, mains harmonics above $100\,{\rm Hz}$ remain. 
-The region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}. 
-Therefore, although a high-pass filter with a higher cut-off removes more harmonics, the played-back signal is unrecognisable. 
-Applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.
+\subsubsection{High-pass filter}
+
+A high-pass filter smoothly attenuates frequencies below some cut-off frequency.
+Applying a high-pass filter, with a cut-off frequency around $150\,{\rm Hz}$, to the signal spectrum works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics.
+However, the mains harmonics above $100\,{\rm Hz}$ remain.
+Using a high-pass filter with a higher cut-off can be used to mitigate this issue.
+However, it makes the played-back signal unrecognisable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
+Often in speech processing the logarithm of the signal is taken since the amplitude information seems to be more important to intelligibility than the phase information.
+However, applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.
 
 %We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass
+
+\subsubsection{Butterworth band-pass filter}
 
@@
 
-We next apply a Butterworth filter. 
-A low-pass Butterworth filter with cut-off frequency $f_c$ and gain $\varepsilon$, has a frequency response (attenuation at each frequency)
+A band-pass filter combines a high-pass filter and a low-pass filter to smoothly attenuate frequencies outside of some band.
+A Butterworth band-pass filter is just a particular choice of the frequency response (an attenuation at each frequency) as to be `maximally flat' within the pass-band region.
+The Butterworth low-pass component is given by Eqn.~\ref{eq:butterworth} and is combined with a similar high-pass filter to form the band-pass filter.
+
 \begin{equation}
 \label{eq:butterworth}
-H(\omega) = \left[1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right]^{-1/2}
+H(\omega) = \left[1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right]^{-1/2},
 \end{equation}
@@
 %\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
-To construct a Butterworth band-pass filter, the low-pass filter in Eqn.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
-The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}).
-The effect of the Butterworth band-pass filter on the power spectral density of the background noise is shown in the bottom panel of Fig.~\ref{fig:psd_noise}, where we see that it reduces the amplitude of the mains harmonics below $150\,{\rm Hz}$
-The response of the filter is shown in Fig.~\ref{fig:butterworth}.
-Band-passing the signal does not address the remaining harmonics above $100\,{\rm Hz}$. 
-However, it does suppress other noise at higher frequencies. 
-In general, the Butterworth filter performs well at maintaining source intelligibility as it is optimised to be `maximally flat' within the band region (prioritising condition ii) above).
-However recordings of speech still have low intelligibility (the speech can't be understood).
-Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique to address this.
+
+In the above expression, $f_c$ represents the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ represents the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
+Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here, a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}, 3\,{\rm kHz})$.
+This high frequency cut-off is chosen since the frequencies important for speech generally lie below $2\,{\rm kHz}$~\cite{speech_intelligibility}.
+
+The effect of applying this filter to the background noise PSD can be seen in the bottom panel of Fig.~\ref{fig:psd_noise}.
+The Butterworth band-pass filter reduces the amplitude of mains harmonics below $150\,{\rm Hz}$ and suppresses unrelated noise sources above $3\,{\rm kHz}$.
+However, it does not address the issue of mains harmonics above $150\,{\rm Hz}$ (i.e., in the pass-band).
+In the following section, we experiment with a cascade notch filter to address this.
+
+\subsubsection{Cascade notch filter}
 
@@
 
-We also consider a combined notch filter and Wiener filter. 
-Here we describe our use of these methods with timeseries and frequency spectrum results shown in Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively. 
-The original signal is identical to that shown in Fig.~\ref{fig:noiseSubtract} where the first and second panels of Fig.~\ref{fig:timeNotchWiener} show the original audio signal and the recording after passing through the optical microphone respectively (i.e.
-the timeseries of the data shown in the first two panels of Fig.~\ref{fig:noiseSubtract}).
-
-\subsubsection{Cascade notched filter}
-A straightforward method for removing hum noise and its harmonics is to use a sequence of notch filters centered at each of the frequency bins we want to remove. 
-A typical transfer function of infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
+One method to remove mains noise and its harmonics is to use a sequence of notch filters centered at each of the frequency bins we want to remove.
+This sequence is known as a cascade of notches.
+Here the notches are smooth in comparison to the naive zeroing of each frequency (which looks like a rectangular comb).
+The complex frequency response of a typical infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
 \begin{equation}
@@
 \end{equation}
-where $\alpha$ and $\beta$ are self-tuned parameters which adjust the locations of poles and zeros and $z$ is the complex frequency. 
-By allowing the numerator and demoninator of $H(z)$ to be zero, we can find the corresponding locations of zeros and poles of the filter. 
-There are zeros are at $e^{\pm jw_0}, \text{where } w_0=\cos^{-1}(\beta)$, and poles are at $re^{\pm j\phi}, \text{where } r=\sqrt{\alpha} \text{ and }\phi=\cos^{-1}[\beta(1+\alpha)/2\sqrt{\alpha}]$.
-At $w_0$, the magnitude response will be zero. 
-Thus, we can calculate $\beta$ setting $w_0$ to be the frequency we want to remove. 
-The parameter $\alpha$ affects the bandwidth $B_w$ of the filter where $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$. 
-%In our paper, we first decide $\beta$ for each filter.
-
-In this work we find a cascade filter with $15$ notches works well. 
-We decide $\beta$ first by building a notch filter at each of the frequency bins located at $50k\,{\rm Hz}$ where the integer has values $k=1,2,\dots,15$. 
-%The mains noise is located at $50k\,{\rm Hz}$ where the integer has values $k=1,2,\dots,15$. 
-%For each $k$, we build a notch filter at the specific frequency bin at $50k \, {\rm Hz}$. 
-We then alter the parameter $\alpha$ to determine the bandwidth of each filter. 
-On one side, the bandwidth should be as narrow as possible to avoid disturbing useful signals. 
-This can be achieved through increasing orders of the filter (add more poles and zeros). 
-On the other side, the bandwidth should have enough residues to account for the uncertainty of our pinpointing the location of hum noise. 
-After considering these two aspects, we find the ideal transfer function for each filter. 
-The whole cascaded filter is achieved by multiplying them together, 
+Where $z$ is the complex frequency and $\alpha$ and $\beta$ are parameters that control the filter.
+$w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed) at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which reflects how quickly the response changes around the notched frequency).
+We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centred on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$.
+When then choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
+The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,
 \begin{equation}
     \label{eqn:notch15}
-    H(z) = H_{k=1}(z) H_{k=2}(z)\dots H_{k=15}(z)
-%    H(z)=H_1(z)H_2(z)\dots H_k(z),k=1,2,\dots 15\,,
+    H(z) = \prod_{k=0}^{14} H_k(z).
 \end{equation}
-where $H(z)$ denotes the transfer function of the full cascaded notch filter, and $H_k(z)$ denotes one single notch filter for removing $50k \, {\rm Hz}$ frequency. 
-We use the built-in MATLAB filter design toolbox to help us realize a well performed cascaded notch filter (each filter has order 6).  
-The magnitude response of the first five notch filters are shown in Fig.~\ref{fig:notchMagResponse}. 
+
+We use the built-in MATLAB filter design toolbox, the (amplitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
+The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}, respectively.
+We see that the mains harmonics are significantly attenuated in comparison to the results shown in Fig.~\ref{fig:noiseSubtract}.
+
 \begin{figure*}
 \begin{center}
-\includegraphics[width=.7\textwidth]{figures/1111}
+\includegraphics[width=.7\textwidth]{figures/notch_filter_response}
 \end{center}
@@
 \end{figure*}
-
-The timeseries and frequency spectrum results of the notch filter are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
-We see that the mains harmonics are significantly attenuated in comparison to the analysis in Fig.~\ref{fig:noiseSubtract}.
 
@@
 
-Here we use a Wiener filter to compute an estimate of the original signal. 
-We label the noisy speech signal sequence as  $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence. 
-It is the sum of the original clean signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$, 
+A Wiener filter is an advanced statistical technique that estimates the injected signal given prior information about the injected spectrum and the reference spectrum of the background noise.
+Let the observed noisy speech signal sequence be given as $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence.
+It is the sum of the original injected (or “clean”) signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$,
 \begin{equation}
@@
 \end{equation}
-Given the observation data $\textbf{x}$, our goal is to filter the signal $\textbf{s}$ from the noise $\textbf{w}$ to find an estimated $\hat{\textbf{s}}$, where the difference between $\hat{\textbf{s}}$ and true signal $\textbf{s}$ is minimized in the mean square error sense, which is represented as
+Given the observed data $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the (Bayesian) mean square error (MSE), as defined in Eq.~\ref{eq:BMSE}, between the two.
 \begin{equation}
+\label{eq:BMSE}
 \text{BMSE}(\hat{\textbf{s}})=E[(\textbf{s}-\hat{\textbf{s}})^2]\,,
 \end{equation}
-where $\text{BMSE}$ is the Bayesian mean square error. 
-Assuming that: i) $\textbf{x}$ is wide sense stationary~\footnote{
-A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
-$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
-$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
-only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} 
-with zero mean; ii) the signal $\textbf{s}$ is zero mean; and iii) the noise processes $\textbf{w}$ are uncorrelated with the signal $\textbf{s}$, we can further express the estimated $\hat{\textbf{s}}$ to be a linear combination of present and past input observed data
+
+If we assume that: i) $\textbf{x}$ is wide sense stationary~\footnote{ A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all $t_1,t_2 \in R$, if (1) it’s mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
 \begin{equation}
@@
 where $\textbf{h}=\{h(0),\dots h(n)\}$ represents the coefficients of an $n$th order Wiener filter.
+These assumptions are to be justified by the efficacy of the Wiener filter.
+
 %Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
@@
 %and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$.
-The famous Wiener-Hopf equation~\citep{noble1959methods} provides us a way to determine $\textbf{h}$, as
-\begin{equation}
-\label{eqn:winer-hopf}
+
+The famous Wiener-Hopf equation~\citep{noble1959methods} allows us to determine $\textbf{h}$, as\begin{equation}
+\label{eqn:wiener-hopf}
 \begin{bmatrix}  
@@
 %where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
-If we do not need to guarantee the causality of the Wiener filter, i.e.
-we can estimate the current signal based on both past and future observations, then Eqn.~\ref{eqn:winer-hopf} can be represented in the frequency domain as
+
+As an instructive aside, if we let the Wiener filter be non-causal (i.e.
+we estimate the current signal based on both past \emph{and future} observations), then we can represent Eqn.~\ref{eqn:wiener-hopf} in the frequency domain as
 \begin{equation}
@@
 \end{equation}
-where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ are the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 
-Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the Wiener filter amplifies the input signal where signal to noise ratio (SNR) is high, and weakens the signal where SNR is low.
-We note that this filter is non-causal, which may not be suitable for real time signal processing. 
-More detailed analysis can be found in Ref.~\citep{10.5555/151045}. 
-In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:winer-hopf}, and set the length of the filter to $n=100$. 
+where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data (i.e., signal + noise), the injected signal, and the background noise, respectively.
+Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high, and attenuates the signal where the SNR is low.
+The causal Wiener filter (that only makes estimates based on past observations) is similar.
+More detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.
+% That this filter is non-causal makes it unsuitable for real time signal processing
+
+In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}, and set the length of the filter to $n=100$. 
 A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
@@
 
-The timeseries and frequency spectrum results of the Wiener filter are are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
-There is improvement which is particularly visible in the timeseries, however a strong noise hum persists. 
+In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}.
+A higher order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required.
+For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
+The effects of applying the Wiener filter to the time series and frequency spectrum are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener}, respectively.
+We see a significant improvement in the time series of the recovered signal, however a strong noise hum persists, audibly.
 %The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
@@
 
-\subsubsection{Combined Notch and Wiener filter}
+\subsubsection{Combined notch and Wiener filter}
+
+Here we experiment with applying a combination of the cascaded notch and the Wiener filter to the recorded speech signal.
+
 The Wiener filter makes use of statistical information from the speech data and noise.
-It amplifies the part of the signal with high SNR, while suppressing the part with low SNR. 
-It is implemented in the form of a finite response filter, which ensures linear phase response and stability at the cost of high orders. 
+It amplifies the part of the signal with high SNR, while suppressing the parts with low SNR.
+It is implemented in the form of a finite response filter, which ensures linear phase response and stability (both desirable), but at the cost of high orders computationally.
 By comparison, the notch filter is based on directly removing the unwanted frequency components.
-It is implemented in the form of an infinite impulse response fileter. 
-Although it significantly decreases the order of the filter, it unavoidably introduces nonlinear phase and instability.
+It is implemented in the form of an infinite impulse response filter.
+Although it significantly decreases the order of the overall filter, it unavoidably introduces nonlinear phase and instability.
 
-Combining the notch and Wiener filter, we have a trade off between the two filters, which shows better performance in filtering out different noise components (white noise plus hum noise). 
-The results from the combination of notch and Wiener filters are visibly cleaner as shown in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} for the timeseries and spectrum respectively
+By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen visibly in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener}.
 
@@
 
-\subsection{logMMSE estimator}
+\subsubsection{logMMSE estimator}
 \label{sec:logmmse}
@@
 
-Speech enhancement for a noisy channel is a classic problem in signal processing. 
-In Ref.~\cite{SubjectiveComparison}, $13$ speech enhancement methods are compared and the best found to be a log minimum means-square error (logMMSE) estimator. 
-This type of estimator is based on work in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
+What we’re attempting here is not new, speech enhancement of noisy channels is a classic problem in signal processing.
+To borrow from that field, Yi Hu and P.C.
+Loizon~\cite{SubjectiveComparison} compared $13$ speech enhancement methods and found the log minimum means-square error (logMMSE) estimator to be the best among them, qualitatively, at recovering speech.
+This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean square error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that is measures the MSE between the logarithm of the Fourier amplitudes.
+This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
+We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}) to the recorded signal.
 %Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
 %This type of estimator is based on work Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
-The logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
-We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}).
 
@@
 %Applying this estimator to the optical microphone output produced dramatically cleaner results.
-
-The timeseries and frequency results of the logMMSE are shown in Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} respectively. 
-The top two panels of each figure again show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison. 
-The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} show the timeseries and corresponding spectra after applying the logMMSE filter.
 
@@
 
-The results show significant attenuation of the mains harmonics as well as a general smoothing of the spectrum.
-However, we find that the intelligibility of speech remains low with the voice sounding mumbled and indistinct. 
-The logMMSE removes most of the background noise, but does not clarify the speech.
+The results of the logMMSE estimator are shown in Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum}.
+The top two panels of each figure show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison.
+The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} show the time series and corresponding spectral amplitude after applying the logMMSE estimator.
+We see significant attenuation of the mains harmonics as well as a general smoothing of the spectrum.
+The logMMSE method indeed removes most of the background noise, but does not significantly enhance the speech as the voice sounds muffled and indistinct, with the words unable to be understood.
 
-We find some improvement with music recordings. 
-After applying the logMMSE, simple chords and drums can be heard clearly, however more complex melodies are sometimes completely absent. 
-From observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
-This could be a perceptual effect or a frequency dependence somewhere in the optical microphone. 
-Speculating, perhaps the speaker-mirror oscillates more at at low frequencies and so instruments like electric bass and drums are louder in the results.
+We find some improvement with music over speech.
+After applying the logMMSE estimator, simple chords and drums can be heard clearly, however more composite sounds and complex melodies cannot be heard clearly.
+Our observations suggest that this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins sometimes can’t be heard at all.
+This could be a perceptual effect or a frequency dependence somewhere in the optical microphone, although the exact cause is yet to be determined.
+Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums are louder in the results.
 
-To address these problems we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-This requires a better understanding of the system as discussed in Section~\ref{sec:future_work}.
+To address these problems, we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
+Or if instead they are present in the noisy data but we need better techniques to extract them.
+To determine this requires a better understanding of the system as discussed in Section~\ref{sec:future_work}.
 

### 2020-10-01 paper/ifo-conclusions.tex
Deeksha's comments - last few sections

@@
 \begin{document}
+
+In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector and demonstrate a range of signal processing techniques used within the GW community and more broadly.
+Gravitational-wave signals are simulated by playing sound recordings through a speaker which is fixed to the back of one of the arm mirrors.
+We consider quasi-monochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music for more general interest.
+These recordings of complex audio could also serve as more distant analogues to un-modelled gravitational-wave bursts, e.g.
+from supernovae.
+
+We demonstrate the recovery of a signal with constant frequency and signal with slowly wandering frequency in Sections~\ref{sec:ifo} and ~\ref{sec:single_tone}, respectively.
+We use a basic webcam with a sampling rate of $30\,{\rm Hz}$ to track the interference pattern.
+For the wandering frequency we successfully recover the injected signal using the Viterbi algorithm.
+These sections directly demonstrate the techniques used in continuous gravitational-wave searches.
+
+In Section~\ref{sec:optical_microphone}, we demonstrate the recovery of more complex signals of music and speech.
+By substituting the webcam with a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone by greatly increasing the sampling rate.
+We demonstrate that the optical microphone is capable of playing back simple recordings well but produces unintelligible recordings of speech.
+The mains signal at $50\,{\rm Hz}$ and its harmonics dominate the photodiode output.
+To recover the signal from under this noise we apply a hierarchy of filters and statistical techniques from signals processing, including a Butterworth band-pass filter and a notch filter, to a Wiener filter and the logMMSE estimator.
+We find that the logMMSE estimator works the best from those tried but still doesn’t enhance the speech to an intelligible level.
+This demonstration allows students to explore a variety of filtering and speech enhancement techniques.
+Further avenues of exploration include even more advanced processing techniques and a full characterisation of the transfer function of the system, with the goal to lead to improvements in intelligibility.
+
+The demonstration presented here has great potential to be adapted for use in the undergraduate laboratory.
+It can be used to demonstrate topics such as interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques to physics and electrical engineering students.
+Given the increased level of excitement and public interest in the field of gravitational-wave astronomy in recent years, this table-top demonstration can be used as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
+The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves, to which this demonstration provides some charming insights.
 
@@
 
-In this paper we present an educational demonstration in which a table-top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
-Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
-We consider quasi-monochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g.
-from supernovae.
-% This serves to demonstrate signal processing techniques.
-
-In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of signals with constant and slowly wandering frequency. 
-A webcam sampling at $30\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered successfully with the Viterbi algorithm. 
-These sections directly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
 %Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer.
 We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.
-
-In Section~\ref{sec:optical_microphone} we demonstrate recovery of more complex signals of music and speech. 
-Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
-We demonstrate that the optical microphone is capable of playing back simple recordings well but produces unintelligible recordings of speech. 
-Mains hum at $50\,{\rm Hz}$ and its harmonics dominates the photodiode output. 
-This demonstration offers students the opportunity to explore a variety of filtering and speech enhancement techniques. 
-Further avenues of exploration include a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.
-
-The demonstration presented here has good potential to be adapted for use in the undergraduate laboratory. 
-It can be used to demonstrate topics of interest to physicists and electrical engineers including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement.
-Given the increased excitement and public interest in gravitational-wave science since the detection in 2015, this table-top demonstration presented here can also be adapted as a tool to explain gravitational-wave research developments for a wider non-specialist audience.
-The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves.
 

### 2020-10-01 paper/ifo-futureWork.tex
Deeksha's comments - last few sections

--- 2020-08-21 paper/ifo-futureWork.tex
@@
 \documentclass[paper-main.tex]{subfiles}
+\begin{document}
 
-\begin{document}
+The optical microphone experiment described here can be used to demonstrate and teach a variety of topics including basic optics and photodiode circuits, to signals processing and speech enhancement.
+Here we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.
+
+An initial improvement could be to better isolate the experiment, especially from the sources of mains noise.
+As the Raspberry Pi demands little power and connects to USB it could instead be powered with a commercial battery rather than the mains supply.
+All other components of the circuit, including the photodiode are connected to the Pi.
+The laser is DC powered through a converter, so should not couple noise from the mains into the interferometer.
+Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}.
+%With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
+
+We can consider improving the stability of the optical microphone components.
+A feedback loop control system can be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}.
+
+We can also benefit from fully characterizing the transfer function of the system as it provides the exact information of how audio signal couples through the interferometer.
+The total transfer function would start with the voltage being sent to the speaker and end with the voltage recorded by the Pi.
+It also would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode.
+The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the signal flow through the system.
+% commented out whilst James is checking this
+% jam: I'm pretty confident this is what Rob said. 
+%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
+%\jam{[check with Rob et al. about the veracity of this]}
+
+A voluminous literature exists on using a hidden Markov model (HMM) trained on phonemes to recognize speech~\cite{HMM_english}.
+These could perform the final stage of speech enhancement for the optical microphone.
+Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a HMM, albeit of a different type.
+Alternatively, machine learning solutions exist throughout the field that are able to compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.
+
+We could also use more traditional techniques such as wavelet transform (WT) \citep{nason1995stationary} to extract the signal from noise, and compare these with the above methods.
+WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time.
+The authors in Refs.~\cite{tufekci2000feature,agbinya1996discrete} also proposed using WT for speech recognition.
+
+Finally, considering the desire to demonstrate continuous GW analysis, further improvement of the optical microphone could lead to more advanced demonstration of techniques such as aiming or locking two or more optical microphones in an array.
 
@@
 	\caption{
-Flowchart for the optical microphone transfer function. 
+Flowchart for the signal through the optical microphone. The transfer function would have to account for each stage. 
 The stage inside the red-dash box is analysed in Appendix~\ref{app:intensity_derivation}. Understanding the other stages is one aim of future work. 
@@
 
-The optical microphone experiment describe here can be used to convey a variety of topics as a teaching tool. 
-Here we consider a selection of further extensions to the experiment touching on noise isolation, transfer function analysis, speech recognition, and control systems.
-
-An initial improvement is to better isolate the experiment from interference, such as the mains supply. 
-The Raspberry Pi is USB powered and requires low power so it could be run off a commercial external phone battery. 
-Everything else in the circuit, including the photodiode, is connected to the Pi. 
-The laser is DC powered through a converter, so should not introduce mains harmonics itself. 
-Transferring the circuit from a breadboard to a printed circuit board will also reduce interference~\cite{elfekey2013design}
-
-%With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
-
-A second improvement is to measure the full model of the transfer function of the system, providing more information about how the audio signal couples with the interferometer.
-% commented out whilst James is checking this
-%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
-%\jam{[check with Rob et al. about the veracity of this]}
-The transfer function starts from the voltage sent to the speaker and ends with the voltage recorded by the Pi.
-It includes any non-linearities in the speaker itself, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode reading.
-A schematic of the full optical microphone transfer function is shown in Figure~\ref{fig:pipeline_highlighted}.
-
-A voluminous literature exists about using a hidden Markov model (HMM) trained against phonemes to recognize speech~\cite{HMM_english}. 
-Interestingly, this connects back to the Viterbi algorithm in Section IV, which is also underpinned by a hidden Markov model, albeit of a different type. 
-Alternatively, machine learning solutions exist throughout the field that are able to compete~\cite{SEGAN} with statistical techniques like the logMMSE estimator.
-
-Other traditional signal processing methods for extracting spectrogram like wavelet transform (WT) \citep{nason1995stationary} can also be implemented and compared with HMM.
-WT provides both time and frequency information, making it easier to pinpoint the origin of noise respect to time.
-In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT was also proposed for speech recognition.
-
-Finally, we can also consider how to improve the stability of the optical microphone components.
-A feedback loop control system can be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}.
-
 %To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.

### 2020-10-01 paper/ifo-wanderingTone.tex
Deeksha's comments - last few sections

@@
 
-Continuous wave searches, such as in Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017,SearchTwoSpecS6:2017,SunEtAlSNR:2018,JonesSun:2020}, are often (not always) performed on long datasets (months to years in duration).
+Continuous wave searches, such as in Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017,SearchTwoSpecS6:2017,SunEtAlSNR:2018,JonesSun:2020}, are often (but not always) performed on long datasets (months to years in duration).
 The frequency of a continuous wave signals may wander significantly over this timescale, due to stochastic internal processes in the superfluid interior of isolated neutron stars~\cite{MelatosDouglassSimula:2015,Jones:2010} or variable accretion flows onto a neutron star from a stellar companion (e.g.
@@
 
-One method to search for a wandering signal is to split the time series into several shorter segments, analyse these individually, then recombine them (under some continuity assumption) to coherently analyse the whole.
-The detection statistics in every segment are combined to form a grid in time and frequency (spectrogram).
-The value at each grid point gives the likelihood that a signal is present at that particular time and frequency.
-In gravitational-wave data analysis, a detection statistic is used to calculate the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
-The detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
+One method to search for a wandering signal is to split the time series data into several shorter segments, analyse these individually, and then recombine them under some continuity assumption to coherently analyse the whole set.
+Most continuous wave searches use detection statistics to quantify the likelihood of a signal being present in the data at each frequency for each time series segment.
+The detection statistic used depends on the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun, and the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
+In this work, we take the discrete Fourier transform of each time series segment and use its amplitude at each frequency as a detection statistic.
+This ensures that we maximize the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}.
+The likelihoods (Fourier amplitudes) from each segment are combined to form a grid in time and frequency (a spectrogram).
+We also normalise the detection statistics for convenience.
 
-In this work the detection statistic is the amplitude (modulus) of the Fourier transform of the data at each time segment, which maximizes the likelihood of detecting a sinusoid in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}.
-The resulting spectrogram gives the Fourier amplitude for each frequency and time bin.
-In continuous wave searches a hidden Markov model is used to describe the system.
-The frequency (the state) of the signal is unknown (hidden) and the signal is Markovian in that the state at any time depends only on the state at the previous time step.
-The hidden Markov model method relates the observable quantity (the detection statistic) to the signal frequency (the hidden state).
-The task is to find the most likely wandering frequency signal given the observable of the spectrogram.
-The Viterbi algorithm~\cite{Viterbi:1967} is an efficient method to find the most likely path of a signal given an observation sequence.
-
-This method is used in continuous wave searches for a range of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,JonesSun:2020,MiddletonEtAlO2LMXBs:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
+We use a hidden Markov model (HMM) to recover the wandering signal from noisy data.
+To do this, we define the frequency of the signal as a hidden (unknown) state which can undergo transitions at discrete times.
+We assume that these transitions are Markovian in that the hidden state (i.e.\ frequency) of the system at any time depends solely on its state at the previous time step.
+The detection statistic (observable) then relates the hidden states of the system to the observed data.
+To perform the task of recovering the signal, we use the Viterbi algorithm~\cite{Viterbi:1967} to efficiently find the most probable path through the spectrogram given the sequence of observations.
+This method is used in continuous wave searches for a variety of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,JonesSun:2020,MiddletonEtAlO2LMXBs:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
 
@@
 
-The spectrogram grid is indexed by time $t$ and frequency $f$ where there are $N_t$ and $N_f$ time and frequency bins respectively, as shown in Fig.~\ref{fig:viterbi}.
-We label the elements of the grid as $t_i$ and $f_j$ with $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-The normalised Fourier amplitude $F(t_i,f_j)$ is the detection statistic indicating the likelihood that a signal is present in each grid element.
+The spectrogram, as shown in Fig.~\ref{fig:viterbi}, has $N_t$ time and $N_f$ frequency bins.
+From here on, we will call these time “steps” instead of time bins, but they should not be confused with the timesteps of the original time series.
+We label each time and frequency step as $t_i$ and $f_j$ with $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$.
+At each state (grid point) the detection statistic (the normalised Fourier amplitude) is notated $F(t_i,f_j)$.
 
-The objective is to find the most likely path through the grid given the data and any probabilistic constraints on how the frequency wanders from $t_i$ to $t_{i+1}$. 
-In continuous wave searches, a physical model is used to decide how much the frequency of the signal is allowed to wander at each time step, which depends on the type of target. 
-In this work, at each time step we allow the frequency to either a) stay in the same bin; b) move up a single frequency bin; or c) move down a single frequency bin. 
-In a hidden Markov model, the probability of transitioning from $f_j$ at $t_i$ to any of $\{f_{j+1},f_j,f_{j-1}\}$ at $t_{i+1}$ is called the transition probability $A(f_k,f_m)$. 
-In this work the three options have equal probability $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
+The objective is to find the most likely path through the grid given the observed data and any probabilistic constraints on how the frequency of the signal can wander between time steps $t_i$ to $t_{i+1}$.
+In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over each time step.
+This information is encoded in the transition probability matrix $A(f_k,f_m)$, which describes the probability of system transitioning from state $f_i$ at $t_i$ to a state $f_{j}$ at $t_{i+1}$.
+Here, we allow the frequency of the signal to either (a) stay in the same bin; (b) move up by a single frequency bin; or (c) move down by a single frequency bin at each time step.
+We assign these three transitions equal probability, i.e.\ $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
 
-A path through the grid has a likelihood dependent on the detection statistic (the observation) and the transition probability.
-If the starting probabilities are all equal, i.e. ${\rm Pr}[f(t_0)] = 1/N_f$, the the probability of the frequency equalling $f(t_{i+1})$ at $t = t_{i+1}$ is defined recursively by
+We define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalised) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
+The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
 \begin{eqnarray}
@@
 
-The Viterbi algorithm is an efficient method to find the Viterbi path. 
-Here we describe the algorithm as pseudo-code referring to the schematic in Fig.~\ref{fig:viterbi}, where the circles indicate the elements of the spectrogram grid (lighter colours correspond to higher likelihood).
+The Viterbi algorithm provides a computationally efficient method for finding the optimal path, at every time step all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}).
+Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}, where the circles indicate the states in the spectrogram with the lighter colours corresponding to a higher likelihood.
 The implementation used in this work is available online (see Appendix~\ref{app:code}).
 \begin{enumerate}
-\item Starting at $t_1$, each $f_j$ element has three possible paths it could originate from at $t_0$, indicated by the lines in Fig.~\ref{fig:viterbi}. 
-At each $f_j$ element, the highest $F[t_0,f_j] A[f(t_0),f(t_1)]$ from the previous time step is selected, highlighted by the black lines in Fig.~\ref{fig:viterbi}. 
-For example for the element labelled (a), the highest value is directly behind it, therefore the transition from $\{t_0,f_0\}$ to $\{t_1,f_0\}$ is selected. 
-Each element has now selected the best path from the previous time step. 
-In Fig.~\ref{fig:viterbi}, black lines show the best paths and grey lines show all other paths.
+\item Starting at time $t_1$, each $f_j$ state can originate from three prior states at time $t_0$ (except for $f_0$ and $f_{N_f}$ which only have two).
+The paths between these states are indicated by the lines in Fig.~\ref{fig:viterbi}.
+At each $f_j$ state, we select the path with the highest $A[f(t_0),f(t_1)] {\rm Pr}[f(t_0)]$ as the most probable path.
+These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the grey lines show the rejected paths.
+For example, the most probable connection to the element labelled (a) is the one directly behind it (i.e., $f_4$).
+Therefore, this path is selected as the best path from $t_0$ to $t_1$.
+For backtracking at the end, the index of the most probable connection along with the current cost of the best path is stored for each node.
 
-\item Moving to $t_2$, the path with the highest value of the recursive relation in Eqn.~\ref{eqn:recursiveViterbi} is selected for each $f_j$ as shown by the solid lines from $t_1$ to $t_2$ in Fig.~\ref{fig:viterbi}.
-The origins are stored for each grid point.
+\item Moving to the third time step, $t_2$, again, we select the path which maximises the recursive quantity in Eqn.~\ref{eqn:recursiveViterbi} for each $f_j$.
+These are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.
 
-\item Steps 2 is repeated until the end of the grid ($t=t_{N_t}$) is reached. 
-Only the best paths are stored.
+\item Step 2 is repeated until the end of the grid ($t=t_{N_t}$) is reached with only the best paths being stored at each iteration.
 
-\item At $t=t_{N_t}$, each $f_j$ has a value ${\rm Pr}[f(t_{N_t})=f_j]$ (Eqn.~\ref{eqn:recursiveViterbi}). 
-The algorithm selects the terminating frequency $f(t_{N_t})$ with the highest value, marked (b) in Fig.~\ref{fig:viterbi}.
+\item Each state $f_j$ at $t=t_{N_t}$ has some value ${\rm Pr}[f(t_{N_t})=f_j]$ (Eqn.~\ref{eqn:recursiveViterbi}), which represents the probability of the most likely path that ends in state $f_j$ at $t_{N_t}$.
+The Viterbi algorithm then selects the terminating frequency bin $f(t_{N_t})$ with the highest probability, marked by (b) in Fig.~\ref{fig:viterbi}.
 
-\item The Viterbi path is then found by backtracking along the stored best paths at each time step (see also Appendix~\ref{app:viterbi}). 
-% come back to this. 
-%\jam{[Andrew mentioned providing a closed form here, which I'm not sure exists.
-The code works by storing three arrays: the weights, the score for the best path to a given vertex, and the index of the previous vertex along that path.
-Backtracking comprises of looping through that third array from the vertex in the last row with the highest score from the second array.
-I don't think we need to explain it any further than that.]}
-In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c). 
-The Viterbi path is highlighted in orange in Fig.~\ref{fig:viterbi}.
+\item The Viterbi path (the overall best path) is then found by backtracking along the stored best connections at each time step (see also Appendix~\ref{app:viterbi}).
+In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c) and is highlighted in orange in Fig.~\ref{fig:viterbi}.
 \end{enumerate}
-The Viterbi algorithm is an example of a dynamical programming algorithm, where a computation can be broken down into a series of sub-computations.
-The Viterbi path for a time series sequences spanning $t_0$ to $t_{N_t}$ contains the Viterbi path for sub-sequences of that time series. 
-The algorithm is computationally efficient, at every time step all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}).
+
+In the following section, we use the Viterbi algorithm to recover a slowly wandering signal.
+
+% The Viterbi algorithm is an example of a dynamical programming algorithm, where a computation can be broken down into a series of sub-computations.
+% The Viterbi path for a time series sequences spanning $t_0$ to $t_{N_t}$ contains the Viterbi path for sub-sequences of that time series. 
+% The algorithm is computationally efficient, at every time step all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}).
 
@@
 
-To test the Viterbi algorithm, a sinusoidal signal with wandering frequency was played through the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}. 
-The injected frequency itself followed a sinusoid with decaying amplitude. 
-This choice was arbitrary, any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
+% throughout this section we are using present tense, should this be changed to past tense?
+
+A sinusoidal signal with a slowly decaying amplitude is injected into the interferometer using the speaker in the setup shown in Fig~\ref{fig:ifo_schematic_webcam} to mimic a slowly wandering signal.
+The particular choice of signal is arbitrary as any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
+The output of the interferometer is recorded via a webcam as in Section~\ref{sec:single_tone}.
+We then test the Viterbi algorithm’s ability to recover the wandering signal.
 
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal.
-The overlaid pink-dot markers show the injected signal. 
-The white-cross markers show the recovered Viterbi path maximised over all terminating frequency bins, which qualitatively provides a good match to the injection. 
-Initially, the injected signal moves faster than the frequency wander limit, leading to the discrepancy observed for $t\lesssim 10\,{\rm s}$. 
-There is also an anomaly at $150\,{\rm s}$, which illustrates what happens when the interferometer is disturbed, e.g. when someone walks nearby. 
-The Viterbi algorithm is robust against such disturbances, recovering the optimal path promptly, as Fig.~\ref{fig:viterbi_overlay} shows.
-
-Quantitatively, the Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time. 
-Another figure of merit is the root mean square (RMS) fractional error $E_{\rm rms}$ along the path defined as 
-\han{}
+The overlaid pink dots show the injected signal while the white crosses show the recovered Viterbi path (the most likely path through the data according to the Viterbi algorithm).
+The recovered Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time.
+We also compute the root mean square (RMS) fractional error $E_{\rm rms}$ along the path which is defined as
 \begin{equation}
@@
 where $I_i$ and $R_i$ are the injected and recovered (Viterbi) frequency paths.
-Computing this for the result shown in Fig.~\ref{fig:viterbi_overlay} we find $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
+Computing this for the result shown in Fig.~\ref{fig:viterbi_overlay}, we find that $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
+
+This may be explained by two anomalies in the recovered path.
+Initially, the injected signal wanders by more than one frequency bin at each time step (i.e.\ faster than the algorithm is allowed to), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
+One may be tempted to increase the allowed frequency wander in the algorithm, however this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.
+There is also an anomaly at $150\,{\rm s}$, which is likely due to a local disturbance, e.g.
+someone walked past the interferometer.
+The Viterbi algorithm is somewhat robust against such disturbances, qualitatively recovering the injected path on the other side, as shown in Fig.~\ref{fig:viterbi_overlay}.
 

### 2020-10-03 paper/ifo-complexAudio.tex
Changrong's quick additions

@@
 
+Although the notch filter removes the mains hum sound (mostly), the filtered recording is not intelligible.
+Qualitatively it sounds more like a drum than a human voice.
+Evidently, this is due to loss of voice information under the filter.
+
+% This effect will become more severe if the original human spectrum overlaps with the hum spectrum by a large amount.
+To overcome this, we turn to more advanced techniques starting with the Wiener filter.
+Instead of just passively filtering different frequencies, this statistical technique optimises an estimate of the original signal.
+% By purely minimizing the "distance" between estimated signal (filtered human voice) and observed signal (original human voice), we hope that we keep signal part undisturbed while filtering out noise.
+
 \begin{figure*}
@@
 By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen visibly in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener}.
+The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone.
+The mains noise is all but removed and more voice information is retained.
+However, the recovered voice still sounds muffled and is not understandable.
 

### 2020-10-05 paper/ifo-complexAudio.tex
pass through text to edit language in a few places, shortened conclusion, added missing exhibition reference

@@
 We describe the additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}.
-Filter and speech enhancement techniques applied to the data are outlined in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse}, respectively.
+Filter and speech enhancement techniques applied to the data are outlined in Section~\ref{sec:filters}.
 This section is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals.
@@
 This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$.
-To overcome this issue, we use a photodiode\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the `reverse' direction.
+To overcome this issue, we use a photodiode\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the reverse direction.
 As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.} at the output of the interferometer to achieve a sampling rate of $16\,{\rm kHz}$.
@@
 
-Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing\footnote{folding of higher frequencies above half the sampling rate} into the detected range.
+Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing\footnote{Folding of frequencies greater than half the sampling rate.} into the detected range.
 We include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
 This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
-We also place a cloth screen over the face of the photodiode to reduce the intensity of the voltage signal as to avoid saturating the ADC.
+We also place a cloth screen over the face of the photodiode to reduce the intensity of the voltage signal and avoid saturating the ADC.
 % It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
@@
 We test the optical microphone with a variety of recordings including speech from different people and music ranging from simple melodies and rhythms to songs.
-During recordings, care must be taken to minimise activity around the interferometer to reduce the amount of environmental noise coupling into the interferometer.
+During recordings, care is taken to minimise activity around the demonstration to reduce the amount of environmental noise coupling into the interferometer.
 The time series data is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function.
@@
 \subsection{A hierarchy of filters}
-\label{sec:simple_filters}
+\label{sec:filters}
 
@@
 The fourth panel of Fig.~\ref{fig:noiseSubtract} shows the spectrum obtained after subtracting the background noise spectrum.
-We see no obvious improvement, which may be attributed to a time-variant noise spectrum.
-Why the noise sources are time-varying is a fact yet to be determined.
+We see no obvious improvement, which may be attributed to a time-variant noise spectrum the cause of which is unidentified.
+%Why the noise sources are time-varying is a fact yet to be determined.
 
-Beyond this naive approach, we need to apply filters to the data.
+Beyond this naive approach, we apply a range of filters to the data.
 In signals processing, the ideal filter would be one that:
@@
 In the above expression, $f_c$ represents the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ represents the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
-Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here, a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}, 3\,{\rm kHz})$.
+Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here, a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}$, $3\,{\rm kHz})$.
 This high frequency cut-off is chosen since the frequencies important for speech generally lie below $2\,{\rm kHz}$~\cite{speech_intelligibility}.
@@
 \subsubsection{Cascade notch filter}
+\begin{figure*}
+\begin{center}
+\includegraphics[width=.9\textwidth]{figures/notch_filter_response}
+\end{center}
+\caption{\label{fig:notchMagResponse}
+Magnitude response for the first five notches of the $15$ cascaded notch filter described by Eqns.~\ref{eqn:notch} and ~\ref{eqn:notch15}. 
+}
+\end{figure*}
+
+One method to remove mains noise and its harmonics is to use a sequence of notch filters centred at each of the frequency bins we want to remove.
+This sequence is known as a cascade of notches.
+Here the notches are smooth in comparison to the naive zeroing of each frequency (which looks like a rectangular comb).
+The complex frequency response of a typical infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
+\begin{equation}
+    \label{eqn:notch}
+    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}
+\end{equation}
+Where $z$ is the complex frequency and $\alpha$ and $\beta$ are parameters that control the filter.
+$w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed) at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which reflects how quickly the response changes around the notched frequency).
+We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centred on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$.
+When then choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
+The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,
+\begin{equation}
+    \label{eqn:notch15}
+    H(z) = \prod_{k=0}^{14} H_k(z).
+\end{equation}
+
+We use the built-in MATLAB filter design toolbox in this work. 
+The amplitude response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
+The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}, respectively.
+We see that the mains harmonics are significantly attenuated in comparison to the results shown in Fig.~\ref{fig:noiseSubtract}.
+
+Although the notch filter removes much of the mains hum sound, the filtered recording is not intelligible.
+Qualitatively it sounds more like a drum than a human voice.
+Evidently, this is due to loss of voice information under the filter.
+
+% This effect will become more severe if the original human spectrum overlaps with the hum spectrum by a large amount.
+To overcome this, we turn to more advanced techniques starting with the Wiener filter.
+Instead of just passively filtering different frequencies, this statistical technique optimises an estimate of the original signal.
+% By purely minimizing the "distance" between estimated signal (filtered human voice) and observed signal (original human voice), we hope that we keep signal part undisturbed while filtering out noise.
 
@@
 
-One method to remove mains noise and its harmonics is to use a sequence of notch filters centered at each of the frequency bins we want to remove.
-This sequence is known as a cascade of notches.
-Here the notches are smooth in comparison to the naive zeroing of each frequency (which looks like a rectangular comb).
-The complex frequency response of a typical infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
-\begin{equation}
-    \label{eqn:notch}
-    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}
-\end{equation}
-Where $z$ is the complex frequency and $\alpha$ and $\beta$ are parameters that control the filter.
-$w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed) at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which reflects how quickly the response changes around the notched frequency).
-We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centred on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$.
-When then choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
-The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,
-\begin{equation}
-    \label{eqn:notch15}
-    H(z) = \prod_{k=0}^{14} H_k(z).
-\end{equation}
-
-We use the built-in MATLAB filter design toolbox, the (amplitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
-The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}, respectively.
-We see that the mains harmonics are significantly attenuated in comparison to the results shown in Fig.~\ref{fig:noiseSubtract}.
-
-Although the notch filter removes the mains hum sound (mostly), the filtered recording is not intelligible.
-Qualitatively it sounds more like a drum than a human voice.
-Evidently, this is due to loss of voice information under the filter.
-
-% This effect will become more severe if the original human spectrum overlaps with the hum spectrum by a large amount.
-To overcome this, we turn to more advanced techniques starting with the Wiener filter.
-Instead of just passively filtering different frequencies, this statistical technique optimises an estimate of the original signal.
-% By purely minimizing the "distance" between estimated signal (filtered human voice) and observed signal (original human voice), we hope that we keep signal part undisturbed while filtering out noise.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=.7\textwidth]{figures/notch_filter_response}
-\end{center}
-\caption{\label{fig:notchMagResponse}
-Magnitude response for the first five notches of the $15$ cascaded notch filter described by Eqns.~\ref{eqn:notch} and ~\ref{eqn:notch15}. 
-}
-\end{figure*}
-
 %A notch filter attenuates signals in specific frequency bands. 
@@
 Let the observed noisy speech signal sequence be given as $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence.
-It is the sum of the original injected (or “clean”) signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$,
+It is the sum of the original injected signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$,
 \begin{equation}
@@
 
-If we assume that: i) $\textbf{x}$ is wide sense stationary~\footnote{ A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all $t_1,t_2 \in R$, if (1) it’s mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
+If we assume that: i) $\textbf{x}$ is wide sense stationary~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, if (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
 \begin{equation}
@@
 where $\textbf{h}=\{h(0),\dots h(n)\}$ represents the coefficients of an $n$th order Wiener filter.
-These assumptions are to be justified by the efficacy of the Wiener filter.
-
+% What this sentence is trying to convey is not clear to me
+%These assumptions are to be justified by the efficacy of the Wiener filter.
 %Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
@@
 %and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
-
 The famous Wiener-Hopf equation~\citep{noble1959methods} allows us to determine $\textbf{h}$, as\begin{equation}
@@
 
-As an instructive aside, if we let the Wiener filter be non-causal (i.e.
+%As an instructive aside, i
+If we let the Wiener filter be non-causal (i.e.
 we estimate the current signal based on both past \emph{and future} observations), then we can represent Eqn.~\ref{eqn:wiener-hopf} in the frequency domain as
@@
 \end{equation}
-where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data (i.e., signal + noise), the injected signal, and the background noise, respectively.
+where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data, the injected signal, and the background noise, respectively.
 Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high, and attenuates the signal where the SNR is low.
@@
 
-What we’re attempting here is not new, speech enhancement of noisy channels is a classic problem in signal processing.
-To borrow from that field, Yi Hu and P.C.
+%probably a bit too causal language 
+%What we’re attempting here is not new, 
+Speech enhancement of noisy channels is a classic problem in signal processing. 
+% shouldn't we make the citations consistent throughout? We don't mention author names up to this point and typically journals have a set style guide. - Hannah 
+%To borrow from that field, Yi Hu and P.C.
 Loizon~\cite{SubjectiveComparison} compared $13$ speech enhancement methods and found the log minimum means-square error (logMMSE) estimator to be the best among them, qualitatively, at recovering speech.
+In Ref.~\cite{SubjectiveComparison} a comparison is made of $13$ speech enhancement methods, finding the log minimum means-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
 This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean square error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that is measures the MSE between the logarithm of the Fourier amplitudes.
@@
 The top two panels of each figure show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison.
-The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} show the time series and corresponding spectral amplitude after applying the logMMSE estimator.
+The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} show the time series and corresponding spectral amplitude respectively after applying the logMMSE estimator.
 We see significant attenuation of the mains harmonics as well as a general smoothing of the spectrum.
-The logMMSE method indeed removes most of the background noise, but does not significantly enhance the speech as the voice sounds muffled and indistinct, with the words unable to be understood.
+The logMMSE method indeed removes most of the background noise, but does not significantly enhance the speech as the voice sounds muffled and indistinct.
+% suggest remove as it's just rephrasing what's already been said
+%, with the words unable to be understood.
 
 We find some improvement with music over speech.
-After applying the logMMSE estimator, simple chords and drums can be heard clearly, however more composite sounds and complex melodies cannot be heard clearly.
+After applying the logMMSE estimator, simple chords and drums can be heard, however more composite sounds and complex melodies cannot be heard clearly.
 Our observations suggest that this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins sometimes can’t be heard at all.
-This could be a perceptual effect or a frequency dependence somewhere in the optical microphone, although the exact cause is yet to be determined.
+This could be a perceptual effect or a frequency dependence somewhere in the optical microphone.
+%, although the exact cause is yet to be determined.  - repetition
 Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums are louder in the results.
@@
 To address these problems, we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-Or if instead they are present in the noisy data but we need better techniques to extract them.
+%Or if instead they are present in the noisy data but we need better techniques to extract them.
+- should probably justify what you mean by better if you are going to say it here, but I think enough is discussed in section VI for it not to be necessary here - hannah
 To determine this requires a better understanding of the system as discussed in Section~\ref{sec:future_work}.

### 2020-10-05 paper/ifo-conclusions.tex
pass through text to edit language in a few places, shortened conclusion, added missing exhibition reference

@@
 
-In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector and demonstrate a range of signal processing techniques used within the GW community and more broadly.
-Gravitational-wave signals are simulated by playing sound recordings through a speaker which is fixed to the back of one of the arm mirrors.
+In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector, demonstrating a range of signal processing techniques used within the gravitational-wave community and more broadly.
+Gravitational-wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors.
 We consider quasi-monochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music for more general interest.
@@
 
-We demonstrate the recovery of a signal with constant frequency and signal with slowly wandering frequency in Sections~\ref{sec:ifo} and ~\ref{sec:single_tone}, respectively.
+% reduced detail for conclusion (trying to reduce overall length)
+In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} we analyse audio signals with constant and wandering frequencies respectively as analogues to continuous gravitational wave signals. 
+The wandering frequency signal is recovered using using the Viterbi algorithm. 
+These sections directly demonstrate the techniques used in current continuous gravitational-wave searches. 
+%We demonstrate the recovery of a signal with constant frequency and signal with slowly wandering frequency in Sections~\ref{sec:ifo} and ~\ref{sec:single_tone}, respectively.
 We use a basic webcam with a sampling rate of $30\,{\rm Hz}$ to track the interference pattern.
@@
 
-In Section~\ref{sec:optical_microphone}, we demonstrate the recovery of more complex signals of music and speech.
+% again reducing overall length as we probably don't need to list everything
+The recovery of more complex music and speech signals are demonstrated in Section~\ref{sec:optical_microphone}, where we treat the interferometer as an optical microphone. 
+Simple music recordings can be recognised when played back, however speech remains unintelligible after applying a selection of filters and statistical techniques from signal processing. 
+We find that the logMMSE estimator produces the best result of those applied here, however it does not enhance the recording to an intelligible level. 
+This demonstration allows students to explore a variety of filtering and speech enhancement techniques. 
+Further avenues of exploration to improve intelligibility are detailed in Section~\ref{sec:future_work}. 
+%In Section~\ref{sec:optical_microphone}, we demonstrate the recovery of more complex signals of music and speech.
 By substituting the webcam with a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone by greatly increasing the sampling rate.
@@
 
-The demonstration presented here has great potential to be adapted for use in the undergraduate laboratory.
+The demonstration presented has potential to be adapted for use in the undergraduate laboratory.
+It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
+The increased excitement and public interest in the field of gravitational-wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
+The field of gravitational-wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals such as continuous waves to which this demonstration provides some charming insights.
+%The demonstration presented here has great potential to be adapted for use in the undergraduate laboratory.
 It can be used to demonstrate topics such as interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques to physics and electrical engineering students.

### 2020-10-05 paper/ifo-futureWork.tex
pass through text to edit language in a few places, shortened conclusion, added missing exhibition reference

@@
 
-An initial improvement could be to better isolate the experiment, especially from the sources of mains noise.
-As the Raspberry Pi demands little power and connects to USB it could instead be powered with a commercial battery rather than the mains supply.
-All other components of the circuit, including the photodiode are connected to the Pi.
+% reduced length
+Isolating the experiment from interference may provide some reduction in noise. 
+For example, the Raspberry Pi could be powered with a commercial battery instead of mains power. 
+%An initial improvement could be to better isolate the experiment, especially from the sources of mains noise. 
+%As the Raspberry Pi demands little power and connects to USB it could instead be powered with a commercial battery rather than the mains supply. 
+All other components of the circuit are connected to the Pi. 
+%All other components of the circuit, including the photodiode are connected to the Pi. 
 The laser is DC powered through a converter, so should not couple noise from the mains into the interferometer. 
@@
 %With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
-
-We can consider improving the stability of the optical microphone components.
+A feedback loop control system can also be used to suppress unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational wave observatories.
+%We can consider improving the stability of the optical microphone components.
 A feedback loop control system can be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}.
-
-We can also benefit from fully characterizing the transfer function of the system as it provides the exact information of how audio signal couples through the interferometer.
-The total transfer function would start with the voltage being sent to the speaker and end with the voltage recorded by the Pi.
-It also would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode.
-The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the signal flow through the system.
-% commented out whilst James is checking this
-% jam: I'm pretty confident this is what Rob said. 
-%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
-%\jam{[check with Rob et al. about the veracity of this]}
-
-A voluminous literature exists on using a hidden Markov model (HMM) trained on phonemes to recognize speech~\cite{HMM_english}.
-These could perform the final stage of speech enhancement for the optical microphone.
-Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a HMM, albeit of a different type.
-Alternatively, machine learning solutions exist throughout the field that are able to compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.
-
-We could also use more traditional techniques such as wavelet transform (WT) \citep{nason1995stationary} to extract the signal from noise, and compare these with the above methods.
-WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time.
-The authors in Refs.~\cite{tufekci2000feature,agbinya1996discrete} also proposed using WT for speech recognition.
-
-Finally, considering the desire to demonstrate continuous GW analysis, further improvement of the optical microphone could lead to more advanced demonstration of techniques such as aiming or locking two or more optical microphones in an array.
 
@@
 
+A thorough characterisation of the systems transfer function would allow us to understand how audio signals couple through the interferometer. 
+%We can also benefit from fully characterizing the transfer function of the system as it provides the exact information of how audio signal couples through the interferometer. 
+The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Pi.
+It also would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode.
+The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the signal flow through the system.
+% commented out whilst James is checking this
+% jam: I'm pretty confident this is what Rob said. 
+%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
+%\jam{[check with Rob et al. about the veracity of this]}
+
+A voluminous literature exists on using HMMs trained on phonemes to recognize speech~\cite{HMM_english}.
+These could perform the final stage of speech enhancement for the optical microphone.
+Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a HMM, albeit of a different type.
+Alternatively, machine learning solutions exist throughout the field that are able to compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.
+
+We could also use more traditional techniques such as wavelet transform (WT) \citep{nason1995stationary} to extract the signal from noise, and compare these with the above methods.
+WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time.
+In Refs.~\cite{tufekci2000feature,agbinya1996discrete} WT methods are proposed for speech recognition.
+
+Returning to the demonstration of continuous gravitational wave analysis, further additions could include analysing data from multiple interferometers to extract directional information for audio signals.
+%Finally, considering the desire to demonstrate continuous wave analysis, further improvements to the optical microphone could lead to more advanced demonstration of techniques such as aiming or locking two or more optical microphones in an array.
+% I think locking is the wrong word to use here as it is a word used a lot in GW instrumentation (an interferometer can be 'locked' on its own).
+I've tried to rephrase based on what I think you are getting at.
+-hannah
+
 %To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.

### 2020-10-05 paper/ifo-introduction.tex
Deeksha's comments: single and wandering tone
pass through text to edit language in a few places, shortened conclusion, added missing exhibition reference

@@
 % What are gws and gw detectors 
-Gravitational waves, predicted by Albert Einstein in his theory of general relativity~\cite{Einstein:1916}, are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
+Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}, 
+They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
 The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects. 
-Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (aLIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
+Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
 Gravitational-wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
@@
 % CW 
-To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of 
-binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018}, 
-binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018} and recently an observation of a black hole with an object that could either have been a neutron star or a black hole~\cite{GW190814}.
+To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018}, binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018}, and recently the merger of black hole with an object that could either have been a neutron star or a black hole~\cite{GW190814}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
@@
 
-In this paper, we present a table-top Michelson interferometer as a `toy' gravitational-wave detector designed to detect sound instead of gravitational waves. 
-The demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
+% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
+In this paper, we use a table-top Michelson interferometer as a `toy' gravitational-wave detector designed to detect sound instead of gravitational waves. 
+% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
+As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement. 
 %% I don't like the use of ``freedom to explore''. 
-Additionally, it allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
+% removed Additionally - I don't think it is neccessary - hannah
+It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
 % This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
-This demonstration also has the potential to be used as an outreach tool alongside other exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational wave research groups around the world.
-These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational wave science to a non-specialist audience.
+% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
+This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world.
+These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to a non-specialist audience.
 %\han{look up Sarah Farmer's project}
 
-The design of the interferometer is detailed in Section~\ref{sec:ifo} and we demonstrate observing a single tone from a speaker in Section~\ref{sec:single_tone}. 
+This paper is laid out as follows. 
+In Section~\ref{sec:ifo} we detail the table-top interferometer design. 
+We demonstrate observing a single tone from a speaker in Section~\ref{sec:single_tone}. 
 In Section~\ref{sec:viterbi_wandering} we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous wave searches.
-In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech, to varying degrees of success.
+In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
+% removed as there's not space to really say what this means here: to varying degrees of success.
 This demonstration of an ‘optical microphone’ (using light to capture sound) serves as a more general exhibition of signal processing. 

### 2020-10-05 paper/ifo-singleTone.tex
pass through text to edit language in a few places, shortened conclusion, added missing exhibition reference

@@
 
-Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone; a signal at a single constant frequency.
+% rearranged as I found this section to switch between method and results too much, I've tried this order now: intro, webcam and data specifics, injection and results.
+
+Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone; a single constant frequency.
 
 As described in Section~\ref{sec:ifo}, the audio signal is played through a speaker fixed to the back of M2. 
-The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam (see Fig.~\ref{fig:ifo_schematic_webcam}).
+The intensity of the interference pattern is measured at a single point on the screen, indicated by the pink star in Fig~\ref{fig:interference_pattern}. 
+The webcam records video in three colour channels; red, blue, and green. 
+We take the green channel as an approximation of the total intensity produced by the monochromatic green laser.
+The webcam samples at a rate of $30\, {\rm Hz}$, which limits the spectral content of observable signals to less than $15\,{\rm Hz}$, the Nyquist frequency.
+
+%The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam (see Fig.~\ref{fig:ifo_schematic_webcam}).
 The webcam samples at a rate of $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
-A tone with a frequency of 2.09Hz is played through the speaker for one minute and the interference pattern is recorded by the webcam.
+A tone with a frequency of $2.09\,{\rm Hz$} is played through the speaker for one minute and the interference pattern is recorded by the webcam.
 
-We take data from an off-centre pixel as indicated by the pink star in Fig.~\ref{fig:interference_pattern}.
-The pixel intensity in the green-channel of the video (the webcam records in three colour channels: red, blue, and green) is used to approximate the total intensity of the interference pattern.
-This is because the laser emits monochromatic green light at $532\,{\rm nm}$.
-The amplitude of the Fourier transform of the recorded signal is shown in Fig.~\ref{fig:webcam_spectrum}.
-We measured a strong signal at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$.
-Although this range doesn't contain our injected frequency, the error is not substantial and cannot be heard by the (untrained) human ear.
+%The pixel intensity in the green-channel of the video (the webcam records in three colour channels: red, blue, and green) is used to approximate the total intensity of the interference pattern. 
+%This is because the laser emits monochromatic green light at $532\,{\rm nm}$.
+A tone with a frequency of $2.09\,{\rm Hz}$ is played through the speaker for one minute. 
+The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}. 
+% removed 'strong' as what does strong mean here?
+We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$. 
+%Although this range doesn't contain our injected frequency, the error is not substantial and cannot be heard by the (untrained) human ear.
 Two harmonics can also be seen at integer multiples of the peak frequency. 

### 2020-10-05 paper/ifo-tableTopGWs.tex
pass through text to edit language in a few places, shortened conclusion, added missing exhibition reference

@@
 
-Gravitational-wave detectors such as aLIGO and aVirgo are large, complex experiments.
+% rephrased to mention the laser as the first component and shifted the arm lengths of LIGO and Virgo to later in the para when we discuss them. 
+Gravitational-wave detectors such as LIGO and Virgo are large, complex experiments. 
+However their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
+In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms (see Fig.~\ref{fig:ifo_schematic_webcam}). 
+Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern, 
+The resulting interference pattern is dependent on the relative distance travelled by the beams. 
+Current generation gravitational-wave observatories have ${\rm km}$-scale arms ($4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
+They are able to measure minuscule changes in distance due to gravitational-waves, e.g.
+the first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller than the width of a proton.
+
+\begin{comment}
+Gravitational-wave detectors such as LIGO and Virgo are large, complex experiments.
 However, their design is based fundamentally on a Michelson interferometer, an optical configuration commonly used in undergraduate laboratories.
-It consists of a beam splitter which splits light from a laser down two long perpendicular ``arms'' ($4\,{\rm km}$ long at aLIGO, $3\,{\rm km}$ at aVirgo).
+It consists of a beam splitter which splits light from a laser down two perpendicular arms. % I suggest moving this to later as we are talking about Michelson interferometers generally here. 
+%($4\,{\rm km}$ and $3\,{\rm km}$ long at LIGO and Virgo respectively).
 Mirrors at the end of each arm reflect the light back towards the beam splitter, where they recombine and produce an interference pattern.
 This pattern is dependent on the relative distance travelled by each beam and thus on the relative length of each arm. 
-The current generation gravitational-wave observatories use this technique, in combination with others, to detect minuscule changes in distance of two arms due to a passing gravitational wave.
-For example, the first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller with the width of a proton.
+Current generation gravitational-wave observatories are able to measure minuscule changes in distance due to passing gravitational waves. 
+The first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller than the width of a proton.
+\end{comment}
 
-The complex detection and analysis techniques used in the gravitational wave community can be demonstrated using a table-top experiment, such as the one presented here.
-Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} provide a selection of other table-top Michelson interferometer designs used for science communication.
-In addition to this, gravitational-wave signals from binary black hole and neutron star mergers are often converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
-For example, gravitational waves can propagate in a vacuum, whereas sound cannot.} to aid in explanations.~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
-This analogue allows scientists to use sound vibrations as a simple means to move the components of a demonstration interferometer.
-The vibrations move the mirrors, changing the length of the interferometer arms, and thus changing in the interference pattern.
-We use audio signals throughout this work to simulate gravitational-wave-like signals and to demonstrate different data analysis and signal processing techniques used to recover such signals.
+Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} provide a selection of other table-top Michelson interferometer designs used for science communication).
+% I rephrased to lead with the analogy here and removed 'often' as GW signals are only really listened to for public engagement interest - hannah
+A commonly-used analogy when explaining gravitational-wave science is sound. 
+Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
+For example, gravitational waves can propagate in a vacuum, whereas sound cannot.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
+% rephrased as this what we are doing, rather than what scientists in general are doing
+%This analogue allows scientists to use sound vibrations as a simple means to move the components of a demonstration interferometer. 
+Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
+The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
+We use audio signals throughout this work to simulate gravitational-wave-like signals.
+% and  to demonstrate different data analysis and signal processing techniques used to recover such signals.
 
@@
 The optical configuration of the Michelson interferometer used in this work is shown in Fig.~\ref{fig:ifo_schematic_webcam}.
-It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses as a source a green eye-safe laser (top-left) with peak emission at $532\,{\rm nm}$.
-The output of the laser is first reflected off a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
-After passing through a converging (plano-convex) lens of focal length (FL = $125.0\,{\rm mm}$), the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2).
+It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green eye-safe laser (top-left) with peak emission at a wavelength of $532\,{\rm nm}$.
+The output of the laser is first reflected by a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
+After passing through a converging (plano-convex) lens with focal length $125.0\,{\rm mm}$, the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2).
 A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations. 
@@
 The other speaker in the pair is kept face-down and as far away from the interferometer as possible to prevent it from acting as a second source.
-The beams are reflected by mirrors M1 and M2, located at the end of (approximately) $7.5\,{\rm cm}$ and $10.0\,{\rm cm}$ long arms.
-The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging (bi-concave) lens of FL = $-25.0\,{\rm mm}$, and projected onto a screen.
+The beams are reflected by mirrors M1 and M2, located at the end of $\approx 7.5\,{\rm cm}$ and $\approx 10.0\,{\rm cm}$ long arms.
+The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging (bi-concave) lens of focal length $-25.0\,{\rm mm}$, and projected onto a screen.
 The interference pattern, as shown by the photograph in Fig.~\ref{fig:interference_pattern}, is a set of concentric light and dark rings (fringes). 
@@
 % Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
-An intensity time series of some point in the interference pattern is recorded by a webcam (or later, a photodiode) for data analysis to extract the signal injected at the speaker.
+% Why remove the sentence above? Rings are chosen rather than happening by default. - Hannah
+The intensity timeseries is recorded via webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
+In Section~\ref{sec:optical_microphone}, which considers more complex audio, the webcam is replaced with a photodiode, which offers a higher sample rate than the webcam. 
+%rephrased as this sentence is confusing I think: An intensity time series of some point in the interference pattern is recorded by a webcam (or a photodiode in Section~\ref{sec:optical_microphone}) for data analysis to extract the signal injected at the speaker.
 
@@
 If the relative change in the length of the arms is kept small enough, the intensity of the interference pattern at any point on the screen oscillates at the same frequency as the injected audio.
-But if the relative change in the length of the arms is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting (artificially) raises the measured frequency.
+But if the relative change in the length of the arms is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting artificially raises the measured frequency.
 As such, any motion of the fringes must be kept small by playing the sound softly through the speaker.
-Even without over-counting, large fringe motions display a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting---phenomena (e.g frequency doubling).
+% too much to have --- and () in the same sentence I think 
+Even without over-counting, large fringe motions display a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting---phenomena like frequency doubling.
 We derive this nonlinear relationship between the length difference of the arms and the intensity of the pattern at some point on the screen in Appendix~\ref{app:intensity_derivation}.
 
-In the following section, we start simply by injecting and recovering a audio signal of constant frequency (a single tone).
+% removed as almost identical to the next sentence
+%In the following section, we start simply by injecting and recovering a audio signal of constant frequency (a single tone).
 

### 2020-10-05 paper/ifo-wanderingTone.tex
pass through text to edit language in a few places, shortened conclusion, added missing exhibition reference

@@
 \begin{document}
+% I've rephrased the first part of this section a lot and provide some justification below
+% I think we need to lead with the details of the signal here (similar to the last section), not the specifics of the data analysis as the issues seem confused here.
+The fact that long data sets are analysed is not because of the signal wandering, if anything that's an argument against analysing long data sets to make detection easier, it's because the signal has low amplitude.
+Also what is the justification of the "(but not always)" statement, do you have a reference?
+- hannah
 
+\begin{comment}
 Continuous wave searches, such as in Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017,SearchTwoSpecS6:2017,SunEtAlSNR:2018,JonesSun:2020}, are often (but not always) performed on long datasets (months to years in duration).
@@
 Here, we adapt the data analysis technique used to search for slowly wandering continuous waves signals from spinning neutron stars in aLIGO and aVirgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
+\end{comment}
 
-One method to search for a wandering signal is to split the time series data into several shorter segments, analyse these individually, and then recombine them under some continuity assumption to coherently analyse the whole set.
-Most continuous wave searches use detection statistics to quantify the likelihood of a signal being present in the data at each frequency for each time series segment.
-The detection statistic used depends on the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun, and the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
-In this work, we take the discrete Fourier transform of each time series segment and use its amplitude at each frequency as a detection statistic.
-This ensures that we maximize the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}.
-The likelihoods (Fourier amplitudes) from each segment are combined to form a grid in time and frequency (a spectrogram).
-We also normalise the detection statistics for convenience.
+% the signal and the analogy we use
+Continuous gravitational waves may not be monochromatic, unlike the signal considered in Section~\ref{sec:single_tone}.  
+A continuous-wave signal may wander slowly in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
+The audio analogue is a tone that wanders in frequency: a note that changes in pitch.
 
-We use a hidden Markov model (HMM) to recover the wandering signal from noisy data.
-To do this, we define the frequency of the signal as a hidden (unknown) state which can undergo transitions at discrete times.
-We assume that these transitions are Markovian in that the hidden state (i.e.\ frequency) of the system at any time depends solely on its state at the previous time step.
-The detection statistic (observable) then relates the hidden states of the system to the observed data.
-To perform the task of recovering the signal, we use the Viterbi algorithm~\cite{Viterbi:1967} to efficiently find the most probable path through the spectrogram given the sequence of observations.
+% the data analysis techniques we draw from 
+The analysis techniques used in this section are inspired by those used to search for continuous waves from spinning neutron stars in LIGO and Virgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
+Continuous wave searches are performed on long datasets, months to years in duration. 
+The frequency of the signal can wander significantly over the observation period. 
+In this context significantly means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
+One method to search for a wandering signal is to split the time series data into several shorter segments which are analysed individually.%, and then recombine them under some continuity assumption to coherently analyse the whole set.
+% I think the following is more accurate. The antenna beam pattern is used to compute a detection statistic. Which detection statistic to use depends on the type of target. -Hannah
+A detection statistic is used to quantify the likelihood of a signal being present in the data at each frequency and timeseries segment. 
+In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
+%Most continuous wave searches use detection statistics to quantify the likelihood of a signal being present in the data at each frequency for each timeseries segment. 
+%The detection statistic used depends on the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun, and the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
+Which detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
+In this work, we take the discrete Fourier transform of each timeseries segment and use its amplitude at each frequency as a detection statistic.
+This maximises the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}. 
+The detection statistic is normalised for convenience. 
+The detection statistics from each segment are combined to form a grid in time and frequency (a spectrogram). 
+% moved one sentence up 
+%We also normalise the detection statistics for convenience.
+
+Following the method of Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, we use a hidden Markov model (HMM) to recover the wandering signal from noisy data. 
+In a HMM, the frequency state of the signal is unknown, or hidden, and can undergo transitions at discrete times. 
+%To do this, we define the frequency of the signal as a hidden (unknown) state which can undergo transitions at discrete times. 
+The transitions are Markovian in that the hidden state (i.e.\ frequency) of the system at any time depends solely on its state at the previous time.
+The detection statistic (observable) relates the hidden states of the system to the observed data.
+The Viterbi algorithm~\cite{Viterbi:1967} is used to find the most probable sequence of states given the sequence of observables.
+%To perform the task of recovering the signal, we use the Viterbi algorithm~\cite{Viterbi:1967} to efficiently find the most probable path through the spectrogram given the sequence of observations.
 This method is used in continuous wave searches for a variety of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,JonesSun:2020,MiddletonEtAlO2LMXBs:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
@@
 In Section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work. 
-In Section~\ref{sec:wanderingResults} we present the results for recovery of a wandering audio frequency.
+In Section~\ref{sec:wanderingResults} we present results for recovery of a wandering audio frequency.
 
@@
 The spectrogram, as shown in Fig.~\ref{fig:viterbi}, has $N_t$ time and $N_f$ frequency bins. 
-From here on, we will call these time “steps” instead of time bins, but they should not be confused with the timesteps of the original time series.
-We label each time and frequency step as $t_i$ and $f_j$ with $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$.
-At each state (grid point) the detection statistic (the normalised Fourier amplitude) is notated $F(t_i,f_j)$.
+% why add to the confusion? why not make things consistent? 
+%From here on, we will call these time “steps” instead of time bins, but they should not be confused with the timesteps of the original time series.
+Reading through it seems we were using both steps and bins in this section anyway.
+We label the elements of the time and frequency  grid as $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
+% state is means frequency not grid point doesn't it? 
+%At each state (grid point) the detection statistic (the normalised Fourier amplitude) is notated $F(t_i,f_j)$.
+The normalised Fourier amplitude $F(t_i,f_j)$ is the detection statistic.
 
-The objective is to find the most likely path through the grid given the observed data and any probabilistic constraints on how the frequency of the signal can wander between time steps $t_i$ to $t_{i+1}$.
-In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over each time step.
+The objective is to find the most likely path through the grid given the observed data and any probabilistic constraints on how the frequency of the signal can wander from $t_i$ to $t_{i+1}$.
+In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over at each time.
 This information is encoded in the transition probability matrix $A(f_k,f_m)$, which describes the probability of system transitioning from state $f_i$ at $t_i$ to a state $f_{j}$ at $t_{i+1}$.
-Here, we allow the frequency of the signal to either (a) stay in the same bin; (b) move up by a single frequency bin; or (c) move down by a single frequency bin at each time step.
+Here, we allow the frequency of the signal to either (a) stay in the same bin; (b) move up by a single frequency bin; or (c) move down by a single frequency bin at each transition.
 We assign these three transitions equal probability, i.e.\ $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
@@
 The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes ${\rm Pr}[f(t_{N_t}) = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
-We can the maximize the latter quantity over $0 \leq j \leq N_f$ to find the optimal Viterbi path overall, i.e. terminating in any frequency bin.
+We can then maximize the latter quantity over $0 \leq j \leq N_f$ to find the optimal Viterbi path overall, i.e. terminating in any frequency bin.
 
-The Viterbi algorithm provides a computationally efficient method for finding the optimal path, at every time step all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}).
-Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}, where the circles indicate the states in the spectrogram with the lighter colours corresponding to a higher likelihood.
+The Viterbi algorithm provides a computationally efficient method for finding the optimal path. 
+At every $t_i$ all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}). 
+Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}, where the circles indicate the elements in the spectrogram with the lighter colours corresponding to a higher likelihood.
 The implementation used in this work is available online (see Appendix~\ref{app:code}).
@@
 
-\item Moving to the third time step, $t_2$, again, we select the path which maximises the recursive quantity in Eqn.~\ref{eqn:recursiveViterbi} for each $f_j$.
+\item Moving to time $t_2$, again, we select the path which maximises the recursive quantity in Eqn.~\ref{eqn:recursiveViterbi} for each $f_j$.
 These are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.
@@
 
-\item The Viterbi path (the overall best path) is then found by backtracking along the stored best connections at each time step (see also Appendix~\ref{app:viterbi}).
-In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c) and is highlighted in orange in Fig.~\ref{fig:viterbi}.
+\item The Viterbi path (the overall best path) is then found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}).
+In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c) and is highlighted in orange.
 \end{enumerate}
@@
 The colour of each circle indicates the likelihood that a signal is present in the box, where a lighter colour corresponds to higher likelihood. 
-AT each time step, the paths can move up one $f$ step, move down one $f$ step, or stay at the same $f$ step (shown by lines in the diagram). 
-Each circles has three paths leading to it indicated by the lines. 
+At each time, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines in the diagram). 
+Each circle has three paths leading to it indicated by the lines. 
 The best path to each circle at each $t_i$ is highlighted in black. 
@@
 
-\subsection{Sample output}
+\subsection{Wandering frequency results}
 \label{sec:wanderingResults}
@@
 Recovery of a wandering tone. 
-The spectrogram shows the observed frequency amplitude at each time-frequency step. 
+The spectrogram shows the observed frequency amplitude in each time-frequency bin. 
 The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path respectively. 
@@
 % throughout this section we are using present tense, should this be changed to past tense?
+% I think we should go for present tense throughout - hannah
 
-A sinusoidal signal with a slowly decaying amplitude is injected into the interferometer using the speaker in the setup shown in Fig~\ref{fig:ifo_schematic_webcam} to mimic a slowly wandering signal.
+A sinusoidal signal with slowly decaying amplitude is played through the speaker to mimic a slowly wandering signal. 
+We use an identical set up to that shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
+%A sinusoidal signal with a slowly decaying amplitude is injected into the interferometer using the speaker in the setup shown in Fig~\ref{fig:ifo_schematic_webcam} to mimic a slowly wandering signal.
 The particular choice of signal is arbitrary as any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
 The output of the interferometer is recorded via a webcam as in Section~\ref{sec:single_tone}.
-We then test the Viterbi algorithm’s ability to recover the wandering signal.
+In this section we test the Viterbi algorithm’s ability to recover the wandering signal.
 
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
-The overlaid pink dots show the injected signal while the white crosses show the recovered Viterbi path (the most likely path through the data according to the Viterbi algorithm).
+The overlaid pink dots show the injected signal and the white crosses show the recovered Viterbi path.% I think this is already defined and seems like repetition:(the most likely path through the data according to the Viterbi algorithm).
 The recovered Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time.
@@
 This may be explained by two anomalies in the recovered path.
-Initially, the injected signal wanders by more than one frequency bin at each time step (i.e.\ faster than the algorithm is allowed to), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
+Initially, the injected signal wanders by more than one frequency bin per time bin (i.e.\ faster than the algorithm is allowed to), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
 One may be tempted to increase the allowed frequency wander in the algorithm, however this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.
 There is also an anomaly at $150\,{\rm s}$, which is likely due to a local disturbance, e.g.
-someone walked past the interferometer.
-The Viterbi algorithm is somewhat robust against such disturbances, qualitatively recovering the injected path on the other side, as shown in Fig.~\ref{fig:viterbi_overlay}.
+someone walking past the interferometer.
+As shown in Fig.~\ref{fig:viterbi_overlay}, the Viterbi algorithm is able to recover and continues to track the signal after the disturbance. 
+%The Viterbi algorithm is somewhat robust against such disturbances, qualitatively recovering the injected path on the other side, as shown in Fig.~\ref{fig:viterbi_overlay}.
 

### 2020-10-06 paper/ifo-complexAudio.tex
combined notch and Wiener plots

@@
 The amplitude response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
-The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the third panel of Fig.~\ref{fig:timeNotchWiener} and the first panel of Fig.~\ref{fig:freqNotchWiener}, respectively.
+The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fourth row of Fig.~\ref{fig:NotchWienerCombined}.
 We see that the mains harmonics are significantly attenuated in comparison to the results shown in Fig.~\ref{fig:noiseSubtract}.
@@
 	\begin{center}
-\includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/timeSeriesNotchWiener.pdf}
+	\includegraphics[width=\textwidth]{figures/notch_and_wiener_superplot.pdf}
 	\end{center}
-\caption{\label{fig:timeNotchWiener}
-Timeseries results for the notch and Wiener filters. 
-The first panel shows the original speech recording which was played from the speaker. 
-The second panel shows the data from the optical microphone before any filters are applied. 
-The third and fourth panels show results after applying the notch and Wiener filter respectively. 
-Finally, the fifth panel shows the result of applying both the notch and Wiener filter. 
+	\caption{\label{fig:NotchWienerCombined}
+	Time series (left column) and frequency spectrum (right column) results for the notch and Wiener filters.
+	The first row shows the original speech recording.
+	The second row shows the raw output from the optical microphone.
+	The third row shows the naive technique of subtracting off the noise spectrum from the recorded spectrum.
+	The fourth, fifth, and sixth rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
 	}
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth,trim={0cm 8.2cm 0cm 8cm},clip]{figures/freqNotchWiener.pdf}
-\end{center}
-\caption{\label{fig:freqNotchWiener}
-Frequency spectrum results for the notch and Wiener filters. 
-The first and second panels show the results of applying the notch and Wiener filter respectively.
-The third panel shows the spectrum after applying both the notch and Wiener filters. 
-The first, second, and third panels show the spectrum of the timeseries show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
-}
-\end{figure*}
+% \begin{figure*}
+% \begin{center}
+% \includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/timeSeriesNotchWiener.pdf}
+% \end{center}
+% \caption{\label{fig:timeNotchWiener}
+% Timeseries results for the notch and Wiener filters. 
+% The first panel shows the original speech recording which was played from the speaker. 
+% The second panel shows the data from the optical microphone before any filters are applied. 
+% The third and fourth panels show results after applying the notch and Wiener filter respectively. 
+% Finally, the fifth panel shows the result of applying both the notch and Wiener filter. 
+% }
+% \end{figure*}
+
+% \begin{figure*}
+% \begin{center}
+% \includegraphics[width=\textwidth,trim={0cm 8.2cm 0cm 8cm},clip]{figures/freqNotchWiener.pdf}
+% \end{center}
+% \caption{\label{fig:freqNotchWiener}
+% Frequency spectrum results for the notch and Wiener filters. 
+% The first and second panels show the results of applying the notch and Wiener filter respectively.
+% The third panel shows the spectrum after applying both the notch and Wiener filters. 
+% The first, second, and third panels show the spectrum of the timeseries show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
+% }
+% \end{figure*}
 
@@
 % What this sentence is trying to convey is not clear to me
+%% just trying to say that we can make the assumptions that we do because the Wiener filter appears to work well under them - J
 %These assumptions are to be justified by the efficacy of the Wiener filter.
@@
 For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
-The effects of applying the Wiener filter to the time series and frequency spectrum are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener}, respectively.
+The time series and frequency spectrum after applying the Wiener filter are shown in the fifth row of Fig.~\ref{fig:NotchWienerCombined}.
 We see a significant improvement in the time series of the recovered signal, however a strong noise hum persists, audibly. 
@@
 
-By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen visibly in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener}.
+By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:NotchWienerCombined}.
 The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone.

### 2020-10-06 paper/paper_main.tex
updating affiliation for James

@@
 
-\usepackage{lineno}
-\linenumbers
+%for line number
+%\usepackage{lineno}
+%\linenumbers
 
@@
 \affiliation{%
-College of Science, Australian National University, Acton, ACT, 2601, Australia 
+ ANU Centre for Gravitational Astrophysics, The Australian National University, Acton, ACT, 2601.
+ %College of Science, Australian National University, Acton, ACT, 2601, Australia 
 }%
+\affiliation{%
+ OzGrav-ANU, Australian Research Council Centre of Excellence for Gravitational-wave Discovery, The Australian National University, Acton, ACT, 2601.
+}
 \affiliation{%

### 2020-10-06 paper/paper_main.tex
affliation change for James

--- 2020-09-28 paper/paper_main.tex
@@
 \email{u6069809@anu.edu.au}
+% \affiliation{%
+% College of Science, Australian National University, Acton, ACT, 2601, Australia 
+% }%
 \affiliation{%
-College of Science, Australian National University, Acton, ACT, 2601, Australia 
-}%
+ANU Centre for Gravitational Astrophysics, The Australian National University, Acton, ACT, 2601, Australia
+}
+\affiliation{%
+OzGrav-ANU, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, The Australian National University, Acton, ACT, 2601, Australia
+} 
 \affiliation{%

### 2020-10-29 paper/ifo-tableTopGWs.tex
adding comments from Lynn

@@
 They are able to measure minuscule changes in distance due to gravitational-waves, e.g.
-the first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller than the width of a proton.
+the first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$, a distance smaller than the width of a proton~\cite{GW150914}.
 
@@
 % I rephrased to lead with the analogy here and removed 'often' as GW signals are only really listened to for public engagement interest - hannah
-A commonly-used analogy when explaining gravitational-wave science is sound. 
+Sound is a commonly-used analogy when explaining gravitational-wave science. 
 Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
-For example, gravitational waves can propagate in a vacuum, whereas sound cannot.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
+For example, gravitational waves can propagate in a vacuum, whereas sound cannot.
+Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
 % rephrased as this what we are doing, rather than what scientists in general are doing

### 2020-10-29 paper/paper_main.tex
adding comments from Lynn

@@
 
+\author{Changrong Liu}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+
 \author{Andrew Melatos}
@@
 OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{Changrong Liu}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
 }
@@
 \begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; and Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction.
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular Lynn Cominsky and Sam Cooper for their helpful suggestions.
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
-This work received financial support towards hardware from the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme and travel support from the Australian National PhB Science program.
+Financial support towards hardware was provided by the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme. 
+Travel support was provided by the Australian National PhB Science program.
 This work has been assigned LIGO document number P2000386.

### 2020-10-30 paper/ifo-complexAudio.tex
adding new version of big filtering plot to include background noise and removing what was fig 9

@@
 
-\begin{figure*}
-	\includegraphics[width=.7\textwidth]{figures/setup_pic2.pdf}
+\begin{figure}
+	\includegraphics[width=.49\textwidth]{figures/setup_pic2.pdf}
 \caption{Photograph of the optical microphone.
@@
 	\label{fig:setup_pic2}
-\end{figure*}
+\end{figure}
 
@@
 	\begin{center}
-% trim order is <left> <lower> <right> <upper>
-\includegraphics[width=\textwidth,trim={1cm 8.5cm 1cm 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
+    %% trim order is <left> <lower> <right> <upper>
+	\includegraphics[width=\textwidth,trim={1cm 4.5cm 1cm 3.5cm},clip]{figures/notch_and_wiener_superplot_v2.pdf}
 	\end{center}
-\caption{\label{fig:noiseSubtract}
-Frequency spectrum of a $1\,{\rm s}$ signal before the use of filters. 
-The first panel shows the original frequency spectrum of the human voice recording (saying ``A cathode ...''). 
-The second panel shows the spectrum recorded after the signal passes through the optical microphone (sampled at $16\,000\,{\rm Hz}$).
-The third panel shows a noise-only recording from the optical microphone when no audio signal is being played. 
-The fourth panel shows the result of subtracting the noise-only spectrum in the third panel from the recording in the second panel. 
-We see no obvious improvement from this method. 
+	\caption{\label{fig:BackgroundNotchWienerCombined}
+	Time series (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
+	The first and second rows show the original speech recording and raw output from the optical microphone respectively. 
+    The third row shows a background noise recording from the optical microphone when no audio signal is played. 
+    The fourth row shows the result of subtracting the background noise spectrum in the third row from the recording in the second row. 
+	%The second row shows the raw output from the optical microphone.
+	%The third row shows the naive technique of subtracting off the noise spectrum from the recorded spectrum.
+	The fifth, sixth, and seventh rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
 	}
@@
 
-The Fourier spectrum of the source signal is shown in the first panel of Fig.~\ref{fig:noiseSubtract}.
+%\begin{figure*}
+%\begin{center}
+%% trim order is <left> <lower> <right> <upper>
+%\includegraphics[width=\textwidth,trim={1cm 8.5cm 1cm 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
+%\end{center}
+%\caption{\label{fig:noiseSubtract}
+%Frequency spectrum of a $1\,{\rm s}$ signal before the use of filters. 
+%The first panel shows the original frequency spectrum of the human voice recording (saying ``A cathode ...''). 
+%The second panel shows the spectrum recorded after the signal passes through the optical microphone (sampled at $16\,000\,{\rm Hz}$).
+%The third panel shows a noise-only recording from the optical microphone when no audio signal is being played. 
+%The fourth panel shows the result of subtracting the noise-only spectrum in the third panel from the recording in the second panel. 
+%We see no obvious improvement from this method. 
+%}
+%\end{figure*}
+
+Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results respectively. 
+The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the source signal and the raw optical microphone recording respectively. 
+The third row shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.
+
+%The Fourier spectrum of the source signal is shown in the first panel of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
 The second panel shows the spectrum of the raw optical microphone recording of the signal.
@@
 Given that we have access to the background noise spectrum, an intuitive way to remove noise is via subtracting the noise spectrum from the recorded spectrum. 
-The fourth panel of Fig.~\ref{fig:noiseSubtract} shows the spectrum obtained after subtracting the background noise spectrum.
-We see no obvious improvement, which may be attributed to a time-variant noise spectrum the cause of which is unidentified.
-%Why the noise sources are time-varying is a fact yet to be determined.
+The fourth panel of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the spectrum obtained after subtracting the background noise spectrum. 
+We see no obvious improvement, which may be attributed to a time-variant noise spectrum the cause of which is unidentified. %Why the noise sources are time-varying is a fact yet to be determined.
 
-Beyond this naive approach, we apply a range of filters to the data.
+Beyond this na\"{i}ve approach, we apply a range of filters to the data.
 In signals processing, the ideal filter would be one that:
@@
 \begin{center}
-\includegraphics[width=.9\textwidth]{figures/notch_filter_response}
+\includegraphics[width=.9\textwidth]{figures/notch_filter_response.jpg}
 \end{center}
@@
 The amplitude response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
-The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fourth row of Fig.~\ref{fig:NotchWienerCombined}.
-We see that the mains harmonics are significantly attenuated in comparison to the results shown in Fig.~\ref{fig:noiseSubtract}.
+The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
+We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
 
@@
 % By purely minimizing the "distance" between estimated signal (filtered human voice) and observed signal (original human voice), we hope that we keep signal part undisturbed while filtering out noise.
-
-\begin{figure*}
-	\begin{center}
-	\includegraphics[width=\textwidth]{figures/notch_and_wiener_superplot.pdf}
-	\end{center}
-	\caption{\label{fig:NotchWienerCombined}
-	Time series (left column) and frequency spectrum (right column) results for the notch and Wiener filters.
-	The first row shows the original speech recording.
-	The second row shows the raw output from the optical microphone.
-	The third row shows the naive technique of subtracting off the noise spectrum from the recorded spectrum.
-	The fourth, fifth, and sixth rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
-	}
-\end{figure*}
 
@@
 For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
-The time series and frequency spectrum after applying the Wiener filter are shown in the fifth row of Fig.~\ref{fig:NotchWienerCombined}.
+The time series and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
 We see a significant improvement in the time series of the recovered signal, however a strong noise hum persists, audibly. 
@@
 
-By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:NotchWienerCombined}.
+By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:BackgroundNotchWienerCombined}..
 The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone.
@@
 \label{sec:logmmse}
-% logMMSE
+
+\begin{figure*}
+	\includegraphics[width=\textwidth]{figures/combined_timeseries_melatos.pdf}
+	\caption{Time series of an adult male voice (saying ``A cathode ...'') showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
+One second of the one minute recording is shown.
+The rise at the start of the bottom panel is expected when filtering a signal of finite duration.}
+	\label{fig:logMMSE_timeseries}
+\end{figure*}
+
+\begin{figure*}
+	\includegraphics[width=\textwidth]{figures/combined_spectrum_melatos.pdf}
+	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
+The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.}
+	\label{fig:logMMSE_spectrum}
+\end{figure*}
 
@@
 %Applying this estimator to the optical microphone output produced dramatically cleaner results.
-
-\begin{figure*}
-	\includegraphics[width=\textwidth]{figures/combined_timeseries_melatos.pdf}
-	\caption{Time series of an adult male voice (saying ``A cathode ...'') showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
-One second of the one minute recording is shown.
-The rise at the start of the bottom panel is expected when filtering a signal of finite duration.}
-	\label{fig:logMMSE_timeseries}
-\end{figure*}
-
-\begin{figure*}
-	\includegraphics[width=\textwidth]{figures/combined_spectrum_melatos.pdf}
-	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
-The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.}
-	\label{fig:logMMSE_spectrum}
-\end{figure*}
 

### 2020-10-30 paper/ifo-introduction.tex
adding reference for GWTC-2

@@
 % CW 
-To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018}, binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018}, and recently the merger of black hole with an object that could either have been a neutron star or a black hole~\cite{GW190814}.
+To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020}, binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, and recently the merger of black hole with an object that could either have been a neutron star or a black hole~\cite{GW190814}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 

### 2020-11-06 paper/ifo-appendix.tex
some general text editing and addressing comments
first pass at addressing p&p comments

@@
 Near a bright fringe, where one has $\cos{[2kd\cos{(\theta)}]} \approx 1$, the nonlinearity is quadratic, with $\delta I \propto (\delta d)^2$ for small $\delta d$. 
-Consequently, an oscillation of the form $\delta d \propto \exp{( -i \Omega t )}$ in the arm length difference produces a DC signal and a frequency-double signal $\propto \exp{(-2i\Omega t)}$ in $\delta I$.
-This is discussed further in Section~\ref{}\han{to do}. \jam{[not quite sure how much further we can take discussion about this, are we wanting results that show the doubling?]}
+Consequently, an oscillation of the form $\delta d \propto \exp{( -i \Omega t )}$ in the arm length difference produces a DC signal and a frequency-doubled signal $\propto \exp{(-2i\Omega t)}$ in $\delta I$.
+This is discussed further in Sec.~\ref{sec:ifo}.
 
@@
                  &=& -\frac{1}{2}(x|x) - \frac{A^2}{2} (\cos[\omega t + \phi] | \cos[\omega t + \phi]) \\ &&+ A (x | \cos[\omega t +\phi])\,, \\ 
-                 &=& -\frac{1}{2}(x|x) - \frac{A^2}{4} + A (x | \cos[\omega t +\phi])\,, \label{eqn:logLike}\\ 
+                 &=& -\frac{1}{2}(x|x) - \frac{A^2}{4} + A (x | \cos[\omega t +\phi])\,, \label{eqn:logLike}
 \end{eqnarray}
@@
 
-\subsection{Pseudocode}
+%\subsection{Pseudocode}
 Here we give pseudocode of the implementation used in Section~\ref{sec:viterbi}.
@@
 The photodiode is connected in reverse-bias across an op-amp. 
-The analog op-amp output is passed through a Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC). 
+The analog op-amp output is passed through a Sallen-Key anti-aliasing filter tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
 The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.
+This diagram is produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online} and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
 }

### 2020-11-06 paper/ifo-complexAudio.tex
first pass at addressing p&p comments

@@
 % aim of optical microphone: play back injected sounds
+In this section we explore how this demonstration can be used to teach a selection of signal processing techniques.
 A natural successor to constant and wandering tones is complex audio, such as music and speech.
@@
 
-\begin{figure}
-	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_photodiode_edit.pdf}
+\begin{figure*}
+	\includegraphics[width=0.44\textwidth]{figures/ifo_schematic_photodiode_edit.pdf}
+	\includegraphics[width=0.52\textwidth]{figures/setup_pic2.pdf}
 	\caption{
-Michelson interferometer schematic. 
-The schematic is identical to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
+Left: Michelson interferometer schematic. 
+The schematic is identica.l to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
+Right: photograph of the optical microphone. 
+The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. 
+In the main image, the photodiode is placed behind the cloth screen. 
+The inset shows a face-on view of the photodiode with the covering screen removed. 
+See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.
 }
 	\label{fig:ifo_schematic_podo}
-\end{figure}
+\end{figure*}
 
+\begin{comment}
 \begin{figure}
@@
 \end{figure}
+\end{comment}
 
@@
 %\end{itemize}
-However, these three conditions are cannot all hold at once and any filter must compromise between them.
+However, these three conditions cannot all hold at once and any filter must compromise between them.
 
@@
 The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is some combination of sinc functions, which significantly corrupt the signal.
+See also Section~\ref{sec:notch} where we explore a notch filter.
 
@@
 \subsubsection{Cascade notch filter}
+\label{sec:notch}
 \begin{figure*}
@@
 \end{equation}
-Given the observed data $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the (Bayesian) mean square error (MSE), as defined in Eq.~\ref{eq:BMSE}, between the two.
+Given the observed data $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean square error (BMSE) between the two, defined as
 \begin{equation}
@@
 
-If we assume that: i) $\textbf{x}$ is wide sense stationary~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, if (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
+If we assume that: i) $\textbf{x}$ is \emph{wide sense stationary}~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
 \begin{equation}
@@
 
-In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}, and set the length of the filter to $n=100$. 
-A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
-The choice of $n=100$ is a reasonable selection to balance smoothing and efficiency.
+%In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}, and set the length of the filter to $n=100$. 
+%A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
+%The choice of $n=100$ is a reasonable selection to balance smoothing and efficiency.
 
@@
 After applying the logMMSE estimator, simple chords and drums can be heard, however more composite sounds and complex melodies cannot be heard clearly.
-Our observations suggest that this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins sometimes can’t be heard at all.
+%Our observations suggest that this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins sometimes can’t be heard at all. 
+Our observations suggest that this is especially true for certain instruments, in particular flutes and violins sometimes can’t be heard at all. 
 This could be a perceptual effect or a frequency dependence somewhere in the optical microphone.

### 2020-11-06 paper/ifo-introduction.tex
first pass at addressing p&p comments

--- 2020-10-05 paper/ifo-introduction.tex
@@
 % CW 
-To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018}, binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018}, and recently the merger of black hole with an object that could either have been a neutron star or a black hole~\cite{GW190814}.
+To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
@@
 The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
-Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
+Scorpius X-1 is a prime LMXB target for continuous wave searches. 
+%Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
 Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.
@@
 % Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
-This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world.
-These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to a non-specialist audience.
+%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world.
+This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
+These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
 %\han{look up Sarah Farmer's project}

### 2020-11-06 paper/ifo-tableTopGWs.tex
first pass at addressing p&p comments

@@
 The resulting interference pattern is dependent on the relative distance travelled by the beams. 
-Current generation gravitational-wave observatories have ${\rm km}$-scale arms ($4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
+Current generation gravitational-wave observatories have ${\rm km}$-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
 They are able to measure minuscule changes in distance due to gravitational-waves, e.g.
-the first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$, a distance smaller than the width of a proton~\cite{GW150914}.
+the first detection of a binary black hole merger produced a strain of $10^{-21}$, corresponding to a change in distance which is a fraction of the width of a proton.
+%changed the relative length of the two arms by $10^{-21}$, a distance smaller than the width of a proton~\cite{GW150914}.
 
@@
 
-Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} provide a selection of other table-top Michelson interferometer designs used for science communication).
 % I rephrased to lead with the analogy here and removed 'often' as GW signals are only really listened to for public engagement interest - hannah
@@
 %This analogue allows scientists to use sound vibrations as a simple means to move the components of a demonstration interferometer. 
+Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
 Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
@@
 A change in the relative length of the arms causes these rings to move radially in or outwards.
+Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
 % Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed. 

### 2020-11-06 paper/ifo-wanderingTone.tex
first pass at addressing p&p comments

@@
 % the signal and the analogy we use
-Continuous gravitational waves may not be monochromatic, unlike the signal considered in Section~\ref{sec:single_tone}.  
+Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational waves may not be monochromatic.
+%Continuous gravitational waves may not be monochromatic, unlike the signal considered in Section~\ref{sec:single_tone}.  
 A continuous-wave signal may wander slowly in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
@@
 The frequency of the signal can wander significantly over the observation period. 
-In this context significantly means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
+In this context \emph{significantly} means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
 One method to search for a wandering signal is to split the time series data into several shorter segments which are analysed individually.%, and then recombine them under some continuity assumption to coherently analyse the whole set.
 % I think the following is more accurate. The antenna beam pattern is used to compute a detection statistic. Which detection statistic to use depends on the type of target. -Hannah
-A detection statistic is used to quantify the likelihood of a signal being present in the data at each frequency and timeseries segment. 
+
+A \emph{detection statistic} is used to quantify the likelihood of a signal being present in the data at each frequency and timeseries segment. 
 In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
@@
 %The detection statistic used depends on the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun, and the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
-Which detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
+The choice of detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
 In this work, we take the discrete Fourier transform of each timeseries segment and use its amplitude at each frequency as a detection statistic.
 This maximises the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}. 
-The detection statistic is normalised for convenience. 
-The detection statistics from each segment are combined to form a grid in time and frequency (a spectrogram). 
+The detection statistic is normalised for convenience and we use them to form a grid in  time and frequency (a spectrogram). 
 % moved one sentence up 
@@
 %To perform the task of recovering the signal, we use the Viterbi algorithm~\cite{Viterbi:1967} to efficiently find the most probable path through the spectrogram given the sequence of observations.
-This method is used in continuous wave searches for a variety of astrophysical targets~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,MillhouseStrangMelatos:2020,JonesSun:2020,MiddletonEtAlO2LMXBs:2020,PostMergerRemnantSearch:2019,SunEtAlSNR:2018,viterbi_application}.
+This method is used in continuous wave searches for a variety of astrophysical targets~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application} and forms the basis of the analysis in this section.
 
@@
 
-\item The Viterbi path (the overall best path) is then found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}).
+\item The last step is to find the Viterbi path (the overall best path that terminates in the frequency bin with the highest probability in step 4).
+The Viterbi path is found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}).
+In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c) and is highlighted in orange.
+%\item The Viterbi path (the overall best path) is then found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}).
 In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c) and is highlighted in orange.
@@
 
-A sinusoidal signal with slowly decaying amplitude is played through the speaker to mimic a slowly wandering signal. 
+To simulate a slowly wandering signal we modulate the frequency sinusoidally with a modulation amplitude that decays with time. 
+%A sinusoidal signal with slowly decaying amplitude is played through the speaker to mimic a slowly wandering signal. 
 We use an identical set up to that shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
@@
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
-The overlaid pink dots show the injected signal and the white crosses show the recovered Viterbi path.% I think this is already defined and seems like repetition:(the most likely path through the data according to the Viterbi algorithm).
+For the purposes of this demonstration we use a signal that can easily be identified  in the spectrogram, however real continuous wave signals have much lower amplitudes. 
+The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.% I think this is already defined and seems like repetition:(the most likely path through the data according to the Viterbi algorithm).
 The recovered Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time.

### 2020-11-06 paper/paper_main.tex
first pass at addressing p&p comments

@@
 These searches use sophisticated statistical techniques to look for weak signals in noisy data. 
-In this paper, we demonstrate these techniques using a table-top model gravitational-wave detector (a Michelson interferometer), where sound plays the role of gravitational waves. 
+In this paper, we demonstrate these techniques using a table-top model gravitational-wave detector (a Michelson interferometer), where sound is used as an analogue for  gravitational waves. 
 Using signal processing techniques from continuous-wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
@@
 \begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular Lynn Cominsky and Sam Cooper for their helpful suggestions.
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular Anna Green, Lynn Cominsky and Sam Cooper for their helpful feedback and suggestions.
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 

### 2020-11-16 paper/ifo-complexAudio.tex
working on pnp comments updating new plot from Changrong with synchronised original signal

@@
 % aim of optical microphone: play back injected sounds
+% some editing in response to pnp
 In this section we explore how this demonstration can be used to teach a selection of signal processing techniques.
-A natural successor to constant and wandering tones is complex audio, such as music and speech.
+The apparatus used in Sections~\ref{sec:ifo} to~\ref{sec:viterbi_wandering} affords the opportunity to explore a range of signal processing techniques and filters of interest to undergraduate students and provide a foundations for many technical applications~\cite{DigitalProcingOfSpeechSignals:1978}.
+We provide a series of example activities using complex audio (such as music and speech) as a natural successor to constant and wandering tones. 
+%A natural successor to constant and wandering tones is complex audio, such as music and speech. 
 As complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not directly applicable here.
-Instead, in this section, we use the interferometer as an `optical microphone' to detect and play back complex audio signals (i.e.\ using light to capture sound).
+Instead, in this section, we use the interferometer as an \emph{optical microphone} to detect and play back complex audio signals (i.e.\ using light to capture sound).
 In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
-The device has precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defence industry and that operate on a variety of related principles.
+A vast literature exists on signal processing with speech, however see Ref.~\cite{DigitalProcingOfSpeechSignals:1978} for introductory material including Fourier analysis and digital representations.
+Optical microphone devices have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defence industry and operate on a variety of related principles.
+
+%An introduction to ..... can be found in Ref.~\cite{}, including 
+%A review of digital processing techniques for speech signals can be found 
 %It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
@@
 
+This section is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals.
+Here, the optical microphone serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories.
 We describe the additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}. 
-Filter and speech enhancement techniques applied to the data are outlined in Section~\ref{sec:filters}.
-This section is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals.
-In addition to this, the optical microphone serves as an interesting independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories.
+A selection of filter and speech enhancement example activities using the optical microphone are outlined in Section~\ref{sec:filters}.
 
@@
     %% trim order is <left> <lower> <right> <upper>
-	\includegraphics[width=\textwidth,trim={1cm 4.5cm 1cm 3.5cm},clip]{figures/notch_and_wiener_superplot_v2.pdf}
+	\includegraphics[width=\textwidth,trim={1cm 4.5cm 0.4cm 3.5cm},clip]{figures/notch_and_wiener_superplot_v2.pdf}
 	\end{center}

### 2020-11-16 paper/ifo-introduction.tex
working on pnp comments updating new plot from Changrong with synchronised original signal

@@
 % changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
-In this paper, we use a table-top Michelson interferometer as a `toy' gravitational-wave detector designed to detect sound instead of gravitational waves. 
+In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
+We then extend its use to an \emph{optical microphone} (using light to capture sound) and present a range of example analysis techniques for educators to use. 
 % re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
@@
 % removed as there's not space to really say what this means here: to varying degrees of success.
-This demonstration of an ‘optical microphone’ (using light to capture sound) serves as a more general exhibition of signal processing. 
+This demonstration of an optical microphone  serves as a more general exhibition of signal processing with a range of examples which can be used in the undergraduate laboratory. 
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in ~\ref{sec:conclusions}. 

### 2020-11-16 paper/paper_main.tex
working on pnp comments updating new plot from Changrong with synchronised original signal

@@
 \begin{abstract}
-Gravitational-wave observatories around the world are searching for continuous waves: persistent signals from spinning neutron stars. 
+Gravitational-wave observatories around the world are searching for continuous waves: persistent signals from sources such as spinning neutron stars. 
 These searches use sophisticated statistical techniques to look for weak signals in noisy data. 
@@
 Using signal processing techniques from continuous-wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
-Finally we use the interferometer as an `optical microphone' to capture music and speech and apply filtering techniques to recover the original signals from the noisy data.
+We explore use of the interferometer as a teaching tool for educators in physics and electrical undergraduate laboratories. 
+The interferometer is used as an \emph{optical microphone} to capture music and speech. 
+A selection of filtering techniques are applied to recover the original signals from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
+Simple chords and drums are easily recovered, but complex music and speech are more challenging. 
+This demonstration has scope for use by undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
+% rearranged in response to Anna's pnp review
+%Finally we use the interferometer as an `optical microphone' to capture music and speech and apply filtering techniques to recover the original signals from the noisy data.
 Such techniques include simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
-Simple chords and drums are easily recovered, but complex music and speech are more challenging. 
-This demonstration has scope for use in physics and electrical engineering undergraduate laboratories. 
-It can also be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences. 
+%Simple chords and drums are easily recovered, but complex music and speech are more challenging. 
+%This demonstration has scope for use in physics and electrical engineering undergraduate laboratories. 
+%It can also be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences. 
 \end{abstract}

### 2020-11-16 paper/ifo-complexAudio.tex
minor changes, commented out footmisc

@@
 
-An updated schematic is shown in Fig.~\ref{fig:ifo_schematic_podo} and a photograph of the entire optical microphone in Fig.~\ref{fig:setup_pic2}.
+An updated schematic and a photograph of the entire optical microphone are shown in the left and right panels of Fig.~\ref{fig:ifo_schematic_podo}, respectively.
 See Appendix~\ref{app:circuit_diagram} for a diagram and photograph of the circuit.
@@
     %% trim order is <left> <lower> <right> <upper>
-	\includegraphics[width=\textwidth,trim={1cm 4.5cm 0.4cm 3.5cm},clip]{figures/notch_and_wiener_superplot_v2.pdf}
+    % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
+	\includegraphics[height=0.85\textheight]{figures/notch_and_wiener_superplot_v2.pdf}
 	\end{center}
@@
 \begin{figure*}
-	\includegraphics[width=\textwidth]{figures/combined_timeseries_melatos.pdf}
+	\includegraphics[width=0.8\textwidth]{figures/combined_timeseries_melatos.pdf}
 	\caption{Time series of an adult male voice (saying ``A cathode ...'') showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
@@
 \begin{figure*}
-	\includegraphics[width=\textwidth]{figures/combined_spectrum_melatos.pdf}
+	\includegraphics[width=0.8\textwidth]{figures/combined_spectrum_melatos.pdf}
 	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).

### 2020-11-16 paper/paper_main.tex
minor changes, commented out footmisc

@@
 %\documentclass[prb,reprint,nofootinbib]{revtex4-1} 
-\documentclass[pra,superscriptaddress,reprint,amsmath,amssymb,nofootinbib]{revtex4-1}
+\documentclass[aps,pra,superscriptaddress,reprint,nofootinbib]{revtex4-1}
 % \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
@@
 \usepackage{subfiles}
-\usepackage[bottom]{footmisc}
+% \usepackage[bottom]{footmisc}
 \usepackage{comment}
@@
 % \usepackage{algorithmic}
+\usepackage{lmodern}
+
+\usepackage{silence}
+\WarningFilter{revtex4-1}{Repair the float}
 
@@
 Financial support towards hardware was provided by the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme. 
-Travel support was provided by the Australian National PhB Science program.
+Travel support was provided by the Australian National University PhB Science program.
 This work has been assigned LIGO document number P2000386.

### 2020-11-18 paper/ifo-complexAudio.tex
updating figures and slight edit of abstract

@@
     % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
-	\includegraphics[height=0.85\textheight]{figures/notch_and_wiener_superplot_v2.pdf}
+	\includegraphics[width=.85\textwidth, trim={1cm 3cm 1cm 1cm}]{figures/notch_and_wiener_superplot_v2.pdf}
 	\end{center}
@@
 \begin{figure*}
-	\includegraphics[width=0.8\textwidth]{figures/combined_timeseries_melatos.pdf}
+	\includegraphics[width=0.8\textwidth]{figures/combined_timeseries_melatos_shift_source.pdf}
 	\caption{Time series of an adult male voice (saying ``A cathode ...'') showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).

### 2020-11-18 paper/ifo-conclusions.tex
reducing repetition in conclusion to shorten paper a little

@@
 Gravitational-wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
-We consider quasi-monochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music for more general interest.
-These recordings of complex audio could also serve as more distant analogues to un-modelled gravitational-wave bursts, e.g.
-from supernovae.
+We consider quasi-monochromatic signals as an analogue to continuous wave signals to directly demonstrate techniques used in current continuous gravitational-wave searches.
+We then consider more complex audio (music and speech) for a more general interest in signal processing and explore use of the interferometer as an optical microphone. 
+A variety of example activities are presented. 
+We find that the logMMSE produces the best results, however it does not enhance the recording to an intelligible level. 
+The complex audio signals can also serve as more distant analogues to un-modelled gravitaitonal-wave burst signals, e.g. from supernova.
 
+\begin{comment}
+% included these paras in the first to avoid repetition and reduce some of the length of the paper before submission
 % reduced detail for conclusion (trying to reduce overall length)
@@
 Further avenues of exploration include even more advanced processing techniques and a full characterisation of the transfer function of the system, with the goal to lead to improvements in intelligibility.
+\end{comment}
 

### 2020-11-18 paper/ifo-futureWork.tex
minor changes, commented out footmisc
removing HMM acroymn

@@
 \begin{figure*}
-	\includegraphics[height=.99\textwidth, angle=-90]{figures/pipeline.pdf}
+	\includegraphics[height=.8\textwidth, angle=-90]{figures/pipeline.pdf}
 	\caption{
@@
 
-A voluminous literature exists on using HMMs trained on phonemes to recognize speech~\cite{HMM_english}.
+A voluminous literature exists on using hidden Markov models trained on phonemes to recognize speech~\cite{HMM_english}.
 These could perform the final stage of speech enhancement for the optical microphone.
-Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a HMM, albeit of a different type.
+Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
 Alternatively, machine learning solutions exist throughout the field that are able to compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.

### 2020-11-18 paper/ifo-wanderingTone.tex
removing HMM acroymn

@@
 
-Following the method of Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, we use a hidden Markov model (HMM) to recover the wandering signal from noisy data. 
-In a HMM, the frequency state of the signal is unknown, or hidden, and can undergo transitions at discrete times. 
+Following the method of Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, we use a hidden Markov model to recover the wandering signal from noisy data. 
+In a hidden Markov model, the frequency state of the signal is unknown, or hidden, and can undergo transitions at discrete times. 
 %To do this, we define the frequency of the signal as a hidden (unknown) state which can undergo transitions at discrete times. 

### 2020-11-19 paper/ifo-appendix.tex
full read-through before submission

@@
 \label{app:code}
-This project is implemented in python3~\cite{python} scripts as well as jupyter notebooks~\cite{jupyter}~\cite{ipython} and used the packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
+This project is implemented in Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}.
+We used the Python packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
 The current build and sample data can be found at:
@@
 
-Part of the transfer function for the optical microphone relates change in intensity, $\delta I$, at a point $P$ in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
+Part of the transfer function for the optical microphone relates the change in intensity, $\delta I$, at a point $P$ in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
 Let $\theta$ be the angle between the optical axis and the ray from the centre of the relevant mirror to the point $P$. 
@@
 Consequently, an oscillation of the form $\delta d \propto \exp{( -i \Omega t )}$ in the arm length difference produces a DC signal and a frequency-doubled signal $\propto \exp{(-2i\Omega t)}$ in $\delta I$.
-This is discussed further in Sec.~\ref{sec:ifo}.
+This is discussed further in Section~\ref{sec:ifo}.
 
@@
 Classic sig proc result useful for teachers to know.'']}
-In this appendix we demonstrate the the modulus of the Fourier Transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
+In this appendix, we demonstrate the the modulus of the Fourier Transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
 We describe the data as
@@
 where $s(t)$ and $n(t)$ are the signal and noise respectively. 
-The we describe the signal as 
+Then we describe the signal as 
 \begin{equation}
@@
 \begin{eqnarray}
-\log \mathcal{L} &=& -\frac{1}{2}(n|n)\,, \\
-                 &=& -\frac{1}{2}(x-s|x-s)\,, \\ 
-                 &=& -\frac{1}{2}(x|x) - \frac{A^2}{2} (\cos[\omega t + \phi] | \cos[\omega t + \phi]) \\ &&+ A (x | \cos[\omega t +\phi])\,, \\ 
-                 &=& -\frac{1}{2}(x|x) - \frac{A^2}{4} + A (x | \cos[\omega t +\phi])\,, \label{eqn:logLike}
+\log \mathcal{L} &=& -\frac{1}{2}\ip{n}{n}\,, \\
+                 &=& -\frac{1}{2}\ip{x-s}{x-s}\,, \\ 
+                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{2} \ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]} \nonumber\\
+                 &&+ A \ip{x}{\cos[\omega t + \phi]}\,,\\ 
+                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{4} + A \ip{x}{\cos[\omega t + \phi]}\,,\label{eqn:logLike}
 \end{eqnarray}
-where we define the inner product between $u(t)$ and $v(t)$ as $(u|v) = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data (and the last line assumes that $T>>1/\omega$ so that $(\cos[\omega t + \phi] | \cos[\omega t + \phi])\approx{1/2}$).
+where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data (and the last line assumes that $T\gg1/\omega$ so that $\ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]}\approx{1/2}$).
 
-Maximising Eqn.~\ref{eqn:logLike} with respect to $A$ shows that the maximum likelihood is when $A = 2\left( x | \cos[\omega t+\phi]\right)$. 
+Maximising Eqn.~\ref{eqn:logLike} with respect to $A$ shows that the maximum likelihood is when $A = 2\ip{x}{\cos[\omega t + \phi]}$. 
 The maximum likelihood is then, 
 \begin{eqnarray}
-\log \mathcal{L}_\mathrm{max} = - \frac{1}{2} (x|x) + ( x | cos[\omega t + \phi)^2\,. \label{eqn:maxLogLike}
+\label{eqn:maxLogLike}
+\log \mathcal{L}_\mathrm{max} = - \frac{1}{2} \ip{x}{x} + \ip{x}{\cos[\omega t + \phi]}^2\,. 
 \end{eqnarray}
-Focusing on the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, baring the constant factor $(x|x)$.
+Focusing on the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, baring the constant factor $\ip{x}{x}$.
 
@@
 More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}.
-We use Fourier amplitudes normalised between $(0, 1)$ as multiplicative weights, but to avoid underflow we take the logarithm and consider them as additive weights.
+We use Fourier amplitudes normalised between $(0, 1)$ as multiplicative weights, but to avoid numerical underflow we take the logarithm and consider them as additive weights.
 
@@
 Let Y and Z be grids of the same shape to store the weight of the best path to each node and the row index of the previous node on that path, respectively.
-Let W be a length $N_t+1$ array to store the final sequence of row indices.
+Let W be a length $N_t+1$ array to store the final sequence of row indices for the optimal path.
 We restrict paths to only move one cell up or down at a time (or stay constant).
@@
 The photodiode is connected in reverse-bias across an op-amp. 
-The analog op-amp output is passed through a Sallen-Key anti-aliasing filter tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
+The op-amp output is passed through a Sallen-Key anti-aliasing filter tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
 The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.
@@
 	\caption{\label{fig:circuit_pic3}
-Photodiode circuit assembled on breadboard.
+Photodiode circuit assembled on breadboard. The leads from the photodiode enter from the left of the picture. The Raspberry Pi is shown at the top of the picture.
 }

### 2020-11-19 paper/ifo-complexAudio.tex
full read-through before submission

@@
 % some editing in response to pnp
-In this section we explore how this demonstration can be used to teach a selection of signal processing techniques.
+In this section, we explore how this demonstration can be used to teach a selection of signal processing techniques.
 The apparatus used in Sections~\ref{sec:ifo} to~\ref{sec:viterbi_wandering} affords the opportunity to explore a range of signal processing techniques and filters of interest to undergraduate students and provide a foundations for many technical applications~\cite{DigitalProcingOfSpeechSignals:1978}.
@@
 As complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not directly applicable here.
-Instead, in this section, we use the interferometer as an \emph{optical microphone} to detect and play back complex audio signals (i.e.\ using light to capture sound).
-In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
-A vast literature exists on signal processing with speech, however see Ref.~\cite{DigitalProcingOfSpeechSignals:1978} for introductory material including Fourier analysis and digital representations.
+Instead, in this section, we use the interferometer as an ``optical microphone'' to detect and play back complex audio signals (i.e.\ using light to capture sound).
+In doing so, we experiment with a hierarchy of passive filters which suppress noise, yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
+A vast literature exists on signal processing with speech, see Ref.~\cite{DigitalProcingOfSpeechSignals:1978} for introductory material including Fourier analysis and digital representations.
 Optical microphone devices have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defence industry and operate on a variety of related principles.
@@
 %A review of digital processing techniques for speech signals can be found 
-%It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
+%It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
 %It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae. [refs; my last sentence can be improved]
 
-This section is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals.
+This section is indirectly related to the advanced signal processing techniques required to extract gravitational wave signals.
 Here, the optical microphone serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories.
@@
 
-Audible frequencies are in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$.
-Speech intelligibility requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$~\cite{speech_intelligibility}.
-Therefore the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music.
-This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$.
+The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$.
+Speech intelligibility (being able to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to (and beyond) $8\,{\rm kHz}$~\cite{speech_intelligibility}.
+Therefore, the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music (adjusting for the Nyquist frequency).
+This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$ and thus can only ``hear'' frequencies below $15\,{\rm Hz}$.
 To overcome this issue, we use a photodiode\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the reverse direction.
@@
 The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
-The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath. 
-The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi (Pi) Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data.
+The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath the screen. 
+The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
 Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. Resources for using the Pi and photodiode are described in Appendix~\ref{app:circuit_diagram}.
@@
 This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
-We also place a cloth screen over the face of the photodiode to reduce the intensity of the voltage signal and avoid saturating the ADC.
+We also place a second cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC.
 % It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
@@
 	\caption{
-Left: Michelson interferometer schematic. 
-The schematic is identica.l to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
-Right: photograph of the optical microphone. 
+Left: the updated Michelson interferometer schematic. 
+The schematic is identical to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
+Right: a photograph of the optical microphone. 
 The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. 
@@
 %The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources). 
-We test the optical microphone with a variety of recordings including speech from different people and music ranging from simple melodies and rhythms to songs.
+We test the optical microphone with a variety of recordings, including speech from different people and music ranging from simple melodies and rhythms to songs.
 During recordings, care is taken to minimise activity around the demonstration to reduce the amount of environmental noise coupling into the interferometer.
-The time series data is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function.
+The time series data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python~\cite{scipy,python}.
 When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.
@@
 	\caption{\label{fig:psd_noise}
-Power spectral density (PSD) of background noise (top) from the optical microphone (with the speaker off), and the PSD after applying a Butterworth bandpass filter (bottom) between the two frequencies marked with red, dashed lines.
-We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise the PSD is fairly white. 
-After filtering we see strong attenuation of all frequencies outside the band, but little change to the harmonics within the band.
+Power spectral density (PSD) of background noise (top panel) from the optical microphone (with the speaker off), and the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines.
+We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
+After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
 }
@@
 
-Here, we describe several filters used to remove the $50\,{\rm Hz}$ mains hum and its harmonics, but also to ultimately improve the speech intelligibility of the recording (to be able to understand the words).
+Here, we describe several filters used to remove the $50\,{\rm Hz}$ mains hum and its harmonics, but also to ultimately improve the speech intelligibility of the recording (to be able to understand the words spoken).
 We start with some naive approaches and then move to traditional signal processing filters like band-passing and cascaded notches.
@@
     % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
-	\includegraphics[width=.85\textwidth, trim={1cm 3cm 1cm 1cm}]{figures/notch_and_wiener_superplot_v2.pdf}
+	\includegraphics[width=.8\textwidth, trim={1cm 3cm 1cm 1cm}]{figures/notch_and_wiener_superplot_v2.pdf}
 	\end{center}
@@
 	Time series (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
-	The first and second rows show the original speech recording and raw output from the optical microphone respectively. 
+	The first and second rows show the original speech recording and raw output from the optical microphone, respectively. 
     The third row shows a background noise recording from the optical microphone when no audio signal is played. 
@@
 
-Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results respectively. 
-The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the source signal and the raw optical microphone recording respectively. 
+Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results, respectively. 
+The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} show the source signal and the raw optical microphone recording, respectively. 
 The third row shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.
@@
 The fourth panel of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the spectrum obtained after subtracting the background noise spectrum. 
-We see no obvious improvement, which may be attributed to a time-variant noise spectrum the cause of which is unidentified. %Why the noise sources are time-varying is a fact yet to be determined.
+We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified. %Why the noise sources are time-varying is a fact yet to be determined.
 
-Beyond this na\"{i}ve approach, we apply a range of filters to the data.
+Beyond this naive approach, we apply a range of filters to the data.
+% \jam{changed na\"{i}ve to naive, if you change this back please do it globally}
 In signals processing, the ideal filter would be one that:
@@
 %\end{itemize}
-However, these three conditions cannot all hold at once and any filter must compromise between them.
+However, these three conditions cannot all hold at once and any filter must compromise between them, as we will see below.
 
-\subsubsection{Comb filter}
+\subsubsection{Rectangular comb filter}
 
@@
 It does remove the mains harmonics, but audibly ruins the rest of the signal due to lack of smoothness.
-This is because applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter.
+This is because applying a filter in frequency space is equivalent to convolving the time-domain signal with the inverse Fourier transform of that filter.
 The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is some combination of sinc functions, which significantly corrupt the signal.
@@
 However, it makes the played-back signal unrecognisable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
-Often in speech processing the logarithm of the signal is taken since the amplitude information seems to be more important to intelligibility than the phase information.
+Often in speech processing, the logarithm of the signal is taken since the amplitude information seems to be more important to intelligibility than the phase information to the human ear~\cite{SubjectiveComparison}.
 However, applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.
@@
 	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000.pdf}
-	\caption{Butterworth bandpass filter frequency response, any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power).
+	\caption{Butterworth bandpass filter frequency response.
+Any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power).
 The red, dashed vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$.
-Note the flat response within the band.}
+Note the flat response within the band characteristic of the Butterworth filter.}
 	\label{fig:butterworth}
@@
 
-A band-pass filter combines a high-pass filter and a low-pass filter to smoothly attenuate frequencies outside of some band.
-A Butterworth band-pass filter is just a particular choice of the frequency response (an attenuation at each frequency) as to be `maximally flat' within the pass-band region.
+A general band-pass filter combines a high-pass filter and a low-pass filter to smoothly attenuate frequencies outside of some band (alt.\ pass-band).
+A Butterworth band-pass filter is a particular band-pass filter such that the frequency response (the attenuation at each frequency) is ``maximally flat'' within the band.
 The Butterworth low-pass component is given by Eqn.~\ref{eq:butterworth} and is combined with a similar high-pass filter to form the band-pass filter.
@@
 
-In the above expression, $f_c$ represents the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ represents the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
+In the above expression, $f_c$ is the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ is the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
 Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here, a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}$, $3\,{\rm kHz})$.
@@
 \caption{\label{fig:notchMagResponse}
-Magnitude response for the first five notches of the $15$ cascaded notch filter described by Eqns.~\ref{eqn:notch} and ~\ref{eqn:notch15}. 
+Amplitude (magnitude) response for the first five notches of fifteen for the cascaded notch filter described by Eqs.~\ref{eqn:notch} and~\ref{eqn:notch15}. 
 }
@@
 One method to remove mains noise and its harmonics is to use a sequence of notch filters centred at each of the frequency bins we want to remove.
-This sequence is known as a cascade of notches.
-Here the notches are smooth in comparison to the naive zeroing of each frequency (which looks like a rectangular comb).
+This sequence is known as a ``cascade'' of notches.
+Here, the notches are smooth in comparison to the naive zeroing of each frequency (which looks like a rectangular comb).
 The complex frequency response of a typical infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
@@
 Where $z$ is the complex frequency and $\alpha$ and $\beta$ are parameters that control the filter.
-$w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed) at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which reflects how quickly the response changes around the notched frequency).
+$w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
 We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centred on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$.
@@
 
-We use the built-in MATLAB filter design toolbox in this work. 
-The amplitude response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
+We use the built-in MATLAB filter design toolbox in this work~\cite{MATLAB}. 
+The amplitude (magnitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
 The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
@@
 % This effect will become more severe if the original human spectrum overlaps with the hum spectrum by a large amount.
-To overcome this, we turn to more advanced techniques starting with the Wiener filter.
+To overcome this, we turn to more advanced techniques, starting with the Wiener filter.
 Instead of just passively filtering different frequencies, this statistical technique optimises an estimate of the original signal.
@@
 
-\subsubsection{Wiener Filter}
+\subsubsection{Wiener filter}
+\label{sec:Wiener}
 
 A Wiener filter is an advanced statistical technique that estimates the injected signal given prior information about the injected spectrum and the reference spectrum of the background noise.
-Let the observed noisy speech signal sequence be given as $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence.
-It is the sum of the original injected signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$,
+Let the observed noisy speech signal sequence be given as $\mathbf{x}=(x(0),\dots, x(N-1))$, where $N$ is the length of the data sequence.
+It is the sum of the original injected signal $\mathbf{s}=(s(0),\dots,s(N-1))$, and the noise sequence $\mathbf{w}=(w(0),\dots,w(N-1))$,
 \begin{equation}
@@
 \end{equation}
-Given the observed data $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean square error (BMSE) between the two, defined as
+Given the observed data $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean-square-error (BMSE) between the two, defined as
 \begin{equation}
@@
 
-If we assume that: i) $\textbf{x}$ is \emph{wide sense stationary}~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
+If we assume that: i) $\textbf{x}$ is ``wide sense stationary''~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
 \begin{equation}
@@
 \end{equation}
-where $\textbf{h}=\{h(0),\dots h(n)\}$ represents the coefficients of an $n$th order Wiener filter.
+where $\textbf{h}=(h(0),\dots h(n))$ represents the coefficients of an $n$th order Wiener filter.
 % What this sentence is trying to convey is not clear to me
@@
 In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}.
-A higher order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required.
+A higher-order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required.
 For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
@@
 
-Here we experiment with applying a combination of the cascaded notch and the Wiener filter to the recorded speech signal.
+Here, we experiment with applying a combination of the cascaded notch and the Wiener filter to the recorded speech signal.
 
 The Wiener filter makes use of statistical information from the speech data and noise.
-It amplifies the part of the signal with high SNR, while suppressing the parts with low SNR.
+It amplifies the part of the signal with high SNR, while suppressing the parts with low SNR (see above Sec.~\ref{sec:Wiener}).
 It is implemented in the form of a finite response filter, which ensures linear phase response and stability (both desirable), but at the cost of high orders computationally.
@@
 	\caption{Time series of an adult male voice (saying ``A cathode ...'') showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
+The source signal is shifted by $0.12$ s to the right to synchronise the manual delay from starting to play the source through the speaker and starting the recording at the Raspberry Pi.
 One second of the one minute recording is shown.
@@
 	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
-The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.}
+The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.
+The change in scale from the top panel is not of concern as the ear hears the relative frequency content and normalisation of the time series fixes any scaling.
+The bottom two panels also show a broad feature from around $750\,{\rm Hz}$ to $1200\,{\rm Hz}$, the origin of which is yet to be determined.}
 	\label{fig:logMMSE_spectrum}
@@
 Loizon~\cite{SubjectiveComparison} compared $13$ speech enhancement methods and found the log minimum means-square error (logMMSE) estimator to be the best among them, qualitatively, at recovering speech.
-In Ref.~\cite{SubjectiveComparison} a comparison is made of $13$ speech enhancement methods, finding the log minimum means-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
-This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean square error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that is measures the MSE between the logarithm of the Fourier amplitudes.
+In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
+This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that is measures the MSE between the logarithm of the Fourier amplitudes.
 This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
@@
 The top two panels of each figure show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison.
-The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} show the time series and corresponding spectral amplitude respectively after applying the logMMSE estimator.
+The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} shows the time series and corresponding spectral amplitude respectively after applying the logMMSE estimator.
 We see significant attenuation of the mains harmonics as well as a general smoothing of the spectrum.
@@
 - should probably justify what you mean by better if you are going to say it here, but I think enough is discussed in section VI for it not to be necessary here - hannah
-To determine this requires a better understanding of the system as discussed in Section~\ref{sec:future_work}.
+To determine this requires a better understanding of the system, as discussed in Section~\ref{sec:future_work}.
 

### 2020-11-19 paper/ifo-conclusions.tex
full read-through before submission

@@
 
-In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector, demonstrating a range of signal processing techniques used within the gravitational-wave community and more broadly.
-Gravitational-wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
-We consider quasi-monochromatic signals as an analogue to continuous wave signals to directly demonstrate techniques used in current continuous gravitational-wave searches.
+In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational wave detector, demonstrating a range of signal processing techniques used within the gravitational wave community and more broadly.
+Gravitational wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
+We consider quasi-monochromatic signals as an analogue to continuous wave signals to directly demonstrate techniques used in current continuous gravitational wave searches.
 We then consider more complex audio (music and speech) for a more general interest in signal processing and explore use of the interferometer as an optical microphone. 
@@
 The wandering frequency signal is recovered using using the Viterbi algorithm. 
-These sections directly demonstrate the techniques used in current continuous gravitational-wave searches. 
+These sections directly demonstrate the techniques used in current continuous gravitational wave searches. 
 %We demonstrate the recovery of a signal with constant frequency and signal with slowly wandering frequency in Sections~\ref{sec:ifo} and ~\ref{sec:single_tone}, respectively.
@@
 For the wandering frequency we successfully recover the injected signal using the Viterbi algorithm.
-These sections directly demonstrate the techniques used in continuous gravitational-wave searches.
+These sections directly demonstrate the techniques used in continuous gravitational wave searches.
 
@@
 The demonstration presented has potential to be adapted for use in the undergraduate laboratory.
-It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
-The increased excitement and public interest in the field of gravitational-wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
-The field of gravitational-wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals such as continuous waves to which this demonstration provides some charming insights.
+It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
+The increased excitement and public interest in the field of gravitational wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational wave research to a wider, non-specialist audience.
+The field of gravitational wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.
+% \jam{``charming'' is very sweet}
 %The demonstration presented here has great potential to be adapted for use in the undergraduate laboratory.
-It can be used to demonstrate topics such as interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques to physics and electrical engineering students.
-Given the increased level of excitement and public interest in the field of gravitational-wave astronomy in recent years, this table-top demonstration can be used as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
+It can be used to demonstrate topics such as interferometer physics; gravitational wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques to physics and electrical engineering students.
+Given the increased level of excitement and public interest in the field of gravitational wave astronomy in recent years, this table-top demonstration can be used as a tool for explaining gravitational wave research to a wider, non-specialist audience.
 The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves, to which this demonstration provides some charming insights.

### 2020-11-19 paper/ifo-futureWork.tex
full read-through before submission

@@
 
-The optical microphone experiment described here can be used to demonstrate and teach a variety of topics including basic optics and photodiode circuits, to signals processing and speech enhancement.
-Here we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.
+The optical microphone experiment described here can be used to demonstrate and teach a variety of topics, including basic optics and photodiode circuits, to signals processing and speech enhancement.
+Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.
 
@@
 %As the Raspberry Pi demands little power and connects to USB it could instead be powered with a commercial battery rather than the mains supply. 
-All other components of the circuit are connected to the Pi. 
+All other components of the circuit draw power from the Pi, so this would isolate the circuit from the mains. 
 %All other components of the circuit, including the photodiode are connected to the Pi. 
@@
 It also would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode.
-The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the signal flow through the system.
+The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the abstract signal flow through the system.
 % commented out whilst James is checking this
@@
 These could perform the final stage of speech enhancement for the optical microphone.
-Interestingly, this connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
+Interestingly, this coincidently connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
 Alternatively, machine learning solutions exist throughout the field that are able to compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.
@@
 WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time.
-In Refs.~\cite{tufekci2000feature,agbinya1996discrete} WT methods are proposed for speech recognition.
+In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT methods are proposed for speech recognition.
 
-Returning to the demonstration of continuous gravitational wave analysis, further additions could include analysing data from multiple interferometers to extract directional information for audio signals.
+Returning to the demonstration of continuous gravitational wave analysis, further extensions could include analysing data from multiple interferometers to extract directional information for audio signals.
+Although, this would require coupling from a separate noise source instead of from a speaker attached directly to one of the mirrors of the interferometer.
 %Finally, considering the desire to demonstrate continuous wave analysis, further improvements to the optical microphone could lead to more advanced demonstration of techniques such as aiming or locking two or more optical microphones in an array.

### 2020-11-19 paper/ifo-introduction.tex
full read-through before submission

@@
 
-In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
+In 2015, the first detection of gravitational waves from the merger of two black holes~\cite{GW150914} marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
 % What are gws and gw detectors 
@@
 They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
-The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects. 
-Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
-Gravitational-wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
+The effect of gravitational waves is a change in lengths: a `stretching and squashing' of the distance between objects. 
+Ground-based gravitational wave observatories such as the Advanced Laser Interferometer Gravitational wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
+Gravitational wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
 Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
@@
 % CW 
-To date the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
+To date the network of gravitational wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
@@
 % changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
-In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
-We then extend its use to an \emph{optical microphone} (using light to capture sound) and present a range of example analysis techniques for educators to use. 
+In this paper, we use a table-top Michelson interferometer as a toy gravitational wave detector designed to detect sound instead of gravitational waves. 
+We then extend its use to an ``optical microphone'' (using light to capture sound) and present a range of example analysis techniques for educators to use. 
 % re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
@@
 % Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
-%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world.
-This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
-These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
+%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational wave research groups around the world.
+This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
+These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational wave science to non-specialist audiences.
 %\han{look up Sarah Farmer's project}
@@
 This paper is laid out as follows. 
-In Section~\ref{sec:ifo} we detail the table-top interferometer design. 
-We demonstrate observing a single tone from a speaker in Section~\ref{sec:single_tone}. 
-In Section~\ref{sec:viterbi_wandering} we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous wave searches.
+In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
+In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
+In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous wave searches.
 In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
@@
 This demonstration of an optical microphone  serves as a more general exhibition of signal processing with a range of examples which can be used in the undergraduate laboratory. 
-We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in ~\ref{sec:conclusions}. 
-Appendix~\ref{app:code} collects the software and scripts used to produce this work.
+We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
+We collect the software and scripts used to produce this work in Appendix~\ref{app:code}.
 

### 2020-11-19 paper/ifo-singleTone.tex
combined notch and Wiener plots
full read-through before submission

@@
 The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has a FWHM of $0.033\,{\rm Hz}$.
-The plot also shows two harmonics at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ with normalised amplitude to the primary peak of $0.103$ and $0.098$ respectively.
+The plot also shows two harmonics at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ with amplitudes $10.3 \%$ and $9.8 \%$ of the amplitude of the primary peak, respectively.
 }	
@@
 
-Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone; a single constant frequency.
+Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone at a single, constant frequency. %should be single constant frequency *signal*?
 
@@
 The webcam records video in three colour channels; red, blue, and green. 
-We take the green channel as an approximation of the total intensity produced by the monochromatic green laser.
+We use the green channel as an approximation of the total intensity produced by the monochromatic green laser.
 The webcam samples at a rate of $30\, {\rm Hz}$, which limits the spectral content of observable signals to less than $15\,{\rm Hz}$, the Nyquist frequency.

### 2020-11-19 paper/ifo-tableTopGWs.tex
minor changes, commented out footmisc
full read-through before submission

@@
 % rephrased to mention the laser as the first component and shifted the arm lengths of LIGO and Virgo to later in the para when we discuss them. 
-Gravitational-wave detectors such as LIGO and Virgo are large, complex experiments. 
+Gravitational wave detectors such as LIGO and Virgo are large, complex experiments. 
 However their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
@@
 The resulting interference pattern is dependent on the relative distance travelled by the beams. 
-Current generation gravitational-wave observatories have ${\rm km}$-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
-They are able to measure minuscule changes in distance due to gravitational-waves, e.g.
+Current generation gravitational wave observatories have kilometre-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
+They are able to measure minuscule changes in distance due to gravitational waves, e.g.
 the first detection of a binary black hole merger produced a strain of $10^{-21}$, corresponding to a change in distance which is a fraction of the width of a proton.
@@
 \begin{comment}
-Gravitational-wave detectors such as LIGO and Virgo are large, complex experiments.
+Gravitational wave detectors such as LIGO and Virgo are large, complex experiments.
 However, their design is based fundamentally on a Michelson interferometer, an optical configuration commonly used in undergraduate laboratories.
@@
 This pattern is dependent on the relative distance travelled by each beam and thus on the relative length of each arm. 
-Current generation gravitational-wave observatories are able to measure minuscule changes in distance due to passing gravitational waves. 
+Current generation gravitational wave observatories are able to measure minuscule changes in distance due to passing gravitational waves. 
 The first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller than the width of a proton.
@@
 % I rephrased to lead with the analogy here and removed 'often' as GW signals are only really listened to for public engagement interest - hannah
-Sound is a commonly-used analogy when explaining gravitational-wave science. 
-Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
+Sound is a commonly-used analogy when explaining gravitational wave science. 
+Gravitational wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
 For example, gravitational waves can propagate in a vacuum, whereas sound cannot.
@@
 %This analogue allows scientists to use sound vibrations as a simple means to move the components of a demonstration interferometer. 
-Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
+Detection and analysis techniques used by the gravitational wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
 Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
 The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
-We use audio signals throughout this work to simulate gravitational-wave-like signals.
+We use audio signals throughout this work to simulate gravitational wave-like signals.
 % and  to demonstrate different data analysis and signal processing techniques used to recover such signals.
@@
 The beam then passes through a converging lens (FL = $125.0\,{\rm mm}$) before reaching the beamsplitter which is $50$/$50$ in reflection and transmission.
-The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and are recombined at the beamsplitter. 
-The output interference pattern is enlarged using a diverging lens (FL = $-25.0\,{\rm mm}$), projected onto a screen, and recorded using a commercial webcam connected to a personal computer (PC). 
+The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and recombine at the beamsplitter. 
+The output interference pattern is enlarged using a diverging lens (FL = $-25.0\,{\rm mm}$), projected onto a screen, and recorded using a commercial webcam connected to a computer (PC). 
 A speaker fixed to the back of M2 is used to inject audio signals from the PC into the interferometer.
@@
 Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
-% Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
+% Ring interference patterns like this are often used in gravitational wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
 % Why remove the sentence above? Rings are chosen rather than happening by default. - Hannah
@@
  \begin{center}
-  \includegraphics[height=.45\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
+  \includegraphics[height=0.3\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
  \end{center}
@@
 The interference pattern produced by the table-top Michelson interferometer.
-This image was taken with the webcam off the beam axis, which may explain why the (predicted spherical) pattern appears elliptical.
+This image was taken with the webcam off the beam axis, which may explain why the (predicted circular) pattern appears elliptical.
 The pink star indicates the point where the intensity time series measurements were taken.
@@
 The amplitude of these motions is given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
-If the relative change in the length of the arms is kept small enough, the intensity of the interference pattern at any point on the screen oscillates at the same frequency as the injected audio.
+If the relative change in the length of the arms is kept small enough, then the intensity of the interference pattern at any point on the screen oscillates at the same frequency as the injected audio.
 But if the relative change in the length of the arms is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting artificially raises the measured frequency.

### 2020-11-19 paper/ifo-wanderingTone.tex
full read-through before submission

@@
 The frequency of the signal can wander significantly over the observation period. 
-In this context \emph{significantly} means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
+In this context, ``significantly'' means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
 One method to search for a wandering signal is to split the time series data into several shorter segments which are analysed individually.%, and then recombine them under some continuity assumption to coherently analyse the whole set.
@@
 
-A \emph{detection statistic} is used to quantify the likelihood of a signal being present in the data at each frequency and timeseries segment. 
-In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
+A ``detection statistic'' is used to quantify the likelihood of a signal being present in the data at each frequency and timeseries segment. 
+In gravitational wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
 %Most continuous wave searches use detection statistics to quantify the likelihood of a signal being present in the data at each frequency for each timeseries segment. 
@@
 This maximises the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}. 
-The detection statistic is normalised for convenience and we use them to form a grid in  time and frequency (a spectrogram). 
+The detection statistic is normalised for convenience and we use it to form a grid in time and frequency (a spectrogram) which displays the frequency content of the signal over time. 
 % moved one sentence up 
@@
 In Section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work. 
-In Section~\ref{sec:wanderingResults} we present results for recovery of a wandering audio frequency.
+In Section~\ref{sec:wanderingResults} we present results for recovery of a wandering frequency audio signal.
 
@@
 The paths between these states are indicated by the lines in Fig.~\ref{fig:viterbi}.
-At each $f_j$ state, we select the path with the highest $A[f(t_0),f(t_1)] {\rm Pr}[f(t_0)]$ as the most probable path.
+At each $f_j$ state, we select the path with the highest $A[f(t_0),f(t_1)] {\rm Pr}[f(t_0)]$ value as the most probable path.
 These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the grey lines show the rejected paths.
@@
 Therefore, this path is selected as the best path from $t_0$ to $t_1$.
-For backtracking at the end, the index of the most probable connection along with the current cost of the best path is stored for each node.
+To allow backtracking at the end, the index of the most probable connection to a node along with the value of the best path to that node is stored, for each node.
 
@@
 A schematic diagram of the Viterbi algorithm. 
-Each box represents an element in the time-frequency grid starting from the instant $t_0$ on the left and ending at $t_{N_t}$ on the right. 
+Each circle represents an element in the time-frequency grid starting from the instant $t_0$ on the left and ending at $t_{N_t}$ on the right. 
 Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
@@
 At each time, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines in the diagram). 
-Each circle has three paths leading to it indicated by the lines. 
+Each circle has three paths leading to it, indicated by the lines. 
 The best path to each circle at each $t_i$ is highlighted in black. 
-Some routes through the grid are found to be dead ends, such as the path ending at $t_1$ marked (a). 
+Some routes through the grid are found to be dead ends, with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a). 
 At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:recursiveViterbi}, marked (b). 
@@
 	\caption{\label{fig:viterbi_overlay}
-Recovery of a wandering tone. 
-The spectrogram shows the observed frequency amplitude in each time-frequency bin. 
-The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path respectively. 
-On the left, before $\sim 15\,{\rm s}$, the signal changes frequency too quickly for the Viterbi algorithm to recover. 
+Recovery of a wandering tone using the Viterbi algorithm. 
+The spectrogram shows the observed frequency (Fourier) amplitude in each time-frequency bin. 
+The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path, respectively. 
+On the left, before $\sim 15\,{\rm s}$, the signal changes frequency too quickly for the Viterbi algorithm to recover, given a parameter for the rate of change of one frequency bin per time bin. 
 At $150\,{\rm s}$ the data appears anomalous, which may be due to some background noise. }
@@
 %A sinusoidal signal with slowly decaying amplitude is played through the speaker to mimic a slowly wandering signal. 
-We use an identical set up to that shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
+We use the same set-up as shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
 %A sinusoidal signal with a slowly decaying amplitude is injected into the interferometer using the speaker in the setup shown in Fig~\ref{fig:ifo_schematic_webcam} to mimic a slowly wandering signal.
@@
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
-For the purposes of this demonstration we use a signal that can easily be identified  in the spectrogram, however real continuous wave signals have much lower amplitudes. 
-The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.% I think this is already defined and seems like repetition:(the most likely path through the data according to the Viterbi algorithm).
+For the purposes of this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous wave signals to have far lower signal-to-noise ratios (SNR's).
+The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.
+% I think this is already defined and seems like repetition:(the most likely path through the data according to the Viterbi algorithm). 
 The recovered Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time.
-We also compute the root mean square (RMS) fractional error $E_{\rm rms}$ along the path which is defined as
+We also compute the root-mean-square (RMS) fractional error $E_{\rm rms}$ along the path, which is defined as
 \begin{equation}

### 2020-11-19 paper/paper_main.tex
updating figures and slight edit of abstract
full read-through before submission

@@
 % \setlength{\parindent}{0pt}
+
+% inner product
+\newcommand{\ip}[2]{\langle #1,#2 \rangle}
 
@@
 \begin{abstract}
-Gravitational-wave observatories around the world are searching for continuous waves: persistent signals from sources such as spinning neutron stars. 
+Gravitational wave observatories around the world are searching for continuous waves: persistent signals from sources such as spinning neutron stars. 
+% \jam{For the hyphen in "gravitational wave" versus "gravitational wave". I prefer the former, but either way, we should be consistent.}
 These searches use sophisticated statistical techniques to look for weak signals in noisy data. 
-In this paper, we demonstrate these techniques using a table-top model gravitational-wave detector (a Michelson interferometer), where sound is used as an analogue for gravitational waves. 
+In this paper, we demonstrate these techniques using a table-top model gravitational wave detector: a Michelson interferometer where sound is used as an analogue for gravitational waves. 
 Using signal processing techniques from continuous-wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
-We explore use of the interferometer as a teaching tool for educators in physics and electrical undergraduate laboratories. 
-The interferometer is used as an \emph{optical microphone} to capture music and speech. 
+We explore use of the interferometer as a teaching tool for educators in physics and electrical undergraduate laboratories with a range of example activities.
+% \jam{We mention undergraduate labs twice, roll this sentence into the end? Also, electrical sounds weird, why the change from electrical engineering?}
+The interferometer is used as an ``optical microphone'' to capture music and speech. 
+% \jam{I changed all the \emph to ``'', feel free to change them back, but I think it is more explicit this way}
 A selection of filtering techniques are applied to recover the original signals from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
-Simple chords and drums are easily recovered, but complex music and speech are more challenging. 
-This demonstration has scope for use by undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
+Recordings of simple chords and drums are easily recovered, but complex music and speech are more challenging. 
+This demonstration has scope for use by undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational wave and signal-processing topics to non-specialist audiences.
 % rearranged in response to Anna's pnp review
@@
 %This demonstration has scope for use in physics and electrical engineering undergraduate laboratories. 
-%It can also be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences. 
+%It can also be adapted for use as an engagement tool for communicating gravitational wave and signal-processing topics to non-specialist audiences. 
 \end{abstract}
@@
 \begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular Anna Green, Lynn Cominsky and Sam Cooper for their helpful feedback and suggestions.
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky and Sam Cooper, for their helpful feedback and suggestions.
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 

### 2020-11-19 paper/ifo-appendix.tex
full grammar check before submission

@@
 Classic sig proc result useful for teachers to know.'']}
-In this appendix, we demonstrate the the modulus of the Fourier Transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
+In this appendix, we demonstrate the modulus of the Fourier transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
 We describe the data as

### 2020-11-19 paper/ifo-complexAudio.tex
full grammar check before submission

@@
 In this section, we explore how this demonstration can be used to teach a selection of signal processing techniques.
-The apparatus used in Sections~\ref{sec:ifo} to~\ref{sec:viterbi_wandering} affords the opportunity to explore a range of signal processing techniques and filters of interest to undergraduate students and provide a foundations for many technical applications~\cite{DigitalProcingOfSpeechSignals:1978}.
+% changed ``affords the opportunity to explore'' to ``allows exploring''
+The apparatus used in Sections~\ref{sec:ifo} to~\ref{sec:viterbi_wandering} allows undergraduate students to explore a range of signal processing techniques and filters of interest and provides a foundation for many technical applications~\cite{DigitalProcingOfSpeechSignals:1978}.
 We provide a series of example activities using complex audio (such as music and speech) as a natural successor to constant and wandering tones. 
@@
 As complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not directly applicable here.
-Instead, in this section, we use the interferometer as an ``optical microphone'' to detect and play back complex audio signals (i.e.\ using light to capture sound).
+Instead, in this section, we use the interferometer as an ``optical microphone'' to detect and playback complex audio signals (i.e.\ using light to capture sound).
 In doing so, we experiment with a hierarchy of passive filters which suppress noise, yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
@@
 
-We place a OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage dependent on the incident intensity. 
+We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage dependent on the incident intensity. 
 The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
 The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath the screen. 
-The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
+The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
 Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. Resources for using the Pi and photodiode are described in Appendix~\ref{app:circuit_diagram}.
@@
 
-Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results, respectively. 
+Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show time series and Fourier spectrum results, respectively. 
 The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} show the source signal and the raw optical microphone recording, respectively. 
@@
 Qualitatively it sounds more like a drum than a human voice.
-Evidently, this is due to loss of voice information under the filter.
+This is due to loss of voice information under the filter.
 
@@
 % \begin{center}
-% \includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/timeSeriesNotchWiener.pdf}
+% \includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/time SeriesNotchWiener.pdf}
 % \end{center}
 % \caption{\label{fig:timeNotchWiener}
-% Timeseries results for the notch and Wiener filters. 
+% Time series results for the notch and Wiener filters. 
 % The first panel shows the original speech recording which was played from the speaker. 
@@
 % The third panel shows the spectrum after applying both the notch and Wiener filters. 
-% The first, second, and third panels show the spectrum of the timeseries show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
+% The first, second, and third panels show the spectrum of the time series show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
 % }
@@
 where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data, the injected signal, and the background noise, respectively.
-Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high, and attenuates the signal where the SNR is low.
+Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high and attenuates the signal where the SNR is low.
 The causal Wiener filter (that only makes estimates based on past observations) is similar.
@@
 The time series and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-We see a significant improvement in the time series of the recovered signal, however a strong noise hum persists, audibly. 
+We see a significant improvement in the time series of the recovered signal, however, a strong noise hum persists, audibly. 
 %The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
@@
 %\han{should we add more details here too?} \jam{[seconded]}
-%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
-%The timeseries is visibly cleaner, however there is a strong noise hum.
+%The time series and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
+%The time series is visibly cleaner, however there is a strong noise hum.
 
@@
 The Wiener filter makes use of statistical information from the speech data and noise.
-It amplifies the part of the signal with high SNR, while suppressing the parts with low SNR (see above Sec.~\ref{sec:Wiener}).
+It amplifies the part of the signal with high SNR while suppressing the parts with low SNR (see above Sec.~\ref{sec:Wiener}).
 It is implemented in the form of a finite response filter, which ensures linear phase response and stability (both desirable), but at the cost of high orders computationally.
@@
 
-By combining the notch and Wiener filter, we can trade off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:BackgroundNotchWienerCombined}..
+By combining the notch and Wiener filter, we can trade-off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
 The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone.
@@
 The source signal is shifted by $0.12$ s to the right to synchronise the manual delay from starting to play the source through the speaker and starting the recording at the Raspberry Pi.
-One second of the one minute recording is shown.
+One second of the one minute long recording is shown.
 The rise at the start of the bottom panel is expected when filtering a signal of finite duration.}
@@
 We see significant attenuation of the mains harmonics as well as a general smoothing of the spectrum.
-The logMMSE method indeed removes most of the background noise, but does not significantly enhance the speech as the voice sounds muffled and indistinct.
+The logMMSE method indeed removes most of the background noise but does not significantly enhance the speech as the voice sounds muffled and indistinct.
 % suggest remove as it's just rephrasing what's already been said

### 2020-11-19 paper/ifo-conclusions.tex
full grammar check before submission

@@
 We consider quasi-monochromatic signals as an analogue to continuous wave signals to directly demonstrate techniques used in current continuous gravitational wave searches.
-We then consider more complex audio (music and speech) for a more general interest in signal processing and explore use of the interferometer as an optical microphone. 
-A variety of example activities are presented. 
-We find that the logMMSE produces the best results, however it does not enhance the recording to an intelligible level. 
-The complex audio signals can also serve as more distant analogues to un-modelled gravitaitonal-wave burst signals, e.g. from supernova.
+We then consider more complex audio (music and speech) for a more general interest in signal processing and explore the use of the interferometer as an optical microphone. 
+A variety of example activities is presented. 
+We find that the logMMSE produces the best results, however, it does not enhance the recording to an intelligible level. 
+The complex audio signals can also serve as more distant analogues to un-modelled gravitational wave burst signals, e.g. from supernovae.
 
@@
 
-The demonstration presented has potential to be adapted for use in the undergraduate laboratory.
+The demonstration presented has the potential to be adapted for use in the undergraduate laboratory.
 It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.

### 2020-11-19 paper/ifo-futureWork.tex
full grammar check before submission

@@
 % reduced length
-Isolating the experiment from interference may provide some reduction in noise. 
+Isolating the experiment from interference may provide some noise reduction. 
 For example, the Raspberry Pi could be powered with a commercial battery instead of mains power. 
@@
 %With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
-A feedback loop control system can also be used to suppress unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational wave observatories.
+A feedback loop control system can also be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational wave observatories.
 %We can consider improving the stability of the optical microphone components.
@@
 Interestingly, this coincidently connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
-Alternatively, machine learning solutions exist throughout the field that are able to compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.
+Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.
 

### 2020-11-19 paper/ifo-introduction.tex
full grammar check before submission

@@
 % CW 
-To date the network of gravitational wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
+To date, the network of gravitational wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
@@
 % removed Additionally - I don't think it is neccessary - hannah
-It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, opto-mechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
+It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
 % This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.

### 2020-11-19 paper/ifo-tableTopGWs.tex
full grammar check before submission

@@
 Gravitational wave detectors such as LIGO and Virgo are large, complex experiments. 
-However their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
+However, their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
 In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms (see Fig.~\ref{fig:ifo_schematic_webcam}). 
@@
 Current generation gravitational wave observatories have kilometre-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
-They are able to measure minuscule changes in distance due to gravitational waves, e.g.
+They can measure minuscule changes in distance due to gravitational waves, e.g.
 the first detection of a binary black hole merger produced a strain of $10^{-21}$, corresponding to a change in distance which is a fraction of the width of a proton.

### 2020-11-19 paper/ifo-wanderingTone.tex
full grammar check before submission

@@
 
-A ``detection statistic'' is used to quantify the likelihood of a signal being present in the data at each frequency and timeseries segment. 
+A ``detection statistic'' is used to quantify the likelihood of a signal being present in the data at each frequency and time series segment. 
 In gravitational wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
@@
 The choice of detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
-In this work, we take the discrete Fourier transform of each timeseries segment and use its amplitude at each frequency as a detection statistic.
+In this work, we take the discrete Fourier transform of each time series segment and use its amplitude at each frequency as a detection statistic.
 This maximises the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}. 
@@
 Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
-The colour of each circle indicates the likelihood that a signal is present in the box, where a lighter colour corresponds to higher likelihood. 
+The colour of each circle indicates the likelihood that a signal is present in the box, where a lighter colour corresponds to a higher likelihood. 
 At each time, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines in the diagram). 
@@
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
-For the purposes of this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous wave signals to have far lower signal-to-noise ratios (SNR's).
+For this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous wave signals to have far lower signal-to-noise ratios (SNR's). 
 The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.
@@
 Initially, the injected signal wanders by more than one frequency bin per time bin (i.e.\ faster than the algorithm is allowed to), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
-One may be tempted to increase the allowed frequency wander in the algorithm, however this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.
+One may be tempted to increase the allowed frequency wander in the algorithm, however, this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.
 There is also an anomaly at $150\,{\rm s}$, which is likely due to a local disturbance, e.g.
 someone walking past the interferometer.
-As shown in Fig.~\ref{fig:viterbi_overlay}, the Viterbi algorithm is able to recover and continues to track the signal after the disturbance. 
+As shown in Fig.~\ref{fig:viterbi_overlay}, the Viterbi algorithm can recover and continues to track the signal after the disturbance. 
 %The Viterbi algorithm is somewhat robust against such disturbances, qualitatively recovering the injected path on the other side, as shown in Fig.~\ref{fig:viterbi_overlay}.

### 2020-11-19 paper/paper_main.tex
full grammar check before submission

@@
 Gravitational wave observatories around the world are searching for continuous waves: persistent signals from sources such as spinning neutron stars. 
-% \jam{For the hyphen in "gravitational wave" versus "gravitational wave". I prefer the former, but either way, we should be consistent.}
+% \jam{For the hyphen in "gravitational wave" versus "gravitational-wave". I prefer the former, but either way, we should be consistent.}
 These searches use sophisticated statistical techniques to look for weak signals in noisy data. 
@@
 Using signal processing techniques from continuous-wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
-We explore use of the interferometer as a teaching tool for educators in physics and electrical undergraduate laboratories with a range of example activities.
+We explore the use of the interferometer as a teaching tool for educators in physics and electrical undergraduate laboratories with a range of example activities.
 % \jam{We mention undergraduate labs twice, roll this sentence into the end? Also, electrical sounds weird, why the change from electrical engineering?}
@@
 % \jam{I changed all the \emph to ``'', feel free to change them back, but I think it is more explicit this way}
-A selection of filtering techniques are applied to recover the original signals from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
+A selection of filtering techniques is applied to recover the original signals from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
 Recordings of simple chords and drums are easily recovered, but complex music and speech are more challenging. 

### 2020-11-19 paper/ifo-introduction.tex
small edits

@@
 
-In 2015, the first detection of gravitational waves from the merger of two black holes~\cite{GW150914} marked a breakthrough in modern astrophysics and revealed a new means to observe the universe.
-% What are gws and gw detectors 
+In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
 Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}, 
@@
 The effect of gravitational waves is a change in lengths: a `stretching and squashing' of the distance between objects. 
-Ground-based gravitational wave observatories such as the Advanced Laser Interferometer Gravitational wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
-Gravitational wave observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
-Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
+Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
+These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
+Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
 
 % CW 
-To date, the network of gravitational wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
+To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
 Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
-Rotating neutron stars are prime candidates for continuous wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
+Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
 %where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
@@
 % changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
-In this paper, we use a table-top Michelson interferometer as a toy gravitational wave detector designed to detect sound instead of gravitational waves. 
+In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
 We then extend its use to an ``optical microphone'' (using light to capture sound) and present a range of example analysis techniques for educators to use. 
@@
 %This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational wave research groups around the world.
-This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
-These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational wave science to non-specialist audiences.
+This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
+These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
 %\han{look up Sarah Farmer's project}
@@
 In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
-In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous wave searches.
+In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches.
 In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.

### 2020-11-19 paper/ifo-singleTone.tex
full grammar check before submission
small edits

@@
 
-Continuous-wave searches look for nearly monochromatic signals. In this section we consider a simple sinusoidal tone at a single, constant frequency. %should be single constant frequency *signal*?
+Continuous-wave searches look for nearly monochromatic signals. In this section, we consider a simple sinusoidal tone at a single, constant frequency. %should be single constant frequency *signal*?
 
@@
 
-%The pixel intensity in the green-channel of the video (the webcam records in three colour channels: red, blue, and green) is used to approximate the total intensity of the interference pattern. 
-%This is because the laser emits monochromatic green light at $532\,{\rm nm}$.
 A tone with a frequency of $2.09\,{\rm Hz}$ is played through the speaker for one minute. 
 The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}. 
-% removed 'strong' as what does strong mean here?
 We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$. 
-%Although this range doesn't contain our injected frequency, the error is not substantial and cannot be heard by the (untrained) human ear.
 Two harmonics can also be seen at integer multiples of the peak frequency. 
@@
 
-In the following section, we use a hidden Markov model technique, commonly used in continuous wave searches, to recover a slowly wandering frequency signal.
+In the following section, we use a hidden Markov model technique, commonly used in continuous-wave searches, to recover a slowly wandering frequency signal.
 

### 2020-11-19 paper/ifo-tableTopGWs.tex
small edits

@@
 
-% rephrased to mention the laser as the first component and shifted the arm lengths of LIGO and Virgo to later in the para when we discuss them. 
-Gravitational wave detectors such as LIGO and Virgo are large, complex experiments. 
+Gravitational-wave detectors such as LIGO and Virgo are large, complex experiments. 
 However, their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
 In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms (see Fig.~\ref{fig:ifo_schematic_webcam}). 
-Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern, 
+Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern.
 The resulting interference pattern is dependent on the relative distance travelled by the beams. 
-Current generation gravitational wave observatories have kilometre-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
+Current generation gravitational-wave observatories have kilometre-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
 They can measure minuscule changes in distance due to gravitational waves, e.g.
 the first detection of a binary black hole merger produced a strain of $10^{-21}$, corresponding to a change in distance which is a fraction of the width of a proton.
-%changed the relative length of the two arms by $10^{-21}$, a distance smaller than the width of a proton~\cite{GW150914}.
 
-\begin{comment}
-Gravitational wave detectors such as LIGO and Virgo are large, complex experiments.
-However, their design is based fundamentally on a Michelson interferometer, an optical configuration commonly used in undergraduate laboratories.
-It consists of a beam splitter which splits light from a laser down two perpendicular arms. % I suggest moving this to later as we are talking about Michelson interferometers generally here. 
-%($4\,{\rm km}$ and $3\,{\rm km}$ long at LIGO and Virgo respectively).
-Mirrors at the end of each arm reflect the light back towards the beam splitter, where they recombine and produce an interference pattern.
-This pattern is dependent on the relative distance travelled by each beam and thus on the relative length of each arm. 
-Current generation gravitational wave observatories are able to measure minuscule changes in distance due to passing gravitational waves. 
-The first detection of a binary black hole merger changed the relative length of the two arms by $10^{-21}$~\cite{GW150914}, smaller than the width of a proton.
-\end{comment}
-
-% I rephrased to lead with the analogy here and removed 'often' as GW signals are only really listened to for public engagement interest - hannah
-Sound is a commonly-used analogy when explaining gravitational wave science. 
-Gravitational wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
+Sound is a commonly-used analogy when explaining gravitational-wave science. 
+Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
 For example, gravitational waves can propagate in a vacuum, whereas sound cannot.
 Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
-% rephrased as this what we are doing, rather than what scientists in general are doing
-%This analogue allows scientists to use sound vibrations as a simple means to move the components of a demonstration interferometer. 
-Detection and analysis techniques used by the gravitational wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
+Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
 Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
 The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
-We use audio signals throughout this work to simulate gravitational wave-like signals.
-% and  to demonstrate different data analysis and signal processing techniques used to recover such signals.
+We use audio signals throughout this work to simulate gravitational-wave-like signals.
 
@@
 Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
-% Ring interference patterns like this are often used in gravitational wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
-% Why remove the sentence above? Rings are chosen rather than happening by default. - Hannah
 The intensity timeseries is recorded via webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
 In Section~\ref{sec:optical_microphone}, which considers more complex audio, the webcam is replaced with a photodiode, which offers a higher sample rate than the webcam.
-%rephrased as this sentence is confusing I think: An intensity time series of some point in the interference pattern is recorded by a webcam (or a photodiode in Section~\ref{sec:optical_microphone}) for data analysis to extract the signal injected at the speaker.
 
@@
 The interference pattern produced by the table-top Michelson interferometer.
-This image was taken with the webcam off the beam axis, which may explain why the (predicted circular) pattern appears elliptical.
+This image was taken with the webcam off the beam axis.%, which may explain why the (predicted circular) pattern appears elliptical.
 The pink star indicates the point where the intensity time series measurements were taken.

### 2020-11-19 paper/ifo-wanderingTone.tex
small edits

@@
 
-% the signal and the analogy we use
-Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational waves may not be monochromatic.
-%Continuous gravitational waves may not be monochromatic, unlike the signal considered in Section~\ref{sec:single_tone}.  
+Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational-waves may not be monochromatic.
 A continuous-wave signal may wander slowly in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
@@
 The analysis techniques used in this section are inspired by those used to search for continuous waves from spinning neutron stars in LIGO and Virgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
-Continuous wave searches are performed on long datasets, months to years in duration. 
+Continuous-wave searches are performed on long datasets, months to years in duration. 
 The frequency of the signal can wander significantly over the observation period. 
@@
 A ``detection statistic'' is used to quantify the likelihood of a signal being present in the data at each frequency and time series segment. 
-In gravitational wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
+In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
 %Most continuous wave searches use detection statistics to quantify the likelihood of a signal being present in the data at each frequency for each timeseries segment. 
@@
 %To perform the task of recovering the signal, we use the Viterbi algorithm~\cite{Viterbi:1967} to efficiently find the most probable path through the spectrogram given the sequence of observations.
-This method is used in continuous wave searches for a variety of astrophysical targets~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application} and forms the basis of the analysis in this section.
+This method is used in continuous-wave searches for a variety of astrophysical targets~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application} and forms the basis of the analysis in this section.
 
@@
 
-The spectrogram, as shown in Fig.~\ref{fig:viterbi}, has $N_t$ time and $N_f$ frequency bins. 
-% why add to the confusion? why not make things consistent? 
-%From here on, we will call these time “steps” instead of time bins, but they should not be confused with the timesteps of the original time series.
-Reading through it seems we were using both steps and bins in this section anyway.
+The spectrogram schematic, as shown in Fig.~\ref{fig:viterbi}, has $N_t$ time and $N_f$ frequency bins. 
 We label the elements of the time and frequency  grid as $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-% state is means frequency not grid point doesn't it? 
-%At each state (grid point) the detection statistic (the normalised Fourier amplitude) is notated $F(t_i,f_j)$.
 The normalised Fourier amplitude $F(t_i,f_j)$ is the detection statistic.
@@
 For example, the most probable connection to the element labelled (a) is the one directly behind it (i.e., $f_4$).
-Therefore, this path is selected as the best path from $t_0$ to $t_1$.
+Therefore, this path is selected as the best path from $t_0$ to $t_1$ for $f_4$.
 To allow backtracking at the end, the index of the most probable connection to a node along with the value of the best path to that node is stored, for each node.
@@
 
-% throughout this section we are using present tense, should this be changed to past tense?
-% I think we should go for present tense throughout - hannah
-
 To simulate a slowly wandering signal we modulate the frequency sinusoidally with a modulation amplitude that decays with time. 
-%A sinusoidal signal with slowly decaying amplitude is played through the speaker to mimic a slowly wandering signal. 
 We use the same set-up as shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
-%A sinusoidal signal with a slowly decaying amplitude is injected into the interferometer using the speaker in the setup shown in Fig~\ref{fig:ifo_schematic_webcam} to mimic a slowly wandering signal.
-The particular choice of signal is arbitrary as any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
-The output of the interferometer is recorded via a webcam as in Section~\ref{sec:single_tone}.
 In this section we test the Viterbi algorithm’s ability to recover the wandering signal.
@@
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
-For this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous wave signals to have far lower signal-to-noise ratios (SNR's). 
+For this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous-wave signals to have far lower signal-to-noise ratios. 
 The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.
-% I think this is already defined and seems like repetition:(the most likely path through the data according to the Viterbi algorithm). 
 The recovered Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time.

### 2020-11-19 paper/paper_main.tex
small edits

@@
 \begin{abstract}
-Gravitational wave observatories around the world are searching for continuous waves: persistent signals from sources such as spinning neutron stars. 
+Gravitational-wave observatories around the world are searching for continuous waves: persistent signals from sources such as spinning neutron stars. 
 % \jam{For the hyphen in "gravitational wave" versus "gravitational-wave". I prefer the former, but either way, we should be consistent.}
 These searches use sophisticated statistical techniques to look for weak signals in noisy data. 
-In this paper, we demonstrate these techniques using a table-top model gravitational wave detector: a Michelson interferometer where sound is used as an analogue for gravitational waves. 
+In this paper, we demonstrate these techniques using a table-top model gravitational-wave detector: a Michelson interferometer where sound is used as an analogue for gravitational waves. 
 Using signal processing techniques from continuous-wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
-We explore the use of the interferometer as a teaching tool for educators in physics and electrical undergraduate laboratories with a range of example activities.
+We explore the use of the interferometer as a teaching tool for educators in physics and electrical engineering by using it as an ``optical microphone to capture music and speech. 
 % \jam{We mention undergraduate labs twice, roll this sentence into the end? Also, electrical sounds weird, why the change from electrical engineering?}
-The interferometer is used as an ``optical microphone'' to capture music and speech. 
+%The interferometer is used as an ``optical microphone'' to capture music and speech. 
 % \jam{I changed all the \emph to ``'', feel free to change them back, but I think it is more explicit this way}
-A selection of filtering techniques is applied to recover the original signals from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
+We present a range of example filtering activities to recover the original signal from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
 Recordings of simple chords and drums are easily recovered, but complex music and speech are more challenging. 
-This demonstration has scope for use by undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational wave and signal-processing topics to non-specialist audiences.
+This demonstration has scope for use by educators in undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
 % rearranged in response to Anna's pnp review

### 2020-11-19 paper/ifo-complexAudio.tex
slight reshuffle of figures and small edits in complex audio and future

@@
 
-% Optical microphone
-% aim of optical microphone: play back injected sounds
-% some editing in response to pnp
 In this section, we explore how this demonstration can be used to teach a selection of signal processing techniques.
-% changed ``affords the opportunity to explore'' to ``allows exploring''
 The apparatus used in Sections~\ref{sec:ifo} to~\ref{sec:viterbi_wandering} allows undergraduate students to explore a range of signal processing techniques and filters of interest and provides a foundation for many technical applications~\cite{DigitalProcingOfSpeechSignals:1978}.
 We provide a series of example activities using complex audio (such as music and speech) as a natural successor to constant and wandering tones. 
-%A natural successor to constant and wandering tones is complex audio, such as music and speech. 
-As complex audio signals are not quasi-monochromatic, unlike continuous gravitational wave signals, the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not directly applicable here.
-Instead, in this section, we use the interferometer as an ``optical microphone'' to detect and playback complex audio signals (i.e.\ using light to capture sound).
+As complex audio signals are not quasi-monochromatic, unlike continuous gravitational-wave signals, the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not directly applicable here.
+Instead, in this section, we use the interferometer as an optical microphone to detect and playback complex audio signals (i.e.\ using light to capture sound).
 In doing so, we experiment with a hierarchy of passive filters which suppress noise, yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
@@
 
-%An introduction to ..... can be found in Ref.~\cite{}, including 
-%A review of digital processing techniques for speech signals can be found 
-%It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae~\cite{}\han{to do - add citation}. 
-%It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae. [refs; my last sentence can be improved]
-
-This section is indirectly related to the advanced signal processing techniques required to extract gravitational wave signals.
+This section is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals.
 Here, the optical microphone serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories.
@@
 \label{sec:photodiode}
-% advanced method: explain how we capture data
+
+\begin{figure*}
+	\includegraphics[width=0.44\textwidth]{figures/ifo_schematic_photodiode_edit.pdf}
+	\includegraphics[width=0.52\textwidth]{figures/setup_pic2.pdf}
+	\caption{
+Left: the updated Michelson interferometer schematic. 
+The schematic is identical to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
+Right: a photograph of the optical microphone. 
+The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. 
+In the main image, the photodiode is placed behind the cloth screen. 
+The inset shows a face-on view of the photodiode with the covering screen removed. 
+See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.
+}
+	\label{fig:ifo_schematic_podo}
+\end{figure*}
 
@@
 
-\begin{figure*}
-	\includegraphics[width=0.44\textwidth]{figures/ifo_schematic_photodiode_edit.pdf}
-	\includegraphics[width=0.52\textwidth]{figures/setup_pic2.pdf}
-	\caption{
-Left: the updated Michelson interferometer schematic. 
-The schematic is identical to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
-Right: a photograph of the optical microphone. 
-The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. 
-In the main image, the photodiode is placed behind the cloth screen. 
-The inset shows a face-on view of the photodiode with the covering screen removed. 
-See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.
-}
-	\label{fig:ifo_schematic_podo}
-\end{figure*}
-
-\begin{comment}
-\begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/setup_pic2.pdf}
-\caption{Photograph of the optical microphone.
-The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right.
-The interferometer is laid out as in Figure~\ref{fig:ifo_schematic_podo}.
-In the main image, the photodiode is placed behind the cloth screen.
-The inset shows a face-on view of the photodiode with the covering screen removed.
-See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.}
-	\label{fig:setup_pic2}
-\end{figure}
-\end{comment}
-
 \subsection{Anti-aliased output}
@@
 	\caption{\label{fig:psd_noise}
-Power spectral density (PSD) of background noise (top panel) from the optical microphone (with the speaker off), and the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines.
+Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
+Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
 We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
@@
 Here, we describe several filters used to remove the $50\,{\rm Hz}$ mains hum and its harmonics, but also to ultimately improve the speech intelligibility of the recording (to be able to understand the words spoken).
-We start with some naive approaches and then move to traditional signal processing filters like band-passing and cascaded notches.
-We finish by combining these with advanced statistical techniques like a Wiener filter and the logMMSE estimator.
+We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
+We finish by combining these with advanced statistical techniques (Wiener filter and the logMMSE estimator). 
 As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
@@
 	Time series (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
-	The first and second rows show the original speech recording and raw output from the optical microphone, respectively. 
+    The source signal, shown in the first row, is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode ...''). 
+    The source signal is shifted by $0.12\,{\rm s}$ to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker. 
+	The second row shows the raw output from the optical microphone when the input signal from the first row is played. 
     The third row shows a background noise recording from the optical microphone when no audio signal is played. 
     The fourth row shows the result of subtracting the background noise spectrum in the third row from the recording in the second row. 
-	%The second row shows the raw output from the optical microphone.
-	%The third row shows the naive technique of subtracting off the noise spectrum from the recorded spectrum.
 	The fifth, sixth, and seventh rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
@@
 Given that we have access to the background noise spectrum, an intuitive way to remove noise is via subtracting the noise spectrum from the recorded spectrum. 
-The fourth panel of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the spectrum obtained after subtracting the background noise spectrum. 
+The fourth row of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the spectrum obtained after subtracting the background noise spectrum. 
 We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified. %Why the noise sources are time-varying is a fact yet to be determined.
 
-Beyond this naive approach, we apply a range of filters to the data.
-% \jam{changed na\"{i}ve to naive, if you change this back please do it globally}
+Beyond this na{\"i}ve approach, we apply a range of filters to the data.
 In signals processing, the ideal filter would be one that:
-%\begin{itemize}
 i) completely attenuates the undesired parts of the spectrum, 
@@
 iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
-%\end{itemize}
 However, these three conditions cannot all hold at once and any filter must compromise between them, as we will see below. 
+The following sections describe a series of example analyses with the optical microphone.
 
@@
 A Butterworth band-pass filter is a particular band-pass filter such that the frequency response (the attenuation at each frequency) is ``maximally flat'' within the band.
-The Butterworth low-pass component is given by Eqn.~\ref{eq:butterworth} and is combined with a similar high-pass filter to form the band-pass filter.
+The Butterworth low-pass component is given by %Eqn.~\ref{eq:butterworth}
 
@@
 %\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
-
-In the above expression, $f_c$ is the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ is the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
+and is combined with a similar high-pass filter to form the band-pass filter.
+In Eqn.~\ref{eq:butterworth}, $f_c$ is the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ is the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
 Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here, a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}$, $3\,{\rm kHz})$.
@@
 This sequence is known as a ``cascade'' of notches.
-Here, the notches are smooth in comparison to the naive zeroing of each frequency (which looks like a rectangular comb).
+Here, the notches are smooth in comparison to the na{\"i}ve zeroing of each frequency (which looks like a rectangular comb).
 The complex frequency response of a typical infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
@@
     \label{eqn:notch}
-    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}
+    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}\,,
 \end{equation}
-Where $z$ is the complex frequency and $\alpha$ and $\beta$ are parameters that control the filter.
+where $z$ is the complex frequency and $\alpha$ and $\beta$ are parameters that control the filter.
 $w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
@@
 This is due to loss of voice information under the filter.
-
-% This effect will become more severe if the original human spectrum overlaps with the hum spectrum by a large amount.
 To overcome this, we turn to more advanced techniques, starting with the Wiener filter.
 Instead of just passively filtering different frequencies, this statistical technique optimises an estimate of the original signal.
-% By purely minimizing the "distance" between estimated signal (filtered human voice) and observed signal (original human voice), we hope that we keep signal part undisturbed while filtering out noise.
 
@@
 A Wiener filter is an advanced statistical technique that estimates the injected signal given prior information about the injected spectrum and the reference spectrum of the background noise. 
-Let the observed noisy speech signal sequence be given as $\mathbf{x}=(x(0),\dots, x(N-1))$, where $N$ is the length of the data sequence.
-It is the sum of the original injected signal $\mathbf{s}=(s(0),\dots,s(N-1))$, and the noise sequence $\mathbf{w}=(w(0),\dots,w(N-1))$,
+The observed noisy speech signal sequence is given as $\mathbf{x}=(x(0),\dots, x(N-1))$, where $N$ is the length of the data sequence and $\mathbf{x}$ is the sum of the original injected signal $\mathbf{s}=(s(0),\dots,s(N-1))$, and the noise sequence $\mathbf{w}=(w(0),\dots,w(N-1))$,
 \begin{equation}
@@
 \end{equation}
-Given the observed data $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean-square-error (BMSE) between the two, defined as
+Given $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean-square-error (BMSE) between the two, defined as 
 \begin{equation}
 \label{eq:BMSE}
-\text{BMSE}(\hat{\textbf{s}})=E[(\textbf{s}-\hat{\textbf{s}})^2]\,,
+\text{BMSE}(\hat{\textbf{s}})=E[(\textbf{s}-\hat{\textbf{s}})^2]\,.
 \end{equation}
-
 If we assume that: i) $\textbf{x}$ is ``wide sense stationary''~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
@@
 \label{eqn:wiener}
-    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}
+    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}\,,
 \end{equation}
@@
 
+% I just combined these into a single figure and also moved some of the intro text about the signal into fig 8 so that it is clear we are all talking about the same data in these plots - hannah
+\begin{figure*}
+	%\includegraphics[width=0.49\textwidth]{figures/combined_timeseries_melatos_shift_source.pdf}
+	%\includegraphics[width=0.49\textwidth]{figures/combined_spectrum_melatos.pdf}
+    \includegraphics[height=0.57\textheight]{figures/combined_timeseries_melatos_shift_source_tall.pdf}
+    \includegraphics[height=0.57\textheight]{figures/combined_spectrum_melatos_tall.pdf}
+    \caption{\label{fig:logMMSE_timeseries_freqspectrum}
+Timeseries (left column) and frequency spectrum (right columns) results with the logMMSE estimator. 
+The original source signal and optical microphone recording in the top and middle panels respectively are for comparison and are identical to those shown in the top two panels in Fig.~\ref{fig:BackgroundNotchWienerCombined} (again, the source signal is shifted by $0.12\,{\rm s}$).
+The bottom-left panel shows the timeseries of the recording after filtering with the logMMSE estimator, where the rise at the start of the timeseries is expected when filtering a signal of finite duration.
+The bottom-right panel shows the corresponding frequency spectrum with the logMMSE estimator. 
+The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude $2$ which otherwise dominate the plot. 
+The change in scale from the top panel is not of concern as the ear hears the relative frequency content and normalisation of the timeseries fixes any scaling. 
+% my feeling is not to include this sentence here as it is hard to be certain what the noise and what is the input signal for this plot without further investigation - hannah
+%The bottom two panels also show a broad feature from around $750\,{\rm Hz}$ to $1200\,{\rm Hz}$, the origin of which is yet to be determined.
+    }
+\end{figure*}
+
 \subsubsection{logMMSE estimator}
@@
 
+\begin{comment}
 \begin{figure*}
@@
 \end{figure*}
+\end{comment}
 
@@
 
-The results of the logMMSE estimator are shown in Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum}.
-The top two panels of each figure show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison.
-The third panel of Figs.~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} shows the time series and corresponding spectral amplitude respectively after applying the logMMSE estimator.
+The timeseries and frequency spectrum results of the logMMSE estimator are shown in Fig.~\ref{fig:logMMSE_timeseries_freqspectrum}. 
+The top two rows show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison (identical to the top two rows in Fig.~\ref{fig:BackgroundNotchWienerCombined}).
+The third row shows the timeseries and corresponding spectral amplitude after applying the logMMSE estimator. 
 We see significant attenuation of the mains harmonics as well as a general smoothing of the spectrum. 
@@
 Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums are louder in the results.
-
 To address these problems, we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.

### 2020-11-19 paper/ifo-futureWork.tex
slight reshuffle of figures and small edits in complex audio and future

@@
 
-% reduced length
 Isolating the experiment from interference may provide some noise reduction. 
 For example, the Raspberry Pi could be powered with a commercial battery instead of mains power. 
-%An initial improvement could be to better isolate the experiment, especially from the sources of mains noise. 
-%As the Raspberry Pi demands little power and connects to USB it could instead be powered with a commercial battery rather than the mains supply. 
 All other components of the circuit draw power from the Pi, so this would isolate the circuit from the mains. 
-%All other components of the circuit, including the photodiode are connected to the Pi. 
 The laser is DC powered through a converter, so should not couple noise from the mains into the interferometer. 
 Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}.
-%With the interferometer in a dark room and under a box to better isolate from any ambient light or air currents, this would hopefully remove (or better separate) all sources of the 50Hz mains hum from the optical microphone.
-A feedback loop control system can also be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational wave observatories.
-%We can consider improving the stability of the optical microphone components.
-A feedback loop control system can be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}.
+A feedback loop control system can also be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational-wave observatories.
 
@@
 
-A thorough characterisation of the systems transfer function would allow us to understand how audio signals couple through the interferometer. 
-%We can also benefit from fully characterizing the transfer function of the system as it provides the exact information of how audio signal couples through the interferometer. 
+A thorough characterisation of the system's transfer function would allow us to understand how audio signals couple through the interferometer. 
 The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Pi.
@@
 
-Returning to the demonstration of continuous gravitational wave analysis, further extensions could include analysing data from multiple interferometers to extract directional information for audio signals.
-Although, this would require coupling from a separate noise source instead of from a speaker attached directly to one of the mirrors of the interferometer.
-%Finally, considering the desire to demonstrate continuous wave analysis, further improvements to the optical microphone could lead to more advanced demonstration of techniques such as aiming or locking two or more optical microphones in an array.
-% I think locking is the wrong word to use here as it is a word used a lot in GW instrumentation (an interferometer can be 'locked' on its own).
-I've tried to rephrase based on what I think you are getting at.
--hannah
-
-%To remove the mains hum we could also monitor it separately from the interferometer and subtract the sinusoids with the right phase and amplitude in real (not frequency) space to avoid the problems of Section~\ref{sec:simple_filters}.
-%This could be done with the existing signals if the recording duration is long enough to precisely measure the phase and amplitude over a large number of cycles.
+Returning to the demonstration of continuous gravitational-wave analysis, further extensions could include analysing data from multiple interferometers to extract directional information for audio signals.
+This would require coupling from a separate noise source instead of from a speaker attached directly to one of the mirrors of the interferometer.
 

### 2020-11-19 paper/ifo-complexAudio.tex
small consistency edits

@@
 During recordings, care is taken to minimise activity around the demonstration to reduce the amount of environmental noise coupling into the interferometer.
-The time series data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python~\cite{scipy,python}.
+The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python~\cite{scipy,python}.
 When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.
@@
 	\caption{\label{fig:BackgroundNotchWienerCombined}
-	Time series (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
+	Timeseries (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
     The source signal, shown in the first row, is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode ...''). 
@@
 
-Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show time series and Fourier spectrum results, respectively. 
+Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results, respectively. 
 The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} show the source signal and the raw optical microphone recording, respectively. 
@@
 For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
-The time series and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-We see a significant improvement in the time series of the recovered signal, however, a strong noise hum persists, audibly. 
+The timeseries and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
+We see a significant improvement in the timeseries of the recovered signal, however, a strong noise hum persists, audibly. 
 %The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 

### 2020-11-19 paper/ifo-conclusions.tex
small consistency edits

@@
 
-In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational wave detector, demonstrating a range of signal processing techniques used within the gravitational wave community and more broadly.
-Gravitational wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
-We consider quasi-monochromatic signals as an analogue to continuous wave signals to directly demonstrate techniques used in current continuous gravitational wave searches.
+In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector, demonstrating a range of signal processing techniques used within the gravitational-wave community and more broadly.
+Gravitational-wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
+We consider quasi-monochromatic audio signals as an analogue to continuous-wave signals to directly demonstrate techniques used in current continuous gravitational-wave searches.
 We then consider more complex audio (music and speech) for a more general interest in signal processing and explore the use of the interferometer as an optical microphone. 
@@
 We find that the logMMSE produces the best results, however, it does not enhance the recording to an intelligible level. 
-The complex audio signals can also serve as more distant analogues to un-modelled gravitational wave burst signals, e.g. from supernovae.
+The complex audio signals can also serve as more distant analogues to un-modelled gravitational-wave burst signals, e.g. from supernovae.
 
@@
 It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
-The increased excitement and public interest in the field of gravitational wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational wave research to a wider, non-specialist audience.
-The field of gravitational wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.
-% \jam{``charming'' is very sweet}
+The increased excitement and public interest in the field of gravitational-wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
+The field of gravitational-wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.
+% \jam{``charming'' is very sweet}     - agreed I wasn't sure at first but it has grown on me a lot! :) - Hannah
 %The demonstration presented here has great potential to be adapted for use in the undergraduate laboratory.

### 2020-11-19 paper/ifo-wanderingTone.tex
small consistency edits

@@
 In this context, ``significantly'' means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
-One method to search for a wandering signal is to split the time series data into several shorter segments which are analysed individually.%, and then recombine them under some continuity assumption to coherently analyse the whole set.
+One method to search for a wandering signal is to split the timeseries data into several shorter segments which are analysed individually.%, and then recombine them under some continuity assumption to coherently analyse the whole set.
 % I think the following is more accurate. The antenna beam pattern is used to compute a detection statistic. Which detection statistic to use depends on the type of target. -Hannah
 
-A ``detection statistic'' is used to quantify the likelihood of a signal being present in the data at each frequency and time series segment. 
+A ``detection statistic'' is used to quantify the likelihood of a signal being present in the data at each frequency and timeseries segment. 
 In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
@@
 The choice of detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
-In this work, we take the discrete Fourier transform of each time series segment and use its amplitude at each frequency as a detection statistic.
+In this work, we take the discrete Fourier transform of each timeseries segment and use its amplitude at each frequency as a detection statistic.
 This maximises the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}. 

### 2020-11-19 paper/ifo-conclusions.tex
ready for submision - I have commented out all author identification (required by the journal for double anonymous review process)

@@
 The demonstration presented has the potential to be adapted for use in the undergraduate laboratory.
-It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
+It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
 The increased excitement and public interest in the field of gravitational-wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.

### 2020-11-19 paper/paper_main.tex
ready for submision - I have commented out all author identification (required by the journal for double anonymous review process)

@@
 \title{Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
+
+\author{\textbf{[Authors removed for double-anonymous review.]}}
+\begin{comment}
 
@@
 
+\end{comment}
+
 %\author{\han{to do: check names - initials?}}
@@
 
-\section{Table-top gravitational wave science}
+\section{Table-top gravitational-wave science}
 \label{sec:ifo}
@@
 \begin{acknowledgments}
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky and Sam Cooper, for their helpful feedback and suggestions.
-This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
-Financial support towards hardware was provided by the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme. 
-Travel support was provided by the Australian National University PhB Science program.
-This work has been assigned LIGO document number P2000386.
+\textbf{[Removed for double-anonymous review.]} 
+%The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky and Sam Cooper, for their helpful feedback and suggestions.
+%This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
+%Financial support towards hardware was provided by the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme. 
+%Travel support was provided by the Australian National University PhB Science program.
+%This work has been assigned LIGO document number P2000386.
 

### 2021-01-25 paper/ifo-complexAudio.tex
starting to move material around for referee response

@@
 
+In this section, we explore how this demonstration can be used to teach a selection of signal processing techniques. 
+We use complex audio signals (such as music and speech) as a natural successor to the constant and wandering tones used in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} respectively.
+As complex audio signals are not quasi-monochromatic, the Viterbi algorithm as applied in Section~\ref{sec:viterbi_wandering} is not directly applicable here. 
+Instead we use a hierarchy of passive filters which suppress noise, yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
+
+Here we treat the Michelson interferometer as an optical microphone, where a laser beam is used to detect sound vibrations. 
+Optical microphone devices have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defence industry and operate on a variety of related principles.
+Our objective is to play a recording of speech or music through the speaker attached to mirror M1 (see Fig.XXXX), record the interference pattern, and then recover the original signal via a selection of signal processing techniques.
+
+This analysis is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals. 
+Instead, the apparatus serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories. 
+We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
+A selection of filters are applied to the data, details of which are presented in Supplementary Material and we present the results of the best performing filters in Section~\ref{sec:opticalMicResults}.
+\han{rephrase depending on what we show.}
+
+\subsection{Hardware modifications for the optical microphone}
+\label{sec:initialResultsOpMic}
+The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
+Speech intelligibility (being able to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$~\cite{speech_intelligibility}.
+Therefore, the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music (adjusting for the Nyquist frequency).
+This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$ and thus can only ``hear'' frequencies below $15\,{\rm Hz}$.
+To overcome this issue, we use a photodiode\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the reverse direction.
+As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.} at the output of the interferometer to achieve a sampling rate of $16\,{\rm kHz}$.
+
+We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage dependent on the incident intensity. 
+The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
+The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath the screen. 
+The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
+Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. Resources for using the Pi and photodiode are described in Appendix~\ref{app:circuit_diagram}.
+
+Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing\footnote{Folding of frequencies greater than half the sampling rate.} into the detected range.
+We include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
+This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
+We also place a second cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC.
+% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
+% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
+
+%An updated schematic and a photograph of the entire optical microphone are shown in the left and right panels of Fig.~\ref{fig:ifo_schematic_podo}, respectively.
+See Appendix~\ref{app:circuit_diagram} for a diagram and photograph of the circuit. 
+\han{[Hannah] I've removed the second schematic and photo for now - could be an option for the photo to go in the appendix / supplement if we want to show the physical set up?}
+
+\subsection{Anti-aliased output}
+\label{sec:initialResultsOpMic}
+
+%The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources). 
+We test the optical microphone with a variety of recordings, including speech from different people and music ranging from simple melodies and rhythms to songs.
+During recordings, care is taken to minimise activity around the demonstration to reduce environmental noise coupling into the interferometer.
+The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python~\cite{scipy,python}. 
+When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.
+
+The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass hum throughout. 
+This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), as shown in the top panel of Fig.~\ref{fig:psd_noise}. 
+The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ mains signal up to and beyond $8$th harmonic thereof. 
+The mains signal is also present (but far weaker) in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 
+Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself. 
+The appearance of harmonics of the mains noise is likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}). 
+The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
+
+\begin{figure}
+	\includegraphics[width=.49\textwidth]{figures/psd_butterworth_14_6.pdf}
+	\caption{\label{fig:psd_noise}
+\han{If we do not cover the butterworth filter in the main text, I'd suggest only showing the top panel of this figure in the main text.
+We can include the twp panel version in the butterworth section of the supplementary material}
+Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
+Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
+We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
+After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
+}
+\end{figure}
+
+\subsection{Optical microphone results}
+\label{sec:opticalMicResults}
+\han{[Hannah] keen to hear other's thoughts, but I think it would be cool to show the notch+wiener combined and the logmmse as the results?
+They could be placed in a single figure for ease of comparison with time \& spectrum columns and rows: 1.
+original signal, 2.
+raw output, 3.
+combined cascaded notch and Wiener, 4.
+logMMSE}
+
+We explore several filters to remove the $50\,{\rm Hz}$ mains hum and harmonics, and also to improve speech intelligibility of the recording. 
+The accompanying supplementary material describes a range of analysis techniques which can be used as examples for the undergraduate laboratory. 
+All filters are tested on the same $1\,{\rm s}$ long speech recording.
+
+In signal processing, the ideal filter would be one that:
+i) completely attenuates the undesired parts of the spectrum, 
+ii) does not change the rest of the spectrum, and 
+iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
+However, these three conditions cannot all hold at once and any filter must compromise between them. \han{[Notes] editor's comments suggest expanding on this?}
+Here we present results of two analysis.
+They are briefly described and the results shown (see Supplementary Material for details).
+
+Firstly, we consider two signal processing techniques used in combination: the cascaded notch and Wiener filter. 
+A notch filter remove signals within a specific frequency range \han{ref}. 
+In our case, we would like to remove the $50\,{\rm Hz}$ mains noise and harmonics. 
+Therefore, we use a cascaded notch filter where each notch is centred at each of the harmonics. 
+The Wiener filter is an advanced statistical technique which makes used of statistical information from the speech data and noise. 
+It amplifies parts of the signal with high SNR while suppressing pares with low SNR.
+
+\han{refer to figure.. } shows our results.
+We find that the combined cascaded notch and Wiener filter produces a better result than either alone. 
+Most of the mains noise is removed, however the recovered voice sounds muffled and is not understandable.
+
+Secondly, we apply a speech enhancement technique. 
+Ref.~\cite{SubjectiveComparison} compares $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech (see also Ref.~\cite{Ephraim1984SpeechEU_logMMSE} for details on the logMMSE estimator).
+We use an existing implementation of the logMMSE~\cite{logmmse}.
+
+\han{refer to figure} 
+After applying the logMMSE estimator we see significant attenuation of the mains harmonics and a general smoothing of the spectrum. 
+Most of the background noise is removed, however the logMMSE does not significantly enhance the speech as the voice sounds muffled and indistinct.
+We find some improvement with music over speech. 
+Simple chords and drums can be heard, however more composite sounds and complex melodies cannot be heard clearly. 
+Our observations suggest that this is especially true for certain instruments, in particular flutes and violins sometimes can’t be heard at all. 
+This could be a perceptual effect or a frequency dependence somewhere in the optical microphone.
+Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums are louder in the results.
+To address these problems, we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
+To determine this requires a better understanding of the system, as discussed in Section~\ref{sec:future_work}.
+
+\begin{comment} 
 In this section, we explore how this demonstration can be used to teach a selection of signal processing techniques.
@@
 \end{figure*}
-\end{comment}
 
@@
 
+\end{comment}
+
 \end{document}

### 2021-01-25 paper/ifo-supplementary-material.tex
starting to move material around for referee response

--- 2020-11-19 paper/ifo-complexAudio.tex
@@
-\documentclass[paper-main.tex]{subfiles}
+% GravExplainSuppMatt
+%\documentclass[prb,reprint,nofootinbib]{revtex4-1} 
+\documentclass[aps,pra,superscriptaddress,reprint,nofootinbib]{revtex4-1}
+% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
+
+% Note that AJP uses the same style as Phys. Rev. B (prb).
+% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
+% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
+% We look forward to receiving your submission to AJP.
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath,amssymb,amsthm}
+\usepackage{amsfonts}
+\usepackage{graphicx}
+\usepackage{float}
+\usepackage{mathtools}
+\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{hyperref}
+\usepackage{textcomp}
+\usepackage{subfiles}
+\usepackage{comment}
+\usepackage{algpseudocode}
+\usepackage{lmodern}
+
+\usepackage{silence}
+\WarningFilter{revtex4-1}{Repair the float}
+
+\newcommand{\jam}{\textcolor{magenta}}
+\newcommand{\han}{\textcolor{orange}}
 
@@
 
-In this section, we explore how this demonstration can be used to teach a selection of signal processing techniques.
-The apparatus used in Sections~\ref{sec:ifo} to~\ref{sec:viterbi_wandering} allows undergraduate students to explore a range of signal processing techniques and filters of interest and provides a foundation for many technical applications~\cite{DigitalProcingOfSpeechSignals:1978}.
-We provide a series of example activities using complex audio (such as music and speech) as a natural successor to constant and wandering tones. 
-As complex audio signals are not quasi-monochromatic, unlike continuous gravitational-wave signals, the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not directly applicable here.
-Instead, in this section, we use the interferometer as an optical microphone to detect and playback complex audio signals (i.e.\ using light to capture sound).
-In doing so, we experiment with a hierarchy of passive filters which suppress noise, yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
-A vast literature exists on signal processing with speech, see Ref.~\cite{DigitalProcingOfSpeechSignals:1978} for introductory material including Fourier analysis and digital representations.
-Optical microphone devices have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defence industry and operate on a variety of related principles.
+\title{Supplementary Material} 
+\maketitle
 
-This section is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals.
-Here, the optical microphone serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories.
-We describe the additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}. 
-A selection of filter and speech enhancement example activities using the optical microphone are outlined in Section~\ref{sec:filters}.
+\section{Optical microphone signal processing}
+\han{Notes - I've just made a start at moving material over. We will need to go through once we make final decisions on what to show in the main text / when new figures ready to check things}
 
-\subsection{Hardware modifications for the optical microphone}
-\label{sec:photodiode}
-
-\begin{figure*}
-	\includegraphics[width=0.44\textwidth]{figures/ifo_schematic_photodiode_edit.pdf}
-	\includegraphics[width=0.52\textwidth]{figures/setup_pic2.pdf}
-	\caption{
-Left: the updated Michelson interferometer schematic. 
-The schematic is identical to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
-Right: a photograph of the optical microphone. 
-The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. 
-In the main image, the photodiode is placed behind the cloth screen. 
-The inset shows a face-on view of the photodiode with the covering screen removed. 
-See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.
-}
-	\label{fig:ifo_schematic_podo}
-\end{figure*}
-
-The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$.
-Speech intelligibility (being able to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to (and beyond) $8\,{\rm kHz}$~\cite{speech_intelligibility}.
-Therefore, the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music (adjusting for the Nyquist frequency).
-This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$ and thus can only ``hear'' frequencies below $15\,{\rm Hz}$.
-To overcome this issue, we use a photodiode\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the reverse direction.
-As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.} at the output of the interferometer to achieve a sampling rate of $16\,{\rm kHz}$.
-
-We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage dependent on the incident intensity. 
-The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
-The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath the screen. 
-The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
-Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. Resources for using the Pi and photodiode are described in Appendix~\ref{app:circuit_diagram}.
-
-Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing\footnote{Folding of frequencies greater than half the sampling rate.} into the detected range.
-We include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
-This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
-We also place a second cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC.
-% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-
-An updated schematic and a photograph of the entire optical microphone are shown in the left and right panels of Fig.~\ref{fig:ifo_schematic_podo}, respectively.
-See Appendix~\ref{app:circuit_diagram} for a diagram and photograph of the circuit.
-
-\subsection{Anti-aliased output}
-\label{sec:initialResultsOpMic}
-
-%The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources). 
-We test the optical microphone with a variety of recordings, including speech from different people and music ranging from simple melodies and rhythms to songs.
-During recordings, care is taken to minimise activity around the demonstration to reduce the amount of environmental noise coupling into the interferometer.
-The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python~\cite{scipy,python}.
-When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.
-
-The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass hum throughout.
-This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), as shown in the top panel of Fig.~\ref{fig:psd_noise}.
-The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ mains signal up to and beyond $8$th harmonic thereof.
-The mains signal is also present (but far weaker) in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution.
-Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself.
-The appearance of harmonics of the mains noise is likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}).
-The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
-
-\begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/psd_butterworth_14_6.pdf}
-	\caption{\label{fig:psd_noise}
-Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
-Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
-We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
-After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
-}
-\end{figure}
-
-\subsection{A hierarchy of filters}
-\label{sec:filters}
-
-Here, we describe several filters used to remove the $50\,{\rm Hz}$ mains hum and its harmonics, but also to ultimately improve the speech intelligibility of the recording (to be able to understand the words spoken).
+In section V of the main article, we use the inerferoeter as an optical microphone. 
+We use signal processing techniques to recover audio signals o speech and music. 
+In this supplementary material we provide futher details of a selection of analysis applied to the data which may be considered for undergraduate laboratory examples.  
 We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
@@
 
-%\begin{figure*}
-%\begin{center}
-%% trim order is <left> <lower> <right> <upper>
-%\includegraphics[width=\textwidth,trim={1cm 8.5cm 1cm 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
-%\end{center}
-%\caption{\label{fig:noiseSubtract}
-%Frequency spectrum of a $1\,{\rm s}$ signal before the use of filters. 
-%The first panel shows the original frequency spectrum of the human voice recording (saying ``A cathode ...''). 
-%The second panel shows the spectrum recorded after the signal passes through the optical microphone (sampled at $16\,000\,{\rm Hz}$).
-%The third panel shows a noise-only recording from the optical microphone when no audio signal is being played. 
-%The fourth panel shows the result of subtracting the noise-only spectrum in the third panel from the recording in the second panel. 
-%We see no obvious improvement from this method. 
-%}
-%\end{figure*}
-
 Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results, respectively. 
 The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} show the source signal and the raw optical microphone recording, respectively. 
-The third row shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.
+The third row shows a background spectrum of the optical microphone. 
+The following sections describe each of the analysis in turn.
 
-%The Fourier spectrum of the source signal is shown in the first panel of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-The second panel shows the spectrum of the raw optical microphone recording of the signal.
-The third panel of Fig.~\ref{fig:noiseSubtract} shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.
+\subsection{Background noise subtraction}
 
@@
 The fourth row of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the spectrum obtained after subtracting the background noise spectrum. 
-We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified. %Why the noise sources are time-varying is a fact yet to be determined.
+We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified.
 
-Beyond this na{\"i}ve approach, we apply a range of filters to the data.
-In signals processing, the ideal filter would be one that:
-i) completely attenuates the undesired parts of the spectrum, 
-ii) does not change the rest of the spectrum, and 
-iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
-However, these three conditions cannot all hold at once and any filter must compromise between them, as we will see below. 
-The following sections describe a series of example analyses with the optical microphone.
-
-\subsubsection{Rectangular comb filter}
+\subsection{Rectangular comb filter}
 
@@
 
-%A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible. 
-% \jam{[re:comb filters.
-When I tried to comb filter it either was too sharp and so jumbled the whole signal in convolution.
-Or if smooth attenuated the signal away into noise.
-At least, that's what I remember.
-Changrong has a comb filter section anyway so I am happy dropping my unsuccessful attempts.]}
-
-\subsubsection{High-pass filter}
+\subsection{High-pass filter}
 
@@
 
-%We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass
-
-\subsubsection{Butterworth band-pass filter}
+\subsection{Butterworth band-pass filter}
 
@@
 \end{equation}
-% commenting to send to co-authors
-%\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
 and is combined with a similar high-pass filter to form the band-pass filter.
@@
 
-\subsubsection{Cascade notch filter}
+\begin{figure}
+	\includegraphics[width=.49\textwidth]{figures/psd_butterworth_14_6.pdf}
+	\caption{\label{fig:psd_noise}
+Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
+Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
+We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
+After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
+}
+\end{figure}
+
+\subsection{Cascade notch filter}
+
 \label{sec:notch}
@@
 
-% \begin{figure*}
-% \begin{center}
-% \includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/time SeriesNotchWiener.pdf}
-% \end{center}
-% \caption{\label{fig:timeNotchWiener}
-% Time series results for the notch and Wiener filters. 
-% The first panel shows the original speech recording which was played from the speaker. 
-% The second panel shows the data from the optical microphone before any filters are applied. 
-% The third and fourth panels show results after applying the notch and Wiener filter respectively. 
-% Finally, the fifth panel shows the result of applying both the notch and Wiener filter. 
-% }
-% \end{figure*}
-
-% \begin{figure*}
-% \begin{center}
-% \includegraphics[width=\textwidth,trim={0cm 8.2cm 0cm 8cm},clip]{figures/freqNotchWiener.pdf}
-% \end{center}
-% \caption{\label{fig:freqNotchWiener}
-% Frequency spectrum results for the notch and Wiener filters. 
-% The first and second panels show the results of applying the notch and Wiener filter respectively.
-% The third panel shows the spectrum after applying both the notch and Wiener filters. 
-% The first, second, and third panels show the spectrum of the time series show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
-% }
-% \end{figure*}
-
-%A notch filter attenuates signals in specific frequency bands. 
-%We apply a $15$ cascaded notched filter to remove the mains noise and its harmonics up to $750\,{\rm Hz}$. 
-%\han{should we include here a mathematical description of the $15$ cascade notch filter?}
-%\jam{[Yes, I have no idea what a cascaded notched filter is and I'm an author!]}
-
-\subsubsection{Wiener filter}
+\subsection{Wiener filter}
 \label{sec:Wiener}
@@
 where $\textbf{h}=(h(0),\dots h(n))$ represents the coefficients of an $n$th order Wiener filter.
-% What this sentence is trying to convey is not clear to me
-%% just trying to say that we can make the assumptions that we do because the Wiener filter appears to work well under them - J
-%These assumptions are to be justified by the efficacy of the Wiener filter.
-%Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
-%\begin{equation}
-%    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
-%\end{equation}
-%Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
-%A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
-%$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
-%$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
-%only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.}
-%and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
 The famous Wiener-Hopf equation~\citep{noble1959methods} allows us to determine $\textbf{h}$, as\begin{equation}
@@
 %\end{equation*}
-%where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
 
-%As an instructive aside, i
 If we let the Wiener filter be non-causal (i.e.
@@
 More detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.
-% That this filter is non-causal makes it unsuitable for real time signal processing
-
-%In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}, and set the length of the filter to $n=100$. 
-%A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
-%The choice of $n=100$ is a reasonable selection to balance smoothing and efficiency.
 
@@
 We see a significant improvement in the timeseries of the recovered signal, however, a strong noise hum persists, audibly.
-%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
-%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
-%\han{should we add more details here too?} \jam{[seconded]}
-%The time series and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
-%The time series is visibly cleaner, however there is a strong noise hum.
 
-\subsubsection{Combined notch and Wiener filter}
+\subsection{Combined notch and Wiener filter}
 
 Here, we experiment with applying a combination of the cascaded notch and the Wiener filter to the recorded speech signal. 
+The results of this analysis are described in the main paper in section \han{VC} and figure \han{XX}.
 
@@
 
-%The combination of the notch filter and the Wiener filter produces a cleaner signal than either individually. 
-%The time series and spectrum results are shown in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively.
+\subsection{logMMSE estimator}
+\label{sec:logmmse}
 
-%The ideal filter would be one that i) completely attenuates the undesired parts of the spectrum, ii) does not change the rest of the spectrum, and iii) has smooth edges so as to not damage the signal after convolution.
-%These three conditions are contradictory and any filter must compromise between them. 
-%The above simple filters fail one or more of i)--iii). 
-%The Butterworth filter performs the best as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
-%\han{see note from Andrew and ask James about this}
-
-%\han{Is it worth also including some plots to show how the different filters perform? E.g. a PSD with comb filter, high pass filter, Butterworth filter?}
-
-% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
-Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-
-% I just combined these into a single figure and also moved some of the intro text about the signal into fig 8 so that it is clear we are all talking about the same data in these plots - hannah
 \begin{figure*}
@@
 The change in scale from the top panel is not of concern as the ear hears the relative frequency content and normalisation of the timeseries fixes any scaling. 
-% my feeling is not to include this sentence here as it is hard to be certain what the noise and what is the input signal for this plot without further investigation - hannah
-%The bottom two panels also show a broad feature from around $750\,{\rm Hz}$ to $1200\,{\rm Hz}$, the origin of which is yet to be determined.
     }
@@
 
-\subsubsection{logMMSE estimator}
-\label{sec:logmmse}
-
-\begin{comment}
-\begin{figure*}
-	\includegraphics[width=0.8\textwidth]{figures/combined_timeseries_melatos_shift_source.pdf}
-	\caption{Time series of an adult male voice (saying ``A cathode ...'') showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
-The source signal is shifted by $0.12$ s to the right to synchronise the manual delay from starting to play the source through the speaker and starting the recording at the Raspberry Pi.
-One second of the one minute long recording is shown.
-The rise at the start of the bottom panel is expected when filtering a signal of finite duration.}
-	\label{fig:logMMSE_timeseries}
-\end{figure*}
-
-\begin{figure*}
-	\includegraphics[width=0.8\textwidth]{figures/combined_spectrum_melatos.pdf}
-	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
-The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.
-The change in scale from the top panel is not of concern as the ear hears the relative frequency content and normalisation of the time series fixes any scaling.
-The bottom two panels also show a broad feature from around $750\,{\rm Hz}$ to $1200\,{\rm Hz}$, the origin of which is yet to be determined.}
-	\label{fig:logMMSE_spectrum}
-\end{figure*}
-\end{comment}
-
-%probably a bit too causal language 
-%What we’re attempting here is not new, 
 Speech enhancement of noisy channels is a classic problem in signal processing. 
-% shouldn't we make the citations consistent throughout? We don't mention author names up to this point and typically journals have a set style guide. - Hannah 
-%To borrow from that field, Yi Hu and P.C.
-Loizon~\cite{SubjectiveComparison} compared $13$ speech enhancement methods and found the log minimum means-square error (logMMSE) estimator to be the best among them, qualitatively, at recovering speech.
 In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
@@
 We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}) to the recorded signal.
-%Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
-%This type of estimator is based on work Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
 
-%In this work we use an existing implementation~\cite{logmmse} of the logMMSE estimator. 
-%Applying this estimator to the optical microphone output produced dramatically cleaner results.
+\han{results for the logmmse estimator are shown in the main paper in section VC and figure X}
 
-The timeseries and frequency spectrum results of the logMMSE estimator are shown in Fig.~\ref{fig:logMMSE_timeseries_freqspectrum}. 
-The top two rows show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison (identical to the top two rows in Fig.~\ref{fig:BackgroundNotchWienerCombined}).
-The third row shows the timeseries and corresponding spectral amplitude after applying the logMMSE estimator. 
-We see significant attenuation of the mains harmonics as well as a general smoothing of the spectrum. 
-The logMMSE method indeed removes most of the background noise but does not significantly enhance the speech as the voice sounds muffled and indistinct.
-% suggest remove as it's just rephrasing what's already been said
-%, with the words unable to be understood.
-
-We find some improvement with music over speech.
-After applying the logMMSE estimator, simple chords and drums can be heard, however more composite sounds and complex melodies cannot be heard clearly.
-%Our observations suggest that this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins sometimes can’t be heard at all. 
-Our observations suggest that this is especially true for certain instruments, in particular flutes and violins sometimes can’t be heard at all. 
-This could be a perceptual effect or a frequency dependence somewhere in the optical microphone.
-%, although the exact cause is yet to be determined.  - repetition
-Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums are louder in the results.
-To address these problems, we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-%Or if instead they are present in the noisy data but we need better techniques to extract them.
-- should probably justify what you mean by better if you are going to say it here, but I think enough is discussed in section VI for it not to be necessary here - hannah
-To determine this requires a better understanding of the system, as discussed in Section~\ref{sec:future_work}.
-
-%Overall, we find that the intelligibility of speech remains low. 
-%The played-back voice sounds mumbled and indistinct. 
-%The estimator removes most of the background noise, but does not clarify the speech. 
-%To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-%This requires a better understanding of the system.
-% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
-However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
-An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
+\bibliographystyle{myunsrt}
+\bibliography{ifoDemoBib}
 

### 2021-01-25 paper/AJPRefereeResponse.txt
a place to keep track of the changes we make in answer to the reviewers

This is to keep track of our changes. We'll use it to write our cover letter to AJP.

###############################################################################
###############################################################################

Editor Comments

This is an addendum to my decision letter and the two reviews of your manuscript.  I am satisfied that the reviewers read your work carefully, and that their recommendations are fair, thoughtful, and potentially useful.  While I sympathize with R1’s concern that the manuscript is better suited to a research journal, I think it could be revised to make a valuable contribution to AJP, which, as you know, is geared toward pedagogy.  So as I read your manuscript, I recorded things that were not sufficiently clear to me, supposing that the same things might not be clear to many AJP readers who mightbe very interested in duplicating your apparatus.  Here are some of my concerns.

1.Simple things: Convert footnotes to endnotes, blending them in with the bibliography.  Also, use US English (color, not colour; defense, not defence, etc.). 
>

2.Section III.  How was the Fourier transform generated?  (software?)
>

3.Section IV.  The term “Markov process” may not be familiar to readers.  Figure 4: The three shades (or are there four?) indicating probability do not show up wellin color, and worse, will be indistinguishable in black/white.  I’m guessing that’s the reasonI didn’t understand why certain paths were grey and not black.  While I applaud your efforts to explain the Viterbi algorithm (VA), I still think there’s room for improvement.  I don’t understand how it improves the signal to noise ratio.4.Figure 5 is described as a heatmap.  Please explain, in the caption, what the color coding indicates.  Can you compare what is shown to what would be detected without using the VA?  Is this “heatmap” the same representation as is often shown in LIGO publications?
>

5.Section V.  What do you mean by an optical microphone device?  “Cloth screen” over the photodiode?  Fig. 6 is unnecessary; the differences from Fig. 1 are explained clearly in the text.  
> optical microphone point - I did some reshuffling so that the optical microphone concept is introduced in a separate paragraph and expanded on it a little to try and improve clarity. [Hannah]

6.Figure 7: why is there a -100 dB difference in the signal level from 7a to 7b?  
>

7.Much of your description of the various filters should be moved to Supplementary Materials and/or referenced.  This is a physics audience.  More interesting is your observation that all filters degrade the signal in some way, i.e., they are all compromises.  You could usefully expand that argument without becoming overly mathematical.  
>

8.Where in your flowchart are the various filters you describe?  Are the codes for them readily available?  How does one implement a filter?  
>

So you see that I’m asking for a major revision of your manuscript.  I hope this does not discourage you.  My goal is to expand the size of the audience reading your article, and to increase its utility for that group of educators.  I look forward to a revised manuscript.

###############################################################################
###############################################################################

R3 Comments

The paper describes a table top interferometer for demonstrating the basic operating principles of gravitational wave (GW) detectors. Audio signals are used to demonstrate gravitational wave data analysis techniques for the detection and estimation of signals in noisy data. The paper is well written and will be of interest to educators who wish to provide some hands-on training to students in GW astronomy. I have some comments below that the authors should address.

-- The data analysis problems discussed in this paper are related to continuous wave searches. However, I did not find a mention of the key feature involved in these searches, namely, Doppler modulation of the signal frequency caused by the rotation of the Earth. This is the principal reason behind the extremely high computational cost of CW searches. A discussion of doppler modulation should be included. (It would be neat if the setup could be extended in the future to show the Doppler effect.)

-- Is the HMM defined properly in Sec. IV A? The observables here are the amplitudes of the spectrogram pixels at a given time index and the hidden markov process is the underlying drift of frequency and amplitude of the signal. I was expecting to see a conditional pdf that connects the state of the underlying (suitably discretized) process to the observables. I don't see it.

-- Is Eq. 1 correct? The LHS is a marginal probability while the RHS is a joint probability. Shouldn't there be a marginalization on the RHS?

-- The normalization used for F(t_i, f_j) should be defined explicitly.

-- Looking at Fig. 3, it appears that the noise is not white and rises in power at lower frequencies. How is the color of the noise taken into account, if at all, in the HMM or the Viterbi algorithm? A discussion of this point should be included although it is fine to assume white noise as a first approximation.

-- The Viterbi algorithm used in this paper falls along the general lines of finding the optimal path through a chirplet graph (which includes the spectrogram) that has been explored extensively in the GW literature. More broadly, the detection and estimation of chirp signals, which leave a track in the time-frequency plane, has been a topic of many papers. For the benefit of the readers, the authors should provide an expanded literature review in Sec. IV A and state clearly if their approach is new and, if so, how it differs from the existing ones. A partial list of papers is given below:     
* E. Chassande-Mottin and A. Pai, Phys. Rev. D 73, 042003 (2006). 
* E. J. Candes, P. R. Charlton, and H. Helgason, Classical Quantum Gravity 25, 184020 (2008).   
* W. G. Anderson and R. Balasubramanian, Phys. Rev. D 60, 102001 
(1999).   
* P. Addesso, M. Longo, S. Marano, V. Matta, I. Pinto, and M. Principe, in 2015 3rd International Workshop on  Compressed Sensing Theory and its Applications to Radar, Sonar and Remote Sensing (CoSeRa) (IEEE, New York, 2015), p. 154.   
* E. Thrane et al., Phys. Rev. D 83, 083004 (2011).   
* E. Thrane and M. Coughlin, Phys. Rev. D 89, 063012 (2014).    
* S. D. Mohanty, Phys. Rev. D  D 96, 102008 (2017).   
* Margaret Millhouse, Neil J. Cornish, and Tyson Littenberg, Phys. Rev. D 97, 104057 (2018).
> [from Hannah] These papers are addressing how to detect an unknown chirp signal from a compact binary merger. I think it may be relevant in some cases, but will take a look at the papers. We should also emphasise that the analysis is not new and that we have referenced the papers for the techniques we use.

-- Sec VC: I feel that this section too long and is trying to fit too much of digital signal processing (DSP) in a limited space. The topic of digital filtering is not something that Physics undergraduates (or even graduates) are usually familiar with and requires a more gentle introduction than provided here. For example, the term "IIR" is introduced without explaining what the impulse response of a filter is (and there is no mention of FIR filters). Similarly, Eq. 4 looks like a Z-transform, and advanced topic in itself, but this is not mentioned and instead the term "complex frequency" is used that is likely to be unfamiliar to most readers. (Incidentally, there seems to be a notational inconsistency across the two sides of Eq.3: 'f' and 'omega'.) To improve the readability of the paper, the authors should consider shortening this section by focussing on only the filtering procedure that worked the best. References to DSP textbooks may be provided for the interested reader to acquire background material. 
> to do

-- While the authors have applied standard filtering methods in Sec. VC, the removal of high power line features from data in GW detectors is done using more sophisticated approaches. For example, cross-channel regression, linear predictive filtering, Kalman filtering etc. There are also some non-linear and non-parametric methods that have been explored in the GW literature. This point should be mentioned in the paper.

-- The axes label for frequency is not consistent across the figures: in some plots it is "f/Hz" and others "f(Hz)".

-- This is not a comment but just my curiosity: How strong is the effect of mechanical vibrations of the camera or photodiode mount on the observed signal?

###############################################################################
###############################################################################

R1 Comments

This is a very interesting paper written in a clear and informativelanguage. However, I don’t think that it suits the style of papers in AJP because it represents a novel idea of teaching and exploring gravitational waves with audio signals. After reading and inspection of the manuscript I came to the  conclusion  that it is a  research  paper  which should pass  through critical discussion  with researchers  from gravitational  wave  communitywho  should  confirm  that  the audio signals produced in the way proposed in the manuscript,canindeed mimicrealgravitational wave signalswithout confusionof students and teachers. Gravitational waves are transverse-traceless while the sound waves are purely longitudinal. In this respect there is no analogy between gravitational waves and audio signals whatsoever. At the same time,the principles of extraction of signal from noise andthe filtering techniqueworked out in the manuscriptare applicable in a wide range of physical disciplines andmay be useful for the gravitational wave communityas well. On the other hand, I cannot imagine how teachers of physics will explain the complexity of the Winer filter, the Markov chain, the Viterbi algorithm to high school students or even to undergraduatestudentsin university college. Signal  processing technique is based  on a  high  level  mathematics (statistics, probability theory, stochastic processes, etc.) and engineering. In my opinion, the paper is too complicated both technically and mathematicallyfor publishing in AJP. I am convinced that this is a very good and high-quality research  paper but  my  fair  opinion  is  thatthe authorsshould submit  it  to  Physical Review D or another peer-review physics journal.

> Thoughts on this [Hannah]: The use of audio as an analogy to gravitational waves is not novel to this work. Sound has been used in numerous situations to convey the concept of gravitational wave signals to general audiences [will add references]. It is of course true that sound is not the same as gravitational waves, however in this work we are focusing on the extraction of a signal from an interferometer to demonstrate a series of analysis techniques, some of which are related to gravitational wave analysis. [will think about this some more and very happy to hear other thoughts on how we respond here -Hannah]

### 2021-03-30 paper/AJPRefereeResponse.txt
working on reviewer and editor comments

@@
 While I sympathize with R1’s concern that the manuscript is better suited to a research journal, I think it could be revised to make a valuable contribution to AJP, which, as you know, is geared toward pedagogy.
-So as I read your manuscript, I recorded things that were not sufficiently clear to me, supposing that the same things might not be clear to many AJP readers who mightbe very interested in duplicating your apparatus.
+So as I read your manuscript, I recorded things that were not sufficiently clear to me, supposing that the same things might not be clear to many AJP readers who might be very interested in duplicating your apparatus.
 Here are some of my concerns.
@@
 The term “Markov process” may not be familiar to readers.
-Figure 4: The three shades (or are there four?) indicating probability do not show up wellin color, and worse, will be indistinguishable in black/white.
-I’m guessing that’s the reasonI didn’t understand why certain paths were grey and not black.
+Figure 4: The three shades (or are there four?) indicating probability do not show up well in color, and worse, will be indistinguishable in black/white.
+I’m guessing that’s the reason I didn’t understand why certain paths were grey and not black.
 While I applaud your efforts to explain the Viterbi algorithm (VA), I still think there’s room for improvement.
-I don’t understand how it improves the signal to noise ratio.4.Figure 5 is described as a heatmap.
+I don’t understand how it improves the signal to noise ratio.
+> [Hannah] on Markov process. added sentence on Markov processes: "In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal)."
+> to do - work on the image colours
+> [Hannah] The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals (rather than an improvement in SNR).
+The single constant tone of section III is recovered easily using a Fourier transform, however if the signal frequency changes randomly over time, this will reduce the Fourier transform's ability to identify the signal as it is spread across multiple frequency bins.
+[Added sentence at start of section IV, para 2 to explain that Fourier transform of whole data set is not well suited here]
+
+4.Figure 5 is described as a heatmap.
 Please explain, in the caption, what the color coding indicates.
@@
 5.Section V.  What do you mean by an optical microphone device?  “Cloth screen” over the photodiode?  Fig. 6 is unnecessary; the differences from Fig. 1 are explained clearly in the text.  
-> optical microphone point - I did some reshuffling so that the optical microphone concept is introduced in a separate paragraph and expanded on it a little to try and improve clarity. [Hannah]
+> [from Hannah] optical microphone point - I did some reshuffling so that the optical microphone concept is introduced in a separate paragraph and expanded on it a little to try and improve clarity. 
+>
 
@@
 -- Is Eq. 1 correct? The LHS is a marginal probability while the RHS is a joint probability. Shouldn't there be a marginalization on the RHS?
+> to do check with Andrew about this, I'm not sure I follow what they are suggesting [Hannah]
 
 -- The normalization used for F(t_i, f_j) should be defined explicitly. 
+> to do check with James
 
@@
 References to DSP textbooks may be provided for the interested reader to acquire background material.
-> to do
+>
 
@@
 
-This is a very interesting paper written in a clear and informativelanguage.
+This is a very interesting paper written in a clear and informative language.
 However, I don’t think that it suits the style of papers in AJP because it represents a novel idea of teaching and exploring gravitational waves with audio signals.
-After reading and inspection of the manuscript I came to the  conclusion  that it is a  research  paper  which should pass  through critical discussion  with researchers  from gravitational  wave  communitywho  should  confirm  that  the audio signals produced in the way proposed in the manuscript,canindeed mimicrealgravitational wave signalswithout confusionof students and teachers.
+After reading and inspection of the manuscript I came to the  conclusion  that it is a  research  paper  which should pass  through critical discussion  with researchers  from gravitational  wave  community who  should  confirm  that  the audio signals produced in the way proposed in the manuscript,can indeed mimic real gravitational wave signalswithout confusionof students and teachers.
 Gravitational waves are transverse-traceless while the sound waves are purely longitudinal.

### 2021-03-30 paper/ifo-wanderingTone.tex
working on reviewer and editor comments

@@
 Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational-waves may not be monochromatic.
-A continuous-wave signal may wander slowly in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
+A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
 The audio analogue is a tone that wanders in frequency: a note that changes in pitch.
 
-% the data analysis techniques we draw from 
+A Fourier transform applied to the whole dataset (as in Section~\ref{todo}) is not well suited to the wandering frequency case, as the signal can be spread across multiple frequency bins making it challenging or perhaps impossible to detect.
+%A different means of detection is needed for a signal which has a randomly wandering frequency. 
 The analysis techniques used in this section are inspired by those used to search for continuous waves from spinning neutron stars in LIGO and Virgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
@@
 Following the method of Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, we use a hidden Markov model to recover the wandering signal from noisy data. 
+In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal). 
 In a hidden Markov model, the frequency state of the signal is unknown, or hidden, and can undergo transitions at discrete times. 
@@
 
-The spectrogram schematic, as shown in Fig.~\ref{fig:viterbi}, has $N_t$ time and $N_f$ frequency bins. 
+We use the schematic in Fig.~\ref{fig:viterbi} to aid our explanation of the method. 
+It represents a spectrogram with $N_t$ time bins and $N_f$ frequency bins. 
 We label the elements of the time and frequency  grid as $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-The normalised Fourier amplitude $F(t_i,f_j)$ is the detection statistic.
+We use the normalised Fourier amplitude $F(t_i,f_j)$ as the detection statistic, which is represented by the shading of each grid point (lighter corresponds to higher value).
+
+%The spectrogram schematic, as shown in Fig.~\ref{fig:viterbi}, has $N_t$ time and $N_f$ frequency bins. 
+%We label the elements of the time and frequency  grid as $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
+%The normalised Fourier amplitude $F(t_i,f_j)$ is the detection statistic.
 
@@
 Here, we allow the frequency of the signal to either (a) stay in the same bin; (b) move up by a single frequency bin; or (c) move down by a single frequency bin at each transition. 
+The possible transitions are shown as lines in Fig.~\ref{fig:viterbi}.
 We assign these three transitions equal probability, i.e.\ $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
@@
 \end{eqnarray}
+for any path through the grid. 
 The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes ${\rm Pr}[f(t_{N_t}) = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
@@
 At every $t_i$ all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}). 
-Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}, where the circles indicate the elements in the spectrogram with the lighter colours corresponding to a higher likelihood.
+Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}.%, where the circles indicate the elements in the spectrogram with the lighter colours corresponding to a higher likelihood.
 The implementation used in this work is available online (see Appendix~\ref{app:code}).

### 2021-03-30 paper/ifo-appendix.tex
ready for submision - I have commented out all author identification (required by the journal for double anonymous review process)
moving footnotes to endnote, starting on changing US spelling

@@
 The current build and sample data can be found at:
-\url{https://github.com/daccordeon/gravexplain}
+\textbf{[Removed github web link for double-anonymous review.]}
+%\url{https://github.com/daccordeon/gravexplain}
 % \begin{verbatim}
@@
 More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}.
-We use Fourier amplitudes normalised between $(0, 1)$ as multiplicative weights, but to avoid numerical underflow we take the logarithm and consider them as additive weights.
+We use Fourier amplitudes normalized between $(0, 1)$ as multiplicative weights, but to avoid numerical underflow we take the logarithm and consider them as additive weights.
 

### 2021-03-30 paper/ifo-complexAudio.tex
moving footnotes to endnote, starting on changing US spelling

@@
 Here we treat the Michelson interferometer as an optical microphone, where a laser beam is used to detect sound vibrations. 
-Optical microphone devices have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defence industry and operate on a variety of related principles.
+Optical microphone devices have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles.
 Our objective is to play a recording of speech or music through the speaker attached to mirror M1 (see Fig.XXXX), record the interference pattern, and then recover the original signal via a selection of signal processing techniques.
@@
 We test the optical microphone with a variety of recordings, including speech from different people and music ranging from simple melodies and rhythms to songs.
-During recordings, care is taken to minimise activity around the demonstration to reduce environmental noise coupling into the interferometer.
+During recordings, care is taken to minimize activity around the demonstration to reduce environmental noise coupling into the interferometer.
 The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python~\cite{scipy,python}. 

### 2021-03-30 paper/ifo-wanderingTone.tex
moving footnotes to endnote, starting on changing US spelling

@@
 In this work, we take the discrete Fourier transform of each timeseries segment and use its amplitude at each frequency as a detection statistic.
-This maximises the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}. 
-The detection statistic is normalised for convenience and we use it to form a grid in time and frequency (a spectrogram) which displays the frequency content of the signal over time. 
+This maximizes the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}. 
+The detection statistic is normalized for convenience and we use it to form a grid in time and frequency (a spectrogram) which displays the frequency content of the signal over time. 
 % moved one sentence up 
-%We also normalise the detection statistics for convenience.
+%We also normalize the detection statistics for convenience.
 
@@
 We label the elements of the time and frequency  grid as $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-We use the normalised Fourier amplitude $F(t_i,f_j)$ as the detection statistic, which is represented by the shading of each grid point (lighter corresponds to higher value).
+We use the normalized Fourier amplitude $F(t_i,f_j)$ as the detection statistic, which is represented by the shading of each grid point (lighter corresponds to higher value).
 
@@
 %We label the elements of the time and frequency  grid as $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-%The normalised Fourier amplitude $F(t_i,f_j)$ is the detection statistic.
+%The normalized Fourier amplitude $F(t_i,f_j)$ is the detection statistic.
 
@@
 
-We define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalised) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
+We define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
 The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
@@
 At each $f_j$ state, we select the path with the highest $A[f(t_0),f(t_1)] {\rm Pr}[f(t_0)]$ value as the most probable path.
-These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the grey lines show the rejected paths.
+These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the gray lines show the rejected paths.
 For example, the most probable connection to the element labelled (a) is the one directly behind it (i.e., $f_4$).
@@
 
-\item Moving to time $t_2$, again, we select the path which maximises the recursive quantity in Eqn.~\ref{eqn:recursiveViterbi} for each $f_j$.
+\item Moving to time $t_2$, again, we select the path which maximizes the recursive quantity in Eqn.~\ref{eqn:recursiveViterbi} for each $f_j$.
 These are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.
@@
 Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
-The colour of each circle indicates the likelihood that a signal is present in the box, where a lighter colour corresponds to a higher likelihood. 
+The color of each circle indicates the likelihood that a signal is present in the box, where a lighter color corresponds to a higher likelihood. 
 At each time, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines in the diagram). 

### 2021-03-30 paper/AJPRefereeResponse.txt
moving footnotes to endnote, starting on changing US spelling
some more notes on comments

@@
 1.Simple things: Convert footnotes to endnotes, blending them in with the bibliography.  Also, use US English (color, not colour; defense, not defence, etc.). 
->
+> Footnotes in bibliography -> Done
+> Done at least some of the US spellings [please correct as you spot them]
 
 2.Section III.  How was the Fourier transform generated?  (software?)
->
+> yes - check with James
 
@@
 Is this “heatmap” the same representation as is often shown in LIGO publications?
->
+> I think this is the Fourier amplitude normalised 0,1 - James is that right?
 
@@
 > [from Hannah] optical microphone point - I did some reshuffling so that the optical microphone concept is introduced in a separate paragraph and expanded on it a little to try and improve clarity. 
->
+> Removed fig 6 - do we want the photo in the supplementary material? 
+> To do: cloth screen.
 
 6.Figure 7: why is there a -100 dB difference in the signal level from 7a to 7b?  
->
+> Check with James and Changrong
 
@@
 You could usefully expand that argument without becoming overly mathematical.
->
+> Agreed, we have moved the majority of this section to Supplementary Material
+> to do - expansion on compromises
 
 8.Where in your flowchart are the various filters you describe?  Are the codes for them readily available?  How does one implement a filter?  
->
+> Check with James / Changrong - add Changrong's scripts to git
 
@@
 -- The normalization used for F(t_i, f_j) should be defined explicitly. 
-> to do check with James
+> to do check with James what was used
 

### 2021-04-01 paper/response/AJPRefereeResponse.tex
editing response, moved to directory

--- 2021-03-30 paper/AJPRefereeResponse.txt
@@
-This is to keep track of our changes. We'll use it to write our cover letter to AJP.
+\documentclass{article}
 
-###############################################################################
-###############################################################################
+\usepackage{fullpage}
+\usepackage{xcolor}
+\newcommand{\todo}{\textcolor{blue}}
+\newcommand{\han}{\textcolor{orange}}
 
-Editor Comments
+\begin{document}
+
+\todo{Opening comments}
+
+\subsection*{Editors comments (Joseph Amato)}
 
@@
 Here are some of my concerns.
+\\
 
-1.Simple things: Convert footnotes to endnotes, blending them in with the bibliography.  Also, use US English (color, not colour; defense, not defence, etc.). 
->
+\noindent 1. Simple things: Convert footnotes to endnotes, blending them in with the bibliography.  Also, use US English (color, not colour; defense, not defence, etc.). 
+\begin{itemize}
+\item Done: we have moved the footnotes to the bibliography
+\item \todo{Work in progress, please correct any that have not been fixed yet}
+\end{itemize}
 
-2.Section III.  How was the Fourier transform generated?  (software?)
->
+\noindent 2. Section III.  How was the Fourier transform generated?  (software?)
+\begin{itemize}
+\item  \todo{...}
+\end{itemize}
 
-3.Section IV.
+\noindent 3.Section IV.
 The term “Markov process” may not be familiar to readers.
@@
 I don’t understand how it improves the signal to noise ratio.
-> [Hannah] on Markov process. added sentence on Markov processes: "In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal)."
-> to do - work on the image colours
-> [Hannah] The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals (rather than an improvement in SNR).
+\begin{itemize}
+\item We have added a sentence to explain Markov process: ``In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal).''
+\item \todo{image colours}
+\item \han{[Hannah] The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals (rather than an improvement in SNR).
 The single constant tone of section III is recovered easily using a Fourier transform, however if the signal frequency changes randomly over time, this will reduce the Fourier transform's ability to identify the signal as it is spread across multiple frequency bins.
-[Added sentence at start of section IV, para 2 to explain that Fourier transform of whole data set is not well suited here]
+[Added sentence at start of section IV, para 2 to explain that Fourier transform of whole data set is not well suited here]}
+\end{itemize}
 
-4.Figure 5 is described as a heatmap.
+\noindent 4.
+Figure 5 is described as a heatmap.
 Please explain, in the caption, what the color coding indicates.
@@
 Is this “heatmap” the same representation as is often shown in LIGO publications?
->
+\begin{itemize}
+\item  \todo{...}
+\end{itemize}
 
-5.Section V.  What do you mean by an optical microphone device?  “Cloth screen” over the photodiode?  Fig. 6 is unnecessary; the differences from Fig. 1 are explained clearly in the text.  
-> [from Hannah] optical microphone point - I did some reshuffling so that the optical microphone concept is introduced in a separate paragraph and expanded on it a little to try and improve clarity. 
->
+\noindent 5. Section V.  What do you mean by an optical microphone device?  “Cloth screen” over the photodiode?  Fig. 6 is unnecessary; the differences from Fig. 1 are explained clearly in the text.
+\begin{itemize}
+\item Optical microphone: we have rearranged the text so that the optical microphone concept is introduced in a separate paragraph and expanded our discission of it to improve clarity.
+(Section V, paragraph 2)
+\item Fig. 6: agreed, we have removed this figure. \todo{decide if we want the photo in supplementary material}
+\item \todo{clarification of cloth screen }
+\end{itemize}
 
-6.Figure 7: why is there a -100 dB difference in the signal level from 7a to 7b?  
->
+\noindent 6. Figure 7: why is there a -100 dB difference in the signal level from 7a to 7b?  
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
-7.Much of your description of the various filters should be moved to Supplementary Materials and/or referenced.
+\noindent 7.
+Much of your description of the various filters should be moved to Supplementary Materials and/or referenced.
 This is a physics audience.
@@
 You could usefully expand that argument without becoming overly mathematical.
->
+\begin{itemize}
+\item Agreed, we have moved the majority of this section to Supplementary Matrial
+\item \todo{decide what we will show in main text}
+\item \todo{expansion on compromises discussion}
+\end{itemize}
 
-8.Where in your flowchart are the various filters you describe?  Are the codes for them readily available?  How does one implement a filter?  
->
+\noindent 8.Where in your flowchart are the various filters you describe?  Are the codes for them readily available?  How does one implement a filter?  
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
-So you see that I’m asking for a major revision of your manuscript.
+\noindent So you see that I’m asking for a major revision of your manuscript.
 I hope this does not discourage you.
@@
 
-###############################################################################
-###############################################################################
-
-R3 Comments
+\subsection*{Comments from R3}
 
@@
 The paper is well written and will be of interest to educators who wish to provide some hands-on training to students in GW astronomy.
-I have some comments below that the authors should address.
+I have some comments below that the authors should address.\\
 
+\noindent 
 -- The data analysis problems discussed in this paper are related to continuous wave searches.
@@
 (It would be neat if the setup could be extended in the future to show the Doppler effect.)
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
+\noindent
 -- Is the HMM defined properly in Sec.
@@
 I don't see it.
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
+\noindent
 -- Is Eq. 1 correct? The LHS is a marginal probability while the RHS is a joint probability. Shouldn't there be a marginalization on the RHS?
-> to do check with Andrew about this, I'm not sure I follow what they are suggesting [Hannah]
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
--- The normalization used for F(t_i, f_j) should be defined explicitly. 
-> to do check with James
+\noindent
+-- The normalization used for $F(t_i, f_j)$ should be defined explicitly. 
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
+\noindent
 -- Looking at Fig.
@@
 A discussion of this point should be included although it is fine to assume white noise as a first approximation.
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
+\noindent
 -- The Viterbi algorithm used in this paper falls along the general lines of finding the optimal path through a chirplet graph (which includes the spectrogram) that has been explored extensively in the GW literature.
@@
 IV A and state clearly if their approach is new and, if so, how it differs from the existing ones.
-A partial list of papers is given below:
-* E. Chassande-Mottin and A. Pai, Phys. Rev. D 73, 042003 (2006). 
-* E. J. Candes, P. R. Charlton, and H. Helgason, Classical Quantum Gravity 25, 184020 (2008).   
-* W. G. Anderson and R. Balasubramanian, Phys. Rev. D 60, 102001 
-(1999).   
+A partial list of papers is given below: \\
+* E. Chassande-Mottin and A. Pai, Phys. Rev. D 73, 042003 (2006). \\
+* E. J. Candes, P. R. Charlton, and H. Helgason, Classical Quantum Gravity 25, 184020 (2008).   \\
+* W. G. Anderson and R. Balasubramanian, Phys. Rev. D 60, 102001 (1999).   \\
 * P.
@@
 154.
-* E. Thrane et al., Phys. Rev. D 83, 083004 (2011).   
-* E. Thrane and M. Coughlin, Phys. Rev. D 89, 063012 (2014).    
-* S. D. Mohanty, Phys. Rev. D  D 96, 102008 (2017).   
-* Margaret Millhouse, Neil J. Cornish, and Tyson Littenberg, Phys. Rev. D 97, 104057 (2018).
-> [from Hannah] These papers are addressing how to detect an unknown chirp signal from a compact binary merger.
+\\
+* E. Thrane et al., Phys. Rev. D 83, 083004 (2011).   \\
+* E. Thrane and M. Coughlin, Phys. Rev. D 89, 063012 (2014).    \\
+* S. D. Mohanty, Phys. Rev. D  D 96, 102008 (2017).   \\
+* Margaret Millhouse, Neil J. Cornish, and Tyson Littenberg, Phys. Rev. D 97, 104057 (2018).\\
+\begin{itemize}
+\item \todo{...}
+\item \han{[from Hannah] These papers are addressing how to detect an unknown chirp signal from a compact binary merger.
 I think it may be relevant in some cases, but will take a look at the papers.
 We should also emphasise that the analysis is not new and that we have referenced the papers for the techniques we use.
+}
+\end{itemize}
 
+\noindent
 -- Sec VC: I feel that this section too long and is trying to fit too much of digital signal processing (DSP) in a limited space.
@@
 References to DSP textbooks may be provided for the interested reader to acquire background material.
->
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
+\noindent
 -- While the authors have applied standard filtering methods in Sec.
@@
 This point should be mentioned in the paper.
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
+\noindent
 -- The axes label for frequency is not consistent across the figures: in some plots it is "f/Hz" and others "f(Hz)".
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
+\noindent
 -- This is not a comment but just my curiosity: How strong is the effect of mechanical vibrations of the camera or photodiode mount on the observed signal?
 
-###############################################################################
-###############################################################################
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
-R1 Comments
+\subsection{Comments from R1 }
 
@@
 I am convinced that this is a very good and high-quality research  paper but  my  fair  opinion  is  thatthe authorsshould submit  it  to  Physical Review D or another peer-review physics journal.
+\begin{itemize}
+\item \todo{...}
+\end{itemize}
 
-> Thoughts on this [Hannah]: The use of audio as an analogy to gravitational waves is not novel to this work.
+\han{> Thoughts on this [Hannah]: The use of audio as an analogy to gravitational waves is not novel to this work.
 Sound has been used in numerous situations to convey the concept of gravitational wave signals to general audiences [will add references].
 It is of course true that sound is not the same as gravitational waves, however in this work we are focusing on the extraction of a signal from an interferometer to demonstrate a series of analysis techniques, some of which are related to gravitational wave analysis.
-[will think about this some more and very happy to hear other thoughts on how we respond here -Hannah]
+[will think about this some more and very happy to hear other thoughts on how we respond here -Hannah]}
+
+\end{document}

### 2021-04-01 paper/ifo-wanderingTone.tex

@@
 In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over at each time. 
-This information is encoded in the transition probability matrix $A(f_k,f_m)$, which describes the probability of system transitioning from state $f_i$ at $t_i$ to a state $f_{j}$ at $t_{i+1}$. 
+This information is encoded in the transition probability matrix $A(f_k,f_m)$, which describes the probability of system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
 Here, we allow the frequency of the signal to either (a) stay in the same bin; (b) move up by a single frequency bin; or (c) move down by a single frequency bin at each transition. 
@@
 We define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
-The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
+A specific path (which may not be optimal) is written as $f(t_0), f(t_1),..., f(t_n)$. 
+The probability of this specific path given the data is 
 \begin{eqnarray}
-{\rm Pr}[f(t_{i+1})=f_j] =~& F[t_{i+1},f(t_{i+1})] \nonumber \\
-                     &\times A[f(t_{i+1}),f(t_i)]  \nonumber \\
-                     &\times {\rm Pr}[f(t_i)].
-\label{eqn:recursiveViterbi}
+{\rm Pr}[ f(t_0), f(t_1),... f(t_n) | {\rm data}] =
+          F[t_n,f(t_n)] A[f(t_n),f(t_n-1)]
+          \times... A[f(t_1),f(t_0)] F[t_0,f(t_0)]\,,
+\label{eqn:pathprob}
 \end{eqnarray}
-for any path through the grid. 
-The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes ${\rm Pr}[f(t_{N_t}) = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
-We can then maximize the latter quantity over $0 \leq j \leq N_f$ to find the optimal Viterbi path overall, i.e. terminating in any frequency bin.
+The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises $Pr[ f(t_0), f(t_1),... f(t_n) | data]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
+We note that there is no marginalization used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:probpath} are both evaluated for a specific path and then we mazimise over all such paths to find the optimal Viterbi path.
+%The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
+%\begin{eqnarray}
+%{\rm Pr}[f(t_{i+1})=f_j] =~& F[t_{i+1},f(t_{i+1})] \nonumber \\
+%                     &\times A[f(t_{i+1}),f(t_i)]  \nonumber \\
+%                     &\times {\rm Pr}[f(t_i)].
+%\label{eqn:recursiveViterbi}
+%\end{eqnarray}
+%for any path through the grid. 
+%The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes ${\rm Pr}[f(t_{N_t}) = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
+%We can then maximize the latter quantity over $0 \leq j \leq N_f$ to find the optimal Viterbi path overall, i.e. terminating in any frequency bin.
 

### 2021-04-07 paper/ifo-appendix.tex
the low-hanging fruit of edits

@@
 Part of the transfer function for the optical microphone relates the change in intensity, $\delta I$, at a point $P$ in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
-Let $\theta$ be the angle between the optical axis and the ray from the centre of the relevant mirror to the point $P$. 
+Let $\theta$ be the angle between the optical axis and the ray from the center of the relevant mirror to the point $P$. 
 %Let the point be at angle $\theta$ on the screen (in the far-field) above the central ray as measured from the mirrors. 
@@
 where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal respectively. 
-The log likelihood assuming Gaussian noise is given by
+The log-likelihood assuming Gaussian noise is given by
 \begin{eqnarray}
@@
 %\subsection{Pseudocode}
-Here we give pseudocode of the implementation used in Section~\ref{sec:viterbi}.
+Here we give some pseudocode of the implementation used in Section~\ref{sec:viterbi}.
 More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}.
@@
 It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. 
-A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}.
+A photo of the breadboard and Raspberry Pi pins is shown in the Supplementary Material.
 The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an ADC to a Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
@@
 The op-amp output is passed through a Sallen-Key anti-aliasing filter tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
-The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.
+The digitized signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.
 This diagram is produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online} and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-}
-\end{figure*}
-
-\begin{figure*}%[H]
-	\includegraphics[width=0.5\textwidth]{figures/circuit_pic3.pdf}
-	\caption{\label{fig:circuit_pic3}
-Photodiode circuit assembled on breadboard. The leads from the photodiode enter from the left of the picture. The Raspberry Pi is shown at the top of the picture.
 }

### 2021-04-07 paper/ifo-complexAudio.tex
the low-hanging fruit of edits

@@
 As complex audio signals are not quasi-monochromatic, the Viterbi algorithm as applied in Section~\ref{sec:viterbi_wandering} is not directly applicable here. 
-Instead we use a hierarchy of passive filters which suppress noise, yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
+Instead, we use a hierarchy of passive filters which suppress noise, yet do not assume any specific form of the signal, unlike the Fourier-based maximum likelihood filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
 
-Here we treat the Michelson interferometer as an optical microphone, where a laser beam is used to detect sound vibrations. 
-Optical microphone devices have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles.
-Our objective is to play a recording of speech or music through the speaker attached to mirror M1 (see Fig.XXXX), record the interference pattern, and then recover the original signal via a selection of signal processing techniques.
+Here we use the Michelson interferometer as an ``optical microphone'', where a laser is used to detect sound instead of the electrical components in a conventional microphone.
+Optical microphones have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles.
+We use the term ``optical microphone'' to distinguish this technology from those previous technologies.
+Our objective is to play a recording of speech or music through the speaker attached to mirror M2 (see Fig.~\ref{fig:ifo_schematic_webcam}), record the resulting interference pattern, and then recover the original signal via a selection of signal processing techniques.
 
@@
 \subsection{Hardware modifications for the optical microphone}
-\label{sec:initialResultsOpMic}
+\label{sec:photodiode}
 The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
@@
 
-We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage dependent on the incident intensity. 
-The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
-The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath the screen. 
+We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage that depends on the incident intensity. 
+The photodiode records the interference pattern at roughly the same off-center position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
+The photodiode is mounted on a piece of cloth stretched over a plastic frame, with the electrical leads connected underneath the cloth.
+This cloth screen mount was the ``grill cloth'' from the front of a dismantled commercial speaker, re-purposed. 
 The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
@@
 This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
-We also place a second cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC.
+We also place a second cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC -- an improvised, physical solution that could instead be replaced by scaling down the voltage electronically.
+This cloth screen was also re-purposed grill cloth from a commercial speaker.
 % It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
@@
 %An updated schematic and a photograph of the entire optical microphone are shown in the left and right panels of Fig.~\ref{fig:ifo_schematic_podo}, respectively.
-See Appendix~\ref{app:circuit_diagram} for a diagram and photograph of the circuit. 
+See Appendix~\ref{app:circuit_diagram} for a diagram of the circuit. 
 \han{[Hannah] I've removed the second schematic and photo for now - could be an option for the photo to go in the appendix / supplement if we want to show the physical set up?}
@@
 %The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources). 
-We test the optical microphone with a variety of recordings, including speech from different people and music ranging from simple melodies and rhythms to songs.
+We test the optical microphone with a variety of recordings, including the speech of different people and music ranging from simple melodies and rhythms to songs.
 During recordings, care is taken to minimize activity around the demonstration to reduce environmental noise coupling into the interferometer.
@@
 
-We explore several filters to remove the $50\,{\rm Hz}$ mains hum and harmonics, and also to improve speech intelligibility of the recording. 
-The accompanying supplementary material describes a range of analysis techniques which can be used as examples for the undergraduate laboratory. 
+We explore several filters to remove the $50\,{\rm Hz}$ mains hum and harmonics, and also to improve the speech intelligibility of the recording. 
+The accompanying supplementary material describes a range of analysis techniques that can be used as examples for the undergraduate laboratory. 
 All filters are tested on the same $1\,{\rm s}$ long speech recording.
@@
 In signal processing, the ideal filter would be one that:
-i) completely attenuates the undesired parts of the spectrum, 
-ii) does not change the rest of the spectrum, and 
-iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
-However, these three conditions cannot all hold at once and any filter must compromise between them. \han{[Notes] editor's comments suggest expanding on this?}
-Here we present results of two analysis.
+(i) completely attenuates the undesired parts of the spectrum, 
+(ii) does not change the rest of the spectrum, and 
+(iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
+However, these three conditions cannot all hold at once.
+For example, if conditions (i) and (ii) hold, then the filter must be discontinuous at the boundary of the undesired region but this implies that the filter has ``infinite latency'' and so will affect (or damage) the time domain signal for infinite time~\cite{10.5555/151045}.
+Therefore, any filter must compromise between these three conditions.
+For speech intelligibility, this means that either: (i) some noise remains in the filtered recording, (ii) some of the speech content is lost as certain important frequencies are attenuated, or (iii) the speech is somewhat distorted in time.
+All three of these cases can, in extreme cases, worsen speech intelligibility beyond the unfiltered recording and so we choose filters that compromise between achieving the three conditions.
+\han{[Notes] editor's comments suggest expanding on this?}
+% https://en.wikipedia.org/wiki/Filter_(signal_processing)
+
+Here we present the results of two analyses using advanced signal processing techniques.
 They are briefly described and the results shown (see Supplementary Material for details).
 
-Firstly, we consider two signal processing techniques used in combination: the cascaded notch and Wiener filter. 
-A notch filter remove signals within a specific frequency range \han{ref}. 
+Firstly, we consider two signal processing techniques used in combination: the cascaded notch and the Wiener filter. 
+A notch filter removes signals within a specific frequency range \han{ref}. 
 In our case, we would like to remove the $50\,{\rm Hz}$ mains noise and harmonics. 
-Therefore, we use a cascaded notch filter where each notch is centred at each of the harmonics. 
-The Wiener filter is an advanced statistical technique which makes used of statistical information from the speech data and noise. 
+Therefore, we use a cascaded notch filter where each notch is centered at each of the harmonics. 
+The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. 
 It amplifies parts of the signal with high SNR while suppressing pares with low SNR.
@@
 We find that the combined cascaded notch and Wiener filter produces a better result than either alone. 
-Most of the mains noise is removed, however the recovered voice sounds muffled and is not understandable.
+Most of the mains noise is removed, however, the recovered voice sounds muffled and is not understandable.
 
@@
 \han{refer to figure} 
-After applying the logMMSE estimator we see significant attenuation of the mains harmonics and a general smoothing of the spectrum. 
-Most of the background noise is removed, however the logMMSE does not significantly enhance the speech as the voice sounds muffled and indistinct.
+After applying the logMMSE estimator we see significant attenuation of the mains harmonics and general smoothing of the spectrum. 
+Most of the background noise is removed, however, the logMMSE does not significantly enhance the speech as the voice sounds muffled and indistinct.
 We find some improvement with music over speech. 

### 2021-04-07 paper/ifo-conclusions.tex
the low-hanging fruit of edits

@@
 
-In this paper, we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector, demonstrating a range of signal processing techniques used within the gravitational-wave community and more broadly.
+In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating a range of signal processing techniques used within the gravitational-wave community and more broadly.
 Gravitational-wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
-We consider quasi-monochromatic audio signals as an analogue to continuous-wave signals to directly demonstrate techniques used in current continuous gravitational-wave searches.
+We consider quasi-monochromatic audio signals as an analog to continuous-wave signals to directly demonstrate techniques used in current continuous gravitational-wave searches.
 We then consider more complex audio (music and speech) for a more general interest in signal processing and explore the use of the interferometer as an optical microphone. 
@@
 We find that the logMMSE produces the best results, however, it does not enhance the recording to an intelligible level. 
-The complex audio signals can also serve as more distant analogues to un-modelled gravitational-wave burst signals, e.g. from supernovae.
+The complex audio signals can also serve as more distant analog to un-modeled gravitational-wave burst signals, e.g. from supernovae.
 
@@
 It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
-The increased excitement and public interest in the field of gravitational-wave research in recent years means that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
+The increased excitement and public interest in the field of gravitational-wave research in recent years mean that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
 The field of gravitational-wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.

### 2021-04-07 paper/ifo-futureWork.tex
moving footnotes to endnote, starting on changing US spelling
the low-hanging fruit of edits

@@
 All other components of the circuit draw power from the Pi, so this would isolate the circuit from the mains. 
-The laser is DC powered through a converter, so should not couple noise from the mains into the interferometer. 
+The laser is DC-powered through a converter, so the laser should not couple noise from the mains into the interferometer. 
 Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}.
@@
 Flowchart for the signal through the optical microphone. The transfer function would have to account for each stage. 
-The stage inside the red-dash box is analysed in Appendix~\ref{app:intensity_derivation}. Understanding the other stages is one aim of future work. 
+The stage inside the red-dash box is analyzed in Appendix~\ref{app:intensity_derivation}.
+Understanding the other stages is one aim of future work.
+Not shown are the signal processing filters that are performed on the PC once the data is transferred off the Pi.
 }
@@
 
-A thorough characterisation of the system's transfer function would allow us to understand how audio signals couple through the interferometer. 
+A thorough characterization of the system's transfer function would allow us to understand how audio signals couple through the interferometer. 
 The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Pi.
@@
 The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the abstract signal flow through the system.
+It does not include other potential pathways for signal flow such as acoustomechanical vibrations of the webcam or photodiode mount which could also be examined.
 % commented out whilst James is checking this
@@
 
-Returning to the demonstration of continuous gravitational-wave analysis, further extensions could include analysing data from multiple interferometers to extract directional information for audio signals.
-This would require coupling from a separate noise source instead of from a speaker attached directly to one of the mirrors of the interferometer.
+Returning to the demonstration of continuous gravitational-wave analysis, further extensions could include analyzing data from multiple interferometers to extract directional information for audio signals.
+This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
+Another extension would be to demonstrate the Doppler effect and the continuous-wave techniques used in the literature to address it by moving the optical table on some kind of circular track with frequency far higher than the wandering signal and placing a distant signal source at different positions above or beside the track -- or perhaps more practically the source could move instead of the detector.
 

### 2021-04-07 paper/ifo-introduction.tex
the low-hanging fruit of edits

@@
 The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
-Scorpius X-1 is a prime LMXB target for continuous wave searches. 
-%Scorpius X-1 is an LMXB which is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
+Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
+%Scorpius X-1 is an LMXB that is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
 Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.
@@
 % Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
-%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational wave research groups around the world.
+%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world.
 This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
@@
 In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
-In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyse the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches.
+In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyze the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches.
 In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
 % removed as there's not space to really say what this means here: to varying degrees of success.
-This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples which can be used in the undergraduate laboratory. 
+This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory. 
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 

### 2021-04-07 paper/ifo-singleTone.tex
moving footnotes to endnote, starting on changing US spelling
the low-hanging fruit of edits

@@
 Recovery of a tone at a constant frequency: Here we plot the Fourier amplitude of the intensity pattern against frequency.
-The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has a FWHM of $0.033\,{\rm Hz}$.
+The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has an FWHM of $0.033\,{\rm Hz}$.
 The plot also shows two harmonics at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ with amplitudes $10.3 \%$ and $9.8 \%$ of the amplitude of the primary peak, respectively.
@@
 
-% rearranged as I found this section to switch between method and results too much, I've tried this order now: intro, webcam and data specifics, injection and results.
+% rearranged as I found this section to switch between method and results too much, I've tried this order now: intro, webcam and data specifics, injection, and results.
 
-Continuous-wave searches look for nearly monochromatic signals. In this section, we consider a simple sinusoidal tone at a single, constant frequency. %should be single constant frequency *signal*?
+Continuous-wave searches look for nearly monochromatic signals. In this section, we consider a simple sinusoidal tone at a single, constant frequency. %should be a single constant frequency *signal*?
 
@@
 The intensity of the interference pattern is measured at a single point on the screen, indicated by the pink star in Fig~\ref{fig:interference_pattern}. 
-The webcam records video in three colour channels; red, blue, and green. 
+The webcam records video in three color channels; red, blue, and green. 
 We use the green channel as an approximation of the total intensity produced by the monochromatic green laser.
@@
 The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}.
+The Discrete Fourier transform was calculated using the NumPy package for Python~\cite{numpy}~\footnote{\url{https://numpy.org/doc/stable/reference/routines.fft.html}}.
 We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$.
@@
 They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo} and Appendix~\ref{app:intensity_derivation}.
+Also note that the noise appears to rise at lower frequencies.
 

### 2021-04-07 paper/ifo-supplementary-material.tex
the low-hanging fruit of edits

@@
 
-In section V of the main article, we use the inerferoeter as an optical microphone. 
-We use signal processing techniques to recover audio signals o speech and music. 
-In this supplementary material we provide futher details of a selection of analysis applied to the data which may be considered for undergraduate laboratory examples.  
+\begin{figure*}
+	\includegraphics[width=0.59\textwidth]{figures/setup_pic2.pdf}
+	\includegraphics[width=0.4\textwidth]{figures/circuit_pic3.pdf}
+	\caption{
+Left: A photograph of the optical microphone with an inset of the photodiode.
+The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right.
+In the main image, the photodiode is placed behind a cloth screen as explained in the main article.
+The inset shows a face-on view of the photodiode with the cloth screen removed.
+Right: A photograph of the photodiode circuit assembled on a breadboard. The leads from the photodiode enter from the left of the picture. The Raspberry Pi is shown at the top of the picture.
+}
+	\label{fig:photographs_of_optical_microphone_and_circuit}
+\end{figure*}
+
+In Section~V of the main article, we use the interferometer as an optical microphone.
+Photographs of the physical apparatus of the optical microphone are shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit}.
+We use signal processing techniques to recover audio signals of speech and music. 
+In this supplementary material we provide further details of a selection of analysis applied to the data which may be considered for undergraduate laboratory examples.  
 We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 

### 2021-04-07 paper/ifo-tableTopGWs.tex
small consistency edits
the low-hanging fruit of edits

@@
 Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern.
-The resulting interference pattern is dependent on the relative distance travelled by the beams. 
-Current generation gravitational-wave observatories have kilometre-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
+The resulting interference pattern is dependent on the relative distance traveled by the beams. 
+Current generation gravitational-wave observatories have kilometer-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
 They can measure minuscule changes in distance due to gravitational waves, e.g.
@@
 The interference pattern produced by the table-top Michelson interferometer.
-This image was taken with the webcam off the beam axis.%, which may explain why the (predicted circular) pattern appears elliptical.
-The pink star indicates the point where the intensity time series measurements were taken.
+This image was taken with the webcam aligned off the beam axis. %, which may explain why the (predicted circular) pattern appears elliptical.
+The pink star indicates the point where the intensity timeseries measurements were taken.
 The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
@@
 % removed as almost identical to the next sentence
-%In the following section, we start simply by injecting and recovering a audio signal of constant frequency (a single tone).
+%In the following section, we start simply by injecting and recovering an audio signal of constant frequency (a single tone).
 

### 2021-04-07 paper/ifo-wanderingTone.tex
the low-hanging fruit of edits

@@
 
-Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational-waves may not be monochromatic.
+Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational waves may not be monochromatic.
 A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
-The audio analogue is a tone that wanders in frequency: a note that changes in pitch.
+The audio analog is a tone that wanders in frequency: a note that changes in pitch.
 
-A Fourier transform applied to the whole dataset (as in Section~\ref{todo}) is not well suited to the wandering frequency case, as the signal can be spread across multiple frequency bins making it challenging or perhaps impossible to detect.
+A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case, as the signal can be spread across multiple frequency bins making it challenging or perhaps impossible to detect.
 %A different means of detection is needed for a signal which has a randomly wandering frequency. 
@@
 In this context, ``significantly'' means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
-One method to search for a wandering signal is to split the timeseries data into several shorter segments which are analysed individually.%, and then recombine them under some continuity assumption to coherently analyse the whole set.
+One method to search for a wandering signal is to split the timeseries data into several shorter segments which are analyzed individually.%, and then recombine them under some continuity assumption to coherently analyze the whole set.
 % I think the following is more accurate. The antenna beam pattern is used to compute a detection statistic. Which detection statistic to use depends on the type of target. -Hannah
@@
 In this work, we take the discrete Fourier transform of each timeseries segment and use its amplitude at each frequency as a detection statistic.
-This maximizes the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood}. 
+This maximizes the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood} -- we assume that that the noise is Gaussian.
 The detection statistic is normalized for convenience and we use it to form a grid in time and frequency (a spectrogram) which displays the frequency content of the signal over time. 
@@
 At every $t_i$ all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}). 
-Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}.%, where the circles indicate the elements in the spectrogram with the lighter colours corresponding to a higher likelihood.
+Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}.%, where the circles indicate the elements in the spectrogram with the lighter colors corresponding to a higher likelihood.
 The implementation used in this work is available online (see Appendix~\ref{app:code}).
@@
 These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the gray lines show the rejected paths.
-For example, the most probable connection to the element labelled (a) is the one directly behind it (i.e., $f_4$).
+For example, the most probable connection to the element labeled (a) is the one directly behind it (i.e., $f_4$).
 Therefore, this path is selected as the best path from $t_0$ to $t_1$ for $f_4$.
@@
 
+The key to the Viterbi algorithm is that we can use the signal's correlation in time to improve the signal-to-noise ratio of the frequency of the wandering signal at a particular time.
+Since the noise is assumed to not be correlated in time, if the detection statistic at a given frequency is high at a particular time and at neighboring frequencies at neighboring times then the likelihood that the given frequency at that particular time is from the wandering signal increases.
+By iterating this process through the grid for a suitable choice of the window of neighboring frequencies and times, the Viterbi algorithm is well-suited to finding wandering signals.
 In the following section, we use the Viterbi algorithm to recover a slowly wandering signal.
@@
 The color of each circle indicates the likelihood that a signal is present in the box, where a lighter color corresponds to a higher likelihood. 
-At each time, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines in the diagram). 
+At each time, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines connecting the circles in the diagram). 
 Each circle has three paths leading to it, indicated by the lines. 
@@
 Recovery of a wandering tone using the Viterbi algorithm.
-The spectrogram shows the observed frequency (Fourier) amplitude in each time-frequency bin. 
+The spectrogram (heatmap) colouring indicates the value of the detection statistic (here, the logarithm of the absolute value of the discrete Fourier transform) in each time-frequency bin -- with brighter colours indicating higher values.
 The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path, respectively. 
 On the left, before $\sim 15\,{\rm s}$, the signal changes frequency too quickly for the Viterbi algorithm to recover, given a parameter for the rate of change of one frequency bin per time bin. 
-At $150\,{\rm s}$ the data appears anomalous, which may be due to some background noise. }
+At $150\,{\rm s}$ the data appears anomalous, which may be due to some transient background noise. }
 \end{figure*}
@@
 We use the same set-up as shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
-In this section we test the Viterbi algorithm’s ability to recover the wandering signal.
+In this section, we test the Viterbi algorithm’s ability to recover the wandering signal. 
+Note that we implicitly approximate the noise at the webcam as white (uniform in frequency) in this implementation of the Viterbi algorithm.
+This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum} but is a sound approximation for the higher-frequency (around 5~Hz) wandering signals that we study.
 

### 2021-04-07 paper/paper_main.tex
moving footnotes to endnote, starting on changing US spelling
some more notes on comments
the low-hanging fruit of edits

@@
 %\documentclass[prb,reprint,nofootinbib]{revtex4-1} 
-\documentclass[aps,pra,superscriptaddress,reprint,nofootinbib]{revtex4-1}
+\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-1}
 % \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
@@
 \usepackage{subfiles}
-% \usepackage[bottom]{footmisc}
 \usepackage{comment}
@@
 
-%\usepackage{dcolumn}
-%\usepackage{bm}
-%\usepackage[pdftex,plainpages=false,colorlinks=true]{hyperref}
-%\usepackage[toc,page,title,titletoc,header]{appendix} 
-%\usepackage{multirow}
-%\usepackage{tabularx}
-%\usepackage[bottom]{footmisc}
-%\usepackage{caption}
-%\usepackage{subcaption}
-%\usepackage[bottom]{footmisc}
-%\usepackage{soul}
-
-%\bibliographystyle{apsrev4-2}
-% \setlength{\parindent}{0pt}
-
 % inner product
@@
 
-%\title{GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer}
 \title{Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
@@
 School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+\affiliation{%
+Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia 
 }
@@
 
-%\author{\han{to do: check names - initials?}}
-%\affiliation{%
-%\han{to do: check all affiliations}
-%}
-
 % Summer 2019/2020
@@
 Gravitational-wave observatories around the world are searching for continuous waves: persistent signals from sources such as spinning neutron stars. 
-% \jam{For the hyphen in "gravitational wave" versus "gravitational-wave". I prefer the former, but either way, we should be consistent.}
 These searches use sophisticated statistical techniques to look for weak signals in noisy data. 
-In this paper, we demonstrate these techniques using a table-top model gravitational-wave detector: a Michelson interferometer where sound is used as an analogue for gravitational waves. 
-Using signal processing techniques from continuous-wave searches we demonstrate the recovery of tones with constant and wandering frequencies. 
-We explore the use of the interferometer as a teaching tool for educators in physics and electrical engineering by using it as an ``optical microphone to capture music and speech. 
+In this paper, we demonstrate these techniques using a table-top model gravitational-wave detector: a Michelson interferometer where sound is used as an analog for gravitational waves. 
+Using signal processing techniques from continuous-wave searches, we demonstrate the recovery of tones with constant and wandering frequencies. 
+We explore the use of the interferometer as a teaching tool for educators in physics and electrical engineering by using it as an ``optical microphone'' to capture music and speech. 
 % \jam{We mention undergraduate labs twice, roll this sentence into the end? Also, electrical sounds weird, why the change from electrical engineering?}
@@
 % \jam{I changed all the \emph to ``'', feel free to change them back, but I think it is more explicit this way}
-We present a range of example filtering activities to recover the original signal from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
+We present a range of example filtering techniques to recover the original signal from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
 Recordings of simple chords and drums are easily recovered, but complex music and speech are more challenging. 

### 2021-04-08 paper/response/AJPRefereeResponse.tex
added response about yesterday's changes

@@
 \begin{itemize}
-\item  \todo{...}
+\item We have added a sentence (``The Discrete Fourier transform was calculated using the NumPy package for Python.''), a citation (to the creators of NumPy), and a footnote (to the documentation).
 \end{itemize}
@@
 \begin{itemize}
-\item  \todo{...}
+\item An explanation of the color coding has been added to the caption.
+\item \todo{compare to without VA}
+\item \todo{compare to LIGO literature}
 \end{itemize}
@@
 (Section V, paragraph 2)
-\item Fig. 6: agreed, we have removed this figure. \todo{decide if we want the photo in supplementary material}
-\item \todo{clarification of cloth screen }
+\item Fig. 6: we have removed this figure and moved the photo to the Supplementary Material.
+\item Cloth screen: we have added an additional sentence explaining the cloth screen mount and the source of the cloth screen cover (it is the ``grill cloth'' and plastic frame from a dismantled commercial speaker).
 \end{itemize}
@@
 \begin{itemize}
+\item Added an explanation in the caption that the filters are performed separately, after the data is gathered.
+\item Most of the code is readily available on the provided GitHub with the exception of the Weiner filtering results.
+The provided code is documented and commented if someone wanted to replicate the filter implementations.
+\item \todo{We may need to have access to Changrong's processing code}
 \item \todo{...}
@@
 \begin{itemize}
+\item A mention has been made in the Future Work as to demonstrating the Doppler effect.
 \item \todo{...}
@@
 \begin{itemize}
-\item \todo{...}
+\item Mention of this is now made both at the introduction of the plot and in the Viterbi algorithm section -- this implementation does assume white noise.
 \end{itemize}
@@
 \begin{itemize}
-\item \todo{...}
+\item This is unknown to us but has now been mentioned in Future Work.
 \end{itemize}

### 2021-04-13 paper/response/AJPRefereeResponse.tex
going through more comments

@@
 \todo{Opening comments}
+Dear Joseph Amato,
+
+We thank you and the referees for your helpful comments and suggestions. 
+We address the comments below in turn.
 
@@
 \item Fig. 6: we have removed this figure and moved the photo to the Supplementary Material.
-\item Cloth screen: we have added an additional sentence explaining the cloth screen mount and the source of the cloth screen cover (it is the ``grill cloth'' and plastic frame from a dismantled commercial speaker).
+\item The cloth screen was used to reduce the intensity of light on the photodiode and avoid saturating the analog-to-digital converter.
+It was indeed an improvised solution using the ``grill cloth'' and plastic frame from the dismantelled commercial speaker used in the demonstration.
+We have added a sentence to Section V.A.
+to clarify.
 \end{itemize}
@@
 \begin{itemize}
-\item Added an explanation in the caption that the filters are performed separately, after the data is gathered.
+\item The filters are performed separately after the data is gathered. We have added a sentence to the figure caption to clarify this. 
 \item Most of the code is readily available on the provided GitHub with the exception of the Weiner filtering results.
@@
 \begin{itemize}
-\item \todo{...}
+\item This equation was indeed incorrect and we thank you for the careful review.
+The equation should give the probability for a specific path, and is not a sum over all possibilities.
+We have corrected Eq.
+1 and added clarification to the text.
 \end{itemize}

### 2021-04-13 paper/ifo-complexAudio.tex
going through more comments

@@
 We use complex audio signals (such as music and speech) as a natural successor to the constant and wandering tones used in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} respectively.
-As complex audio signals are not quasi-monochromatic, the Viterbi algorithm as applied in Section~\ref{sec:viterbi_wandering} is not directly applicable here. 
+As complex audio signals are not quasi-monochromatic, the Viterbi algorithm as used in Section~\ref{sec:viterbi_wandering} is not directly applicable here. 
 Instead, we use a hierarchy of passive filters which suppress noise, yet do not assume any specific form of the signal, unlike the Fourier-based maximum likelihood filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
@@
 The photodiode records the interference pattern at roughly the same off-center position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
-The photodiode is mounted on a piece of cloth stretched over a plastic frame, with the electrical leads connected underneath the cloth.
-This cloth screen mount was the ``grill cloth'' from the front of a dismantled commercial speaker, re-purposed. 
+% I think this is probably not needed as I believe the editor comment is referring to the cloth screen over the photodiode, which is addressed below. 
+%The photodiode is mounted on a piece of cloth stretched over a plastic frame, with the electrical leads connected underneath the cloth.
+%This cloth screen mount was the ``grill cloth'' from the front of a dismantled commercial speaker, re-purposed. 
 The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
-Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. Resources for using the Pi and photodiode are described in Appendix~\ref{app:circuit_diagram}.
+Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. 
+Resources for using the Pi and photodiode, including a circuit diagram, are described in Appendix~\ref{app:circuit_diagram}.
 
@@
 This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
-We also place a second cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC -- an improvised, physical solution that could instead be replaced by scaling down the voltage electronically.
-This cloth screen was also re-purposed grill cloth from a commercial speaker.
+We also place a cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC -- an improvised, physical solution that could instead be replaced by scaling down the voltage electronically.
+This cloth screen was re-purposed grill cloth from a commercial speaker.
 % It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
@@
 %An updated schematic and a photograph of the entire optical microphone are shown in the left and right panels of Fig.~\ref{fig:ifo_schematic_podo}, respectively.
-See Appendix~\ref{app:circuit_diagram} for a diagram of the circuit. 
-\han{[Hannah] I've removed the second schematic and photo for now - could be an option for the photo to go in the appendix / supplement if we want to show the physical set up?}
+%See Appendix~\ref{app:circuit_diagram} for a diagram of the circuit.
 
@@
 The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass hum throughout. 
-This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), as shown in the top panel of Fig.~\ref{fig:psd_noise}. 
+This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), \fixme{as shown in the top panel} of Fig.~\ref{fig:psd_noise}. 
 The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ mains signal up to and beyond $8$th harmonic thereof. 
@@
 
-Here we present the results of two analyses using advanced signal processing techniques.
-They are briefly described and the results shown (see Supplementary Material for details).
+In this section we present results of two advanced signal processing analysis techniques applied to the optical microphone recordings. 
+The techniques are only briefly described here and we refer the reader to the accompanying Supplementary Material for further details and some other analysis techniques.
 
@@
 A notch filter removes signals within a specific frequency range \han{ref}. 
-In our case, we would like to remove the $50\,{\rm Hz}$ mains noise and harmonics. 
-Therefore, we use a cascaded notch filter where each notch is centered at each of the harmonics. 
-The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. 
+In this case, our aim is to remove the $50\,{\rm Hz}$ mains noise and harmonics, therefore we use a cascaded notch filter where each notch is centered at each of the harmonics. 
+The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. \han{ref}
 It amplifies parts of the signal with high SNR while suppressing pares with low SNR. 
-
-\han{refer to figure.. } shows our results.
-We find that the combined cascaded notch and Wiener filter produces a better result than either alone. 
+The results of the combined cascaded notch and Wiener filter are shown in the third row of Fig.~\fixme{XXXXX} where the timeseries and spectrum are shown by the left and right panels respectively. 
 Most of the mains noise is removed, however, the recovered voice sounds muffled and is not understandable.
@@
 Ref.~\cite{SubjectiveComparison} compares $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech (see also Ref.~\cite{Ephraim1984SpeechEU_logMMSE} for details on the logMMSE estimator).
-We use an existing implementation of the logMMSE~\cite{logmmse}.
-
-\han{refer to figure} 
-After applying the logMMSE estimator we see significant attenuation of the mains harmonics and general smoothing of the spectrum. 
+We use an existing implementation of the logMMSE from~\cite{logmmse}.  
+The logMMSE estimator results are shown in the bottom panel of Fig.~\fixme{XXXXX}. 
+We see significant attenuation of the mains harmonics and general smoothing of the spectrum. 
 Most of the background noise is removed, however, the logMMSE does not significantly enhance the speech as the voice sounds muffled and indistinct.

### 2021-04-16 paper/ifo-complexAudio.tex
fixing plots - part 1

@@
 \begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/psd_butterworth_14_6.pdf}
+	\includegraphics[width=.49\textwidth]{figures/psd_podo_14_6.pdf}
 	\caption{\label{fig:psd_noise}
@@
 We can include the twp panel version in the butterworth section of the supplementary material}
-Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
-Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
-We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
-After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
+Power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
+% Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
+We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling).
+Otherwise, the PSD is fairly white except for a peak at around 0.75~kHz.
+% After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
 }

### 2021-04-16 paper/ifo-supplementary-material.tex
going through more comments
fixing and combining plots - part 2

@@
 \begin{figure*}
-	\includegraphics[width=0.59\textwidth]{figures/setup_pic2.pdf}
-	\includegraphics[width=0.4\textwidth]{figures/circuit_pic3.pdf}
+	\includegraphics[height=0.41\textwidth]{figures/setup_pic2.pdf}
+	\includegraphics[height=0.41\textwidth]{figures/circuit_pic3.pdf}
 	\caption{
@@
 	\caption{\label{fig:BackgroundNotchWienerCombined}
+	\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}
 	Timeseries (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
@@
 \begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/psd_butterworth_14_6.pdf}
+	\includegraphics[width=.45\textwidth]{figures/psd_butterworth_14_6.pdf}
 	\caption{\label{fig:psd_noise}
@@
 \label{sec:notch}
-\begin{figure*}
+\begin{figure}
 \begin{center}
-\includegraphics[width=.9\textwidth]{figures/notch_filter_response.jpg}
+\includegraphics[width=.45\textwidth]{figures/cascaded_notch_response_plot.pdf}
 \end{center}
@@
 }
-\end{figure*}
+\end{figure}
 
@@
     \caption{\label{fig:logMMSE_timeseries_freqspectrum}
+\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}    
 Timeseries (left column) and frequency spectrum (right columns) results with the logMMSE estimator. 

### 2021-04-22 paper/ifo-supplementary-material.tex
work in progress on supp mat

@@
 \section{Optical microphone signal processing}
-\han{Notes - I've just made a start at moving material over. We will need to go through once we make final decisions on what to show in the main text / when new figures ready to check things}
+
+In Section~V in the main article, we use the interferometer as an optical microphone; audio signals are used to vibrate the interferometer mirrors and the signals are recovered from a timeseries of the interference pattern changing.
+In this Supplementary Material we include a selection of signal processing techniques applied to the data which may be considered for undergraduate laboratory examples. 
+We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
+We finish by combining these with advanced statistical techniques (Wiener filter and the logMMSE estimator). 
+As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
 
@@
 
-In Section~V of the main article, we use the interferometer as an optical microphone.
-Photographs of the physical apparatus of the optical microphone are shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit}.
-We use signal processing techniques to recover audio signals of speech and music. 
-In this supplementary material we provide further details of a selection of analysis applied to the data which may be considered for undergraduate laboratory examples.  
-We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
-We finish by combining these with advanced statistical techniques (Wiener filter and the logMMSE estimator). 
-As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
+All analysis is done using the photodiode set-up as described in Section~V.A. of the main text. 
+A photograph of the equipment is shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit} with the interferometer and circuit shown on the left and right respectively.
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
@@
     % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
-	\includegraphics[width=.8\textwidth, trim={1cm 3cm 1cm 1cm}]{figures/notch_and_wiener_superplot_v2.pdf}
+	\includegraphics[width=.5\textwidth]
+% trim={1cm 3cm 1cm 1cm}]
+{figures/combined_results_full_melatos_labelled.pdf}
+%notch_and_wiener_superplot_v2.pdf}
 	\end{center}
-	\caption{\label{fig:BackgroundNotchWienerCombined}
+	\caption{\label{fig:op_mic_results}
 	\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}
@@
 	}
-\end{figure*}
+\end{figure}
 
-Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results, respectively. 
-The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} show the source signal and the raw optical microphone recording, respectively. 
+Results for a selection of filters are collated in Fig.~\ref{fig:op_mic_results}, which re refer to throughout the text. 
+The left and right columns show timeseries and Fourier spectrum results, respectively. 
+The first and second rows of Fig.~\ref{fig:op_mic_results} show the source signal and the raw optical microphone recording, respectively. 
 The third row shows a background spectrum of the optical microphone. 
@@
 
-Given that we have access to the background noise spectrum, an intuitive way to remove noise is via subtracting the noise spectrum from the recorded spectrum. 
-The fourth row of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the spectrum obtained after subtracting the background noise spectrum. 
+An inituitive way to remove noise is by directly subtracting the background noise (third row of Fig.~\ref{fig:op_mic_results}) from the recorded spectrum. 
+The fourth row of Fig.~\ref{fig:op_mic_results} shows the spectrum obtained after subtracting the background noise spectrum. 
 We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified.
@@
 The amplitude (magnitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
-The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
+The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row of Fig.~\ref{fig:op_mic_results}.
+We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row of Fig.~\ref{fig:op_mic_results}.
 
@@
 For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
-The timeseries and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
+The timeseries and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:op_mic_results}.
 We see a significant improvement in the timeseries of the recovered signal, however, a strong noise hum persists, audibly.
@@
 
-By combining the notch and Wiener filter, we can trade-off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
+By combining the notch and Wiener filter, we can trade-off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:op_mic_results}.
 The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone.

### 2021-04-23 paper/ifo-supplementary-material.tex
adding text based on Changrongs notes

@@
 \label{eq:butterworth}
-H(\omega) = \left[1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right]^{-1/2},
+H(f) = \left[1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right]^{-1/2},
 \end{equation}
@@
 Here, the notches are smooth in comparison to the na{\"i}ve zeroing of each frequency (which looks like a rectangular comb).
-The complex frequency response of a typical infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
+
+\han{The impulse response describes the reaction of a system to an input signal as a function of time. 
+If a system is linear and time-invarient, the output signals are completely determined for any input signal using the impulse response. 
+Finite inpulse response (FIRs) filters have finite length (they equal zero outside of a finite range). 
+By contrast an infinite impulse response (IIR) filter has infinite length due to feedback. 
+Typically FIR filters outperform IIR filters as they are always stable. 
+However, IIR filters normally require fewer coefficients which in turn speeds up signal processing in comparison to FIR filters. 
+}
+
+\han{In this work we opt for a IIR notch filter. 
+We write the filter in the form of a Z-transform, in which a discrete-time signal is converted to the complex frequency domain $z=e^{jf}$. 
+The IIR notch filter we use here is given by~\citep{10.5555/541204} }
 \begin{equation}
@@
 \end{equation}
-where $z$ is the complex frequency and $\alpha$ and $\beta$ are parameters that control the filter.
-$w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
+where $\alpha$ and $\beta$ are parameters that control the filter. 
+In Eqn.~\ref{eqn:notch}, $w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
 We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centred on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$. 
-When then choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
+We choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
 The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,

### 2021-04-23 paper/ifo-complexAudio.tex
fixing and combining plots - part 2
adding new plot from James into text, fixing the placeholders

@@
 \begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/psd_podo_14_6.pdf}
+	\includegraphics[width=.45\textwidth]{figures/psd_podo_14_6.pdf}
 	\caption{\label{fig:psd_noise}
@@
 
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{figures/combined_highlight_results_melatos_labelled.pdf}
+\caption{\label{fig:notchWienerLogMMSEResults}
+Timeseries (left column) and frequency spectrum (right column) results with the optical microphone. 
+The original input signal in the first row is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode''). 
+\han{In the previous version of the plot we shifted the top timeseires by 0.12 seconds and wrote: 
+``The source signal is shifted by 0.12 s to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker.'' - need to correct}
+The second row shows the raw output from the optical microphone when the input from the first row is played. 
+The third row shows the result of applying the notch and Wiener filter combined. 
+The fourth row shows the result of applying the logMMSE estimator, where the rise at the start of the timeseries is an expected effect when filtering a signal of finite duration. 
+}
+\end{center}
+\end{figure*}
+
 We explore several filters to remove the $50\,{\rm Hz}$ mains hum and harmonics, and also to improve the speech intelligibility of the recording.
-The accompanying supplementary material describes a range of analysis techniques that can be used as examples for the undergraduate laboratory. 
+The accompanying Supplementary Material describes a range of analysis techniques that can be used as examples for the undergraduate laboratory. 
 All filters are tested on the same $1\,{\rm s}$ long speech recording.
+The results of this section are shown in Fig.~\ref{fig:notchWienerLogMMSEResults} which we refer to throughout. 
+In Fig.~\ref{fig:notchWienerLogMMSEResults} the left and right columns show the timeseries and frequency spectrum respectively. 
+The first row shows the input signal played through the speaker (see Fig.~\ref{fig:ifo_schematic_webcam}). 
+The second row shows the raw output from the photodiode recording.
 
@@
 All three of these cases can, in extreme cases, worsen speech intelligibility beyond the unfiltered recording and so we choose filters that compromise between achieving the three conditions.
-\han{[Notes] editor's comments suggest expanding on this?}
 % https://en.wikipedia.org/wiki/Filter_(signal_processing)
@@
 Firstly, we consider two signal processing techniques used in combination: the cascaded notch and the Wiener filter. 
-A notch filter removes signals within a specific frequency range \han{ref}. 
+A notch filter removes signals within a specific frequency range. 
 In this case, our aim is to remove the $50\,{\rm Hz}$ mains noise and harmonics, therefore we use a cascaded notch filter where each notch is centered at each of the harmonics. 
-The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. \han{ref}
-It amplifies parts of the signal with high SNR while suppressing pares with low SNR. 
-The results of the combined cascaded notch and Wiener filter are shown in the third row of Fig.~\fixme{XXXXX} where the timeseries and spectrum are shown by the left and right panels respectively. 
+The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. 
+It amplifies parts of the signal with high signal-to-noise (SNR) ratio while suppressing parts with low SNR. 
+The results of the combined cascaded notch and Wiener filter are shown in the third row of Fig.~\ref{fig:notchWienerLogMMSEResults} where the timeseries and spectrum are shown by the left and right panels respectively.
 Most of the mains noise is removed, however, the recovered voice sounds muffled and is not understandable.
@@
 We use an existing implementation of the logMMSE from~\cite{logmmse}.  
-The logMMSE estimator results are shown in the bottom panel of Fig.~\fixme{XXXXX}. 
+The logMMSE estimator results are shown in the bottom panel of Fig.~\ref{fig:notchWienerLogMMSEResults}. 
 We see significant attenuation of the mains harmonics and general smoothing of the spectrum. 

### 2021-04-23 paper/response/AJPRefereeResponse.tex
fixing plots - part 1
updating response with progress on comments so far

@@
 \begin{itemize}
-\item \todo{...}
+\item The Butterworth filter has strongly attenuated the signal outside the passband.
+\item \todo{They will also expect a change to the paper as it is a point of confusing - e.g. clarify the reason in the figure caption.}
 \end{itemize}
@@
 \begin{itemize}
-\item \todo{...}
+\item We have move the majority of Section VC to Supplementary Material and now only include highlight results in the main text (the combined notch and Wiener filter and the logMMSE estimator).
+In the Supplementary Material we have expanded the description of the impulse response of a filter mentioning both IIR and FIR.
+We have also added some text saying that Eq.
+2 in the Supplementary Material (previously Eq.
+4 in the main text) is a Z-tranform.
+We have corrected the inconsistency in Eq.
+1 in the Supplementary Material (previously Eq.
+3 in the main text).
+\todo{add some references to DSP textbooks.}
 \end{itemize}
@@
 \begin{itemize}
-\item \todo{...}
+\item Thank you, we have corrected the axis labels and they are now consistent.
 \end{itemize}
@@
 -- This is not a comment but just my curiosity: How strong is the effect of mechanical vibrations of the camera or photodiode mount on the observed signal?
-
 \begin{itemize}
@@
 
-\subsection{Comments from R1 }
+\subsection{Comments from R1}
 

### 2021-04-23 paper/ifo-supplementary-material.tex
updating text for new figures

@@
 
-\title{Supplementary Material} 
+\title{Supplementary material for \\Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone} 
 \maketitle
@@
 	\caption{\label{fig:op_mic_results}
-	\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}
+	%\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}
 	Timeseries (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
-    The source signal, shown in the first row, is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode ...''). 
-    The source signal is shifted by $0.12\,{\rm s}$ to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker. 
-	The second row shows the raw output from the optical microphone when the input signal from the first row is played. 
+    The top two rows are the input signal and raw output (identical to the top two rows in Fig.~7 in the main article). 
+    %The source signal, shown in the first row, is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode ...''). 
+    %The source signal is shifted by $0.12\,{\rm s}$ to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker.
+	%The second row shows the raw output from the optical microphone when the input signal from the first row is played. 
     The third row shows a background noise recording from the optical microphone when no audio signal is played. 
     The fourth row shows the result of subtracting the background noise spectrum in the third row from the recording in the second row. 
-	The fifth, sixth, and seventh rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
+	The fifth and sixth rows show the results of applying the cascaded-notch filter and Wiener filter respectively.
+    The seventh shows the results of applying both the cascaded-notch and Wiener filter (identical to the third row in Fig.~7 in the main article).
+    The eight row shows the results of applying the logMMSE estimator (identical to the fourth row in Fig.~7 in the main article). 
 	}
@@
     \caption{\label{fig:logMMSE_timeseries_freqspectrum}
-\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}    
+\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}    - done 
 Timeseries (left column) and frequency spectrum (right columns) results with the logMMSE estimator. 

### 2021-04-23 paper/ifo-complexAudio.tex
updating text for new figures
editing figure caption

@@
 \label{sec:opticalMicResults}
-\han{[Hannah] keen to hear other's thoughts, but I think it would be cool to show the notch+wiener combined and the logmmse as the results?
-They could be placed in a single figure for ease of comparison with time \& spectrum columns and rows: 1.
-original signal, 2.
-raw output, 3.
-combined cascaded notch and Wiener, 4.
-logMMSE}
 
@@
 The original input signal in the first row is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode''). 
-\han{In the previous version of the plot we shifted the top timeseires by 0.12 seconds and wrote: 
-``The source signal is shifted by 0.12 s to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker.'' - need to correct}
+ The input signal is shifted by 0.12 s to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker.
+\han{check this is still true for new versions}
 The second row shows the raw output from the optical microphone when the input from the first row is played. 

### 2021-04-23 paper/ifo-complexAudio.tex
fixing placeholder text in section V, putting authors back in for now (had commented for double blind review)

@@
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
-A selection of filters are applied to the data, details of which are presented in Supplementary Material and we present the results of the best performing filters in Section~\ref{sec:opticalMicResults}.
-\han{rephrase depending on what we show.}
+A selection of filters are applied to the data, details of which are presented in Supplementary Material. 
+We present the two best performing analyses from our Supplementary Material in Section~\ref{sec:opticalMicResults}.
 
@@
 	\caption{\label{fig:psd_noise}
-\han{If we do not cover the butterworth filter in the main text, I'd suggest only showing the top panel of this figure in the main text.
-We can include the twp panel version in the butterworth section of the supplementary material}
 Power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 

### 2021-04-28 paper/ifo-supplementary-material.tex
editing figure caption
finishing first pass through supplementary material with new plot references

@@
 
-In Section~V in the main article, we use the interferometer as an optical microphone; audio signals are used to vibrate the interferometer mirrors and the signals are recovered from a timeseries of the interference pattern changing.
+In Section~V in the main article, we use the interferometer as an optical microphone; audio signals are used to vibrate the interferometer mirrors, the changing interference pattern is recorded, and we aim to recover the original signal from the resulting timeseries data.
 In this Supplementary Material we include a selection of signal processing techniques applied to the data which may be considered for undergraduate laboratory examples. 
@@
 
-All analysis is done using the photodiode set-up as described in Section~V.A. of the main text. 
+All analysis is done using the photodiode set-up as described in Section~V.A. in the main text. 
 A photograph of the equipment is shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit} with the interferometer and circuit shown on the left and right respectively.
@@
     The top two rows are the input signal and raw output (identical to the top two rows in Fig.~7 in the main article).  
-    %The source signal, shown in the first row, is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode ...''). 
-    %The source signal is shifted by $0.12\,{\rm s}$ to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker.
-	%The second row shows the raw output from the optical microphone when the input signal from the first row is played. 
     The third row shows a background noise recording from the optical microphone when no audio signal is played. 
@@
 
-Results for a selection of filters are collated in Fig.~\ref{fig:op_mic_results}, which re refer to throughout the text. 
+Results for a selection of filters are collated in Fig.~\ref{fig:op_mic_results}, which we refer to throughout this Supplementary Material. 
 The left and right columns show timeseries and Fourier spectrum results, respectively. 
-The first and second rows of Fig.~\ref{fig:op_mic_results} show the source signal and the raw optical microphone recording, respectively. 
-The third row shows a background spectrum of the optical microphone. 
-The following sections describe each of the analysis in turn.
+The first and second rows of Fig.~\ref{fig:op_mic_results} show the original source signal and the raw optical microphone recording, respectively. 
+The following sections describe a sequence of analysis techniques and a selection of results are presented in Fig.~\ref{fig:op_mic_results}.
 
@@
 
-An inituitive way to remove noise is by directly subtracting the background noise (third row of Fig.~\ref{fig:op_mic_results}) from the recorded spectrum. 
-The fourth row of Fig.~\ref{fig:op_mic_results} shows the spectrum obtained after subtracting the background noise spectrum. 
+An inituitive way to remove noise is by directly subtracting the background noise from the recorded spectrum. 
+The third row in Fig.~\ref{fig:op_mic_results} shows the interferometer background noise when there is no input signal.
+The fourth row shows the spectrum obtained after subtracting the background noise spectrum, i.e. the result of subtracting the third row from the second row. 
 We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified.
@@
 
-\han{The impulse response describes the reaction of a system to an input signal as a function of time. 
+The impulse response describes the reaction of a system to an input signal as a function of time. 
 If a system is linear and time-invarient, the output signals are completely determined for any input signal using the impulse response. 
@@
 However, IIR filters normally require fewer coefficients which in turn speeds up signal processing in comparison to FIR filters.
-}
 
-\han{In this work we opt for a IIR notch filter. 
+In this work we opt for a IIR notch filter. 
 We write the filter in the form of a Z-transform, in which a discrete-time signal is converted to the complex frequency domain $z=e^{jf}$. 
-The IIR notch filter we use here is given by~\citep{10.5555/541204} }
+The IIR notch filter we use here is given by~\citep{10.5555/541204}, 
 \begin{equation}
@@
 Here, we experiment with applying a combination of the cascaded notch and the Wiener filter to the recorded speech signal. 
-The results of this analysis are described in the main paper in section \han{VC} and figure \han{XX}.
+The results of this analysis also are described in Section V.C. in the main text.
 
@@
 
-By combining the notch and Wiener filter, we can trade-off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:op_mic_results}.
+By combining the notch and Wiener filter, we can trade-off between the two and achieve an overall better performance, as can be seen in the seventh row in Fig.~\ref{fig:op_mic_results}.
 The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone. 
@@
 In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
-This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that is measures the MSE between the logarithm of the Fourier amplitudes.
+This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that it measures the MSE between the logarithm of the Fourier amplitudes.
 This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
@@
 
-\han{results for the logmmse estimator are shown in the main paper in section VC and figure X}
+The final row in Fig.~\ref{fig:op_mic_results} shows the results of the logMMSE estimator.
+Again there is significant noise reduction, however the speech remains indistict.
+The logMMSE results are also presented in Section V.C. in the main text.
 

### 2021-04-28 paper/ifo-supplementary-material.tex
typo fix

@@
 The left and right columns show timeseries and Fourier spectrum results, respectively. 
-The first and second rows of Fig.~\ref{fig:op_mic_results} show the original source signal and the raw optical microphone recording, respectively. 
+The first and second rows in Fig.~\ref{fig:op_mic_results} show the original source signal and the raw optical microphone recording, respectively. 
 The following sections describe a sequence of analysis techniques and a selection of results are presented in Fig.~\ref{fig:op_mic_results}.
@@
 
-The effect of applying this filter to the background noise PSD can be seen in the bottom panel of Fig.~\ref{fig:psd_noise}.
+The effect of applying this filter to the background noise PSD can be seen in the bottom panel in Fig.~\ref{fig:psd_noise}.
 The Butterworth band-pass filter reduces the amplitude of mains harmonics below $150\,{\rm Hz}$ and suppresses unrelated noise sources above $3\,{\rm kHz}$.
@@
 The amplitude (magnitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
-The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row of Fig.~\ref{fig:op_mic_results}.
-We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row of Fig.~\ref{fig:op_mic_results}.
+The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row in Fig.~\ref{fig:op_mic_results}.
+We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row in Fig.~\ref{fig:op_mic_results}.
 
@@
 For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
-The timeseries and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:op_mic_results}.
+The timeseries and frequency spectrum after applying the Wiener filter are shown in the sixth row in Fig.~\ref{fig:op_mic_results}.
 We see a significant improvement in the timeseries of the recovered signal, however, a strong noise hum persists, audibly.

### 2021-04-28 paper/ifo-complexAudio.tex
removing commented out text from previous version, fixing reference to figure 6

@@
 The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass hum throughout. 
-This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), \fixme{as shown in the top panel} of Fig.~\ref{fig:psd_noise}. 
+This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), as shown in Fig.~\ref{fig:psd_noise}. 
 The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ mains signal up to and beyond $8$th harmonic thereof. 
@@
  The input signal is shifted by 0.12 s to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker. 
-\han{check this is still true for new versions}
 The second row shows the raw output from the optical microphone when the input from the first row is played. 
@@
 
-\begin{comment} 
-In this section, we explore how this demonstration can be used to teach a selection of signal processing techniques.
-The apparatus used in Sections~\ref{sec:ifo} to~\ref{sec:viterbi_wandering} allows undergraduate students to explore a range of signal processing techniques and filters of interest and provides a foundation for many technical applications~\cite{DigitalProcingOfSpeechSignals:1978}.
-We provide a series of example activities using complex audio (such as music and speech) as a natural successor to constant and wandering tones. 
-As complex audio signals are not quasi-monochromatic, unlike continuous gravitational-wave signals, the Viterbi algorithm as implemented in Section~\ref{sec:viterbi_wandering} is not directly applicable here.
-Instead, in this section, we use the interferometer as an optical microphone to detect and playback complex audio signals (i.e.\ using light to capture sound).
-In doing so, we experiment with a hierarchy of passive filters which suppress noise, yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
-A vast literature exists on signal processing with speech, see Ref.~\cite{DigitalProcingOfSpeechSignals:1978} for introductory material including Fourier analysis and digital representations.
-Optical microphone devices have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defence industry and operate on a variety of related principles.
-
-This section is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals.
-Here, the optical microphone serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories.
-We describe the additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and initial results in Section~\ref{sec:initialResultsOpMic}. 
-A selection of filter and speech enhancement example activities using the optical microphone are outlined in Section~\ref{sec:filters}.
-
-\subsection{Hardware modifications for the optical microphone}
-\label{sec:photodiode}
-
-\begin{figure*}
-	\includegraphics[width=0.44\textwidth]{figures/ifo_schematic_photodiode_edit.pdf}
-	\includegraphics[width=0.52\textwidth]{figures/setup_pic2.pdf}
-	\caption{
-Left: the updated Michelson interferometer schematic. 
-The schematic is identical to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
-Right: a photograph of the optical microphone. 
-The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right. 
-In the main image, the photodiode is placed behind the cloth screen. 
-The inset shows a face-on view of the photodiode with the covering screen removed. 
-See Appendix~\ref{app:circuit_diagram} for a circuit diagram and detailed photograph of the circuit.
-}
-	\label{fig:ifo_schematic_podo}
-\end{figure*}
-
-The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$.
-Speech intelligibility (being able to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to (and beyond) $8\,{\rm kHz}$~\cite{speech_intelligibility}.
-Therefore, the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music (adjusting for the Nyquist frequency).
-This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$ and thus can only ``hear'' frequencies below $15\,{\rm Hz}$.
-To overcome this issue, we use a photodiode\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the reverse direction.
-As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.} at the output of the interferometer to achieve a sampling rate of $16\,{\rm kHz}$.
-
-We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage dependent on the incident intensity. 
-The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
-The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath the screen. 
-The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
-Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. Resources for using the Pi and photodiode are described in Appendix~\ref{app:circuit_diagram}.
-
-Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing\footnote{Folding of frequencies greater than half the sampling rate.} into the detected range.
-We include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
-This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
-We also place a second cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC.
-% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-
-An updated schematic and a photograph of the entire optical microphone are shown in the left and right panels of Fig.~\ref{fig:ifo_schematic_podo}, respectively.
-See Appendix~\ref{app:circuit_diagram} for a diagram and photograph of the circuit.
-
-\subsection{Anti-aliased output}
-\label{sec:initialResultsOpMic}
-
-%The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources). 
-We test the optical microphone with a variety of recordings, including speech from different people and music ranging from simple melodies and rhythms to songs.
-During recordings, care is taken to minimise activity around the demonstration to reduce the amount of environmental noise coupling into the interferometer.
-The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python~\cite{scipy,python}.
-When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.
-
-The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass hum throughout.
-This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), as shown in the top panel of Fig.~\ref{fig:psd_noise}.
-The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ mains signal up to and beyond $8$th harmonic thereof.
-The mains signal is also present (but far weaker) in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution.
-Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself.
-The appearance of harmonics of the mains noise is likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}).
-The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
-
-\begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/psd_butterworth_14_6.pdf}
-	\caption{\label{fig:psd_noise}
-Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
-Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
-We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
-After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
-}
-\end{figure}
-
-\subsection{A hierarchy of filters}
-\label{sec:filters}
-
-Here, we describe several filters used to remove the $50\,{\rm Hz}$ mains hum and its harmonics, but also to ultimately improve the speech intelligibility of the recording (to be able to understand the words spoken).
-We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
-We finish by combining these with advanced statistical techniques (Wiener filter and the logMMSE estimator). 
-As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
-
-\begin{figure*}
-	\begin{center}
-    %% trim order is <left> <lower> <right> <upper>
-    % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
-	\includegraphics[width=.8\textwidth, trim={1cm 3cm 1cm 1cm}]{figures/notch_and_wiener_superplot_v2.pdf}
-	\end{center}
-	\caption{\label{fig:BackgroundNotchWienerCombined}
-	Timeseries (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
-    The source signal, shown in the first row, is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode ...''). 
-    The source signal is shifted by $0.12\,{\rm s}$ to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker. 
-	The second row shows the raw output from the optical microphone when the input signal from the first row is played. 
-    The third row shows a background noise recording from the optical microphone when no audio signal is played. 
-    The fourth row shows the result of subtracting the background noise spectrum in the third row from the recording in the second row. 
-	The fifth, sixth, and seventh rows show the results of applying the notch filter, the Wiener filter, and both combined, respectively.
-	}
-\end{figure*}
-
-%\begin{figure*}
-%\begin{center}
-%% trim order is <left> <lower> <right> <upper>
-%\includegraphics[width=\textwidth,trim={1cm 8.5cm 1cm 8cm},clip]{figures/freqSpectrumOriginalNoiseSubtracted.pdf}
-%\end{center}
-%\caption{\label{fig:noiseSubtract}
-%Frequency spectrum of a $1\,{\rm s}$ signal before the use of filters. 
-%The first panel shows the original frequency spectrum of the human voice recording (saying ``A cathode ...''). 
-%The second panel shows the spectrum recorded after the signal passes through the optical microphone (sampled at $16\,000\,{\rm Hz}$).
-%The third panel shows a noise-only recording from the optical microphone when no audio signal is being played. 
-%The fourth panel shows the result of subtracting the noise-only spectrum in the third panel from the recording in the second panel. 
-%We see no obvious improvement from this method. 
-%}
-%\end{figure*}
-
-Results for a selection of filters are collated in Fig.~\ref{fig:BackgroundNotchWienerCombined} where the left and right columns show timeseries and Fourier spectrum results, respectively. 
-The first and second rows of Fig.~\ref{fig:BackgroundNotchWienerCombined} show the source signal and the raw optical microphone recording, respectively. 
-The third row shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.
-
-%The Fourier spectrum of the source signal is shown in the first panel of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-The second panel shows the spectrum of the raw optical microphone recording of the signal.
-The third panel of Fig.~\ref{fig:noiseSubtract} shows a background spectrum of the optical microphone, as shown previously in the top panel Fig.~\ref{fig:psd_noise}.
-
-Given that we have access to the background noise spectrum, an intuitive way to remove noise is via subtracting the noise spectrum from the recorded spectrum. 
-The fourth row of Fig.~\ref{fig:BackgroundNotchWienerCombined} shows the spectrum obtained after subtracting the background noise spectrum. 
-We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified. %Why the noise sources are time-varying is a fact yet to be determined.
-
-Beyond this na{\"i}ve approach, we apply a range of filters to the data.
-In signals processing, the ideal filter would be one that:
-i) completely attenuates the undesired parts of the spectrum, 
-ii) does not change the rest of the spectrum, and 
-iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
-However, these three conditions cannot all hold at once and any filter must compromise between them, as we will see below. 
-The following sections describe a series of example analyses with the optical microphone.
-
-\subsubsection{Rectangular comb filter}
-
-Simply zeroing the frequency bins corresponding to the harmonics of the $50\,{\rm Hz}$ mains signal is unsuccessful.
-This effectively multiplies the spectrum by a rectangular comb filter.
-It does remove the mains harmonics, but audibly ruins the rest of the signal due to lack of smoothness.
-This is because applying a filter in frequency space is equivalent to convolving the time-domain signal with the inverse Fourier transform of that filter.
-The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is some combination of sinc functions, which significantly corrupt the signal.
-See also Section~\ref{sec:notch} where we explore a notch filter.
-
-%A smoother comb filter significantly attenuates the spectrum between each harmonic so that the signal is no longer audible. 
-% \jam{[re:comb filters.
-When I tried to comb filter it either was too sharp and so jumbled the whole signal in convolution.
-Or if smooth attenuated the signal away into noise.
-At least, that's what I remember.
-Changrong has a comb filter section anyway so I am happy dropping my unsuccessful attempts.]}
-
-\subsubsection{High-pass filter}
-
-A high-pass filter smoothly attenuates frequencies below some cut-off frequency.
-Applying a high-pass filter, with a cut-off frequency around $150\,{\rm Hz}$, to the signal spectrum works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics.
-However, the mains harmonics above $100\,{\rm Hz}$ remain.
-Using a high-pass filter with a higher cut-off can be used to mitigate this issue.
-However, it makes the played-back signal unrecognisable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
-Often in speech processing, the logarithm of the signal is taken since the amplitude information seems to be more important to intelligibility than the phase information to the human ear~\cite{SubjectiveComparison}.
-However, applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.
-
-%We also tried applying a high-pass filter to the logarithm of the signal spectrum and then exponentiating back, but this didn’t significantly improve on the above simple high-pass
-
-\subsubsection{Butterworth band-pass filter}
-
-\begin{figure}
-	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000.pdf}
-	\caption{Butterworth bandpass filter frequency response.
-Any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power).
-The red, dashed vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$.
-Note the flat response within the band characteristic of the Butterworth filter.}
-	\label{fig:butterworth}
-\end{figure}
-
-A general band-pass filter combines a high-pass filter and a low-pass filter to smoothly attenuate frequencies outside of some band (alt.\ pass-band).
-A Butterworth band-pass filter is a particular band-pass filter such that the frequency response (the attenuation at each frequency) is ``maximally flat'' within the band.
-The Butterworth low-pass component is given by %Eqn.~\ref{eq:butterworth}
-
-\begin{equation}
-\label{eq:butterworth}
-H(\omega) = \left[1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right]^{-1/2},
-\end{equation}
-% commenting to send to co-authors
-%\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
-and is combined with a similar high-pass filter to form the band-pass filter.
-In Eqn.~\ref{eq:butterworth}, $f_c$ is the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ is the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
-Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here, a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}$, $3\,{\rm kHz})$.
-This high frequency cut-off is chosen since the frequencies important for speech generally lie below $2\,{\rm kHz}$~\cite{speech_intelligibility}.
-
-The effect of applying this filter to the background noise PSD can be seen in the bottom panel of Fig.~\ref{fig:psd_noise}.
-The Butterworth band-pass filter reduces the amplitude of mains harmonics below $150\,{\rm Hz}$ and suppresses unrelated noise sources above $3\,{\rm kHz}$.
-However, it does not address the issue of mains harmonics above $150\,{\rm Hz}$ (i.e., in the pass-band).
-In the following section, we experiment with a cascade notch filter to address this.
-
-\subsubsection{Cascade notch filter}
-\label{sec:notch}
-\begin{figure*}
-\begin{center}
-\includegraphics[width=.9\textwidth]{figures/notch_filter_response.jpg}
-\end{center}
-\caption{\label{fig:notchMagResponse}
-Amplitude (magnitude) response for the first five notches of fifteen for the cascaded notch filter described by Eqs.~\ref{eqn:notch} and~\ref{eqn:notch15}. 
-}
-\end{figure*}
-
-One method to remove mains noise and its harmonics is to use a sequence of notch filters centred at each of the frequency bins we want to remove.
-This sequence is known as a ``cascade'' of notches.
-Here, the notches are smooth in comparison to the na{\"i}ve zeroing of each frequency (which looks like a rectangular comb).
-The complex frequency response of a typical infinite impulse response (IIR) notch filter can be written as \citep{10.5555/541204}
-\begin{equation}
-    \label{eqn:notch}
-    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}\,,
-\end{equation}
-where $z$ is the complex frequency and $\alpha$ and $\beta$ are parameters that control the filter.
-$w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
-We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centred on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$.
-When then choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
-The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,
-\begin{equation}
-    \label{eqn:notch15}
-    H(z) = \prod_{k=0}^{14} H_k(z).
-\end{equation}
-
-We use the built-in MATLAB filter design toolbox in this work~\cite{MATLAB}. 
-The amplitude (magnitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
-The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-
-Although the notch filter removes much of the mains hum sound, the filtered recording is not intelligible.
-Qualitatively it sounds more like a drum than a human voice.
-This is due to loss of voice information under the filter.
-To overcome this, we turn to more advanced techniques, starting with the Wiener filter.
-Instead of just passively filtering different frequencies, this statistical technique optimises an estimate of the original signal.
-
-% \begin{figure*}
-% \begin{center}
-% \includegraphics[width=\textwidth,trim={1cm 8.2cm 1 8.3cm},clip]{figures/time SeriesNotchWiener.pdf}
-% \end{center}
-% \caption{\label{fig:timeNotchWiener}
-% Time series results for the notch and Wiener filters. 
-% The first panel shows the original speech recording which was played from the speaker. 
-% The second panel shows the data from the optical microphone before any filters are applied. 
-% The third and fourth panels show results after applying the notch and Wiener filter respectively. 
-% Finally, the fifth panel shows the result of applying both the notch and Wiener filter. 
-% }
-% \end{figure*}
-
-% \begin{figure*}
-% \begin{center}
-% \includegraphics[width=\textwidth,trim={0cm 8.2cm 0cm 8cm},clip]{figures/freqNotchWiener.pdf}
-% \end{center}
-% \caption{\label{fig:freqNotchWiener}
-% Frequency spectrum results for the notch and Wiener filters. 
-% The first and second panels show the results of applying the notch and Wiener filter respectively.
-% The third panel shows the spectrum after applying both the notch and Wiener filters. 
-% The first, second, and third panels show the spectrum of the time series show in the third, fourth, and fifth panels of Fig.~\ref{fig:timeNotchWiener} respectively. 
-% }
-% \end{figure*}
-
-%A notch filter attenuates signals in specific frequency bands. 
-%We apply a $15$ cascaded notched filter to remove the mains noise and its harmonics up to $750\,{\rm Hz}$. 
-%\han{should we include here a mathematical description of the $15$ cascade notch filter?}
-%\jam{[Yes, I have no idea what a cascaded notched filter is and I'm an author!]}
-
-\subsubsection{Wiener filter}
-\label{sec:Wiener}
-
-A Wiener filter is an advanced statistical technique that estimates the injected signal given prior information about the injected spectrum and the reference spectrum of the background noise. 
-The observed noisy speech signal sequence is given as $\mathbf{x}=(x(0),\dots, x(N-1))$, where $N$ is the length of the data sequence and $\mathbf{x}$ is the sum of the original injected signal $\mathbf{s}=(s(0),\dots,s(N-1))$, and the noise sequence $\mathbf{w}=(w(0),\dots,w(N-1))$,
-\begin{equation}
-    \mathbf{x}=\mathbf{s}+\mathbf{w}\,.
-\end{equation}
-Given $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean-square-error (BMSE) between the two, defined as 
-\begin{equation}
-\label{eq:BMSE}
-\text{BMSE}(\hat{\textbf{s}})=E[(\textbf{s}-\hat{\textbf{s}})^2]\,.
-\end{equation}
-If we assume that: i) $\textbf{x}$ is ``wide sense stationary''~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
-\begin{equation}
-\hat{{s}}[n]=\sum_{k=0}^{n}h[k]x[n-k]\,,
-\end{equation}
-where $\textbf{h}=(h(0),\dots h(n))$ represents the coefficients of an $n$th order Wiener filter.
-% What this sentence is trying to convey is not clear to me
-%% just trying to say that we can make the assumptions that we do because the Wiener filter appears to work well under them - J
-%These assumptions are to be justified by the efficacy of the Wiener filter.
-%Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
-%\begin{equation}
-%    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
-%\end{equation}
-%Assuming that $\textbf{x}$ is wide-sense stationary~\footnote{
-%A random process $\{x(t)\}$ is wide sense stationary (WSS) if for all
-%$t_1,t_2 \in R$, it satisfies (1) mean is time invariant, i.e.,
-%$\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; (2) autocorrelation depends
-%only on time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.}
-%and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
-The famous Wiener-Hopf equation~\citep{noble1959methods} allows us to determine $\textbf{h}$, as\begin{equation}
-\label{eqn:wiener-hopf}
-\begin{bmatrix}  
-r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
-r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
-\vdots&\vdots&\ddots&\vdots\\
-r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
-\end{bmatrix}
-\begin{bmatrix}
-h[0]\\
-h[1]\\
-\vdots\\
-h[n]
-\end{bmatrix}=
-\begin{bmatrix}
-r_{ss}[0]\\
-r_{ss}[1]\\
-\vdots\\
-r_{ss}[n]
-\end{bmatrix}\,,
-\end{equation}
-where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$, 
-%\begin{equation*}
-%\left\{ 
-\begin{eqnarray} 
-r_{xx}[n] &~=~& E[x(i)~x(i+n)] \,,\\
-r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
-%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
-%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
-\end{eqnarray} 
-%\right\}  
-%\end{equation*}
-%where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
-
-%As an instructive aside, i
-If we let the Wiener filter be non-causal (i.e.
-we estimate the current signal based on both past \emph{and future} observations), then we can represent Eqn.~\ref{eqn:wiener-hopf} in the frequency domain as
-\begin{equation}
-\label{eqn:wiener}
-    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}\,,
-\end{equation}
-where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data, the injected signal, and the background noise, respectively.
-Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high and attenuates the signal where the SNR is low.
-The causal Wiener filter (that only makes estimates based on past observations) is similar.
-More detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.
-% That this filter is non-causal makes it unsuitable for real time signal processing
-
-%In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}, and set the length of the filter to $n=100$. 
-%A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
-%The choice of $n=100$ is a reasonable selection to balance smoothing and efficiency.
-
-In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}.
-A higher-order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required.
-For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
-The timeseries and frequency spectrum after applying the Wiener filter are shown in the sixth row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-We see a significant improvement in the timeseries of the recovered signal, however, a strong noise hum persists, audibly. 
-%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
-%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
-%\han{should we add more details here too?} \jam{[seconded]}
-%The time series and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively.
-%The time series is visibly cleaner, however there is a strong noise hum.
-
-\subsubsection{Combined notch and Wiener filter}
-
-Here, we experiment with applying a combination of the cascaded notch and the Wiener filter to the recorded speech signal.
-
-The Wiener filter makes use of statistical information from the speech data and noise.
-It amplifies the part of the signal with high SNR while suppressing the parts with low SNR (see above Sec.~\ref{sec:Wiener}).
-It is implemented in the form of a finite response filter, which ensures linear phase response and stability (both desirable), but at the cost of high orders computationally.
-By comparison, the notch filter is based on directly removing the unwanted frequency components.
-It is implemented in the form of an infinite impulse response filter.
-Although it significantly decreases the order of the overall filter, it unavoidably introduces nonlinear phase and instability.
-
-By combining the notch and Wiener filter, we can trade-off between the two and achieve an overall better performance, as can be seen in the bottom row of Fig.~\ref{fig:BackgroundNotchWienerCombined}.
-The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone.
-The mains noise is all but removed and more voice information is retained.
-However, the recovered voice still sounds muffled and is not understandable.
-
-%The combination of the notch filter and the Wiener filter produces a cleaner signal than either individually. 
-%The time series and spectrum results are shown in the final panels of Figs.~\ref{fig:timeNotchWiener} and ~\ref{fig:freqNotchWiener} respectively.
-
-%The ideal filter would be one that i) completely attenuates the undesired parts of the spectrum, ii) does not change the rest of the spectrum, and iii) has smooth edges so as to not damage the signal after convolution.
-%These three conditions are contradictory and any filter must compromise between them. 
-%The above simple filters fail one or more of i)--iii). 
-%The Butterworth filter performs the best as it is optimised to be `maximally flat' within the band region, prioritising the second condition.
-%\han{see note from Andrew and ask James about this}
-
-%\han{Is it worth also including some plots to show how the different filters perform? E.g. a PSD with comb filter, high pass filter, Butterworth filter?}
-
-% Note that we also replicated the Viterbi algorithm results from before with the photodiode, but only with the mains hum removed and high-passing above 100Hz, else the algorithm never selected the signal.
-Except for being over audible frequencies, there isn’t much difference in the tracking once these filters are applied.
-
-% I just combined these into a single figure and also moved some of the intro text about the signal into fig 8 so that it is clear we are all talking about the same data in these plots - hannah
-\begin{figure*}
-	%\includegraphics[width=0.49\textwidth]{figures/combined_timeseries_melatos_shift_source.pdf}
-	%\includegraphics[width=0.49\textwidth]{figures/combined_spectrum_melatos.pdf}
-    \includegraphics[height=0.57\textheight]{figures/combined_timeseries_melatos_shift_source_tall.pdf}
-    \includegraphics[height=0.57\textheight]{figures/combined_spectrum_melatos_tall.pdf}
-    \caption{\label{fig:logMMSE_timeseries_freqspectrum}
-Timeseries (left column) and frequency spectrum (right columns) results with the logMMSE estimator. 
-The original source signal and optical microphone recording in the top and middle panels respectively are for comparison and are identical to those shown in the top two panels in Fig.~\ref{fig:BackgroundNotchWienerCombined} (again, the source signal is shifted by $0.12\,{\rm s}$).
-The bottom-left panel shows the timeseries of the recording after filtering with the logMMSE estimator, where the rise at the start of the timeseries is expected when filtering a signal of finite duration.
-The bottom-right panel shows the corresponding frequency spectrum with the logMMSE estimator. 
-The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude $2$ which otherwise dominate the plot. 
-The change in scale from the top panel is not of concern as the ear hears the relative frequency content and normalisation of the timeseries fixes any scaling. 
-% my feeling is not to include this sentence here as it is hard to be certain what the noise and what is the input signal for this plot without further investigation - hannah
-%The bottom two panels also show a broad feature from around $750\,{\rm Hz}$ to $1200\,{\rm Hz}$, the origin of which is yet to be determined.
-    }
-\end{figure*}
-
-\subsubsection{logMMSE estimator}
-\label{sec:logmmse}
-
-\begin{comment}
-\begin{figure*}
-	\includegraphics[width=0.8\textwidth]{figures/combined_timeseries_melatos_shift_source.pdf}
-	\caption{Time series of an adult male voice (saying ``A cathode ...'') showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
-The source signal is shifted by $0.12$ s to the right to synchronise the manual delay from starting to play the source through the speaker and starting the recording at the Raspberry Pi.
-One second of the one minute long recording is shown.
-The rise at the start of the bottom panel is expected when filtering a signal of finite duration.}
-	\label{fig:logMMSE_timeseries}
-\end{figure*}
-
-\begin{figure*}
-	\includegraphics[width=0.8\textwidth]{figures/combined_spectrum_melatos.pdf}
-	\caption{Spectrum of optical microphone recording of the same adult male voice in Fig.~\ref{fig:logMMSE_timeseries} showing the source (top panel), the optical microphone recording (middle panel), and the recording after filtering with the logMMSE estimator (bottom panel).
-The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude 2 which otherwise dominate the plot.
-The change in scale from the top panel is not of concern as the ear hears the relative frequency content and normalisation of the time series fixes any scaling.
-The bottom two panels also show a broad feature from around $750\,{\rm Hz}$ to $1200\,{\rm Hz}$, the origin of which is yet to be determined.}
-	\label{fig:logMMSE_spectrum}
-\end{figure*}
-
-%probably a bit too causal language 
-%What we’re attempting here is not new, 
-Speech enhancement of noisy channels is a classic problem in signal processing. 
-% shouldn't we make the citations consistent throughout? We don't mention author names up to this point and typically journals have a set style guide. - Hannah 
-%To borrow from that field, Yi Hu and P.C.
-Loizon~\cite{SubjectiveComparison} compared $13$ speech enhancement methods and found the log minimum means-square error (logMMSE) estimator to be the best among them, qualitatively, at recovering speech.
-In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
-This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that is measures the MSE between the logarithm of the Fourier amplitudes.
-This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
-We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}) to the recorded signal.
-%Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
-%This type of estimator is based on work Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
-
-%In this work we use an existing implementation~\cite{logmmse} of the logMMSE estimator. 
-%Applying this estimator to the optical microphone output produced dramatically cleaner results.
-
-The timeseries and frequency spectrum results of the logMMSE estimator are shown in Fig.~\ref{fig:logMMSE_timeseries_freqspectrum}. 
-The top two rows show the original speech signal and anti-aliased reading from the optical microphone for ease of comparison (identical to the top two rows in Fig.~\ref{fig:BackgroundNotchWienerCombined}).
-The third row shows the timeseries and corresponding spectral amplitude after applying the logMMSE estimator. 
-We see significant attenuation of the mains harmonics as well as a general smoothing of the spectrum. 
-The logMMSE method indeed removes most of the background noise but does not significantly enhance the speech as the voice sounds muffled and indistinct.
-% suggest remove as it's just rephrasing what's already been said
-%, with the words unable to be understood.
-
-We find some improvement with music over speech.
-After applying the logMMSE estimator, simple chords and drums can be heard, however more composite sounds and complex melodies cannot be heard clearly.
-%Our observations suggest that this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins sometimes can’t be heard at all. 
-Our observations suggest that this is especially true for certain instruments, in particular flutes and violins sometimes can’t be heard at all. 
-This could be a perceptual effect or a frequency dependence somewhere in the optical microphone.
-%, although the exact cause is yet to be determined.  - repetition
-Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums are louder in the results.
-To address these problems, we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-%Or if instead they are present in the noisy data but we need better techniques to extract them.
-- should probably justify what you mean by better if you are going to say it here, but I think enough is discussed in section VI for it not to be necessary here - hannah
-To determine this requires a better understanding of the system, as discussed in Section~\ref{sec:future_work}.
-
-%Overall, we find that the intelligibility of speech remains low. 
-%The played-back voice sounds mumbled and indistinct. 
-%The estimator removes most of the background noise, but does not clarify the speech. 
-%To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.
-%This requires a better understanding of the system.
-% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio.
-However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference.
-An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.
-
-\end{comment}
-
 \end{document}

### 2021-04-29 paper/response/AJPRefereeResponse.tex
working on the Viterbi text, updating the schematic, adding referee response text

@@
 \item We have added a sentence to explain Markov process: ``In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal).''
-\item \todo{image colours}
-\item \han{[Hannah] The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals (rather than an improvement in SNR).
-The single constant tone of section III is recovered easily using a Fourier transform, however if the signal frequency changes randomly over time, this will reduce the Fourier transform's ability to identify the signal as it is spread across multiple frequency bins.
-[Added sentence at start of section IV, para 2 to explain that Fourier transform of whole data set is not well suited here]}
+\item We have lightened the colour scheme used in the Fig.
+4 schematic to aid with black and white printing.
+It is also now a similar colour scheme to the spectrogram in Fig.
+5 in order to make it clearer that we are simulating a spectrogram with this schematic.
+When we introduce Fig.
+5 we now also refer back to Fig.
+4 to indicate that the schematic represents a similified version of the results shown in Fig.
+5.
+\item \han{[Hannah] The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals.
+The single constant tone of section III is recovered easily using a Fourier transform, however if the signal frequency changes randomly over time, this will reduce the Fourier transform's ability to identify the signal as the signal power iss spread across multiple frequency bins.
+We have added some text at the start of Section IV (paragraph 2) and at the end of Section IV.
+B.
+to explain that the Fourier transform of the whole dataset is not well suited to this type of signal.
+\item We have also revied the entire Section IV to improve clarity in our Viterbi algorithm explanation and motivation.
 \end{itemize}
@@
 \begin{itemize}
+\item We have including some details on this in the description of detection statistics in section IV.A.
+and have also added that Doppler modulation of the source also needs to be taken into account if searching for a neutron star in a binary.
+\item \todo{...}
 \item A mention has been made in the Future Work as to demonstrating the Doppler effect.
-\item \todo{...}
+ \todo{...}
 \end{itemize}
@@
 \begin{itemize}
-\item \todo{...}
+\item We have clarified this in the text.
+The detection statistic captures the likelihood of a signal being present in each time-frequency bin given the observable data.
+In Section IV.A.
+we have expanded our description og gravitational wave analysis and refered to both the transission matrix and emission matix by name to reduce confusion.
+\todo{...}
 \end{itemize}
@@
 \begin{itemize}
-\item Mention of this is now made both at the introduction of the plot and in the Viterbi algorithm section -- this implementation does assume white noise.
+\item We do assume that the noise is white in our analysis. We have added some text to clarify this at the start of Section IV.C.
 \end{itemize}
@@
 \begin{itemize}
-\item \todo{...}
+\item The Viterbi algorithm is not a new approach for gravitational wave analysis.
+It is used in many searches for continuous waves to date \todo{refs} and the methods papers are Refs \todo{refs}.
+We use these existing methods as inspiration and apply them to this table-top experiment in order to demonstrate the real gravitaitonal wave searches using sound.
+It is important to clarify that we are not searching for chirp signals in this work, we are simulating a signal with wandering frequency as an analoge to continuous wave signals whose frequencies can wander stochastically over time.
+\item [what we have done].
+We have clarified that the method is not new in a new subsection in section IV (section IV A) reviewing how these methods are used in gravitational wave analysis before describing our application of the methods in this work.
+\todo{We also add references to a ......}
 \item \han{[from Hannah] These papers are addressing how to detect an unknown chirp signal from a compact binary merger.

### 2021-04-29 paper/ifo-wanderingTone.tex
going through more comments
working on the Viterbi text, updating the schematic, adding referee response text

@@
 %A different means of detection is needed for a signal which has a randomly wandering frequency.
-The analysis techniques used in this section are inspired by those used to search for continuous waves from spinning neutron stars in LIGO and Virgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
+
+In continuous wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which is able to track the signal's frequency over time~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
+The analysis described and presented in this section is directly inspired by real continuous wave searches~\cite{} 
+In section~\ref{sec:realCWSearches} we breifly review the analysis method used by LIGO and Virgo, in section~\ref{} we describe the methods as applied in this work, and in section~\cite{} we our results of recovering a wandering frequency signal using the table-top interferometer.
+
+\subsection{Continuous wave analysis with real data}
+\label{sec:realCWSearches}
+
+The methods used here are directly inspired by LIGO-Virgo continuous wave searches. 
+For further details on these methods see Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017} and for searches performed using them see Refs.\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application}.
+\han{check for new papers}
+
 Continuous-wave searches are performed on long datasets, months to years in duration. 
@@
 In this context, ``significantly'' means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
-One method to search for a wandering signal is to split the timeseries data into several shorter segments which are analyzed individually.%, and then recombine them under some continuity assumption to coherently analyze the whole set.
-% I think the following is more accurate. The antenna beam pattern is used to compute a detection statistic. Which detection statistic to use depends on the type of target. -Hannah
+%One method to search for a wandering signal is to split the timeseries data into several shorter segments which are analyzed individually.
 
-A ``detection statistic'' is used to quantify the likelihood of a signal being present in the data at each frequency and timeseries segment. 
-In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
-%Most continuous wave searches use detection statistics to quantify the likelihood of a signal being present in the data at each frequency for each timeseries segment. 
-%The detection statistic used depends on the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun, and the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
-The choice of detection statistic used depends on the type of target~\cite{JKS:1998,SuvorovaEtAl:2017}.
-In this work, we take the discrete Fourier transform of each timeseries segment and use its amplitude at each frequency as a detection statistic.
-This maximizes the likelihood of detecting a sinusoidal signal in Gaussian noise, as described in Appendix~\ref{app:sinusoid_likelihood} -- we assume that that the noise is Gaussian.
-The detection statistic is normalized for convenience and we use it to form a grid in time and frequency (a spectrogram) which displays the frequency content of the signal over time. 
-% moved one sentence up 
-%We also normalize the detection statistics for convenience.
-
-Following the method of Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, we use a hidden Markov model to recover the wandering signal from noisy data. 
+In Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, a hidden Markov model is used to search for continuous gravitational waves. 
 In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal). 
 In a hidden Markov model, the frequency state of the signal is unknown, or hidden, and can undergo transitions at discrete times. 
-%To do this, we define the frequency of the signal as a hidden (unknown) state which can undergo transitions at discrete times. 
 The transitions are Markovian in that the hidden state (i.e.\ frequency) of the system at any time depends solely on its state at the previous time.
-The detection statistic (observable) relates the hidden states of the system to the observed data.
+
+A detection statistic relates the observed data to the hidden state and quantifies the likelihood of a signal being present in the data at each frequency and time bin.
+This likelihood is also called the emission matrix in gravitational-wave literature.  
+In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
+If searching for continuous waves from a neutron star in a binary (such as a LMXB), the Doppler modulation of the source also needs to be taken into account and a different detection statistic is used~\cite{SuvorovaEtAl:2017}.
+
+In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over at each time. 
+This is called the transition probability matrix. 
+In LMXB searches, the transition matrix allows the signal frequency to (a) stay in the same frequency bin, (b) moup a single frequency bin, or (c) move down a frequency bin at each time step~\cite{}. 
+In supernova remnanat searches, source frequency is expected to decrease over time, therefore the allowable transitions are to either remain in the same frequency bin or move down one frequency bin~\cite{}.
+
 The Viterbi algorithm~\cite{Viterbi:1967} is used to find the most probable sequence of states given the sequence of observables.
-%To perform the task of recovering the signal, we use the Viterbi algorithm~\cite{Viterbi:1967} to efficiently find the most probable path through the spectrogram given the sequence of observations.
-This method is used in continuous-wave searches for a variety of astrophysical targets~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application} and forms the basis of the analysis in this section.
-
-In Section~\ref{sec:viterbi} we review the Viterbi algorithm and its application to this work. 
-In Section~\ref{sec:wanderingResults} we present results for recovery of a wandering frequency audio signal.
+In the next section we describe our application of the Viterbi method and our choice of detection statistic.
 
@@
 
-We use the schematic in Fig.~\ref{fig:viterbi} to aid our explanation of the method. 
+\begin{figure}
+\includegraphics[width=0.49\textwidth]{figures/viterbiDiagramBlue.pdf}
+\caption{\label{fig:viterbi}
+Schematic diagram of the Viterbi algorithm. 
+The circles represent elements in a time-frequency grid, labelled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
+The color of each circle represents the likelihood of a signal being present in each time-frequency element.
+The dark blue to light green gradient corresponds to low to high likelihood values in arbitrary units for the purpose of this schematic. 
+The objective is to find the most probable path of a signal through the grid from left to right.
+All possible paths are shown by the lines. 
+The black lines show the best path to each circle and gray lines show rejected paths. 
+Some routes though the grid result in dead ends with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a).
+At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:pathprobability}, marked (b). 
+The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
+%A schematic diagram of the Viterbi algorithm. 
+%Each circle represents an element in the time-frequency grid starting from the instant $t_0$ on the left and ending at $t_{N_t}$ on the right. 
+%Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
+%The color of each circle indicates the likelihood that a signal is present in the box (the detection statistic), where a lighter color corresponds to a higher likelihood. 
+%At each time step, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines connecting the circles in the diagram). 
+%Each circle has three paths leading to it, indicated by the lines. 
+%The best path to each circle at each $t_i$ is highlighted in black. 
+%Some routes through the grid are found to be dead ends, with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a). 
+%At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:recursiveViterbi}, marked (b). 
+%The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
+}
+\end{figure}
+
+The timeseries data from the interferometer is first split into segments. 
+We take the discrete Fourier transform of each timeseries segment and use the normalised Fourier amplitude $F(t_i,f_j)$ as the detection statistic (this is the emission matrix).
+Assuming Gaussian noise, this detection statistic choice maximises the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.
+The detection statistic is normalized for convenience and we use it to form a grid in time and frequency (a spectrogram). \han{add normalization - ask James}
+
+%The Viterbi algorithm is then applied to the spectrogram data. 
+We use the schematic in Fig.~\ref{fig:viterbi} to aid our explanation of the method throughout this section. 
 It represents a spectrogram with $N_t$ time bins and $N_f$ frequency bins.
-We label the elements of the time and frequency  grid as $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-We use the normalized Fourier amplitude $F(t_i,f_j)$ as the detection statistic, which is represented by the shading of each grid point (lighter corresponds to higher value).
+The circles represent the elements of the time and frequency grid, which are labelled as  
+$t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
+The color of the circles represents the detection statistic $F(t_i,f_j)$ at each grid point where dark blue corresponds to lower values and light green to higher values.
 
@@
 The objective is to find the most likely path through the grid given the observed data and any probabilistic constraints on how the frequency of the signal can wander from $t_i$ to $t_{i+1}$. 
-In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over at each time. 
-This information is encoded in the transition probability matrix $A(f_k,f_m)$, which describes the probability of system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
-Here, we allow the frequency of the signal to either (a) stay in the same bin; (b) move up by a single frequency bin; or (c) move down by a single frequency bin at each transition. 
-The possible transitions are shown as lines in Fig.~\ref{fig:viterbi}.
+The transition probability matrix, $A(f_k,f_m)$, describes the probability of system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
+Here, we allow the frequency of the signal to either (a) stay in the same bin; (b) move up by a single frequency bin; or (c) move down by a single frequency bin at each transition A 
 We assign these three transitions equal probability, i.e.\ $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
+All possible transitions are shown as lines in Fig.~\ref{fig:viterbi}.
+The different colorings of the lines are explained below.
 
+The signal is equally likely to start in any frequency bin (i.e. the prior is flat). 
 We define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
 A specific path (which may not be optimal) is written as $f(t_0), f(t_1),..., f(t_n)$. 
-The probability of this specific path given the data is 
+The probability of a specific path given the data is 
 \begin{eqnarray}
-{\rm Pr}[ f(t_0), f(t_1),... f(t_n) | {\rm data}] =
-          F[t_n,f(t_n)] A[f(t_n),f(t_n-1)]
-          \times... A[f(t_1),f(t_0)] F[t_0,f(t_0)]\,,
-\label{eqn:pathprob}
+{\rm Pr}[ f(t_0), f(t_1),... && f(t_n) | {\rm data}] = \nonumber\\
+          && F[t_n,f(t_n)] A[f(t_n),f(t_n-1)] \nonumber \\
+          && \times... A[f(t_1),f(t_0)] F[t_0,f(t_0)] \,,
+\label{eqn:pathprobability}
 \end{eqnarray}
 The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises $Pr[ f(t_0), f(t_1),... f(t_n) | data]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
-We note that there is no marginalization used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:probpath} are both evaluated for a specific path and then we mazimise over all such paths to find the optimal Viterbi path.
+We note that there is no marginalization used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we mazimise over all such paths to find the optimal Viterbi path.
 %The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
@@
 \begin{enumerate}
-\item Starting at time $t_1$, each $f_j$ state can originate from three prior states at time $t_0$ (except for $f_0$ and $f_{N_f}$ which only have two).
+\item Starting at time $t_1$, each $f_j$ state can originate from three prior states at time $t_0$ (except for the edge cases $f_0$ and $f_{N_f}$ which only have two). 
 The paths between these states are indicated by the lines in Fig.~\ref{fig:viterbi}. 
@@
 These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the gray lines show the rejected paths. 
-For example, the most probable connection to the element labeled (a) is the one directly behind it (i.e., $f_4$).
+For example, the most probable connection to the element labelled (a) is the one directly behind it (i.e., $f_4$). 
 Therefore, this path is selected as the best path from $t_0$ to $t_1$ for $f_4$.
@@
 
-\item Moving to time $t_2$, again, we select the path which maximizes the recursive quantity in Eqn.~\ref{eqn:recursiveViterbi} for each $f_j$.
-These are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.
+\item Moving to time $t_2$, again, we select the path which maximizes the recursive quantity in Eqn.~\ref{eqn:pathprobability} for each $f_j$. 
+These paths are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.
+Rejected paths are again shown by gray lines.
 
@@
 
-\item Each state $f_j$ at $t=t_{N_t}$ has some value ${\rm Pr}[f(t_{N_t})=f_j]$ (Eqn.~\ref{eqn:recursiveViterbi}), which represents the probability of the most likely path that ends in state $f_j$ at $t_{N_t}$.
-The Viterbi algorithm then selects the terminating frequency bin $f(t_{N_t})$ with the highest probability, marked by (b) in Fig.~\ref{fig:viterbi}.
+\item We have now found the most probable path to each $f_j$ at $t=t_{N_t}$ and its probability (Eqn.~\ref{eqn:pathprobability}). 
+We select the terminating frequency bin $f(t_{N_t})$ with the highest probability labelled as (b) in Fig.~\ref{fig:viterbi}.
 
-\item The last step is to find the Viterbi path (the overall best path that terminates in the frequency bin with the highest probability in step 4).
+\item The final step is to find the Viterbi path (the overall best path that terminates in the frequency bin with the highest probability in step 4). 
 The Viterbi path is found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}). 
-In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c) and is highlighted in orange.
+In Fig.~\ref{fig:viterbi}, it is the path ending at (b) that started at (c) highlighted in orange.
 %\item The Viterbi path (the overall best path) is then found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}).
@@
 
-The key to the Viterbi algorithm is that we can use the signal's correlation in time to improve the signal-to-noise ratio of the frequency of the wandering signal at a particular time.
+In continuous wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency could be changing unpredictably over time.
+A Fourier transform of the entire dataset as described in Section~\ref{sec:single_tone} would split the signal power between multiple frequency bins and could result it in being undetectable under the noise.
+The Viterbi algorithm's strength lies in its ability to track such signals through the data.
+In the following section we present the results of using the Viterbi algorithm with the table-top interferometer data. 
+% - just simiplifed this discussion a little and added a reference to the CW searches
+%The key to the Viterbi algorithm is that we can use the signal's correlation in time to improve the signal-to-noise ratio of the frequency of the wandering signal at a particular time.
 Since the noise is assumed to not be correlated in time, if the detection statistic at a given frequency is high at a particular time and at neighboring frequencies at neighboring times then the likelihood that the given frequency at that particular time is from the wandering signal increases.
 By iterating this process through the grid for a suitable choice of the window of neighboring frequencies and times, the Viterbi algorithm is well-suited to finding wandering signals.
-In the following section, we use the Viterbi algorithm to recover a slowly wandering signal.
 
@@
 
-\begin{figure}
-\includegraphics[width=0.49\textwidth]{figures/viterbiDiagram.pdf}
-\caption{\label{fig:viterbi}
-A schematic diagram of the Viterbi algorithm. 
-Each circle represents an element in the time-frequency grid starting from the instant $t_0$ on the left and ending at $t_{N_t}$ on the right. 
-Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
-The color of each circle indicates the likelihood that a signal is present in the box, where a lighter color corresponds to a higher likelihood. 
-At each time, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines connecting the circles in the diagram). 
-Each circle has three paths leading to it, indicated by the lines. 
-The best path to each circle at each $t_i$ is highlighted in black. 
-Some routes through the grid are found to be dead ends, with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a). 
-At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:recursiveViterbi}, marked (b). 
-The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
-}
-\end{figure}
-
-\subsection{Wandering frequency results}
+\subsection{Wandering frequency signal results}
 \label{sec:wanderingResults}
@@
 Recovery of a wandering tone using the Viterbi algorithm.
-The spectrogram (heatmap) colouring indicates the value of the detection statistic (here, the logarithm of the absolute value of the discrete Fourier transform) in each time-frequency bin -- with brighter colours indicating higher values.
+The spectrogram (heatmap) coloring indicates the value of the detection statistic (here, the logarithm of the absolute value of the discrete Fourier transform) in each time-frequency bin -- with brighter colours indicating higher values.
 The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path, respectively. 
@@
 
-To simulate a slowly wandering signal we modulate the frequency sinusoidally with a modulation amplitude that decays with time. 
+We simulate a slowly wandering signal by modulating the frequency sinusoidally with a modulation amplitude that decays with time. 
 We use the same set-up as shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
@@
 Note that we implicitly approximate the noise at the webcam as white (uniform in frequency) in this implementation of the Viterbi algorithm. 
-This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum} but is a sound approximation for the higher-frequency (around 5~Hz) wandering signals that we study.
+This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum} but is a resonable approximation for the higher-frequency (around 5~Hz) wandering frequency signals that we use.
 
-The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal. 
-For this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous-wave signals to have far lower signal-to-noise ratios. 
+The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal (similar to that represented by the schematic in Fig.~\ref{fig:viterbi}). 
+In this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous-wave signals to have far lower signal-to-noise ratios. 
 The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.
@@
 \begin{equation}
-E_{\rm rms} = \left[ \frac{1}{N_t+1} \sum_{i=0}^{N_t} \frac{(I_i - R_i)^2}{{I_i}^2} \right] ^{1/2}
+E_{\rm rms} = \left[ \frac{1}{N_t+1} \sum_{i=0}^{N_t} \frac{(I_i - R_i)^2}{{I_i}^2} \right] ^{1/2}\,,
 \end{equation}
-where $I_i$ and $R_i$ are the injected and recovered (Viterbi) frequency paths.
-Computing this for the result shown in Fig.~\ref{fig:viterbi_overlay}, we find that $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
+where $I_i$ and $R_i$ are the injected and recovered (Viterbi) frequency paths, respectively. 
+For the result shown in Fig.~\ref{fig:viterbi_overlay}, we find that $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
 

### 2021-04-29 paper/response/AJPRefereeResponse.tex
working on the referee response

@@
 
-\subsection*{Editors comments (Joseph Amato)}
+\subsection*{Comments from editor, Joseph Amato}
 
@@
 \begin{itemize}
-\item Done: we have moved the footnotes to the bibliography
+\item We have moved the footnotes to the bibliography
 \item \todo{Work in progress, please correct any that have not been fixed yet}
@@
 
-\noindent 3.Section IV.
+\noindent 3.
+Section IV.
 The term “Markov process” may not be familiar to readers.
@@
 4 schematic to aid with black and white printing.
-It is also now a similar colour scheme to the spectrogram in Fig.
-5 in order to make it clearer that we are simulating a spectrogram with this schematic.
-When we introduce Fig.
-5 we now also refer back to Fig.
-4 to indicate that the schematic represents a similified version of the results shown in Fig.
+Our intention is that the Fig.
+4 schematic be interpreted as a simplified version of the Fig.
+5 spectrogram result.
+To clarify this we have made the colour scheme of the schmatic more consistent with that of Fig.
+5 and we have refered back to the schematic when we first introduce Fig.
 5.
-\item \han{[Hannah] The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals.
-The single constant tone of section III is recovered easily using a Fourier transform, however if the signal frequency changes randomly over time, this will reduce the Fourier transform's ability to identify the signal as the signal power iss spread across multiple frequency bins.
+\item The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals.
+The single constant tone in section III is recovered easily using a Fourier transform, however if the signal frequency changes randomly over time, this will reduce the Fourier transform's ability to identify the signal as the signal power is spread across multiple frequency bins.
 We have added some text at the start of Section IV (paragraph 2) and at the end of Section IV.
@@
 to explain that the Fourier transform of the whole dataset is not well suited to this type of signal.
-\item We have also revied the entire Section IV to improve clarity in our Viterbi algorithm explanation and motivation.
+\item We have also revised the entire Section IV to improve clarity in our Viterbi algorithm explanation and motivation.
 \end{itemize}
@@
 \begin{itemize}
-\item An explanation of the color coding has been added to the caption.
-\item \todo{compare to without VA}
-\item \todo{compare to LIGO literature}
+\item The color coding represents the logarithm of the absolute value of the discrete Fourier transform. We have clarified this in the caption.
+\item \todo{compare to without VA?}
+\item The heatmap is similar to the figures in LIGO-Virgo-KAGRA publications which show transient gravitational-wave signals, but not identical.
+Those figures normally show Q-tranforms [described in Chatterji, et al., Classical Quantum Gravity 21, S1809 (2004)] whereas we show the Fourier transform in this work.
 \end{itemize}
@@
 \item The Butterworth filter has strongly attenuated the signal outside the passband.
-\item \todo{They will also expect a change to the paper as it is a point of confusing - e.g. clarify the reason in the figure caption.}
+\item \todo{They will also expect a change to the paper as it is a point of confusion - e.g. clarify the reason in the figure caption.}
 \end{itemize}
@@
 \begin{itemize}
-\item Agreed, we have moved the majority of this section to Supplementary Matrial
-\item \todo{decide what we will show in main text}
-\item \todo{expansion on compromises discussion}
+\item Agreed, we have moved the majority of this section to Supplementary Material.
+We have chosen to show two highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator), but we refer the reader to the Supplementary Material for the details of these methods.
+We have also expanded the discussion on filters more generally in the main text.
 \end{itemize}
@@
 
-\noindent So you see that I’m asking for a major revision of your manuscript.
+\noindent So you see that I'm asking for a major revision of your manuscript.
 I hope this does not discourage you.
@@
 \begin{itemize}
-\item We have clarified this in the text.
+\item We have clarified the explanation of the HMM in the text.
 The detection statistic captures the likelihood of a signal being present in each time-frequency bin given the observable data.
 In Section IV.A.
-we have expanded our description og gravitational wave analysis and refered to both the transission matrix and emission matix by name to reduce confusion.
-\todo{...}
+we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission matrix by name to reduce confusion.
 \end{itemize}
@@
 \begin{itemize}
-\item We have move the majority of Section VC to Supplementary Material and now only include highlight results in the main text (the combined notch and Wiener filter and the logMMSE estimator).
+\item We have move the majority of Section VC to Supplementary Material and now only include highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator).
 In the Supplementary Material we have expanded the description of the impulse response of a filter mentioning both IIR and FIR.
-We have also added some text saying that Eq.
-2 in the Supplementary Material (previously Eq.
+We have also added some text saying that Eqn.
+2 in the Supplementary Material (previously Eqn.
 4 in the main text) is a Z-tranform.
-We have corrected the inconsistency in Eq.
-1 in the Supplementary Material (previously Eq.
+We have corrected the inconsistency in Eqn.
+1 in the Supplementary Material (previously Eqn.
 3 in the main text).

### 2021-04-29 paper/response/AJPRefereeResponse.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 \newcommand{\han}{\textcolor{orange}}
+\newcommand{\jam}{\textcolor{magenta}}
 
@@
 
-\todo{Opening comments}
+\todo{Opening comments}\\
+
 Dear Joseph Amato,
@@
 \subsection*{Comments from editor, Joseph Amato}
-
+\begin{quote}
 This is an addendum to my decision letter and the two reviews of your manuscript.
@@
 Here are some of my concerns.
-\\
+\end{quote}
 
@@
 \item The color coding represents the logarithm of the absolute value of the discrete Fourier transform. We have clarified this in the caption.
-\item \todo{compare to without VA?}
+\item Without using the VA the signal becomes spread out over the frequency bins and might not rise above the noise.
+We have now explained this and added a plot that shows the Fourier spectrum of the entire timeseries.
 \item The heatmap is similar to the figures in LIGO-Virgo-KAGRA publications which show transient gravitational-wave signals, but not identical.
@@
 \begin{itemize}
-\item The Butterworth filter has strongly attenuated the signal outside the passband.
-\item \todo{They will also expect a change to the paper as it is a point of confusion - e.g. clarify the reason in the figure caption.}
+\item The Butterworth filter has strongly attenuated the signal outside the passband, this sharp drop-off at low frequencies would appear shallower when shown against a logarithmic scale for frequency as is sometimes done.
+A note has been made in the Supplementary Material where this plot has been moved to.
+% (this sharp drop-off would appear shallower when shown against a logarithmic scale for frequency)
+% \item \todo{They will also expect a change to the paper as it is a point of confusion - e.g. clarify the reason in the figure caption.}
 \end{itemize}
@@
 
+\begin{quote}
 The paper describes a table top interferometer for demonstrating the basic operating principles of gravitational wave (GW) detectors.
@@
 The paper is well written and will be of interest to educators who wish to provide some hands-on training to students in GW astronomy.
-I have some comments below that the authors should address.\\
+I have some comments below that the authors should address.
+\end{quote}
 
@@
 \begin{itemize}
-\item \todo{...}
+\item This has now been done. Each value is divided by the maximum value in the grid such that $\max_{i,j} F(t_i, f_j) = 1$.
 \end{itemize}
@@
 \begin{itemize}
-\item Thank you, we have corrected the axis labels and they are now consistent.
+\item We have corrected the axis labels and they are now consistent.
 \end{itemize}
@@
 
+\begin{quote}
 This is a very interesting paper written in a clear and informative language.
 However, I don’t think that it suits the style of papers in AJP because it represents a novel idea of teaching and exploring gravitational waves with audio signals.
-After reading and inspection of the manuscript I came to the  conclusion  that it is a  research  paper  which should pass  through critical discussion  with researchers  from gravitational  wave  community who  should  confirm  that  the audio signals produced in the way proposed in the manuscript,can indeed mimic real gravitational wave signalswithout confusionof students and teachers.
+After reading and inspection of the manuscript I came to the  conclusion  that it is a  research  paper  which should pass  through critical discussion  with researchers  from gravitational  wave  community who  should  confirm  that  the audio signals produced in the way proposed in the manuscript, can indeed mimic real gravitational wave signals without confusion of students and teachers.
 Gravitational waves are transverse-traceless while the sound waves are purely longitudinal.
 In this respect there is no analogy between gravitational waves and audio signals whatsoever.
-At the same time,the principles of extraction of signal from noise andthe filtering techniqueworked out in the manuscriptare applicable in a wide range of physical disciplines andmay be useful for the gravitational wave communityas well.
-On the other hand, I cannot imagine how teachers of physics will explain the complexity of the Winer filter, the Markov chain, the Viterbi algorithm to high school students or even to undergraduatestudentsin university college.
+At the same time, the principles of extraction of signal from noise and the filtering technique worked out in the manuscript are applicable in a wide range of physical disciplines and may be useful for the gravitational wave community as well.
+On the other hand, I cannot imagine how teachers of physics will explain the complexity of the Wiener filter, the Markov chain, the Viterbi algorithm to high school students or even to undergraduate students in university college.
 Signal processing technique is based  on a  high  level  mathematics (statistics, probability theory, stochastic processes, etc.) and engineering.
-In my opinion, the paper is too complicated both technically and mathematicallyfor publishing in AJP.
-I am convinced that this is a very good and high-quality research  paper but  my  fair  opinion  is  thatthe authorsshould submit  it  to  Physical Review D or another peer-review physics journal.
+In my opinion, the paper is too complicated both technically and mathematically for publishing in AJP.
+I am convinced that this is a very good and high-quality research  paper but  my  fair  opinion  is  that the authors should submit  it  to  Physical Review D or another peer-review physics journal.
+\end{quote}
+
 \begin{itemize}
 \item \todo{...}
-\end{itemize}
-
-\han{> Thoughts on this [Hannah]: The use of audio as an analogy to gravitational waves is not novel to this work.
+\item \han{> Thoughts on this [Hannah]: The use of audio as an analogy to gravitational waves is not novel to this work.
 Sound has been used in numerous situations to convey the concept of gravitational wave signals to general audiences [will add references].
@@
 [will think about this some more and very happy to hear other thoughts on how we respond here -Hannah]}
+\item \jam{(1) The audio analogy is not novel, (2) the interferometer is a toy model for demonstration and teaching not for research, i.e.\ I doubt our analysis techniques are useful to the gravitational wave community in any novel way, (3) as long as we are clear in the text (todo added below) about the difference and that we are demonstrating the analysis techniques then I think that there isn't an issue, (4) we have heavily edited the signal processing section and worked on the Viterbi algorithm section which I believe (the Viterbi algorithm) is understandable to even high-school students if the explanation without using Markov models or Fourier transforms is given, i.e.
+the pseudocode.
+- James}
+\item \todo{Clearly state the differences between gravitational waves and sound waves in the text but reiterate that despite these differences our demonstration is still useful to explain signal analysis techniques used for gravitational waves.}
+\end{itemize}
 

### 2021-04-29 paper/ifo-appendix.tex
fixing plots - part 1
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 The current build and sample data can be found at:
-\textbf{[Removed github web link for double-anonymous review.]}
+\textbf{[Removed GitHub web-link for double-anonymous review.]}
 %\url{https://github.com/daccordeon/gravexplain}
@@
 \end{eqnarray}
-Focusing on the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, baring the constant factor $\ip{x}{x}$.
+Focusing on the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log-likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, baring the constant factor $\ip{x}{x}$.
 
@@
 More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}.
-We use Fourier amplitudes normalized between $(0, 1)$ as multiplicative weights, but to avoid numerical underflow we take the logarithm and consider them as additive weights.
+We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights, but to avoid numerical underflow we take the logarithm and consider them as additive weights.
 
@@
 \begin{figure*}
-	\includegraphics[width=.65\textwidth]{figures/circuit_diagram_2.pdf}
+	\includegraphics[width=.8\textwidth]{figures/circuit_diagram_2.pdf}
 	\caption{\label{fig:circuit_diagram}

### 2021-04-29 paper/ifo-complexAudio.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
-A selection of filters are applied to the data, details of which are presented in Supplementary Material. 
-We present the two best performing analyses from our Supplementary Material in Section~\ref{sec:opticalMicResults}.
+A selection of filters is applied to the data, details of which are presented in Supplementary Material. 
+We present the two best-performing analyses from our Supplementary Material in Section~\ref{sec:opticalMicResults}.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/combined_highlight_results_melatos_labelled.pdf}
+\includegraphics[width=0.8\textwidth]{figures/combined_highlight_results_melatos_labelled.pdf}
 \caption{\label{fig:notchWienerLogMMSEResults}
@@
 The original input signal in the first row is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode''). 
- The input signal is shifted by 0.12 s to the right to synchronise the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker. 
+ The input signal is shifted by 0.12 s to the right to synchronize the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker. 
 The second row shows the raw output from the optical microphone when the input from the first row is played. 
@@
 
-In this section we present results of two advanced signal processing analysis techniques applied to the optical microphone recordings. 
+In this section, we present results of two advanced signal processing analysis techniques applied to the optical microphone recordings. 
 The techniques are only briefly described here and we refer the reader to the accompanying Supplementary Material for further details and some other analysis techniques.
@@
 A notch filter removes signals within a specific frequency range. 
-In this case, our aim is to remove the $50\,{\rm Hz}$ mains noise and harmonics, therefore we use a cascaded notch filter where each notch is centered at each of the harmonics. 
+We want to remove the $50\,{\rm Hz}$ mains noise and harmonics, therefore we use a cascaded notch filter where each notch is centered at each of the harmonics. 
 The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. 

### 2021-04-29 paper/ifo-conclusions.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 We find that the logMMSE produces the best results, however, it does not enhance the recording to an intelligible level. 
-The complex audio signals can also serve as more distant analog to un-modeled gravitational-wave burst signals, e.g. from supernovae.
+The complex audio signals can also serve as a more distant analog to un-modeled gravitational-wave burst signals, e.g. from supernovae.
 
@@
 The increased excitement and public interest in the field of gravitational-wave research in recent years mean that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
-The field of gravitational-wave astrophysics continuous to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.
+The field of gravitational-wave astrophysics continues to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.
 % \jam{``charming'' is very sweet}     - agreed I wasn't sure at first but it has grown on me a lot! :) - Hannah

### 2021-04-29 paper/ifo-futureWork.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 Flowchart for the signal through the optical microphone. The transfer function would have to account for each stage. 
-The stage inside the red-dash box is analyzed in Appendix~\ref{app:intensity_derivation}.
+The stage inside the blue-dashed box is analyzed in Appendix~\ref{app:intensity_derivation}.
 Understanding the other stages is one aim of future work.
@@
 This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
-Another extension would be to demonstrate the Doppler effect and the continuous-wave techniques used in the literature to address it by moving the optical table on some kind of circular track with frequency far higher than the wandering signal and placing a distant signal source at different positions above or beside the track -- or perhaps more practically the source could move instead of the detector.
+Another extension would be to demonstrate the Doppler effect and the continuous-wave techniques used in the literature to address it by moving the optical table on some kind of circular track with period far shorter than the time it takes the signal to significantly wander in frequency.
+Then, by placing a distant signal source at different positions above or beside the track -- or perhaps more practically the source could move instead of the detector, an analogous Doppler effect could be observed.
 

### 2021-04-29 paper/ifo-introduction.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
-Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}, 
+Gravitational waves are a prediction of Albert Einstein's theory of General Relativity~\cite{Einstein:1916}, 
 They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
@@
 % we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
-The rotation frequency of the neutron star can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}. 
+The rotation frequency of the neutron star in an LMXB can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}.
 Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
@@
 In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
-We then extend its use to an ``optical microphone'' (using light to capture sound) and present a range of example analysis techniques for educators to use. 
+We then extend its use to an ``optical microphone'', using light to capture sound, and present a range of example analysis techniques for educators to use. 
 % re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
@@
 In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
-In Section~\ref{sec:viterbi_wandering}, we replace the single tone with a wandering frequency signal and analyze the signal using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches.
+In Section~\ref{sec:viterbi_wandering}, we observe a wandering frequency signal and analyze it using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches. 
 In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
@@
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
-We collect the software and scripts used to produce this work in Appendix~\ref{app:code}.
+We present the software and scripts used to produce this work in Appendix~\ref{app:code}.
 

### 2021-04-29 paper/ifo-supplementary-material.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 In Section~V in the main article, we use the interferometer as an optical microphone; audio signals are used to vibrate the interferometer mirrors, the changing interference pattern is recorded, and we aim to recover the original signal from the resulting timeseries data.
-In this Supplementary Material we include a selection of signal processing techniques applied to the data which may be considered for undergraduate laboratory examples. 
+In this Supplementary Material, we include a selection of signal processing techniques applied to the data which may be considered for undergraduate laboratory examples. 
 We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
@@
     The seventh shows the results of applying both the cascaded-notch and Wiener filter (identical to the third row in Fig.~7 in the main article).
-    The eight row shows the results of applying the logMMSE estimator (identical to the fourth row in Fig.~7 in the main article). 
+    The eighth row shows the results of applying the logMMSE estimator (identical to the fourth row in Fig.~7 in the main article). 
 	}
@@
 
-An inituitive way to remove noise is by directly subtracting the background noise from the recorded spectrum. 
+An intuitive way to remove noise is by directly subtracting the background noise from the recorded spectrum. 
 The third row in Fig.~\ref{fig:op_mic_results} shows the interferometer background noise when there is no input signal.
-The fourth row shows the spectrum obtained after subtracting the background noise spectrum, i.e. the result of subtracting the third row from the second row. 
+The fourth row shows the spectrum obtained after subtracting the background noise spectrum, i.e.\ the result of subtracting the third row from the second row. 
 We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified.
@@
 Using a high-pass filter with a higher cut-off can be used to mitigate this issue.
-However, it makes the played-back signal unrecognisable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
+However, it makes the played-back signal unrecognizable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
 Often in speech processing, the logarithm of the signal is taken since the amplitude information seems to be more important to intelligibility than the phase information to the human ear~\cite{SubjectiveComparison}.
@@
 We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
-After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
+After filtering we see significant attenuation (at least 3~dB) of all frequencies outside the band, including around 100~dB near 0~Hz due to even stronger attenuation far from the band (this sharp drop-off at low frequencies would be more similar to the high-frequency (3~kHz) behaviour if shown against a logarithmic scale for frequency as is sometimes done).
+However, the filter has little change on the harmonics within the band.
 }
@@
 
-One method to remove mains noise and its harmonics is to use a sequence of notch filters centred at each of the frequency bins we want to remove.
+One method to remove mains noise and its harmonics is to use a sequence of notch filters centered at each of the frequency bins we want to remove.
 This sequence is known as a ``cascade'' of notches.
@@
 The impulse response describes the reaction of a system to an input signal as a function of time. 
-If a system is linear and time-invarient, the output signals are completely determined for any input signal using the impulse response. 
-Finite inpulse response (FIRs) filters have finite length (they equal zero outside of a finite range). 
-By contrast an infinite impulse response (IIR) filter has infinite length due to feedback. 
+If a system is linear and time-invariant, the output signals are completely determined for any input signal using the impulse response. 
+Finite impulse response (FIRs) filters have finite length (they equal zero outside of a finite range). 
+In contrast, an infinite impulse response (IIR) filter has infinite length due to feedback. 
 Typically FIR filters outperform IIR filters as they are always stable. 
@@
 
-In this work we opt for a IIR notch filter. 
+In this work, we opt for an IIR notch filter. 
 We write the filter in the form of a Z-transform, in which a discrete-time signal is converted to the complex frequency domain $z=e^{jf}$. 
@@
 In Eqn.~\ref{eqn:notch}, $w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
-We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centred on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$. 
+We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centered on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$. 
 We choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
@@
 The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row in Fig.~\ref{fig:op_mic_results}.
-We see that the mains harmonics are significantly attenuated in comparison background subtracted results shown in the fourth row in Fig.~\ref{fig:op_mic_results}.
+We see that the mains harmonics are significantly attenuated compared to the background subtracted results shown in the fourth row in Fig.~\ref{fig:op_mic_results}.
 
@@
 Qualitatively it sounds more like a drum than a human voice.
-This is due to loss of voice information under the filter.
+This is due to the loss of voice information under the filter.
 To overcome this, we turn to more advanced techniques, starting with the Wiener filter.
-Instead of just passively filtering different frequencies, this statistical technique optimises an estimate of the original signal.
+Instead of just passively filtering different frequencies, this statistical technique optimizes an estimate of the original signal.
 
@@
 The causal Wiener filter (that only makes estimates based on past observations) is similar.
-More detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.
+A more detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.
 
@@
 In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
-This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimises the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that it measures the MSE between the logarithm of the Fourier amplitudes.
+This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimizes the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that it measures the MSE between the logarithm of the Fourier amplitudes.
 This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
@@
 The final row in Fig.~\ref{fig:op_mic_results} shows the results of the logMMSE estimator.
-Again there is significant noise reduction, however the speech remains indistict.
+Again there is significant noise reduction, however, the speech remains indistinct.
 The logMMSE results are also presented in Section V.C. in the main text.

### 2021-04-29 paper/ifo-tableTopGWs.tex
interference pattern image slightly larger
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 
-Sound is a commonly-used analogy when explaining gravitational-wave science. 
+Sound is a commonly used analogy when explaining gravitational-wave science. 
 Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound.
@@
  \begin{center}
-  \includegraphics[height=0.3\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
+  \includegraphics[width=0.35\textwidth, angle=-90]{figures/webcam_still0_star.pdf}
  \end{center}

### 2021-04-29 paper/ifo-wanderingTone.tex
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 - hannah
+% Hmm, not sure where I got "(but not always)" from but if I didn't note it down then then I have no idea why now.
+Re: long time sensing, I was trying to justify that frequencies can wander but I prefer your rewrite.
+- James
 
@@
 
+\begin{figure}
+	\includegraphics[width=.45\textwidth]{figures/viterbi_comparison_entire_spectrum_viterbi_test_webcam.pdf}
+	\caption{\label{fig:viteri_comparison_webcam_spectrum}
+Why we use the Viterbi algorithm to recover wandering frequency signals: Here we plot the Fourier amplitude of the intensity pattern against frequency as in Fig.~\ref{fig:webcam_spectrum} for the entire timeseries of the wandering frequency signal in Fig.~\ref{fig:viterbi_overlay}.
+Instead of resolving definite peaks like the constant frequency signal shown in Fig.~\ref{fig:webcam_spectrum}, we only see inconclusive peaks at 5.7~Hz and 10.7~Hz that correspond to the long amount of time that the wandering frequency spends near those values and/or its amplitude there, as seen in the pink dots in Fig.~\ref{fig:viterbi_overlay}.
+With the signal spread out over the frequency bins, we are unable to reliably detect it above the noise and therefore we use the Viterbi algorithm instead.
+The large values at low frequencies below 0.1~Hz are similar to those in Fig.~\ref{fig:webcam_spectrum} and can be ignored as they are likely inaccurate due to the short length of the signal relative to their period (i.e.\ 300~s only contains 30~periods of a 0.1~Hz signal), but this feature would require further research to fully determine its origin.
+%maybe could be from the long period (100~s) variation?
+}
+\end{figure}
+
 Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational waves may not be monochromatic.
@@
 
-A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case, as the signal can be spread across multiple frequency bins making it challenging or perhaps impossible to detect.
+A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case, as the signal can be spread across multiple frequency bins making it challenging or perhaps impossible to detect, as shown in Fig.~\ref{fig:viteri_comparison_webcam_spectrum}.
 %A different means of detection is needed for a signal which has a randomly wandering frequency.
 
-In continuous wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which is able to track the signal's frequency over time~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
-The analysis described and presented in this section is directly inspired by real continuous wave searches~\cite{} 
-In section~\ref{sec:realCWSearches} we breifly review the analysis method used by LIGO and Virgo, in section~\ref{} we describe the methods as applied in this work, and in section~\cite{} we our results of recovering a wandering frequency signal using the table-top interferometer.
+In continuous wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
+The analysis described and presented in this section is directly inspired by real continuous-wave searches~\cite{} 
+In section~\ref{sec:realCWSearches} we briefly review the analysis method used by LIGO and Virgo, in section~\ref{} we describe the methods as applied in this work, and in section~\cite{} we our results of recovering a wandering frequency signal using the table-top interferometer.
 
@@
 In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
-If searching for continuous waves from a neutron star in a binary (such as a LMXB), the Doppler modulation of the source also needs to be taken into account and a different detection statistic is used~\cite{SuvorovaEtAl:2017}.
+If searching for continuous waves from a neutron star in a binary (such as an LMXB), the Doppler modulation of the source also needs to be taken into account and a different detection statistic is used~\cite{SuvorovaEtAl:2017}.
 
@@
 This is called the transition probability matrix. 
-In LMXB searches, the transition matrix allows the signal frequency to (a) stay in the same frequency bin, (b) moup a single frequency bin, or (c) move down a frequency bin at each time step~\cite{}. 
-In supernova remnanat searches, source frequency is expected to decrease over time, therefore the allowable transitions are to either remain in the same frequency bin or move down one frequency bin~\cite{}.
+In LMXB searches, the transition matrix allows the signal frequency to (a) stay in the same frequency bin, (b) move up a single frequency bin, or (c) move down a frequency bin at each time step~\cite{}.
+In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to either remain in the same frequency bin or move down one frequency bin~\cite{}.
 
 The Viterbi algorithm~\cite{Viterbi:1967} is used to find the most probable sequence of states given the sequence of observables.
-In the next section we describe our application of the Viterbi method and our choice of detection statistic.
+In the next section, we describe our application of the Viterbi method and our choice of detection statistic.
 
@@
 Schematic diagram of the Viterbi algorithm. 
-The circles represent elements in a time-frequency grid, labelled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
+The circles represent elements in a time-frequency grid, labeled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
 The color of each circle represents the likelihood of a signal being present in each time-frequency element.
-The dark blue to light green gradient corresponds to low to high likelihood values in arbitrary units for the purpose of this schematic. 
+The dark blue to light green gradient corresponds to low to high likelihood values in arbitrary units for this schematic. 
 The objective is to find the most probable path of a signal through the grid from left to right.
 All possible paths are shown by the lines. 
-The black lines show the best path to each circle and gray lines show rejected paths. 
-Some routes though the grid result in dead ends with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a).
+The black lines show the best path to each circle and the gray lines show rejected paths. 
+Some routes through the grid result in dead-ends with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a).
 At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:pathprobability}, marked (b). 
@@
 The timeseries data from the interferometer is first split into segments. 
-We take the discrete Fourier transform of each timeseries segment and use the normalised Fourier amplitude $F(t_i,f_j)$ as the detection statistic (this is the emission matrix).
-Assuming Gaussian noise, this detection statistic choice maximises the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.
-The detection statistic is normalized for convenience and we use it to form a grid in time and frequency (a spectrogram). \han{add normalization - ask James}
+We then take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude $F(t_i,f_j)$ as the detection statistic (this is the emission matrix).
+Assuming Gaussian noise, this detection statistic choice maximizes the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.
+The detection statistic is normalized for convenience by dividing each value by the maximum Fourier amplitude in the grid (such that $\max_{i,j} F(t_i,f_j) = 1$).
 
@@
 The transition probability matrix, $A(f_k,f_m)$, describes the probability of system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
-Here, we allow the frequency of the signal to either (a) stay in the same bin; (b) move up by a single frequency bin; or (c) move down by a single frequency bin at each transition A 
+Here, we allow the frequency of the signal to either (a) stay in the same bin, (b) move up by a single frequency bin, or (c) move down by a single frequency bin at each transition A. 
 We assign these three transitions equal probability, i.e.\ $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
@@
 The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises $Pr[ f(t_0), f(t_1),... f(t_n) | data]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
-We note that there is no marginalization used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we mazimise over all such paths to find the optimal Viterbi path.
+We note that there is no marginalization used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
 %The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
@@
 These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the gray lines show the rejected paths. 
-For example, the most probable connection to the element labelled (a) is the one directly behind it (i.e., $f_4$). 
+For example, the most probable connection to the element labeled (a) is the one directly behind it (i.e., $f_4$). 
 Therefore, this path is selected as the best path from $t_0$ to $t_1$ for $f_4$.
@@
 \item We have now found the most probable path to each $f_j$ at $t=t_{N_t}$ and its probability (Eqn.~\ref{eqn:pathprobability}). 
-We select the terminating frequency bin $f(t_{N_t})$ with the highest probability labelled as (b) in Fig.~\ref{fig:viterbi}.
+We select the terminating frequency bin $f(t_{N_t})$ with the highest probability labeled as (b) in Fig.~\ref{fig:viterbi}.
 
@@
 In continuous wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency could be changing unpredictably over time.
-A Fourier transform of the entire dataset as described in Section~\ref{sec:single_tone} would split the signal power between multiple frequency bins and could result it in being undetectable under the noise.
+A Fourier transform of the entire dataset as described in Section~\ref{sec:single_tone} would split the signal power between multiple frequency bins and could result in the signal being undetectable under the noise.
 The Viterbi algorithm's strength lies in its ability to track such signals through the data.
-In the following section we present the results of using the Viterbi algorithm with the table-top interferometer data. 
+In the following section, we present the results of using the Viterbi algorithm with the table-top interferometer data. 
 % - just simiplifed this discussion a little and added a reference to the CW searches
@@
 Recovery of a wandering tone using the Viterbi algorithm.
-The spectrogram (heatmap) coloring indicates the value of the detection statistic (here, the logarithm of the absolute value of the discrete Fourier transform) in each time-frequency bin -- with brighter colours indicating higher values.
+The spectrogram (heat-map) coloring indicates the value of the detection statistic (here, the logarithm of the absolute value of the discrete Fourier transform) in each time-frequency bin -- with brighter colors indicating higher values.
 The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path, respectively. 
@@
 Note that we implicitly approximate the noise at the webcam as white (uniform in frequency) in this implementation of the Viterbi algorithm. 
-This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum} but is a resonable approximation for the higher-frequency (around 5~Hz) wandering frequency signals that we use.
+This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum} but is a reasonable approximation for the higher-frequency (around 5~Hz) wandering frequency signals that we use.
 
-The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heatmap shows the spectrogram of the observed signal (similar to that represented by the schematic in Fig.~\ref{fig:viterbi}). 
+The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heat-map shows the spectrogram of the observed signal (similar to that represented by the schematic in Fig.~\ref{fig:viterbi}). 
 In this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous-wave signals to have far lower signal-to-noise ratios. 

### 2021-04-29 paper/paper_main.tex
going through more comments
fixing placeholder text in section V, putting authors back in for now (had commented for double blind review)
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check

@@
 \newcommand{\han}{\textcolor{orange}}
+\newcommand{\fixme}{\textcolor{blue}}
 
@@
 \author{\textbf{[Authors removed for double-anonymous review.]}}
-\begin{comment}
 
@@
 
-\end{comment}
-
 % Summer 2019/2020
@@
 Using signal processing techniques from continuous-wave searches, we demonstrate the recovery of tones with constant and wandering frequencies. 
-We explore the use of the interferometer as a teaching tool for educators in physics and electrical engineering by using it as an ``optical microphone'' to capture music and speech. 
+We also explore the use of the interferometer as a teaching tool for educators in physics and electrical engineering by using it as an ``optical microphone'' to capture music and speech. 
 % \jam{We mention undergraduate labs twice, roll this sentence into the end? Also, electrical sounds weird, why the change from electrical engineering?}
@@
 % \jam{I changed all the \emph to ``'', feel free to change them back, but I think it is more explicit this way}
-We present a range of example filtering techniques to recover the original signal from noisy data, including simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
+We present a range of example filtering techniques to recover the original signal from noisy data, including a bandpass filter, notch filters, a Wiener filter, and the statistical log minimum mean-square-error (logMMSE) estimator.
+% log minimum mean-square-error (logMMSE) estimator
 Recordings of simple chords and drums are easily recovered, but complex music and speech are more challenging. 
-This demonstration has scope for use by educators in undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
+This demonstration can be used by educators in undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
+% This demonstration has scope for use by educators in undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
 % rearranged in response to Anna's pnp review

### 2021-05-03 paper/response/AJPRefereeResponse.tex
working on single and wandering tone text and response

@@
 \begin{itemize}
-\item We have added a sentence (``The Discrete Fourier transform was calculated using the NumPy package for Python.''), a citation (to the creators of NumPy), and a footnote (to the documentation).
+\item Yes, the Fourier transform was generated using software.
+We have added a sentence to Section III (``The Discrete Fourier transform was calculated using the NumPy package for Python.''), a citation (to the creators of NumPy), and a footnote (to the documentation).
 \end{itemize}
@@
 5 spectrogram result.
-To clarify this we have made the colour scheme of the schmatic more consistent with that of Fig.
+To clarify this we have made the colour scheme of the schematic more consistent with that of Fig.
 5 and we have refered back to the schematic when we first introduce Fig.
@@
 \item The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals.
-The single constant tone in section III is recovered easily using a Fourier transform, however if the signal frequency changes randomly over time, this will reduce the Fourier transform's ability to identify the signal as the signal power is spread across multiple frequency bins.
-We have added some text at the start of Section IV (paragraph 2) and at the end of Section IV.
+The single constant tone in section III is recovered easily using a Fourier transform, however if the signal frequency changes over time, the signal power is spread across multiple frequency bins.
+We have added a new figure to illustrate this (Fig.
+4) and some text at the start of Section IV (paragraph 2) and at the end of Section IV.
 B.
@@
 Can you compare what is shown to what would be detected without using the VA?
-Is this “heatmap” the same representation as is often shown in LIGO publications?
+Is this ``heatmap'' the same representation as is often shown in LIGO publications?
 \begin{itemize}
@@
 \item The heatmap is similar to the figures in LIGO-Virgo-KAGRA publications which show transient gravitational-wave signals, but not identical.
-Those figures normally show Q-tranforms [described in Chatterji, et al., Classical Quantum Gravity 21, S1809 (2004)] whereas we show the Fourier transform in this work.
+Those figures normally show Q-transforms [described in Chatterji, et al., Classical Quantum Gravity 21, S1809 (2004)] whereas we show the Fourier transform in this work.
+We have not drawn comparisons the the Q-transform plots in order to avoid confusion.
 \end{itemize}
@@
 (Section V, paragraph 2)
-\item Fig. 6: we have removed this figure and moved the photo to the Supplementary Material.
+\item Fig. 6: agreed, we have removed this figure and moved the photo to the Supplementary Material.
 \item The cloth screen was used to reduce the intensity of light on the photodiode and avoid saturating the analog-to-digital converter.
@@
 \begin{itemize}
-\item This has now been done. Each value is divided by the maximum value in the grid such that $\max_{i,j} F(t_i, f_j) = 1$.
+\item Each value is divided by the maximum value in the grid such that $\max_{i,j} F(t_i, f_j) = 1$. We have added this information to section IV.B.
 \end{itemize}

### 2021-05-03 paper/ifo-singleTone.tex
fixing plots - part 1
fixing and combining plots - part 2
defined normalisation, added comparison spectrum of entire timeseries for VA discussion, grammar/spelling check
working on single and wandering tone text and response

@@
 \begin{figure}
-	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
+	\includegraphics[width=.45\textwidth]{figures/webcam_spectrum_expt_4_0209.pdf}
 	\caption{\label{fig:webcam_spectrum}
@@
 
-% rearranged as I found this section to switch between method and results too much, I've tried this order now: intro, webcam and data specifics, injection, and results.
+% rearranged as I found this section to not switch between method and results too much, I've tried this order now: intro, webcam and data specifics, injection, and results.
 
@@
 
-As described in Section~\ref{sec:ifo}, the audio signal is played through a speaker fixed to the back of M2. 
-The intensity of the interference pattern is measured at a single point on the screen, indicated by the pink star in Fig~\ref{fig:interference_pattern}. 
+As described in Section~\ref{sec:ifo}, the audio signal is played through a speaker fixed to the back of M2 (see Fig.~\ref{fig:ifo_schematic_webcam}). 
+The intensity of the interference pattern is measured at a single point on the screen, indicated by the pink star in Fig.~\ref{fig:interference_pattern}. 
 The webcam records video in three color channels; red, blue, and green. 
@@
 They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo} and Appendix~\ref{app:intensity_derivation}.
-Also note that the noise appears to rise at lower frequencies.
+Also, note that the noise appears to rise at lower frequencies.
 
-In the following section, we use a hidden Markov model technique, commonly used in continuous-wave searches, to recover a slowly wandering frequency signal.
+In the following section, we use a hidden Markov model technique, commonly used in continuous-wave searches, to recover a signal with a slowly changing frequency.
 

### 2021-05-03 paper/ifo-wanderingTone.tex
working on single and wandering tone text and response

@@
 
-\begin{figure}
-	\includegraphics[width=.45\textwidth]{figures/viterbi_comparison_entire_spectrum_viterbi_test_webcam.pdf}
-	\caption{\label{fig:viteri_comparison_webcam_spectrum}
-Why we use the Viterbi algorithm to recover wandering frequency signals: Here we plot the Fourier amplitude of the intensity pattern against frequency as in Fig.~\ref{fig:webcam_spectrum} for the entire timeseries of the wandering frequency signal in Fig.~\ref{fig:viterbi_overlay}.
-Instead of resolving definite peaks like the constant frequency signal shown in Fig.~\ref{fig:webcam_spectrum}, we only see inconclusive peaks at 5.7~Hz and 10.7~Hz that correspond to the long amount of time that the wandering frequency spends near those values and/or its amplitude there, as seen in the pink dots in Fig.~\ref{fig:viterbi_overlay}.
-With the signal spread out over the frequency bins, we are unable to reliably detect it above the noise and therefore we use the Viterbi algorithm instead.
-The large values at low frequencies below 0.1~Hz are similar to those in Fig.~\ref{fig:webcam_spectrum} and can be ignored as they are likely inaccurate due to the short length of the signal relative to their period (i.e.\ 300~s only contains 30~periods of a 0.1~Hz signal), but this feature would require further research to fully determine its origin.
-%maybe could be from the long period (100~s) variation?
-}
-\end{figure}
-
 Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational waves may not be monochromatic.
@@
 
-A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case, as the signal can be spread across multiple frequency bins making it challenging or perhaps impossible to detect, as shown in Fig.~\ref{fig:viteri_comparison_webcam_spectrum}.
+We first demonstrate why a Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case.
+As shown in Fig.~\ref{fig:viterbi_comparison_webcam_spectrum}, the signal is spread across multiple frequency bins, making it challenging or in some cases impossible to detect.
 %A different means of detection is needed for a signal which has a randomly wandering frequency.
 
-In continuous wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
-The analysis described and presented in this section is directly inspired by real continuous-wave searches~\cite{} 
-In section~\ref{sec:realCWSearches} we briefly review the analysis method used by LIGO and Virgo, in section~\ref{} we describe the methods as applied in this work, and in section~\cite{} we our results of recovering a wandering frequency signal using the table-top interferometer.
+In continuous wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time.
+The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
+In section~\ref{sec:realCWSearches} we briefly review the analysis method used by LIGO and Virgo, in Section~\ref{sec:viterbi} we describe the methods as applied in this work, and in section~\ref{sec:wanderingResults} we show results for recovering a wandering frequency signal using the table-top interferometer.
+
+\begin{figure}
+	\includegraphics[width=.45\textwidth]{figures/viterbi_comparison_entire_spectrum_viterbi_test_webcam.pdf}
+	\caption{\label{fig:viterbi_comparison_webcam_spectrum}
+Fourier amplitude identical to that shown in Fig.~\ref{fig:webcam_spectrum}, except using a signal which has a slowly changing frequency instead of a constant frequency signal.
+The signal is spread between many frequency bins. 
+We see peaks at $5.7\,{\rm Hz}$ and $10.7\,{\rm Hz}$ and refer the reader forward to Fig.~\ref{fig:viterbi_overlay} to see the frequency evolution of this signal. 
+A single Fourier transform of the entire dataset is not well suited to detecting signals with wandering frequencies.
+% just making this a little more concise 
+%Why we use the Viterbi algorithm to recover wandering frequency signals: Here we plot the Fourier amplitude of the intensity pattern against frequency as in Fig.~\ref{fig:webcam_spectrum} for the entire timeseries of the wandering frequency signal in Fig.~\ref{fig:viterbi_overlay}.
+Instead of resolving definite peaks like the constant frequency signal shown in Fig.~\ref{fig:webcam_spectrum}, we only see inconclusive peaks at 5.7~Hz and 10.7~Hz that correspond to the long amount of time that the wandering frequency spends near those values and/or its amplitude there, as seen in the pink dots in Fig.~\ref{fig:viterbi_overlay}.
+With the signal spread out over the frequency bins, we are unable to reliably detect it above the noise and therefore we use the Viterbi algorithm instead.
+%The large values at low frequencies below 0.1~Hz are similar to those in Fig.~\ref{fig:webcam_spectrum} and can be ignored as they are likely inaccurate due to the short length of the signal relative to their period (i.e.\ 300~s only contains 30~periods of a 0.1~Hz signal), but this feature would require further research to fully determine its origin.
+%maybe could be from the long period (100~s) variation?
+}
+\end{figure}
 
@@
 For further details on these methods see Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017} and for searches performed using them see Refs.\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application}.
-\han{check for new papers}
+\han{check for new papers - add Deeksha's paper}
 
@@
 In continuous wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency could be changing unpredictably over time.
-A Fourier transform of the entire dataset as described in Section~\ref{sec:single_tone} would split the signal power between multiple frequency bins and could result in the signal being undetectable under the noise.
+A Fourier transform of the entire dataset splits the signal power between multiple frequency bins and could result in the signal being undetectable under the noise (see Fig.~\ref{fig:viterbi_comparison_webcam_spectrum}).
 The Viterbi algorithm's strength lies in its ability to track such signals through the data.

### 2021-05-04 paper/response/AJPRefereeResponse.tex
x :q

@@
 * E. Chassande-Mottin and A. Pai, Phys. Rev. D 73, 042003 (2006). \\
+% Detection of chirps with model of physically realistic GW mergers. Parametric model 
 * E. J. Candes, P. R. Charlton, and H. Helgason, Classical Quantum Gravity 25, 184020 (2008).   \\
+% dynamical programming to find chirping signals. Finding chirp-like signals that are otherwise unmodelled. Non parametric model. 
 * W. G. Anderson and R. Balasubramanian, Phys. Rev. D 60, 102001 (1999).   \\
+% Real-time detection of poorly modelled and unmodelled sources
 * P.
@@
 \\
+% unmodelled transient detection for chirps and burst signals 
 * E. Thrane et al., Phys. Rev. D 83, 083004 (2011).   \\
+% long transient search (signals lasting many seconds or weeks). Pattern recognition, not match filter
 * E. Thrane and M. Coughlin, Phys. Rev. D 89, 063012 (2014).    \\
+% unmodelled transient searches - all sky
 * S. D. Mohanty, Phys. Rev. D  D 96, 102008 (2017).   \\
+% chirp searches
 * Margaret Millhouse, Neil J. Cornish, and Tyson Littenberg, Phys. Rev. D 97, 104057 (2018).\\
+% chirplet search
+
 \begin{itemize}
-\item The Viterbi algorithm is not a new approach for gravitational wave analysis.
-It is used in many searches for continuous waves to date \todo{refs} and the methods papers are Refs \todo{refs}.
-We use these existing methods as inspiration and apply them to this table-top experiment in order to demonstrate the real gravitaitonal wave searches using sound.
-It is important to clarify that we are not searching for chirp signals in this work, we are simulating a signal with wandering frequency as an analoge to continuous wave signals whose frequencies can wander stochastically over time.
-\item [what we have done].
+\item The hidden Markov model (HMM) and Viterbi algorithm is not a new approach for gravitational wave analysis.
+It is used in many searches for continuous waves to date (see Refs.\todo{refs}) and the methods papers are described in Refs \todo{refs}.
+We use these existing methods as inspiration and apply them to this table-top experiment in order to demonstrate a gravitational wave analysis using sound.
+The HMM and Viterbi method enable a search for persistent gravitational waves where the signal model allows for the continuous wave frequency to change unpredictably over months and years.
+We note that many of the papers listed above are about searching for transient signals rather than the persistent signals considered here.
+\item We have added a new subsection (Section IV.
+A.
+in the revised draft) to emphasise that these techniques are not new and that they are inspired by real continuous wave searches.
+In this section we have pointed the reader to the search papers and methods papers that use HMM and Viterbi.
+We have also added references to other (non-Viterbi) continuous wave search methods.
+Finally we point the interested reader, who would like to know more about chirp or transient searches, to the papers listed above.
+%It is important to clarify that we are not searching for chirp signals in this work, we are simulating a signal with wandering frequency as an analoge to continuous wave signals whose frequencies can wander stochastically over time.
+%\item [what we have done].
 We have clarified that the method is not new in a new subsection in section IV (section IV A) reviewing how these methods are used in gravitational wave analysis before describing our application of the methods in this work.
-\todo{We also add references to a ......}
-\item \han{[from Hannah] These papers are addressing how to detect an unknown chirp signal from a compact binary merger.
+%\todo{We also add references to a ......}
+%\item \han{[from Hannah] These papers are addressing how to detect an unknown chirp signal from a compact binary merger.
 I think it may be relevant in some cases, but will take a look at the papers.

### 2021-05-04 paper/ifo-wanderingTone.tex
x :q

@@
 
-The methods used here are directly inspired by LIGO-Virgo continuous wave searches. 
+The methods used here are  inspired by LIGO-Virgo continuous wave searches. 
 For further details on these methods see Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017} and for searches performed using them see Refs.\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application}.
 \han{check for new papers - add Deeksha's paper}
+For other continuous wave search analyses not discussed here, see for example Refs.\cite{} and for further information on searching for transient gravitational wave signals, see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
 

### 2021-05-04 paper/ifo-wanderingTone.tex
working on adding suggested references and response

--- 2021-05-03 paper/ifo-wanderingTone.tex
@@
 
-The methods used here are directly inspired by LIGO-Virgo continuous wave searches. 
+The methods used here are  inspired by LIGO-Virgo continuous wave searches. 
 For further details on these methods see Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017} and for searches performed using them see Refs.\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application}.
 \han{check for new papers - add Deeksha's paper}
+For other continuous wave search analyses not discussed here, see for example Refs.\han{todo}\cite{} and for further information on searching for transient gravitational wave signals, see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
 

### 2021-05-04 paper/ifo-complexAudio.tex
adding noise subtraction summary and references

@@
 This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), as shown in Fig.~\ref{fig:psd_noise}. 
-The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ mains signal up to and beyond $8$th harmonic thereof. 
+The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ Australian mains power grid signal up to and beyond $8$th harmonic thereof. 
 The mains signal is also present (but far weaker) in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 
@@
 The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
+
+Environmental noise reduction is an active area of research for gravitational-wave detectors~\cite{EfflerEtAl:2015}.
+In some cases environmental noise sources can be removed or mitigated on site~\cite{CovasEtAl:2018}. 
+Where this is not possible, noise subtraction can be done after data collection~\cite{DriggersEtAl:2019, DavisEtAl:2019}. 
+One example is removal of spectral noise lines from the United States power grid operating at $60\,{\rm Hz}$~\cite{VajenteEtAl:2020}.
 

### 2021-05-05 paper/response/AJPRefereeResponse.tex
working on adding suggested references and response
adding noise subtraction summary and references
working on response text

@@
 
-\todo{Opening comments}\\
+\noindent Dear Joseph Amato, \\
 
-Dear Joseph Amato,
-
-We thank you and the referees for your helpful comments and suggestions. 
+\noindent We thank you and the referees for your helpful comments and suggestions. 
 We address the comments below in turn. 
+Ad we have made many changes throughout the manuscript, we have not highlighted them within the manuscript, instead we have pointed to each of the changes in our responses below.
 
@@
 \item We have moved the footnotes to the bibliography
-\item \todo{Work in progress, please correct any that have not been fixed yet}
+\item We have checked through the manuscript to correct spellings to US English. 
+\item[]
 \end{itemize}
@@
 We have added a sentence to Section III (``The Discrete Fourier transform was calculated using the NumPy package for Python.''), a citation (to the creators of NumPy), and a footnote (to the documentation).
+\item[]
 \end{itemize}
@@
 \begin{itemize}
-\item We have added a sentence to explain Markov process: ``In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal).''
-\item We have lightened the colour scheme used in the Fig.
-4 schematic to aid with black and white printing.
-Our intention is that the Fig.
-4 schematic be interpreted as a simplified version of the Fig.
-5 spectrogram result.
-To clarify this we have made the colour scheme of the schematic more consistent with that of Fig.
-5 and we have refered back to the schematic when we first introduce Fig.
-5.
+\item Markov process: We have added a sentence to explain Markov process in Section IV A, ``In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal).''
+\item Figure 4: We have lightened the color scheme used in the Figure 4 schematic to aid with black and white printing.
+The colors represent a gradient rather than a discrete shading and we now use a blue/green color palette to clarify this.
+Our intention is that the Figure 4 schematic be interpreted as a simplified version of the Figure 6 (previously Figure 5) spectrogram result.
+To clarify this we have made the colour scheme of the schematic more consistent with that of Figure 6 and we have referred back to the schematic when we first introduce Figure 6.
+\item The grey paths in Figure 4 are those that have been rejected at each timestep. We have clarified this in the text and caption. 
 \item The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals.
 The single constant tone in section III is recovered easily using a Fourier transform, however if the signal frequency changes over time, the signal power is spread across multiple frequency bins.
-We have added a new figure to illustrate this (Fig.
-4) and some text at the start of Section IV (paragraph 2) and at the end of Section IV.
-B.
-to explain that the Fourier transform of the whole dataset is not well suited to this type of signal.
+We have added a new figure to illustrate this (Figure 4) and some text at the start of Section IV (paragraph 2) and at the end of Section IV B to explain that the Fourier transform of the whole dataset is not well suited to this type of signal.
 \item We have also revised the entire Section IV to improve clarity in our Viterbi algorithm explanation and motivation.
+\item[]
 \end{itemize}
@@
 \begin{itemize}
-\item The color coding represents the logarithm of the absolute value of the discrete Fourier transform. We have clarified this in the caption.
-\item Without using the VA the signal becomes spread out over the frequency bins and might not rise above the noise.
-We have now explained this and added a plot that shows the Fourier spectrum of the entire timeseries.
-\item The heatmap is similar to the figures in LIGO-Virgo-KAGRA publications which show transient gravitational-wave signals, but not identical.
-Those figures normally show Q-transforms [described in Chatterji, et al., Classical Quantum Gravity 21, S1809 (2004)] whereas we show the Fourier transform in this work.
-We have not drawn comparisons the the Q-transform plots in order to avoid confusion.
+\item Figure 5 (now Figure 6) colors: The color coding represents the logarithm of the absolute value of the discrete Fourier transform. We have added this to the caption.
+\item Comparison without VA: Without using the Viterbi algorithm the signal becomes spread out over the frequency bins and might not rise above the noise.
+We have now explained this and added a plot that shows the Fourier spectrum of the entire timeseries to illustrate that the Fourier transofrm is not well suited here.
+(See also our answer to point 3 above).
+\item Heatmap representation: The heatmap is similar to the figures in LIGO-Virgo-KAGRA publications which show transient gravitational-wave signals, but not identical.
+Those figures normally show Q-transforms (described in Chatterji, et al., Classical Quantum Gravity 21, S1809 2004).
+Our work instead shows the Fourier transform at each time-frequency grid point in the spectrogram.
+We have decided not to draw comparisons to the Q-transform plots in our manuscript order to avoid confusion.
+\item[]
 \end{itemize}
@@
 \begin{itemize}
-\item Optical microphone: we have rearranged the text so that the optical microphone concept is introduced in a separate paragraph and expanded our discission of it to improve clarity.
+\item Optical microphone: we have rearranged the text so that the optical microphone concept is introduced in a separate paragraph and expanded our discussion of it to improve clarity.
 (Section V, paragraph 2)
-\item Fig. 6: agreed, we have removed this figure and moved the photo to the Supplementary Material.
 \item The cloth screen was used to reduce the intensity of light on the photodiode and avoid saturating the analog-to-digital converter.
-It was indeed an improvised solution using the ``grill cloth'' and plastic frame from the dismantelled commercial speaker used in the demonstration.
-We have added a sentence to Section V.A.
-to clarify.
+It was indeed an improvised solution using the ``grill cloth'' and plastic frame from the dismantled commercial speaker used in the demonstration.
+We have added a sentence to the end of Section V A to clarify.
+\item Figure 6: agreed, we have removed this figure and moved the photo to the Supplementary Material.
+\item []
 \end{itemize}
@@
 A note has been made in the Supplementary Material where this plot has been moved to.
-% (this sharp drop-off would appear shallower when shown against a logarithmic scale for frequency)
-% \item \todo{They will also expect a change to the paper as it is a point of confusion - e.g. clarify the reason in the figure caption.}
+\item[]
 \end{itemize}
@@
 \item Agreed, we have moved the majority of this section to Supplementary Material.
-We have chosen to show two highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator), but we refer the reader to the Supplementary Material for the details of these methods.
+We have chosen to show two highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator), but we refer the reader to the Supplementary Material for the details of these methods as well as a selection of relevant textbooks.
 We have also expanded the discussion on filters more generally in the main text.
+\item[]
 \end{itemize}
 
-\noindent 8.Where in your flowchart are the various filters you describe?  Are the codes for them readily available?  How does one implement a filter?  
+\noindent 8. Where in your flowchart are the various filters you describe?  Are the codes for them readily available?  How does one implement a filter?  
 \begin{itemize}
 \item The filters are performed separately after the data is gathered. We have added a sentence to the figure caption to clarify this. 
-\item Most of the code is readily available on the provided GitHub with the exception of the Weiner filtering results.
-The provided code is documented and commented if someone wanted to replicate the filter implementations.
-\item \todo{We may need to have access to Changrong's processing code}
-\item \todo{...}
+\item Most of the code is readily available on the provided GitHub with the exception of the Wiener filtering results.
+The provided code is documented and commented to ease reproduceability.
+For the purposes of the double-blind peer review, we removed the link to our git repository from the manuscript.
+\todo{[todo] Ask Changrong whether we can include scripts in the git repo}
+\item[]
 \end{itemize}
@@
 I look forward to a revised manuscript.
+
+\hspace{1em}\vspace{2em}\\
 
@@
 \begin{itemize}
-\item We have including some details on this in the description of detection statistics in section IV.A.
-and have also added that Doppler modulation of the source also needs to be taken into account if searching for a neutron star in a binary.
-\item \todo{...}
-\item A mention has been made in the Future Work as to demonstrating the Doppler effect.
- \todo{...}
+\item We have including some details on how the detection statistic is used to take account of the Earth's motion around the Sun in Section IV A.
+We and have also noted that Doppler modulation of the source also needs to be taken into account if searching for a neutron star in a binary.
+\item It would indeed be very nice to modify the setup to show the Doppler effect. We have added some possible ways to explore this topic in our future work discussion in Section VI. 
+\item[]
 \end{itemize}
@@
 The detection statistic captures the likelihood of a signal being present in each time-frequency bin given the observable data.
-In Section IV.A.
-we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission matrix by name to reduce confusion.
+In Section IV A we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission matrix by name to reduce confusion.
+\item[]
 \end{itemize}
@@
 The equation should give the probability for a specific path, and is not a sum over all possibilities.
-We have corrected Eq.
-1 and added clarification to the text.
+We have corrected equation 1 and clarified what it represents in the text.
+\item[]
 \end{itemize}
@@
 \begin{itemize}
-\item Each value is divided by the maximum value in the grid such that $\max_{i,j} F(t_i, f_j) = 1$. We have added this information to section IV.B.
+\item Each value is divided by the maximum value in the grid such that $\max_{i,j} F(t_i, f_j) = 1$. We have added this information to section IV B.
+\item[]
 \end{itemize}
@@
 \begin{itemize}
-\item We do assume that the noise is white in our analysis. We have added some text to clarify this at the start of Section IV.C.
+\item We do indeed assume that the noise is white in our analysis. We have added some text to clarify this assumption in the first paragraph of Section IV C.
+\item[]
 \end{itemize}
@@
 A partial list of papers is given below: \\
-* E. Chassande-Mottin and A. Pai, Phys. Rev. D 73, 042003 (2006). \\
+$\ast$ E. Chassande-Mottin and A. Pai, Phys. Rev. D 73, 042003 (2006). \\
 % Detection of chirps with model of physically realistic GW mergers. Parametric model 
-* E. J. Candes, P. R. Charlton, and H. Helgason, Classical Quantum Gravity 25, 184020 (2008).   \\
+$\ast$ E. J. Candes, P. R. Charlton, and H. Helgason, Classical Quantum Gravity 25, 184020 (2008).   \\
 % dynamical programming to find chirping signals. Finding chirp-like signals that are otherwise unmodelled. Non parametric model. 
-* W. G. Anderson and R. Balasubramanian, Phys. Rev. D 60, 102001 (1999).   \\
+$\ast$ W. G. Anderson and R. Balasubramanian, Phys. Rev. D 60, 102001 (1999).   \\
 % Real-time detection of poorly modelled and unmodelled sources
-* P.
+$\ast$ P.
 Addesso, M.
@@
 % unmodelled transient detection for chirps and burst signals 
-* E. Thrane et al., Phys. Rev. D 83, 083004 (2011).   \\
+$\ast$ E. Thrane et al., Phys. Rev. D 83, 083004 (2011).   \\
 % long transient search (signals lasting many seconds or weeks). Pattern recognition, not match filter
-* E. Thrane and M. Coughlin, Phys. Rev. D 89, 063012 (2014).    \\
+$\ast$ E. Thrane and M. Coughlin, Phys. Rev. D 89, 063012 (2014).    \\
 % unmodelled transient searches - all sky
-* S. D. Mohanty, Phys. Rev. D  D 96, 102008 (2017).   \\
+$\ast$ S. D. Mohanty, Phys. Rev. D  D 96, 102008 (2017).   \\
 % chirp searches
-* Margaret Millhouse, Neil J. Cornish, and Tyson Littenberg, Phys. Rev. D 97, 104057 (2018).\\
+$\ast$ Margaret Millhouse, Neil J. Cornish, and Tyson Littenberg, Phys. Rev. D 97, 104057 (2018).\\
 % chirplet search
@@
 \item The hidden Markov model (HMM) and Viterbi algorithm is not a new approach for gravitational wave analysis.
-It is used in many searches for continuous waves to date (see Refs.\todo{refs}) and the methods papers are described in Refs \todo{refs}.
-We use these existing methods as inspiration and apply them to this table-top experiment in order to demonstrate a gravitational wave analysis using sound.
+It is used in many searches for continuous waves to date (for example Abbott et al., Phys.
+Rev.
+D, 100(12):122002, Dec 2019) and the methods papers are described in Suvorova et al., Phys.
+Rev.
+D, 96(10):102006, November 2017 and Suvorova et al.
+Phys.
+Rev.
+D, 93(12):123009, June 2016.
+We have clarified throughout the paper that these methods are not new and that we use real continuous wave searches as inspiration for our table-top demonstration.
 The HMM and Viterbi method enable a search for persistent gravitational waves where the signal model allows for the continuous wave frequency to change unpredictably over months and years.
 We note that many of the papers listed above are about searching for transient signals rather than the persistent signals considered here.
-\item We have added a new subsection (Section IV.
-A.
-in the revised draft) to emphasise that these techniques are not new and that they are inspired by real continuous wave searches.
+\item We have added a new subsection (Section IV A in the revised draft) to emphasise that these techniques are not new and that they are inspired by real continuous wave searches.
 In this section we have pointed the reader to the search papers and methods papers that use HMM and Viterbi.
 We have also added references to other (non-Viterbi) continuous wave search methods.
-Finally we point the interested reader, who would like to know more about chirp or transient searches, to the papers listed above.
+Finally we point the interested reader, who would like to know more about chirp or transient searches, to the papers suggested above.
+\item[]
 %It is important to clarify that we are not searching for chirp signals in this work, we are simulating a signal with wandering frequency as an analoge to continuous wave signals whose frequencies can wander stochastically over time.
@@
 The topic of digital filtering is not something that Physics undergraduates (or even graduates) are usually familiar with and requires a more gentle introduction than provided here.
-For example, the term "IIR" is introduced without explaining what the impulse response of a filter is (and there is no mention of FIR filters).
+For example, the term ``IIR'' is introduced without explaining what the impulse response of a filter is (and there is no mention of FIR filters).
 Similarly, Eq.
-4 looks like a Z-transform, and advanced topic in itself, but this is not mentioned and instead the term "complex frequency" is used that is likely to be unfamiliar to most readers.
-(Incidentally, there seems to be a notational inconsistency across the two sides of Eq.3: 'f' and 'omega'.) To improve the readability of the paper, the authors should consider shortening this section by focussing on only the filtering procedure that worked the best.
+4 looks like a Z-transform, and advanced topic in itself, but this is not mentioned and instead the term ``complex frequency'' is used that is likely to be unfamiliar to most readers.
+(Incidentally, there seems to be a notational inconsistency across the two sides of Eq.3: `f' and `omega'.) To improve the readability of the paper, the authors should consider shortening this section by focusing on only the filtering procedure that worked the best.
 References to DSP textbooks may be provided for the interested reader to acquire background material.
 \begin{itemize}
-\item We have move the majority of Section VC to Supplementary Material and now only include highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator).
-In the Supplementary Material we have expanded the description of the impulse response of a filter mentioning both IIR and FIR.
-We have also added some text saying that Eqn.
-2 in the Supplementary Material (previously Eqn.
-4 in the main text) is a Z-tranform.
-We have corrected the inconsistency in Eqn.
-1 in the Supplementary Material (previously Eqn.
-3 in the main text).
-\todo{add some references to DSP textbooks.}
+\item Length of Section V: Agreed, we have moved the majority of Section V C to Supplementary Material and now only include two highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator).
+\item IIR, FIR, and Z-transform: In the Supplementary Material we have expanded the description of the impulse response of a filter mentioning both IIR and FIR.
+We have also added some text clarifying that equation 2 in the Supplementary Material (previously equation 4 in the main text) is a Z-transform.
+\item Notation inconsistency: We have corrected the inconsistency in equation 1 in the Supplementary Material (previously equation 3 in the main text). 
+\item DSP textbooks: We have added references to a selection of Digital Signal Processing textbooks at the start of the Supplementary Material and also in the main text Section V.
+\item[]
 \end{itemize}
@@
 \begin{itemize}
-\item \todo{...}
+\item Agreed we should mention this work.
+We have added a short paragraph at the end of section V B where we discuss environmental noise with the table-top interferometer.
+We draw parallels to the real detectors and briefly mentioning the importance of noise mitigation and subtraction for gravitational-wave data with references to a selection of papers.
+\item[]
 \end{itemize}
@@
 \noindent
--- The axes label for frequency is not consistent across the figures: in some plots it is "f/Hz" and others "f(Hz)".
+-- The axes label for frequency is not consistent across the figures: in some plots it is ``f/Hz'' and others ``f(Hz)''.
 \begin{itemize}
-\item We have corrected the axis labels and they are now consistent.
+\item Thank you, we have corrected the axis labels and they are now consistent throughout.
+\item[]
 \end{itemize}
@@
 \begin{itemize}
-\item This is unknown to us but has now been mentioned in Future Work.
+\item This is an interesting question which does require further investigation. We have added this idea as a possible future investigation in Section VI. 
+\item[]
 \end{itemize}
 
-\subsection{Comments from R1}
+\hspace{1em}\vspace{1em}\\
+
+\subsection*{Comments from R1}
 
@@
 \begin{itemize}
-\item \todo{...}
-\item \han{> Thoughts on this [Hannah]: The use of audio as an analogy to gravitational waves is not novel to this work.
+\item Suitability for AJP: We thank you for the recommendations regarding the suitability and complexity of this manuscript.
+Following the advice of the editor and referee R3, we have moved the majority of Section V to Supplementary Material so that the main text is now less technical and the interested reader can still see our selection of analysis examples.
+We now only focus on two highlight results in the main text Section V.
+We have also clarified our description of the Viterbi method following recommendations from the editor and R3.
+\item Use of audio signals: We agree that there is indeed a fundamental difference between sound waves and gravitational waves.
+In the Section II we have clarified that we are using sound as an analogy.
+For the purposes of this table-top experiment, we require a signal that behaves similar to gravitational waves in that it alters the interference pattern produced by a Michelson interferometer.
+Using a speaker and audio signals provides a simple and relatively cost effective means to do this, compared to for example, a piezo-electric crystal fixed to the back of a mirror as used in some gravitational wave demonstrations (e.g.
+Cooper et al., arXivv e-prints 2004.03052, 2020).
+The main focus of this study is the extraction of a signal from interferometer data using a series of analysis techniques, some of which are inspired by continuous gravitational wave searches.
+We also note that the use of sound as an analogy to gravitational-wave signals is not novel and is used by many gravitational-wave research groups to aid in their explanations (for example Sounds of Spacetime \texttt{https://www.
+soundsofspacetime.org/} and Black Hole Hunter \texttt{blackholehunter.org})
+\item[]
+%\item \han{> Thoughts on this [Hannah]: The use of audio as an analogy to gravitational waves is not novel to this work.
 Sound has been used in numerous situations to convey the concept of gravitational wave signals to general audiences [will add references].
@@
 [will think about this some more and very happy to hear other thoughts on how we respond here -Hannah]}
-\item \jam{(1) The audio analogy is not novel, (2) the interferometer is a toy model for demonstration and teaching not for research, i.e.\ I doubt our analysis techniques are useful to the gravitational wave community in any novel way, (3) as long as we are clear in the text (todo added below) about the difference and that we are demonstrating the analysis techniques then I think that there isn't an issue, (4) we have heavily edited the signal processing section and worked on the Viterbi algorithm section which I believe (the Viterbi algorithm) is understandable to even high-school students if the explanation without using Markov models or Fourier transforms is given, i.e.
+%\item \jam{(1) The audio analogy is not novel, (2) the interferometer is a toy model for demonstration and teaching not for research, i.e.\ I doubt our analysis techniques are useful to the gravitational wave community in any novel way, (3) as long as we are clear in the text (todo added below) about the difference and that we are demonstrating the analysis techniques then I think that there isn't an issue, (4) we have heavily edited the signal processing section and worked on the Viterbi algorithm section which I believe (the Viterbi algorithm) is understandable to even high-school students if the explanation without using Markov models or Fourier transforms is given, i.e.
 the pseudocode.
 - James}
-\item \todo{Clearly state the differences between gravitational waves and sound waves in the text but reiterate that despite these differences our demonstration is still useful to explain signal analysis techniques used for gravitational waves.}
+%\item \todo{Clearly state the differences between gravitational waves and sound waves in the text but reiterate that despite these differences our demonstration is still useful to explain signal analysis techniques used for gravitational waves.}
 \end{itemize}

### 2021-05-05 paper/ifo-introduction.tex
small typo fix addded reference to supplementary material to intro

@@
 
-In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the universe~\cite{GW150914}.
+In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the Universe~\cite{GW150914}.
 Gravitational waves are a prediction of Albert Einstein's theory of General Relativity~\cite{Einstein:1916}, 
 They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
-The effect of gravitational waves is a change in lengths: a `stretching and squashing' of the distance between objects. 
+The effect of gravitational waves is a change in lengths: a ``stretching and squashing'' of the distance between objects. 
 Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
@@
 % removed as there's not space to really say what this means here: to varying degrees of success.
-This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory. 
+This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory (described in the accompanying Supplementary Material).
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 

### 2021-05-05 paper/ifo-complexAudio.tex
adding references to signal processing textbooks from Changrong

@@
 Optical microphones have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles. 
-We use the term ``optical microphone'' to distinguish this technology from those previous technologies.
+%We use the term ``optical microphone'' to distinguish this technology from those previous technologies.
 Our objective is to play a recording of speech or music through the speaker attached to mirror M2 (see Fig.~\ref{fig:ifo_schematic_webcam}), record the resulting interference pattern, and then recover the original signal via a selection of signal processing techniques.
@@
 A selection of filters is applied to the data, details of which are presented in Supplementary Material. 
-We present the two best-performing analyses from our Supplementary Material in Section~\ref{sec:opticalMicResults}.
+For introductory information on digital signal processing, we refer the reader to a selection of textbooks~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045}.
+In Section~\ref{sec:opticalMicResults} we present the two best-performing analyses from our Supplementary Material.
 
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{figures/combined_highlight_results_melatos_labelled.pdf}
+\includegraphics[width=0.95\textwidth]{figures/combined_highlight_results_melatos_labelled.pdf}
 \caption{\label{fig:notchWienerLogMMSEResults}

### 2021-05-05 paper/paper_main.tex
modifying abstract to mention supplementary material

@@
 
-\author{\textbf{[Authors removed for double-anonymous review.]}}
+%\author{\textbf{[Authors removed for double-anonymous review.]}}
 
@@
 % \jam{I changed all the \emph to ``'', feel free to change them back, but I think it is more explicit this way}
-We present a range of example filtering techniques to recover the original signal from noisy data, including a bandpass filter, notch filters, a Wiener filter, and the statistical log minimum mean-square-error (logMMSE) estimator.
+In the accompanying Supplementary Material we consider a range of example filtering techniques to recover the original signal from noisy data and present highlight results here. 
+%We present a range of example filtering techniques to recover the original signal from noisy data, including a bandpass filter, notch filters, a Wiener filter, and the statistical log minimum mean-square-error (logMMSE) estimator.
 % log minimum mean-square-error (logMMSE) estimator

### 2021-05-05 paper/ifo-introduction.tex
some reference fixes

@@
 % removed Additionally - I don't think it is neccessary - hannah
-It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019, SuvorovaEtAl:2017,SuvorovaEtAl:2016}.
+It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~
+\cite{SuvorovaEtAl:2017,SuvorovaEtAl:2016, ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, BeniwalEtAl:2021, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, BayleyEtAlSOAP:2019, SunEtAlSNR:2018}.
 % This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.

### 2021-05-05 paper/ifo-complexAudio.tex
minor changes from read through

@@
 
-Here we use the Michelson interferometer as an ``optical microphone'', where a laser is used to detect sound instead of the electrical components in a conventional microphone.
+We use the Michelson interferometer as an ``optical microphone'', where a laser is used to detect sound instead of the electrical components in a conventional microphone.
 Optical microphones have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles. 
@@
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
-A selection of filters is applied to the data, details of which are presented in Supplementary Material. 
-For introductory information on digital signal processing, we refer the reader to a selection of textbooks~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045}.
+We consider a selection of filter techniques, details of which can be found in the Supplementary Material and we refer the reader to Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045} for further information on digital signal processing.
 In Section~\ref{sec:opticalMicResults} we present the two best-performing analyses from our Supplementary Material.
@@
 \begin{figure}
-	\includegraphics[width=.45\textwidth]{figures/psd_podo_14_6.pdf}
+	\includegraphics[width=.5\textwidth]{figures/psd_podo_14_6.pdf}
 	\caption{\label{fig:psd_noise}
@@
 
-Firstly, we consider two signal processing techniques used in combination: the cascaded notch and the Wiener filter. 
+Firstly, we consider two signal processing techniques used in combination: the cascaded notch~\cite{10.5555/541204} and the Wiener filter~\cite{10.5555/151045}. 
 A notch filter removes signals within a specific frequency range. 

### 2021-05-05 paper/ifo-conclusions.tex
minor changes from read through

@@
 
-In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating a range of signal processing techniques used within the gravitational-wave community and more broadly.
+In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community and more broadly.
 Gravitational-wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
@@
 We then consider more complex audio (music and speech) for a more general interest in signal processing and explore the use of the interferometer as an optical microphone. 
-A variety of example activities is presented. 
+A variety of example activities are considered in the Supplementary Material. 
 We find that the logMMSE produces the best results, however, it does not enhance the recording to an intelligible level. 
-The complex audio signals can also serve as a more distant analog to un-modeled gravitational-wave burst signals, e.g. from supernovae.
+The complex audio signals can also serve as a more distant analog to un-modeled gravitational-wave burst signals, e.g. from supernovae~\cite{MillhouseEtAl:2018,SuvorovaPowellMelatos:2019}.
 

### 2021-05-05 paper/ifo-futureWork.tex
minor changes from read through

@@
 \begin{document}
+
+\begin{figure*}
+	\includegraphics[height=.8\textwidth, angle=-90]{figures/pipeline.pdf}
+	\caption{\label{fig:pipeline_highlighted}
+Flowchart for the signal through the optical microphone. The transfer function would have to account for each stage. 
+The stage inside the blue-dashed box is analyzed in Appendix~\ref{app:intensity_derivation}.
+Understanding the other stages is one aim of future work.
+Not shown are the signal processing filters that are performed on the PC once the data is transferred off the Pi.
+}
+\end{figure*}
 
@@
 A feedback loop control system can also be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational-wave observatories.
-
-\begin{figure*}
-	\includegraphics[height=.8\textwidth, angle=-90]{figures/pipeline.pdf}
-	\caption{
-Flowchart for the signal through the optical microphone. The transfer function would have to account for each stage. 
-The stage inside the blue-dashed box is analyzed in Appendix~\ref{app:intensity_derivation}.
-Understanding the other stages is one aim of future work.
-Not shown are the signal processing filters that are performed on the PC once the data is transferred off the Pi.
-}
-	\label{fig:pipeline_highlighted}
-\end{figure*}
 
@@
 This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
-Another extension would be to demonstrate the Doppler effect and the continuous-wave techniques used in the literature to address it by moving the optical table on some kind of circular track with period far shorter than the time it takes the signal to significantly wander in frequency.
+Another extension would be to demonstrate the Doppler effect of the Earth's motion around the Sun, which needs to be considered in continuous-wave searches (see Section~\ref{sec:realCWSearches} and Ref.~\cite{JKS:1998}).
+One approach could be to move either the interferometer or the source signal on a circular track or modify the input audio signal to simulate Doppler modulation. 
+%techniques used in the literature to address it by moving the optical table on some kind of circular track with period far shorter than the time it takes the signal to significantly wander in frequency.
 Then, by placing a distant signal source at different positions above or beside the track -- or perhaps more practically the source could move instead of the detector, an analogous Doppler effect could be observed.

### 2021-05-05 paper/ifo-supplementary-material.tex
adding references to signal processing textbooks from Changrong
minor changes from read through

@@
 \title{Supplementary material for \\Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
+
+\author{James W. Gardner}
+\email{u6069809@anu.edu.au}
+% \affiliation{%
+% College of Science, Australian National University, Acton, ACT, 2601, Australia 
+% }%
+\affiliation{%
+ANU Centre for Gravitational Astrophysics, The Australian National University, Acton, ACT, 2601, Australia
+}
+\affiliation{%
+OzGrav-ANU, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, The Australian National University, Acton, ACT, 2601, Australia
+} 
+\affiliation{%
+School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+
+\author{Hannah Middleton}
+\email{hannah.middleton@unimelb.edu.au}
+\affiliation{%
+School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+\affiliation{%
+Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia 
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
+}
+
+\author{Changrong Liu}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+
+\author{Andrew Melatos}
+\email{amelatos@unimelb.edu.au}
+\affiliation{%
+School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
+}
+
+\author{Robin Evans}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
+}
+
+\author{William Moran}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+
+\author{Deeksha Beniwal}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Huy Tuong Cao}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Craig Ingram}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Daniel Brown}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Sebastian Ng}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
 \maketitle
@@
 As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
+For more information about the techniques used here see also Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045}.
 
@@
 \begin{figure}
-	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000.pdf}
+	\includegraphics[width=0.5\textwidth]{figures/butterworth_150_3000.pdf}
 	\caption{Butterworth bandpass filter frequency response.
@@
 \begin{figure}
-	\includegraphics[width=.45\textwidth]{figures/psd_butterworth_14_6.pdf}
+	\includegraphics[width=.5\textwidth]{figures/psd_butterworth_14_6.pdf}
 	\caption{\label{fig:psd_noise}
@@
 \begin{center}
-\includegraphics[width=.45\textwidth]{figures/cascaded_notch_response_plot.pdf}
+\includegraphics[width=.5\textwidth]{figures/cascaded_notch_response_plot.pdf}
 \end{center}

### 2021-05-05 paper/ifo-tableTopGWs.tex
minor changes from read through

@@
 They can measure minuscule changes in distance due to gravitational waves, e.g.
-the first detection of a binary black hole merger produced a strain of $10^{-21}$, corresponding to a change in distance which is a fraction of the width of a proton.
+the first detection of a binary black hole merger produced a strain of $10^{-21}$~\cite{GW150914}, corresponding to a change in distance which is a fraction of the width of a proton.
 
@@
 \begin{figure}
-	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam_edit.pdf}
+	\includegraphics[width=0.5\textwidth]{figures/ifo_schematic_webcam_edit.pdf}
 	\caption{\label{fig:ifo_schematic_webcam}
@@
 The output of a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO} (top-left) is reflected from a mirror at a right angle.
-The beam then passes through a converging lens (FL = $125.0\,{\rm mm}$) before reaching the beamsplitter which is $50$/$50$ in reflection and transmission.
+The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter which is $50$/$50$ in reflection and transmission.
 The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and recombine at the beamsplitter. 
-The output interference pattern is enlarged using a diverging lens (FL = $-25.0\,{\rm mm}$), projected onto a screen, and recorded using a commercial webcam connected to a computer (PC). 
+The output interference pattern is enlarged using a diverging lens (focal length: $-25.0\,{\rm mm}$), projected onto a screen, and recorded using a commercial webcam connected to a computer (PC). 
 A speaker fixed to the back of M2 is used to inject audio signals from the PC into the interferometer.
@@
 The optical configuration of the Michelson interferometer used in this work is shown in Fig.~\ref{fig:ifo_schematic_webcam}.
-It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green eye-safe laser (top-left) with peak emission at a wavelength of $532\,{\rm nm}$.
+It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green laser (top-left) with peak emission at a wavelength of $532\,{\rm nm}$.
 The output of the laser is first reflected by a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
@@
 A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations. 
-This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a PC. 
+This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a personal computer (PC). 
 The other speaker in the pair is kept face-down and as far away from the interferometer as possible to prevent it from acting as a second source.

### 2021-05-05 paper/ifo-wanderingTone.tex
minor changes from read through

--- 2021-05-03 paper/ifo-wanderingTone.tex
@@
 \begin{figure}
-	\includegraphics[width=.45\textwidth]{figures/viterbi_comparison_entire_spectrum_viterbi_test_webcam.pdf}
+	\includegraphics[width=.5\textwidth]{figures/viterbi_comparison_entire_spectrum_viterbi_test_webcam.pdf}
 	\caption{\label{fig:viterbi_comparison_webcam_spectrum}
@@
 
-The methods used here are directly inspired by LIGO-Virgo continuous wave searches. 
-For further details on these methods see Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017} and for searches performed using them see Refs.\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, SunEtAlSNR:2018, viterbi_application}.
-\han{check for new papers - add Deeksha's paper}
+The methods used here are  inspired by LIGO-Virgo continuous wave searches. 
+For further details on these methods and searches see Refs.~\cite{SuvorovaEtAl:2016, SuvorovaEtAl:2017, ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, BeniwalEtAl:2021, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, BayleyEtAlSOAP:2019, SunEtAlSNR:2018}.
+For other continuous wave search analyses not discussed here, see for example Refs.\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, MSPSearchO3O2O1:2020, RadiometerO1O2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} and for further information on searching for transient gravitational wave signals, see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
 
@@
 In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal). 
-In a hidden Markov model, the frequency state of the signal is unknown, or hidden, and can undergo transitions at discrete times. 
+In a hidden Markov model, the frequency state of the signal is unknown (hidden) and can undergo transitions at discrete times. 
 The transitions are Markovian in that the hidden state (i.e.\ frequency) of the system at any time depends solely on its state at the previous time.
@@
 This is called the transition probability matrix. 
-In LMXB searches, the transition matrix allows the signal frequency to (a) stay in the same frequency bin, (b) move up a single frequency bin, or (c) move down a frequency bin at each time step~\cite{}.
-In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to either remain in the same frequency bin or move down one frequency bin~\cite{}.
+In LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin at each time step~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017,  MiddletonEtAlO2LMXBs:2020}.
+In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to (i) remain in the same frequency bin, or (ii) move down one frequency bin~\cite{MillhouseStrangMelatos:2020, SunEtAlSNR:2018}.
 
-The Viterbi algorithm~\cite{Viterbi:1967} is used to find the most probable sequence of states given the sequence of observables.
+The Viterbi algorithm~\cite{Viterbi:1967} is used to find the most probable sequence of hidden frequency states given the sequence of observables.
 In the next section, we describe our application of the Viterbi method and our choice of detection statistic.
@@
 \begin{figure}
-\includegraphics[width=0.49\textwidth]{figures/viterbiDiagramBlue.pdf}
+\includegraphics[width=0.5\textwidth]{figures/viterbiDiagramBlue.pdf}
 \caption{\label{fig:viterbi}
@@
 The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
-%A schematic diagram of the Viterbi algorithm. 
-%Each circle represents an element in the time-frequency grid starting from the instant $t_0$ on the left and ending at $t_{N_t}$ on the right. 
-%Vertically, the grid starts from frequency $f_0$ at the bottom and ending at $f_{N_f}$ at the top. 
-%The color of each circle indicates the likelihood that a signal is present in the box (the detection statistic), where a lighter color corresponds to a higher likelihood. 
-%At each time step, the paths can move up one $f$ bin, move down one $f$ bin, or stay at the same $f$ bin (shown by lines connecting the circles in the diagram). 
-%Each circle has three paths leading to it, indicated by the lines. 
-%The best path to each circle at each $t_i$ is highlighted in black. 
-%Some routes through the grid are found to be dead ends, with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a). 
-%At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:recursiveViterbi}, marked (b). 
-%The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
 }
@@
 
-%The Viterbi algorithm is then applied to the spectrogram data. 
 We use the schematic in Fig.~\ref{fig:viterbi} to aid our explanation of the method throughout this section. 
@@
 
-%The spectrogram schematic, as shown in Fig.~\ref{fig:viterbi}, has $N_t$ time and $N_f$ frequency bins. 
-%We label the elements of the time and frequency  grid as $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-%The normalized Fourier amplitude $F(t_i,f_j)$ is the detection statistic.
-
 The objective is to find the most likely path through the grid given the observed data and any probabilistic constraints on how the frequency of the signal can wander from $t_i$ to $t_{i+1}$. 
 The transition probability matrix, $A(f_k,f_m)$, describes the probability of system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
-Here, we allow the frequency of the signal to either (a) stay in the same bin, (b) move up by a single frequency bin, or (c) move down by a single frequency bin at each transition A. 
+Here, we allow the frequency of the signal to either (a) stay in the same bin, (b) move up by a single frequency bin, or (c) move down by a single frequency bin at each transition. 
 We assign these three transitions equal probability, i.e.\ $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
 All possible transitions are shown as lines in Fig.~\ref{fig:viterbi}.
-The different colorings of the lines are explained below.
+The different line colors are explained below.
 
@@
 \end{eqnarray}
-The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises $Pr[ f(t_0), f(t_1),... f(t_n) | data]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
+The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises $Pr[ f(t_0), f(t_1),... f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
 We note that there is no marginalization used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
@@
 
-\item Moving to time $t_2$, again, we select the path which maximizes the recursive quantity in Eqn.~\ref{eqn:pathprobability} for each $f_j$. 
+\item Moving to time $t_2$, again, we select the path which maximizes Eqn.~\ref{eqn:pathprobability} for each $f_j$. 
 These paths are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.

### 2021-05-05 paper/paper_main.tex
minor changes from read through

@@
 \begin{acknowledgments}
-\textbf{[Removed for double-anonymous review.]} 
-%The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky and Sam Cooper, for their helpful feedback and suggestions.
-%This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
-%Financial support towards hardware was provided by the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme. 
-%Travel support was provided by the Australian National University PhB Science program.
-%This work has been assigned LIGO document number P2000386.
+%\textbf{[Removed for double-anonymous review.]} 
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky and Sam Cooper, for their helpful feedback and suggestions.
+This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
+Financial support towards hardware was provided by the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme. 
+Travel support was provided by the Australian National University PhB Science program.
+This work has been assigned LIGO document number P2000386.
 

### 2021-05-06 paper/response/AJPRefereeResponse.tex
working on comments from Andrew on the response

@@
 \noindent We thank you and the referees for your helpful comments and suggestions. 
-We address the comments below in turn. 
-Ad we have made many changes throughout the manuscript, we have not highlighted them within the manuscript, instead we have pointed to each of the changes in our responses below.
+We agree with all the recommendations and address the comments below in turn. 
+As we have made many changes throughout the manuscript, we have not highlighted them within the manuscript, instead we have pointed to each of the changes in our responses below.
 
@@
 \item The grey paths in Figure 4 are those that have been rejected at each timestep. We have clarified this in the text and caption. 
-\item The hidden Markov model and Viterbi algorithm's strength is in it's ability to find randomly wandering signals.
+\item The hidden Markov model and Viterbi algorithm's strength is in its ability to find randomly wandering signals.
 The single constant tone in section III is recovered easily using a Fourier transform, however if the signal frequency changes over time, the signal power is spread across multiple frequency bins.
@@
 \begin{itemize}
-\item Optical microphone: we have rearranged the text so that the optical microphone concept is introduced in a separate paragraph and expanded our discussion of it to improve clarity.
-(Section V, paragraph 2)
+\item Optical microphone: we have rearranged the text so that the optical microphone concept is introduced in a separate paragraph and expanded our discussion of it to improve clarity (Section V, paragraph 2).
 \item The cloth screen was used to reduce the intensity of light on the photodiode and avoid saturating the analog-to-digital converter.
@@
 We have added a sentence to the end of Section V A to clarify.
+\han{clarify cloth screen given Andrew's comment}
 \item Figure 6: agreed, we have removed this figure and moved the photo to the Supplementary Material.
@@
 We and have also noted that Doppler modulation of the source also needs to be taken into account if searching for a neutron star in a binary.
+\han{gramma}
 \item It would indeed be very nice to modify the setup to show the Doppler effect. We have added some possible ways to explore this topic in our future work discussion in Section VI. 
@@
 \item We have clarified the explanation of the HMM in the text.
-The detection statistic captures the likelihood of a signal being present in each time-frequency bin given the observable data.
+The detection statistic captures the likelihood of a signal being present in each time-frequency bin conditional on the observed data.
 In Section IV A we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission matrix by name to reduce confusion.
+\han{from Andrew: p4, top point - "given the observable data" -> "conditional on the observed data" - and just mention that the conditionality is dropped from the notation [if true] to avoid clutter}
 \item[]
@@
 \begin{itemize}
-\item We do indeed assume that the noise is white in our analysis. We have added some text to clarify this assumption in the first paragraph of Section IV C.
+\item We do indeed assume that the noise is white in our analysis.
+This assumption is consistent with the use of Viterbi for continuous wave searches, where white noise is assumed even though LIGO noise is not white\han{todo get refs for this}.
+We have added some text to clarify this assumption in the first paragraph of Section IV C.
 \item[]
@@
 \begin{itemize}
+\item We agree and have expanded the text to include the suggested references (see below). 
 \item The hidden Markov model (HMM) and Viterbi algorithm is not a new approach for gravitational wave analysis.
@@
 The main focus of this study is the extraction of a signal from interferometer data using a series of analysis techniques, some of which are inspired by continuous gravitational wave searches.
+However we emphasize that both the simulated signal and the extraction approaches in the table-top experiment differ from real gravitational-wave experiments in many important ways, e.g.
+diurnal amplitude modulation, Doppler frequency modulation due to proper motion of the source, colored noise, nonstationary glitches in the operation of the interferometer, quantum effects in the measurement, and so on.
+The table-top system represents a simplified analogy of a research-level system and real signals.
 We also note that the use of sound as an analogy to gravitational-wave signals is not novel and is used by many gravitational-wave research groups to aid in their explanations (for example Sounds of Spacetime \texttt{https://www.

### 2021-05-06 paper/ifo-complexAudio.tex
working through comments from  Andrew

@@
 This analysis is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals. 
-Instead, the apparatus serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories. 
+In addition, the apparatus serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories. 
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
@@
 Most of the background noise is removed, however, the logMMSE does not significantly enhance the speech as the voice sounds muffled and indistinct.
+
 We find some improvement with music over speech. 

### 2021-05-06 paper/ifo-futureWork.tex
working through comments from  Andrew

@@
 These could perform the final stage of speech enhancement for the optical microphone.
-Interestingly, this coincidently connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
+Interestingly, this coincidentally connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
 Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.

### 2021-05-06 paper/ifo-wanderingTone.tex
working through comments from  Andrew

@@
 \end{comment}
+
+\han{from Andrew:
+sec IV, para 1, 1st sentence - seems to contradict sec III - I think what we really mean is "Although continuous grav wave signals are nearly monochromatic, the long observing times (~ 1 yr) mean that searches are sensitive to very small frequency drifts.
+A typical observation involves ~1e10 wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation."
+}
 
@@
 
+\han{from Andrew: sec IVA, first para - since ref mentioned other searches, it would be good to have 1-2 brief sentences accompanying the references which mention what those "other searches" are (broad categories, for instance, just so that there is some guidance for the reader)}
+
 The methods used here are  inspired by LIGO-Virgo continuous wave searches. 
@@
 A detection statistic relates the observed data to the hidden state and quantifies the likelihood of a signal being present in the data at each frequency and time bin.
-This likelihood is also called the emission matrix in gravitational-wave literature.  
+This likelihood is also called the emission probability in gravitational-wave literature.  
 In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
@@
 The timeseries data from the interferometer is first split into segments. 
-We then take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude $F(t_i,f_j)$ as the detection statistic (this is the emission matrix).
+We then take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude $F(t_i,f_j)$ as the detection statistic (this is the emission probability).
 Assuming Gaussian noise, this detection statistic choice maximizes the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.

### 2021-05-06 paper/ifo-wanderingTone.tex
comments from Andrew

@@
 
-\han{from Andrew:
-sec IV, para 1, 1st sentence - seems to contradict sec III - I think what we really mean is "Although continuous grav wave signals are nearly monochromatic, the long observing times (~ 1 yr) mean that searches are sensitive to very small frequency drifts.
-A typical observation involves ~1e10 wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation."
-}
-
-Unlike the signal considered in Section~\ref{sec:single_tone}, continuous gravitational waves may not be monochromatic.
+Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. 
+A typical observation involves $1 \times 10^{10}$ wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation. 
 A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
@@
 This may be explained by two anomalies in the recovered path.
-Initially, the injected signal wanders by more than one frequency bin per time bin (i.e.\ faster than the algorithm is allowed to), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
+Initially, the injected signal wanders by more than one frequency bin per time bin (i.e.\ faster than the our choice of $A(f_k,f_m)$ allows), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
 One may be tempted to increase the allowed frequency wander in the algorithm, however, this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.

### 2021-05-06 paper/paper_main.tex
rephrased abstract sentence about sup mat

@@
 % \jam{I changed all the \emph to ``'', feel free to change them back, but I think it is more explicit this way}
-In the accompanying Supplementary Material we consider a range of example filtering techniques to recover the original signal from noisy data and present highlight results here. 
+A range of filtering techniques used to recover signals from noisy data are detailed in the Supplementary Material. 
+Here we present result highlights using a combined notch plus Wiener filter and the statistical log minimum mean-square-error (logMMSE) estimator. 
 %We present a range of example filtering techniques to recover the original signal from noisy data, including a bandpass filter, notch filters, a Wiener filter, and the statistical log minimum mean-square-error (logMMSE) estimator.

### 2021-05-06 paper/ifo-appendix.tex
minor changes from read through
comments from Andrew
adding Changrong's code, updating README, misc comments from Andrew

@@
 The current build and sample data can be found at:
-\textbf{[Removed GitHub web-link for double-anonymous review.]}
-%\url{https://github.com/daccordeon/gravexplain}
+%\textbf{[Removed GitHub web-link for double-anonymous review.]}
+\url{https://github.com/daccordeon/gravexplain}
 % \begin{verbatim}
@@
 \begin{equation}
-s(t) = A \cos{\omega t + \phi}\,,
+s(t) = A \cos\left[{\omega t + \phi}\right]\,,
 \end{equation}
@@
 \end{eqnarray}
-Focusing on the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log-likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, baring the constant factor $\ip{x}{x}$.
+From the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log-likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.
+%baring the constant factor $\ip{x}{x}$.
 
@@
 %\subsection{Pseudocode}
-Here we give some pseudocode of the implementation used in Section~\ref{sec:viterbi}.
+Here we present some pseudocode of the implementation used in Section~\ref{sec:viterbi}.
 More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}.
-We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights, but to avoid numerical underflow we take the logarithm and consider them as additive weights.
+We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights.
+To avoid numerical underflow we take the logarithm and consider them instead as additive weights.
 
@@
 We restrict paths to only move one cell up or down at a time (or stay constant).
+For the edge cases of $j=0,N_f$, we only search over $k \in \{0,1\}$ and $k \in \{-1,0\}$, respectively, to stay inside the grid (this is not shown in the pseudocode).
 
@@
 
-	    	\State $Y_{i,j} \gets \underset{k \in \{-1,0,1\}}{\max} (Y_{i-1,j+k} + X_{i,j})$
-	    	\State $Z_{i,j} \gets j + \underset{k \in \{-1,0,1\}}{\arg\max} (Y_{i-1,j+k} + X_{i,j})$
+	    	\State $Y_{i,j} \gets X_{i,j} + \underset{k \in \{-1,0,1\}}{\max} (Y_{i-1,j+k})$
+	    	\State $Z_{i,j} \gets j + \underset{k \in \{-1,0,1\}}{\arg\max} (Y_{i-1,j+k})$
 

### 2021-05-06 paper/ifo-complexAudio.tex
adding Changrong's code, updating README, misc comments from Andrew

@@
 
+% https://en.wikipedia.org/wiki/Filter_(signal_processing)
 In signal processing, the ideal filter would be one that:
@@
 For speech intelligibility, this means that either: (i) some noise remains in the filtered recording, (ii) some of the speech content is lost as certain important frequencies are attenuated, or (iii) the speech is somewhat distorted in time.
-All three of these cases can, in extreme cases, worsen speech intelligibility beyond the unfiltered recording and so we choose filters that compromise between achieving the three conditions.
-% https://en.wikipedia.org/wiki/Filter_(signal_processing)
+All three of these cases can, when taken to the extreme, make the speech intelligibility worse than the unfiltered recording, i.e.\ the results can be worse than doing no filtering.
+Therefore, we choose filters that compromise between achieving the three conditions.
+% All three of these cases can, in extreme cases, worsen speech intelligibility beyond the unfiltered recording and so we choose filters that compromise between achieving the three conditions.
 

### 2021-05-06 source/noiseclear.m

@@ -1,0 +1,233 @@
+close all
+clc
+clear all
+
+% read sample files from optical microphone
+aa_track = readtable('aa_melatos.csv');
+aa = table2array(aa_track);
+% choose the the first second to analyze (sampling frequency 16kHz)
+
+fs = 16000;  % sampling frequency of photodiode
+aa = aa(1:fs,2);
+aa_inv =aa;
+Fs = 44100;  % default sampling frequency of Matlab
+% original voice sample (first second)
+[melatos_source,FS] = audioread('source_melatos.wav',[1,Fs]);
+melatos_source = melatos_source(:,1);
+%melatos_source = melatos_source(1:length(aa),1);   % keep one voice chanel and "a cathode"
+%audiowrite('melatos_source2.wav',melatos_source,fs)
+%[melatos_source2,FS2] = audioread('melatos_source2.wav');
+
+% pure system noise
+podo = readtable('podo_14_6.csv');
+podo1 = table2array(podo(1:fs,2));
+%%
+N = 15; % number of cascade notch filter
+% spectrum of original human voice
+T = 1/(FS); % sampling period
+L = length(melatos_source);
+figure(1)
+subplot(7,2,1)
+t = (0:L-1)*T;
+% source with time shift
+melatos_source_shift = [melatos_source(end-0.12*FS+1:end);melatos_source(1:end-0.12*FS)];
+plot(t,melatos_source_shift)
+title('Original voice')
+xlabel('time(s)')
+Y = fft(melatos_source,2^nextpow2(L));
+P2 = abs(Y/L);
+P1 = P2(1:L/2+1);
+P1(2:end-1) = 2*P1(2:end-1);
+f = FS*(0:(L/2))/L;
+
+subplot(7,2,2)
+plot(f,P1,'Linewidth',2) 
+title('Original voice (frequency domain)')
+xlabel('f (Hz)')
+ylabel('|P1(f)|')
+xlim([0,2500])
+%% spectrum of recorded voice (with sampling frequency 16KHz)
+
+T = 1/(fs); % sampling period
+L = length(aa);
+t = (0:L-1)*T;
[…]

### 2021-05-06 source/wienerFilt.m

@@ -1,0 +1,16 @@
+function [xest,B,MSE] = wienerFilt(x,y,N)
+
+X = 1/N .* fft(x(1:N));
+Y = 1/N .* fft(y(1:N));
+X = X(:);
+Y = Y(:);
+Rxx = N .* real(ifft(X .* conj(X))); % Autocorrelation function Rxx
+Rxy = N .* real(ifft(X .* conj(Y))); % Crosscorrelation function Rxy
+Rxx = toeplitz(Rxx); %[Rxx[0] Rxx[1]...;]
+Rxy = Rxy';
+B = Rxy / Rxx; B = B(:); % Wiener-Hopf eq. B = inv(Rxx) Rxy
+xest = fftfilt(B,x);
+xest = xest(N+1:end); % cut first N samples due to distorsion during filtering operation
+MSE = mean(y(N+1:end) - xest) .^2; % mean squared error
+
+end

### 2021-05-06 paper/ifo-supplementary-material.tex
responding to Andrew's comments on Supp. Material

@@
 
-\title{Supplementary material for \\Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
+\title{Supplementary Material for \\Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
 
@@
 However, the mains harmonics above $100\,{\rm Hz}$ remain.
-Using a high-pass filter with a higher cut-off can be used to mitigate this issue.
+A high-pass filter with a higher cut-off can be used to mitigate this issue.
 However, it makes the played-back signal unrecognizable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
@@
 	\caption{Butterworth bandpass filter frequency response.
-Any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power).
+Any amplitude below $-3\,{\rm dB}$ is significant attenuation (less than half the power at that frequency remains).
 The red, dashed vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$.
@@
 We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
-After filtering we see significant attenuation (at least 3~dB) of all frequencies outside the band, including around 100~dB near 0~Hz due to even stronger attenuation far from the band (this sharp drop-off at low frequencies would be more similar to the high-frequency (3~kHz) behaviour if shown against a logarithmic scale for frequency as is sometimes done).
+After filtering, we see significant attenuation (at least 3~dB) of all frequencies outside the band.
 However, the filter has little change on the harmonics within the band.
+The sharp drop-off at low frequencies, around 100~dB near 0~Hz, is due to stronger attenuation far from the band and would appear shallower if shown against a logarithmic scale for frequency as in Fig.~\ref{fig:butterworth}.
 }
@@
 where $\alpha$ and $\beta$ are parameters that control the filter. 
-In Eqn.~\ref{eqn:notch}, $w_0=\cos^{-1}(\beta)$ is the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$ is the bandwidth of the notch (which determines how quickly the response changes around the notched frequency).
+These parameters determine $w_0=\cos^{-1}(\beta)$, the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch, and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$, the bandwidth of the notch which determines how quickly the response changes around the notched frequency.
 We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centered on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$. 
@@
 \end{equation}
-If we assume that: i) $\textbf{x}$ is ``wide sense stationary''~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in R$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
+If we assume that: i) $\textbf{x}$ is ``wide sense stationary''~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in \mathbb{R}$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
 \begin{equation}

### 2021-05-06 paper/ifo-appendix.tex
another set of edits on all three documents

@@
 \begin{equation}
-I(t) = \lvert A e^{i (k L - \omega t)}  + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \rvert^2\,.
+I(t) = \left| A e^{i (k L - \omega t)}  + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \right|^2\,.
 \label{eqn:intensity1}    
@@
 %\subsection{Pseudocode}
-Here we present some pseudocode of the implementation used in Section~\ref{sec:viterbi}.
+Here, we present some pseudocode of the implementation used in Section~\ref{sec:viterbi}.
 More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}.
@@
 We restrict paths to only move one cell up or down at a time (or stay constant).
-For the edge cases of $j=0,N_f$, we only search over $k \in \{0,1\}$ and $k \in \{-1,0\}$, respectively, to stay inside the grid (this is not shown in the pseudocode).
+For the edge cases of $j=0,N_f$, we only search over $k \in \{0,1\}$ and $k \in \{-1,0\}$, respectively, to stay inside the grid (this is not shown in the pseudocode below).
 
@@
 Circuit diagram for reading the photodiode. 
-The photodiode is connected in reverse-bias across an op-amp. 
-The op-amp output is passed through a Sallen-Key anti-aliasing filter tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
-The digitized signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.
+The photodiode is connected in reverse-bias across an op-amp (pale-red box). 
+The op-amp output is passed through a Sallen-Key anti-aliasing filter (pale-blue box) tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
+The digitized signal is then read by the special-purpose input (SPI) pins of a Raspberry Pi in standard configuration.
 This diagram is produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online} and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.

### 2021-05-06 paper/ifo-complexAudio.tex
responding to Andrew's comments on Supp. Material
another set of edits on all three documents

@@
 We consider a selection of filter techniques, details of which can be found in the Supplementary Material and we refer the reader to Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045} for further information on digital signal processing.
-In Section~\ref{sec:opticalMicResults} we present the two best-performing analyses from our Supplementary Material.
+In Section~\ref{sec:opticalMicResults}, we present the two best-performing analyses from our Supplementary Material.
 
@@
 This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$ and thus can only ``hear'' frequencies below $15\,{\rm Hz}$.
-To overcome this issue, we use a photodiode\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the reverse direction.
+To overcome this issue, we use a photodiode~\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the reverse direction.
 As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.} at the output of the interferometer to achieve a sampling rate of $16\,{\rm kHz}$.
@@
 
-Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing\footnote{Folding of frequencies greater than half the sampling rate.} into the detected range.
+Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing~\footnote{Folding of frequencies greater than half the sampling rate.} into the detected range.
 We include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
-This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
+This component attenuates any frequencies above $8\,{\rm kHz}$ before they are digitally sampled.
 We also place a cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC -- an improvised, physical solution that could instead be replaced by scaling down the voltage electronically.
@@
 Environmental noise reduction is an active area of research for gravitational-wave detectors~\cite{EfflerEtAl:2015}.
-In some cases environmental noise sources can be removed or mitigated on site~\cite{CovasEtAl:2018}. 
+In some cases, environmental noise sources can be removed or mitigated on site~\cite{CovasEtAl:2018}. 
 Where this is not possible, noise subtraction can be done after data collection~\cite{DriggersEtAl:2019, DavisEtAl:2019}. 
-One example is removal of spectral noise lines from the United States power grid operating at $60\,{\rm Hz}$~\cite{VajenteEtAl:2020}.
+One example is the removal of spectral noise lines from the United States power grid operating at $60\,{\rm Hz}$~\cite{VajenteEtAl:2020}. %United States's?
 
@@
 \begin{center}
-\includegraphics[width=0.95\textwidth]{figures/combined_highlight_results_melatos_labelled.pdf}
+\includegraphics[width=0.8\textwidth]{figures/combined_highlight_results_melatos_labelled.pdf}
 \caption{\label{fig:notchWienerLogMMSEResults}
@@
 
-In this section, we present results of two advanced signal processing analysis techniques applied to the optical microphone recordings. 
+In this section, we present the results of two advanced signal processing analysis techniques applied to the optical microphone recordings. 
 The techniques are only briefly described here and we refer the reader to the accompanying Supplementary Material for further details and some other analysis techniques.
@@
 Secondly, we apply a speech enhancement technique. 
-Ref.~\cite{SubjectiveComparison} compares $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech (see also Ref.~\cite{Ephraim1984SpeechEU_logMMSE} for details on the logMMSE estimator).
+Ref.~\cite{SubjectiveComparison} compares $13$ speech enhancement methods, finding the log minimum mean-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech (see also Ref.~\cite{Ephraim1984SpeechEU_logMMSE} for details on the logMMSE estimator).
 We use an existing implementation of the logMMSE from~\cite{logmmse}.  

### 2021-05-06 paper/ifo-conclusions.tex
another set of edits on all three documents

@@
 We then consider more complex audio (music and speech) for a more general interest in signal processing and explore the use of the interferometer as an optical microphone. 
-A variety of example activities are considered in the Supplementary Material. 
+A variety of example activities is considered in the Supplementary Material. 
 We find that the logMMSE produces the best results, however, it does not enhance the recording to an intelligible level. 
@@
 The demonstration presented has the potential to be adapted for use in the undergraduate laboratory.
-It can be used to teach topics of interest to physics and electrical engineering students including, interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
+It can be used to teach topics of interest to physics and electrical engineering students including interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
 The increased excitement and public interest in the field of gravitational-wave research in recent years mean that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.

### 2021-05-06 paper/ifo-supplementary-material.tex
another set of edits on all three documents

@@
 We finish by combining these with advanced statistical techniques (Wiener filter and the logMMSE estimator). 
-As an illustrative example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
-For more information about the techniques used here see also Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045}.
+As an example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
+For more information about the techniques used here see Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045}.
 
@@
 
-All analysis is done using the photodiode set-up as described in Section~V.A. in the main text. 
-A photograph of the equipment is shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit} with the interferometer and circuit shown on the left and right respectively.
+All analysis is done using the photodiode set-up as described in Section~V.A. in the main article. 
+A photograph of the equipment is shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit} with the interferometer and circuit shown on the left and right, respectively.
 
@@
 The first and second rows in Fig.~\ref{fig:op_mic_results} show the original source signal and the raw optical microphone recording, respectively. 
-The following sections describe a sequence of analysis techniques and a selection of results are presented in Fig.~\ref{fig:op_mic_results}.
+The following sections describe a sequence of analysis techniques and a selection of their results is presented in Fig.~\ref{fig:op_mic_results}.
 
@@
 
-An intuitive way to remove noise is by directly subtracting the background noise from the recorded spectrum. 
+The first technique we consider is quite intuitive: we try to remove noise by directly subtracting the background noise from the recorded spectrum. 
 The third row in Fig.~\ref{fig:op_mic_results} shows the interferometer background noise when there is no input signal.
@@
 
-Simply zeroing the frequency bins corresponding to the harmonics of the $50\,{\rm Hz}$ mains signal is unsuccessful. 
+Since the mains harmonics are present in the signal, we try to selectively remove them.
+However, simply zeroing the frequency bins corresponding to the harmonics of the $50\,{\rm Hz}$ mains signal is unsuccessful. 
 This effectively multiplies the spectrum by a rectangular comb filter. 
@@
 In Eqn.~\ref{eq:butterworth}, $f_c$ is the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ is the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
-Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here, a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}$, $3\,{\rm kHz})$.
-This high frequency cut-off is chosen since the frequencies important for speech generally lie below $2\,{\rm kHz}$~\cite{speech_intelligibility}.
+Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here: a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}$, $3\,{\rm kHz})$.
+This high frequency ($3\,{\rm kHz}$) cut-off is chosen since the frequencies important for speech generally lie below $2\,{\rm kHz}$~\cite{speech_intelligibility}.
 
@@
 The Butterworth band-pass filter reduces the amplitude of mains harmonics below $150\,{\rm Hz}$ and suppresses unrelated noise sources above $3\,{\rm kHz}$.
-However, it does not address the issue of mains harmonics above $150\,{\rm Hz}$ (i.e., in the pass-band).
+However, it does not address the issue of mains harmonics above $150\,{\rm Hz}$ (i.e.\ those in the pass-band).
 In the following section, we experiment with a cascade notch filter to address this.
@@
 We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centered on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$. 
-We choose the bandwidth and order (here equal to six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
+We choose the bandwidth and order (here order six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
 The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,
@@
 The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row in Fig.~\ref{fig:op_mic_results}.
-We see that the mains harmonics are significantly attenuated compared to the background subtracted results shown in the fourth row in Fig.~\ref{fig:op_mic_results}.
+We see that the mains harmonics are significantly attenuated compared to the background-subtracted results shown in the fourth row in Fig.~\ref{fig:op_mic_results}.
 
 Although the notch filter removes much of the mains hum sound, the filtered recording is not intelligible.
-Qualitatively it sounds more like a drum than a human voice.
+Qualitatively, it sounds more like a drum than a human voice.
 This is due to the loss of voice information under the filter.
@@
 \end{equation}
-Given $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean-square-error (BMSE) between the two, defined as 
+Given $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean-square error (BMSE) between the two, defined as 
 \begin{equation}
@@
 Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high and attenuates the signal where the SNR is low.
-The causal Wiener filter (that only makes estimates based on past observations) is similar.
+The causal Wiener filter (which only makes estimates based on past observations) is similar.
 A more detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.
@@
 Here, we experiment with applying a combination of the cascaded notch and the Wiener filter to the recorded speech signal. 
-The results of this analysis also are described in Section V.C. in the main text.
+The results of this analysis also are described in Section V.C. in the main article.
 
@@
 Speech enhancement of noisy channels is a classic problem in signal processing. 
-In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square-error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
-This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimizes the mean-square-error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that it measures the MSE between the logarithm of the Fourier amplitudes.
+In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
+This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimizes the mean-square error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that it measures the MSE between the logarithm of the Fourier amplitudes.
 This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
@@
 Again there is significant noise reduction, however, the speech remains indistinct.
-The logMMSE results are also presented in Section V.C. in the main text.
+The logMMSE results are also presented in Section V.C. in the main article.
 

### 2021-05-06 paper/ifo-tableTopGWs.tex
another set of edits on all three documents

@@
 However, their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
-In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms (see Fig.~\ref{fig:ifo_schematic_webcam}). 
+In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms, shown in Fig.~\ref{fig:ifo_schematic_webcam}. 
 Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern.
 The resulting interference pattern is dependent on the relative distance traveled by the beams. 
-Current generation gravitational-wave observatories have kilometer-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo respectively). 
-They can measure minuscule changes in distance due to gravitational waves, e.g.
-the first detection of a binary black hole merger produced a strain of $10^{-21}$~\cite{GW150914}, corresponding to a change in distance which is a fraction of the width of a proton.
+Current generation gravitational-wave observatories have kilometer-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo, respectively). 
+They can measure minuscule changes in distance due to gravitational waves, e.g.\ the first detection of a binary black hole merger produced a strain of $10^{-21}$~\cite{GW150914}, corresponding to a change in distance which is a fraction of the width of a proton.
 
@@
 The output of a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO} (top-left) is reflected from a mirror at a right angle.
-The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter which is $50$/$50$ in reflection and transmission.
+The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter which is approximately $50$/$50$ in reflection and transmission.
 The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and recombine at the beamsplitter. 

### 2021-05-06 paper/ifo-wanderingTone.tex
another set of edits on all three documents

@@
 The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
-In section~\ref{sec:realCWSearches} we briefly review the analysis method used by LIGO and Virgo, in Section~\ref{sec:viterbi} we describe the methods as applied in this work, and in section~\ref{sec:wanderingResults} we show results for recovering a wandering frequency signal using the table-top interferometer.
+In Section~\ref{sec:realCWSearches}, we briefly review the analysis method used by LIGO and Virgo, in Section~\ref{sec:viterbi}, we describe the methods as applied in this work, and in Section~\ref{sec:wanderingResults}, we show results for recovering a wandering frequency signal using the table-top interferometer.
 
@@
 \end{eqnarray}
-The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises $Pr[ f(t_0), f(t_1),... f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
+The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises ${\rm Pr}[ f(t_0), f(t_1),... f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
 We note that there is no marginalization used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
@@
 At every $t_i$ all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}). 
-Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}.%, where the circles indicate the elements in the spectrogram with the lighter colors corresponding to a higher likelihood.
+Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}.
+%, where the circles indicate the elements in the spectrogram with the lighter colors corresponding to a higher likelihood.
 The implementation used in this work is available online (see Appendix~\ref{app:code}).
@@
 The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path, respectively. 
-On the left, before $\sim 15\,{\rm s}$, the signal changes frequency too quickly for the Viterbi algorithm to recover, given a parameter for the rate of change of one frequency bin per time bin. 
-At $150\,{\rm s}$ the data appears anomalous, which may be due to some transient background noise. }
+On the left, before $\sim 15\,{\rm s}$, the signal changes frequency too quickly for the Viterbi algorithm to recover, given that the algorithm is restricted to only change by one frequency bin per time bin.
+At $150\,{\rm s}$, the data appears anomalous, which may be due to some transient background noise. }
 \end{figure*}
@@
 This may be explained by two anomalies in the recovered path.
-Initially, the injected signal wanders by more than one frequency bin per time bin (i.e.\ faster than the our choice of $A(f_k,f_m)$ allows), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
+Initially, the injected signal wanders by more than one frequency bin per time bin (i.e.\ faster than our choice of $A(f_k,f_m)$ allows), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
 One may be tempted to increase the allowed frequency wander in the algorithm, however, this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.

### 2021-05-06 paper/paper_main.tex
another set of edits on all three documents

@@
 A range of filtering techniques used to recover signals from noisy data are detailed in the Supplementary Material. 
-Here we present result highlights using a combined notch plus Wiener filter and the statistical log minimum mean-square-error (logMMSE) estimator. 
+Here, we present highlights of our results using a combined notch plus Wiener filter and the statistical log minimum mean-square error (logMMSE) estimator. 
 %We present a range of example filtering techniques to recover the original signal from noisy data, including a bandpass filter, notch filters, a Wiener filter, and the statistical log minimum mean-square-error (logMMSE) estimator.
 % log minimum mean-square-error (logMMSE) estimator
-Recordings of simple chords and drums are easily recovered, but complex music and speech are more challenging. 
+Recordings of simple chords and drums are easily recovered using these techniques, but complex music and speech are more challenging. 
 This demonstration can be used by educators in undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
@@
 %\textbf{[Removed for double-anonymous review.]} 
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky and Sam Cooper, for their helpful feedback and suggestions.
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky, and Sam Cooper, for their helpful feedback and suggestions.
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
@@
 
-\nocite{*}
+% \nocite{*}
 \bibliographystyle{myunsrt}

### 2021-05-06 paper/response/AJPRefereeResponse.tex

@@
 \noindent We thank you and the referees for your helpful comments and suggestions. 
-We agree with all the recommendations and address the comments below in turn. 
-As we have made many changes throughout the manuscript, we have not highlighted them within the manuscript, instead we have pointed to each of the changes in our responses below.
+We agree with all the recommendations and address the comments below in turn. % surely, we only agree with ``most of'' the recommendations?
+As we have made many changes throughout the manuscript, we have not highlighted them within the manuscript, instead, we have pointed to each of the changes in our responses below.
 
@@
 \item We have moved the footnotes to the bibliography
-\item We have checked through the manuscript to correct spellings to US English. 
+\item We have checked the manuscript to correct spellings to US English. 
 \item[]
@@
 \begin{itemize}
-\item Markov process: We have added a sentence to explain Markov process in Section IV A, ``In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal).''
+\item Markov process: We have added a sentence to explain Markov processes in Section IV A, ``In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal).''
 \item Figure 4: We have lightened the color scheme used in the Figure 4 schematic to aid with black and white printing.
 The colors represent a gradient rather than a discrete shading and we now use a blue/green color palette to clarify this.
-Our intention is that the Figure 4 schematic be interpreted as a simplified version of the Figure 6 (previously Figure 5) spectrogram result.
-To clarify this we have made the colour scheme of the schematic more consistent with that of Figure 6 and we have referred back to the schematic when we first introduce Figure 6.
+We intend that the Figure 4 schematic be interpreted as a simplified version of the Figure 6 (previously Figure 5) spectrogram result.
+To clarify this we have made the color scheme of the schematic more consistent with that of Figure 6 and we have referred back to the schematic when we first introduce Figure 6.
 \item The grey paths in Figure 4 are those that have been rejected at each timestep. We have clarified this in the text and caption. 
@@
 The single constant tone in section III is recovered easily using a Fourier transform, however if the signal frequency changes over time, the signal power is spread across multiple frequency bins.
-We have added a new figure to illustrate this (Figure 4) and some text at the start of Section IV (paragraph 2) and at the end of Section IV B to explain that the Fourier transform of the whole dataset is not well suited to this type of signal.
+We have added a new figure to illustrate this (Figure 4) and some text at the start of Section IV (paragraph 2) and the end of Section IV B to explain that the Fourier transform of the whole dataset is not well suited to this type of signal.
 \item We have also revised the entire Section IV to improve clarity in our Viterbi algorithm explanation and motivation.
@@
 \item Comparison without VA: Without using the Viterbi algorithm the signal becomes spread out over the frequency bins and might not rise above the noise.
-We have now explained this and added a plot that shows the Fourier spectrum of the entire timeseries to illustrate that the Fourier transofrm is not well suited here.
+We have now explained this and added a plot that shows the Fourier spectrum of the entire timeseries to illustrate that the Fourier transform is not well suited here.
 (See also our answer to point 3 above).
@@
 Our work instead shows the Fourier transform at each time-frequency grid point in the spectrogram.
-We have decided not to draw comparisons to the Q-transform plots in our manuscript order to avoid confusion.
+We have decided not to draw comparisons to the Q-transform plots in our manuscript to avoid confusion.
 \item[]
@@
 \begin{itemize}
-\item The Butterworth filter has strongly attenuated the signal outside the passband, this sharp drop-off at low frequencies would appear shallower when shown against a logarithmic scale for frequency as is sometimes done.
+\item The Butterworth filter has strongly attenuated the signal outside the passband, this sharp drop-off at low frequencies would appear shallower when shown against a logarithmic scale for frequency as in Fig.~3 in the Supplementary Material.
 A note has been made in the Supplementary Material where this plot has been moved to.
@@
 \item The filters are performed separately after the data is gathered. We have added a sentence to the figure caption to clarify this. 
-\item Most of the code is readily available on the provided GitHub with the exception of the Wiener filtering results.
-The provided code is documented and commented to ease reproduceability.
-For the purposes of the double-blind peer review, we removed the link to our git repository from the manuscript.
-\todo{[todo] Ask Changrong whether we can include scripts in the git repo}
+\item All of the code is readily available on the provided GitHub.
+The provided code is documented and commented to ease reproducing the results.
+For the double-blind peer review, we removed the link to our git repository from the manuscript.
 \item[]
@@
 \begin{itemize}
-\item We have including some details on how the detection statistic is used to take account of the Earth's motion around the Sun in Section IV A.
-We and have also noted that Doppler modulation of the source also needs to be taken into account if searching for a neutron star in a binary.
-\han{gramma}
+\item We have including some details on how the detection statistic is used to account for the Earth's motion around the Sun in Section IV A.
+We have also noted that Doppler modulation of the source also needs to be accounted for when searching for a neutron star in a binary.
 \item It would indeed be very nice to modify the setup to show the Doppler effect. We have added some possible ways to explore this topic in our future work discussion in Section VI. 
@@
 \item We do indeed assume that the noise is white in our analysis.
-This assumption is consistent with the use of Viterbi for continuous wave searches, where white noise is assumed even though LIGO noise is not white\han{todo get refs for this}.
+This assumption is consistent with the use of Viterbi for continuous-wave searches, where white noise is assumed even though LIGO noise is not white\han{todo get refs for this}.
 We have added some text to clarify this assumption in the first paragraph of Section IV C.
@@
 We note that many of the papers listed above are about searching for transient signals rather than the persistent signals considered here.
-\item We have added a new subsection (Section IV A in the revised draft) to emphasise that these techniques are not new and that they are inspired by real continuous wave searches.
-In this section we have pointed the reader to the search papers and methods papers that use HMM and Viterbi.
-We have also added references to other (non-Viterbi) continuous wave search methods.
-Finally we point the interested reader, who would like to know more about chirp or transient searches, to the papers suggested above.
+\item We have added a new subsection (Section IV A in the revised draft) to emphasize that these techniques are not new and that they are inspired by real continuous wave searches.
+In this section, we have pointed the reader to the search papers and methods papers that use HMM and Viterbi.
+We have also added references to other (non-Viterbi) continuous-wave search methods.
+Finally, we point the interested reader, who would like to know more about chirp or transient searches, to the papers suggested above.
 \item[]
@@
 \item Length of Section V: Agreed, we have moved the majority of Section V C to Supplementary Material and now only include two highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator).
-\item IIR, FIR, and Z-transform: In the Supplementary Material we have expanded the description of the impulse response of a filter mentioning both IIR and FIR.
+\item IIR, FIR, and Z-transform: In the Supplementary Material, we have expanded the description of the impulse response of a filter mentioning both IIR and FIR.
 We have also added some text clarifying that equation 2 in the Supplementary Material (previously equation 4 in the main text) is a Z-transform.
@@
 \begin{itemize}
-\item Agreed we should mention this work.
+\item Agreed, we should mention this work.
 We have added a short paragraph at the end of section V B where we discuss environmental noise with the table-top interferometer.
@@
 \item Use of audio signals: We agree that there is indeed a fundamental difference between sound waves and gravitational waves.
-In the Section II we have clarified that we are using sound as an analogy.
-For the purposes of this table-top experiment, we require a signal that behaves similar to gravitational waves in that it alters the interference pattern produced by a Michelson interferometer.
-Using a speaker and audio signals provides a simple and relatively cost effective means to do this, compared to for example, a piezo-electric crystal fixed to the back of a mirror as used in some gravitational wave demonstrations (e.g.
+In Section II, we have clarified that we are using sound as an analogy.
+For this table-top experiment, we require a signal that behaves similar to gravitational waves in that it alters the interference pattern produced by a Michelson interferometer.
+Using a speaker and audio signals provides a simple and relatively cost-effective means to do this, compared to, for example, a piezo-electric crystal fixed to the back of a mirror as used in some gravitational wave demonstrations (e.g.
 Cooper et al., arXivv e-prints 2004.03052, 2020).
-The main focus of this study is the extraction of a signal from interferometer data using a series of analysis techniques, some of which are inspired by continuous gravitational wave searches.
+The main focus of this study is the extraction of a signal from interferometer data using a series of analysis techniques, some of which are inspired by continuous gravitational-wave searches.
 However we emphasize that both the simulated signal and the extraction approaches in the table-top experiment differ from real gravitational-wave experiments in many important ways, e.g.
@@
 The table-top system represents a simplified analogy of a research-level system and real signals.
-We also note that the use of sound as an analogy to gravitational-wave signals is not novel and is used by many gravitational-wave research groups to aid in their explanations (for example Sounds of Spacetime \texttt{https://www.
-soundsofspacetime.org/} and Black Hole Hunter \texttt{blackholehunter.org})
+We also note that the use of sound as an analogy to gravitational-wave signals is not novel and is used by many gravitational-wave research groups to aid in their explanations (for example Sounds of Spacetime \texttt{https://www.soundsofspacetime.org/} and Black Hole Hunter \texttt{blackholehunter.org})
 \item[]

### 2021-05-06 paper/ifo-futureWork.tex

@@
 
-We could also use more traditional techniques such as wavelet transform (WT) \citep{nason1995stationary} to extract the signal from noise, and compare these with the above methods.
-WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time.
+We could also use more traditional techniques such as a wavelet transform (WT) \citep{nason1995stationary} to extract the signal from noise, and compare these with the above methods.
+A WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time.
 In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT methods are proposed for speech recognition.

### 2021-05-06 paper/ifo-introduction.tex

@@
 Gravitational waves are a prediction of Albert Einstein's theory of General Relativity~\cite{Einstein:1916}, 
-They are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
+they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
 The effect of gravitational waves is a change in lengths: a ``stretching and squashing'' of the distance between objects. 
@@
 %This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world.
-This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online,chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
+This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online, chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
 These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.

### 2021-05-06 paper/ifo-singleTone.tex
minor changes from read through
working through comments from  Andrew

@@
 \begin{figure}
-	\includegraphics[width=.45\textwidth]{figures/webcam_spectrum_expt_4_0209.pdf}
+	\includegraphics[width=.5\textwidth]{figures/webcam_spectrum_expt_4_0209.pdf}
 	\caption{\label{fig:webcam_spectrum}
-Recovery of a tone at a constant frequency: Here we plot the Fourier amplitude of the intensity pattern against frequency.
+Recovery of a tone at a constant frequency: Here, we plot the Fourier amplitude of the intensity pattern against frequency.
 The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has an FWHM of $0.033\,{\rm Hz}$.
@@
 % rearranged as I found this section to not switch between method and results too much, I've tried this order now: intro, webcam and data specifics, injection, and results. 
-
-Continuous-wave searches look for nearly monochromatic signals. In this section, we consider a simple sinusoidal tone at a single, constant frequency. %should be a single constant frequency *signal*?
+Continuous-wave searches look for nearly monochromatic signals~\cite{JKS:1998}.
+In this section, we consider a simple sinusoidal tone at a single, constant frequency.
+%should be a single constant frequency *signal*?
 

### 2021-05-06 paper/ifo-complexAudio.tex
adding text and references for active/passive isolation and noise sources in gw observatories

@@
 
-Environmental noise reduction is an active area of research for gravitational-wave detectors~\cite{EfflerEtAl:2015}.
+Environmental noise reduction for gravitational-wave detectors is an active area of research~\cite{EfflerEtAl:2015}. 
+One such noise source is due to seismic ground motion from the Earth. 
+In gravitational-wave observatories, the optical components are suspended from a series of pendulums to achieve passive isolation, and feedback control loops provide active isolation\cite{MatichardEtAl:2015, AdvancedVirgo:2015}.
+Such a system is beyond the scope of this article, however feedback control has been demonstrated in table-top Michelson interferometers designed for the undergraduate laboratory~\cite{LibbrechtBlack:2015}.
+Other sources of noise include acoustic sound waves (for example from nearby vehicles, wind, and components such as electric fans) and magnetic noise (for example from high-voltage power lines, building heaters and lights, or any electrical power circuit)~\cite{EfflerEtAl:2015,CironeEtAl:2019}.
 In some cases, environmental noise sources can be removed or mitigated on site~\cite{CovasEtAl:2018}. 

### 2021-05-06 paper/ifo-introduction.tex
adding text and references for active/passive isolation and noise sources in gw observatories

@@
 These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
-Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
+Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019,LibbrechtBlack:2015} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
 

### 2021-05-11 paper/ifo-complexAudio.tex
working on comments, tidying refs

@@
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
-We consider a selection of filter techniques, details of which can be found in the Supplementary Material and we refer the reader to Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045} for further information on digital signal processing.
+We consider a selection of filter techniques, details of which can be found in the Supplementary Material and we refer the reader to Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045, DigitalProcingOfSpeechSignals:1978} for further information on digital signal processing.
 In Section~\ref{sec:opticalMicResults}, we present the two best-performing analyses from our Supplementary Material.

### 2021-05-11 paper/ifo-wanderingTone.tex
removing comment
working on comments, tidying refs

@@
 
-\han{from Andrew: sec IVA, first para - since ref mentioned other searches, it would be good to have 1-2 brief sentences accompanying the references which mention what those "other searches" are (broad categories, for instance, just so that there is some guidance for the reader)}
+%\han{from Andrew: sec IVA, first para - since ref mentioned other searches, it would be good to have 1-2 brief sentences accompanying the references which mention what those "other searches" are (broad categories, for instance, just so that there is some guidance for the reader)}
 
@@
 For further details on these methods and searches see Refs.~\cite{SuvorovaEtAl:2016, SuvorovaEtAl:2017, ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, BeniwalEtAl:2021, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, BayleyEtAlSOAP:2019, SunEtAlSNR:2018}.
-For other continuous wave search analyses not discussed here, see for example Refs.\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, MSPSearchO3O2O1:2020, RadiometerO1O2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} and for further information on searching for transient gravitational wave signals, see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
+For a range of other continuous wave searches we refer the reader to Refs.~\cite{MSPSearchO3O2O1:2020, RadiometerO1O2:2019, KnownPulsarsO2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} for searches that consider particular targets or sky locations, and to Refs.~\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, StackSlide:2012} for search methods that cover the entire sky.
+For further information on searching for transient gravitational wave signals see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
 

### 2021-05-11 paper/response/AJPRefereeResponse.tex
removing comments

@@
 We have added a sentence to the end of Section V A to clarify.
-\han{clarify cloth screen given Andrew's comment}
+%\han{clarify cloth screen given Andrew's comment}
 \item Figure 6: agreed, we have removed this figure and moved the photo to the Supplementary Material.
@@
 The detection statistic captures the likelihood of a signal being present in each time-frequency bin conditional on the observed data.
-In Section IV A we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission matrix by name to reduce confusion.
-\han{from Andrew: p4, top point - "given the observable data" -> "conditional on the observed data" - and just mention that the conditionality is dropped from the notation [if true] to avoid clutter}
+In Section IV A we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission matrix by name to reduce confusion.% \han{from Andrew: p4, top point - "given the observable data" -> "conditional on the observed data" - and just mention that the conditionality is dropped from the notation [if true] to avoid clutter}
 \item[]
@@
 \item We do indeed assume that the noise is white in our analysis.
-This assumption is consistent with the use of Viterbi for continuous-wave searches, where white noise is assumed even though LIGO noise is not white\han{todo get refs for this}.
+This assumption is consistent with the use of Viterbi for continuous-wave searches, where white noise is assumed even though LIGO noise is not white%\han{todo get refs for this}.
 We have added some text to clarify this assumption in the first paragraph of Section IV C.

### 2021-05-11 paper/response/AJPRefereeResponse.tex
adding reference for noise assumption in LIGO data for HMM

@@
 \item We do indeed assume that the noise is white in our analysis.
-This assumption is consistent with the use of Viterbi for continuous-wave searches, where white noise is assumed even though LIGO noise is not white%\han{todo get refs for this}.
+This assumption is consistent with the use of Viterbi for continuous-wave searches, where white noise is assumed even though LIGO noise is not white (for example, see Suvorova et al.
+Phys.
+Rev.
+D, Vol 96, Issue 10, 102006 (2017).
 We have added some text to clarify this assumption in the first paragraph of Section IV C.

### 2021-05-11 paper/response/AJPRefereeResponse.tex
minor change

@@
 \item We have clarified the explanation of the HMM in the text.
-The detection statistic captures the likelihood of a signal being present in each time-frequency bin conditional on the observed data.
-In Section IV A we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission matrix by name to reduce confusion.% \han{from Andrew: p4, top point - "given the observable data" -> "conditional on the observed data" - and just mention that the conditionality is dropped from the notation [if true] to avoid clutter}
+The detection statistic $F(t_i, f_j)$ captures the likelihood of a signal being present in each time-frequency bin conditional on the observed data, where we have dropped the conditionally from the notation to avoid clutter.
+In Section IV A we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission matrix by name to reduce confusion.
 \item[]

### 2021-05-20 paper/paper_main.tex
getting manuscript ready for resubmission (removing identifying information for double anoymous review)

@@
 
-%\author{\textbf{[Authors removed for double-anonymous review.]}}
+\author{\textbf{[Authors removed for double-anonymous review.]}}
+
+\begin{comment}
 
@@
 Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
 }
@@
 
+\end{comment}
+
 % Summer 2019/2020
@@
 We also explore the use of the interferometer as a teaching tool for educators in physics and electrical engineering by using it as an ``optical microphone'' to capture music and speech. 
-% \jam{We mention undergraduate labs twice, roll this sentence into the end? Also, electrical sounds weird, why the change from electrical engineering?}
-%The interferometer is used as an ``optical microphone'' to capture music and speech. 
-% \jam{I changed all the \emph to ``'', feel free to change them back, but I think it is more explicit this way}
 A range of filtering techniques used to recover signals from noisy data are detailed in the Supplementary Material. 
 Here, we present highlights of our results using a combined notch plus Wiener filter and the statistical log minimum mean-square error (logMMSE) estimator. 
-%We present a range of example filtering techniques to recover the original signal from noisy data, including a bandpass filter, notch filters, a Wiener filter, and the statistical log minimum mean-square-error (logMMSE) estimator.
-% log minimum mean-square-error (logMMSE) estimator
 Recordings of simple chords and drums are easily recovered using these techniques, but complex music and speech are more challenging. 
 This demonstration can be used by educators in undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
-% This demonstration has scope for use by educators in undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
-% rearranged in response to Anna's pnp review
-%Finally we use the interferometer as an `optical microphone' to capture music and speech and apply filtering techniques to recover the original signals from the noisy data.
-Such techniques include simple bandpass filtering, notch filters, a Weiner filter, and the statistical logMMSE method.
-%Simple chords and drums are easily recovered, but complex music and speech are more challenging. 
-%This demonstration has scope for use in physics and electrical engineering undergraduate laboratories. 
-%It can also be adapted for use as an engagement tool for communicating gravitational wave and signal-processing topics to non-specialist audiences. 
 \end{abstract}
@@
 \begin{acknowledgments}
-%\textbf{[Removed for double-anonymous review.]} 
+\textbf{[Removed for double-anonymous review.]}
+
+\begin{comment}
 The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky, and Sam Cooper, for their helpful feedback and suggestions.
@@
 This work has been assigned LIGO document number P2000386.
+\end{comment}
 

### 2021-05-20 paper/response/AJPResubmissionCoverLetter.tex
renaming to be cover letter

\documentclass{article}

\usepackage{fullpage}
\usepackage{xcolor}

\begin{document}

\noindent Dear Joseph Amato, \\

\noindent We thank you and the referees for your helpful comments and suggestions. 
We agree with all the recommendations and address the comments below in turn.
As we have made many changes throughout the manuscript, we have not highlighted them within the manuscript, instead, we have pointed to each of the changes in our responses below.

\subsection*{Comments from editor, Joseph Amato}
This is an addendum to my decision letter and the two reviews of your manuscript.  I am satisfied that the reviewers read your work carefully, and that their recommendations are fair, thoughtful, and potentially useful.  While I sympathize with R1’s concern that the manuscript is better suited to a research journal, I think it could be revised to make a valuable contribution to AJP, which, as you know, is geared toward pedagogy.  So as I read your manuscript, I recorded things that were not sufficiently clear to me, supposing that the same things might not be clear to many AJP readers who might be very interested in duplicating your apparatus.  Here are some of my concerns. \\

\noindent 1. Simple things: Convert footnotes to endnotes, blending them in with the bibliography.  Also, use US English (color, not colour; defense, not defence, etc.). 
\begin{itemize}
\item We have moved the footnotes to the bibliography
\item We have checked the manuscript to correct spellings to US English. 
\item[]
\end{itemize}

\noindent 2. Section III.  How was the Fourier transform generated?  (software?)
\begin{itemize}
\item Yes, the Fourier transform was generated using software. We have added a sentence to Section III (``The Discrete Fourier transform was calculated using the NumPy package for Python.''), a citation (to the creators of NumPy), and a footnote (to the documentation).
\item[]
\end{itemize}

\noindent 3. Section IV.  The term “Markov process” may not be familiar to readers.  Figure 4: The three shades (or are there four?) indicating probability do not show up well in color, and worse, will be indistinguishable in black/white.  I’m guessing that’s the reason I didn’t understand why certain paths were grey and not black.  While I applaud your efforts to explain the Viterbi algorithm (VA), I still think there’s room for improvement.  I don’t understand how it improves the signal to noise ratio. 
\begin{itemize}
\item Markov process: We have added a sentence to explain Markov processes in Section IV A, ``In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal).''
\item Figure 5 schematic (previously Figure 4): We have lightened the color scheme used in the Figure 5 schematic to aid with black and white printing. The colors represent a gradient rather than a discrete shading and we now use a blue/green color palette to clarify this. We intend that the Figure 5 schematic be interpreted as a simplified version of the Figure 6 (previously Figure 5) spectrogram result. To clarify this we have made the color scheme of the schematic more consistent with that of Figure 6 and we have referred back to the schematic when we first introduce Figure 6. 
\item The grey paths in Figure 5 are those that have been rejected at each timestep. We have clarified this in the text and caption. 
\item The hidden Markov model and Viterbi algorithm's strength is in its ability to find randomly wandering signals. The single constant tone in section III is recovered easily using a Fourier transform, however if the signal frequency changes over time, the signal power is spread across multiple frequency bins. We have added a new figure to illustrate this (Figure 4) and some text at the start of Section IV (paragraph 2) and the end of Section IV B to explain that the Fourier transform of the whole dataset is not well suited to this type of signal.
\item We have also revised the entire Section IV to improve clarity in our Viterbi algorithm explanation and motivation.
\item[]
\end{itemize}

\noindent 4. Figure 5 is described as a heatmap.  Please explain, in the caption, what the color coding indicates.  Can you compare what is shown to what would be detected without using the VA?  Is this ``heatmap'' the same representation as is often shown in LIGO publications?
\begin{itemize}
\item Figure 6 (previously Figure 5) colors: The color coding represents the logarithm of the absolute value of the discrete Fourier transform. We have added this to the caption.
\item Comparison without VA: Without using the Viterbi algorithm the signal becomes spread out over the frequency bins and might not rise above the noise. We have now explained this and added a new figure (Figure 4) that shows the Fourier spectrum of the entire timeseries to illustrate that the Fourier transform is not well suited here. (See also our answer to point 3 above).
\item Heatmap representation: The heatmap is similar to the figures in LIGO-Virgo-KAGRA publications which show transient gravitational-wave signals, but not identical. Those figures normally show Q-transforms (described in Chatterji, et al., Classical Quantum Gravity 21, S1809 2004). Our work instead shows the Fourier transform at each time-frequency grid point in the spectrogram. We have decided not to draw comparisons to the Q-transform plots in our manuscript to avoid confusion.
\item[]
\end{itemize}

\noindent 5. Section V.  What do you mean by an optical microphone device?  “Cloth screen” over the photodiode?  Fig. 6 is unnecessary; the differences from Fig. 1 are explained clearly in the text.
\begin{itemize}
\item Optical microphone: we have rearranged the text so that the optical microphone concept is introduced in a separate paragraph and expanded our discussion of it to improve clarity (Section V, paragraph 2).
\item The cloth screen was used to reduce the intensity of light on the photodiode and avoid saturating the analog-to-digital converter. It was indeed an improvised solution using the ``grill cloth'' and plastic frame from the dismantled commercial speaker used in the demonstration. We have added a sentence to the end of Section V A to clarify. 
\item Figure 6: agreed, we have removed this figure and moved the photo to the Supplementary Material.
\item []
\end{itemize}

\noindent 6. Figure 7: why is there a -100 dB difference in the signal level from 7a to 7b?  
\begin{itemize}
\item The Butterworth filter has strongly attenuated the signal outside the passband, this sharp drop-off at low frequencies would appear shallower when shown against a logarithmic scale for frequency as in Figure~3 in the Supplementary Material. We have made a note of this in the Supplementary Material where this plot has been moved to.
\item[]
\end{itemize}

\noindent 7. Much of your description of the various filters should be moved to Supplementary Materials and/or referenced.  This is a physics audience.  More interesting is your observation that all filters degrade the signal in some way, i.e., they are all compromises.  You could usefully expand that argument without becoming overly mathematical.
\begin{itemize}
\item Agreed, we have moved the majority of this section to Supplementary Material. We have chosen to show two highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator), but we refer the reader to the Supplementary Material for the details of these methods as well as a selection of relevant textbooks. We have also expanded the discussion on filters more generally in the main text. 
\item[]
\end{itemize}

\noindent 8. Where in your flowchart are the various filters you describe?  Are the codes for them readily available?  How does one implement a filter?  
\begin{itemize}
\item The filters are performed separately after the data is gathered. We have added a sentence to the figure caption to clarify this. 
\item All of the code is readily available on the provided GitHub. The provided code is documented and commented to ease reproducing the results. For the double-anonymous peer review, we removed the link to our git repository from the manuscript. 
\item[]
\end{itemize}

\noindent So you see that I'm asking for a major revision of your manuscript.  I hope this does not discourage you.  My goal is to expand the size of the audience reading your article, and to increase its utility for that group of educators.  I look forward to a revised manuscript.

\hspace{1em}\vspace{2em}\\

\subsection*{Comments from R3}

The paper describes a table top interferometer for demonstrating the basic operating principles of gravitational wave (GW) detectors. Audio signals are used to demonstrate gravitational wave data analysis techniques for the detection and estimation of signals in noisy data. The paper is well written and will be of interest to educators who wish to provide some hands-on training to students in GW astronomy. I have some comments below that the authors should address. \\

\noindent 
-- The data analysis problems discussed in this paper are related to continuous wave searches. However, I did not find a mention of the key feature involved in these searches, namely, Doppler modulation of the signal frequency caused by the rotation of the Earth. This is the principal reason behind the extremely high computational cost of CW searches. A discussion of doppler modulation should be included. (It would be neat if the setup could be extended in the future to show the Doppler effect.)
\begin{itemize}
\item We have including some details on how the detection statistic is used to account for the Earth's motion around the Sun in Section IV A. We have also noted that Doppler modulation of the source also needs to be accounted for when searching for a neutron star in a binary.
\item It would indeed be very nice to modify the setup to show the Doppler effect. We have added some possible ways to explore this topic in our future work discussion in Section VI. 
\item[]
\end{itemize}

\noindent
-- Is the HMM defined properly in Sec. IV A? The observables here are the amplitudes of the spectrogram pixels at a given time index and the hidden markov process is the underlying drift of frequency and amplitude of the signal. I was expecting to see a conditional pdf that connects the state of the underlying (suitably discretized) process to the observables. I don't see it.
\begin{itemize}
\item We have clarified the explanation of the HMM in the text. The detection statistic $F(t_i, f_j)$ captures the likelihood of a signal being present in each time-frequency bin conditional on the observed data, where we have dropped the conditionally from the notation to avoid clutter. In Section IV A we have expanded our description of gravitational wave analysis and referred to both the transmission matrix and emission probability by name to reduce confusion.
\item[]
\end{itemize}

\noindent
-- Is Eq. 1 correct? The LHS is a marginal probability while the RHS is a joint probability. Shouldn't there be a marginalization on the RHS?
\begin{itemize}
\item This equation was indeed incorrect and we thank you for the careful review. The equation should give the probability for a specific path, and is not a sum over all possibilities. We have corrected equation 1 and clarified what it represents in the text.
\item[]
\end{itemize}

\noindent
-- The normalization used for $F(t_i, f_j)$ should be defined explicitly. 
\begin{itemize}
\item Agreed, each value is divided by the maximum value in the grid such that $\max_{i,j} F(t_i, f_j) = 1$. We have added this information to section IV B.
\item[]
\end{itemize}

\noindent
-- Looking at Fig. 3, it appears that the noise is not white and rises in power at lower frequencies. How is the color of the noise taken into account, if at all, in the HMM or the Viterbi algorithm? A discussion of this point should be included although it is fine to assume white noise as a first approximation. 
\begin{itemize}
\item We do indeed assume that the noise is white in our analysis. This assumption is consistent with the use of Viterbi for continuous-wave searches, where white noise is assumed even though LIGO noise is not white (for example, see Suvorova et al. Phys. Rev. D, Vol 96, Issue 10, 102006 (2017). We have added some text to clarify this assumption in the first paragraph of Section IV C.
\item[]
\end{itemize}

\noindent
-- The Viterbi algorithm used in this paper falls along the general lines of finding the optimal path through a chirplet graph (which includes the spectrogram) that has been explored extensively in the GW literature. More broadly, the detection and estimation of chirp signals, which leave a track in the time-frequency plane, has been a topic of many papers. For the benefit of the readers, the authors should provide an expanded literature review in Sec. IV A and state clearly if their approach is new and, if so, how it differs from the existing ones. A partial list of papers is given below: \\    
$\ast$ E. Chassande-Mottin and A. Pai, Phys. Rev. D 73, 042003 (2006). \\
% Detection of chirps with model of physically realistic GW mergers. Parametric model 
$\ast$ E. J. Candes, P. R. Charlton, and H. Helgason, Classical Quantum Gravity 25, 184020 (2008).   \\
% dynamical programming to find chirping signals. Finding chirp-like signals that are otherwise unmodelled. Non parametric model. 
$\ast$ W. G. Anderson and R. Balasubramanian, Phys. Rev. D 60, 102001 (1999).   \\
% Real-time detection of poorly modelled and unmodelled sources
$\ast$ P. Addesso, M. Longo, S. Marano, V. Matta, I. Pinto, and M. Principe, in 2015 3rd International Workshop on  Compressed Sensing Theory and its Applications to Radar, Sonar and Remote Sensing (CoSeRa) (IEEE, New York, 2015), p. 154.   \\
% unmodelled transient detection for chirps and burst signals 
$\ast$ E. Thrane et al., Phys. Rev. D 83, 083004 (2011).   \\
% long transient search (signals lasting many seconds or weeks). Pattern recognition, not match filter
$\ast$ E. Thrane and M. Coughlin, Phys. Rev. D 89, 063012 (2014).    \\
% unmodelled transient searches - all sky
$\ast$ S. D. Mohanty, Phys. Rev. D  D 96, 102008 (2017).   \\
% chirp searches
$\ast$ Margaret Millhouse, Neil J. Cornish, and Tyson Littenberg, Phys. Rev. D 97, 104057 (2018).\\
% chirplet search

\begin{itemize}
\item We agree and have expanded the text to include the suggested references (see below). 
\item The hidden Markov model (HMM) and Viterbi algorithm is not a new approach for gravitational wave analysis. It is used in many searches for continuous waves to date (for example Abbott et al., Phys. Rev. D, 100(12):122002, Dec 2019) and the methods papers are described in Suvorova et al., Phys. Rev. D, 96(10):102006, November 2017 and Suvorova et al. Phys. Rev. D, 93(12):123009, June 2016. We have clarified throughout the paper that these methods are not new and that we use real continuous wave searches as inspiration for our table-top demonstration. The HMM and Viterbi method enable a search for persistent gravitational waves where the signal model allows for the continuous wave frequency to change unpredictably over months and years. We note that many of the papers listed above are about searching for transient signals rather than the persistent signals considered here. 
\item We have added a new subsection (Section IV A in the revised draft) to emphasize that these techniques are not new and that they are inspired by real continuous wave searches. In this section, we have pointed the reader to the search papers and methods papers that use HMM and Viterbi. We have also added references to other (non-Viterbi) continuous-wave search methods. Finally, we point the interested reader, who would like to know more about chirp or transient searches, to the papers suggested above. 
\item[]
\end{itemize}

\noindent
-- Sec VC: I feel that this section too long and is trying to fit too much of digital signal processing (DSP) in a limited space. The topic of digital filtering is not something that Physics undergraduates (or even graduates) are usually familiar with and requires a more gentle introduction than provided here. For example, the term ``IIR'' is introduced without explaining what the impulse response of a filter is (and there is no mention of FIR filters). Similarly, Eq. 4 looks like a Z-transform, and advanced topic in itself, but this is not mentioned and instead the term ``complex frequency'' is used that is likely to be unfamiliar to most readers. (Incidentally, there seems to be a notational inconsistency across the two sides of Eq.3: `f' and `omega'.) To improve the readability of the paper, the authors should consider shortening this section by focusing on only the filtering procedure that worked the best. References to DSP textbooks may be provided for the interested reader to acquire background material. 
\begin{itemize}
\item Length of Section V: Agreed, we have moved the majority of Section V C to Supplementary Material and now only include two highlight results in the main text (the combined notch plus Wiener filter and the logMMSE estimator). 
\item IIR, FIR, and Z-transform: In the Supplementary Material, we have expanded the description of the impulse response of a filter mentioning both IIR and FIR. We have also added some text clarifying that equation 2 in the Supplementary Material (previously equation 4 in the main text) is a Z-transform. 
\item Notation inconsistency: We have corrected the inconsistency in equation 1 in the Supplementary Material (previously equation 3 in the main text). 
\item DSP textbooks: We have added references to a selection of Digital Signal Processing textbooks at the start of the Supplementary Material and also in the main text Section V.
\item[]
\end{itemize}

\noindent
-- While the authors have applied standard filtering methods in Sec. VC, the removal of high power line features from data in GW detectors is done using more sophisticated approaches. For example, cross-channel regression, linear predictive filtering, Kalman filtering etc. There are also some non-linear and non-parametric methods that have been explored in the GW literature. This point should be mentioned in the paper.
\begin{itemize}
\item Agreed, we should mention this work. We have added a short paragraph at the end of section V B where we discuss environmental noise with the table-top interferometer. We draw parallels to the real detectors and briefly mentioning the importance of noise mitigation and subtraction for gravitational-wave data with references to a selection of papers.
\item[]
\end{itemize}

\noindent
-- The axes label for frequency is not consistent across the figures: in some plots it is ``f/Hz'' and others ``f(Hz)''.
\begin{itemize}
\item Thank you, we have corrected the axis labels and they are now consistent throughout.
\item[]
\end{itemize}

\noindent
-- This is not a comment but just my curiosity: How strong is the effect of mechanical vibrations of the camera or photodiode mount on the observed signal?
\begin{itemize}
\item This is an interesting question which does require further investigation. We have added this idea as a possible future investigation in Section VI. 
\item[]
\end{itemize}

\hspace{1em}\vspace{1em}\\

\subsection*{Comments from R1}

This is a very interesting paper written in a clear and informative language. However, I don’t think that it suits the style of papers in AJP because it represents a novel idea of teaching and exploring gravitational waves with audio signals. After reading and inspection of the manuscript I came to the  conclusion  that it is a  research  paper  which should pass  through critical discussion  with researchers  from gravitational  wave  community who  should  confirm  that  the audio signals produced in the way proposed in the manuscript, can indeed mimic real gravitational wave signals without confusion of students and teachers. Gravitational waves are transverse-traceless while the sound waves are purely longitudinal. In this respect there is no analogy between gravitational waves and audio signals whatsoever. At the same time, the principles of extraction of signal from noise and the filtering technique worked out in the manuscript are applicable in a wide range of physical disciplines and may be useful for the gravitational wave community as well. On the other hand, I cannot imagine how teachers of physics will explain the complexity of the Wiener filter, the Markov chain, the Viterbi algorithm to high school students or even to undergraduate students in university college. Signal processing technique is based  on a  high  level  mathematics (statistics, probability theory, stochastic processes, etc.) and engineering. In my opinion, the paper is too complicated both technically and mathematically for publishing in AJP. I am convinced that this is a very good and high-quality research  paper but  my  fair  opinion  is  that the authors should submit  it  to  Physical Review D or another peer-review physics journal. \\

\begin{itemize}
\item Suitability for AJP: We thank you for the recommendations regarding the suitability and complexity of this manuscript. Following the advice of the editor and referee R3, we have moved the majority of Section V to Supplementary Material so that the main text is now less technical and the interested reader can still see our selection of analysis examples. We now only focus on two highlight results in the main text Section V. We have also clarified our description of the Viterbi method following recommendations from the editor and R3. 
\item Use of audio signals: We agree that there is indeed a fundamental difference between sound waves and gravitational waves. In Section II, we have clarified that we are using sound as an analogy. For this table-top experiment, we require a signal that behaves similar to gravitational waves in that it alters the interference pattern produced by a Michelson interferometer. Using a speaker and audio signals provides a simple and relatively cost-effective means to do this, compared to, for example, a piezo-electric crystal fixed to the back of a mirror as used in some gravitational wave demonstrations (e.g. Cooper et al., arXivv e-prints 2004.03052, 2020). The main focus of this study is the extraction of a signal from interferometer data using a series of analysis techniques, some of which are inspired by continuous gravitational-wave searches. However we emphasize that both the simulated signal and the extraction approaches in the table-top experiment differ from real gravitational-wave experiments in many important ways, e.g. diurnal amplitude modulation, Doppler frequency modulation due to proper motion of the source, colored noise, nonstationary glitches in the operation of the interferometer, quantum effects in the measurement, and so on. The table-top system represents a simplified analogy of a research-level system and real signals.  We also note that the use of sound as an analogy to gravitational-wave signals is not novel and is used by many gravitational-wave research groups to aid in their explanations (for example Sounds of Spacetime \texttt{https://www.soundsofspacetime.org/} and Black Hole Hunter \texttt{blackholehunter.org})
\item[]
\end{itemize}

\end{document}

### 2021-05-20 paper/ifo-appendix.tex
getting manuscript ready for resubmission (removing identifying information for double anoymous review)
added Viterbi reference here as well

@@
 The current build and sample data can be found at:
-%\textbf{[Removed GitHub web-link for double-anonymous review.]}
-\url{https://github.com/daccordeon/gravexplain}
+\textbf{[Removed GitHub web-link for double-anonymous review.]}
+%\url{https://github.com/daccordeon/gravexplain}
 % \begin{verbatim}
@@
 This appendix contains some details regarding the implementation of the Viterbi algorithm described in Section~\ref{sec:viterbi}. 
-The Viterbi algorithm is a classic method in signal processing, whose theoretical underpinnings and implementation are easily accessible to undergraduate students. 
+The Viterbi algorithm~\cite{Viterbi:1967} is a classic method in signal processing, whose theoretical underpinnings and implementation are accessible to undergraduate students. 
 The interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001} to learn more.

### 2021-06-30 paper/ifo-introduction.tex
editor comments for abstract, intro and secion ii

@@
 In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the Universe~\cite{GW150914}.
-Gravitational waves are a prediction of Albert Einstein's theory of General Relativity~\cite{Einstein:1916}, 
-they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
+Gravitational waves are a prediction of Albert Einstein's theory of General Relativity~\cite{Einstein:1916}; they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
 The effect of gravitational waves is a change in lengths: a ``stretching and squashing'' of the distance between objects. 

### 2021-06-30 paper/ifo-tableTopGWs.tex
editor comments for abstract, intro and secion ii

@@
 Current generation gravitational-wave observatories have kilometer-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo, respectively). 
-They can measure minuscule changes in distance due to gravitational waves, e.g.\ the first detection of a binary black hole merger produced a strain of $10^{-21}$~\cite{GW150914}, corresponding to a change in distance which is a fraction of the width of a proton.
+They can measure minuscule changes in distance due to gravitational waves; e.g., the first detection of a binary black hole merger produced a strain of $10^{-21}$~\cite{GW150914}, corresponding to a change in distance equal to a fraction of the width of a proton.
 
@@
 Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
-Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
-The vibrations move the mirrors, changing the length of the interferometer arms, and therefore the interference pattern.
+Sound vibrations provide a simple means to move the components of a demonstration interferometer, changing the length of the interferometer arms, and therefore the interference pattern.
 We use audio signals throughout this work to simulate gravitational-wave-like signals.
@@
 Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
-The intensity timeseries is recorded via webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
-In Section~\ref{sec:optical_microphone}, which considers more complex audio, the webcam is replaced with a photodiode, which offers a higher sample rate than the webcam.
+\textbf{The intensity timeseries recorded by the webcam is discussed in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.}
+When capturing more complex audio (Section~\ref{sec:optical_microphone}), the webcam is replaced by a photodiode, which offers a higher sample rate than the webcam.
 
@@
 If the relative change in the length of the arms is kept small enough, then the intensity of the interference pattern at any point on the screen oscillates at the same frequency as the injected audio.
-But if the relative change in the length of the arms is large enough for multiple fringes to pass over the measured point in a single speaker deformation, then over-counting artificially raises the measured frequency.
+For larger relative length changes, multiple fringes will pass through the detection point during a single speaker oscillation, raising the measured frequency artificially. 
 As such, any motion of the fringes must be kept small by playing the sound softly through the speaker.

### 2021-06-30 paper/paper_main.tex
editor comments for abstract, intro and secion ii

@@
 Recordings of simple chords and drums are easily recovered using these techniques, but complex music and speech are more challenging. 
-This demonstration can be used by educators in undergraduate laboratories and can be adapted for use as an engagement tool for communicating gravitational-wave and signal-processing topics to non-specialist audiences.
+This demonstration can be used by educators in undergraduate laboratories and can be adapted for communicating gravitational-wave and signal-processing topics to non-specialist audiences.  
 \end{abstract}

### 2021-06-30 paper/ifo-singleTone.tex
editor comments section iii

@@
 The intensity of the interference pattern is measured at a single point on the screen, indicated by the pink star in Fig.~\ref{fig:interference_pattern}. 
-The webcam records video in three color channels; red, blue, and green. 
-We use the green channel as an approximation of the total intensity produced by the monochromatic green laser.
+The webcam records video in three color channels: red, blue, and green. 
+We use the green channel as an approximation of the total intensity produced by the green laser.
 The webcam samples at a rate of $30\, {\rm Hz}$, which limits the spectral content of observable signals to less than $15\,{\rm Hz}$, the Nyquist frequency.
@@
 
-In the following section, we use a hidden Markov model technique, commonly used in continuous-wave searches, to recover a signal with a slowly changing frequency.
-
 \end{document}

### 2021-07-01 paper/ifo-tableTopGWs.tex
small changes from editor comments

@@
 A change in the relative length of the arms causes these rings to move radially in or outwards.
+The intensity timeseries is recorded by webcam (Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}) or photodiode, which offers a higher sampling rate suitable for capturing more complex audio (Section~\ref{sec:optical_microphone}).
 Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
-\textbf{The intensity timeseries recorded by the webcam is discussed in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.}
-When capturing more complex audio (Section~\ref{sec:optical_microphone}), the webcam is replaced by a photodiode, which offers a higher sample rate than the webcam.
 

### 2021-07-01 paper/ifo-wanderingTone.tex
small changes from editor comments

@@
 A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
-The audio analog is a tone that wanders in frequency: a note that changes in pitch.
+The audio analog is a tone that wanders in frequency.
 
+%%%%%%%%%%%%%%%%%%%%%% Comment from editor to remove this figure
 We first demonstrate why a Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case.
 As shown in Fig.~\ref{fig:viterbi_comparison_webcam_spectrum}, the signal is spread across multiple frequency bins, making it challenging or in some cases impossible to detect.
+
 %A different means of detection is needed for a signal which has a randomly wandering frequency.
@@
 The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
-In Section~\ref{sec:realCWSearches}, we briefly review the analysis method used by LIGO and Virgo, in Section~\ref{sec:viterbi}, we describe the methods as applied in this work, and in Section~\ref{sec:wanderingResults}, we show results for recovering a wandering frequency signal using the table-top interferometer.
+In Section~\ref{sec:realCWSearches}, we briefly review the analysis method used by LIGO and Virgo; in Section~\ref{sec:viterbi}, we describe the methods as applied in this work; and in Section~\ref{sec:wanderingResults}, we show results for recovering a wandering frequency signal using the table-top interferometer.
 
@@
 
-In Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, a hidden Markov model is used to search for continuous gravitational waves. 
-In a Markov process, the current state depends only on the previous state (in this case the state is the frequency of the signal). 
+In Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, a ``hidden Markov model'' is used to search for continuous gravitational waves. 
+In a Markov process, the current state depends only on the previous state (in this case the ``state'' is the frequency of the signal). 
 In a hidden Markov model, the frequency state of the signal is unknown (hidden) and can undergo transitions at discrete times. 
@@
 In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
-If searching for continuous waves from a neutron star in a binary (such as an LMXB), the Doppler modulation of the source also needs to be taken into account and a different detection statistic is used~\cite{SuvorovaEtAl:2017}.
+When searching for continuous waves from a neutron star in a binary (such as an LMXB), the Doppler modulation of the source also needs to be taken into account and a different detection statistic is used~\cite{SuvorovaEtAl:2017}.
 
@@
 This is called the transition probability matrix. 
-In LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin at each time step~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017,  MiddletonEtAlO2LMXBs:2020}.
+For example, in LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin in each time step~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017,  MiddletonEtAlO2LMXBs:2020}.
 In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to (i) remain in the same frequency bin, or (ii) move down one frequency bin~\cite{MillhouseStrangMelatos:2020, SunEtAlSNR:2018}.

### 2021-07-01 paper/ifo-complexAudio.tex
shifting further reading into supp mat, shortening conclusions, minor edits throughout

--- 2021-05-06 paper/ifo-complexAudio.tex
@@
 
-In this section, we explore how this demonstration can be used to teach a selection of signal processing techniques. 
-We use complex audio signals (such as music and speech) as a natural successor to the constant and wandering tones used in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} respectively.
-As complex audio signals are not quasi-monochromatic, the Viterbi algorithm as used in Section~\ref{sec:viterbi_wandering} is not directly applicable here. 
+In this section, we explore how our apparatus can be used to teach a selection of signal processing techniques. 
+We use complex audio signals (such as music and speech) as a natural successor to the constant and wandering tones used in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, respectively.
+As complex audio signals are not quasi-monochromatic, the Viterbi algorithm used in Section~\ref{sec:viterbi_wandering} is not directly applicable here. 
 Instead, we use a hierarchy of passive filters which suppress noise, yet do not assume any specific form of the signal, unlike the Fourier-based maximum likelihood filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.
@@
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
-We consider a selection of filter techniques, details of which can be found in the Supplementary Material and we refer the reader to Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045} for further information on digital signal processing.
+We consider a selection of filter techniques, details of which can be found in the Supplementary Material, and we refer the reader to Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045, DigitalProcingOfSpeechSignals:1978} for further information on digital signal processing.
 In Section~\ref{sec:opticalMicResults}, we present the two best-performing analyses from our Supplementary Material.
@@
 
-We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together make a photo-detector that produces a voltage that depends on the incident intensity. 
+We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together produce a voltage that depends on the incident intensity. 
 The photodiode records the interference pattern at roughly the same off-center position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily.
@@
 
-The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass hum throughout. 
-This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), as shown in Fig.~\ref{fig:psd_noise}. 
-The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ Australian mains power grid signal up to and beyond $8$th harmonic thereof. 
-The mains signal is also present (but far weaker) in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 
+The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass continuous hum. 
+This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), shown in Fig.~\ref{fig:psd_noise}. 
+The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ Australian mains power grid signal up to and beyond $8$th harmonic. 
+The mains signal is also present but far weaker in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 
 Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself. 
@@
 The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
-
-Environmental noise reduction is an active area of research for gravitational-wave detectors~\cite{EfflerEtAl:2015}.
-In some cases, environmental noise sources can be removed or mitigated on site~\cite{CovasEtAl:2018}. 
-Where this is not possible, noise subtraction can be done after data collection~\cite{DriggersEtAl:2019, DavisEtAl:2019}. 
-One example is the removal of spectral noise lines from the United States power grid operating at $60\,{\rm Hz}$~\cite{VajenteEtAl:2020}. %United States's?
+Environmental noise reduction for gravitational-wave detectors is an active area of research (see the accompanying Supplementary Material for further information and resources on this topic).
 
@@
 All filters are tested on the same $1\,{\rm s}$ long speech recording.
-The results of this section are shown in Fig.~\ref{fig:notchWienerLogMMSEResults} which we refer to throughout. 
-In Fig.~\ref{fig:notchWienerLogMMSEResults} the left and right columns show the timeseries and frequency spectrum respectively. 
+The results of this section are shown in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
+In the figure the left and right columns show the timeseries and frequency spectrum respectively. 
 The first row shows the input signal played through the speaker (see Fig.~\ref{fig:ifo_schematic_webcam}). 
@@
 For speech intelligibility, this means that either: (i) some noise remains in the filtered recording, (ii) some of the speech content is lost as certain important frequencies are attenuated, or (iii) the speech is somewhat distorted in time.
-All three of these cases can, when taken to the extreme, make the speech intelligibility worse than the unfiltered recording, i.e.\ the results can be worse than doing no filtering.
+All three of these cases can, when taken to the extreme, make the speech intelligibility worse than the unfiltered recording. 
 Therefore, we choose filters that compromise between achieving the three conditions.
@@
 In this section, we present the results of two advanced signal processing analysis techniques applied to the optical microphone recordings. 
-The techniques are only briefly described here and we refer the reader to the accompanying Supplementary Material for further details and some other analysis techniques.
+The techniques are only briefly described here and we refer the reader to the accompanying Supplementary Material for further details and other analysis techniques.
 
@@
 It amplifies parts of the signal with high signal-to-noise (SNR) ratio while suppressing parts with low SNR. 
-The results of the combined cascaded notch and Wiener filter are shown in the third row of Fig.~\ref{fig:notchWienerLogMMSEResults} where the timeseries and spectrum are shown by the left and right panels respectively.
-Most of the mains noise is removed, however, the recovered voice sounds muffled and is not understandable.
+The results of the combined cascaded notch and Wiener filter are shown in the third row of Fig.~\ref{fig:notchWienerLogMMSEResults}. 
+Most of the mains noise is removed, however the recovered voice sounds muffled and is not understandable.
 
@@
 We use an existing implementation of the logMMSE from~\cite{logmmse}.  
-The logMMSE estimator results are shown in the bottom panel of Fig.~\ref{fig:notchWienerLogMMSEResults}. 
+The logMMSE estimator results are shown in the bottom panels of Fig.~\ref{fig:notchWienerLogMMSEResults}. 
 We see significant attenuation of the mains harmonics and general smoothing of the spectrum. 

### 2021-07-01 paper/ifo-conclusions.tex
shifting further reading into supp mat, shortening conclusions, minor edits throughout

@@
 
-In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community and more broadly.
-Gravitational-wave signals are simulated by playing sound recordings through a speaker fixed to the back of one of the arm mirrors. 
-We consider quasi-monochromatic audio signals as an analog to continuous-wave signals to directly demonstrate techniques used in current continuous gravitational-wave searches.
-We then consider more complex audio (music and speech) for a more general interest in signal processing and explore the use of the interferometer as an optical microphone. 
-A variety of example activities is considered in the Supplementary Material. 
-We find that the logMMSE produces the best results, however, it does not enhance the recording to an intelligible level. 
-The complex audio signals can also serve as a more distant analog to un-modeled gravitational-wave burst signals, e.g. from supernovae~\cite{MillhouseEtAl:2018,SuvorovaPowellMelatos:2019}.
+% reduced conclusion
+
+In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community.
+We explore the use of the interferometer as an optical microphone, considering a more general treatment of signal processing with complex audio signals, which can also serve as a more distant analog to un-modeled gravitational-wave burst signals, e.g., from supernova~\cite{MillhouseEtAl:2018,SuvorovaPowellMelatos:2019}.
+The demonstration can be adapted for use in the undergraduate laboratory and as an tool for explaining gravitational-wave research to a wider, non-specialist audience.
 
 \begin{comment}
-% included these paras in the first to avoid repetition and reduce some of the length of the paper before submission
-% reduced detail for conclusion (trying to reduce overall length)
-In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} we analyse audio signals with constant and wandering frequencies respectively as analogues to continuous gravitational wave signals. 
-The wandering frequency signal is recovered using using the Viterbi algorithm. 
-These sections directly demonstrate the techniques used in current continuous gravitational wave searches. 
-%We demonstrate the recovery of a signal with constant frequency and signal with slowly wandering frequency in Sections~\ref{sec:ifo} and ~\ref{sec:single_tone}, respectively.
-We use a basic webcam with a sampling rate of $30\,{\rm Hz}$ to track the interference pattern.
-For the wandering frequency we successfully recover the injected signal using the Viterbi algorithm.
-These sections directly demonstrate the techniques used in continuous gravitational wave searches.
-
-% again reducing overall length as we probably don't need to list everything
-The recovery of more complex music and speech signals are demonstrated in Section~\ref{sec:optical_microphone}, where we treat the interferometer as an optical microphone. 
-Simple music recordings can be recognised when played back, however speech remains unintelligible after applying a selection of filters and statistical techniques from signal processing. 
-We find that the logMMSE estimator produces the best result of those applied here, however it does not enhance the recording to an intelligible level. 
-This demonstration allows students to explore a variety of filtering and speech enhancement techniques. 
-Further avenues of exploration to improve intelligibility are detailed in Section~\ref{sec:future_work}. 
-%In Section~\ref{sec:optical_microphone}, we demonstrate the recovery of more complex signals of music and speech.
-By substituting the webcam with a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone by greatly increasing the sampling rate.
-We demonstrate that the optical microphone is capable of playing back simple recordings well but produces unintelligible recordings of speech.
-The mains signal at $50\,{\rm Hz}$ and its harmonics dominate the photodiode output.
-To recover the signal from under this noise we apply a hierarchy of filters and statistical techniques from signals processing, including a Butterworth band-pass filter and a notch filter, to a Wiener filter and the logMMSE estimator.
-We find that the logMMSE estimator works the best from those tried but still doesn’t enhance the speech to an intelligible level.
-This demonstration allows students to explore a variety of filtering and speech enhancement techniques.
-Further avenues of exploration include even more advanced processing techniques and a full characterisation of the transfer function of the system, with the goal to lead to improvements in intelligibility.
-\end{comment}
 
@@
 The increased excitement and public interest in the field of gravitational-wave research in recent years mean that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
-The field of gravitational-wave astrophysics continues to grow and the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.
-% \jam{``charming'' is very sweet}     - agreed I wasn't sure at first but it has grown on me a lot! :) - Hannah
-%The demonstration presented here has great potential to be adapted for use in the undergraduate laboratory.
-It can be used to demonstrate topics such as interferometer physics; gravitational wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques to physics and electrical engineering students.
-Given the increased level of excitement and public interest in the field of gravitational wave astronomy in recent years, this table-top demonstration can be used as a tool for explaining gravitational wave research to a wider, non-specialist audience.
-The future will bring many more detections, including further observations of transient signals from binary black hole and binary neutron star mergers, as well as the anticipated first detections of other classes of signal, including continuous waves, to which this demonstration provides some charming insights.
 
-% Continuous gravitational wave searches use sophisticated statistical techniques to search for weak signals in noisy interferometric data.
-We demonstrate some of these techniques by observing sound with a table-top Michelson interferometer.
-In particular, we successfully use the Viterbi algorithm to recover a tone with wandering frequency from a webcam video of the interference pattern.
+\end{comment}
 
-%Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer.
-We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.
-
-%These demonstrations have good potential to be adapted into the undergraduate laboratory.
-There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches.
-Most of all, the signal processing of speech enhancement is an interesting problem that’s relevant to digital communications.
+As the field of gravitational-wave astrophysics continues to grow, the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.
 

### 2021-07-01 paper/ifo-supplementary-material.tex
caption edit
getting manuscript ready for resubmission (removing identifying information for double anoymous review)
moving footnotes into bib for supplementary material as well
shifting further reading into supp mat, shortening conclusions, minor edits throughout

@@
 % GravExplainSuppMatt
-%\documentclass[prb,reprint,nofootinbib]{revtex4-1} 
-\documentclass[aps,pra,superscriptaddress,reprint,nofootinbib]{revtex4-1}
-% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
 
-% Note that AJP uses the same style as Phys. Rev. B (prb).
-% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
-% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
-% We look forward to receiving your submission to AJP.
+\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-1}
 \usepackage[utf8]{inputenc}
@@
 \title{Supplementary Material for \\Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
+
+\author{\textbf{[Authors removed for double-anonymous review.]}}
+
+\begin{comment}
 
@@
 }
+\end{comment}
 
 \maketitle
+
+\section{Further reading for gravitational wave search....}
+
+Here we provide a selection of resources and reading material..
+
+\subsection{Environmental noise in gravitational-wave detectors}
+In Section~V~B in the main article, we see evidence for environmental noise in our data. 
+In gravitational-wave detectors reducing and mitigating environmental noise is an active area of research~\cite{EfflerEtAl:2015}. 
+One such noise source is due to seismic ground motion from the Earth. 
+In gravitational-wave observatories, the optical components are suspended from a series of pendulums to achieve passive isolation, and feedback control loops provide active isolation\cite{MatichardEtAl:2015, AdvancedVirgo:2015}.
+Such a system is beyond the scope of this article, however feedback control has been demonstrated in table-top Michelson interferometers designed for the undergraduate laboratory~\cite{LibbrechtBlack:2015}.
+Other sources of noise include acoustic sound waves (for example from nearby vehicles, wind, and components such as electric fans) and magnetic noise (for example from high-voltage power lines, building heaters and lights, or any electrical power circuit)~\cite{EfflerEtAl:2015,CironeEtAl:2019}.
+In some cases, environmental noise sources can be removed or mitigated on site~\cite{CovasEtAl:2018}. 
+Where this is not possible, noise subtraction can be done after data collection~\cite{DriggersEtAl:2019, DavisEtAl:2019}. 
+One example is the removal of spectral noise lines from the United States power grid operating at $60\,{\rm Hz}$~\cite{VajenteEtAl:2020}. %United States's?
+
+\subsection{test}
 
@@
 	\includegraphics[width=0.5\textwidth]{figures/butterworth_150_3000.pdf}
-	\caption{Butterworth bandpass filter frequency response.
-Any amplitude below $-3\,{\rm dB}$ is significant attenuation (less than half the power at that frequency remains).
-The red, dashed vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$.
-Note the flat response within the band characteristic of the Butterworth filter.}
+	\caption{
+Butterworth bandpass filter frequency response. 
+%Any amplitude below $-3\,{\rm dB}$ is significant attenuation (less than half the power at that frequency remains). 
+% suggest removing as -3db is hard to judge on the figure and I think the attenuation outside the frequency range is the key point to get across here
+The red, dashed vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$. Note the flat response within the band characteristic of the Butterworth filter.}
 	\label{fig:butterworth}

### 2021-07-01 paper/ifo-wanderingTone.tex
shifting further reading into supp mat, shortening conclusions, minor edits throughout

@@
 \begin{document}
-% I've rephrased the first part of this section a lot and provide some justification below
-% I think we need to lead with the details of the signal here (similar to the last section), not the specifics of the data analysis as the issues seem confused here.
-The fact that long data sets are analysed is not because of the signal wandering, if anything that's an argument against analysing long data sets to make detection easier, it's because the signal has low amplitude.
-Also what is the justification of the "(but not always)" statement, do you have a reference?
-- hannah
-% Hmm, not sure where I got "(but not always)" from but if I didn't note it down then then I have no idea why now.
-Re: long time sensing, I was trying to justify that frequencies can wander but I prefer your rewrite.
-- James
 
-\begin{comment}
-Continuous wave searches, such as in Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017,SearchTwoSpecS6:2017,SunEtAlSNR:2018,JonesSun:2020}, are often (but not always) performed on long datasets (months to years in duration).
-The frequency of a continuous wave signals may wander significantly over this timescale, due to stochastic internal processes in the superfluid interior of isolated neutron stars~\cite{MelatosDouglassSimula:2015,Jones:2010} or variable accretion flows onto a neutron star from a stellar companion (e.g.
-as in LMXBs)~\cite{BildstenTB:1998}.
-In this context, wandering ``significantly'' means ``across multiple frequency bins'', where the typical width of a frequency bin is the reciprocal of the total observing time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
-The audio analogue to these continuous wave signals is a tone that wanders in frequency; a note that changes pitch.
-Here, we adapt the data analysis technique used to search for slowly wandering continuous waves signals from spinning neutron stars in aLIGO and aVirgo data~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}.
-\end{comment}
-
+A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
 Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. 
 A typical observation involves $1 \times 10^{10}$ wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation. 
-A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
-The audio analog is a tone that wanders in frequency.
+Here we consider the audio analog of a tone that wanders in frequency.
 
 %%%%%%%%%%%%%%%%%%%%%% Comment from editor to remove this figure
-We first demonstrate why a Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case.
-As shown in Fig.~\ref{fig:viterbi_comparison_webcam_spectrum}, the signal is spread across multiple frequency bins, making it challenging or in some cases impossible to detect.
+%We first demonstrate why a Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case.
+%As shown in Fig.~\ref{fig:viterbi_comparison_webcam_spectrum}, the signal is spread across multiple frequency bins, making it challenging or in some cases impossible to detect.
 
-%A different means of detection is needed for a signal which has a randomly wandering frequency.
-
+A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case as the signal is spread across multiple frequency bin. 
 In continuous wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time.
@@
 
+% Removing figure after editor request. 
+\begin{comment}
 \begin{figure}
@@
 
-In Refs.~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}, a ``hidden Markov model'' is used to search for continuous gravitational waves. 
+A ``hidden Markov model'' can be used to search for continuous gravitational waves~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
 In a Markov process, the current state depends only on the previous state (in this case the ``state'' is the frequency of the signal). 
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{figures/viterbiDiagramBlue.pdf}
+\includegraphics[width=0.5\textwidth]{figures/viterbiDiagramGrey.pdf}
 \caption{\label{fig:viterbi}
@@
 The circles represent elements in a time-frequency grid, labeled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
-The color of each circle represents the likelihood of a signal being present in each time-frequency element.
-The dark blue to light green gradient corresponds to low to high likelihood values in arbitrary units for this schematic. 
+The shading of each circle represents the likelihood of a signal being present in each time-frequency element.
+The dark to light gradient corresponds to low to high likelihood values in arbitrary units for this schematic. 
 The objective is to find the most probable path of a signal through the grid from left to right.
@@
 The timeseries data from the interferometer is first split into segments. 
-We then take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude $F(t_i,f_j)$ as the detection statistic (this is the emission probability).
+We then take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude at frequency $f_j$ and time $t_i$, $F(t_i,f_j)$ as the detection statistic (this is the emission probability).
 Assuming Gaussian noise, this detection statistic choice maximizes the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.
@@
 
-We use the schematic in Fig.~\ref{fig:viterbi} to aid our explanation of the method throughout this section. 
-It represents a spectrogram with $N_t$ time bins and $N_f$ frequency bins.
+Fig.~\ref{fig:viterbi} represents a spectrogram with $N_t$ time bins and $N_f$ frequency bins.
 The circles represent the elements of the time and frequency grid, which are labelled as  
 $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-The color of the circles represents the detection statistic $F(t_i,f_j)$ at each grid point where dark blue corresponds to lower values and light green to higher values.
+The shading of the circles represents the detection statistic $F(t_i,f_j)$ at each grid point, where dark corresponds to lower values and light to higher values.
 
-The objective is to find the most likely path through the grid given the observed data and any probabilistic constraints on how the frequency of the signal can wander from $t_i$ to $t_{i+1}$. 
-The transition probability matrix, $A(f_k,f_m)$, describes the probability of system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
+The objective is to find the most likely path through the grid, given the observed data and any probabilistic constraints on how the frequency of the signal can wander from $t_i$ to $t_{i+1}$. 
+The transition probability matrix, $A(f_k,f_m)$, describes the probability of the system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
 Here, we allow the frequency of the signal to either (a) stay in the same bin, (b) move up by a single frequency bin, or (c) move down by a single frequency bin at each transition. 
@@
 
+%Before beginning the analysis, we assume that the signal is equally likely to start in any frequency bin (i.e. the prior is flat). 
+%The first step of the analysis is to define the probability of the system having frequency $f_j$ at 
+\han{rephrase: `The first two sentences seem to contradict each other. Can you clarify?'}
 The signal is equally likely to start in any frequency bin (i.e. the prior is flat). 
@@
 The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises ${\rm Pr}[ f(t_0), f(t_1),... f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
-We note that there is no marginalization used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
+We note that there is no marginalization \han([What does marginalization mean?
+Do you mean no hard barriers?]) used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
 %The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
@@
 \begin{enumerate}
-\item Starting at time $t_1$, each $f_j$ state can originate from three prior states at time $t_0$ (except for the edge cases $f_0$ and $f_{N_f}$ which only have two). 
-The paths between these states are indicated by the lines in Fig.~\ref{fig:viterbi}. 
-At each $f_j$ state, we select the path with the highest $A[f(t_0),f(t_1)] {\rm Pr}[f(t_0)]$ value as the most probable path. 
+\item Starting at time $t_1$, each $f_j$ node (circles in Fig~\ref{fig:viterbi}) can originate from three prior nodes at time $t_0$ (except for the edge cases $f_0$ and $f_{N_f}$ which only have two).
+The paths between these nodes are indicated by the lines in Fig.~\ref{fig:viterbi}. 
+At each $f_j$ node, we select the path with the highest $A[f(t_0),f(t_1)] {\rm Pr}[f(t_0)]$ value as the most probable path. 
 These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the gray lines show the rejected paths. 
-For example, the most probable connection to the element labeled (a) is the one directly behind it (i.e., $f_4$). 
+For example, the most probable connection to the node labeled (a) is the one directly behind it (i.e., $f_4$). 
 Therefore, this path is selected as the best path from $t_0$ to $t_1$ for $f_4$.
@@
 
-\item Moving to time $t_2$, again, we select the path which maximizes Eqn.~\ref{eqn:pathprobability} for each $f_j$. 
+\item Moving to time $t_2$, again we select the path which maximizes Eqn.~\ref{eqn:pathprobability} for each $f_j$. 
 These paths are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.
@@
 
-In continuous wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency could be changing unpredictably over time.
-A Fourier transform of the entire dataset splits the signal power between multiple frequency bins and could result in the signal being undetectable under the noise (see Fig.~\ref{fig:viterbi_comparison_webcam_spectrum}).
+In continuous gravitational wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency can change unpredictably over time.
 The Viterbi algorithm's strength lies in its ability to track such signals through the data.

### 2021-07-01 paper/ifo-complexAudio.tex
updating figure, working on supp mat, added NSBH paper

@@
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
-We consider a selection of filter techniques, details of which can be found in the Supplementary Material, and we refer the reader to Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045, DigitalProcingOfSpeechSignals:1978} for further information on digital signal processing.
+We consider a selection of filter techniques, details of which can be found in the Supplementary Material, where we also refer the reader to a selection of digital signal processing resources for further information.
 In Section~\ref{sec:opticalMicResults}, we present the two best-performing analyses from our Supplementary Material.
@@
 Simple chords and drums can be heard, however more composite sounds and complex melodies cannot be heard clearly. 
-Our observations suggest that this is especially true for certain instruments, in particular flutes and violins sometimes can’t be heard at all. 
+Our observations suggest that this is especially true for certain instruments; in particular flutes and violins sometimes can’t be heard at all. 
 This could be a perceptual effect or a frequency dependence somewhere in the optical microphone.

### 2021-07-01 paper/ifo-futureWork.tex
updating figure, working on supp mat, added NSBH paper

@@
 
-The optical microphone experiment described here can be used to demonstrate and teach a variety of topics, including basic optics and photodiode circuits, to signals processing and speech enhancement.
+The optical microphone experiment described here can be used to demonstrate and teach a variety of topics, including basic optics and photodiode circuits, to signal processing and speech enhancement. 
 Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.
@@
 The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Pi. 
-It also would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode.
+It would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode. 
 The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the abstract signal flow through the system.
 It does not include other potential pathways for signal flow such as acoustomechanical vibrations of the webcam or photodiode mount which could also be examined.
-% commented out whilst James is checking this
-% jam: I'm pretty confident this is what Rob said. 
-%With perfect knowledge of the transfer function, signal recovery is limited by the signal-to-noise ratio as an optimal matched filter could be constructed.
-%\jam{[check with Rob et al. about the veracity of this]}
 
@@
 These could perform the final stage of speech enhancement for the optical microphone. 
-Interestingly, this coincidentally connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
+This coincidentally connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
 Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.
 
-We could also use more traditional techniques such as a wavelet transform (WT) \citep{nason1995stationary} to extract the signal from noise, and compare these with the above methods.
+More traditional techniques such as a wavelet transform (WT) \citep{nason1995stationary} could be used to extract the signal from noise, and compared with the above methods. 
 A WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. 
@@
 
-Returning to the demonstration of continuous gravitational-wave analysis, further extensions could include analyzing data from multiple interferometers to extract directional information for audio signals.
+Returning to the demonstration of gravitational wave analysis, we could analyze data from two interferometer to extract directional information for audio signals based on the relative arrival time of the signal at each interferometer~\cite{GW150914}.
 This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
 Another extension would be to demonstrate the Doppler effect of the Earth's motion around the Sun, which needs to be considered in continuous-wave searches (see Section~\ref{sec:realCWSearches} and Ref.~\cite{JKS:1998}).
-One approach could be to move either the interferometer or the source signal on a circular track or modify the input audio signal to simulate Doppler modulation. 
-%techniques used in the literature to address it by moving the optical table on some kind of circular track with period far shorter than the time it takes the signal to significantly wander in frequency.
-Then, by placing a distant signal source at different positions above or beside the track -- or perhaps more practically the source could move instead of the detector, an analogous Doppler effect could be observed.
+One approach could be to move the source signal on a circular track or modify the input audio signal to simulate Doppler modulation.
 

### 2021-07-01 paper/ifo-introduction.tex
updating figure, working on supp mat, added NSBH paper

@@
 % CW 
-To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
+To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020,GWTC-1:2018}
+%{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} 
+and binary neutron stars~\cite{GW170817,GWTC-1:2018,GWTC-2:2020}
+%\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020},
+and neutron star-black hole mergers~\cite{NSBH:2021}.
+%as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}. 
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 

### 2021-07-01 paper/ifo-supplementary-material.tex
updating figure, working on supp mat, added NSBH paper

@@
 
+\subsection{Continuous wave searches}
+The methods used in Section~IV are inspired by hidden Markov model searches for continuous gravitational waves. 
+A range of other continuous wave searches exist and we refer the reader to Refs.~\cite{MSPSearchO3O2O1:2020, RadiometerO1O2:2019, KnownPulsarsO2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} for searches that consider particular targets or sky locations, and to Refs.~\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, StackSlide:2012} for search methods that cover the entire sky.
+For further information on searching for transient gravitational wave signals see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
+
 \subsection{Environmental noise in gravitational-wave detectors}
@@
 As an example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
-For more information about the techniques used here see Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045}.
+For more information about the techniques used here see Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045, DigitalProcingOfSpeechSignals:1978}.
 

### 2021-07-01 paper/ifo-wanderingTone.tex
updating figure, working on supp mat, added NSBH paper

@@
 For further details on these methods and searches see Refs.~\cite{SuvorovaEtAl:2016, SuvorovaEtAl:2017, ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, BeniwalEtAl:2021, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, BayleyEtAlSOAP:2019, SunEtAlSNR:2018}.
-For a range of other continuous wave searches we refer the reader to Refs.~\cite{MSPSearchO3O2O1:2020, RadiometerO1O2:2019, KnownPulsarsO2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} for searches that consider particular targets or sky locations, and to Refs.~\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, StackSlide:2012} for search methods that cover the entire sky.
-For further information on searching for transient gravitational wave signals see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
+For a range of other continuous wave search methods we refer the reader to references in the Supplementary Material.
 

### 2021-07-19 paper/ifo-appendix.tex
moving references to further reading section after review

@@
 \label{app:code}
-This project is implemented in Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}.
+This project is implemented in Python 3 scripts and jupyter notebooks and MATLAB. 
+We refer the reader to the Supplementary Material for software references. 
+%This project is implemented in Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}.
 We used the Python packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
@@
 
+\begin{comment}
 \section{Interferometric relation between path length and optical intensity}
@@
 %This non-linearity in part of the transfer function will make the reverse problem of extracting the original signal more difficult.
+\end{comment}
 
@@
 
+\begin{comment} % move to supp matt due to large number of references needed here. 
 \section{Photodiode circuit design}

### 2021-07-19 paper/ifo-complexAudio.tex
moving references to further reading section after review

@@
 During recordings, care is taken to minimize activity around the demonstration to reduce environmental noise coupling into the interferometer. 
-The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python~\cite{scipy,python}. 
+The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python (see Appendix~\ref{app:code}).
+% moved refs \cite{scipy,python}.
 When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.

### 2021-07-19 paper/ifo-introduction.tex
moving references to further reading section after review

--- 2021-06-30 paper/ifo-introduction.tex
@@
 These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
-Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019,LibbrechtBlack:2015} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,AMIGO:online,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.
+Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019,LibbrechtBlack:2015} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{TTExhibit:2021}.
 
 % CW 
-To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020} and binary neutron stars~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020}, as well as a black hole merger with an object that could either have been a neutron star or another black hole~\cite{GW190814}.
+To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and neutron star-black hole mergers~\cite{GWTC-1:2018,GWTC-2:2020,NSBH:2021}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
@@
 %Scorpius X-1 is an LMXB that is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
-Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, RadiometerO1O2:2019, SearchRadiometerO1:2017, SearchCrossCorrO1:2017, ScoX1ViterbiO1:2017, SearchTwoSpecS6:2017, MiddletonEtAlO2LMXBs:2020}.
+Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019}.
 
@@
 It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~
-\cite{SuvorovaEtAl:2017,SuvorovaEtAl:2016, ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, BeniwalEtAl:2021, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, BayleyEtAlSOAP:2019, SunEtAlSNR:2018}.
+\cite{SuvorovaEtAl:2017, ScoX1O2Viterbi:2019}. 
 % This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
@@
 %This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world.
-This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online, chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
+This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world.
 These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.

### 2021-07-19 paper/ifo-supplementary-material.tex
moving references to further reading section after review

@@
 
-Here we provide a selection of resources and reading material..
+Here we provide a selection of resources and reading material on the following topics: Section A \han{will list topics when finalised}
+
+\subsection{Gravitational wave obesrvations}
+Gravitational waves have been observed from compact binary coalescences. 
+For more information on observations of binary black hole mergers see Refs~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020,GWTC-1:2018}, for binary neutron star mergers see Refs~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020} and for neutron star-black hole mergers~\cite{NSBH:2021,GW190814}.
 
 \subsection{Continuous wave searches}
-The methods used in Section~IV are inspired by hidden Markov model searches for continuous gravitational waves. 
-A range of other continuous wave searches exist and we refer the reader to Refs.~\cite{MSPSearchO3O2O1:2020, RadiometerO1O2:2019, KnownPulsarsO2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} for searches that consider particular targets or sky locations, and to Refs.~\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, StackSlide:2012} for search methods that cover the entire sky.
+\label{sec:GWSsearches}
+The methods used in Section~IV are inspired by hidden Markov model (HMM) searches for continuous gravitational waves. 
+The HMM method is described in Refs~\cite{SuvorovaEtAl:2017,SuvorovaEtAl:2016,BayleyEtAlSOAP:2019,Viterbi:1967}. 
+HMM searches have targeted low mass X-ray binaries~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017,MiddletonEtAlO2LMXBs:2020}, young supernova remnants~\cite{SNRO3a:2021,MillhouseStrangMelatos:2020,SunEtAlSNR:2018}, post merger remnants~\cite{PostMergerRemnantSearch:2019}, Fomalhaut b~\cite{JonesSun:2020}, and high energy TeV sources~\cite{BeniwalEtAl:2021}.
+
+A range of other continuous wave searches exist. 
+We refer the reader to Refs.~\cite{MSPSearchO3O2O1:2020, RadiometerO1O2:2019, SearchRadiometerO1:2017, KnownPulsarsO2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} for searches that consider particular targets or sky locations, and to Refs.~\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, StackSlide:2012} for search methods that cover the entire sky.
 For further information on searching for transient gravitational wave signals see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
+
+\subsection{Gravitational-wave engagement}
+Numerous engagement activities have been developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online, chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
+Specific examples of using table-top interferometers as engagement or teaching tools include Refs.~\cite{TTExhibit:2021,UgoliniEtAl:2019,LibbrechtBlack:2015,AMIGO:online,FoxEtAl:1999} and parts lists or instructions can be found in Refs.~\cite{TTExhibit:online,LIGOIFOGlue,LIGOIFOMagnets,NikhefIFO,ThorLabsIFO}.
 
@@
 
-\subsection{test}
+\subsection{Software}
+This work is implemented in Python Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}. 
+We make use of Python packages: numpy~\cite{numpy}\footnote{\url{https://numpy.org/doc/stable/reference/routines.fft.html}}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
 
@@
 A photograph of the equipment is shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit} with the interferometer and circuit shown on the left and right, respectively. 
+% schematic of breadboard and connections to pi
+% https://www.circuit-diagram.org/editor/
+% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
+The photodiode circuit diagram is shown in Figure~\ref{fig:circuit_diagram}
+The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an analogy-to-digital converter ADC to a Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
+Useful throughout was the Raspberry Pi documentation~\cite{pidocumentation:online} and the pinoutXYZ reference~\cite{pipinout:online}.
+
+\begin{figure*}
+	\includegraphics[width=.8\textwidth]{figures/circuit_diagram_2.pdf}
+	\caption{\label{fig:circuit_diagram}
+Circuit diagram for reading the photodiode. 
+The photodiode is connected in reverse-bias across an op-amp (pale-red box). 
+The op-amp output is passed through a Sallen-Key anti-aliasing filter (pale-blue box) tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
+The digitized signal is then read by the special-purpose input (SPI) pins of a Raspberry Pi in standard configuration.
+This diagram is produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online} and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
+}
+\end{figure*}
 

### 2021-07-19 paper/ifo-tableTopGWs.tex
moving references to further reading section after review

@@
 Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
-Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Refs.~\cite{TTExhibit:2020,TTExhibit:online,AMIGO:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999} for a selection of other table-top Michelson interferometer designs used for science communication).
+Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see also the further reading section in supplementary material for a selection of other table-top Michelson interferometer designs used for science communication).
 Sound vibrations provide a simple means to move the components of a demonstration interferometer, changing the length of the interferometer arms, and therefore the interference pattern.
@@
 Schematic of the table-top Michelson interferometer. 
-The output of a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO} (top-left) is reflected from a mirror at a right angle.
+The output of a Class 2 $532\,{\rm nm}$ laser (top-left) is reflected from a mirror at a right angle.
 The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter which is approximately $50$/$50$ in reflection and transmission.

### 2021-07-19 paper/ifo-wanderingTone.tex
moving references to further reading section after review

@@
 In continuous wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time.
-The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
+The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2017}. 
 In Section~\ref{sec:realCWSearches}, we briefly review the analysis method used by LIGO and Virgo; in Section~\ref{sec:viterbi}, we describe the methods as applied in this work; and in Section~\ref{sec:wanderingResults}, we show results for recovering a wandering frequency signal using the table-top interferometer.
@@
 The methods used here are inspired by LIGO-Virgo continuous wave searches. 
-For further details on these methods and searches see Refs.~\cite{SuvorovaEtAl:2016, SuvorovaEtAl:2017, ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017, BeniwalEtAl:2021, MillhouseStrangMelatos:2020, JonesSun:2020, MiddletonEtAlO2LMXBs:2020, PostMergerRemnantSearch:2019, BayleyEtAlSOAP:2019, SunEtAlSNR:2018}.
-For a range of other continuous wave search methods we refer the reader to references in the Supplementary Material.
+For further details on these methods and continuous wave searches generally we refer the reader to references in the Supplementary Material.
 
@@
 
-A ``hidden Markov model'' can be used to search for continuous gravitational waves~\cite{SuvorovaEtAl:2016,SuvorovaEtAl:2017}. 
+A ``hidden Markov model'' can be used to search for continuous gravitational waves~\cite{SuvorovaEtAl:2017}. 
 In a Markov process, the current state depends only on the previous state (in this case the ``state'' is the frequency of the signal). 
@@
 This is called the transition probability matrix. 
-For example, in LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin in each time step~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017,  MiddletonEtAlO2LMXBs:2020}.
-In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to (i) remain in the same frequency bin, or (ii) move down one frequency bin~\cite{MillhouseStrangMelatos:2020, SunEtAlSNR:2018}.
+For example, in LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin in each time step~\cite{ScoX1O2Viterbi:2019}.
+In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to (i) remain in the same frequency bin, or (ii) move down one frequency bin (see the Supplementary Material for supernova remnant search references).
 

### 2021-07-19 paper/ifo-complexAudio.tex
moving some more references

@@
 The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
-Speech intelligibility (being able to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$~\cite{speech_intelligibility}.
+Speech intelligibility (being able to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$. 
 Therefore, the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music (adjusting for the Nyquist frequency). 
@@
 %This cloth screen mount was the ``grill cloth'' from the front of a dismantled commercial speaker, re-purposed. 
-The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online} (henceforth: Pi), which provides a convenient means to record the photodiode data.
+The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2 (henceforth: Pi), which provides a convenient means to record the photodiode data.
 Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. 
@@
 
-Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing~\footnote{Folding of frequencies greater than half the sampling rate.} into the detected range.
+Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing (folding of frequencies greater than half the sampling rate) into the detected range.
 We include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.

### 2021-07-19 paper/ifo-futureWork.tex
moving some more references

@@
 \begin{figure*}
-	\includegraphics[height=.8\textwidth, angle=-90]{figures/pipeline.pdf}
+    \includegraphics[width=.8\textwidth]{figures/pipeline_nobox.pdf}
+	%\includegraphics[height=.8\textwidth, angle=-90]{figures/pipeline.pdf}
 	\caption{\label{fig:pipeline_highlighted}
-Flowchart for the signal through the optical microphone. The transfer function would have to account for each stage. 
-The stage inside the blue-dashed box is analyzed in Appendix~\ref{app:intensity_derivation}.
-Understanding the other stages is one aim of future work.
+Flowchart for the signal through the optical microphone. 
+A transfer function of the system would have to account for each stage and an aim of future work. 
 Not shown are the signal processing filters that are performed on the PC once the data is transferred off the Pi.
@@
 
-The optical microphone experiment described here can be used to demonstrate and teach a variety of topics, including basic optics and photodiode circuits, to signal processing and speech enhancement. 
+The experiment described here can be used to demonstrate and teach a variety of topics, from basic optics and photodiode circuits, to signal processing and speech enhancement. 
 Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.
@@
 The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Pi. 
-It would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation (Appendix~\ref{app:intensity_derivation}), and the photodiode. 
+It would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation, and the photodiode. 
 The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the abstract signal flow through the system.

### 2021-07-19 paper/ifo-introduction.tex
moving some more references

@@
 In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the Universe~\cite{GW150914}.
-Gravitational waves are a prediction of Albert Einstein's theory of General Relativity~\cite{Einstein:1916}; they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
+Gravitational waves are a prediction of Albert Einstein's theory of General Relativity; they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
 The effect of gravitational waves is a change in lengths: a ``stretching and squashing'' of the distance between objects. 
-Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013} use the interference of laser light to measure changes in distance.
+Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO), Advanced Virgo, GEO600, and KAGRA use the interference of laser light to measure changes in distance.
 These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
-Table-top interferometers are commonly used in undergraduate laboratory experiments~\cite{UgoliniEtAl:2019,LibbrechtBlack:2015} and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{TTExhibit:2021}.
+Table-top interferometers are commonly used in undergraduate laboratory experiments and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{TTExhibit:2021}.
 
 % CW 
-To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and neutron star-black hole mergers~\cite{GWTC-1:2018,GWTC-2:2020,NSBH:2021}.
+To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and neutron star-black hole mergers~\cite{GWTC-2:2020,NSBH:2021}.
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
@@
 %Scorpius X-1 is an LMXB that is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
-Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~\cite{ScoX1O2Viterbi:2019}.
+Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~(e.g., Ref~\cite{ScoX1O2Viterbi:2019}).
 
@@
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
-We present the software and scripts used to produce this work in Appendix~\ref{app:code}.
+Further reading and resources are provided in the accompanying Supplementary Material and we present the software and scripts used to produce this work in Appendix~\ref{app:code}.
 

### 2021-07-19 paper/ifo-supplementary-material.tex
moving some more references

@@
 
-\subsection{Gravitational wave obesrvations}
+\subsection{Gravitational-wave observatories and observations}
+The current network of ground-based gravitational-wave observatories includes the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013}.
 Gravitational waves have been observed from compact binary coalescences. 
 For more information on observations of binary black hole mergers see Refs~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020,GWTC-1:2018}, for binary neutron star mergers see Refs~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020} and for neutron star-black hole mergers~\cite{NSBH:2021,GW190814}.
+More information about gravitational waves can be found in Refs.~\cite{Maggiore:2007,CBCs:2014,ObsScenarioDoc:2020,Einstein:1916}.
 
@@
 
-\subsection{Software}
+\subsection{Hardware and software}
+In Section~V in the main text, we make use of a Raspberry Pi to record data from the interferometer. 
+For further information and documentation see Refs~\cite{RaspberryPi:online,pidocumentation:online}.
+
 This work is implemented in Python Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}. 
 We make use of Python packages: numpy~\cite{numpy}\footnote{\url{https://numpy.org/doc/stable/reference/routines.fft.html}}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
+
+\subsection{speech}
+For more information on speech intelligibility see Refs.~\cite{speech_intelligibility}
 

### 2021-07-19 paper/ifo-wanderingTone.tex
moving some more references

@@
 The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises ${\rm Pr}[ f(t_0), f(t_1),... f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
-We note that there is no marginalization \han([What does marginalization mean?
-Do you mean no hard barriers?]) used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
+We note that there is no marginalization \han{[What does marginalization mean?
+Do you mean no hard barriers?]} used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
 %The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
@@
 The Viterbi algorithm provides a computationally efficient method for finding the optimal path. 
-At every $t_i$ all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}). 
+At every $t_i$ all but $N_f$ possible paths are eliminated. 
 Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}.

### 2021-07-20 paper/ifo-appendix.tex
moving some more references
further edits, moving references over to supplementary

@@
 \label{app:viterbi}
-
 This appendix contains some details regarding the implementation of the Viterbi algorithm described in Section~\ref{sec:viterbi}. 
 The Viterbi algorithm~\cite{Viterbi:1967} is a classic method in signal processing, whose theoretical underpinnings and implementation are accessible to undergraduate students. 
-The interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001} to learn more.
+See the Supplementary Material for further resources and other pseudocode examples available online.
+% moved to supp. matt. The interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001} to learn more.
 
@@
 Here, we present some pseudocode of the implementation used in Section~\ref{sec:viterbi}.
-More general pseudocode complete with technical language is available elsewhere online~\cite{viterbipseudocode:online}.
 We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights.

### 2021-07-20 paper/ifo-complexAudio.tex
further edits, moving references over to supplementary

@@
 Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. 
-Resources for using the Pi and photodiode, including a circuit diagram, are described in Appendix~\ref{app:circuit_diagram}.
+Resources for using the Pi and photodiode, including a circuit diagram, are described the Supplementary Material.
 
@@
 Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself. 
-The appearance of harmonics of the mains noise is likely due to the non-linearity discussed in Appendix~\ref{app:intensity_derivation} (see also Ref.~\cite{feynman}). 
+The appearance of harmonics of the mains noise is likely due to non-linearity in the system (see also Ref.~\cite{feynman}). 
 The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.

### 2021-07-20 paper/ifo-conclusions.tex
further edits, moving references over to supplementary

@@
 In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community.
-We explore the use of the interferometer as an optical microphone, considering a more general treatment of signal processing with complex audio signals, which can also serve as a more distant analog to un-modeled gravitational-wave burst signals, e.g., from supernova~\cite{MillhouseEtAl:2018,SuvorovaPowellMelatos:2019}.
+We explore the use of the interferometer as an optical microphone, considering a more general treatment of signal processing with complex audio signals, which can also serve as a more distant analog to un-modeled gravitational-wave burst signals, e.g., from supernova.
 The demonstration can be adapted for use in the undergraduate laboratory and as an tool for explaining gravitational-wave research to a wider, non-specialist audience.

### 2021-07-20 paper/ifo-futureWork.tex
further edits, moving references over to supplementary

@@
 Isolating the experiment from interference may provide some noise reduction. 
-For example, the Raspberry Pi could be powered with a commercial battery instead of mains power. 
-All other components of the circuit draw power from the Pi, so this would isolate the circuit from the mains. 
-The laser is DC-powered through a converter, so the laser should not couple noise from the mains into the interferometer. 
+The Raspberry Pi could be powered with a commercial battery instead of mains power, which may help to isolate the circuit from the mains as all other components draw power from the Pi (the laser is DC-powered through a converter).
 Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}.
@@
 
-More traditional techniques such as a wavelet transform (WT) \citep{nason1995stationary} could be used to extract the signal from noise, and compared with the above methods. 
+More traditional techniques such as a wavelet transform (WT)~\citep{nason1995stationary} could be used to extract the signal from noise, and compared with the above methods. 
 A WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. 

### 2021-07-20 paper/ifo-singleTone.tex
moving references to further reading section after review
further edits, moving references over to supplementary

@@
 The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}.
-The Discrete Fourier transform was calculated using the NumPy package for Python~\cite{numpy}~\footnote{\url{https://numpy.org/doc/stable/reference/routines.fft.html}}.
+The Discrete Fourier transform was calculated using the NumPy package for Python (see Appendix~\ref{app:code}).
+% moved refs~\cite{numpy}~\footnote{\url{https://numpy.org/doc/stable/reference/routines.fft.html}}.
 We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$.
@@
 The peaks at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ have amplitudes of $0.103$ and $0.098$ as a fraction of the height of the main peak, respectively.
-They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo} and Appendix~\ref{app:intensity_derivation}.
+They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo}.
 Also, note that the noise appears to rise at lower frequencies.

### 2021-07-20 paper/ifo-supplementary-material.tex
further edits, moving references over to supplementary

@@
 
-\subsection{Continuous wave searches}
+\subsection{Continuous wave sources and searches}
 \label{sec:GWSsearches}
+Continuous waves are anticipated from rotating neutron stars. 
+For further information about continuous waves from isolated neutron stars see for example Refs.~\cite{MelatosDouglassSimula:2015,Jones:2010} and from neutron stars in binaries see for example Ref.~\cite{BildstenTB:1998}.
+
 The methods used in Section~IV are inspired by hidden Markov model (HMM) searches for continuous gravitational waves. 
-The HMM method is described in Refs~\cite{SuvorovaEtAl:2017,SuvorovaEtAl:2016,BayleyEtAlSOAP:2019,Viterbi:1967}. 
-HMM searches have targeted low mass X-ray binaries~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017,MiddletonEtAlO2LMXBs:2020}, young supernova remnants~\cite{SNRO3a:2021,MillhouseStrangMelatos:2020,SunEtAlSNR:2018}, post merger remnants~\cite{PostMergerRemnantSearch:2019}, Fomalhaut b~\cite{JonesSun:2020}, and high energy TeV sources~\cite{BeniwalEtAl:2021}.
+The HMM method is described in Refs~\cite{SuvorovaEtAl:2017,SuvorovaEtAl:2016,BayleyEtAlSOAP:2019,Viterbi:1967} and further information in Ref.~\cite{JKS:1998}. 
+For more information on the Viterbi algorithm, the interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001} to learn more.
+Alternative pseudocode examples (to that described the appendix) can be found online at Ref.~\cite{viterbipseudocode:online}.
+
+HMM continuous-wave searches have targeted low mass X-ray binaries~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017,MiddletonEtAlO2LMXBs:2020}, young supernova remnants~\cite{SNRO3a:2021,MillhouseStrangMelatos:2020,SunEtAlSNR:2018}, post merger remnants~\cite{PostMergerRemnantSearch:2019}, Fomalhaut b~\cite{JonesSun:2020}, and high energy TeV sources~\cite{BeniwalEtAl:2021}.
 
@@
 We refer the reader to Refs.~\cite{MSPSearchO3O2O1:2020, RadiometerO1O2:2019, SearchRadiometerO1:2017, KnownPulsarsO2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} for searches that consider particular targets or sky locations, and to Refs.~\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, StackSlide:2012} for search methods that cover the entire sky.
-For further information on searching for transient gravitational wave signals see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
+For further information on searching for transient gravitational wave signals see for example Refs.~\cite{SuvorovaPowellMelatos:2019,MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999,}.
 
@@
 Specific examples of using table-top interferometers as engagement or teaching tools include Refs.~\cite{TTExhibit:2021,UgoliniEtAl:2019,LibbrechtBlack:2015,AMIGO:online,FoxEtAl:1999} and parts lists or instructions can be found in Refs.~\cite{TTExhibit:online,LIGOIFOGlue,LIGOIFOMagnets,NikhefIFO,ThorLabsIFO}.
+The analogy between sound and gravitational-wave signals is explored in Refs.~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
 

### 2021-07-20 paper/ifo-tableTopGWs.tex
further edits, moving references over to supplementary

@@
 For example, gravitational waves can propagate in a vacuum, whereas sound cannot.
-Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}.
+Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations.
 Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see also the further reading section in supplementary material for a selection of other table-top Michelson interferometer designs used for science communication).
@@
 The intensity timeseries is recorded by webcam (Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}) or photodiode, which offers a higher sampling rate suitable for capturing more complex audio (Section~\ref{sec:optical_microphone}).
-Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2020}.
+Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2021}.
 
@@
 Even without over-counting, large fringe motions display a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting---phenomena like frequency doubling.
-We derive this nonlinear relationship between the length difference of the arms and the intensity of the pattern at some point on the screen in Appendix~\ref{app:intensity_derivation}.
 

### 2021-07-20 paper/ifo-wanderingTone.tex
further edits, moving references over to supplementary

@@
 
-A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars\cite{MelatosDouglassSimula:2015,Jones:2010}, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs~\cite{BildstenTB:1998}.
+A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs.
 Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. 

### 2021-07-20 paper/paper_main.tex
further edits, moving references over to supplementary

@@
 
-\author{\textbf{[Authors removed for double-anonymous review.]}}
-
-\begin{comment}
+%\author{\textbf{[Authors removed for double-anonymous review.]}}
 
@@
 \begin{acknowledgments}
-\textbf{[Removed for double-anonymous review.]}
+%\textbf{[Removed for double-anonymous review.]}
 
-\begin{comment}
 The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky, and Sam Cooper, for their helpful feedback and suggestions.
@@
 This work has been assigned LIGO document number P2000386.
-\end{comment}
 

### 2021-07-30 paper/responseRound2/AJPResponseLetter.tex
adding first version of response letter

\documentclass[a4paper, 10pt]{letter}

\usepackage{fullpage}

\begin{document}

Dear Joseph,

Thank you for the comments and feedback. 
We attach a new version of the manuscript.

We completely understand about the reference list and we have moved the majority of them into a ``Further Reading'' section in the Supplementary Material.

You also asked about whether we have had any experience using this apparatus in the classroom. Unfortuntately we have not and we are unsure when this will be possible as we are in and out of lockdown in Melbourne. Before COVID, we used the interferometer at several science fairs and schools, however it did not have the audio set-up at that stage, so I think that experience is not directly relevant to this work.

We have also made changes based on the annotated PDF from you and below answer the annotations where clarifications, etc. were needed.

Thank you again for the feedback.

Best wishes, \\
Hannah (for the co-authors)

\begin{itemize}
\item p4 Figure 4 
\begin{itemize}
\item We are happy to remove figure 4. We added it only after the inital review in response to one suggestion of adding a comparison of what would be detected without using Viterbi, but I agree that it does not add much to the explanation. We have kept a shortened version of the higlighted text to explain and we have removed the figure. 
\end{itemize}

\item p4 Section IV A
\begin{itemize}
\item The expanded literature review here was one of the requests from Referee 3. We agree that they are tangentally related to the work presented, so we have instead moved them to the Further Reading section in the Supplementary Material. 
\end{itemize}

\item p5 Figure 5 [about Figure colour scheme]
\begin{itemize}
\item We found it very difficult to make a tidy figure using different sized circles. Instead, we have tried a gray color scheme, which looks okay and should be fine in black and white print. 
\end{itemize}

\item p5, right column, para 2, line 1
\begin{itemize}
\item I see what you mean. The first sentence refers to our ``prior beliefs'' before we look at any data. The second sentence is about the first step of the analysis once we have the data.  
[to do]
\end{itemize}

\item p5, right column, list item 1
\begin{itemize}
\item Yes, the nodes does mean the circles. Thank you, this was confusing and we've made the language more consistent now and refered to the circles when we first mention nodes.
\end{itemize}

\item p5, right column, list item 3: ``In the exampler of Fig. 5, since A=1/3 or 0 for each transition, it looks like the selected path just links the states with the highest Fourier amplitude. Is it always that simple?''
\begin{itemize}
\item That is a good point and it is not always so simple. In our Figure 5, we have made an example where the signal is particularly loud and can easily be seen by eye, but it is not always the case. In continuous-wave searches we expect that the signal cannot be seen by eye. It might also be the case that a loud single frequency-time bin is not part of Viterbi path recovered. [what to change?]
\end{itemize}

\item p6, left column, after itemized list: ``delete `expeced to be' ''
\begin{itemize}
\item I think we should keep ``expected to be''. The reason is that continuous waves have not yet been detected and although of course we believe they are worth searching for, a more pessimistic person might argue that the signals might not be there at all! 
\end{itemize}

\item p6, left column: ``How does the assumption of white noise get inserted into the VA? How does this assumption affect the results?'' 
\begin{itemize}
\item [to do]
\end{itemize}

\item p6 right column: ``If you elimintate the start-up points from the calculation, does E change significantly?''
\begin{itemize}
\item We have also computed the $E_{\rm rms}$ value when the first four time bins are ignored. We find a slight impovement in recovery and have added a sentence about this to the manuscript at the end of this paragraph. 
\end{itemize}

\item p7, Fig. 6: ``It looks like the frequency bins are 0.3 Hz wide and 3s long. Can you estimate the S/N ratio in this data set to illustrate the power of the VA?''
\begin{itemize}
\item We could estimate S/N for single time blocks of data, however we do not believe that a representative S/N can be estimated for the entire data set. 
\end{itemize}

\item p7, right column, Section V, para 2. ``Just to be clear, what have you changed from the apparatus in earlier sections (if anything), other than the photodiode described below? It reads like you have changed something.''
\begin{itemize}
\item The only thing we have changed is replacing the webcam with the photodiode. As we now using the interferometer to capture more complicated speech and music, we draw comparisons to laser microphones used in other fields.  [to do ]
\end{itemize}

\item p9 Figure 8: ``It's hard to judge the effectiveness of the filtering on speech intelligibility from the graphs. Here's an idea: we could attach a direct link to an audio clip for each row, perhaps a 10 second recitation of a familiar phrase. It would be available to online readers via s simple click. his ould enhance the discuation sigificantly.''
\begin{itemize}
\item [to do ]
\end{itemize}

\item p10 left column, para 2
\begin{itemize}
\item I've tried to clarify the extension involving two interferometers. It would enable us to estimate the direction of the source using the relative time of arrival of the signal at each interferometer. This would help to demonstrate how gravitational-wave detectors constrain the direction of transient gravitational wave signals. 
\end{itemize}

\item p11, appendix B: ``The result you derive is a dc signal, which carries no informatoin, plus a frequency-doubled component for small delta-d. But shouldn't there also be a component at the initial frequency, akin to frequency modulation? Otherwise, Fig. 3 cannot be explained.''
\begin{itemize}
\item Thank you very much for the careful reading and we agree completely - this was an error on our part in the derivation. We have decided to remove this appendix. 
\end{itemize}

\item p11, appendix C: ``This needs more explanation. I certainly don't understand it.''
\begin{itemize}
\item [to do]
\end{itemize}

\end{itemize}

\end{document}

### 2021-07-30 paper/responseRound2/AJPResponseLetter.tex
working on the response and minor changes to text

@@
 \begin{itemize}
-\item p4 Figure 4 
+\item p4 (Figure 4 comments):  I think it is obvious that a single FT of a frequency-varying signal cannot capture the signal, so I don't see the need for Figure 4. 
 \begin{itemize}
 \item We are happy to remove figure 4.
-We added it only after the inital review in response to one suggestion of adding a comparison of what would be detected without using Viterbi, but I agree that it does not add much to the explanation.
-We have kept a shortened version of the higlighted text to explain and we have removed the figure.
+We added it only after the inital review in response to one of the suggestions to add a comparison of what would be detected without using Viterbi.
+We agree, however, that it does not add much to the explanation.
+We have kept a shortened version of the higlighted text to explain why we cannot use the Fourier transform in this case and we have removed the figure.
 \end{itemize}
 
-\item p4 Section IV A
+\item p4 Section IV A: I appreciate your desire to be all-inclusive, but is the highlighted text necessary? It is not directly related to the topic of your paper. 
 \begin{itemize}
-\item The expanded literature review here was one of the requests from Referee 3.
+\item This expanded literature review was one of the requests from Referee 3.
 We agree that they are tangentally related to the work presented, so we have instead moved them to the Further Reading section in the Supplementary Material.
@@
 
-\item p5 Figure 5 [about Figure colour scheme]
+\item p5 Figure 5: This won't show up in B/W. How about indicating the probability by different sizes of the circles? 
 \begin{itemize}
-\item We found it very difficult to make a tidy figure using different sized circles. Instead, we have tried a gray color scheme, which looks okay and should be fine in black and white print. 
+\item We found it very difficult to make a tidy figure using different sized circles. Instead, we now use a gray color scheme, which looks clear and should be fine in black and white print. 
 \end{itemize}
 
-\item p5, right column, para 2, line 1
+\item p5, left column, para 1: Just to make sure I understand: the color of each circle is what you observe?
 \begin{itemize}
-\item I see what you mean. The first sentence refers to our ``prior beliefs'' before we look at any data. The second sentence is about the first step of the analysis once we have the data.  
-[to do]
+\item Yes, that's correct. We have added some clarification to the paragraph.
 \end{itemize}
 
-\item p5, right column, list item 1
+\item p5, right column, para 2: The first two sentences seem to contradict each other. Can you clarify? 
+\begin{itemize}
+\item The first sentence refers to our prior beleifs before we begin any analysis or look at the data.
+Without any prior knowledge, we believe that the signal is equally likely to have started in any frequency bin within the range considered.
+The second sentence is the first stage of the algorithm once we have begun the analysis and looked at the data.
+We have added clarification to the text to explain this.
+\end{itemize}
+
+\item p5, right column, para 1: What does ``marginalisation'' mean? Do you mean no hard barriers?
+\item This statement was added in response to referee 3, who asked about marginalization and spotted an error in this equation in the initial version of the manuscript.
+If we have a probability distribution in multiple dimensions, but we only want to consider the probability distribution of a single dimension, we would sum over (or marginalize) all of the other unwanted dimensions.
+This is not what we are doing in this instance.
+The equation corresponds to a single probability related to a unique path.
+We added this statement about marginalization to ensure no confusion given the referee's question, however we believe the meaning is still maintained without it.
+We have removed the first part of this sentence, keeping only the second part.
+
+\item p5, right column, list item 1: ``node'' meaning one of the circles in Fig. 5? 
 \begin{itemize}

### 2021-07-30 paper/ifo-wanderingTone.tex
working on the response and minor changes to text

@@
 
-In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over at each time. 
+In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over time. 
 This is called the transition probability matrix. 
@@
 The timeseries data from the interferometer is first split into segments. 
-We then take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude at frequency $f_j$ and time $t_i$, $F(t_i,f_j)$ as the detection statistic (this is the emission probability).
+We then take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude $F(t_i,f_j)$ at frequency $f_j$ and time $t_i$, as the detection statistic (this is the emission probability).
 Assuming Gaussian noise, this detection statistic choice maximizes the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.
@@
 $t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-The shading of the circles represents the detection statistic $F(t_i,f_j)$ at each grid point, where dark corresponds to lower values and light to higher values.
+The shading of the circles represents the detection statistic $F(t_i,f_j)$ (the observable) at each grid point, where dark corresponds to lower values and light to higher values.
 
@@
 %The first step of the analysis is to define the probability of the system having frequency $f_j$ at
-\han{rephrase: `The first two sentences seem to contradict each other. Can you clarify?'}
-The signal is equally likely to start in any frequency bin (i.e. the prior is flat). 
-We define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
+
+Before we begin analysing the data, we have no prior knowledge as to which frequency bin the signal starts from (i.e. the prior is flat between the minimum and maximum frequency bins).
+
+At the first stage of the analysis, we define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
 A specific path (which may not be optimal) is written as $f(t_0), f(t_1),..., f(t_n)$. 
@@
 The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises ${\rm Pr}[ f(t_0), f(t_1),... f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
-We note that there is no marginalization \han{[What does marginalization mean?
-Do you mean no hard barriers?]} used in this method; the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
+We note that the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
+
+%We note that there is no marginalization the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
 %The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
@@
 For the result shown in Fig.~\ref{fig:viterbi_overlay}, we find that $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
+If we remove the first four time bins from the $E_{\rm rms}$ calculation, we find a slight improvement in recovery with $E_{\rm rms} = 0.044$.
 

### 2021-07-30 paper/responseRound2/AJPResponseLetter.tex
working on response

@@
 
-\item p5, right column, list item 3: ``In the exampler of Fig.
+\item p5, right column, list item 3: In the example of Fig.
 5, since A=1/3 or 0 for each transition, it looks like the selected path just links the states with the highest Fourier amplitude.
-Is it always that simple?''
+Is it always that simple?
 \begin{itemize}
@@
 In continuous-wave searches we expect that the signal cannot be seen by eye.
-It might also be the case that a loud single frequency-time bin is not part of Viterbi path recovered.
-[what to change?]
+It might also be the case that there are loud noise fluctuations in time-frequency bins.
+In this case they may not be selected by the Viterbi algorithm if the overall value of equation 5 is higher for the signal path than the paths containing the noise.
 \end{itemize}
 
-\item p6, left column, after itemized list: ``delete `expeced to be' ''
+\item p6, left column, after itemized list: delete `expeced to be' 
 \begin{itemize}
-\item I think we should keep ``expected to be''.
-The reason is that continuous waves have not yet been detected and although of course we believe they are worth searching for, a more pessimistic person might argue that the signals might not be there at all!
+\item We believe it is important to keep ``expected to be''.
+The reason is that continuous waves have not yet been detected and although of course we believe they are worth searching for, a more pessimistic person might argue that the signals may not be there at all!
 \end{itemize}
 
-\item p6, left column: ``How does the assumption of white noise get inserted into the VA? How does this assumption affect the results?'' 
+\item p6, left column: How does the assumption of white noise get inserted into the VA? How does this assumption affect the results?
 \begin{itemize}
-\item [to do]
+\item The white noise assumption is implicit in that we do not build any knowledge of coloured noise into the analysis.
+This statement was added in response to referee 3's observation that the noise in our data is not white and they asked how or if we take account of this.
+We do not take account of non-white noise in our analysis (which is also consistent with LIGO analyses which assume white noise even though LIGO noise is not white).
+We also note that the narrow frequency range we consider is less impacted by the increase in noise at low frequency.
+We have added some details to clarify this in the text.
 \end{itemize}
 
-\item p6 right column: ``If you elimintate the start-up points from the calculation, does E change significantly?''
+\item p6 right column: If you elimintate the start-up points from the calculation, does E change significantly?
 \begin{itemize}
@@
 
-\item p7, Fig. 6: ``It looks like the frequency bins are 0.3 Hz wide and 3s long. Can you estimate the S/N ratio in this data set to illustrate the power of the VA?''
+\item p7, Fig. 6: It looks like the frequency bins are 0.3 Hz wide and 3s long. Can you estimate the S/N ratio in this data set to illustrate the power of the VA?
 \begin{itemize}
-\item We could estimate S/N for single time blocks of data, however we do not believe that a representative S/N can be estimated for the entire data set. 
+\item We can only estimate S/N for single time blocks of data, however we do not believe that a representative S/N can be estimated for the entire data set. 
 \end{itemize}
@@
 \begin{itemize}
-\item The only thing we have changed is replacing the webcam with the photodiode.
-As we now using the interferometer to capture more complicated speech and music, we draw comparisons to laser microphones used in other fields.
-[to do ]
+\item The only thing we have changed is replacing the webcam with the photodiode so that the apparatus can be used to capture higher frequency speech and music.
+In this section we use the interferometer to capture more complicated signals, so we draw comparisons to laser microphones used in other fields.
+We have added a sentence to this paragraph to clarify the apparatus changes.
+\end{itemize}
+
+\item p8, left column: This paragraph adds very little new information to the topic of Section V, and interrupts the flow of the paper. I recommend shortening it. 
+\begin{itemize}
+\item Agreed.
+This paragraph was added in response to reviewer 3's request for us to mention other methods of noise reduction from the gravitational-wave literature.
+This seemed like the most appropriate place for it to go, however we agree that it is not closely related to the work presented here.
+We have shortened this discussion to a single sentence and moved the literature review to the Further Reading section in the Supplementary Material.
 \end{itemize}
@@
 \begin{itemize}
-\item [to do ]
+\item This sounds a great idea! We have collated the four sounds files in this directory \textcolor{orange}{add a link here}. Please let us know if you need any further details from us. 
 \end{itemize}

### 2021-07-30 paper/ifo-complexAudio.tex
minor edits

@@
 We use the Michelson interferometer as an ``optical microphone'', where a laser is used to detect sound instead of the electrical components in a conventional microphone.
+The only change to our apparatus is replacing the webcam with a photodiode to allow us to capture higher frequencies neccessary for speech and music (see Section~\ref{sec:photodiode}). 
 Optical microphones have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles. 
@@
 The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ Australian mains power grid signal up to and beyond $8$th harmonic. 
-The mains signal is also present but far weaker in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 
+The mains signal is also present, but far weaker, in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 
 Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself. 
@@
 It amplifies parts of the signal with high signal-to-noise (SNR) ratio while suppressing parts with low SNR. 
-The results of the combined cascaded notch and Wiener filter are shown in the third row of Fig.~\ref{fig:notchWienerLogMMSEResults}. 
+The results of the combined cascaded notch and Wiener filter are shown in the third row in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
 Most of the mains noise is removed, however the recovered voice sounds muffled and is not understandable.
@@
 We use an existing implementation of the logMMSE from~\cite{logmmse}.  
-The logMMSE estimator results are shown in the bottom panels of Fig.~\ref{fig:notchWienerLogMMSEResults}. 
+The logMMSE estimator results are shown in the bottom panels in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
 We see significant attenuation of the mains harmonics and general smoothing of the spectrum. 

### 2021-07-30 paper/ifo-wanderingTone.tex
minor edits

@@
 Note that we implicitly approximate the noise at the webcam as white (uniform in frequency) in this implementation of the Viterbi algorithm. 
-This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum} but is a reasonable approximation for the higher-frequency (around 5~Hz) wandering frequency signals that we use.
+This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum}.
+However, White noise is a reasonable approximation for the narrow frequency range (around $3$--$11\,{\rm Hz}$) we use in this analysis.
 

### 2021-07-30 paper/responseRound2/AJPResponseLetter.tex
editing appendix and working on editor comments

@@
 
-\item p9 Figure 8: ``It's hard to judge the effectiveness of the filtering on speech intelligibility from the graphs.
+\item p9 Figure 8: It's hard to judge the effectiveness of the filtering on speech intelligibility from the graphs.
 Here's an idea: we could attach a direct link to an audio clip for each row, perhaps a 10 second recitation of a familiar phrase.
 It would be available to online readers via s simple click.
-his ould enhance the discuation sigificantly.''
+his ould enhance the discuation sigificantly.
 \begin{itemize}
@@
 
-\item p10 left column, para 2
+\item p10 left column, para 2: This seems like an unlikely approach. Why not simply increase the sensitivity, as described in the following sentence, and rotate the apparatus relative to the source?
 \begin{itemize}
-\item I've tried to clarify the extension involving two interferometers.
-It would enable us to estimate the direction of the source using the relative time of arrival of the signal at each interferometer.
-This would help to demonstrate how gravitational-wave detectors constrain the direction of transient gravitational wave signals.
+\item We have clarified this in the text.
+The gain in having data from two interferometer is that directional information can be estimated based on the time delay between the signal arriving at each interferometer, in a similar way to how we are able to tel the direction of a sound with our ears.
+This is also how gravitational-wave detectors estimate the direction of compact binary mergers based on the time-delay of the signal reaching two or more detectors.
+\end{itemize}
+
+\item p10, left column, para 2: Again, this seems like an unlikely strategy, given the sensitivity of the apparatus to mechanical noise. 
+\begin{itemize}
+\item This is a fair assessment. We have removed the part about moving the interferometer. 
+\end{itemize}
+
+\item p10, conclusions comments
+\begin{itemize}
+\item We have removed the repetition and shorted the conclusion. 
 \end{itemize}
@@
 \begin{itemize}
-\item Thank you very much for the careful reading and we agree completely - this was an error on our part in the derivation. We have decided to remove this appendix. 
+\item Thank you for the careful reading and we agree completely - this was an error on our part in the derivation. We have decided to remove this appendix. 
 \end{itemize}
@@
 \begin{itemize}
-\item [to do]
+\item We have expanded the text to clarify the explanation. 
+\end{itemize}
+
+\item Other appendix changes
+\begin{itemize}
+\item In order to reduce references, we have moved the software references in Appendix A and the whole of Appendix E into the Supplementary Material. 
 \end{itemize}

### 2021-07-30 paper/ifo-appendix.tex
editing appendix and working on editor comments

@@
 x(t) = s(t) + n(t)\,, 
+\label{eqn:GNdata}
 \end{equation}
-where $s(t)$ and $n(t)$ are the signal and noise respectively. 
-Then we describe the signal as 
+where $s(t)$ and $n(t)$ are the signal and noise, respectively and  
 \begin{equation}
 s(t) = A \cos\left[{\omega t + \phi}\right]\,,
+\label{eqn:GNmodel}
 \end{equation}
-where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal respectively. 
+where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal, respectively.
+
+The likelihood is proportional to the probability of observing the data given an assumed model of the signal~\cite{Jaynes:2003}.
+In this example the model is given by Eqn.~\ref{eqn:GNmodel}.
+Under the assumption of Gaussian noise, the log-likelihood is 
+\begin{equation}
+\log \mathcal{L} = -\frac{1}{2}\ip{n}{n}\,, 
+\end{equation}
+where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data. 
+We replace $n$ by $x-s$ (Eqn.~\ref{eqn:GNdata}), so that 
+\begin{eqnarray}
+\log \mathcal{L} &=& -\frac{1}{2}\ip{x-s}{x-s}\,, \\ 
+                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{2} \ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]} \nonumber\\
+                 &&+ A \ip{x}{\cos[\omega t + \phi]}\,,\\ 
+                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{4} + A \ip{x}{\cos[\omega t + \phi]}\,,\label{eqn:logLike}
+\end{eqnarray}
+
+\begin{comment}
 The log-likelihood assuming Gaussian noise is given by
@@
 where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data (and the last line assumes that $T\gg1/\omega$ so that $\ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]}\approx{1/2}$).
+\end{comment}
 
@@
 \end{eqnarray}
-From the second term of Eqn~\ref{eqn:maxLogLike}, we see that the maximum log-likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.
+From the second term of Eqn.~\ref{eqn:maxLogLike}, we see that the maximum log-likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.
 %baring the constant factor $\ip{x}{x}$.

### 2021-07-30 paper/ifo-futureWork.tex
editing appendix and working on editor comments

@@
 
-Returning to the demonstration of gravitational wave analysis, we could analyze data from two interferometer to extract directional information for audio signals based on the relative arrival time of the signal at each interferometer~\cite{GW150914}.
+Returning to the demonstration of gravitational wave analysis, single interferometers cannot detect directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal being recorded at two or more detectors~\cite{GW150914}.
+We could therefore extend this analysis to include data from two interferometers to extract directional information.
 This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
 Another extension would be to demonstrate the Doppler effect of the Earth's motion around the Sun, which needs to be considered in continuous-wave searches (see Section~\ref{sec:realCWSearches} and Ref.~\cite{JKS:1998}).
-One approach could be to move the source signal on a circular track or modify the input audio signal to simulate Doppler modulation.
+One approach could be to modify the input audio signal to simulate Doppler modulation.
 

### 2021-07-31 paper/responseRound2/AJPResponseLetter.tex
small edits
spell check

@@
 You also asked about whether we have had any experience using this apparatus in the classroom.
-Unfortuntately we have not and we are unsure when this will be possible as we are in and out of lockdown in Melbourne.
+Unfortunately we have not and we are unsure when this will be possible as we are in and out of lockdown in Melbourne.
 Before COVID, we used the interferometer at several science fairs and schools, however it did not have the audio set-up at that stage, so I think that experience is not directly relevant to this work.
@@
 \begin{itemize}
-\item p4 (Figure 4 comments):  I think it is obvious that a single FT of a frequency-varying signal cannot capture the signal, so I don't see the need for Figure 4. 
+\item p.4 (Figure 4 comments):  I think it is obvious that a single FT of a frequency-varying signal cannot capture the signal, so I don't see the need for Figure 4. 
 \begin{itemize}
 \item We are happy to remove figure 4.
-We added it only after the inital review in response to one of the suggestions to add a comparison of what would be detected without using Viterbi.
+We added it only after the initial review in response to one of the suggestions to add a comparison of what would be detected without using Viterbi.
 We agree, however, that it does not add much to the explanation.
-We have kept a shortened version of the higlighted text to explain why we cannot use the Fourier transform in this case and we have removed the figure.
+We have kept a shortened version of the highlighted text to explain why we cannot use the Fourier transform in this case and we have removed the figure.
 \end{itemize}
 
-\item p4 Section IV A: I appreciate your desire to be all-inclusive, but is the highlighted text necessary? It is not directly related to the topic of your paper. 
+\item p.4 Section IV A: I appreciate your desire to be all-inclusive, but is the highlighted text necessary? It is not directly related to the topic of your paper. 
 \begin{itemize}
 \item This expanded literature review was one of the requests from Referee 3.
-We agree that they are tangentally related to the work presented, so we have instead moved them to the Further Reading section in the Supplementary Material.
+We agree that they are tangentially related to the work presented, so we have instead moved them to the Further Reading section in the Supplementary Material.
 \end{itemize}
 
-\item p5 Figure 5: This won't show up in B/W. How about indicating the probability by different sizes of the circles? 
+\item p.5 Figure 5: This won't show up in B/W. How about indicating the probability by different sizes of the circles? 
 \begin{itemize}
-\item We found it very difficult to make a tidy figure using different sized circles. Instead, we now use a gray color scheme, which looks clear and should be fine in black and white print. 
+\item We found it very difficult to make a tidy figure using different sized circles. Instead, we now use a grey colour scheme, which looks clear and should be fine in black and white print. 
 \end{itemize}
 
-\item p5, left column, para 1: Just to make sure I understand: the color of each circle is what you observe?
+\item p.5, left column, para 1: Just to make sure I understand: the color of each circle is what you observe?
 \begin{itemize}
@@
 
-\item p5, right column, para 2: The first two sentences seem to contradict each other. Can you clarify? 
+\item p.5, right column, para 2: The first two sentences seem to contradict each other. Can you clarify? 
 \begin{itemize}
-\item The first sentence refers to our prior beleifs before we begin any analysis or look at the data.
+\item The first sentence refers to our prior beliefs before we begin any analysis or look at the data.
 Without any prior knowledge, we believe that the signal is equally likely to have started in any frequency bin within the range considered.
@@
 
-\item p5, right column, para 1: What does ``marginalisation'' mean? Do you mean no hard barriers?
-\item This statement was added in response to referee 3, who asked about marginalization and spotted an error in this equation in the initial version of the manuscript.
+\item p.5, right column, para 1: What does ``marginalization'' mean? Do you mean no hard barriers?
+\item This statement was added in response to referee 3, who asked about marginalisation and spotted an error in this equation in the initial version of the manuscript.
 If we have a probability distribution in multiple dimensions, but we only want to consider the probability distribution of a single dimension, we would sum over (or marginalize) all of the other unwanted dimensions.
@@
 
-\item p5, right column, list item 1: ``node'' meaning one of the circles in Fig. 5? 
+\item p.5, right column, list item 1: ``node'' meaning one of the circles in Fig. 5? 
 \begin{itemize}
-\item Yes, the nodes does mean the circles. Thank you, this was confusing and we've made the language more consistent now and refered to the circles when we first mention nodes.
+\item Yes, the nodes does mean the circles. Thank you, this was confusing and we've made the language more consistent now and referred to the circles when we first mention nodes.
 \end{itemize}
 
-\item p5, right column, list item 3: In the example of Fig.
+\item p.5, right column, list item 3: In the example of Fig.
 5, since A=1/3 or 0 for each transition, it looks like the selected path just links the states with the highest Fourier amplitude.
@@
 
-\item p6, left column, after itemized list: delete `expeced to be' 
+\item p.6, left column, after itemized list: delete `expected to be' 
 \begin{itemize}
@@
 
-\item p6, left column: How does the assumption of white noise get inserted into the VA? How does this assumption affect the results?
+\item p.6, left column: How does the assumption of white noise get inserted into the VA? How does this assumption affect the results?
 \begin{itemize}
@@
 
-\item p6 right column: If you elimintate the start-up points from the calculation, does E change significantly?
+\item p.6, right column: If you eliminate the start-up points from the calculation, does E change significantly?
 \begin{itemize}
 \item We have also computed the $E_{\rm rms}$ value when the first four time bins are ignored.
-We find a slight impovement in recovery and have added a sentence about this to the manuscript at the end of this paragraph.
+We find a slight improvement in recovery and have added a sentence about this to the manuscript at the end of this paragraph.
 \end{itemize}
 
-\item p7, Fig. 6: It looks like the frequency bins are 0.3 Hz wide and 3s long. Can you estimate the S/N ratio in this data set to illustrate the power of the VA?
+\item p.7, Fig. 6: It looks like the frequency bins are 0.3 Hz wide and 3s long. Can you estimate the S/N ratio in this data set to illustrate the power of the VA?
 \begin{itemize}
@@
 
-\item p7, right column, Section V, para 2.
+\item p.7, right column, Section V, para 2.
 ``Just to be clear, what have you changed from the apparatus in earlier sections (if anything), other than the photodiode described below?
@@
 
-\item p8, left column: This paragraph adds very little new information to the topic of Section V, and interrupts the flow of the paper. I recommend shortening it. 
+\item p.8, left column: This paragraph adds very little new information to the topic of Section V, and interrupts the flow of the paper. I recommend shortening it. 
 \begin{itemize}
@@
 
-\item p9 Figure 8: It's hard to judge the effectiveness of the filtering on speech intelligibility from the graphs.
+\item p.9, Figure 8: It's hard to judge the effectiveness of the filtering on speech intelligibility from the graphs.
 Here's an idea: we could attach a direct link to an audio clip for each row, perhaps a 10 second recitation of a familiar phrase.
 It would be available to online readers via s simple click.
-his ould enhance the discuation sigificantly.
+This would enhance the discussion significantly.
 \begin{itemize}
-\item This sounds a great idea! We have collated the four sounds files in this directory \textcolor{orange}{add a link here}. Please let us know if you need any further details from us. 
+\item This sounds a great idea! We have collated the four sounds files and attached them in the email. Please let us know if you need any further details from us. 
 \end{itemize}
 
-\item p10 left column, para 2: This seems like an unlikely approach. Why not simply increase the sensitivity, as described in the following sentence, and rotate the apparatus relative to the source?
+\item p.10, left column, para 2: This seems like an unlikely approach. Why not simply increase the sensitivity, as described in the following sentence, and rotate the apparatus relative to the source?
 \begin{itemize}
@@
 
-\item p10, left column, para 2: Again, this seems like an unlikely strategy, given the sensitivity of the apparatus to mechanical noise. 
+\item p.10, left column, para 2: Again, this seems like an unlikely strategy, given the sensitivity of the apparatus to mechanical noise. 
 \begin{itemize}
-\item This is a fair assessment. We have removed the part about moving the interferometer. 
+\item We agree, this is likely to be a fair assessment about the noise of the system. We have removed the part about moving the interferometer. 
 \end{itemize}
 
-\item p10, conclusions comments
+\item p.10, conclusions comments
 \begin{itemize}
@@
 
-\item p11, appendix B: ``The result you derive is a dc signal, which carries no informatoin, plus a frequency-doubled component for small delta-d.
+\item p.11, appendix B: ``The result you derive is a dc signal, which carries no information, plus a frequency-doubled component for small delta-d.
 But shouldn't there also be a component at the initial frequency, akin to frequency modulation?
@@
 
-\item p11, appendix C: ``This needs more explanation. I certainly don't understand it.''
+\item p/11, appendix C: ``This needs more explanation. I certainly don't understand it.''
 \begin{itemize}

### 2021-08-02 paper/ifo-tableTopGWs.tex
proof reading, minor typo fixes and edit

@@
 Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations.
-Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see also the further reading section in supplementary material for a selection of other table-top Michelson interferometer designs used for science communication).
+Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Ref.~\cite{TTExhibit:2021} as well as the further reading section in Supplementary Material for a selection of other table-top Michelson interferometer designs used for science communication).
 Sound vibrations provide a simple means to move the components of a demonstration interferometer, changing the length of the interferometer arms, and therefore the interference pattern.
@@
 A change in the relative length of the arms causes these rings to move radially in or outwards.
-The intensity timeseries is recorded by webcam (Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}) or photodiode, which offers a higher sampling rate suitable for capturing more complex audio (Section~\ref{sec:optical_microphone}).
+The intensity timeseries is recorded by either a webcam (in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}) or a photodiode, which offers a higher sampling rate suitable for capturing more complex audio (in Section~\ref{sec:optical_microphone}).
 Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2021}.
@@
 The interference pattern produced by the table-top Michelson interferometer.
-This image was taken with the webcam aligned off the beam axis. %, which may explain why the (predicted circular) pattern appears elliptical.
+This image was taken with the webcam aligned off the beam axis. 
 The pink star indicates the point where the intensity timeseries measurements were taken.
@@
 As such, any motion of the fringes must be kept small by playing the sound softly through the speaker.
-% too much to have --- and () in the same sentence I think 
 Even without over-counting, large fringe motions display a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting---phenomena like frequency doubling.
 
-% removed as almost identical to the next sentence
-%In the following section, we start simply by injecting and recovering an audio signal of constant frequency (a single tone).
-
-% Data can be recorded via a webcam (as shown in Fig.~\ref{fig:ifo_schematic_webcam}) or photodiode. 
-% In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used. 
-% In Section~\ref{sec:optical_microphone}, which considers more complex audio, we turn to a photodiode because it offers a higher sample rate than the webcam.
-
 \end{document}

### 2021-08-02 paper/ifo-wanderingTone.tex
proof reading, minor typo fixes and edit

@@
 
-%%%%%%%%%%%%%%%%%%%%%% Comment from editor to remove this figure
-%We first demonstrate why a Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case.
-%As shown in Fig.~\ref{fig:viterbi_comparison_webcam_spectrum}, the signal is spread across multiple frequency bins, making it challenging or in some cases impossible to detect.
-
-A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case as the signal is spread across multiple frequency bin. 
-In continuous wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time.
+A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case as the signal is spread across multiple frequency bins. 
+In continuous-wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time.
 The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2017}. 
 In Section~\ref{sec:realCWSearches}, we briefly review the analysis method used by LIGO and Virgo; in Section~\ref{sec:viterbi}, we describe the methods as applied in this work; and in Section~\ref{sec:wanderingResults}, we show results for recovering a wandering frequency signal using the table-top interferometer.
-
-% Removing figure after editor request. 
-\begin{comment}
-\begin{figure}
-	\includegraphics[width=.5\textwidth]{figures/viterbi_comparison_entire_spectrum_viterbi_test_webcam.pdf}
-	\caption{\label{fig:viterbi_comparison_webcam_spectrum}
-Fourier amplitude identical to that shown in Fig.~\ref{fig:webcam_spectrum}, except using a signal which has a slowly changing frequency instead of a constant frequency signal.
-The signal is spread between many frequency bins. 
-We see peaks at $5.7\,{\rm Hz}$ and $10.7\,{\rm Hz}$ and refer the reader forward to Fig.~\ref{fig:viterbi_overlay} to see the frequency evolution of this signal. 
-A single Fourier transform of the entire dataset is not well suited to detecting signals with wandering frequencies.
-% just making this a little more concise 
-%Why we use the Viterbi algorithm to recover wandering frequency signals: Here we plot the Fourier amplitude of the intensity pattern against frequency as in Fig.~\ref{fig:webcam_spectrum} for the entire timeseries of the wandering frequency signal in Fig.~\ref{fig:viterbi_overlay}.
-Instead of resolving definite peaks like the constant frequency signal shown in Fig.~\ref{fig:webcam_spectrum}, we only see inconclusive peaks at 5.7~Hz and 10.7~Hz that correspond to the long amount of time that the wandering frequency spends near those values and/or its amplitude there, as seen in the pink dots in Fig.~\ref{fig:viterbi_overlay}.
-With the signal spread out over the frequency bins, we are unable to reliably detect it above the noise and therefore we use the Viterbi algorithm instead.
-%The large values at low frequencies below 0.1~Hz are similar to those in Fig.~\ref{fig:webcam_spectrum} and can be ignored as they are likely inaccurate due to the short length of the signal relative to their period (i.e.\ 300~s only contains 30~periods of a 0.1~Hz signal), but this feature would require further research to fully determine its origin.
-%maybe could be from the long period (100~s) variation?
-}
-\end{figure}
-\end{comment}
 
@@
 
-%\han{from Andrew: sec IVA, first para - since ref mentioned other searches, it would be good to have 1-2 brief sentences accompanying the references which mention what those "other searches" are (broad categories, for instance, just so that there is some guidance for the reader)}
-
-The methods used here are inspired by LIGO-Virgo continuous wave searches. 
-For further details on these methods and continuous wave searches generally we refer the reader to references in the Supplementary Material.
+The methods used here are inspired by LIGO and Virgo continuous-wave searches. 
+For further details on these methods and continuous-wave searches generally we refer the reader to references in the Supplementary Material.
 
@@
 In this context, ``significantly'' means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
-%One method to search for a wandering signal is to split the timeseries data into several shorter segments which are analyzed individually.
 
@@
 
-%Before beginning the analysis, we assume that the signal is equally likely to start in any frequency bin (i.e. the prior is flat). 
-%The first step of the analysis is to define the probability of the system having frequency $f_j$ at
-
-Before we begin analysing the data, we have no prior knowledge as to which frequency bin the signal starts from (i.e. the prior is flat between the minimum and maximum frequency bins).
+Before we begin analyzing the data, we have no prior knowledge as to which frequency bin the signal starts from (i.e. the prior is flat between the minimum and maximum frequency bins).
 
@@
 
-%We note that there is no marginalization the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
-%The probability of the system having frequency $f_j$ at any later time $t_{i+1}$ is then defined recursively by
-%\begin{eqnarray}
-%{\rm Pr}[f(t_{i+1})=f_j] =~& F[t_{i+1},f(t_{i+1})] \nonumber \\
-%                     &\times A[f(t_{i+1}),f(t_i)]  \nonumber \\
-%                     &\times {\rm Pr}[f(t_i)].
-%\label{eqn:recursiveViterbi}
-%\end{eqnarray}
-%for any path through the grid. 
-%The sequence $f(t_0),f(t_1),\dots,f(t_{N_t})$ that maximizes ${\rm Pr}[f(t_{N_t}) = f_j]$ is the optimal Viterbi path terminating in the frequency bin $f_j$. 
-%We can then maximize the latter quantity over $0 \leq j \leq N_f$ to find the optimal Viterbi path overall, i.e. terminating in any frequency bin.
-
 The Viterbi algorithm provides a computationally efficient method for finding the optimal path. 
@@
 Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}.  
-%, where the circles indicate the elements in the spectrogram with the lighter colors corresponding to a higher likelihood.
 The implementation used in this work is available online (see Appendix~\ref{app:code}).
@@
 In Fig.~\ref{fig:viterbi}, it is the path ending at (b) that started at (c) highlighted in orange.
-%\item The Viterbi path (the overall best path) is then found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}).
-In Fig.~\ref{fig:viterbi}, we see that the path ending at (b) started at (c) and is highlighted in orange.
 \end{enumerate}
@@
 In the following section, we present the results of using the Viterbi algorithm with the table-top interferometer data.
-% - just simiplifed this discussion a little and added a reference to the CW searches
-%The key to the Viterbi algorithm is that we can use the signal's correlation in time to improve the signal-to-noise ratio of the frequency of the wandering signal at a particular time.
-Since the noise is assumed to not be correlated in time, if the detection statistic at a given frequency is high at a particular time and at neighboring frequencies at neighboring times then the likelihood that the given frequency at that particular time is from the wandering signal increases.
-By iterating this process through the grid for a suitable choice of the window of neighboring frequencies and times, the Viterbi algorithm is well-suited to finding wandering signals.
-
-% The Viterbi algorithm is an example of a dynamical programming algorithm, where a computation can be broken down into a series of sub-computations.
-% The Viterbi path for a time series sequences spanning $t_0$ to $t_{N_t}$ contains the Viterbi path for sub-sequences of that time series. 
-% The algorithm is computationally efficient, at every time step all but $N_f$ possible paths are eliminated (see also Ref.~\cite{ScoX1ViterbiO1:2017}).
 
@@
 As shown in Fig.~\ref{fig:viterbi_overlay}, the Viterbi algorithm can recover and continues to track the signal after the disturbance.
-%The Viterbi algorithm is somewhat robust against such disturbances, qualitatively recovering the injected path on the other side, as shown in Fig.~\ref{fig:viterbi_overlay}.
 

### 2021-08-02 paper/ifo-complexAudio.tex
proof read

@@
 Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing (folding of frequencies greater than half the sampling rate) into the detected range.
-We include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
+We include an anti-aliasing Sallen-Key filter tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
 This component attenuates any frequencies above $8\,{\rm kHz}$ before they are digitally sampled.
@@
 Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself. 
-The appearance of harmonics of the mains noise is likely due to non-linearity in the system (see also Ref.~\cite{feynman}). 
+The appearance of harmonics of the mains noise is likely due to non-linearity in the system. 
 The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
@@
 
-Firstly, we consider two signal processing techniques used in combination: the cascaded notch~\cite{10.5555/541204} and the Wiener filter~\cite{10.5555/151045}. 
+Firstly, we consider two signal processing techniques used in combination: the cascaded notch and the Wiener filter (see also the Supplementary Material). 
 A notch filter removes signals within a specific frequency range. 
@@
 Secondly, we apply a speech enhancement technique. 
-Ref.~\cite{SubjectiveComparison} compares $13$ speech enhancement methods, finding the log minimum mean-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech (see also Ref.~\cite{Ephraim1984SpeechEU_logMMSE} for details on the logMMSE estimator).
+Ref.~\cite{SubjectiveComparison} compares $13$ speech enhancement methods, finding the log minimum mean-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech (see also the Supplementary Material).
 We use an existing implementation of the logMMSE from~\cite{logmmse}.  

### 2021-08-02 paper/ifo-wanderingTone.tex
proof read

@@
 
-In continuous wave searches, a physical model of the target informs how far the frequency of the signal can wander over time. 
+In continuous-wave searches, a physical model of the target informs how far the frequency of the signal can wander over time. 
 This is called the transition probability matrix. 
 For example, in LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin in each time step~\cite{ScoX1O2Viterbi:2019}.
-In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to (i) remain in the same frequency bin, or (ii) move down one frequency bin (see the Supplementary Material for supernova remnant search references).
+In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to (i) remain in the same frequency bin, or (ii) move down one frequency bin (see the Supplementary Material for LMXB and supernova remnant search references).
 
@@
 Schematic diagram of the Viterbi algorithm. 
-The circles represent elements in a time-frequency grid, labeled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
-The shading of each circle represents the likelihood of a signal being present in each time-frequency element.
+The circular nodes represent elements in a time-frequency grid, labeled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
+The shading of each node represents the likelihood of a signal being present in each time-frequency element.
 The dark to light gradient corresponds to low to high likelihood values in arbitrary units for this schematic. 
@@
 All possible paths are shown by the lines. 
-The black lines show the best path to each circle and the gray lines show rejected paths. 
+The black lines show the best path to each node and the gray lines show rejected paths. 
 Some routes through the grid result in dead-ends with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a).
-At $t=t_{N_t}$ the algorithm chooses the terminating frequency circle which has the highest value given by Eqn.~\ref{eqn:pathprobability}, marked (b). 
-The Viterbi path is the path leading to this circle, highlighted in orange from (b) to the start at (c). 
+At $t=t_{N_t}$ the algorithm chooses the terminating frequency node which has the highest value given by Eqn.~\ref{eqn:pathprobability}, marked (b). 
+The Viterbi path is the path leading to this node, highlighted in orange, from (b) to the start at (c). 
 }
@@
 Fig.~\ref{fig:viterbi} represents a spectrogram with $N_t$ time bins and $N_f$ frequency bins.
-The circles represent the elements of the time and frequency grid, which are labelled as  
-$t_i$ and $f_j$ respectively, where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
-The shading of the circles represents the detection statistic $F(t_i,f_j)$ (the observable) at each grid point, where dark corresponds to lower values and light to higher values.
+The circlular nodes represent the elements of the time and frequency grid, which are labelled as  
+$t_i$ and $f_j$, respectively where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
+The shading of the nodes represents the detection statistic $F(t_i,f_j)$ (the observable) at each grid point, where dark corresponds to lower values and light to higher values.
 
@@
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heat-map shows the spectrogram of the observed signal (similar to that represented by the schematic in Fig.~\ref{fig:viterbi}). 
-In this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous-wave signals to have far lower signal-to-noise ratios. 
+In this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous-wave signals to have far lower signal-to-noise ratios at each time step. 
 The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.
@@
 For the result shown in Fig.~\ref{fig:viterbi_overlay}, we find that $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.
-If we remove the first four time bins from the $E_{\rm rms}$ calculation, we find a slight improvement in recovery with $E_{\rm rms} = 0.044$.
 
@@
 Initially, the injected signal wanders by more than one frequency bin per time bin (i.e.\ faster than our choice of $A(f_k,f_m)$ allows), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$.
+If we remove the first four time bins from the $E_{\rm rms}$ calculation, we find a slight improvement in recovery with $E_{\rm rms} = 0.044$.
 One may be tempted to increase the allowed frequency wander in the algorithm, however, this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.

### 2021-08-02 paper/responseRound2/AJPResponseLetter.tex
small edits

@@
 \item We have also computed the $E_{\rm rms}$ value when the first four time bins are ignored.
-We find a slight improvement in recovery and have added a sentence about this to the manuscript at the end of this paragraph.
+We find a slight improvement in recovery and have added a sentence about this to the manuscript to our results discussion.
 \end{itemize}
@@
 \item We can only estimate S/N for single time blocks of data, however we do not believe that a representative S/N can be estimated for the entire data set.
+We have clarified this in the text were we refer to the signal to noise ratio of real continuous-wave signals in Section IV.C.
 \end{itemize}
 
-\item p.7, right column, Section V, para 2.
-``Just to be clear, what have you changed from the apparatus in earlier sections (if anything), other than the photodiode described below?
-It reads like you have changed something.''
+\item p.7, right column, Section V, para 2: Just to be clear, what have you changed from the apparatus in earlier sections (if anything), other than the photodiode described below?
+It reads like you have changed something.
 \begin{itemize}

### 2021-08-02 paper/ifo-supplementary-material.tex
small edits

@@
 
-\section{Further reading for gravitational wave search....}
+\section{Further reading and resources}
 
-Here we provide a selection of resources and reading material on the following topics: Section A \han{will list topics when finalised}
+Here we provide a selection of resources and reading material on topics related to this work.
 
-\subsection{Gravitational-wave observatories and observations}
-The current network of ground-based gravitational-wave observatories includes the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013}.
-Gravitational waves have been observed from compact binary coalescences. 
+\subsection{Gravitational-wave observatories}
+The current network of ground-based gravitational-wave observatories includes the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015} which includes two United States sites and third LIGO India site under construction in India~\cite{LIGOIndia:2021}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013}.
+
+\subsection{Gravitational-wave observations}
+Gravitational waves have been observed from compact binary coalescences~\cite{GW150914}. 
 For more information on observations of binary black hole mergers see Refs~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020,GWTC-1:2018}, for binary neutron star mergers see Refs~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020} and for neutron star-black hole mergers~\cite{NSBH:2021,GW190814}.
@@
 We refer the reader to Refs.~\cite{MSPSearchO3O2O1:2020, RadiometerO1O2:2019, SearchRadiometerO1:2017, KnownPulsarsO2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} for searches that consider particular targets or sky locations, and to Refs.~\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, StackSlide:2012} for search methods that cover the entire sky.
-For further information on searching for transient gravitational wave signals see for example Refs.~\cite{SuvorovaPowellMelatos:2019,MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999,}.
+For further information on searching for transient gravitational wave signals see for example Refs.~\cite{SuvorovaPowellMelatos:2019, MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
 
 \subsection{Gravitational-wave engagement}
-Numerous engagement activities have been developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online, chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
+Numerous engagement activities have been developed by gravitational-wave research groups around the world~\cite{TTExhibit:2021,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online, chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
 Specific examples of using table-top interferometers as engagement or teaching tools include Refs.~\cite{TTExhibit:2021,UgoliniEtAl:2019,LibbrechtBlack:2015,AMIGO:online,FoxEtAl:1999} and parts lists or instructions can be found in Refs.~\cite{TTExhibit:online,LIGOIFOGlue,LIGOIFOMagnets,NikhefIFO,ThorLabsIFO}.
@@
 
-\subsection{speech}
-For more information on speech intelligibility see Refs.~\cite{speech_intelligibility}
+\subsection{Signal processing and speech}
+Information on the Sallen-Key filter used in Section~V.A. can be found in Ref.~\cite{sallen_key_filter}. 
+We refer the reader to Ref.~\cite{feynman} for details of the harmonics and non-linearity we observe in Section~V.B. 
+The cascaded notch and Wiener filter are described in Section~\ref{sec:opmic} in this Supplementary Material and also in  Refs.~\cite{10.5555/541204,10.5555/151045}.
+Further details about the logMMSE estimator can be found in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and for speech intelligibility see Refs.~\cite{speech_intelligibility,SubjectiveComparison}
 
 \section{Optical microphone signal processing}
+\label{sec:opmic}
 

### 2021-08-02 paper/authorList.tex
moving author list to separate input so that it is consistent between the paper and the supp mat

--- 2021-07-19 paper/ifo-supplementary-material.tex
@@
-% GravExplainSuppMatt
-
-\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-1}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb,amsthm}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{mathtools}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{hyperref}
-\usepackage{textcomp}
-\usepackage{subfiles}
-\usepackage{comment}
-\usepackage{algpseudocode}
-\usepackage{lmodern}
-
-\usepackage{silence}
-\WarningFilter{revtex4-1}{Repair the float}
-
-\newcommand{\jam}{\textcolor{magenta}}
-\newcommand{\han}{\textcolor{orange}}
-
-\begin{document}
-
-\title{Supplementary Material for \\Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
-
-\author{\textbf{[Authors removed for double-anonymous review.]}}
-
-\begin{comment}
-
 \author{James W. Gardner}
@@
 }
-\end{comment}
-
-\maketitle
-
-\section{Further reading for gravitational wave search....}
-
-Here we provide a selection of resources and reading material on the following topics: Section A \han{will list topics when finalised}
-
-\subsection{Gravitational-wave observatories and observations}
-The current network of ground-based gravitational-wave observatories includes the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013}.
-Gravitational waves have been observed from compact binary coalescences. 
-For more information on observations of binary black hole mergers see Refs~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020,GWTC-1:2018}, for binary neutron star mergers see Refs~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020} and for neutron star-black hole mergers~\cite{NSBH:2021,GW190814}.
-More information about gravitational waves can be found in Refs.~\cite{Maggiore:2007,CBCs:2014,ObsScenarioDoc:2020,Einstein:1916}.
-
-\subsection{Continuous wave searches}
-\label{sec:GWSsearches}
-The methods used in Section~IV are inspired by hidden Markov model (HMM) searches for continuous gravitational waves. 
-The HMM method is described in Refs~\cite{SuvorovaEtAl:2017,SuvorovaEtAl:2016,BayleyEtAlSOAP:2019,Viterbi:1967}. 
-HMM searches have targeted low mass X-ray binaries~\cite{ScoX1O2Viterbi:2019, ScoX1ViterbiO1:2017,MiddletonEtAlO2LMXBs:2020}, young supernova remnants~\cite{SNRO3a:2021,MillhouseStrangMelatos:2020,SunEtAlSNR:2018}, post merger remnants~\cite{PostMergerRemnantSearch:2019}, Fomalhaut b~\cite{JonesSun:2020}, and high energy TeV sources~\cite{BeniwalEtAl:2021}.
-
-A range of other continuous wave searches exist. 
-We refer the reader to Refs.~\cite{MSPSearchO3O2O1:2020, RadiometerO1O2:2019, SearchRadiometerO1:2017, KnownPulsarsO2:2019, SearchCrossCorrO1:2017, SearchTwoSpecS6:2017, VirgoCrabVelaSearch:2015, StackSlide:2012} for searches that consider particular targets or sky locations, and to Refs.~\cite{UnknownNSSearchO3a:2021, EinsteinAtHomeO2:2021, StackSlide:2012} for search methods that cover the entire sky.
-For further information on searching for transient gravitational wave signals see for example Refs.~\cite{MillhouseEtAl:2018, Mohanty:2017, AddessoEtAl:2016, ThraneCoughlin:2014, ThraneEtAl:2011, CandesEtAl:2008, ChassandeMottinArchana:2006, AndersonBalasubramanian:1999}.
-
-\subsection{Gravitational-wave engagement}
-Numerous engagement activities have been developed by gravitational-wave research groups around the world~\cite{TTExhibit:2020,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online, chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
-Specific examples of using table-top interferometers as engagement or teaching tools include Refs.~\cite{TTExhibit:2021,UgoliniEtAl:2019,LibbrechtBlack:2015,AMIGO:online,FoxEtAl:1999} and parts lists or instructions can be found in Refs.~\cite{TTExhibit:online,LIGOIFOGlue,LIGOIFOMagnets,NikhefIFO,ThorLabsIFO}.
-
-\subsection{Environmental noise in gravitational-wave detectors}
-In Section~V~B in the main article, we see evidence for environmental noise in our data. 
-In gravitational-wave detectors reducing and mitigating environmental noise is an active area of research~\cite{EfflerEtAl:2015}. 
-One such noise source is due to seismic ground motion from the Earth. 
-In gravitational-wave observatories, the optical components are suspended from a series of pendulums to achieve passive isolation, and feedback control loops provide active isolation\cite{MatichardEtAl:2015, AdvancedVirgo:2015}.
-Such a system is beyond the scope of this article, however feedback control has been demonstrated in table-top Michelson interferometers designed for the undergraduate laboratory~\cite{LibbrechtBlack:2015}.
-Other sources of noise include acoustic sound waves (for example from nearby vehicles, wind, and components such as electric fans) and magnetic noise (for example from high-voltage power lines, building heaters and lights, or any electrical power circuit)~\cite{EfflerEtAl:2015,CironeEtAl:2019}.
-In some cases, environmental noise sources can be removed or mitigated on site~\cite{CovasEtAl:2018}. 
-Where this is not possible, noise subtraction can be done after data collection~\cite{DriggersEtAl:2019, DavisEtAl:2019}. 
-One example is the removal of spectral noise lines from the United States power grid operating at $60\,{\rm Hz}$~\cite{VajenteEtAl:2020}. %United States's?
-
-\subsection{Hardware and software}
-In Section~V in the main text, we make use of a Raspberry Pi to record data from the interferometer. 
-For further information and documentation see Refs~\cite{RaspberryPi:online,pidocumentation:online}.
-
-This work is implemented in Python Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}. 
-We make use of Python packages: numpy~\cite{numpy}\footnote{\url{https://numpy.org/doc/stable/reference/routines.fft.html}}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
-
-\subsection{speech}
-For more information on speech intelligibility see Refs.~\cite{speech_intelligibility}
-
-\section{Optical microphone signal processing}
-
-In Section~V in the main article, we use the interferometer as an optical microphone; audio signals are used to vibrate the interferometer mirrors, the changing interference pattern is recorded, and we aim to recover the original signal from the resulting timeseries data.
-In this Supplementary Material, we include a selection of signal processing techniques applied to the data which may be considered for undergraduate laboratory examples. 
-We start with some na{\"i}ve approaches and then move to traditional signal processing filters (band-passing and cascaded notches). 
-We finish by combining these with advanced statistical techniques (Wiener filter and the logMMSE estimator). 
-As an example, we test all of these filters on the same $1\,{\rm s}$ long speech recording.
-For more information about the techniques used here see Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045, DigitalProcingOfSpeechSignals:1978}.
-
-\begin{figure*}
-	\includegraphics[height=0.41\textwidth]{figures/setup_pic2.pdf}
-	\includegraphics[height=0.41\textwidth]{figures/circuit_pic3.pdf}
-	\caption{
-Left: A photograph of the optical microphone with an inset of the photodiode.
-The Michelson interferometer is shown on the left and the circuit and Raspberry Pi on the right.
-In the main image, the photodiode is placed behind a cloth screen as explained in the main article.
-The inset shows a face-on view of the photodiode with the cloth screen removed.
-Right: A photograph of the photodiode circuit assembled on a breadboard. The leads from the photodiode enter from the left of the picture. The Raspberry Pi  is shown at the top of the picture.
-}
-	\label{fig:photographs_of_optical_microphone_and_circuit}
-\end{figure*}
-
-All analysis is done using the photodiode set-up as described in Section~V.A. in the main article. 
-A photograph of the equipment is shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit} with the interferometer and circuit shown on the left and right, respectively. 
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-The photodiode circuit diagram is shown in Figure~\ref{fig:circuit_diagram}
-The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an analogy-to-digital converter ADC to a Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
-Useful throughout was the Raspberry Pi documentation~\cite{pidocumentation:online} and the pinoutXYZ reference~\cite{pipinout:online}.
-
-\begin{figure*}
-	\includegraphics[width=.8\textwidth]{figures/circuit_diagram_2.pdf}
-	\caption{\label{fig:circuit_diagram}
-Circuit diagram for reading the photodiode. 
-The photodiode is connected in reverse-bias across an op-amp (pale-red box). 
-The op-amp output is passed through a Sallen-Key anti-aliasing filter (pale-blue box) tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
-The digitized signal is then read by the special-purpose input (SPI) pins of a Raspberry Pi in standard configuration.
-This diagram is produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online} and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-}
-\end{figure*}
-
-\begin{figure}
-	\begin{center}
-    %% trim order is <left> <lower> <right> <upper>
-    % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
-	\includegraphics[width=.5\textwidth]
-% trim={1cm 3cm 1cm 1cm}]
-{figures/combined_results_full_melatos_labelled.pdf}
-%notch_and_wiener_superplot_v2.pdf}
-	\end{center}
-	\caption{\label{fig:op_mic_results}
-	%\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}
-	Timeseries (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
-    The top two rows are the input signal and raw output (identical to the top two rows in Fig.~7 in the main article).  
-    The third row shows a background noise recording from the optical microphone when no audio signal is played. 
-    The fourth row shows the result of subtracting the background noise spectrum in the third row from the recording in the second row. 
-	The fifth and sixth rows show the results of applying the cascaded-notch filter and Wiener filter respectively.
-    The seventh shows the results of applying both the cascaded-notch and Wiener filter (identical to the third row in Fig.~7 in the main article).
-    The eighth row shows the results of applying the logMMSE estimator (identical to the fourth row in Fig.~7 in the main article). 
-	}
-\end{figure}
-
-Results for a selection of filters are collated in Fig.~\ref{fig:op_mic_results}, which we refer to throughout this Supplementary Material. 
-The left and right columns show timeseries and Fourier spectrum results, respectively. 
-The first and second rows in Fig.~\ref{fig:op_mic_results} show the original source signal and the raw optical microphone recording, respectively. 
-The following sections describe a sequence of analysis techniques and a selection of their results is presented in Fig.~\ref{fig:op_mic_results}.
-
-\subsection{Background noise subtraction}
-
-The first technique we consider is quite intuitive: we try to remove noise by directly subtracting the background noise from the recorded spectrum. 
-The third row in Fig.~\ref{fig:op_mic_results} shows the interferometer background noise when there is no input signal.
-The fourth row shows the spectrum obtained after subtracting the background noise spectrum, i.e.\ the result of subtracting the third row from the second row. 
-We see no obvious improvement, which may be attributed to a time-variant noise spectrum, the cause of which is unidentified.
-
-\subsection{Rectangular comb filter}
-
-Since the mains harmonics are present in the signal, we try to selectively remove them.
-However, simply zeroing the frequency bins corresponding to the harmonics of the $50\,{\rm Hz}$ mains signal is unsuccessful. 
-This effectively multiplies the spectrum by a rectangular comb filter. 
-It does remove the mains harmonics, but audibly ruins the rest of the signal due to lack of smoothness. 
-This is because applying a filter in frequency space is equivalent to convolving the time-domain signal with the inverse Fourier transform of that filter. 
-The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is some combination of sinc functions, which significantly corrupt the signal. 
-See also Section~\ref{sec:notch} where we explore a notch filter.
-
-\subsection{High-pass filter}
-
-A high-pass filter smoothly attenuates frequencies below some cut-off frequency. 
-Applying a high-pass filter, with a cut-off frequency around $150\,{\rm Hz}$, to the signal spectrum works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics.
-However, the mains harmonics above $100\,{\rm Hz}$ remain.
-A high-pass filter with a higher cut-off can be used to mitigate this issue.
-However, it makes the played-back signal unrecognizable as the region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}.
-Often in speech processing, the logarithm of the signal is taken since the amplitude information seems to be more important to intelligibility than the phase information to the human ear~\cite{SubjectiveComparison}.
-However, applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.
-
-\subsection{Butterworth band-pass filter}
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{figures/butterworth_150_3000.pdf}
-	\caption{
-Butterworth bandpass filter frequency response. 
-%Any amplitude below $-3\,{\rm dB}$ is significant attenuation (less than half the power at that frequency remains). 
-% suggest removing as -3db is hard to judge on the figure and I think the attenuation outside the frequency range is the key point to get across here
-The red, dashed vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$. Note the flat response within the band characteristic of the Butterworth filter.}
-	\label{fig:butterworth}
-\end{figure}
-
-A general band-pass filter combines a high-pass filter and a low-pass filter to smoothly attenuate frequencies outside of some band (alt.\ pass-band).
-A Butterworth band-pass filter is a particular band-pass filter such that the frequency response (the attenuation at each frequency) is ``maximally flat'' within the band.
-The Butterworth low-pass component is given by %Eqn.~\ref{eq:butterworth}
-
-\begin{equation}
-\label{eq:butterworth}
-H(f) = \left[1+\varepsilon^2 \left( \frac{f}{f_c} \right)^{2n}\right]^{-1/2},
-\end{equation}
-and is combined with a similar high-pass filter to form the band-pass filter.
-In Eqn.~\ref{eq:butterworth}, $f_c$ is the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ is the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
-Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here: a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}$, $3\,{\rm kHz})$.
-This high frequency ($3\,{\rm kHz}$) cut-off is chosen since the frequencies important for speech generally lie below $2\,{\rm kHz}$~\cite{speech_intelligibility}.
-
-The effect of applying this filter to the background noise PSD can be seen in the bottom panel in Fig.~\ref{fig:psd_noise}.
-The Butterworth band-pass filter reduces the amplitude of mains harmonics below $150\,{\rm Hz}$ and suppresses unrelated noise sources above $3\,{\rm kHz}$.
-However, it does not address the issue of mains harmonics above $150\,{\rm Hz}$ (i.e.\ those in the pass-band).
-In the following section, we experiment with a cascade notch filter to address this.
-
-\begin{figure}
-	\includegraphics[width=.5\textwidth]{figures/psd_butterworth_14_6.pdf}
-	\caption{\label{fig:psd_noise}
-Top: power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
-Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
-We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white. 
-After filtering, we see significant attenuation (at least 3~dB) of all frequencies outside the band.
-However, the filter has little change on the harmonics within the band.
-The sharp drop-off at low frequencies, around 100~dB near 0~Hz, is due to stronger attenuation far from the band and would appear shallower if shown against a logarithmic scale for frequency as in Fig.~\ref{fig:butterworth}.
-}
-\end{figure}
-
-\subsection{Cascade notch filter}
-
-\label{sec:notch}
-\begin{figure}
-\begin{center}
-\includegraphics[width=.5\textwidth]{figures/cascaded_notch_response_plot.pdf}
-\end{center}
-\caption{\label{fig:notchMagResponse}
-Amplitude (magnitude) response for the first five notches of fifteen for the cascaded notch filter described by Eqs.~\ref{eqn:notch} and~\ref{eqn:notch15}. 
-}
-\end{figure}
-
-One method to remove mains noise and its harmonics is to use a sequence of notch filters centered at each of the frequency bins we want to remove.
-This sequence is known as a ``cascade'' of notches.
-Here, the notches are smooth in comparison to the na{\"i}ve zeroing of each frequency (which looks like a rectangular comb).
-
-The impulse response describes the reaction of a system to an input signal as a function of time. 
-If a system is linear and time-invariant, the output signals are completely determined for any input signal using the impulse response. 
-Finite impulse response (FIRs) filters have finite length (they equal zero outside of a finite range). 
-In contrast, an infinite impulse response (IIR) filter has infinite length due to feedback. 
-Typically FIR filters outperform IIR filters as they are always stable. 
-However, IIR filters normally require fewer coefficients which in turn speeds up signal processing in comparison to FIR filters.
-
-In this work, we opt for an IIR notch filter. 
-We write the filter in the form of a Z-transform, in which a discrete-time signal is converted to the complex frequency domain $z=e^{jf}$. 
-The IIR notch filter we use here is given by~\citep{10.5555/541204}, 
-\begin{equation}
-    \label{eqn:notch}
-    H(z)=\frac{1+\alpha}{2}\frac{1-2\beta z^{-1}+z^{-2}}{1-\beta(1+\alpha)z^{-1}+\alpha z^{-2}}\,,
-\end{equation}
-where $\alpha$ and $\beta$ are parameters that control the filter. 
-These parameters determine $w_0=\cos^{-1}(\beta)$, the frequency that is completely attenuated (zeroed or ``notched'') at the centre of the notch, and $B_w=\cos^{-1}[2\alpha/(1+\alpha^2)]$, the bandwidth of the notch which determines how quickly the response changes around the notched frequency.
-We find that a sequence of $15$ notches works well here, with the $k^\mathrm{th}$ notch centered on the $k^\mathrm{th}$ harmonic of the $50\,{\rm Hz}$ mains signal, where $k=0,2,\dots,14$. 
-We choose the bandwidth and order (here order six) of each notch to avoid disturbing useful signals while still allowing for uncertainty in the location of each harmonic of the mains signal.
-The response of this cascaded notch filter $H(z)$ is the product of the responses of each of the individual $H_k(z)$ notches,
-\begin{equation}
-    \label{eqn:notch15}
-    H(z) = \prod_{k=0}^{14} H_k(z).
-\end{equation}
-
-We use the built-in MATLAB filter design toolbox in this work~\cite{MATLAB}. 
-The amplitude (magnitude) response of the first five notch filters is shown in Fig.~\ref{fig:notchMagResponse}. 
-The time series and spectrum obtained after applying the cascade notch filter to the speech recording are shown in the fifth row in Fig.~\ref{fig:op_mic_results}.
-We see that the mains harmonics are significantly attenuated compared to the background-subtracted results shown in the fourth row in Fig.~\ref{fig:op_mic_results}.
-
-Although the notch filter removes much of the mains hum sound, the filtered recording is not intelligible.
-Qualitatively, it sounds more like a drum than a human voice.
-This is due to the loss of voice information under the filter.
-To overcome this, we turn to more advanced techniques, starting with the Wiener filter.
-Instead of just passively filtering different frequencies, this statistical technique optimizes an estimate of the original signal.
-
-\subsection{Wiener filter}
-\label{sec:Wiener}
-
-A Wiener filter is an advanced statistical technique that estimates the injected signal given prior information about the injected spectrum and the reference spectrum of the background noise. 
-The observed noisy speech signal sequence is given as $\mathbf{x}=(x(0),\dots, x(N-1))$, where $N$ is the length of the data sequence and $\mathbf{x}$ is the sum of the original injected signal $\mathbf{s}=(s(0),\dots,s(N-1))$, and the noise sequence $\mathbf{w}=(w(0),\dots,w(N-1))$,
-\begin{equation}
-    \mathbf{x}=\mathbf{s}+\mathbf{w}\,.
-\end{equation}
-Given $\textbf{x}$, our goal is to make an estimate $\hat{\textbf{s}}$ of the original signal $\textbf{s}$ such that we minimise the Bayesian mean-square error (BMSE) between the two, defined as 
-\begin{equation}
-\label{eq:BMSE}
-\text{BMSE}(\hat{\textbf{s}})=E[(\textbf{s}-\hat{\textbf{s}})^2]\,.
-\end{equation}
-If we assume that: i) $\textbf{x}$ is ``wide sense stationary''~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in \mathbb{R}$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
-\begin{equation}
-\hat{{s}}[n]=\sum_{k=0}^{n}h[k]x[n-k]\,,
-\end{equation}
-where $\textbf{h}=(h(0),\dots h(n))$ represents the coefficients of an $n$th order Wiener filter.
-The famous Wiener-Hopf equation~\citep{noble1959methods} allows us to determine $\textbf{h}$, as\begin{equation}
-\label{eqn:wiener-hopf}
-\begin{bmatrix}  
-r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
-r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
-\vdots&\vdots&\ddots&\vdots\\
-r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
-\end{bmatrix}
-\begin{bmatrix}
-h[0]\\
-h[1]\\
-\vdots\\
-h[n]
-\end{bmatrix}=
-\begin{bmatrix}
-r_{ss}[0]\\
-r_{ss}[1]\\
-\vdots\\
-r_{ss}[n]
-\end{bmatrix}\,,
-\end{equation}
-where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$, 
-%\begin{equation*}
-%\left\{ 
-\begin{eqnarray} 
-r_{xx}[n] &~=~& E[x(i)~x(i+n)] \,,\\
-r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
-%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
-%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
-\end{eqnarray} 
-%\right\}  
-%\end{equation*}
-
-If we let the Wiener filter be non-causal (i.e.
-we estimate the current signal based on both past \emph{and future} observations), then we can represent Eqn.~\ref{eqn:wiener-hopf} in the frequency domain as
-\begin{equation}
-\label{eqn:wiener}
-    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}\,,
-\end{equation}
-where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data, the injected signal, and the background noise, respectively.
-Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high and attenuates the signal where the SNR is low.
-The causal Wiener filter (which only makes estimates based on past observations) is similar.
-A more detailed analysis of both kinds of Wiener filters can be found in Ref.~\citep{10.5555/151045}.
-
-In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}.
-A higher-order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required.
-For this work, we choose a Wiener filter of order $n=100$ as it provides a reasonable balance between smoothing and efficiency.
-The timeseries and frequency spectrum after applying the Wiener filter are shown in the sixth row in Fig.~\ref{fig:op_mic_results}.
-We see a significant improvement in the timeseries of the recovered signal, however, a strong noise hum persists, audibly.
-
-\subsection{Combined notch and Wiener filter}
-
-Here, we experiment with applying a combination of the cascaded notch and the Wiener filter to the recorded speech signal. 
-The results of this analysis also are described in Section V.C. in the main article.
-
-The Wiener filter makes use of statistical information from the speech data and noise.
-It amplifies the part of the signal with high SNR while suppressing the parts with low SNR (see above Sec.~\ref{sec:Wiener}).
-It is implemented in the form of a finite response filter, which ensures linear phase response and stability (both desirable), but at the cost of high orders computationally.
-By comparison, the notch filter is based on directly removing the unwanted frequency components.
-It is implemented in the form of an infinite impulse response filter.
-Although it significantly decreases the order of the overall filter, it unavoidably introduces nonlinear phase and instability.
-
-By combining the notch and Wiener filter, we can trade-off between the two and achieve an overall better performance, as can be seen in the seventh row in Fig.~\ref{fig:op_mic_results}.
-The filtered voice after the combined notch and Wiener filter is enhanced compared to either alone. 
-The mains noise is all but removed and more voice information is retained. 
-However, the recovered voice still sounds muffled and is not understandable.
-
-\subsection{logMMSE estimator}
-\label{sec:logmmse}
-
-\begin{comment}
-\begin{figure*}
-	%\includegraphics[width=0.49\textwidth]{figures/combined_timeseries_melatos_shift_source.pdf}
-	%\includegraphics[width=0.49\textwidth]{figures/combined_spectrum_melatos.pdf}
-    \includegraphics[height=0.57\textheight]{figures/combined_timeseries_melatos_shift_source_tall.pdf}
-    \includegraphics[height=0.57\textheight]{figures/combined_spectrum_melatos_tall.pdf}
-    \caption{\label{fig:logMMSE_timeseries_freqspectrum}
-\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}    - done 
-Timeseries (left column) and frequency spectrum (right columns) results with the logMMSE estimator. 
-The original source signal and optical microphone recording in the top and middle panels respectively are for comparison and are identical to those shown in the top two panels in Fig.~\ref{fig:BackgroundNotchWienerCombined} (again, the source signal is shifted by $0.12\,{\rm s}$).
-The bottom-left panel shows the timeseries of the recording after filtering with the logMMSE estimator, where the rise at the start of the timeseries is expected when filtering a signal of finite duration.
-The bottom-right panel shows the corresponding frequency spectrum with the logMMSE estimator. 
-The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude $2$ which otherwise dominate the plot. 
-The change in scale from the top panel is not of concern as the ear hears the relative frequency content and normalisation of the timeseries fixes any scaling. 
-    }
-\end{figure*}
-\end{comment}
-
-Speech enhancement of noisy channels is a classic problem in signal processing. 
-In Ref.~\cite{SubjectiveComparison}, a comparison is made of $13$ speech enhancement methods, finding the log minimum mean-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech.
-This estimator is based on speech enhancement techniques discussed in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and minimizes the mean-square error (MSE) of the estimate from the injected signal, like the Wiener filter above, except that it measures the MSE between the logarithm of the Fourier amplitudes.
-This is motivated by the fact that the logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.
-We apply an existing implementation of the logMMSE estimator (see Ref.~\cite{logmmse}) to the recorded signal.
-
-The final row in Fig.~\ref{fig:op_mic_results} shows the results of the logMMSE estimator.
-Again there is significant noise reduction, however, the speech remains indistinct.
-The logMMSE results are also presented in Section V.C. in the main article.
-
-\bibliographystyle{myunsrt}
-\bibliography{ifoDemoBib}
-
-\end{document}

### 2021-08-02 paper/ifo-supplementary-material.tex
moving author list to separate input so that it is consistent between the paper and the supp mat

@@
 
-\author{\textbf{[Authors removed for double-anonymous review.]}}
-
-\begin{comment}
-
-\author{James W. Gardner}
-\email{u6069809@anu.edu.au}
-% \affiliation{%
-% College of Science, Australian National University, Acton, ACT, 2601, Australia 
-% }%
-\affiliation{%
-ANU Centre for Gravitational Astrophysics, The Australian National University, Acton, ACT, 2601, Australia
-}
-\affiliation{%
-OzGrav-ANU, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, The Australian National University, Acton, ACT, 2601, Australia
-} 
-\affiliation{%
-School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
-}
-
-\author{Hannah Middleton}
-\email{hannah.middleton@unimelb.edu.au}
-\affiliation{%
-School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
-}
-\affiliation{%
-Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia 
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{Changrong Liu}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{Andrew Melatos}
-\email{amelatos@unimelb.edu.au}
-\affiliation{%
-School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{Robin Evans}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{William Moran}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
-
-\author{Deeksha Beniwal}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-
-\author{Huy Tuong Cao}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-
-\author{Craig Ingram}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-
-\author{Daniel Brown}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-
-\author{Sebastian Ng}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-\end{comment}
+\input{authorList.tex}
 

### 2021-08-02 paper/paper_main.tex
moving author list to separate input so that it is consistent between the paper and the supp mat

@@
 
-\author{James W. Gardner}
-\email{u6069809@anu.edu.au}
-% \affiliation{%
-% College of Science, Australian National University, Acton, ACT, 2601, Australia 
-% }%
-\affiliation{%
-ANU Centre for Gravitational Astrophysics, The Australian National University, Acton, ACT, 2601, Australia
-}
-\affiliation{%
-OzGrav-ANU, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, The Australian National University, Acton, ACT, 2601, Australia
-} 
-\affiliation{%
-School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
-}
-
-\author{Hannah Middleton}
-\email{hannah.middleton@unimelb.edu.au}
-\affiliation{%
-School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
-}
-\affiliation{%
-Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia 
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{Changrong Liu}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{Andrew Melatos}
-\email{amelatos@unimelb.edu.au}
-\affiliation{%
-School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{Robin Evans}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
-\affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
-}
-
-\author{William Moran}
-\affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
-}
-
-\author{Deeksha Beniwal}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-
-\author{Huy Tuong Cao}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-
-\author{Craig Ingram}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-
-\author{Daniel Brown}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
-
-\author{Sebastian Ng}
-\affiliation{%
-Department of Physics, The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
-}
-\affiliation{%
-OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
-}
+\input{authorList.tex}
 

### 2021-08-02 paper/authorList.tex
adding commas

@@
 \affiliation{%
-Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia 
+Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria, 3122, Australia 
 }
 \affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria, 3010, Australia
 }
@@
 \affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria, 3010, Australia
 }
 \affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria, 3010, Australia
 }
@@
 \affiliation{%
-OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria 3010, Australia
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria, 3010, Australia
 }
@@
 \affiliation{%
-Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria 3010, Australia
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria, 3010, Australia
 }

### 2021-08-03 paper/ifo-complexAudio.tex
proof-read after integration of 2nd round of referee comments

@@
 We use the Michelson interferometer as an ``optical microphone'', where a laser is used to detect sound instead of the electrical components in a conventional microphone.
-The only change to our apparatus is replacing the webcam with a photodiode to allow us to capture higher frequencies neccessary for speech and music (see Section~\ref{sec:photodiode}). 
+The only change to our apparatus is replacing the webcam with a photodiode to allow us to capture higher frequencies necessary for speech and music (see Section~\ref{sec:photodiode}). 
 Optical microphones have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles. 
@@
 
-This analysis is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals. 
+This analysis is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals. % citation?
 In addition, the apparatus serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories. 
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
-We consider a selection of filter techniques, details of which can be found in the Supplementary Material, where we also refer the reader to a selection of digital signal processing resources for further information.
-In Section~\ref{sec:opticalMicResults}, we present the two best-performing analyses from our Supplementary Material.
+We consider a selection of filter techniques, details of which along with a summary of digital signal processing resources can be found in the Supplementary Material.
+ % We also refer the reader to a selection of digital signal processing resources for further information. 
+In Section~\ref{sec:opticalMicResults}, we present the two best-performing techniques from the Supplementary Material.
 
@@
 The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
-Speech intelligibility (being able to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$. 
+Speech intelligibility (the ability to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$. 
 Therefore, the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music (adjusting for the Nyquist frequency). 
@@
 %This cloth screen mount was the ``grill cloth'' from the front of a dismantled commercial speaker, re-purposed. 
-The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2 (henceforth: Pi), which provides a convenient means to record the photodiode data.
+The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2, which provides a convenient means to record the photodiode data.
 Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. 
-Resources for using the Pi and photodiode, including a circuit diagram, are described the Supplementary Material.
+Resources for using the Raspberry~Pi and photodiode, including a circuit diagram, are described in the Supplementary Material.
 
@@
 
-The raw output of the optical microphone (with anti-aliasing) is noisy with a loud bass continuous hum. 
-This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., with the speaker switched off), shown in Fig.~\ref{fig:psd_noise}. 
+The raw output of the optical microphone (with anti-aliasing) is noisy with a loud, continuous bass hum. 
+This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., the output with the speaker switched off), shown in Fig.~\ref{fig:psd_noise}. 
 The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ Australian mains power grid signal up to and beyond $8$th harmonic. 
@@
 Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself. 
-The appearance of harmonics of the mains noise is likely due to non-linearity in the system. 
+The appearance of harmonics of the mains noise might be due to the non-linearity in the system discussed in Section~\ref{sec:ifo}. 
 The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
-Environmental noise reduction for gravitational-wave detectors is an active area of research (see the accompanying Supplementary Material for further information and resources on this topic).
+Environmental noise reduction for gravitational-wave detectors is an active area of research (see the Supplementary Material for further information and resources on this topic).
 
@@
 Timeseries (left column) and frequency spectrum (right column) results with the optical microphone. 
-The original input signal in the first row is a $1\,{\rm s}$ recording of an adult male voice (saying ``A cathode''). 
+The original input signal in the first row is a $1\,{\rm s}$ recording of an adult male voice (saying ``a cathode''). 
  The input signal is shifted by 0.12 s to the right to synchronize the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker. 
 The second row shows the raw output from the optical microphone when the input from the first row is played. 
-The third row shows the result of applying the notch and Wiener filter combined. 
+The third row shows the result of applying the notch and Wiener filters combined. 
 The fourth row shows the result of applying the logMMSE estimator, where the rise at the start of the timeseries is an expected effect when filtering a signal of finite duration. 
@@
 
-We explore several filters to remove the $50\,{\rm Hz}$ mains hum and harmonics, and also to improve the speech intelligibility of the recording.
-The accompanying Supplementary Material describes a range of analysis techniques that can be used as examples for the undergraduate laboratory. 
+We explore several filters to remove the $50\,{\rm Hz}$ mains hum and harmonics and improve the speech intelligibility of the recording.
+The Supplementary Material describes a range of analysis techniques that can be used as examples for the undergraduate laboratory. 
 All filters are tested on the same $1\,{\rm s}$ long speech recording.
 The results of this section are shown in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
-In the figure the left and right columns show the timeseries and frequency spectrum respectively. 
+In the figure, the left and right columns show the timeseries and frequency spectrum respectively. 
 The first row shows the input signal played through the speaker (see Fig.~\ref{fig:ifo_schematic_webcam}). 
@@
 In this section, we present the results of two advanced signal processing analysis techniques applied to the optical microphone recordings. 
-The techniques are only briefly described here and we refer the reader to the accompanying Supplementary Material for further details and other analysis techniques.
+The techniques are only briefly described here and we refer the reader to the Supplementary Material for further details and other analysis techniques.
 
@@
 The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. 
-It amplifies parts of the signal with high signal-to-noise (SNR) ratio while suppressing parts with low SNR. 
+It amplifies parts of the signal with high signal-to-noise ratio while suppressing parts with low signal-to-noise ratio. 
 The results of the combined cascaded notch and Wiener filter are shown in the third row in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
-Most of the mains noise is removed, however the recovered voice sounds muffled and is not understandable.
+Most of the mains noise is removed, however, the recovered voice sounds muffled and is not understandable.
 
@@
 Ref.~\cite{SubjectiveComparison} compares $13$ speech enhancement methods, finding the log minimum mean-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech (see also the Supplementary Material).
-We use an existing implementation of the logMMSE from~\cite{logmmse}.  
+We use an existing implementation of the logMMSE from Ref.~\cite{logmmse}.  
 The logMMSE estimator results are shown in the bottom panels in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
 We see significant attenuation of the mains harmonics and general smoothing of the spectrum. 
-Most of the background noise is removed, however, the logMMSE does not significantly enhance the speech as the voice sounds muffled and indistinct.
+Most of the background noise is removed, however, the logMMSE still does not significantly enhance the speech as the voice sounds muffled and indistinct.
 
 We find some improvement with music over speech. 
-Simple chords and drums can be heard, however more composite sounds and complex melodies cannot be heard clearly. 
+Simple chords and drums can be heard after filtering, however, more composite sounds and complex melodies cannot be heard clearly. 
 Our observations suggest that this is especially true for certain instruments; in particular flutes and violins sometimes can’t be heard at all. 

### 2021-08-03 paper/ifo-conclusions.tex
proof-read after integration of 2nd round of referee comments

@@
 In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community.
-We explore the use of the interferometer as an optical microphone, considering a more general treatment of signal processing with complex audio signals, which can also serve as a more distant analog to un-modeled gravitational-wave burst signals, e.g., from supernova.
-The demonstration can be adapted for use in the undergraduate laboratory and as an tool for explaining gravitational-wave research to a wider, non-specialist audience.
+We explore the use of the interferometer as an optical microphone and consider a more general treatment of signal processing with complex audio signals, which can also serve as a distant analog to un-modeled gravitational-wave burst signals, e.g., from supernovae.
+% justify ``un-modeled''? - james
+The demonstration can be adapted for use in the undergraduate laboratory and as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
 

### 2021-08-03 paper/ifo-futureWork.tex
proof-read after integration of 2nd round of referee comments

@@
 Flowchart for the signal through the optical microphone. 
-A transfer function of the system would have to account for each stage and an aim of future work. 
-Not shown are the signal processing filters that are performed on the PC once the data is transferred off the Pi.
+A transfer function of the system would have to account for each stage and is an aim of future work. 
+Not shown are the signal processing filters that are performed on the PC once the data is transferred off the Raspberry~Pi.
 }
@@
 Isolating the experiment from interference may provide some noise reduction. 
-The Raspberry Pi could be powered with a commercial battery instead of mains power, which may help to isolate the circuit from the mains as all other components draw power from the Pi (the laser is DC-powered through a converter).
+The Raspberry Pi could be powered with a commercial battery instead of mains power, which may help to isolate the circuit from the mains as all other components draw power from the Raspberry~Pi (the laser is DC-powered through a converter).
 Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}.
@@
 A thorough characterization of the system's transfer function would allow us to understand how audio signals couple through the interferometer. 
-The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Pi. 
+The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Raspberry~Pi. 
 It would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation, and the photodiode. 
@@
 
-More traditional techniques such as a wavelet transform (WT)~\citep{nason1995stationary} could be used to extract the signal from noise, and compared with the above methods. 
-A WT provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. 
-In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, WT methods are proposed for speech recognition.
+More traditional techniques such as a wavelet transform~\citep{nason1995stationary} could be used to extract the signal from noise, and compared with the above methods. 
+A wavelet transform provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. 
+In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, wavelet transform methods are proposed for speech recognition.
 
@@
 One approach could be to modify the input audio signal to simulate Doppler modulation. 
+% did we cut the turn-table idea? that was fun! - james
 

### 2021-08-03 paper/ifo-singleTone.tex
proof reading, minor typo fixes and edit
proof-read after integration of 2nd round of referee comments

@@
 	\caption{\label{fig:webcam_spectrum}
-Recovery of a tone at a constant frequency: Here, we plot the Fourier amplitude of the intensity pattern against frequency.
+Recovery of a tone at a constant frequency: the Fourier amplitude of the intensity pattern against frequency.
 The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has an FWHM of $0.033\,{\rm Hz}$.
@@
 
-% rearranged as I found this section to not switch between method and results too much, I've tried this order now: intro, webcam and data specifics, injection, and results. 
-Continuous-wave searches look for nearly monochromatic signals~\cite{JKS:1998}.
-In this section, we consider a simple sinusoidal tone at a single, constant frequency.
-%should be a single constant frequency *signal*?
+Continuous-wave searches look for nearly monochromatic signals~\cite{JKS:1998}. In this section, we consider a simple sinusoidal tone at a single, constant frequency.
 
@@
 
-%The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam (see Fig.~\ref{fig:ifo_schematic_webcam}).
-The webcam samples at a rate of $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
-A tone with a frequency of $2.09\,{\rm Hz$} is played through the speaker for one minute and the interference pattern is recorded by the webcam.
-
 A tone with a frequency of $2.09\,{\rm Hz}$ is played through the speaker for one minute. 
 The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}.
-The Discrete Fourier transform was calculated using the NumPy package for Python (see Appendix~\ref{app:code}).
-% moved refs~\cite{numpy}~\footnote{\url{https://numpy.org/doc/stable/reference/routines.fft.html}}.
+The Discrete Fourier transform is calculated using the NumPy package for Python (see Appendix~\ref{app:code}).
 We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$.

### 2021-08-03 paper/ifo-wanderingTone.tex
proof-read after integration of 2nd round of referee comments

@@
 
-A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion flows from a stellar companion for neutron stars in binaries as in LMXBs.
+A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries as in LMXBs.
 Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. 
@@
 The methods used here are inspired by LIGO and Virgo continuous-wave searches. 
-For further details on these methods and continuous-wave searches generally we refer the reader to references in the Supplementary Material.
+For further details on these methods and continuous-wave searches, we refer the reader to the references in the Supplementary Material.
 
@@
 
-The timeseries data from the interferometer is first split into segments. 
-We then take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude $F(t_i,f_j)$ at frequency $f_j$ and time $t_i$, as the detection statistic (this is the emission probability).
-Assuming Gaussian noise, this detection statistic choice maximizes the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.
+First, we split the timeseries data from the interferometer into segments. 
+Then, we take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude $F(t_i,f_j)$ at frequency $f_j$ and time $t_i$, as the detection statistic (this is the emission probability).
+Assuming Gaussian noise, this detection statistic maximizes the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.
 The detection statistic is normalized for convenience by dividing each value by the maximum Fourier amplitude in the grid (such that $\max_{i,j} F(t_i,f_j) = 1$).
@@
 The circlular nodes represent the elements of the time and frequency grid, which are labelled as  
-$t_i$ and $f_j$, respectively where $i=0,1,2,...N_t$ and $j=0,1,2,...,N_f$. 
+$t_i$ and $f_j$, respectively where $i=0,1,2,\mathellipsis,N_t$ and $j=0,1,2,\mathellipsis,N_f$. 
 The shading of the nodes represents the detection statistic $F(t_i,f_j)$ (the observable) at each grid point, where dark corresponds to lower values and light to higher values.
@@
 The transition probability matrix, $A(f_k,f_m)$, describes the probability of the system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
-Here, we allow the frequency of the signal to either (a) stay in the same bin, (b) move up by a single frequency bin, or (c) move down by a single frequency bin at each transition. 
+Here, we allow the frequency of the signal to either (i) stay in the same bin, (ii) move up by a single frequency bin, or (iii) move down by a single frequency bin at each transition. 
 We assign these three transitions equal probability, i.e.\ $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
@@
 
-Before we begin analyzing the data, we have no prior knowledge as to which frequency bin the signal starts from (i.e. the prior is flat between the minimum and maximum frequency bins).
+Before we begin analyzing the data, we have no prior knowledge as to which frequency bin the signal starts in (i.e. the prior is flat between the minimum and maximum frequency bins).
 
 At the first stage of the analysis, we define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
-A specific path (which may not be optimal) is written as $f(t_0), f(t_1),..., f(t_n)$. 
+A specific path (which may not be optimal) is written as $f(t_0), f(t_1),\mathellipsis, f(t_n)$. 
 The probability of a specific path given the data is 
 \begin{eqnarray}
-{\rm Pr}[ f(t_0), f(t_1),... && f(t_n) | {\rm data}] = \nonumber\\
-          && F[t_n,f(t_n)] A[f(t_n),f(t_n-1)] \nonumber \\
-          && \times... A[f(t_1),f(t_0)] F[t_0,f(t_0)] \,,
+{\rm Pr}[ f(t_0), f(t_1),&&\mathellipsis, f(t_n) | {\rm data}] \nonumber\\
+          && =F[t_n,f(t_n)] A[f(t_n),f(t_n-1)] \nonumber \\
+          && \times\mathellipsis\times A[f(t_1),f(t_0)] F[t_0,f(t_0)] \,,
 \label{eqn:pathprobability}
 \end{eqnarray}
-The path $f^\ast(t_0),...,f^\ast(t_n)$ that maximises ${\rm Pr}[ f(t_0), f(t_1),... f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
+The path $f^\ast(t_0),\mathellipsis,f^\ast(t_n)$ that maximises ${\rm Pr}[ f(t_0), f(t_1),\mathellipsis, f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
 We note that the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
@@
 At every $t_i$ all but $N_f$ possible paths are eliminated. 
-Here we describe the algorithm in pseudocode while referring to the schematic in Fig.~\ref{fig:viterbi}.  
-The implementation used in this work is available online (see Appendix~\ref{app:code}).
+Here we describe the algorithm while referring to the schematic in Fig.~\ref{fig:viterbi}.  
+The implementation used in this work is available online (see Appendix~\ref{app:code}) and we provide further information (including pseudocode) in Appendix~\ref{app:viterbi}.
 \begin{enumerate}
@@
 This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum}.
-However, White noise is a reasonable approximation for the narrow frequency range (around $3$--$11\,{\rm Hz}$) we use in this analysis.
+However, white noise is a reasonable approximation for the narrow frequency range (around $3$--$11\,{\rm Hz}$) we use in this analysis.
 
@@
 The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.
-The recovered Viterbi path is within one frequency bin ($\approx 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time.
+The recovered Viterbi path is within one frequency bin ($\sim 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time.
 We also compute the root-mean-square (RMS) fractional error $E_{\rm rms}$ along the path, which is defined as
@@
 \end{equation}
-where $I_i$ and $R_i$ are the injected and recovered (Viterbi) frequency paths, respectively. 
+where $I_i$ and $R_i$ are the injected and recovered (Viterbi) frequencies at time $t_i$, respectively. 
 For the result shown in Fig.~\ref{fig:viterbi_overlay}, we find that $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.

### 2021-08-03 paper/paper_main.tex
proof-read after integration of 2nd round of referee comments

@@
 
-\title{Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
+\title{Continuous gravitational waves in the lab: recovering audio signals with a table-top~optical microphone}
 
@@
 Here, we present highlights of our results using a combined notch plus Wiener filter and the statistical log minimum mean-square error (logMMSE) estimator. 
-Recordings of simple chords and drums are easily recovered using these techniques, but complex music and speech are more challenging. 
+% Recordings of simple chords and drums are easily recovered using these techniques, but complex music and speech are more challenging. 
+Using these techniques, we easily recover recordings of simple chords and drums, but complex music and speech are more challenging.
 This demonstration can be used by educators in undergraduate laboratories and can be adapted for communicating gravitational-wave and signal-processing topics to non-specialist audiences.  

### 2021-08-03 paper/responseRound2/AJPResponseLetter.tex
small edits
proof-read after integration of 2nd round of referee comments
picking up two comments that were missed

@@
 
-Dear Joseph,
+Dear Joseph Amato,
 
@@
 \begin{itemize}
-\item p.4 (Figure 4 comments):  I think it is obvious that a single FT of a frequency-varying signal cannot capture the signal, so I don't see the need for Figure 4. 
+\item p.3, left column: Is this 1/f noise attributable to ...?
+\begin{itemize}
+\item The origin of this noise is uncertain. We have extended this sentence to clarify. 
+\end{itemize}
+
+\item p.4, (Figure 4 comments):  I think it is obvious that a single FT of a frequency-varying signal cannot capture the signal, so I don't see the need for Figure 4. 
 \begin{itemize}
@@
 
-\item p.4 Section IV A: I appreciate your desire to be all-inclusive, but is the highlighted text necessary? It is not directly related to the topic of your paper. 
+\item p.4, Section IV A: I appreciate your desire to be all-inclusive, but is the highlighted text necessary? It is not directly related to the topic of your paper. 
 \begin{itemize}
@@
 
-\item p.5 Figure 5: This won't show up in B/W. How about indicating the probability by different sizes of the circles? 
+\item p.5, Figure 5: This won't show up in B/W. How about indicating the probability by different sizes of the circles? 
 \begin{itemize}

### 2021-08-03 paper/ifo-singleTone.tex
picking up two comments that were missed

@@
 	\caption{\label{fig:webcam_spectrum}
-Recovery of a tone at a constant frequency: the Fourier amplitude of the intensity pattern against frequency.
+Recovery of a tone at a constant frequency: Here, we plot the Fourier amplitude of the intensity pattern against frequency.
 The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has an FWHM of $0.033\,{\rm Hz}$.
@@
 They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo}.
-Also, note that the noise appears to rise at lower frequencies.
+Also, note that the noise appears to rise at lower frequencies, however the origin of this noise is not known.
 

### 2021-08-03 paper/ifo-appendix.tex

@@
 Classic sig proc result useful for teachers to know.'']}
-In this appendix, we demonstrate the modulus of the Fourier transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
+In this appendix, we demonstrate that the modulus of the Fourier transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
 We describe the data as
@@
 where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data. 
-We replace $n$ by $x-s$ (Eqn.~\ref{eqn:GNdata}), so that 
+% We replace $n$ by $x-s$ (Eqn.~\ref{eqn:GNdata}), so that 
+Using Eqns.~\ref{eqn:GNdata} and~\ref{eqn:GNmodel}, we find that
 \begin{eqnarray}
@@
 %\subsection{Pseudocode}
-Here, we present some pseudocode of the implementation used in Section~\ref{sec:viterbi}.
+Here, we present some pseudocode (below) of the implementation used in Section~\ref{sec:viterbi}.
 We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights.
@@
 We restrict paths to only move one cell up or down at a time (or stay constant).
-For the edge cases of $j=0,N_f$, we only search over $k \in \{0,1\}$ and $k \in \{-1,0\}$, respectively, to stay inside the grid (this is not shown in the pseudocode below).
+For the boundary cases of $j=0,N_f$, we only search over $k \in \{0,1\}$ and $k \in \{-1,0\}$, respectively, to stay inside the grid (this is not shown in the pseudocode below).
 

### 2021-08-03 paper/ifo-introduction.tex
working on the response and minor changes to text

@@
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
-Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
+%Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
 Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
@@
 As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement.
+% I don't think it is novel to use an optical device as a microphone, should we make this clearer? - james
 %% I don't like the use of ``freedom to explore''. 
@@
 % removed as there's not space to really say what this means here: to varying degrees of success.
-This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory (described in the accompanying Supplementary Material).
+This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory (described in the Supplementary Material).
+% deleted ``accompaning'', the material is already supplementary - james
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
-Further reading and resources are provided in the accompanying Supplementary Material and we present the software and scripts used to produce this work in Appendix~\ref{app:code}.
+Further reading and resources are provided in the Supplementary Material and we present the software and scripts used to produce this work in Appendix~\ref{app:code}.
 

### 2021-08-03 paper/ifo-supplementary-material.tex

@@
 
-\title{Supplementary Material for \\Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone}
+\title{Supplementary Material for \texorpdfstring{\\}{}Continuous gravitational waves in the lab: recovering audio signals with a table-top~optical microphone}
 
 \input{authorList.tex}
+
+\date{\today}
 
@@
 For more information on the Viterbi algorithm, the interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001} to learn more.
-Alternative pseudocode examples (to that described the appendix) can be found online at Ref.~\cite{viterbipseudocode:online}.
+Alternative pseudocode examples (to that described in Appendix~C of the main paper) can be found online at Ref.~\cite{viterbipseudocode:online}.
 
@@
 In gravitational-wave observatories, the optical components are suspended from a series of pendulums to achieve passive isolation, and feedback control loops provide active isolation\cite{MatichardEtAl:2015, AdvancedVirgo:2015}.
-Such a system is beyond the scope of this article, however feedback control has been demonstrated in table-top Michelson interferometers designed for the undergraduate laboratory~\cite{LibbrechtBlack:2015}.
+Such a system is beyond the scope of this article, however, feedback control has been demonstrated in table-top Michelson interferometers designed for the undergraduate laboratory~\cite{LibbrechtBlack:2015}.
 Other sources of noise include acoustic sound waves (for example from nearby vehicles, wind, and components such as electric fans) and magnetic noise (for example from high-voltage power lines, building heaters and lights, or any electrical power circuit)~\cite{EfflerEtAl:2015,CironeEtAl:2019}.
@@
 
-This work is implemented in Python Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}. 
+This work is implemented in Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}. 
 We make use of Python packages: numpy~\cite{numpy}\footnote{\url{https://numpy.org/doc/stable/reference/routines.fft.html}}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
@@
 The photodiode circuit diagram is shown in Figure~\ref{fig:circuit_diagram}
-The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an analogy-to-digital converter ADC to a Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
+The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an analogy-to-digital converter ADC to a Raspberry~Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
 Useful throughout was the Raspberry Pi documentation~\cite{pidocumentation:online} and the pinoutXYZ reference~\cite{pipinout:online}.
@@
     % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
-	\includegraphics[width=.5\textwidth]
+	\includegraphics[width=.48\textwidth]
 % trim={1cm 3cm 1cm 1cm}]
@@
 The left and right columns show timeseries and Fourier spectrum results, respectively. 
-The first and second rows in Fig.~\ref{fig:op_mic_results} show the original source signal and the raw optical microphone recording, respectively. 
+The first and second rows in Fig.~\ref{fig:op_mic_results} show the original, source signal and the raw optical microphone recording, respectively. 
 The following sections describe a sequence of analysis techniques and a selection of their results is presented in Fig.~\ref{fig:op_mic_results}.
@@
 \end{equation}
-If we assume that: i) $\textbf{x}$ is ``wide sense stationary''~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in \mathbb{R}$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; ii) the signal $\textbf{s}$ has a mean of zero; and iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
+If we assume that: (i) $\textbf{x}$ is ``wide sense stationary''~\footnote{ A random process $\{x(t)\}$ is wide sense stationary if, for all $t_1,t_2 \in \mathbb{R}$, (1) its mean is time invariant, i.e., $\mu_x(t_1)=\mu_x(t_2)=\text{constant}$; and (2) the autocorrelation depends only on the time difference, i.e., $R_x(t_1,t_2)=R_x(\tau),\tau=t_1-t_2$.} with zero mean; (ii) the signal $\textbf{s}$ has a mean of zero; and (iii) the noise $\textbf{w}$ is uncorrelated with the signal $\textbf{s}$, we can further express $\hat{\textbf{s}}$ to be a linear combination of present and past observed data
 \begin{equation}
@@
 where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data, the injected signal, and the background noise, respectively.
-Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal to noise ratio (SNR) is high and attenuates the signal where the SNR is low.
+Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal-to-noise ratio (SNR) is high and attenuates the signal where the SNR is low.
 The causal Wiener filter (which only makes estimates based on past observations) is similar.

### 2021-08-03 paper/ifo-tableTopGWs.tex

@@
 Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations.
-Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Ref.~\cite{TTExhibit:2021} as well as the further reading section in Supplementary Material for a selection of other table-top Michelson interferometer designs used for science communication).
+Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Ref.~\cite{TTExhibit:2021} as well as the further reading section in the Supplementary Material for a selection of other table-top Michelson interferometer designs used for science communication).
 Sound vibrations provide a simple means to move the components of a demonstration interferometer, changing the length of the interferometer arms, and therefore the interference pattern.
-We use audio signals throughout this work to simulate gravitational-wave-like signals.
+We use audio signals throughout this work to simulate gravitational wave--like signals.
 
@@
 The output of a Class 2 $532\,{\rm nm}$ laser (top-left) is reflected from a mirror at a right angle.
-The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter which is approximately $50$/$50$ in reflection and transmission.
+The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter which reflects approximately $50\%$ of the beam and transmits the remaining $\sim50\%$.
 The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and recombine at the beamsplitter. 
@@
 The optical configuration of the Michelson interferometer used in this work is shown in Fig.~\ref{fig:ifo_schematic_webcam}.
-It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green laser (top-left) with peak emission at a wavelength of $532\,{\rm nm}$.
+It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green laser with peak emission at a wavelength of $532\,{\rm nm}$.
 The output of the laser is first reflected by a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
@@
 A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations. 
-This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a personal computer (PC). 
+This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a computer (PC). 
 The other speaker in the pair is kept face-down and as far away from the interferometer as possible to prevent it from acting as a second source.
-The beams are reflected by mirrors M1 and M2, located at the end of $\approx 7.5\,{\rm cm}$ and $\approx 10.0\,{\rm cm}$ long arms.
+The beams are reflected by mirrors M1 and M2, located at the end of $\sim 7.5\,{\rm cm}$ and $\sim 10.0\,{\rm cm}$ long arms.
 The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging (bi-concave) lens of focal length $-25.0\,{\rm mm}$, and projected onto a screen.
 The interference pattern, as shown by the photograph in Fig.~\ref{fig:interference_pattern}, is a set of concentric light and dark rings (fringes). 
-A change in the relative length of the arms causes these rings to move radially in or outwards.
+A change in the relative length of the arms causes these rings to move radially inwards or outwards.
 The intensity timeseries is recorded by either a webcam (in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}) or a photodiode, which offers a higher sampling rate suitable for capturing more complex audio (in Section~\ref{sec:optical_microphone}).

### 2021-08-03 paper/responseRound2/AJPResponseLetter.tex
adding final missing comment

@@
 
+\noindent\rule{\textwidth}{0.5pt}
+
+Response to comments:
 \begin{itemize}
@@
 
+\item pg. 5, right column, para 2: In a working GW detector, are there other constraints (hinted at earlier in this paragraph)?
+\begin{itemize}
+\item Yes, in the GW analysis, the detection statistic also needs to take into account the Doppler modulation from the Earth's orbit around the Sun as well as the Doppler motion of the source if it is in a binary.
+Section IV.A.
+includes some discussion on this topic.
+\end{itemize}
+
 \item p.5, right column, para 2: The first two sentences seem to contradict each other. Can you clarify? 
@@
 \item p.5, right column, para 1: What does ``marginalization'' mean? Do you mean no hard barriers?
+\begin{itemize}
 \item This statement was added in response to referee 3, who asked about marginalisation and spotted an error in this equation in the initial version of the manuscript.
@@
 We have removed the first part of this sentence, keeping only the second part.
+\end{itemize}
 

### 2021-08-27 paper/ifo-supplementary-material.tex
comments from Andrew on supplementary material

@@
 \subsection{Gravitational-wave observatories}
-The current network of ground-based gravitational-wave observatories includes the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015} which includes two United States sites and third LIGO India site under construction in India~\cite{LIGOIndia:2021}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013}.
+The current network of ground-based gravitational-wave observatories includes the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO~\cite{AdvancedLIGO:2015} which includes LIGO-Hanford and LIGO-Livingston in the United States and LIGO India which is under construction in India~\cite{LIGOIndia:2021}), Advanced Virgo~\cite{AdvancedVirgo:2015}, GEO600~\cite{GEO600:2010}, and KAGRA~\cite{KAGRA:2013}.
 
@@
 Gravitational waves have been observed from compact binary coalescences~\cite{GW150914}. 
-For more information on observations of binary black hole mergers see Refs~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020,GWTC-1:2018}, for binary neutron star mergers see Refs~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020} and for neutron star-black hole mergers~\cite{NSBH:2021,GW190814}.
+For more information on observations of binary black hole mergers see Refs~\cite{GW150914,GW151226,GW170104,GW170814,GW190521,GWTC-1:2018,GWTC-2:2020,GWTC-1:2018}, for binary neutron star mergers see Refs~\cite{GW170817,GW170817multi,GW190425,GWTC-1:2018,GWTC-2:2020} and for neutron star-black hole mergers see Refs.~\cite{NSBH:2021,GW190814}.
 More information about gravitational waves can be found in Refs.~\cite{Maggiore:2007,CBCs:2014,ObsScenarioDoc:2020,Einstein:1916}.
@@
 \label{sec:GWSsearches}
-Continuous waves are anticipated from rotating neutron stars. 
+The network of ground-based gravitational-wave observatories is also searching for continuous gravitational waves. 
+Rotating neutron stars are prime targets for continuous wave searches. 
 For further information about continuous waves from isolated neutron stars see for example Refs.~\cite{MelatosDouglassSimula:2015,Jones:2010} and from neutron stars in binaries see for example Ref.~\cite{BildstenTB:1998}.
@@
 The methods used in Section~IV are inspired by hidden Markov model (HMM) searches for continuous gravitational waves. 
-The HMM method is described in Refs~\cite{SuvorovaEtAl:2017,SuvorovaEtAl:2016,BayleyEtAlSOAP:2019,Viterbi:1967} and further information in Ref.~\cite{JKS:1998}. 
-For more information on the Viterbi algorithm, the interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001} to learn more.
+The HMM method is described in Refs~\cite{SuvorovaEtAl:2017,SuvorovaEtAl:2016,BayleyEtAlSOAP:2019,Viterbi:1967} and further information can be found in Ref.~\cite{JKS:1998}. 
+For more information on the Viterbi algorithm, the interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001}. 
 Alternative pseudocode examples (to that described in Appendix~C of the main paper) can be found online at Ref.~\cite{viterbipseudocode:online}.
@@
 In gravitational-wave detectors reducing and mitigating environmental noise is an active area of research~\cite{EfflerEtAl:2015}. 
-One such noise source is due to seismic ground motion from the Earth. 
+One such noise source is seismic ground motion from the Earth. 
 In gravitational-wave observatories, the optical components are suspended from a series of pendulums to achieve passive isolation, and feedback control loops provide active isolation\cite{MatichardEtAl:2015, AdvancedVirgo:2015}.

### 2021-08-31 paper/ifo-supplementary-material.tex
adding teaching references and paragrph in sup mat

@@
 For more information about the techniques used here see Refs.~\cite{Mitra:2011, Lyons:2011, Stein:2000, OpenheimSchaferBuck:1999, PrakisManolakis:1996,10.5555/151045, DigitalProcingOfSpeechSignals:1978}.
+
+The experimental apparatus presented here provides an opportunity to bring together physics and electrical engineering topics in the lab environment. 
+Several examples of cross-disciplinary teaching can be found in the literature, for example see Refs.~\cite{OConnerSibrayForinash:2001, PiersonGurlandCrawford:2002, CafarelliEtAl:2012, MakanKopaszGingl:2014}.
+Cross-disciplinary experimental projects between physics and engineering students can increase student understanding of overlapping concepts through contextual learning and also improve student motivation~\cite{PiersonGurlandCrawford:2002}.
+Realistic and open-ended projects spanning both topics allow students a better understanding of real-world application of these topics and of the dependency between scientists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
+The optical microphone presented here can similarly be used as a teaching tool for students to conduct investigations into signal processing techniques with a physical system.
 

### 2021-08-31 paper/responseRound2/AJPResponseLetter.tex
response comments from Andrew

@@
 In this case they may not be selected by the Viterbi algorithm if the overall value of equation 5 is higher for the signal path than the paths containing the noise.
+Even in the low signal-to-noise ratio regime, where the algorithm might be ``sidetracked'' by a loud burst of noise in a frequency-time bin, Viterbi still finds the optimal path.
 \end{itemize}
@@
 We have added some details to clarify this in the text.
+In principle in a real continuous wave search you can work out thresholds for a detection in colored noise, and sometimes we do that for LIGO searches (e.g.
+Middleton et al., PRD, 2020), but it is too advanced for the demonstration project in this paper.
 \end{itemize}
@@
 \begin{itemize}
-\item We can only estimate S/N for single time blocks of data, however we do not believe that a representative S/N can be estimated for the entire data set.
-We have clarified this in the text were we refer to the signal to noise ratio of real continuous-wave signals in Section IV.C.
+\item 
+We do not believe that a representative signal-to-noise ratio can be computed for the Viterbi algorithm path. 
+The signal-to-noise ratio could be computed for a single block of time or for the whole data set if we used a coherent matched filter over the whole data (e.g.
+an FFT), but it doesn't translate simply into a signal-to-noise ratio relevant to the Viterbi algorithm.
+We have clarified this in the text were we refer to the signal-to-noise ratio of real continuous-wave signals (in single time steps) in Section IV.C. 
 \end{itemize}

### 2021-08-31 paper/ifo-supplementary-material.tex
rephrasing

@@
 Cross-disciplinary experimental projects between physics and engineering students can increase student understanding of overlapping concepts through contextual learning and also improve student motivation~\cite{PiersonGurlandCrawford:2002}.
-Realistic and open-ended projects spanning both topics allow students a better understanding of real-world application of these topics and of the dependency between scientists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
+Realistic and open-ended projects spanning both topics allow students a better understanding of real-world application of these topics and of the dependency between physicists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
 The optical microphone presented here can similarly be used as a teaching tool for students to conduct investigations into signal processing techniques with a physical system.

### 2021-08-31 paper/ifo-complexAudio.tex
adding comments from Andrew

@@
 
-This analysis is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals. % citation?
-In addition, the apparatus serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories. 
+%This analysis is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals. % citation?
+The apparatus serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories. 
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 

### 2021-08-31 paper/ifo-conclusions.tex
adding comments from Andrew

@@
 % justify ``un-modeled''? - james
-The demonstration can be adapted for use in the undergraduate laboratory and as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
+The demonstration can be adapted for use in both the physics and engineering undergraduate laboratory, providing opportunities for cross-disciplinary teaching. 
+Additionally, it can be used as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
 

### 2021-08-31 paper/ifo-wanderingTone.tex
adding comments from Andrew

@@
 
-A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries as in LMXBs.
+A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries as in LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
 Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. 
@@
 In continuous-wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time.
-The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2017}. 
+The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2017} yet is pitched at a level appropriate for an undergraduate laboratory setting.
 In Section~\ref{sec:realCWSearches}, we briefly review the analysis method used by LIGO and Virgo; in Section~\ref{sec:viterbi}, we describe the methods as applied in this work; and in Section~\ref{sec:wanderingResults}, we show results for recovering a wandering frequency signal using the table-top interferometer.
@@
 {\rm Pr}[ f(t_0), f(t_1),&&\mathellipsis, f(t_n) | {\rm data}] \nonumber\\
-          && =F[t_n,f(t_n)] A[f(t_n),f(t_n-1)] \nonumber \\
+          && =F[t_n,f(t_n)] A[f(t_n),f(t_{n-1})] \nonumber \\
           && \times\mathellipsis\times A[f(t_1),f(t_0)] F[t_0,f(t_0)] \,,

### 2021-08-31 paper/ifo-appendix.tex
definition correction in app b

@@
 \end{equation}
-where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data. 
+where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \frac{1}{T} \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data. 
 % We replace $n$ by $x-s$ (Eqn.~\ref{eqn:GNdata}), so that 

### 2021-08-31 paper/ifo-appendix.tex
adding JKS reference

--- 2021-08-03 paper/ifo-appendix.tex
@@
 \end{equation}
-where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data. 
+where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \frac{1}{T} \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data~\cite{JKS:1998}. 
 % We replace $n$ by $x-s$ (Eqn.~\ref{eqn:GNdata}), so that 

### 2021-08-31 paper/ifo-futureWork.tex
adding comments from Andrew
comments and rephrased a sentence

@@
 Flowchart for the signal through the optical microphone. 
-A transfer function of the system would have to account for each stage and is an aim of future work. 
+A transfer function of the system would have to account for each stage and exploration of this is an aim of future work. 
 Not shown are the signal processing filters that are performed on the PC once the data is transferred off the Raspberry~Pi.
@@
 The experiment described here can be used to demonstrate and teach a variety of topics, from basic optics and photodiode circuits, to signal processing and speech enhancement. 
+Cross-disciplinary undergraduate laboratory experiment between physics and engineering courses have been shown to be beneficial to student learning (see Section II in the Supplementary Material for further information).
 Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.

### 2021-08-31 paper/responseRound2/AJPResponseLetter.tex
typo fix

@@
 \item We have clarified this in the text.
-The gain in having data from two interferometer is that directional information can be estimated based on the time delay between the signal arriving at each interferometer, in a similar way to how we are able to tel the direction of a sound with our ears.
+The gain in having data from two interferometer is that directional information can be estimated based on the time delay between the signal arriving at each interferometer, in a similar way to how we are able to tell the direction of a sound with our ears.
 This is also how gravitational-wave detectors estimate the direction of compact binary mergers based on the time-delay of the signal reaching two or more detectors.

### 2021-09-23 paper/ifo-introduction.tex
moving citations to after punctuation at editor request

@@
 
-In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the Universe~\cite{GW150914}.
+In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the Universe.\cite{GW150914}
 Gravitational waves are a prediction of Albert Einstein's theory of General Relativity; they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
@@
 These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
-Table-top interferometers are commonly used in undergraduate laboratory experiments and to demonstrate the science of gravitational-wave detection to non-specialist audiences~\cite{TTExhibit:2021}.
+Table-top interferometers are commonly used in undergraduate laboratory experiments and to demonstrate the science of gravitational-wave detection to non-specialist audiences.\cite{TTExhibit:2021}
 
 % CW 
-To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and neutron star-black hole mergers~\cite{GWTC-2:2020,NSBH:2021}.
+To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and neutron star-black hole mergers.\cite{GWTC-2:2020,NSBH:2021}
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
-%Continuous waves may be emitted at a constant frequency or may wander slowly in frequency over time. 
 Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
-%where a compact object such as a black hole or a neutron star is in orbit with a low mass stellar companion.
-% we're talking about neutron stars in LMXB's, so we don't need to mention black holes in LMXB's.
-The rotation frequency of the neutron star in an LMXB can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion~\cite{xraybinaries:1997}.
+The rotation frequency of the neutron star in an LMXB can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion.\cite{xraybinaries:1997}
 Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
-%Scorpius X-1 is an LMXB that is particularly bright in X-rays, making it a prime target for continuous wave searches. %from LMXBs. 
-Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~(e.g., Ref~\cite{ScoX1O2Viterbi:2019}).
+Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~(e.g., Ref.~\cite{ScoX1O2Viterbi:2019}).
 
-% changed present to use, as this is not the first time a Michelson has been presented as a tpy GW detector - hannah
 In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
 We then extend its use to an ``optical microphone'', using light to capture sound, and present a range of example analysis techniques for educators to use. 
-% re-added "as an undergraduate experiment to be clear we are not talking about school teaching - hannah
 As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement.
-% I don't think it is novel to use an optical device as a microphone, should we make this clearer? - james
-%% I don't like the use of ``freedom to explore''. 
-% removed Additionally - I don't think it is neccessary - hannah
-It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data~
-\cite{SuvorovaEtAl:2017, ScoX1O2Viterbi:2019}. 
-% This demonstration can be adapted to an effective outreach tool to explain gravitational wave science to a non-specialist audience.
-% Rephrased as this kind of sounds like we are an exhibition too [hannah] - This demonstration also has the potential to be used as an outreach tool alongside other exhibitions
-%This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities including; hands-on demonstrations and exhibitions~\cite{TTExhibit:2020,CavagliaExhibit:2009,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, phone applications~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online} which have been developed by gravitational-wave research groups around the world.
+It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data.\cite{SuvorovaEtAl:2017, ScoX1O2Viterbi:2019}
 This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world.
 These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.
-%\han{look up Sarah Farmer's project}
 
@@
 In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
-% removed as there's not space to really say what this means here: to varying degrees of success.
 This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory (described in the Supplementary Material).
-% deleted ``accompaning'', the material is already supplementary - james
 We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 

### 2021-09-23 paper/ifo-complexAudio.tex
moving citations to after puctuation

@@
 Optical microphones have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles. 
-%We use the term ``optical microphone'' to distinguish this technology from those previous technologies.
 Our objective is to play a recording of speech or music through the speaker attached to mirror M2 (see Fig.~\ref{fig:ifo_schematic_webcam}), record the resulting interference pattern, and then recover the original signal via a selection of signal processing techniques.
 
-%This analysis is indirectly related to the advanced signal processing techniques required to extract gravitational-wave signals. % citation?
 The apparatus serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories. 
@@
 We consider a selection of filter techniques, details of which along with a summary of digital signal processing resources can be found in the Supplementary Material. 
- % We also refer the reader to a selection of digital signal processing resources for further information. 
 In Section~\ref{sec:opticalMicResults}, we present the two best-performing techniques from the Supplementary Material.
@@
 The photodiode records the interference pattern at roughly the same off-center position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily. 
-% I think this is probably not needed as I believe the editor comment is referring to the cloth screen over the photodiode, which is addressed below. 
-%The photodiode is mounted on a piece of cloth stretched over a plastic frame, with the electrical leads connected underneath the cloth.
-%This cloth screen mount was the ``grill cloth'' from the front of a dismantled commercial speaker, re-purposed. 
 The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2, which provides a convenient means to record the photodiode data.
@@
 This cloth screen was re-purposed grill cloth from a commercial speaker.
-% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
-% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).
-
-%An updated schematic and a photograph of the entire optical microphone are shown in the left and right panels of Fig.~\ref{fig:ifo_schematic_podo}, respectively.
-%See Appendix~\ref{app:circuit_diagram} for a diagram of the circuit.
 
@@
 
-%The source audio was played through the speaker with the optical microphone recording for the full duration of the audio (around a minute long for most sources). 
 We test the optical microphone with a variety of recordings, including the speech of different people and music ranging from simple melodies and rhythms to songs. 
@@
 The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python (see Appendix~\ref{app:code}).
-% moved refs \cite{scipy,python}.
 When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.
@@
 Power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
-% Bottom: the PSD after applying a Butterworth bandpass filter (bottom panel) between the two frequencies marked with red, dashed lines. 
 We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling).
 Otherwise, the PSD is fairly white except for a peak at around 0.75~kHz.
-% After filtering we see strong attenuation (at least 3~dB) of all frequencies outside the band, but little change to the harmonics within the band.
 }
@@
 
-% https://en.wikipedia.org/wiki/Filter_(signal_processing)
 In signal processing, the ideal filter would be one that:
@@
 However, these three conditions cannot all hold at once. 
-For example, if conditions (i) and (ii) hold, then the filter must be discontinuous at the boundary of the undesired region but this implies that the filter has ``infinite latency'' and so will affect (or damage) the time domain signal for infinite time~\cite{10.5555/151045}.
+For example, if conditions (i) and (ii) hold, then the filter must be discontinuous at the boundary of the undesired region but this implies that the filter has ``infinite latency'' and so will affect (or damage) the time domain signal for infinite time.\cite{10.5555/151045}
 Therefore, any filter must compromise between these three conditions. 
@@
 Therefore, we choose filters that compromise between achieving the three conditions.
-% All three of these cases can, in extreme cases, worsen speech intelligibility beyond the unfiltered recording and so we choose filters that compromise between achieving the three conditions.
 

### 2021-09-23 paper/ifo-conclusions.tex
moving citations to after puctuation

@@
 In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community.
-We explore the use of the interferometer as an optical microphone and consider a more general treatment of signal processing with complex audio signals, which can also serve as a distant analog to un-modeled gravitational-wave burst signals, e.g., from supernovae.
+We explore the use of the interferometer as an optical microphone and consider a more general treatment of signal processing with complex audio signals, which can also serve as a distant analog to minimally gravitational-wave burst signals, e.g., from supernovae.
 % justify ``un-modeled''? - james
+% fair point, it's the lingo used in LIGO, but really it's more accurate to say minimally modelled - Han
 The demonstration can be adapted for use in both the physics and engineering undergraduate laboratory, providing opportunities for cross-disciplinary teaching. 
 Additionally, it can be used as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
-
-\begin{comment}
-
-The demonstration presented has the potential to be adapted for use in the undergraduate laboratory.
-It can be used to teach topics of interest to physics and electrical engineering students including interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement techniques.
-The increased excitement and public interest in the field of gravitational-wave research in recent years mean that this demonstration may also be adapted as a tool for explaining gravitational-wave research to a wider, non-specialist audience.
-
-\end{comment}
 

### 2021-09-23 paper/ifo-futureWork.tex
moving citations to after puctuation

@@
     \includegraphics[width=.8\textwidth]{figures/pipeline_nobox.pdf}
-	%\includegraphics[height=.8\textwidth, angle=-90]{figures/pipeline.pdf}
 	\caption{\label{fig:pipeline_highlighted}
@@
 The Raspberry Pi could be powered with a commercial battery instead of mains power, which may help to isolate the circuit from the mains as all other components draw power from the Raspberry~Pi (the laser is DC-powered through a converter).
-Transferring the circuit from a breadboard to a printed circuit board can also reduce interference~\cite{elfekey2013design}.
-A feedback loop control system can also be used to suppress any unwanted movement of the interferometer components~\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust}, similar to methods used to isolate gravitational-wave observatories.
+Transferring the circuit from a breadboard to a printed circuit board can also reduce interference.\cite{elfekey2013design}
+A feedback loop control system can also be used to suppress any unwanted movement of the interferometer components,\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust} similar to methods used to isolate gravitational-wave observatories.
 
@@
 
-A voluminous literature exists on using hidden Markov models trained on phonemes to recognize speech~\cite{HMM_english}. 
+A voluminous literature exists on using hidden Markov models trained on phonemes to recognize speech.\cite{HMM_english}
 These could perform the final stage of speech enhancement for the optical microphone. 
 This coincidentally connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type.
-Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator~\cite{SEGAN}.
+Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator.\cite{SEGAN}
 
@@
 
-Returning to the demonstration of gravitational wave analysis, single interferometers cannot detect directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal being recorded at two or more detectors~\cite{GW150914}.
+Returning to the demonstration of gravitational wave analysis, single interferometers cannot detect directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal being recorded at two or more detectors.\cite{GW150914}
 We could therefore extend this analysis to include data from two interferometers to extract directional information.
@@
 % did we cut the turn-table idea? that was fun! - james
+% yeah afraid so - it was nice, but the editor did not like it! :( - Hannah
 

### 2021-09-23 paper/ifo-singleTone.tex
moving citations to after puctuation

--- 2021-08-03 paper/ifo-singleTone.tex
@@
 
-Continuous-wave searches look for nearly monochromatic signals~\cite{JKS:1998}. In this section, we consider a simple sinusoidal tone at a single, constant frequency.
+Continuous-wave searches look for nearly monochromatic signals.\cite{JKS:1998} 
+In this section, we consider a simple sinusoidal tone at a single, constant frequency.
 

### 2021-09-23 paper/ifo-tableTopGWs.tex
moving citations to after puctuation

@@
 Current generation gravitational-wave observatories have kilometer-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo, respectively). 
-They can measure minuscule changes in distance due to gravitational waves; e.g., the first detection of a binary black hole merger produced a strain of $10^{-21}$~\cite{GW150914}, corresponding to a change in distance equal to a fraction of the width of a proton.
+They can measure minuscule changes in distance due to gravitational waves; e.g., the first detection of a binary black hole merger produced a strain of $10^{-21}$,\cite{GW150914} corresponding to a change in distance equal to a fraction of the width of a proton.
 

### 2021-09-23 paper/ifo-appendix.tex
Andrews rewrite on the app b

@@
 \end{equation}
-where $s(t)$ and $n(t)$ are the signal and noise, respectively and  
+where $s(t)$ and $n(t)$ are the signal and noise, respectively.
+The signal takes the form
 \begin{equation}
@@
 where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal, respectively. 
+The noise $n(t)$ is a fluctuating zero-mean time series with the following property: if we define an inner product between two time series $u(t)$ and $v(t)$ as 
+\begin{equation}
+\ip{u}{v} = \frac{1}{T} \int_0^T \mathrm{d} t \, u(t) v(t)\,,
+\label{eqn:ipuv}
+\end{equation}
+where $T$ is the total time of the observation, then the likelihood $\mathcal{L}$ of measuring the noise-related value $\ip{n}{n}$ is given by 
+\begin{equation}
+\mathcal{L} = \exp\left( -\frac{1}{2}\ip{n}{n}\right)\,.
+\label{eqn:ipnn}
+\end{equation}
+Equations~\ref{eqn:ipuv} and~\ref{eqn:ipnn} define Gaussian noise through the fundamental measurement of $\ip{n}{n}$.
 
+\begin{comment}
 The likelihood is proportional to the probability of observing the data given an assumed model of the signal~\cite{Jaynes:2003}.
@@
 where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \frac{1}{T} \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data~\cite{JKS:1998}. 
+\end{comment}
+
 % We replace $n$ by $x-s$ (Eqn.~\ref{eqn:GNdata}), so that 
-Using Eqns.~\ref{eqn:GNdata} and~\ref{eqn:GNmodel}, we find that
+%Using Eqns.~\ref{eqn:GNdata} and~\ref{eqn:GNmodel}, we find that
+The likelihood of measuring the signal $s(t)$ in the presence of noise follows from Eqns.~\ref{eqn:GNdata} and~\ref{eqn:ipnn} by replacing $n(t)$ in Eqn.~\ref{eqn:ipnn} with $x(t) - s(t) = n(t)$ from Eqn.~\ref{eqn:GNdata}.~\cite{JKS:1998,Jaynes:2003}
+The result is
 \begin{eqnarray}
-\log \mathcal{L} &=& -\frac{1}{2}\ip{x-s}{x-s}\,, \\ 
-                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{2} \ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]} \nonumber\\
-                 &&+ A \ip{x}{\cos[\omega t + \phi]}\,,\\ 
-                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{4} + A \ip{x}{\cos[\omega t + \phi]}\,,\label{eqn:logLike}
+\mathcal{L} &=& \exp\left( -\frac{1}{2}\ip{x-s}{x-s} \right) \label{eqn:likeOne}\,, \\ 
+            &=& \exp\left( -\frac{1}{2}\ip{x}{x} - \frac{A^2}{2} \ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]} \right. \nonumber\\
+                 &&\hspace{3em} \left. +  A \ip{x}{\cos[\omega t + \phi]} \vphantom{\frac12}\right) \,,\\ 
+                 &=& \exp \left( -\frac{1}{2}\ip{x}{x} - \frac{A^2}{4} + A \ip{x}{\cos[\omega t + \phi]} \right)\,.\label{eqn:logLike}
 \end{eqnarray}
+We then maximise Eqn~.\ref{eqn:logLike} with respect to $A$, obtaining 
+\begin{eqnarray}
+\label{eqn:maxLogLike}
+\mathcal{L}_\mathrm{max} = \exp \left( - \frac{1}{2} \ip{x}{x} + \ip{x}{\cos[\omega t + \phi]}^2 \right)\,,
+\end{eqnarray}
+for $A = 2\ip{x}{\cos[\omega t + \phi]}$.
+From the second term of Eqn.~\ref{eqn:maxLogLike}, we see that the maximum likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.
+
+Two important points must be made about the above procedure. 
+(i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eqn.~\ref{eqn:likeOne} by varying the signal.
+For the special case of the Gaussian likelihood (Eqn.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the distance between $x(t)$ and $s(t)$ defined through Eqn.~\ref{eqn:ipuv}.
+In general, however, it is not equivalent to minimizing the distance between $x(t)$ and $s(t)$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
+(ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eqn.~\ref{eqn:maxLogLike} defines the detection statistic. 
+When its value exceeds a threshold (chosen freely by the analyst) at some value of $\omega$, a signal is deemed to have been deteccted at that value of $\omega$.
+Therefore the specific functional form of Eqn.~\ref{eqn:maxLogLike} matters, which is a second reason why one must start from Eqn.~\ref{eqn:ipnn} rather than Eqn.~\ref{eqn:likeOne}, in addition to reason (i).
 
@@
 where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data (and the last line assumes that $T\gg1/\omega$ so that $\ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]}\approx{1/2}$).
-\end{comment}
 
@@
 %baring the constant factor $\ip{x}{x}$.  
+\end{comment}
 

### 2021-09-23 paper/ifo-wanderingTone.tex
altering figure to use sizes

@@
 A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case as the signal is spread across multiple frequency bins. 
-In continuous-wave searches, the wandering frequency problem is solved using the Viterbi algorithm~\cite{Viterbi:1967}, which can track the signal's frequency over time.
+In continuous-wave searches, the wandering frequency problem is solved using the Viterbi algorithm,\cite{Viterbi:1967} which can track the signal's frequency over time.
 The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2017} yet is pitched at a level appropriate for an undergraduate laboratory setting.
@@
 The frequency of the signal can wander significantly over the observation period. 
-In this context, ``significantly'' means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time~\cite{JKS:1998,ScoX1O2Viterbi:2019}.
+In this context, ``significantly'' means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time.\cite{JKS:1998,ScoX1O2Viterbi:2019}
 
-A ``hidden Markov model'' can be used to search for continuous gravitational waves~\cite{SuvorovaEtAl:2017}. 
+A ``hidden Markov model'' can be used to search for continuous gravitational waves.\cite{SuvorovaEtAl:2017} 
 In a Markov process, the current state depends only on the previous state (in this case the ``state'' is the frequency of the signal). 
@@
 This likelihood is also called the emission probability in gravitational-wave literature.  
-In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun~\cite{JKS:1998}.
-When searching for continuous waves from a neutron star in a binary (such as an LMXB), the Doppler modulation of the source also needs to be taken into account and a different detection statistic is used~\cite{SuvorovaEtAl:2017}.
+In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun.\cite{JKS:1998}
+When searching for continuous waves from a neutron star in a binary (such as an LMXB), the Doppler modulation of the source also needs to be taken into account and a different detection statistic is used.\cite{SuvorovaEtAl:2017}
 
@@
 This is called the transition probability matrix. 
-For example, in LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin in each time step~\cite{ScoX1O2Viterbi:2019}.
+For example, in LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin in each time step.\cite{ScoX1O2Viterbi:2019}
 In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to (i) remain in the same frequency bin, or (ii) move down one frequency bin (see the Supplementary Material for LMXB and supernova remnant search references).
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{figures/viterbiDiagramGrey.pdf}
+\begin{center}
+\includegraphics[width=0.48\textwidth]{figures/viterbiDiagramSizes.pdf}
 \caption{\label{fig:viterbi}
@@
 The circular nodes represent elements in a time-frequency grid, labeled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
-The shading of each node represents the likelihood of a signal being present in each time-frequency element.
-The dark to light gradient corresponds to low to high likelihood values in arbitrary units for this schematic. 
+The size of the nodes represents the likelihood of a signal being present in each time-frequency element. 
+The small to large sizing corresponds to low to high likelihood values in arbitary units for this schematic. 
+The color gradient (online only) also represents the likelihood values where the dark blue corresponds to low likelihood and light green corresponds to high likelihood. 
 The objective is to find the most probable path of a signal through the grid from left to right.
@@
 At $t=t_{N_t}$ the algorithm chooses the terminating frequency node which has the highest value given by Eqn.~\ref{eqn:pathprobability}, marked (b). 
-The Viterbi path is the path leading to this node, highlighted in orange, from (b) to the start at (c). 
+The Viterbi path is the path leading to this node, highlighted by the thick-black lines, from (b) to the start at (c). 
 }
+\end{center}
 \end{figure}
@@
 The Viterbi path is found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}). 
-In Fig.~\ref{fig:viterbi}, it is the path ending at (b) that started at (c) highlighted in orange.
+In Fig.~\ref{fig:viterbi}, it is the path ending at (b) that started at (c) highlighted by the thick-black lines.
 \end{enumerate}

### 2021-09-25 paper/ifo-singleTone.tex
editor comments

@@
 The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}.
-The Discrete Fourier transform is calculated using the NumPy package for Python (see Appendix~\ref{app:code}).
+The discrete Fourier transform is calculated using the NumPy package for Python (see Appendix~\ref{app:code}).
 We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$.
@@
 They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo}.
-Also, note that the noise appears to rise at lower frequencies, however the origin of this noise is not known.
+Also, note that the noise appears to rise at lower frequencies; however the origin of this noise is not known.
 

### 2021-09-25 paper/ifo-tableTopGWs.tex
editor comments

@@
 Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations.
-Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Ref.~\cite{TTExhibit:2021} as well as the further reading section in the Supplementary Material for a selection of other table-top Michelson interferometer designs used for science communication).
+Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Ref.~\cite{TTExhibit:2021} as well as the further reading section in the Supplementary Material for a selection of other table-top Michelson interferometer designs used for science outreach).
 Sound vibrations provide a simple means to move the components of a demonstration interferometer, changing the length of the interferometer arms, and therefore the interference pattern.

### 2021-09-25 paper/ifo-wanderingTone.tex
editor comments

@@
 
-A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries as in LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
-Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. 
+A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries such as LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
+Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. \han{note}
 A typical observation involves $1 \times 10^{10}$ wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation. 
@@
 The circlular nodes represent the elements of the time and frequency grid, which are labelled as  
-$t_i$ and $f_j$, respectively where $i=0,1,2,\mathellipsis,N_t$ and $j=0,1,2,\mathellipsis,N_f$. 
+$t_i$ and $f_j$, respectively, where $i=0,1,2,\mathellipsis,N_t$ and $j=0,1,2,\mathellipsis,N_f$. 
 The shading of the nodes represents the detection statistic $F(t_i,f_j)$ (the observable) at each grid point, where dark corresponds to lower values and light to higher values.

### 2021-09-26 paper/ifo-complexAudio.tex
working on editors monor comments and adding reminders for larger rephrasings

@@
 
-We use the Michelson interferometer as an ``optical microphone'', where a laser is used to detect sound instead of the electrical components in a conventional microphone.
-The only change to our apparatus is replacing the webcam with a photodiode to allow us to capture higher frequencies necessary for speech and music (see Section~\ref{sec:photodiode}). 
+\han{We use the Michelson interferometer as an ``optical mocrophone'' to detect sound, replacing the components of a conventional microphone with }
+\st{We use the Michelson interferometer as an ``optical microphone'', where a laser is used to detect sound instead of the electrical components in a conventional microphone.}
+The only change to our apparatus is replacing the webcam with a photodiode to allow us to capture the higher frequencies necessary for speech and music (see Section~\ref{sec:photodiode}). 
 Optical microphones have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles. 
@@
 We include an anti-aliasing Sallen-Key filter tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
+\han{note}
 This component attenuates any frequencies above $8\,{\rm kHz}$ before they are digitally sampled.
@@
 The results of this section are shown in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
-In the figure, the left and right columns show the timeseries and frequency spectrum respectively. 
+In the figure, the left and right columns show the timeseries and frequency spectrum, respectively. 
 The first row shows the input signal played through the speaker (see Fig.~\ref{fig:ifo_schematic_webcam}). 
@@
 
-In this section, we present the results of two advanced signal processing analysis techniques applied to the optical microphone recordings. 
+In this section, we present the results of two advanced signal processing techniques applied to the optical microphone recordings. 
 The techniques are only briefly described here and we refer the reader to the Supplementary Material for further details and other analysis techniques.
@@
 A notch filter removes signals within a specific frequency range. 
-We want to remove the $50\,{\rm Hz}$ mains noise and harmonics, therefore we use a cascaded notch filter where each notch is centered at each of the harmonics. 
+We want to remove the $50\,{\rm Hz}$ mains noise and harmonics, therefore we use a cascaded notch filter where each notch is centered on one of the harmonics. 
 The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. 
@@
 The results of the combined cascaded notch and Wiener filter are shown in the third row in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
-Most of the mains noise is removed, however, the recovered voice sounds muffled and is not understandable.
+Most of the mains noise is removed; however, the recovered voice sounds muffled and is not understandable.
 
@@
 We see significant attenuation of the mains harmonics and general smoothing of the spectrum. 
-Most of the background noise is removed, however, the logMMSE still does not significantly enhance the speech as the voice sounds muffled and indistinct.
+Most of the background noise is removed; however, the logMMSE still does not significantly enhance the speech as the voice sounds muffled and indistinct.
 
 We find some improvement with music over speech. 
-Simple chords and drums can be heard after filtering, however, more composite sounds and complex melodies cannot be heard clearly. 
+Simple chords and drums can be heard after filtering, but more composite sounds and complex melodies cannot be heard clearly. 
 Our observations suggest that this is especially true for certain instruments; in particular flutes and violins sometimes can’t be heard at all. 
 This could be a perceptual effect or a frequency dependence somewhere in the optical microphone.
-Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums are louder in the results.
+Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums sound louder in the results.
 To address these problems, we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all.

### 2021-09-26 paper/ifo-futureWork.tex
working on editors monor comments and adding reminders for larger rephrasings

@@
 The experiment described here can be used to demonstrate and teach a variety of topics, from basic optics and photodiode circuits, to signal processing and speech enhancement. 
-Cross-disciplinary undergraduate laboratory experiment between physics and engineering courses have been shown to be beneficial to student learning (see Section II in the Supplementary Material for further information).
+Cross-disciplinary undergraduate laboratory experiments between physics and engineering courses have been shown to be beneficial to student learning (see Section II in the Supplementary Material for further information).
 Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.
@@
 This coincidentally connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type. 
-Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator.\cite{SEGAN}
+Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator.\cite{SEGAN} \han{note}
 
@@
 
-Returning to the demonstration of gravitational wave analysis, single interferometers cannot detect directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal being recorded at two or more detectors.\cite{GW150914}
+Returning to the demonstration of gravitational wave analysis, single interferometers cannot yield directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal\han{note} being recorded at two or more detectors.\cite{GW150914}
 We could therefore extend this analysis to include data from two interferometers to extract directional information.

### 2021-09-26 paper/ifo-wanderingTone.tex
working on editors monor comments and adding reminders for larger rephrasings

@@
 
-At the first stage of the analysis, we define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic at that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
+\han{arrow?}At the first stage of the analysis, we define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic of that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
 A specific path (which may not be optimal) is written as $f(t_0), f(t_1),\mathellipsis, f(t_n)$. 
@@
           && =F[t_n,f(t_n)] A[f(t_n),f(t_{n-1})] \nonumber \\
-          && \times\mathellipsis\times A[f(t_1),f(t_0)] F[t_0,f(t_0)] \,,
+          && \times\mathellipsis\times A[f(t_1),f(t_0)] F[t_0,f(t_0)] \,.
 \label{eqn:pathprobability}
@@
 In continuous gravitational wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency can change unpredictably over time.
-The Viterbi algorithm's strength lies in its ability to track such signals through the data.
+The Viterbi algorithm's strength lies in its ability to track such signals through the data \han{even in the presence of comparatively loud noise fluctuations in the time-frequency bins}.
 In the following section, we present the results of using the Viterbi algorithm with the table-top interferometer data.
@@
 One may be tempted to increase the allowed frequency wander in the algorithm, however, this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.
-There is also an anomaly at $150\,{\rm s}$, which is likely due to a local disturbance, e.g.
-someone walking past the interferometer.
-As shown in Fig.~\ref{fig:viterbi_overlay}, the Viterbi algorithm can recover and continues to track the signal after the disturbance.
+There is also an anomaly at $150\,{\rm s}$, which is likely due to a local disturbance, e.g. someone walking past the interferometer. 
+As shown in Fig.~\ref{fig:viterbi_overlay}, the Viterbi algorithm can recover and continue to track the signal after the disturbance.
 

### 2021-09-26 paper/ifo-complexAudio.tex
suggested rephrasing on op mic

@@
 
-\han{We use the Michelson interferometer as an ``optical mocrophone'' to detect sound, replacing the components of a conventional microphone with }
-\st{We use the Michelson interferometer as an ``optical microphone'', where a laser is used to detect sound instead of the electrical components in a conventional microphone.}
+We use the Michelson interferometer as an ``optical microphone'' to detect sound, replacing the components of a conventional microphone with a laser interferometer.
 The only change to our apparatus is replacing the webcam with a photodiode to allow us to capture the higher frequencies necessary for speech and music (see Section~\ref{sec:photodiode}). 

### 2021-09-26 paper/ifo-appendix.tex
working on editors monor comments and adding reminders for larger rephrasings
comments from Andrew on appendix

@@
 \label{app:sinusoid_likelihood}
+
 %\jam{[to do!
@@
 where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal, respectively. 
-The noise $n(t)$ is a fluctuating zero-mean time series with the following property: if we define an inner product between two time series $u(t)$ and $v(t)$ as 
+The noise $n(t)$ is a fluctuating zero-mean time series with the following property: if we define an inner product between two arbitrary time series $u(t)$ and $v(t)$ as 
 \begin{equation}
@@
 \end{equation}
-where $T$ is the total time of the observation, then the likelihood $\mathcal{L}$ of measuring the noise-related value $\ip{n}{n}$ is given by 
+where $T$ is the total time of the observation, then the likelihood $\mathcal{L}$ of measuring the noise-noise product $\ip{n}{n}$ is given by 
 \begin{equation}
@@
 \end{equation}
-Equations~\ref{eqn:ipuv} and~\ref{eqn:ipnn} define Gaussian noise through the fundamental measurement of $\ip{n}{n}$.
+Equations~\ref{eqn:ipuv} and~\ref{eqn:ipnn} define what it meand for noise to be Gaussian through the fundamental measurement of $\ip{n}{n}$.
 
@@
 Two important points must be made about the above procedure. 
-(i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eqn.~\ref{eqn:likeOne} by varying the signal.
-For the special case of the Gaussian likelihood (Eqn.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the distance between $x(t)$ and $s(t)$ defined through Eqn.~\ref{eqn:ipuv}.
-In general, however, it is not equivalent to minimizing the distance between $x(t)$ and $s(t)$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
-(ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eqn.~\ref{eqn:maxLogLike} defines the detection statistic. 
+(i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eqn.~\ref{eqn:likeOne} by varying $s(t)$ through $A$.
+For the special case of the Gaussian likelihood (Eqn.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the distance between $x(t)$ and $s(t)$ as evident in Eqn.~\ref{eqn:ipuv}.
+In general, however, minimizing the distance between $x(t)$ and $s(t)$ is not equivalent always to the fundamental goal of maximising $\mathcal{L}$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
+(ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eqn.~\ref{eqn:maxLogLike} (or equivalently its logarithm) defines the detection statistic. 
 When its value exceeds a threshold (chosen freely by the analyst) at some value of $\omega$, a signal is deemed to have been deteccted at that value of $\omega$.
@@
 We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights.
-To avoid numerical underflow we take the logarithm and consider them instead as additive weights.
+To avoid numerical underflow we take the logarithm of the Fourier amplitudes, which we can equivalently use as additive weights.
 

### 2021-09-27 paper/ifo-futureWork.tex
updating revtex and woring on editor comments

@@
 This coincidentally connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type. 
-Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator.\cite{SEGAN} \han{note}
+Alternatively, machine learning solutions exist throughout the field that can provide results competitive with statistical techniques (such as the logMMSE estimator used here).\cite{SEGAN}
+%Alternatively, machine learning solutions exist throughout the field that can compete with statistical techniques such as the logMMSE estimator.\cite{SEGAN} \han{note}
 

### 2021-09-27 paper/ifo-wanderingTone.tex
updating revtex and woring on editor comments

@@
 Before we begin analyzing the data, we have no prior knowledge as to which frequency bin the signal starts in (i.e. the prior is flat between the minimum and maximum frequency bins).
-
-\han{arrow?}At the first stage of the analysis, we define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic of that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
+At the first stage of the analysis, we define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic of that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
 A specific path (which may not be optimal) is written as $f(t_0), f(t_1),\mathellipsis, f(t_n)$. 
@@
 In continuous gravitational wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency can change unpredictably over time.
-The Viterbi algorithm's strength lies in its ability to track such signals through the data \han{even in the presence of comparatively loud noise fluctuations in the time-frequency bins}.
+The Viterbi algorithm's strength lies in its ability to track such signals through the data even in the presence of comparatively loud noise fluctuations in the time-frequency bins.
 In the following section, we present the results of using the Viterbi algorithm with the table-top interferometer data.

### 2021-10-05 paper/ifo-complexAudio.tex
correcting text 16kHz -> 8kHz

@@
 Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing (folding of frequencies greater than half the sampling rate) into the detected range.
-We include an anti-aliasing Sallen-Key filter tuned to $16\,{\rm kHz}$ before the ADC to prevent this from happening.
-\han{note}
+We include an anti-aliasing Sallen-Key filter tuned to $8\,{\rm kHz}$ before the ADC to prevent this from happening.
 This component attenuates any frequencies above $8\,{\rm kHz}$ before they are digitally sampled.

### 2021-10-05 paper/ifo-supplementary-material.tex
updating revtex and woring on editor comments
corrected circuit diagram, cut-off frequency is indeed 8kHz
correcting text 16kHz -> 8kHz

@@
 
-\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-1}
+\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-2}
 \usepackage[utf8]{inputenc}
@@
 \begin{figure*}
-	\includegraphics[width=.8\textwidth]{figures/circuit_diagram_2.pdf}
+	\includegraphics[angle=-90,width=.8\textwidth]{figures/circuit_diagram_2.pdf}
 	\caption{\label{fig:circuit_diagram}
@@
 The photodiode is connected in reverse-bias across an op-amp (pale-red box). 
-The op-amp output is passed through a Sallen-Key anti-aliasing filter (pale-blue box) tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
+The op-amp output is passed through a Sallen-Key anti-aliasing filter (pale-blue box) tuned to $8\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
 The digitized signal is then read by the special-purpose input (SPI) pins of a Raspberry Pi in standard configuration.

### 2021-10-05 paper/ifo-complexAudio.tex
tiny change to avoid ''tuned''

@@
 Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing (folding of frequencies greater than half the sampling rate) into the detected range.
-We include an anti-aliasing Sallen-Key filter tuned to $8\,{\rm kHz}$ before the ADC to prevent this from happening.
+We include an anti-aliasing Sallen-Key filter with a cut-off frequency of $8\,{\rm kHz}$ before the ADC to prevent this from happening.
 This component attenuates any frequencies above $8\,{\rm kHz}$ before they are digitally sampled.
@@
 This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., the output with the speaker switched off), shown in Fig.~\ref{fig:psd_noise}. 
-The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ Australian mains power grid signal up to and beyond $8$th harmonic. 
+The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ Australian mains power grid signal up to and beyond the $8$th harmonic. 
 The mains signal is also present, but far weaker, in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 

### 2021-10-05 paper/ifo-supplementary-material.tex
tiny change to avoid ''tuned''

@@
 The photodiode is connected in reverse-bias across an op-amp (pale-red box). 
-The op-amp output is passed through a Sallen-Key anti-aliasing filter (pale-blue box) tuned to $8\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
+The op-amp output is passed through a Sallen-Key anti-aliasing filter (pale-blue box) with a cut-off frequency of $8\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
 The digitized signal is then read by the special-purpose input (SPI) pins of a Raspberry Pi in standard configuration.
-This diagram is produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online} and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
+This diagram was produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online}.
+% and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
+this url has the incorrect resistors and capacitors - James
 }

### 2021-10-05 paper/ifo-appendix.tex
grammar and spelling

@@
 (ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eqn.~\ref{eqn:maxLogLike} (or equivalently its logarithm) defines the detection statistic. 
-When its value exceeds a threshold (chosen freely by the analyst) at some value of $\omega$, a signal is deemed to have been deteccted at that value of $\omega$.
+When its value exceeds a threshold (chosen freely by the analyst) at some value of $\omega$, a signal is deemed to have been detected at that value of $\omega$.
 Therefore the specific functional form of Eqn.~\ref{eqn:maxLogLike} matters, which is a second reason why one must start from Eqn.~\ref{eqn:ipnn} rather than Eqn.~\ref{eqn:likeOne}, in addition to reason (i).

### 2021-10-05 paper/ifo-conclusions.tex
grammar and spelling

@@
 In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community.
-We explore the use of the interferometer as an optical microphone and consider a more general treatment of signal processing with complex audio signals, which can also serve as a distant analog to minimally gravitational-wave burst signals, e.g., from supernovae.
+We explore the use of the interferometer as an optical microphone and consider a more general treatment of signal processing with complex audio signals, which can also serve as a distant analog to minimally modelled gravitational-wave burst signals, e.g., from supernovae.
 % justify ``un-modeled''? - james

### 2021-10-05 paper/ifo-futureWork.tex
grammar and spelling

@@
 The experiment described here can be used to demonstrate and teach a variety of topics, from basic optics and photodiode circuits, to signal processing and speech enhancement. 
-Cross-disciplinary undergraduate laboratory experiments between physics and engineering courses have been shown to be beneficial to student learning (see Section II in the Supplementary Material for further information).
+Cross-disciplinary undergraduate laboratory experiments between physics and engineering courses are beneficial to student learning (see Section II in the Supplementary Material for further information).
+% ``have been shown to be beneficial'' changed to ``are beneficial'' since we justify it elsewhere - james
 Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.
@@
 
-More traditional techniques such as a wavelet transform~\citep{nason1995stationary} could be used to extract the signal from noise, and compared with the above methods. 
+More traditional techniques such as a wavelet transform~\citep{nason1995stationary} could be used to extract the signal from noise and compared with the above methods. 
 A wavelet transform provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. 
@@
 % yeah afraid so - it was nice, but the editor did not like it! :( - Hannah
+% haha, okay. - james
 

### 2021-10-05 paper/ifo-supplementary-material.tex
grammar and spelling

@@
 The experimental apparatus presented here provides an opportunity to bring together physics and electrical engineering topics in the lab environment. 
-Several examples of cross-disciplinary teaching can be found in the literature, for example see Refs.~\cite{OConnerSibrayForinash:2001, PiersonGurlandCrawford:2002, CafarelliEtAl:2012, MakanKopaszGingl:2014}.
+Several examples of cross-disciplinary teaching can be found in the literature, for example, see Refs.~\cite{OConnerSibrayForinash:2001, PiersonGurlandCrawford:2002, CafarelliEtAl:2012, MakanKopaszGingl:2014}.
 Cross-disciplinary experimental projects between physics and engineering students can increase student understanding of overlapping concepts through contextual learning and also improve student motivation~\cite{PiersonGurlandCrawford:2002}.
-Realistic and open-ended projects spanning both topics allow students a better understanding of real-world application of these topics and of the dependency between physicists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
+Realistic and open-ended projects spanning both topics allow students a better understanding of real-world applications of these topics and of the dependency between physicists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
 The optical microphone presented here can similarly be used as a teaching tool for students to conduct investigations into signal processing techniques with a physical system.

### 2021-10-08 paper/ifo-futureWork.tex
minor edits

@@
 
-Returning to the demonstration of gravitational wave analysis, single interferometers cannot yield directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the signal\han{note} being recorded at two or more detectors.\cite{GW150914}
+Returning to the demonstration of gravitational wave analysis, single interferometers cannot yield directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the observation being recorded at two or more detectors.\cite{GW150914}
 We could therefore extend this analysis to include data from two interferometers to extract directional information.

### 2021-10-08 paper/ifo-wanderingTone.tex
grammar and spelling
minor edits

@@
 A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries such as LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
-Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches are sensitive to very small frequency drifts. \han{note}
+Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches can be impacted by very small frequency drifts. 
 A typical observation involves $1 \times 10^{10}$ wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation. 
@@
 The size of the nodes represents the likelihood of a signal being present in each time-frequency element. 
-The small to large sizing corresponds to low to high likelihood values in arbitary units for this schematic. 
+The small to large sizing corresponds to low to high likelihood values in arbitrary units for this schematic. 
 The color gradient (online only) also represents the likelihood values where the dark blue corresponds to low likelihood and light green corresponds to high likelihood. 
@@
 In continuous gravitational wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency can change unpredictably over time.
+% how to hyphenate ``continuous gravitational wave searches''?
 The Viterbi algorithm's strength lies in its ability to track such signals through the data even in the presence of comparatively loud noise fluctuations in the time-frequency bins.

### 2021-10-14 paper/ifo-appendix.tex
typo fix in app

@@
 \end{equation}
-Equations~\ref{eqn:ipuv} and~\ref{eqn:ipnn} define what it meand for noise to be Gaussian through the fundamental measurement of $\ip{n}{n}$.
+Equations~\ref{eqn:ipuv} and~\ref{eqn:ipnn} define what it means for noise to be Gaussian through the fundamental measurement of $\ip{n}{n}$.
 

### 2021-10-27 paper/ifo-appendix.tex
addressing comments from editors and removing our commented out notes to get ready to share tex files

@@
 We refer the reader to the Supplementary Material for software references. 
-%This project is implemented in Python 3~\cite{python} scripts and jupyter notebooks~\cite{jupyter,ipython} as well as in MATLAB~\cite{MATLAB}.
-We used the Python packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
 The current build and sample data can be found at:
-%\textbf{[Removed GitHub web-link for double-anonymous review.]}
 \url{https://github.com/daccordeon/gravexplain}
-% \begin{verbatim}
-% README
-% \end{verbatim}
-
-\begin{comment}
-\section{Interferometric relation between path length and optical intensity}
-\label{app:intensity_derivation}
-
-Part of the transfer function for the optical microphone relates the change in intensity, $\delta I$, at a point $P$ in the interference pattern, given a small change, $\delta d$, in the difference between the arm lengths of the mirrors.
-Let $\theta$ be the angle between the optical axis and the ray from the center of the relevant mirror to the point $P$. 
-%Let the point be at angle $\theta$ on the screen (in the far-field) above the central ray as measured from the mirrors. 
-%Noting circular symmetry as to not specify exactly where in the fringe the point is.
-At $P$, two rays meet, one from each mirror, with total electric field amplitude $E_1(t) + E_2(t)$ and 
-\begin{align}
-    E_1(t) &= A e^{i (k L - \omega t)},                       \label{eqn:E1} \\
-    E_2(t) &= A e^{i (k (L + 2 d \cos{\theta}) - \omega t)}\,  \label{eqn:E2}, 
-\end{align}
-where $A$ is the amplitude, $L$ is the unperturbed arm length, $k$ is the wavenumber and $\omega$ is the angular frequency of the laser. 
-If the unperturbed difference between the arm lengths is $d$, then the path difference between the two rays is $2 d \cos{\theta}$~\cite{fringes:online}.
-
-The intensity $I(t)$ at $P$ is then the norm squared of the total electric field amplitude,
-\begin{equation}
-I(t) = \left| A e^{i (k L - \omega t)}  + A e^{i (k (L + 2 d \cos{\theta}) - \omega t)} \right|^2\,.
-\label{eqn:intensity1}    
-\end{equation}
-Some elementary trigonometry yields
-\begin{equation}
-%I(t) = 2 A^2 (1 + \cos{\phi})
-I(t) = 2 A^2 \left[ 1 + \cos{\left( 2 k d(t) \cos{\theta} \right) } \right]  \label{eqn:intensity2}\,.
-\end{equation}
-Equation~\ref{eqn:intensity2} shows that $I(t)$ and $d(t)$ are related nonlinearly. 
-Near a bright fringe, where one has $\cos{[2kd\cos{(\theta)}]} \approx 1$, the nonlinearity is quadratic, with $\delta I \propto (\delta d)^2$ for small $\delta d$. 
-Consequently, an oscillation of the form $\delta d \propto \exp{( -i \Omega t )}$ in the arm length difference produces a DC signal and a frequency-doubled signal $\propto \exp{(-2i\Omega t)}$ in $\delta I$.
-This is discussed further in Section~\ref{sec:ifo}.
-
-%Collapse the phase difference to some value $\phi = 2 k d \cos{\theta}$.
-Then expand the norm squared using Euler’s relation, $e^{i \theta} = \cos{\theta} + i \sin{\theta}$.
-Here we stop considering intensity as a function of time as the time dependence will vanish.
-
-%\begin{align}    
-%    I =& A^2 \lvert e^{i (k L - \omega t)} + e^{i (k L + \phi - \omega t)} \rvert^2,\, \phi = 2 k d \cos{\theta} \\
-%    I =& A^2 ((\cos{(k L - \omega t)} + \cos{(k L + \phi - \omega t)})^2 \nonumber\\ &+ (\sin{(k L - \omega t)} + \sin{(k L + \phi - \omega t)})^2)
-%\end{align}
-
-%Expand and collect like terms, pull out a factor of two, and then recognise the form of $\cos{(\alpha-\beta)}$.
-
-%\begin{align}
-%    I =& A^2 (1 + 2 \cos{(k L - \omega t)} \cos{(k L + \phi - \omega t)} \nonumber\\&+ 1 + 2 \sin{(k L - \omega t)} \sin{(k L + \phi - \omega t)}) \\
-%    I =& 2 A^2 (1 + \cos{\phi}),\, \phi = \frac{4 \pi d \cos{\theta}}{\lambda}
-%\end{align}
-
-%Finally, changing derivatives and expand $\phi$ to find that the change in intensity given a change in distance is non-linear, that ${\delta I}/{\delta d}$ is non-constant.
-
-%\begin{align}    
-%    \frac{\delta I}{\delta d} &= \frac{\delta I}{\delta \phi}\; \frac{\delta \phi}{\delta d}\\
-%    \frac{\delta I}{\delta\phi} &= - 2 A^2 \sin{\phi}\\
-%    \frac{\delta\phi}{\delta d} &= \frac{4 \pi \cos{\theta}}{\lambda}\\
-%    \therefore \frac{\delta I}{\delta d} &= \frac{- 8 \pi A^2 \cos{\theta}}{\lambda} \sin{(\frac{4 \pi \cos{\theta}}{\lambda} d)}
-%\end{align}
-
-%This non-linearity in part of the transfer function will make the reverse problem of extracting the original signal more difficult.
-\end{comment}
 
@@
 
-%\jam{[to do!
-from Andrew: ``Write a short appendix proving that maximizing the likelihood exp(-n||n/2) with respect to A given a signal A cos (wt + phi) is the Fourier transform.
-Classic sig proc result useful for teachers to know.'']}
 In this appendix, we demonstrate that the modulus of the Fourier transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
@@
 
-\begin{comment}
-The likelihood is proportional to the probability of observing the data given an assumed model of the signal~\cite{Jaynes:2003}.
-In this example the model is given by Eqn.~\ref{eqn:GNmodel}.
-Under the assumption of Gaussian noise, the log-likelihood is 
-\begin{equation}
-\log \mathcal{L} = -\frac{1}{2}\ip{n}{n}\,, 
-\end{equation}
-where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \frac{1}{T} \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data~\cite{JKS:1998}. 
-\end{comment}
-
-% We replace $n$ by $x-s$ (Eqn.~\ref{eqn:GNdata}), so that 
-%Using Eqns.~\ref{eqn:GNdata} and~\ref{eqn:GNmodel}, we find that
 The likelihood of measuring the signal $s(t)$ in the presence of noise follows from Eqns.~\ref{eqn:GNdata} and~\ref{eqn:ipnn} by replacing $n(t)$ in Eqn.~\ref{eqn:ipnn} with $x(t) - s(t) = n(t)$ from Eqn.~\ref{eqn:GNdata}.~\cite{JKS:1998,Jaynes:2003}
@@
 (i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eqn.~\ref{eqn:likeOne} by varying $s(t)$ through $A$.
-For the special case of the Gaussian likelihood (Eqn.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the distance between $x(t)$ and $s(t)$ as evident in Eqn.~\ref{eqn:ipuv}.
-In general, however, minimizing the distance between $x(t)$ and $s(t)$ is not equivalent always to the fundamental goal of maximising $\mathcal{L}$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
+For the special case of the Gaussian likelihood (Eqn.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the difference between $x(t)$ and $s(t)$ as evident in Eqn.~\ref{eqn:ipuv}.
+In general, however, minimizing the difference between $x(t)$ and $s(t)$ is not equivalent always to the fundamental goal of maximising $\mathcal{L}$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
 (ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eqn.~\ref{eqn:maxLogLike} (or equivalently its logarithm) defines the detection statistic. 
@@
 Therefore the specific functional form of Eqn.~\ref{eqn:maxLogLike} matters, which is a second reason why one must start from Eqn.~\ref{eqn:ipnn} rather than Eqn.~\ref{eqn:likeOne}, in addition to reason (i).
-
-\begin{comment}
-The log-likelihood assuming Gaussian noise is given by
-\begin{eqnarray}
-\log \mathcal{L} &=& -\frac{1}{2}\ip{n}{n}\,, \\
-                 &=& -\frac{1}{2}\ip{x-s}{x-s}\,, \\ 
-                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{2} \ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]} \nonumber\\
-                 &&+ A \ip{x}{\cos[\omega t + \phi]}\,,\\ 
-                 &=& -\frac{1}{2}\ip{x}{x} - \frac{A^2}{4} + A \ip{x}{\cos[\omega t + \phi]}\,,\label{eqn:logLike}
-\end{eqnarray}
-where we define the inner product between $u(t)$ and $v(t)$ as $\ip{u}{v} = \int_0^T \mathrm{d} t \, u(t) v(t) $ and $T$ is the total time of the data (and the last line assumes that $T\gg1/\omega$ so that $\ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]}\approx{1/2}$).
-
-Maximising Eqn.~\ref{eqn:logLike} with respect to $A$ shows that the maximum likelihood is when $A = 2\ip{x}{\cos[\omega t + \phi]}$. 
-The maximum likelihood is then, 
-\begin{eqnarray}
-\label{eqn:maxLogLike}
-\log \mathcal{L}_\mathrm{max} = - \frac{1}{2} \ip{x}{x} + \ip{x}{\cos[\omega t + \phi]}^2\,. 
-\end{eqnarray}
-From the second term of Eqn.~\ref{eqn:maxLogLike}, we see that the maximum log-likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.
-%baring the constant factor $\ip{x}{x}$.  
-\end{comment}
 
@@
 See the Supplementary Material for further resources and other pseudocode examples available online.
-% moved to supp. matt. The interested reader is referred to the excellent review paper by Rabiner~\cite{Rabiner:1989} and textbook by Quinn and Hannam~\cite{QuinnEtAl:2001} to learn more.
 
-%\subsection{Pseudocode}
 Here, we present some pseudocode (below) of the implementation used in Section~\ref{sec:viterbi}.
@@
 
-Let $X$ be grid of $j=0,\ldots,N_f$ rows and $i=0,\ldots,N_t$ columns of additive weights for each node.
+Let $X$ be a grid of $j=0,\ldots,N_f$ rows and $i=0,\ldots,N_t$ columns of additive weights for each node.
 Let Y and Z be grids of the same shape to store the weight of the best path to each node and the row index of the previous node on that path, respectively.
@@
 
-%We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time.
-These are normalised between $(0, 1)$ to use as multiplicative weights.
-In fact, we take the logarithm and so the sum of all weights to avoid underflow.
-The best path has the highest weight among all paths.
-
-%With the grid prepared, the algorithm starts in the second column.
-In each iteration, the algorithm looks for the best path to each cell from the cells in the previous column, selecting the previous cell with the greatest cumulative weight among those ‘nearby’ (the value of which was discovered in the previous iteration).
-Where the cumulative weight is the weight of the best path to reach that previous cell.
-The nearness of a previous cell is given by the limit on the allowed frequency change over time.
-
-%For each cell in the current column, the algorithm calculates the cumulative weight to reach that cell by adding its weight to the cumulative weight to reach it.
-The algorithm also notes the row index of which cell in the previous column it best connects to.
-%Continuing like this until it reaches the end of the grid, the algorithm then looks for the greatest weight in the last column and retraces the indices of the steps it took until it reaches the start of the grid.
-That discovered best path is the Viterbi path.
-
-%\section{A common mistake for those new to wandering frequency signals}
-%\label{app:phase_gotcha}
-%We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$.
-A common mistake is to appeal to the standard form $\sin{(2 \pi f t)}$, where f is a constant frequency, and guess $\sin{(2 \pi f(t) t)}$.
-However, this fails to address the phase accumulated earlier in the signal.
-Instead, the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$.
-Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
-%% This formula is also useful when performing frequency modulation.
-
-\begin{comment} % move to supp matt due to large number of references needed here. 
-\section{Photodiode circuit design}
-\label{app:circuit_diagram}
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-
-Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
-It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. 
-A photo of the breadboard and Raspberry Pi pins is shown in the Supplementary Material.
-The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an ADC to a Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
-Useful throughout was the Pi documentation~\cite{pidocumentation:online} and the pinoutXYZ reference~\cite{pipinout:online}.
-
-\begin{figure*}
-	\includegraphics[width=.8\textwidth]{figures/circuit_diagram_2.pdf}
-	\caption{\label{fig:circuit_diagram}
-Circuit diagram for reading the photodiode. 
-The photodiode is connected in reverse-bias across an op-amp (pale-red box). 
-The op-amp output is passed through a Sallen-Key anti-aliasing filter (pale-blue box) tuned to $16\,{\rm kHz}$, then into an analog-to-digital converter (ADC). 
-The digitized signal is then read by the special-purpose input (SPI) pins of a Raspberry Pi in standard configuration.
-This diagram is produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online} and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-}
-\end{figure*}
-\end{comment}
-
-%\section{Optical Components List}
-%\label{app:opComponents}
-%In this appendix we include a parts list for the optical components of the Michelson interferometer. 
-%The interferometer described in this paper uses parts from ThorLabs~\cite{}, with a total price of AU\$ 1120 at the time of purchase (July 2019). 
-%We also refer the reader to other examples of table-top interferometers at various price points~\cite{TTExhibit:2020,TTExhibit:online,ThorLabsIFO,LIGOIFOGlue,LIGOIFOMagnets,FoxEtAl:1999}.
-
-%\begin{table*}
-%\begin{tabular}{llll}
-%\hline \hline
-%Item & Component & Quantity & Part number \\ 
-%\hline
-%1    & Aluminium optical breadboard $450\,{\rm mm} \times 450\,{\rm mm}$ & 1 & MB4545/M \\
-%\hline \hline
-%\end{tabular}
-%\end{table*}
-
 \end{document}

### 2021-10-27 paper/ifo-complexAudio.tex
grammar and spelling
addressing comments from editors and removing our commented out notes to get ready to share tex files

@@
 In this section, we explore how our apparatus can be used to teach a selection of signal processing techniques. 
-We use complex audio signals (such as music and speech) as a natural successor to the constant and wandering tones used in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, respectively.
+We use complex audio signals (such as music and speech) as natural successors to the constant and wandering tones used in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, respectively.
 As complex audio signals are not quasi-monochromatic, the Viterbi algorithm used in Section~\ref{sec:viterbi_wandering} is not directly applicable here. 
@@
 We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
-We consider a selection of filter techniques, details of which along with a summary of digital signal processing resources can be found in the Supplementary Material. 
+We consider a selection of filter techniques, details of which, along with a summary of digital signal processing resources, can be found in the Supplementary Material. 
 In Section~\ref{sec:opticalMicResults}, we present the two best-performing techniques from the Supplementary Material.
@@
 The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python (see Appendix~\ref{app:code}).
-When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here.
+When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here. \han{NOTE!}
 
+\han{NOTES!}
 The raw output of the optical microphone (with anti-aliasing) is noisy with a loud, continuous bass hum. 
@@
 The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. 
-It amplifies parts of the signal with high signal-to-noise ratio while suppressing parts with low signal-to-noise ratio. 
+It amplifies parts of the signal with a high signal-to-noise ratio while suppressing parts with a low signal-to-noise ratio. 
 The results of the combined cascaded notch and Wiener filter are shown in the third row in Fig.~\ref{fig:notchWienerLogMMSEResults}. 

### 2021-10-27 paper/ifo-introduction.tex
addressing comments from editors and removing our commented out notes to get ready to share tex files

@@
 
-% CW 
-To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and neutron star-black hole mergers.\cite{GWTC-2:2020,NSBH:2021}
+To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and binaries consisting of a neutron star and a black hole.\cite{GWTC-2:2020,NSBH:2021}
 However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
@@
 Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
-Numerous, as yet unsuccessful, searches have been performed for Scorpius X-1 and other LMXBs~(e.g., Ref.~\cite{ScoX1O2Viterbi:2019}).
+Numerous searches, as yet unsuccessful, have been performed for Scorpius X-1 and other LMXBs~(e.g., Ref.~\cite{ScoX1O2Viterbi:2019}).
 
@@
 We then extend its use to an ``optical microphone'', using light to capture sound, and present a range of example analysis techniques for educators to use. 
-As an undergraduate lab experiment, the demonstration can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement.
+As an undergraduate lab experiment, the apparatus can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement.
 It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data.\cite{SuvorovaEtAl:2017, ScoX1O2Viterbi:2019}

### 2021-10-27 paper/ifo-singleTone.tex
grammar and spelling
addressing comments from editors and removing our commented out notes to get ready to share tex files

@@
 	\caption{\label{fig:webcam_spectrum}
-Recovery of a tone at a constant frequency: the Fourier amplitude of the intensity pattern against frequency.
+Recovery of a tone at a constant frequency: the Fourier amplitude of the intensity pattern plotted against frequency.
 The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has an FWHM of $0.033\,{\rm Hz}$.
@@
 They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo}.
-Also, note that the noise appears to rise at lower frequencies; however the origin of this noise is not known.
+Also, note that the noise appears to rise at lower frequencies; however, the origin of this noise is not known.
 

### 2021-10-27 paper/ifo-tableTopGWs.tex
addressing comments from editors and removing our commented out notes to get ready to share tex files

@@
 Current generation gravitational-wave observatories have kilometer-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo, respectively). 
-They can measure minuscule changes in distance due to gravitational waves; e.g., the first detection of a binary black hole merger produced a strain of $10^{-21}$,\cite{GW150914} corresponding to a change in distance equal to a fraction of the width of a proton.
+They can measure minuscule changes in distance due to gravitational waves; e.g., the first detection of a binary black hole merger produced a strain of $10^{-21}$,\cite{GW150914} corresponding to a change in distance equal to a fraction of the diameter of a proton.
 
@@
 	\caption{\label{fig:ifo_schematic_webcam}
+\han{Notes!}
 Schematic of the table-top Michelson interferometer. 
 The output of a Class 2 $532\,{\rm nm}$ laser (top-left) is reflected from a mirror at a right angle.
-The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter which reflects approximately $50\%$ of the beam and transmits the remaining $\sim50\%$.
+The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter, which reflects approximately $50\%$ of the beam and transmits the remaining $\sim50\%$.
 The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and recombine at the beamsplitter. 
@@
 
-The motion of the interference fringes follows the motion of M2, and therefore of the speaker.
+\han{NOTE!} The motion of the interference fringes follows the motion of M2, and therefore of the speaker.
 The amplitude of these motions is given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.

### 2021-10-27 paper/ifo-wanderingTone.tex
addressing comments from editors and removing our commented out notes to get ready to share tex files

@@
 
-A continuous-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries such as LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
+A continuous gravitational-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries such as LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
 Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches can be impacted by very small frequency drifts. 
@@
 In continuous-wave searches, the wandering frequency problem is solved using the Viterbi algorithm,\cite{Viterbi:1967} which can track the signal's frequency over time.
-The analysis described and presented in this section is directly inspired by real continuous-wave search methods~\cite{SuvorovaEtAl:2017} yet is pitched at a level appropriate for an undergraduate laboratory setting.
+The analysis described and presented in this section is directly inspired by real continuous-wave search methods,~\cite{SuvorovaEtAl:2017} yet is pitched at a level appropriate for an undergraduate laboratory setting.
 In Section~\ref{sec:realCWSearches}, we briefly review the analysis method used by LIGO and Virgo; in Section~\ref{sec:viterbi}, we describe the methods as applied in this work; and in Section~\ref{sec:wanderingResults}, we show results for recovering a wandering frequency signal using the table-top interferometer.
@@
 The small to large sizing corresponds to low to high likelihood values in arbitrary units for this schematic. 
-The color gradient (online only) also represents the likelihood values where the dark blue corresponds to low likelihood and light green corresponds to high likelihood. 
 The objective is to find the most probable path of a signal through the grid from left to right.
@@
 Fig.~\ref{fig:viterbi} represents a spectrogram with $N_t$ time bins and $N_f$ frequency bins.
-The circlular nodes represent the elements of the time and frequency grid, which are labelled as  
+The circular nodes represent the elements of the time and frequency grid, which are labelled as  
 $t_i$ and $f_j$, respectively, where $i=0,1,2,\mathellipsis,N_t$ and $j=0,1,2,\mathellipsis,N_f$. 
@@
 We simulate a slowly wandering signal by modulating the frequency sinusoidally with a modulation amplitude that decays with time. 
-We use the same set-up as shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
+We use the same apparatus as shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
 In this section, we test the Viterbi algorithm’s ability to recover the wandering signal. 
@@
 The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heat-map shows the spectrogram of the observed signal (similar to that represented by the schematic in Fig.~\ref{fig:viterbi}). 
-In this demonstration we use a signal that can easily be identified in the spectrogram, however, we expect real continuous-wave signals to have far lower signal-to-noise ratios at each time step. 
+In this demonstration we use a signal that can easily be identified in the spectrogram; however, we expect real continuous-wave signals to have far lower signal-to-noise ratios at each time step. 
 The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.
@@
 If we remove the first four time bins from the $E_{\rm rms}$ calculation, we find a slight improvement in recovery with $E_{\rm rms} = 0.044$.
-One may be tempted to increase the allowed frequency wander in the algorithm, however, this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.
+One may be tempted to increase the allowed frequency wander in the algorithm; however, this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise.
 There is also an anomaly at $150\,{\rm s}$, which is likely due to a local disturbance, e.g. someone walking past the interferometer. 

### 2021-10-27 paper/ifo-supplementary-material.tex
editor comments on supp matt and removing our commented items before sharing files

@@
-% GravExplainSuppMatt
-
 \documentclass[aps,pra,superscriptaddress,reprint]{revtex4-2}
@@
 \subsection{Gravitational-wave engagement}
-Numerous engagement activities have been developed by gravitational-wave research groups around the world~\cite{TTExhibit:2021,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online, chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
+Numerous outreach activities have been developed by gravitational-wave research groups around the world~\cite{TTExhibit:2021,CavagliaExhibit:2009, LIGOScienceEducationCentre:online, GravityDiscoveryCentre:online, GWOSC:online, LOSC:2015, LaserLabs:online, SciVR:online, chirp:online, BlackHoleHunter:online, ArthurJeffesMusic:online, GravitySynthLeonTrimble:online}.
 Specific examples of using table-top interferometers as engagement or teaching tools include Refs.~\cite{TTExhibit:2021,UgoliniEtAl:2019,LibbrechtBlack:2015,AMIGO:online,FoxEtAl:1999} and parts lists or instructions can be found in Refs.~\cite{TTExhibit:online,LIGOIFOGlue,LIGOIFOMagnets,NikhefIFO,ThorLabsIFO}.
@@
 In gravitational-wave observatories, the optical components are suspended from a series of pendulums to achieve passive isolation, and feedback control loops provide active isolation\cite{MatichardEtAl:2015, AdvancedVirgo:2015}.
-Such a system is beyond the scope of this article, however, feedback control has been demonstrated in table-top Michelson interferometers designed for the undergraduate laboratory~\cite{LibbrechtBlack:2015}.
+Such a system is beyond the scope of this article. 
+However, feedback control has been demonstrated in table-top Michelson interferometers designed for the undergraduate laboratory~\cite{LibbrechtBlack:2015}.
 Other sources of noise include acoustic sound waves (for example from nearby vehicles, wind, and components such as electric fans) and magnetic noise (for example from high-voltage power lines, building heaters and lights, or any electrical power circuit)~\cite{EfflerEtAl:2015,CironeEtAl:2019}.
@@
 Where this is not possible, noise subtraction can be done after data collection~\cite{DriggersEtAl:2019, DavisEtAl:2019}. 
-One example is the removal of spectral noise lines from the United States power grid operating at $60\,{\rm Hz}$~\cite{VajenteEtAl:2020}. %United States's?
+One example is the removal of spectral noise lines from the United States power grid operating at $60\,{\rm Hz}$~\cite{VajenteEtAl:2020}.
 
@@
 The cascaded notch and Wiener filter are described in Section~\ref{sec:opmic} in this Supplementary Material and also in  Refs.~\cite{10.5555/541204,10.5555/151045}.
-Further details about the logMMSE estimator can be found in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and for speech intelligibility see Refs.~\cite{speech_intelligibility,SubjectiveComparison}
+Further details about the logMMSE estimator can be found in Ref.~\cite{Ephraim1984SpeechEU_logMMSE} and for speech intelligibility see Refs.~\cite{speech_intelligibility,SubjectiveComparison}.
 
@@
 Cross-disciplinary experimental projects between physics and engineering students can increase student understanding of overlapping concepts through contextual learning and also improve student motivation~\cite{PiersonGurlandCrawford:2002}.
-Realistic and open-ended projects spanning both topics allow students a better understanding of real-world applications of these topics and of the dependency between physicists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
+Realistic open-ended projects spanning both topics allow students to better understand real-world applications of these topics and the interdependency between physicists and engineers (as shown by the collaborative project presented in Ref.~\cite{OConnerSibrayForinash:2001}).
 The optical microphone presented here can similarly be used as a teaching tool for students to conduct investigations into signal processing techniques with a physical system.
@@
 A photograph of the equipment is shown in Fig.~\ref{fig:photographs_of_optical_microphone_and_circuit} with the interferometer and circuit shown on the left and right, respectively. 
-% schematic of breadboard and connections to pi
-% https://www.circuit-diagram.org/editor/
-% https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce
-The photodiode circuit diagram is shown in Figure~\ref{fig:circuit_diagram}
-The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an analogy-to-digital converter ADC to a Raspberry~Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
-Useful throughout was the Raspberry Pi documentation~\cite{pidocumentation:online} and the pinoutXYZ reference~\cite{pipinout:online}.
+The photodiode circuit diagram is shown in Figure~\ref{fig:circuit_diagram}.
+The design was based on standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an analogy-to-digital converter ADC to a Raspberry~Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}.
+The Raspberry Pi documentation~\cite{pidocumentation:online} and the pinoutXYZ reference~\cite{pipinout:online} were also useful.
 
@@
 This diagram was produced using the Circuit Diagram online tool~\cite{CircuitDiagram:online}.
-% and is available from~\url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}.
-this url has the incorrect resistors and capacitors - James
 }
@@
 	\begin{center}
-    %% trim order is <left> <lower> <right> <upper>
-    % height=0.5\textheight,trim={1cm 4.5cm 0.4cm 3.5cm},clip
 	\includegraphics[width=.48\textwidth]
-% trim={1cm 3cm 1cm 1cm}]
 {figures/combined_results_full_melatos_labelled.pdf}
-%notch_and_wiener_superplot_v2.pdf}
 	\end{center}
 	\caption{\label{fig:op_mic_results}
-	%\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}
 	Timeseries (left column) and frequency spectrum (right column) results for the background subtraction, notch, and Wiener filters.
@@
 Butterworth bandpass filter frequency response. 
-%Any amplitude below $-3\,{\rm dB}$ is significant attenuation (less than half the power at that frequency remains). 
-% suggest removing as -3db is hard to judge on the figure and I think the attenuation outside the frequency range is the key point to get across here
 The red, dashed vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$. Note the flat response within the band characteristic of the Butterworth filter.}
@@
 A Butterworth band-pass filter is a particular band-pass filter such that the frequency response (the attenuation at each frequency) is ``maximally flat'' within the band.
-The Butterworth low-pass component is given by %Eqn.~\ref{eq:butterworth}
-
+The Butterworth low-pass component is given by
 \begin{equation}
@@
 where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$,  
-%\begin{equation*}
-%\left\{ 
 \begin{eqnarray} 
@@
 r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
-%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
-%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
 \end{eqnarray}
-%\right\}  
-%\end{equation*}
 
@@
 
-\begin{comment}
-\begin{figure*}
-	%\includegraphics[width=0.49\textwidth]{figures/combined_timeseries_melatos_shift_source.pdf}
-	%\includegraphics[width=0.49\textwidth]{figures/combined_spectrum_melatos.pdf}
-    \includegraphics[height=0.57\textheight]{figures/combined_timeseries_melatos_shift_source_tall.pdf}
-    \includegraphics[height=0.57\textheight]{figures/combined_spectrum_melatos_tall.pdf}
-    \caption{\label{fig:logMMSE_timeseries_freqspectrum}
-\jam{Fig.2 should be combined with Fig.6 using combined\_results\_full\_melatos\_labelled.pdf}    - done 
-Timeseries (left column) and frequency spectrum (right columns) results with the logMMSE estimator. 
-The original source signal and optical microphone recording in the top and middle panels respectively are for comparison and are identical to those shown in the top two panels in Fig.~\ref{fig:BackgroundNotchWienerCombined} (again, the source signal is shifted by $0.12\,{\rm s}$).
-The bottom-left panel shows the timeseries of the recording after filtering with the logMMSE estimator, where the rise at the start of the timeseries is expected when filtering a signal of finite duration.
-The bottom-right panel shows the corresponding frequency spectrum with the logMMSE estimator. 
-The spectrum shows only a detail of the total frequency domain (there is little activity at higher frequencies) and has truncated peaks at amplitude $2$ which otherwise dominate the plot. 
-The change in scale from the top panel is not of concern as the ear hears the relative frequency content and normalisation of the timeseries fixes any scaling. 
-    }
-\end{figure*}
-\end{comment}
-
 Speech enhancement of noisy channels is a classic problem in signal processing. 

### 2021-11-01 paper/ifo-complexAudio.tex
removing our comments to share source files

@@
 \label{sec:photodiode}
+
 The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
@@
 The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python (see Appendix~\ref{app:code}).
-When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second here. \han{NOTE!}
+When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second in our results (Fig.~\ref{fig:notchWienerLogMMSEResults}).
 
-\han{NOTES!}
 The raw output of the optical microphone (with anti-aliasing) is noisy with a loud, continuous bass hum. 
 This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., the output with the speaker switched off), shown in Fig.~\ref{fig:psd_noise}. 
-The spectrum is dominated by AC mains noise with power from the fundamental $50\,{\rm Hz}$ Australian mains power grid signal up to and beyond the $8$th harmonic. 
+The spectrum is dominated by AC eletrical power grid noise from the fundamental $50\,{\rm Hz}$ Australian mains electricity grid signal up to and beyond the $8$th harmonic (at $400\,{\rm Hz}$). 
 The mains signal is also present, but far weaker, in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 

### 2021-11-01 paper/ifo-tableTopGWs.tex
removing our comments to share source files

@@
 	\caption{\label{fig:ifo_schematic_webcam}
-\han{Notes!}
 Schematic of the table-top Michelson interferometer. 
@@
 
-\han{NOTE!} The motion of the interference fringes follows the motion of M2, and therefore of the speaker.
+The motion of the interference fringes follows the motion of M2, and therefore of the speaker.
 The amplitude of these motions is given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.

### 2021-11-01 paper/paper_main.tex
changing to new revtex4-2
removing our comments to share source files

@@
-% GravExplain
-%\documentclass[prb,reprint,nofootinbib]{revtex4-1} 
-\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-1}
-% \documentclass[prb,preprint,letterpaper,noeprint,longbibliography,nodoi,footinbib]{revtex4-1}
+\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-2}
 
-% Note that AJP uses the same style as Phys. Rev. B (prb).
-% The AIP Style Manual\cite{AIPstylemanual} is an indispensable reference on good physics writing, covering everything from planning and organization to standard spellings and abbreviations.
-% Most important of all, please familiarize yourself with the AJP Statement of Editorial Policy,\cite{editorsite} which describes the types of manuscripts that AJP publishes and the audience for which AJP authors are expected to write.
-% We look forward to receiving your submission to AJP.
 \usepackage[utf8]{inputenc}
@@
 \usepackage{hyperref}
-% \usepackage{siunitx}
 \usepackage{textcomp}
@@
 \usepackage{algpseudocode}
-% \usepackage{algorithm}
-% \usepackage{algorithmic}
 \usepackage{lmodern}
@@
 
-%for line number
 %\usepackage{lineno}
@@
 
-% inner product
 \newcommand{\ip}[2]{\langle #1,#2 \rangle}
-
-\newcommand{\jam}{\textcolor{magenta}}
-\newcommand{\han}{\textcolor{orange}}
-\newcommand{\fixme}{\textcolor{blue}}
 
@@
 
-%\author{\textbf{[Authors removed for double-anonymous review.]}}
-
 \input{authorList.tex}
 
-% Summer 2019/2020
 \date{\today}
@@
 Here, we present highlights of our results using a combined notch plus Wiener filter and the statistical log minimum mean-square error (logMMSE) estimator. 
-% Recordings of simple chords and drums are easily recovered using these techniques, but complex music and speech are more challenging. 
 Using these techniques, we easily recover recordings of simple chords and drums, but complex music and speech are more challenging.
@@
 \begin{acknowledgments}
-%\textbf{[Removed for double-anonymous review.]}
 
-The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky, and Sam Cooper, for their helpful feedback and suggestions.
+The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky, Sam Cooper, and Martin Hendry, for their helpful feedback and suggestions.
 This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
@@
 
-% \nocite{*}
 \bibliographystyle{myunsrt}

### 2021-11-01 paper/ifo-appendix.tex
removing our comments to share source files
changing eqn to eq for AJP style guide

@@
 
-The likelihood of measuring the signal $s(t)$ in the presence of noise follows from Eqns.~\ref{eqn:GNdata} and~\ref{eqn:ipnn} by replacing $n(t)$ in Eqn.~\ref{eqn:ipnn} with $x(t) - s(t) = n(t)$ from Eqn.~\ref{eqn:GNdata}.~\cite{JKS:1998,Jaynes:2003}
+The likelihood of measuring the signal $s(t)$ in the presence of noise follows from Eqs.~\ref{eqn:GNdata} and~\ref{eqn:ipnn} by replacing $n(t)$ in Eq.~\ref{eqn:ipnn} with $x(t) - s(t) = n(t)$ from Eq.~\ref{eqn:GNdata}.~\cite{JKS:1998,Jaynes:2003}
 The result is
@@
 \end{eqnarray}
-We then maximise Eqn~.\ref{eqn:logLike} with respect to $A$, obtaining 
+We then maximise Eq.~\ref{eqn:logLike} with respect to $A$, obtaining 
 \begin{eqnarray}
@@
 for $A = 2\ip{x}{\cos[\omega t + \phi]}$.
-From the second term of Eqn.~\ref{eqn:maxLogLike}, we see that the maximum likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.
+From the second term of Eq.~\ref{eqn:maxLogLike}, we see that the maximum likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.
 
 Two important points must be made about the above procedure. 
-(i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eqn.~\ref{eqn:likeOne} by varying $s(t)$ through $A$.
-For the special case of the Gaussian likelihood (Eqn.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the difference between $x(t)$ and $s(t)$ as evident in Eqn.~\ref{eqn:ipuv}.
+(i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eq.~\ref{eqn:likeOne} by varying $s(t)$ through $A$.
+For the special case of the Gaussian likelihood (Eq.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the difference between $x(t)$ and $s(t)$ as evident in Eq.~\ref{eqn:ipuv}.
 In general, however, minimizing the difference between $x(t)$ and $s(t)$ is not equivalent always to the fundamental goal of maximising $\mathcal{L}$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
-(ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eqn.~\ref{eqn:maxLogLike} (or equivalently its logarithm) defines the detection statistic. 
+(ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eq.~\ref{eqn:maxLogLike} (or equivalently its logarithm) defines the detection statistic. 
 When its value exceeds a threshold (chosen freely by the analyst) at some value of $\omega$, a signal is deemed to have been detected at that value of $\omega$.
-Therefore the specific functional form of Eqn.~\ref{eqn:maxLogLike} matters, which is a second reason why one must start from Eqn.~\ref{eqn:ipnn} rather than Eqn.~\ref{eqn:likeOne}, in addition to reason (i).
+Therefore the specific functional form of Eq.~\ref{eqn:maxLogLike} matters, which is a second reason why one must start from Eq.~\ref{eqn:ipnn} rather than Eq.~\ref{eqn:likeOne}, in addition to reason (i).
 
@@
 \label{app:viterbi}
+
 This appendix contains some details regarding the implementation of the Viterbi algorithm described in Section~\ref{sec:viterbi}. 
@@
 \Function{Viterbi}{$X$}
-    % \State $a \gets a+1$
-    % \State \Return $a$
 

### 2021-11-01 paper/ifo-supplementary-material.tex
removing our comments to share source files
changing eqn to eq for AJP style guide

@@
 \WarningFilter{revtex4-1}{Repair the float}
-
-\newcommand{\jam}{\textcolor{magenta}}
-\newcommand{\han}{\textcolor{orange}}
 
@@
 and is combined with a similar high-pass filter to form the band-pass filter.
-In Eqn.~\ref{eq:butterworth}, $f_c$ is the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ is the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
+In Eq.~\ref{eq:butterworth}, $f_c$ is the cut-off frequency of the low-pass Butterworth filter, $\varepsilon$ is the gain, and $n$ is the order of the filter which determines how quickly the response rolls off past the cut-off frequency.
 Fig.~\ref{fig:butterworth} shows the frequency response of the filter used here: a fifth order ($n = 5$) Butterworth filter with a pass-band of $(150\,{\rm Hz}$, $3\,{\rm kHz})$.
@@
 If we let the Wiener filter be non-causal (i.e.
-we estimate the current signal based on both past \emph{and future} observations), then we can represent Eqn.~\ref{eqn:wiener-hopf} in the frequency domain as
+we estimate the current signal based on both past \emph{and future} observations), then we can represent Eq.~\ref{eqn:wiener-hopf} in the frequency domain as
 \begin{equation}
@@
 where $P_{xx}(f), P_{ss}(f), P_{ww}(f)$ are the spectra of the observed noisy data, the injected signal, and the background noise, respectively.
-Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal-to-noise ratio (SNR) is high and attenuates the signal where the SNR is low.
+Intuitively, from Eq.~\ref{eqn:wiener}, we can see that the non-causal Wiener filter amplifies the input signal where the signal-to-noise ratio (SNR) is high and attenuates the signal where the SNR is low.
 The causal Wiener filter (which only makes estimates based on past observations) is similar.
@@
 
-In this work, we construct a higher-order causal Wiener filter based on Eqn.~\ref{eqn:wiener-hopf}.
+In this work, we construct a higher-order causal Wiener filter based on Eq.~\ref{eqn:wiener-hopf}.
 A higher-order Wiener filter provides greater smoothing of the input signal but also increases the computational memory required.

### 2021-11-01 paper/ifo-wanderingTone.tex
removing our comments to share source files
changing eqn to eq for AJP style guide

@@
 Some routes through the grid result in dead-ends with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a).
-At $t=t_{N_t}$ the algorithm chooses the terminating frequency node which has the highest value given by Eqn.~\ref{eqn:pathprobability}, marked (b). 
+At $t=t_{N_t}$ the algorithm chooses the terminating frequency node which has the highest value given by Eq.~\ref{eqn:pathprobability}, marked (b). 
 The Viterbi path is the path leading to this node, highlighted by the thick-black lines, from (b) to the start at (c). 
@@
 The path $f^\ast(t_0),\mathellipsis,f^\ast(t_n)$ that maximises ${\rm Pr}[ f(t_0), f(t_1),\mathellipsis, f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
-We note that the left-hand-side and right-hand-side of Eqn.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
+We note that the left-hand-side and right-hand-side of Eq.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.
 
@@
 
-\item Moving to time $t_2$, again we select the path which maximizes Eqn.~\ref{eqn:pathprobability} for each $f_j$. 
+\item Moving to time $t_2$, again we select the path which maximizes Eq.~\ref{eqn:pathprobability} for each $f_j$. 
 These paths are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.
@@
 
-\item We have now found the most probable path to each $f_j$ at $t=t_{N_t}$ and its probability (Eqn.~\ref{eqn:pathprobability}). 
+\item We have now found the most probable path to each $f_j$ at $t=t_{N_t}$ and its probability (Eq.~\ref{eqn:pathprobability}). 
 We select the terminating frequency bin $f(t_{N_t})$ with the highest probability labeled as (b) in Fig.~\ref{fig:viterbi}.
@@
 In continuous gravitational wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency can change unpredictably over time. 
-% how to hyphenate ``continuous gravitational wave searches''?
 The Viterbi algorithm's strength lies in its ability to track such signals through the data even in the presence of comparatively loud noise fluctuations in the time-frequency bins.

### 2021-11-01 paper/finalSubmission/JGardnerManuscript/JGardner.tex
getting things ready to submit

%\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-2}
\documentclass[prb,preprint]{revtex4-2}

\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{float}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{hyperref}
\usepackage{textcomp}
\usepackage{subfiles}
\usepackage{comment}
\usepackage{algpseudocode}
\usepackage{lmodern}

\usepackage{silence}
\WarningFilter{revtex4-1}{Repair the float}

%\usepackage{lineno}
%\linenumbers

\newcommand{\ip}[2]{\langle #1,#2 \rangle}

\begin{document}

\title{Continuous gravitational waves in the lab: recovering audio signals with a table-top~optical microphone}

\input{authorList.tex}

\date{\today}

\begin{abstract}
Gravitational-wave observatories around the world are searching for continuous waves: persistent signals from sources such as spinning neutron stars. 
These searches use sophisticated statistical techniques to look for weak signals in noisy data. 
In this paper, we demonstrate these techniques using a table-top model gravitational-wave detector: a Michelson interferometer where sound is used as an analog for gravitational waves. 
Using signal processing techniques from continuous-wave searches, we demonstrate the recovery of tones with constant and wandering frequencies. 
We also explore the use of the interferometer as a teaching tool for educators in physics and electrical engineering by using it as an ``optical microphone'' to capture music and speech. 
A range of filtering techniques used to recover signals from noisy data are detailed in the Supplementary Material. 
Here, we present highlights of our results using a combined notch plus Wiener filter and the statistical log minimum mean-square error (logMMSE) estimator. 
Using these techniques, we easily recover recordings of simple chords and drums, but complex music and speech are more challenging.
This demonstration can be used by educators in undergraduate laboratories and can be adapted for communicating gravitational-wave and signal-processing topics to non-specialist audiences.  
\end{abstract}

\maketitle

\section{Introduction}
\label{sec:introduction}

In 2015, the first detection of gravitational waves from the merger of two black holes marked a breakthrough in modern astrophysics and revealed a new means to observe the Universe.\cite{GW150914}
Gravitational waves are a prediction of Albert Einstein's theory of General Relativity; they are disturbances in spacetime caused by the acceleration of asymmetric massive objects.
The effect of gravitational waves is a change in lengths: a ``stretching and squashing'' of the distance between objects. 
Ground-based gravitational-wave observatories such as the Advanced Laser Interferometer Gravitational-wave Observatory (LIGO), Advanced Virgo, GEO600, and KAGRA use the interference of laser light to measure changes in distance. 
These observatories are extremely complex but are fundamentally based on the Michelson interferometer. 
Table-top interferometers are commonly used in undergraduate laboratory experiments and to demonstrate the science of gravitational-wave detection to non-specialist audiences.\cite{TTExhibit:2021}

To date, the network of gravitational-wave observatories has observed short-duration transient signals from the mergers of binary black holes, binary neutron stars, and binaries consisting of a neutron star and a black hole.\cite{GWTC-2:2020,NSBH:2021}
However, the network is also searching for continuous gravitational waves: persistent, periodic, near-monochromatic signals, which are yet to be detected. 
Rotating neutron stars are prime candidates for continuous-wave emission, especially those in low mass X-ray binaries (LMXB), where the neutron star is in orbit with a low mass stellar companion.
The rotation frequency of the neutron star in an LMXB can wander over time due to variable accretion of matter (and hence angular momentum transfer) from the stellar companion.\cite{xraybinaries:1997}
Scorpius X-1 is a prime LMXB target for continuous-wave searches. 
Numerous searches, as yet unsuccessful, have been performed for Scorpius X-1 and other LMXBs~(e.g., Ref.~\cite{ScoX1O2Viterbi:2019}).

In this paper, we use a table-top Michelson interferometer as a toy gravitational-wave detector designed to detect sound instead of gravitational waves. 
We then extend its use to an ``optical microphone'', using light to capture sound, and present a range of example analysis techniques for educators to use. 
As an undergraduate lab experiment, the apparatus can be used to teach topics ranging from continuous-wave detection and analysis to electronics, signal processing, and speech enhancement.
It allows students in courses such as physics and electrical engineering to explore the response of an accessible, yet nontrivial, optomechanical system using a hierarchy of data analysis techniques of increasing complexity, including those used in the search for continuous waves in LIGO and Virgo data.\cite{SuvorovaEtAl:2017, ScoX1O2Viterbi:2019} 
This demonstration also has the potential to be used as an outreach tool alongside a range of other public engagement demonstrations and activities developed by gravitational-wave research groups around the world. 
These tools allow scientists to cater to the increased public and media interest in this field and explain gravitational-wave science to non-specialist audiences.

This paper is laid out as follows. 
In Section~\ref{sec:ifo}, we detail the table-top interferometer design. 
In Section~\ref{sec:single_tone}, we demonstrate observing a single tone from a speaker. 
In Section~\ref{sec:viterbi_wandering}, we observe a wandering frequency signal and analyze it using a hidden Markov model technique (the Viterbi algorithm) from continuous-wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capture and playback of complex audio, such as music and speech.
This demonstration of an optical microphone serves as a more general exhibition of signal processing with a range of examples that can be used in the undergraduate laboratory (described in the Supplementary Material).
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}. 
Further reading and resources are provided in the Supplementary Material and we present the software and scripts used to produce this work in Appendix~\ref{app:code}.

\section{Table-top gravitational-wave science}
\label{sec:ifo}

Gravitational-wave detectors such as LIGO and Virgo are large, complex experiments. 
However, their design is fundamentally based on the Michelson interferometer, an optical configuration commonly used in undergraduate laboratories. 
In a Michelson interferometer, laser light is split by a beamsplitter into two perpendicular arms, shown in Fig.~\ref{fig:ifo_schematic_webcam}. 
Mirrors at the end of each arm reflect the two beams back to the beamsplitter where they recombine to produce an interference pattern.
The resulting interference pattern is dependent on the relative distance traveled by the beams. 
Current generation gravitational-wave observatories have kilometer-scale arms (with arm lengths $4\,{\rm km}$ and $3\,{\rm km}$ at LIGO and Virgo, respectively). 
They can measure minuscule changes in distance due to gravitational waves; e.g., the first detection of a binary black hole merger produced a strain of $10^{-21}$,\cite{GW150914} corresponding to a change in distance equal to a fraction of the diameter of a proton.

Sound is a commonly used analogy when explaining gravitational-wave science. 
Gravitational-wave signals from binary black hole and neutron star mergers can be converted to audio signals~\footnote{When introducing the acoustic analogy to lay audiences, it is important to emphasize that gravitational waves are not sound. For example, gravitational waves can propagate in a vacuum, whereas sound cannot. Gravitational waves also travel at the speed of light and are not longitudinal waves.} to aid in explanations.
Detection and analysis techniques used by the gravitational-wave community can be demonstrated using table-top equipment (see Ref.~\cite{TTExhibit:2021} as well as the further reading section in the Supplementary Material for a selection of other table-top Michelson interferometer designs used for science outreach).
Sound vibrations provide a simple means to move the components of a demonstration interferometer, changing the length of the interferometer arms, and therefore the interference pattern.
We use audio signals throughout this work to simulate gravitational wave--like signals.

\begin{figure}
	\includegraphics[width=0.5\textwidth]{figures/fig01_ifo_schematic_webcam.pdf}
	\caption{\label{fig:ifo_schematic_webcam}
Schematic of the table-top Michelson interferometer. 
The output of a Class 2 $532\,{\rm nm}$ laser (top-left) is reflected from a mirror at a right angle. The beam then passes through a converging lens (focal length: $125.0\,{\rm mm}$) before reaching the beamsplitter, which reflects approximately $50\%$ of the beam and transmits the remaining $\sim50\%$.
The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and recombine at the beamsplitter. 
The output interference pattern is enlarged using a diverging lens (focal length: $-25.0\,{\rm mm}$), projected onto a screen, and recorded using a commercial webcam connected to a computer (PC). 
A speaker fixed to the back of M2 is used to inject audio signals from the PC into the interferometer.
    }
\end{figure}

The optical configuration of the Michelson interferometer used in this work is shown in Fig.~\ref{fig:ifo_schematic_webcam}.
It is assembled on a $450 \times 450 \,{\rm mm} $ optical breadboard and uses a green laser with peak emission at a wavelength of $532\,{\rm nm}$.
The output of the laser is first reflected by a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard and to have greater control of the alignment of the interferometer).
After passing through a converging (plano-convex) lens with focal length $125.0\,{\rm mm}$, the laser beam is incident on a beamsplitter, which reflects half of the light towards mirror 1 (M1) and transmits the other half to mirror 2 (M2). 
A $0.5\,{\rm W}$ speaker is fixed to the back of M2 using commercial adhesive putty and serves as a controllable source of vibrations. 
This speaker is one of a pair of commercial, USB-powered speakers, fed by a $3.5\,{\rm mm}$ jack and driven by a computer (PC). 
The other speaker in the pair is kept face-down and as far away from the interferometer as possible to prevent it from acting as a second source.
The beams are reflected by mirrors M1 and M2, located at the end of $\sim 7.5\,{\rm cm}$ and $\sim 10.0\,{\rm cm}$ long arms.
The beams recombine at the beamsplitter and produce an interference pattern that is then enlarged using a diverging (bi-concave) lens of focal length $-25.0\,{\rm mm}$, and projected onto a screen.
The interference pattern, as shown by the photograph in Fig.~\ref{fig:interference_pattern}, is a set of concentric light and dark rings (fringes). 
A change in the relative length of the arms causes these rings to move radially inwards or outwards.
The intensity timeseries is recorded by either a webcam (in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}) or a photodiode, which offers a higher sampling rate suitable for capturing more complex audio (in Section~\ref{sec:optical_microphone}).
Further design details for Michelson interferometers can be found in Ref.~\cite{TTExhibit:2021}.

\begin{figure}
 \begin{center}
  \includegraphics[width=0.35\textwidth, angle=-90]{figures/fig02_webcam_still_star.pdf}
 \end{center}
 \caption{\label{fig:interference_pattern}
The interference pattern produced by the table-top Michelson interferometer.
This image was taken with the webcam aligned off the beam axis. 
The pink star indicates the point where the intensity timeseries measurements were taken.
The central bright fringe of the interference pattern is approximately $5\,{\rm mm}$ in diameter. 
}
\end{figure}

The motion of the interference fringes follows the motion of M2, and therefore of the speaker.
The amplitude of these motions is given by a transfer function that accounts for the coupling and resonance of the speaker-mirror system.
If the relative change in the length of the arms is kept small enough, then the intensity of the interference pattern at any point on the screen oscillates at the same frequency as the injected audio.
For larger relative length changes, multiple fringes will pass through the detection point during a single speaker oscillation, raising the measured frequency artificially. 
As such, any motion of the fringes must be kept small by playing the sound softly through the speaker.
Even without over-counting, large fringe motions display a nonlinear relation between the intensity fluctuations and injected audio, leading to troublesome---but physically interesting---phenomena like frequency doubling.

\section{Constant frequency signal}
\label{sec:single_tone}

\begin{figure}
	\includegraphics[width=.5\textwidth]{figures/fig03_webcam_spectrum.pdf}
	\caption{\label{fig:webcam_spectrum}
Recovery of a tone at a constant frequency: the Fourier amplitude of the intensity pattern plotted against frequency.
The injected signal has a frequency of $2.09\,{\rm Hz}$, while the recovered signal peaks at $2.099\,{\rm Hz}$ and has an FWHM of $0.033\,{\rm Hz}$.
The plot also shows two harmonics at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ with amplitudes $10.3 \%$ and $9.8 \%$ of the amplitude of the primary peak, respectively.
}	
\end{figure}

Continuous-wave searches look for nearly monochromatic signals.\cite{JKS:1998} 
In this section, we consider a simple sinusoidal tone at a single, constant frequency.

As described in Section~\ref{sec:ifo}, the audio signal is played through a speaker fixed to the back of M2 (see Fig.~\ref{fig:ifo_schematic_webcam}). 
The intensity of the interference pattern is measured at a single point on the screen, indicated by the pink star in Fig.~\ref{fig:interference_pattern}. 
The webcam records video in three color channels: red, blue, and green. 
We use the green channel as an approximation of the total intensity produced by the green laser.
The webcam samples at a rate of $30\, {\rm Hz}$, which limits the spectral content of observable signals to less than $15\,{\rm Hz}$, the Nyquist frequency.

A tone with a frequency of $2.09\,{\rm Hz}$ is played through the speaker for one minute. 
The amplitude of the Fourier transform of the recovered signal is shown in Fig.~\ref{fig:webcam_spectrum}. The discrete Fourier transform is calculated using the NumPy package for Python (see Appendix~\ref{app:code}).
We measured a peak amplitude at $2.099\,{\rm Hz}$ with a full width half maximum (FWHM) of $0.033\,{\rm Hz}$.
Two harmonics can also be seen at integer multiples of the peak frequency. 
The peaks at $4.19\,{\rm Hz}$ and $6.28\,{\rm Hz}$ have amplitudes of $0.103$ and $0.098$ as a fraction of the height of the main peak, respectively. They are likely due to the nonlinear response of the system discussed in Section~\ref{sec:ifo}.
Also, note that the noise appears to rise at lower frequencies; however, the origin of this noise is not known.

\section{Wandering frequency signal}
\label{sec:viterbi_wandering}
A continuous gravitational-wave signal may wander slowly (and randomly) in frequency over time, due to stochastic internal processes in the superfluid interior of isolated neutron stars, or variable accretion from a stellar companion for neutron stars in binaries such as LMXBs (see Ref.~\cite{SuvorovaEtAl:2017} and references therein).
Although continuous gravitational waves are nearly monochromatic, the long observing times ($\sim 1\,{\rm yr}$) mean that searches can be impacted by very small frequency drifts. 
A typical observation involves $1 \times 10^{10}$ wave cycles, and a coherent search must track the phase to better than half a cycle over the full observation. 
Here we consider the audio analog of a tone that wanders in frequency.

A Fourier transform applied to the whole dataset (as in Section~\ref{sec:single_tone}) is not well suited to the wandering frequency case as the signal is spread across multiple frequency bins. 
In continuous-wave searches, the wandering frequency problem is solved using the Viterbi algorithm,\cite{Viterbi:1967} which can track the signal's frequency over time.
The analysis described and presented in this section is directly inspired by real continuous-wave search methods,~\cite{SuvorovaEtAl:2017} yet is pitched at a level appropriate for an undergraduate laboratory setting. 
In Section~\ref{sec:realCWSearches}, we briefly review the analysis method used by LIGO and Virgo; in Section~\ref{sec:viterbi}, we describe the methods as applied in this work; and in Section~\ref{sec:wanderingResults}, we show results for recovering a wandering frequency signal using the table-top interferometer.

\subsection{Continuous wave analysis with real data}
\label{sec:realCWSearches}

The methods used here are inspired by LIGO and Virgo continuous-wave searches. 
For further details on these methods and continuous-wave searches, we refer the reader to the references in the Supplementary Material.

Continuous-wave searches are performed on long datasets, months to years in duration. 
The frequency of the signal can wander significantly over the observation period. 
In this context, ``significantly'' means across multiple frequency bins, where the typical width of a frequency bin is the reciprocal of the total observation time.\cite{JKS:1998,ScoX1O2Viterbi:2019}

A ``hidden Markov model'' can be used to search for continuous gravitational waves.\cite{SuvorovaEtAl:2017} 
In a Markov process, the current state depends only on the previous state (in this case the ``state'' is the frequency of the signal). 
In a hidden Markov model, the frequency state of the signal is unknown (hidden) and can undergo transitions at discrete times. 
The transitions are Markovian in that the hidden state (i.e.\ frequency) of the system at any time depends solely on its state at the previous time.

A detection statistic relates the observed data to the hidden state and quantifies the likelihood of a signal being present in the data at each frequency and time bin.
This likelihood is also called the emission probability in gravitational-wave literature.  
In gravitational-wave data analysis, the detection statistic gives the likelihood of a signal given the antenna beam pattern of the detector, which varies as the Earth rotates and orbits the Sun.\cite{JKS:1998}
When searching for continuous waves from a neutron star in a binary (such as an LMXB), the Doppler modulation of the source also needs to be taken into account and a different detection statistic is used.\cite{SuvorovaEtAl:2017}

In continuous-wave searches, a physical model of the target informs how far the frequency of the signal can wander over time. 
This is called the transition probability matrix. 
For example, in LMXB searches, the transition matrix allows the signal frequency to (i) stay in the same frequency bin, (ii) move up a single frequency bin, or (iii) move down a frequency bin in each time step.\cite{ScoX1O2Viterbi:2019}
In supernova remnant searches, source frequency is expected to decrease over time, therefore the allowable transitions are to (i) remain in the same frequency bin, or (ii) move down one frequency bin (see the Supplementary Material for LMXB and supernova remnant search references).

The Viterbi algorithm~\cite{Viterbi:1967} is used to find the most probable sequence of hidden frequency states given the sequence of observables.
In the next section, we describe our application of the Viterbi method and our choice of detection statistic.

\subsection{The hidden Markov model and Viterbi algorithm}
\label{sec:viterbi}

\begin{figure}
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/fig04_viterbi_diagram_sizes.pdf}
\caption{\label{fig:viterbi}
Schematic diagram of the Viterbi algorithm. 
The circular nodes represent elements in a time-frequency grid, labeled $t_0$ to $t_{N_t}$ (left to right) in time and $f_0$ to $f_{N_f}$ (bottom to top) in frequency. 
The size of the nodes represents the likelihood of a signal being present in each time-frequency element. 
The small to large sizing corresponds to low to high likelihood values in arbitrary units for this schematic. 
The objective is to find the most probable path of a signal through the grid from left to right.
All possible paths are shown by the lines. 
The black lines show the best path to each node and the gray lines show rejected paths. 
Some routes through the grid result in dead-ends with respect to optimally reaching the other side, such as the path ending at $t_1$ marked (a).
At $t=t_{N_t}$ the algorithm chooses the terminating frequency node which has the highest value given by Eq.~\ref{eqn:pathprobability}, marked (b). 
The Viterbi path is the path leading to this node, highlighted by the thick-black lines, from (b) to the start at (c). 
}
\end{center}
\end{figure}

First, we split the timeseries data from the interferometer into segments. 
Then, we take the discrete Fourier transform of each timeseries segment to form a grid in time and frequency (a spectrogram) with the Fourier amplitude $F(t_i,f_j)$ at frequency $f_j$ and time $t_i$, as the detection statistic (this is the emission probability).
Assuming Gaussian noise, this detection statistic maximizes the likelihood of detecting a sinusoidal signal as described in Appendix~\ref{app:sinusoid_likelihood}.
The detection statistic is normalized for convenience by dividing each value by the maximum Fourier amplitude in the grid (such that $\max_{i,j} F(t_i,f_j) = 1$).

Fig.~\ref{fig:viterbi} represents a spectrogram with $N_t$ time bins and $N_f$ frequency bins.
The circular nodes represent the elements of the time and frequency grid, which are labelled as  
$t_i$ and $f_j$, respectively, where $i=0,1,2,\mathellipsis,N_t$ and $j=0,1,2,\mathellipsis,N_f$. 
The shading of the nodes represents the detection statistic $F(t_i,f_j)$ (the observable) at each grid point, where dark corresponds to lower values and light to higher values.

The objective is to find the most likely path through the grid, given the observed data and any probabilistic constraints on how the frequency of the signal can wander from $t_i$ to $t_{i+1}$. 
The transition probability matrix, $A(f_k,f_m)$, describes the probability of the system transitioning from state $f_k$ at $t_i$ to a state $f_{m}$ at $t_{i+1}$. 
Here, we allow the frequency of the signal to either (i) stay in the same bin, (ii) move up by a single frequency bin, or (iii) move down by a single frequency bin at each transition. 
We assign these three transitions equal probability, i.e.\ $A(f_k,f_m)=1/3$ for $k=m+1,m,m-1$ and $A(f_k,f_m)=0$ otherwise.
All possible transitions are shown as lines in Fig.~\ref{fig:viterbi}.
The different line colors are explained below.

Before we begin analyzing the data, we have no prior knowledge as to which frequency bin the signal starts in (i.e. the prior is flat between the minimum and maximum frequency bins).
At the first stage of the analysis, we define the probability of the system having frequency $f_j$ at the initial time $t_0$ to be equal to the (normalized) detection statistic of that state (i.e.\  ${\rm Pr}[f(t_0) = f_j] = F[t_0,f_j]$).
A specific path (which may not be optimal) is written as $f(t_0), f(t_1),\mathellipsis, f(t_n)$. 
The probability of a specific path given the data is 
\begin{eqnarray}
{\rm Pr}[ f(t_0), f(t_1),&&\mathellipsis, f(t_n) | {\rm data}] \nonumber\\
          && =F[t_n,f(t_n)] A[f(t_n),f(t_{n-1})] \nonumber \\
          && \times\mathellipsis\times A[f(t_1),f(t_0)] F[t_0,f(t_0)] \,.
\label{eqn:pathprobability}
\end{eqnarray}
The path $f^\ast(t_0),\mathellipsis,f^\ast(t_n)$ that maximises ${\rm Pr}[ f(t_0), f(t_1),\mathellipsis, f(t_n) | {\rm data}]$ is the optimal path terminating in the frequency bin $f^\ast(t_n)$. 
We note that the left-hand-side and right-hand-side of Eq.~\ref{eqn:pathprobability} are both evaluated for a specific path and then we maximize over all such paths to find the optimal Viterbi path.

The Viterbi algorithm provides a computationally efficient method for finding the optimal path. 
At every $t_i$ all but $N_f$ possible paths are eliminated. 
Here we describe the algorithm while referring to the schematic in Fig.~\ref{fig:viterbi}.  
The implementation used in this work is available online (see Appendix~\ref{app:code}) and we provide further information (including pseudocode) in Appendix~\ref{app:viterbi}.
\begin{enumerate}
\item Starting at time $t_1$, each $f_j$ node (circles in Fig~\ref{fig:viterbi}) can originate from three prior nodes at time $t_0$ (except for the edge cases $f_0$ and $f_{N_f}$ which only have two). 
The paths between these nodes are indicated by the lines in Fig.~\ref{fig:viterbi}. 
At each $f_j$ node, we select the path with the highest $A[f(t_0),f(t_1)] {\rm Pr}[f(t_0)]$ value as the most probable path. 
These choices are highlighted using the black lines in Fig.~\ref{fig:viterbi} while the gray lines show the rejected paths. 
For example, the most probable connection to the node labeled (a) is the one directly behind it (i.e., $f_4$). 
Therefore, this path is selected as the best path from $t_0$ to $t_1$ for $f_4$.
To allow backtracking at the end, the index of the most probable connection to a node along with the value of the best path to that node is stored, for each node.

\item Moving to time $t_2$, again we select the path which maximizes Eq.~\ref{eqn:pathprobability} for each $f_j$. 
These paths are again shown by the solid black lines between the nodes at $t_1$ and $t_2$ in Fig.~\ref{fig:viterbi}.
Rejected paths are again shown by gray lines.

\item Step 2 is repeated until the end of the grid ($t=t_{N_t}$) is reached with only the best paths being stored at each iteration.

\item We have now found the most probable path to each $f_j$ at $t=t_{N_t}$ and its probability (Eq.~\ref{eqn:pathprobability}). 
We select the terminating frequency bin $f(t_{N_t})$ with the highest probability labeled as (b) in Fig.~\ref{fig:viterbi}.

\item The final step is to find the Viterbi path (the overall best path that terminates in the frequency bin with the highest probability in step 4). 
The Viterbi path is found by backtracking along the stored best connections at each $t_i$ (see also Appendix~\ref{app:viterbi}). 
In Fig.~\ref{fig:viterbi}, it is the path ending at (b) that started at (c) highlighted by the thick-black lines.
\end{enumerate}

In continuous gravitational wave searches, the signal amplitude is expected to be small in comparison to the noise and its frequency can change unpredictably over time. 
The Viterbi algorithm's strength lies in its ability to track such signals through the data even in the presence of comparatively loud noise fluctuations in the time-frequency bins.
In the following section, we present the results of using the Viterbi algorithm with the table-top interferometer data.

\subsection{Wandering frequency signal results}
\label{sec:wanderingResults}

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/fig05_overlay_viterbi_webcam.pdf}
	\caption{\label{fig:viterbi_overlay}
Recovery of a wandering tone using the Viterbi algorithm.
The spectrogram (heat-map) coloring indicates the value of the detection statistic (here, the logarithm of the absolute value of the discrete Fourier transform) in each time-frequency bin -- with brighter colors indicating higher values. 
The overlaid pink-dot and white-cross markers show the injected signal and recovered Viterbi path, respectively. 
On the left, before $\sim 15\,{\rm s}$, the signal changes frequency too quickly for the Viterbi algorithm to recover, given that the algorithm is restricted to only change by one frequency bin per time bin. 
At $150\,{\rm s}$, the data appears anomalous, which may be due to some transient background noise. }
\end{figure*}

We simulate a slowly wandering signal by modulating the frequency sinusoidally with a modulation amplitude that decays with time. 
We use the same apparatus as shown in Fig.~\ref{fig:ifo_schematic_webcam} and the output interference pattern is recorded via webcam as in Section~\ref{sec:single_tone}. 
In this section, we test the Viterbi algorithm’s ability to recover the wandering signal. 
Note that we implicitly approximate the noise at the webcam as white (uniform in frequency) in this implementation of the Viterbi algorithm. 
This approximation ignores the increase in noise at low frequencies (below 1~Hz) shown in Fig.~\ref{fig:webcam_spectrum}.
However, white noise is a reasonable approximation for the narrow frequency range (around $3$--$11\,{\rm Hz}$) we use in this analysis.

The results are shown in Fig.~\ref{fig:viterbi_overlay}, where the heat-map shows the spectrogram of the observed signal (similar to that represented by the schematic in Fig.~\ref{fig:viterbi}). 
In this demonstration we use a signal that can easily be identified in the spectrogram; however, we expect real continuous-wave signals to have far lower signal-to-noise ratios at each time step. 
The overlaid pink dots in Fig.~\ref{fig:viterbi_overlay} show the injected signal and the white crosses show the recovered Viterbi path.
The recovered Viterbi path is within one frequency bin ($\sim 0.3\,{\rm Hz}$) of the injected signal for $94\%$ of the time. We also compute the root-mean-square (RMS) fractional error $E_{\rm rms}$ along the path, which is defined as 
\begin{equation}
E_{\rm rms} = \left[ \frac{1}{N_t+1} \sum_{i=0}^{N_t} \frac{(I_i - R_i)^2}{{I_i}^2} \right] ^{1/2}\,,
\end{equation}
where $I_i$ and $R_i$ are the injected and recovered (Viterbi) frequencies at time $t_i$, respectively. 
For the result shown in Fig.~\ref{fig:viterbi_overlay}, we find that $E_{\rm rms} = 0.082$, which indicates fair but not total recovery of the injected signal.

This may be explained by two anomalies in the recovered path. Initially, the injected signal wanders by more than one frequency bin per time bin (i.e.\ faster than our choice of $A(f_k,f_m)$ allows), thus leading to a discrepancy between the injected and recovered paths for $t\lesssim 10\,{\rm s}$. 
If we remove the first four time bins from the $E_{\rm rms}$ calculation, we find a slight improvement in recovery with $E_{\rm rms} = 0.044$.
One may be tempted to increase the allowed frequency wander in the algorithm; however, this leads to an overall decrease in the above statistics, as the algorithm is prone to jump briefly to nearby spots of noise. 
There is also an anomaly at $150\,{\rm s}$, which is likely due to a local disturbance, e.g. someone walking past the interferometer. 
As shown in Fig.~\ref{fig:viterbi_overlay}, the Viterbi algorithm can recover and continue to track the signal after the disturbance.

\section{Complex audio: music and speech}
\label{sec:optical_microphone}

In this section, we explore how our apparatus can be used to teach a selection of signal processing techniques. 
We use complex audio signals (such as music and speech) as natural successors to the constant and wandering tones used in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, respectively.
As complex audio signals are not quasi-monochromatic, the Viterbi algorithm used in Section~\ref{sec:viterbi_wandering} is not directly applicable here. 
Instead, we use a hierarchy of passive filters which suppress noise, yet do not assume any specific form of the signal, unlike the Fourier-based maximum likelihood filter which is tuned to the sinusoidal signals in Section~\ref{sec:single_tone} and Appendix~\ref{app:sinusoid_likelihood}.

We use the Michelson interferometer as an ``optical microphone'' to detect sound, replacing the components of a conventional microphone with a laser interferometer.
The only change to our apparatus is replacing the webcam with a photodiode to allow us to capture the higher frequencies necessary for speech and music (see Section~\ref{sec:photodiode}). 
Optical microphones have precedence in the laser microphones~\cite{laser_microphone} which are (or were historically) used in the defense industry and operate on a variety of related principles. 
Our objective is to play a recording of speech or music through the speaker attached to mirror M2 (see Fig.~\ref{fig:ifo_schematic_webcam}), record the resulting interference pattern, and then recover the original signal via a selection of signal processing techniques.

The apparatus serves as an independent demonstration for a broader physics and engineering audience, particularly in undergraduate laboratories. 
We describe additional hardware components required for this demonstration in Section~\ref{sec:photodiode} and the initial results in Section~\ref{sec:initialResultsOpMic}. 
We consider a selection of filter techniques, details of which, along with a summary of digital signal processing resources, can be found in the Supplementary Material. 
In Section~\ref{sec:opticalMicResults}, we present the two best-performing techniques from the Supplementary Material.

\subsection{Hardware modifications for the optical microphone}
\label{sec:photodiode}

The human ear can hear frequencies in the range of $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
Speech intelligibility (the ability to understand speech) requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$. 
Therefore, the optical microphone requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music (adjusting for the Nyquist frequency). 
This cannot be achieved with the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering} as it has a sampling rate of $30\,{\rm Hz}$ and thus can only ``hear'' frequencies below $15\,{\rm Hz}$.
To overcome this issue, we use a photodiode~\footnote{A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the reverse direction. As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction.} at the output of the interferometer to achieve a sampling rate of $16\,{\rm kHz}$.

We place an OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp which together produce a voltage that depends on the incident intensity. 
The photodiode records the interference pattern at roughly the same off-center position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}, again chosen arbitrarily. 
The voltage signal from the photo-detector is captured by an MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2, which provides a convenient means to record the photodiode data.
Together, the circuit samples the signal at $\sim 16\,{\rm kHz}$. 
Resources for using the Raspberry~Pi and photodiode, including a circuit diagram, are described in the Supplementary Material.

Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing (folding of frequencies greater than half the sampling rate) into the detected range. We include an anti-aliasing Sallen-Key filter with a cut-off frequency of $8\,{\rm kHz}$ before the ADC to prevent this from happening. 
This component attenuates any frequencies above $8\,{\rm kHz}$ before they are digitally sampled. We also place a cloth screen over the face of the photodiode to reduce the incident intensity and avoid saturating the ADC -- an improvised, physical solution that could instead be replaced by scaling down the voltage electronically. This cloth screen was re-purposed grill cloth from a commercial speaker.

\subsection{Anti-aliased output}
\label{sec:initialResultsOpMic}

We test the optical microphone with a variety of recordings, including the speech of different people and music ranging from simple melodies and rhythms to songs. 
During recordings, care is taken to minimize activity around the demonstration to reduce environmental noise coupling into the interferometer. 
The timeseries data is then directly converted to a .wav file and played as an audio recording using the \texttt{scipy.io.wavfile.write} function in Python (see Appendix~\ref{app:code}).
When processing the results, we restrict our analysis to only the first $10\,{\rm s}$ of each observation (for efficiency), and only plot the first second in our results (Fig.~\ref{fig:notchWienerLogMMSEResults}).

The raw output of the optical microphone (with anti-aliasing) is noisy with a loud, continuous bass hum. 
This can be explained by looking at the power spectral density (PSD) of the background noise (i.e., the output with the speaker switched off), shown in Fig.~\ref{fig:psd_noise}. 
The spectrum is dominated by AC eletrical power grid noise from the fundamental $50\,{\rm Hz}$ Australian mains electricity grid signal up to and beyond the $8$th harmonic (at $400\,{\rm Hz}$). 
The mains signal is also present, but far weaker, in the background spectrum taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution. 
Besides lighting, other possible contributions to the mains signal include air conditioning and the photodiode circuit itself. 
The appearance of harmonics of the mains noise might be due to the non-linearity in the system discussed in Section~\ref{sec:ifo}. 
The spectrum in Fig.~\ref{fig:psd_noise} also has a broad feature at around $750\,{\rm Hz}$, the origin of which is yet to be determined.
Environmental noise reduction for gravitational-wave detectors is an active area of research (see the Supplementary Material for further information and resources on this topic).

\begin{figure}
	\includegraphics[width=.5\textwidth]{figures/fig06_psd_podo_14_6.pdf}
	\caption{\label{fig:psd_noise}
Power spectral density (PSD) of background noise from the optical microphone (with the speaker off). 
We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). Otherwise, the PSD is fairly white except for a peak at around 0.75~kHz.
}
\end{figure}

\subsection{Optical microphone results}
\label{sec:opticalMicResults}

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/fig07_combined_highlight_results.pdf}
\caption{\label{fig:notchWienerLogMMSEResults}
Timeseries (left column) and frequency spectrum (right column) results with the optical microphone. 
The original input signal in the first row is a $1\,{\rm s}$ recording of an adult male voice (saying ``a cathode''). 
 The input signal is shifted by 0.12 s to the right to synchronize the manual delay from starting the recording with the Raspberry Pi and starting to play the source through the speaker. 
The second row shows the raw output from the optical microphone when the input from the first row is played. 
The third row shows the result of applying the notch and Wiener filters combined. 
The fourth row shows the result of applying the logMMSE estimator, where the rise at the start of the timeseries is an expected effect when filtering a signal of finite duration. 
}
\end{center}
\end{figure*}

We explore several filters to remove the $50\,{\rm Hz}$ mains hum and harmonics and improve the speech intelligibility of the recording.
The Supplementary Material describes a range of analysis techniques that can be used as examples for the undergraduate laboratory. 
All filters are tested on the same $1\,{\rm s}$ long speech recording.
The results of this section are shown in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
In the figure, the left and right columns show the timeseries and frequency spectrum, respectively. 
The first row shows the input signal played through the speaker (see Fig.~\ref{fig:ifo_schematic_webcam}). 
The second row shows the raw output from the photodiode recording.

In signal processing, the ideal filter would be one that:
(i) completely attenuates the undesired parts of the spectrum, 
(ii) does not change the rest of the spectrum, and 
(iii) smoothly transitions between these regions, as to not damage the time domain signal when seen under convolution. 
However, these three conditions cannot all hold at once. 
For example, if conditions (i) and (ii) hold, then the filter must be discontinuous at the boundary of the undesired region but this implies that the filter has ``infinite latency'' and so will affect (or damage) the time domain signal for infinite time.\cite{10.5555/151045}
Therefore, any filter must compromise between these three conditions. 
For speech intelligibility, this means that either: (i) some noise remains in the filtered recording, (ii) some of the speech content is lost as certain important frequencies are attenuated, or (iii) the speech is somewhat distorted in time.
All three of these cases can, when taken to the extreme, make the speech intelligibility worse than the unfiltered recording. 
Therefore, we choose filters that compromise between achieving the three conditions.

In this section, we present the results of two advanced signal processing techniques applied to the optical microphone recordings. 
The techniques are only briefly described here and we refer the reader to the Supplementary Material for further details and other analysis techniques.

Firstly, we consider two signal processing techniques used in combination: the cascaded notch and the Wiener filter (see also the Supplementary Material). 
A notch filter removes signals within a specific frequency range. 
We want to remove the $50\,{\rm Hz}$ mains noise and harmonics, therefore we use a cascaded notch filter where each notch is centered on one of the harmonics. 
The Wiener filter is an advanced statistical technique that makes use of statistical information from the speech data and noise. 
It amplifies parts of the signal with a high signal-to-noise ratio while suppressing parts with a low signal-to-noise ratio. 
The results of the combined cascaded notch and Wiener filter are shown in the third row in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
Most of the mains noise is removed; however, the recovered voice sounds muffled and is not understandable.

Secondly, we apply a speech enhancement technique. 
Ref.~\cite{SubjectiveComparison} compares $13$ speech enhancement methods, finding the log minimum mean-square error (logMMSE) estimator to be the best, qualitatively, at recovering speech (see also the Supplementary Material). 
We use an existing implementation of the logMMSE from Ref.~\cite{logmmse}.  
The logMMSE estimator results are shown in the bottom panels in Fig.~\ref{fig:notchWienerLogMMSEResults}. 
We see significant attenuation of the mains harmonics and general smoothing of the spectrum. 
Most of the background noise is removed; however, the logMMSE still does not significantly enhance the speech as the voice sounds muffled and indistinct.

We find some improvement with music over speech. 
Simple chords and drums can be heard after filtering, but more composite sounds and complex melodies cannot be heard clearly. 
Our observations suggest that this is especially true for certain instruments; in particular flutes and violins sometimes can’t be heard at all. 
This could be a perceptual effect or a frequency dependence somewhere in the optical microphone.
Speculating, perhaps the speaker-mirror coupling is stronger at low frequencies and thus instruments like electric bass and drums sound louder in the results.
To address these problems, we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all. 
To determine this requires a better understanding of the system, as discussed in Section~\ref{sec:future_work}.

\section{Future work}
\label{sec:future_work}

\begin{figure*}
    \includegraphics[width=.8\textwidth]{figures/fig08_pipeline_nobox.pdf}
	\caption{\label{fig:pipeline_highlighted}
Flowchart for the signal through the optical microphone. 
A transfer function of the system would have to account for each stage and exploration of this is an aim of future work. 
Not shown are the signal processing filters that are performed on the PC once the data is transferred off the Raspberry~Pi.
}
\end{figure*}

The experiment described here can be used to demonstrate and teach a variety of topics, from basic optics and photodiode circuits, to signal processing and speech enhancement. 
Cross-disciplinary undergraduate laboratory experiments between physics and engineering courses are beneficial to student learning (see Section II in the Supplementary Material for further information). 
Here, we consider further extensions and improvements to the experiment such as noise isolation, transfer function analysis, speech recognition, and control systems.

Isolating the experiment from interference may provide some noise reduction. 
The Raspberry Pi could be powered with a commercial battery instead of mains power, which may help to isolate the circuit from the mains as all other components draw power from the Raspberry~Pi (the laser is DC-powered through a converter).
Transferring the circuit from a breadboard to a printed circuit board can also reduce interference.\cite{elfekey2013design}
A feedback loop control system can also be used to suppress any unwanted movement of the interferometer components,\citep{abbott2017exploring, Sekiguchi:2016bmv, verhoeven2009robust} similar to methods used to isolate gravitational-wave observatories.

A thorough characterization of the system's transfer function would allow us to understand how audio signals couple through the interferometer. 
The total transfer function starts with the voltage being sent to the speaker and ends with the voltage recorded by the Raspberry~Pi. 
It would include any inherent non-linearities in the speaker, the speaker-mirror coupling, the path length to intensity relation, and the photodiode. 
The schematic in Fig.~\ref{fig:pipeline_highlighted} shows the abstract signal flow through the system.
It does not include other potential pathways for signal flow such as acoustomechanical vibrations of the webcam or photodiode mount which could also be examined.

A voluminous literature exists on using hidden Markov models trained on phonemes to recognize speech.\cite{HMM_english}
These could perform the final stage of speech enhancement for the optical microphone. 
This coincidentally connects back to the Viterbi algorithm in Section~\ref{sec:viterbi_wandering}, which is also underpinned by a hidden Markov model, albeit of a different type. 
Alternatively, machine learning solutions exist throughout the field that can provide results competitive with statistical techniques (such as the logMMSE estimator used here).\cite{SEGAN}

More traditional techniques such as a wavelet transform~\citep{nason1995stationary} could be used to extract the signal from noise and compared with the above methods. 
A wavelet transform provides both time and frequency information, making it easier to pinpoint the origin of noise with respect to time. 
In Refs.~\cite{tufekci2000feature,agbinya1996discrete}, wavelet transform methods are proposed for speech recognition.

Returning to the demonstration of gravitational wave analysis, single interferometers cannot yield directional information for signals; a large proportion of the directional information in gravitational-wave detections comes from the time delay offset between the observation being recorded at two or more detectors.\cite{GW150914}
We could therefore extend this analysis to include data from two interferometers to extract directional information.
This would require increasing the sensitivity of the optical microphone to pick up the signal from a distant source instead of from a speaker attached directly to one of the mirrors of the interferometer.
Another extension would be to demonstrate the Doppler effect of the Earth's motion around the Sun, which needs to be considered in continuous-wave searches (see Section~\ref{sec:realCWSearches} and Ref.~\cite{JKS:1998}). 
One approach could be to modify the input audio signal to simulate Doppler modulation.

\section{Conclusions}
\label{sec:conclusions}

In this paper, we use a table-top Michelson interferometer as an analog to a gravitational-wave detector, demonstrating signal processing techniques used within the gravitational-wave community.
We explore the use of the interferometer as an optical microphone and consider a more general treatment of signal processing with complex audio signals, which can also serve as a distant analog to minimally modelled gravitational-wave burst signals, e.g., from supernovae.
The demonstration can be adapted for use in both the physics and engineering undergraduate laboratory, providing opportunities for cross-disciplinary teaching. 
Additionally, it can be used as a tool for explaining gravitational-wave research to a wider, non-specialist audience.

As the field of gravitational-wave astrophysics continues to grow, the future will bring many more detections of binary black holes and neutron stars, as well as the anticipated first detection of other classes of signals, such as continuous waves, to which this demonstration provides some charming insights.

\begin{acknowledgments}

The authors are grateful to Jude Prezens, Alex Tolotchkoc, and Blake Molyneux for their technical advice and generous assistance; Patrick Meyers, Margaret Millhouse, Sofia Suvorova for useful discussions; Patrick Clearwater, Patrick Meyers, Suk Yee Yong, Lucy Strang, Julian Carlin, Sanjaykumar Patil, and Alex Cameron for early work on the interferometer design requirements and construction; and the LIGO Education Public Outreach working group, in particular, Anna Green, Lynn Cominsky, Sam Cooper, and Martin Hendry, for their helpful feedback and suggestions.  
This research is supported by the Australian Research Council Centre of Excellence for Gravitational Wave Discovery (OzGrav) (project number CE170100004). 
Financial support towards hardware was provided by the Institute of Physics International Member Grant and the OzGrav Outreach Support scheme. 
Travel support was provided by the Australian National University PhB Science program.
This work has been assigned LIGO document number P2000386.

\end{acknowledgments}

\appendix

\section{Open-source code}
\label{app:code}
This project is implemented in Python 3 scripts and jupyter notebooks and MATLAB. 
We refer the reader to the Supplementary Material for software references. 
The current build and sample data can be found at:
\url{https://github.com/daccordeon/gravexplain}

\section{Detecting a sinusoidal signal in Gaussian noise}
\label{app:sinusoid_likelihood}

In this appendix, we demonstrate that the modulus of the Fourier transform is an appropriate detection statistic when searching for a sinusoidal signal in Gaussian noise. 
We describe the data as
\begin{equation}
x(t) = s(t) + n(t)\,, 
\label{eqn:GNdata}
\end{equation}
where $s(t)$ and $n(t)$ are the signal and noise, respectively.
The signal takes the form
\begin{equation}
s(t) = A \cos\left[{\omega t + \phi}\right]\,,
\label{eqn:GNmodel}
\end{equation}
where $A$, $\omega$, and $\phi$ are the amplitude, angular frequency and phase of the signal, respectively. 
The noise $n(t)$ is a fluctuating zero-mean time series with the following property: if we define an inner product between two arbitrary time series $u(t)$ and $v(t)$ as 
\begin{equation}
\ip{u}{v} = \frac{1}{T} \int_0^T \mathrm{d} t \, u(t) v(t)\,,
\label{eqn:ipuv}
\end{equation}
where $T$ is the total time of the observation, then the likelihood $\mathcal{L}$ of measuring the noise-noise product $\ip{n}{n}$ is given by 
\begin{equation}
\mathcal{L} = \exp\left( -\frac{1}{2}\ip{n}{n}\right)\,.
\label{eqn:ipnn}
\end{equation}
Equations~\ref{eqn:ipuv} and~\ref{eqn:ipnn} define what it means for noise to be Gaussian through the fundamental measurement of $\ip{n}{n}$.

The likelihood of measuring the signal $s(t)$ in the presence of noise follows from Eqs.~\ref{eqn:GNdata} and~\ref{eqn:ipnn} by replacing $n(t)$ in Eq.~\ref{eqn:ipnn} with $x(t) - s(t) = n(t)$ from Eq.~\ref{eqn:GNdata}.~\cite{JKS:1998,Jaynes:2003}
The result is
\begin{eqnarray}
\mathcal{L} &=& \exp\left( -\frac{1}{2}\ip{x-s}{x-s} \right) \label{eqn:likeOne}\,, \\ 
            &=& \exp\left( -\frac{1}{2}\ip{x}{x} - \frac{A^2}{2} \ip{\cos[\omega t + \phi]}{\cos[\omega t + \phi]} \right. \nonumber\\
                 &&\hspace{3em} \left. +  A \ip{x}{\cos[\omega t + \phi]} \vphantom{\frac12}\right) \,,\\ 
                 &=& \exp \left( -\frac{1}{2}\ip{x}{x} - \frac{A^2}{4} + A \ip{x}{\cos[\omega t + \phi]} \right)\,.\label{eqn:logLike}
\end{eqnarray}
We then maximise Eq.~\ref{eqn:logLike} with respect to $A$, obtaining 
\begin{eqnarray}
\label{eqn:maxLogLike}
\mathcal{L}_\mathrm{max} = \exp \left( - \frac{1}{2} \ip{x}{x} + \ip{x}{\cos[\omega t + \phi]}^2 \right)\,,
\end{eqnarray}
for $A = 2\ip{x}{\cos[\omega t + \phi]}$.
From the second term of Eq.~\ref{eqn:maxLogLike}, we see that the maximum likelihood of a sinusoidal signal in Gaussian noise is the modulus of the cosine Fourier transform, plus the term $\ip{x}{x}$, which is independent of $\omega$ and $\phi$ and can therefore be ignored when searching over $\omega$.

Two important points must be made about the above procedure. 
(i) Fundamentally the goal is to maximize $\mathcal{L}$ in Eq.~\ref{eqn:likeOne} by varying $s(t)$ through $A$.
For the special case of the Gaussian likelihood (Eq.~\ref{eqn:ipnn}), which peaks at $\ip{n}{n}= 0$, this is equivalent to minimizing the difference between $x(t)$ and $s(t)$ as evident in Eq.~\ref{eqn:ipuv}.
In general, however, minimizing the difference between $x(t)$ and $s(t)$ is not equivalent always to the fundamental goal of maximising $\mathcal{L}$, for example if $\mathcal{L}$ peaks at $\ip{n}{n} \neq 0$, or if $\mathcal{L}$ has multiple maxima.
(ii) The maximum likelihood $\mathcal{L}_{\rm max}$ in Eq.~\ref{eqn:maxLogLike} (or equivalently its logarithm) defines the detection statistic. 
When its value exceeds a threshold (chosen freely by the analyst) at some value of $\omega$, a signal is deemed to have been detected at that value of $\omega$.
Therefore the specific functional form of Eq.~\ref{eqn:maxLogLike} matters, which is a second reason why one must start from Eq.~\ref{eqn:ipnn} rather than Eq.~\ref{eqn:likeOne}, in addition to reason (i).

\section{Viterbi algorithm}
\label{app:viterbi}

This appendix contains some details regarding the implementation of the Viterbi algorithm described in Section~\ref{sec:viterbi}. 
The Viterbi algorithm~\cite{Viterbi:1967} is a classic method in signal processing, whose theoretical underpinnings and implementation are accessible to undergraduate students. 
See the Supplementary Material for further resources and other pseudocode examples available online.

Here, we present some pseudocode (below) of the implementation used in Section~\ref{sec:viterbi}. We use Fourier amplitudes, normalized between $(0, 1)$ by dividing by the maximum value in the grid, as multiplicative weights. 
To avoid numerical underflow we take the logarithm of the Fourier amplitudes, which we can equivalently use as additive weights.

Let $X$ be a grid of $j=0,\ldots,N_f$ rows and $i=0,\ldots,N_t$ columns of additive weights for each node. Let Y and Z be grids of the same shape to store the weight of the best path to each node and the row index of the previous node on that path, respectively. Let W be a length $N_t+1$ array to store the final sequence of row indices for the optimal path. We restrict paths to only move one cell up or down at a time (or stay constant). For the boundary cases of $j=0,N_f$, we only search over $k \in \{0,1\}$ and $k \in \{-1,0\}$, respectively, to stay inside the grid (this is not shown in the pseudocode below).

\begin{algorithmic}
\Function{Viterbi}{$X$}

    \For{each row $j=0,\ldots,N_f$}
		\State $Y_{0,j} \gets X_{0,j}$
    \EndFor

    \For{each column $i=1,\ldots,N_t$}
	    \For{each row $j=0,\ldots,N_f$}

	    	\State $Y_{i,j} \gets X_{i,j} + \underset{k \in \{-1,0,1\}}{\max} (Y_{i-1,j+k})$
	    	\State $Z_{i,j} \gets j + \underset{k \in \{-1,0,1\}}{\arg\max} (Y_{i-1,j+k})$

	    \EndFor
    \EndFor

    \State $W_{N_t} \gets \underset{j=0,\ldots,N_f}{\arg\max} (Y_{N_t,j})$

    \For{each col $i=N_t-1,\ldots,0$}

		\State $W_i \gets Z_{i+1, W_{i+1}}$

    \EndFor

    \State \Return $W$
\EndFunction
\end{algorithmic}

\bibliographystyle{myunsrt}
\bibliography{ifoDemoBib}

\end{document}

### 2021-11-15 paper/finalSubmission/JGardnerManuscript/JGardner.tex
moving author list into main tex adding build script for submission

@@
-%\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-2}
-\documentclass[prb,preprint]{revtex4-2}
+\documentclass[aps,pra,superscriptaddress,reprint]{revtex4-2}
 
@@
 
-\input{authorList.tex}
+\author{James~W.~Gardner}
+\email{james.gardner@anu.edu.au}
+\affiliation{%
+Centre for Gravitational Astrophysics, The Australian National University, Acton, ACT, 2601, Australia
+}
+\affiliation{%
+OzGrav-ANU, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, The Australian National University, Acton, ACT, 2601, Australia
+} 
+\affiliation{%
+School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+
+\author{Hannah~Middleton}
+\email{hannah.middleton@unimelb.edu.au}
+\affiliation{%
+School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+\affiliation{%
+Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria, 3122, Australia 
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria, 3010, Australia
+}
+
+\author{Changrong~Liu}
+\email{changrongl1@student.unimelb.edu.au}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria, 3010, Australia
+}
+
+\author{Andrew~Melatos}
+\email{amelatos@unimelb.edu.au}
+\affiliation{%
+School of Physics, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria, 3010, Australia
+}
+
+\author{Robin~Evans}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+\affiliation{%
+OzGrav-Melbourne, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, Parkville, Victoria, 3010, Australia
+}
+
+\author{William~Moran}
+\affiliation{%
+Department of Electrical and Electronic Engineering, University of Melbourne, Parkville, Victoria, 3010, Australia
+}
+
+\author{Deeksha~Beniwal}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Huy~Tuong~Cao}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Craig~Ingram}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Daniel~Brown}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
+
+\author{Sebastian~Ng}
+\affiliation{%
+Department of Physics, The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+The Institute of Photonics and Advanced Sensing (IPAS), The University of Adelaide, South Australia, 5005, Australia
+}
+\affiliation{%
+OzGrav-Adelaide, Australian Research Council Centre of Excellence for Gravitational Wave Discovery, South Australia, 5005, Australia
+}
 
@@
 \end{equation}
-where $T$ is the total time of the observation, then the likelihood $\mathcal{L}$ of measuring the noise-noise product $\ip{n}{n}$ is given by 
+where $T$ is the total time of the observation, then the probability $\mathcal{L}$ of measuring the noise-noise product $\ip{n}{n}$ is given by 
 \begin{equation}

### 2021-11-15 paper/finalSubmission/JGardnerManuscript/pdfbuild.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+##########################
+# build the pdf by doing #
+#     bash pdfbuild.sh   #
+##########################
+
+pdflatex JGardner.tex
+bibtex JGardner
+pdflatex JGardner.tex
+pdflatex JGardner.tex

### 2021-11-16 paper/finalSubmission/JGardnerSupplementaryMaterials/JGardnerSupplementaryMaterial.tex

--- 2021-11-01 paper/ifo-supplementary-material.tex
@@
 \begin{figure*}
-	\includegraphics[height=0.41\textwidth]{figures/setup_pic2.pdf}
-	\includegraphics[height=0.41\textwidth]{figures/circuit_pic3.pdf}
+	\includegraphics[height=0.41\textwidth]{figures/fig01a_setup_pic2.pdf}
+	\includegraphics[height=0.41\textwidth]{figures/fig01b_circuit_pic3.pdf}
 	\caption{
@@
 \begin{figure*}
-	\includegraphics[angle=-90,width=.8\textwidth]{figures/circuit_diagram_2.pdf}
+	\includegraphics[angle=-90,width=.8\textwidth]{figures/fig02_circuit_diagram_2.pdf}
 	\caption{\label{fig:circuit_diagram}
@@
 	\includegraphics[width=.48\textwidth]
-{figures/combined_results_full_melatos_labelled.pdf}
+{figures/fig03_combined_results_full_melatos_labelled.pdf}
 	\end{center}
@@
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{figures/butterworth_150_3000.pdf}
+	\includegraphics[width=0.5\textwidth]{figures/fig04_butterworth_150_3000.pdf}
 	\caption{
@@
 \begin{figure}
-	\includegraphics[width=.5\textwidth]{figures/psd_butterworth_14_6.pdf}
+	\includegraphics[width=.5\textwidth]{figures/fig05_psd_butterworth_14_6.pdf}
 	\caption{\label{fig:psd_noise}
@@
 \begin{center}
-\includegraphics[width=.5\textwidth]{figures/cascaded_notch_response_plot.pdf}
+\includegraphics[width=.5\textwidth]{figures/fig06_cascaded_notch_response_plot.pdf}
 \end{center}
